### 2025-04-11 article/muon.tex
added statement

%% ============================================
%% ================ Preambule =================
%% ============================================
\documentclass[]{scrartcl}
\usepackage[margin = 0.5in]{geometry}

\usepackage[pdftex,unicode, 
colorlinks=true,
linkcolor = blue]{hyperref}	% нумерование страниц, ссылки!!!!ИМЕННО В ТАКОМ ПОРЯДКЕ СО СЛЕДУЮЩИМ ПАКЕТОМ
%\usepackage[warn]{mathtext}				% Поддержка русского текста в формулах
\usepackage[T1, T2A]{fontenc}			% Пакет выбора кодировки и шрифтов
\usepackage[utf8]{inputenc} 			% любая желаемая кодировка
\usepackage[english]{babel}		% поддержка русского языка
\usepackage{wrapfig}					% Плавающие картинки
\usepackage{amssymb, amsmath}			% стилевой пакет для формул
\usepackage{algorithm}
\usepackage{algorithmic}

\ifpdf
\usepackage{cmap} 				% чтобы работал поиск по PDF
\usepackage[pdftex]{graphicx}
%\usepackage{pgfplotstable}		% Для вставки таблиц.
\pdfcompresslevel=9 			% сжимать PDF
\else
\usepackage{graphicx}
\fi

\graphicspath{{./figures/}}
\usepackage{subcaption}
%% ============================================
%% ================ Info =================
%% ============================================
\title{Neon: Nuclear Norm to Beat Muon}

\author{
  Alexey Kravatskiy\\
  \texttt{kravtskii.aiu@phystech.edu}
  \and
  Ivan Kozyrev\\
  \texttt{kozyrev.i?@phystech.edu}
  \and
  Nikolay Kozlov \\
  \texttt{kozlov.n?@phystech.edu}
}

%\author{\begin{tabular}{c c}
%	  	 Alexey Kravatsky & Nikolay Kozlov & Ivan Kozyrev \\
%		 \texttt{kravatskii.aiu@phystech.edu} & \texttt{kozlov.n?@phystech.edu} & \texttt{kozyrev.i?@phystech.edu}  
%		\end{tabular}}
\date{\today}

\begin{document}

\maketitle

\begin{abstract}
In this paper, we develop a new algorithm for optimization of functions of weight matrices, which are typical for training large language models. Changing spectral norm, which was used to derive Muon, to nuclear norm, we pose a new optimization problem for an update matrix, solution of which defines a novel method we name Neon. After providing theoretical guarantees of Neon convergence, we compare performances of Neon, Muon, and Adam on training multilayer perceptron and BERT vectorizer.

\end{abstract}

\section{Idea}
The goal of the project is to make variations on Muon to speed it up. Recently, authors of \cite{bernstein2024oldoptimizernewnorm} have proposed to look at different optimizers as the solution of the optimization problem for an update. This approach can be utilized to derive Muon \cite{jordan2024muon}, a novel algorithm for fast training of neural networks. Instead of using spectral norm in the problem, we use nuclear norm to produce a new problem. Then, we add momentum to update. After finalizing the algorithm, we test it on MLP, and, if results are satisfactory, on transformers (probably something from huggingface). The result will be a fast algorithm, which we will convert into a new optimizer class for PyTorch, as was done with Muon.

\bibliographystyle{unsrt}
\bibliography{muon}

\end{document}

### 2025-04-22 article/muon.tex
emails updated
new author
add gitignore
added sections we need to fill until saturday
completed the problem subsection

@@
 \usepackage{subcaption}
+
+%custom commands
+\newcommand{\norm}[1]{\Vert{#1}\Vert}
+
 %% ============================================
@@
   Ivan Kozyrev\\
-  \texttt{kozyrev.i?@phystech.edu}
+  \texttt{kozyrev.in@phystech.edu}
   \and
   Nikolay Kozlov \\
-  \texttt{kozlov.n?@phystech.edu}
+  \texttt{kozlov.na@phystech.edu}
+  \and
+  Alexander Vinogradov \\
+  \texttt{vinogradov.am@phystech.edu}
 }
@@
 Instead of using spectral norm in the problem, we use nuclear norm to produce a new problem.
-Then, we add momentum to update.
+%Then, we add momentum to update.
 After finalizing the algorithm, we test it on MLP, and, if results are satisfactory, on transformers (probably something from huggingface).
-The result will be a fast algorithm, which we will convert into a new optimizer class for PyTorch, as was done with Muon.
+The result will be a fast algorithm, which we will convert into a new optimizer class for PyTorch, as was done with Muon.%
+
+\subsection{Problem}
+
+In this subsection, we provide a more detailed description of our idea and formulate it as a mathematical problem.
+The authors of \cite{bernstein2024oldoptimizernewnorm} suggest obtaining the update step as a solution to the optimization problem:
+\begin{equation}
+    \langle g, \delta w \rangle + \lambda \norm{\delta w}^2 \to \min_{\delta w}\,,
+\end{equation}
+where $w$ is the weight vector, $g$ is a gradient-like vector (e.g., obtained via momentum SGD), and $\norm{\cdot}$ represents a certain norm.
+Many popular optimizers, such as Adam (with exponential moving average disabled) and vanilla SGD, can be cast within this framework \cite{bernstein2024oldoptimizernewnorm}.
+
+In large language models, most weights are structured as matrices, which offers additional opportunities for optimization.
+Let $W$ be the weight matrix of a linear layer, and $G$ be a gradient-like matrix.
+Then, the update step $\delta W$ can be obtained as a solution to the optimization problem:
+\begin{equation}\label{eqn:opt_problem_mat}
+  \langle G, \delta W \rangle + \lambda \norm{\delta W}^2 \to \min_{\delta W}\,,
+\end{equation}
+where $\norm{\cdot}$ denotes a certain matrix norm.
+By setting this norm to the RMS-to-RMS norm (a scaled version of the spectral norm), we recover the Muon optimizer \cite{bernstein2025deriving, bernstein2024oldoptimizernewnorm} with an update step defined by:
+\begin{equation}
+\delta W = - \frac{1}{\lambda}\sqrt{\frac{n}{m}}UV^T\,,
+\end{equation}
+where $m$ is the input dimension of the layer, $n$ is the output dimension, and $U$ and $V$ are obtained from the singular value decomposition of the gradient matrix $G = U \Sigma V$.
+
+Motivated by the recent achievements of the Muon optimizer (e.g., \cite{liu2025muon}), we consider alternative choices of norms, specifically the kernel norm $\norm{\cdot}_*$ and a custom $F*$ norm, given by $\norm{X}_{F*}^2 = (\norm{X}_F + \norm{X}_*)/2$, where $\norm{\cdot}_F$ denotes the Frobenius norm.
+
+Using the kernel norm in \eqref{eqn:opt_problem_mat} leads to a rank-one update of the weight matrices:
+\begin{equation}\label{eqn:update_star}
+  \delta W = \frac{1}{\lambda} u_1 \sigma_1 v_1^T\,,
+\end{equation}
+where $\sigma_1$ is the largest singular value, and $u_1$ and $v_1$ are the corresponding singular vectors.
+We expect one iteration of this method to be significantly faster than one iteration of Muon.
+
+Another choice is the $F*$ norm. With this choice, \eqref{eqn:opt_problem_mat} yields 
+\begin{equation}\label{eqn:update_F_star}
+\delta W = UDV^T
+\end{equation} 
+with $D = \text{diag}(d_i)$, where $d_i = [\sigma_i - \tau]_+$ and $\tau$ is given by:
+\begin{equation}
+    \sum_{i=1}^n [\sigma_i - \tau]_+ = \tau\,.
+\end{equation}
+We anticipate that the method with this update step will perform well with large batch sizes.
+
+In this article we show how one can quickly compute weight updates defined by \eqref{eqn:update_star} and \eqref{eqn:update_F_star}.
+Then we finilize the methods by adding momentum and test their performance against those of Muon ant training multilayer perceptron and transformer.
+The results will be fast algorithms, which we will convert into a new optimizer classes for PyTorch, as was done with Muon.
+
+\section{Outcomes}
+
+\section{Literature review}
+
+\section{Quality metrics}
+
+\section{Preliminary plan}
+
+\section{Prototyping phase report}
 

### 2025-04-25 article/muon.tex
added missed 1/\lambda
Added plans, quality metrics etc.

@@
 In this paper, we develop a new algorithm for optimization of functions of weight matrices, which are typical for training large language models.
-Changing spectral norm, which was used to derive Muon, to nuclear norm, we pose a new optimization problem for an update matrix, solution of which defines a novel method we name Neon.
-After providing theoretical guarantees of Neon convergence, we compare performances of Neon, Muon, and Adam on training multilayer perceptron and BERT vectorizer.
+Changing spectral norm, which was used to derive Muon, to nuclear norm, we pose a new optimization problem for an update matrix, solution of which defines a novel algorithm we name Neon.
+To make it feasible, we use Lanzos algorithm to find a required step.
+After providing theoretical guarantees of Neon convergence, we compare performances of Neon, Muon, and Adam on training multilayer perceptron, convolutional neural network and NanoGPT.
 
@@
 
-\subsection{Problem}
+\subsection{Problem (Project description)}
 
@@
 \begin{equation}\label{eqn:update_F_star}
-\delta W = UDV^T
+\delta W = \frac{1}{\lambda}UDV^T
 \end{equation} 
@@
 \section{Outcomes}
+The main results of the project are expected to be:
+\begin{enumerate}
+    \item Deriving of Neon and argumentation of the numerical methods which will be used in the algorithm
+    \item Code implementation of Neon as PyTorch optimizer class
+    \item Numerical experiments comparing performance of training with Neon, Muon, SVD and Adam on several architectures: MLP, CNN, and NanoGPT.
+    \item Participation in CIFAR-10 challenge with Neon (if the alrothims turns out to be competitive)
+    \item Article describing all the results. If they are satisfactory, we intend to be in time to apply for NeurIPS 2025. It is a challenge, but I think we must give it a try.
+\end{enumerate}
 
 \section{Literature review}
-
+TO-DO.
 \section{Quality metrics}
+\begin{enumerate}
+    \item The derivation is theoretically solid
+    \item The numerical procedure used to compute a step is grounded and has estimated time overhead (say, in FLOPS)
+    \item The code with Neon trains MLP and CNN (and NanoGPT, but it's a bonus) less than 3 times slower than Adam
+    \item Instruction of setting the parameters of the algorithm are presented and justified
+    \item The announced article has full structure (Abstract, Introduction, Theory, Experiments, Conclusion, Appendix)
+    \item If results are positive, it is written with NeurIPS template.
+\end{enumerate}
 
 \section{Preliminary plan}
+\paragraph{Week April 28 - May 4}
+\begin{itemize}
+    \item For Alexey: solve how to tune the algorithms for MLP and CNN, try formulating theory (and an appropriate model of the problem) why Muon and Neon are so successful, and create the drafts of the proofs.
+Register at NeurIPS site.
+
+    \item For Ivan: write the theory for an update from the algebra point of view (as for an article)
+
+    \item For Nikolay: write the theory for computing an update, and implement the method, if required
+
+    \item For Alexander: reproduce results of Jordan on NanoGPT and ResNet (CIFAR-10), learn to train both models with Neon.
+\end{itemize}
+
+\paragraph{Week May 5 - May 11}
+\begin{itemize}
+    \item For Alexey: finalize the proofs. Verify them via small experiments on MLP and CNN. Write with Alexander Experiments for the article.
+
+    \item For Ivan: join Nikolay to finalize algebra part of the article. Estimate FLOPS, memory and other overheads (produce O bounds)
+
+    \item For Nikolay: write a draft of the poster (before May 6), and work with Ivan
+
+    \item For Alexander: agressively test algorithms, prove that Neon outperforms competitors and prepare the results for the article.
+
+    \item For everybody: write and edit the article
+
+    \item May 11: submit an abstract to NeurIPS.
+
+    \item May 12-14: the article is being polished.
+
+    \item May 15: the article must be sent.
+\end{itemize}
 
 \section{Prototyping phase report}
+\begin{enumerate}
+    \item Update rule is derived, see idea
+    \item Update rule methods are tested: power iteration vs Lanzos (see playground)
+    \item NanoGPT is tested on Muon and Adam (add fig). The results are strange, because the methods do not converge at the expected rate (Muon is slower)
+    \item Neon, Muon, AdamW and SGD are compared on MLP and CNN (add fig). All methods work correctly, but again there is the problem with which one is the fastest (for now, it's SGD).
+\end{enumerate}
 

### 2025-04-26 article/muon.tex
fixed references
changed sentence structure in first paragraph of the section

@@
 The goal of the project is to make variations on Muon to speed it up.
-Recently, authors of \cite{bernstein2024oldoptimizernewnorm} have proposed to look at different optimizers as the solution of the optimization problem for an update.
+Recently, authors of \cite{bernstein2024oldoptimizernewnorm} have proposed a framework, which allows to describe update step of many optimizers as the solution of the certain optimization problem.
 This approach can be utilized to derive Muon \cite{jordan2024muon}, a novel algorithm for fast training of neural networks.
-Instead of using spectral norm in the problem, we use nuclear norm to produce a new problem.
+We change optimization problem defining update step of Muon slightly.
+Instead of using spectral norm in the optimization problem, we use nuclear norm to produce a new problem and a new algorithm.
 %Then, we add momentum to update.
@@
 \bibliographystyle{unsrt}
-\bibliography{muon}
+\bibliography{muon.bib}
 

### 2025-04-26 article/muon.tex
Prototype almost done

--- 2025-04-25 article/muon.tex
@@
 linkcolor = blue]{hyperref}	% нумерование страниц, ссылки!!!!ИМЕННО В ТАКОМ ПОРЯДКЕ СО СЛЕДУЮЩИМ ПАКЕТОМ
+\usepackage{natbib}  % Add natbib for better citation handling
 %\usepackage[warn]{mathtext}				% Поддержка русского текста в формулах
@@
 \section{Literature review}
-TO-DO.
+TO-DO
+
 \section{Quality metrics}
@@
     \item Update rule methods are tested: power iteration vs Lanzos (see playground)
-    \item NanoGPT is tested on Muon and Adam (add fig). The results are strange, because the methods do not converge at the expected rate (Muon is slower)
-    \item Neon, Muon, AdamW and SGD are compared on MLP and CNN (add fig). All methods work correctly, but again there is the problem with which one is the fastest (for now, it's SGD).
+    \item Recorded the distribution of singular values of gradients during NanoGPT training (add fig).
+    \item NanoGPT is tested on Muon and Adam (add fig). The results are strange, because the methods do not converge at the expected rate (Muon is slower). Besides, Neon doe not converge (add fig).
+    \item Neon, Muon, AdamW and SGD are compared on MLP and CNN (see Figures~\ref{fig:mlp_epochs}, \ref{fig:cnn_epochs}, \ref{fig:mlp_time}, and~\ref{fig:cnn_time}).
+All methods work correctly, but again there is the problem with which one is the fastest (for now, it's SGD).
+    \begin{figure}[h!]
+        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/accuracy_vs_epochs_simple.png}}
+        \caption{MLP: self.linear1 = nn.Linear(32*32*3, 512), self.linear2 = nn.Linear(512, 10), self.activ = nn.GELU()}
+        \label{fig:mlp_epochs}
+    \end{figure}
+    \begin{figure}[h!]
+        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/accuracy_vs_epochs_moderate.png}}
+        \caption{CNN: 2 convolutional blocks, 2 fully connected layers, activation + dropout}
+        \label{fig:cnn_epochs}
+    \end{figure}
+    \begin{figure}[h!]
+        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/accuracy_vs_time_simple.png}}
+        \caption{MLP: wallclock time measurements}
+        \label{fig:mlp_time}
+    \end{figure}
+    \begin{figure}[h!]
+        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/accuracy_vs_time_moderate.png}}
+        \caption{CNN: wallclock time measurements}
+        \label{fig:cnn_time}
+    \end{figure}
 \end{enumerate}
 
-\bibliographystyle{unsrt}
-\bibliography{muon}
+\bibliographystyle{plainnat}  % Change to plainnat style
+\bibliography{muon}  % Make sure to include the bibliography
 

### 2025-04-26 Nano_gpt_exp/config/finetune_tiniy_stories_Adam.py
Refactor

@@ -21,5 +21,5 @@ gradient_accumulation_steps = 32
 max_iters = 200
 
 # finetune at constant LR
-learning_rate = 3e-5
+learning_rate = 0.05
 decay_lr = False

### 2025-04-26 Nano_gpt_exp/config/finetune_tiniy_stories_Neon.py

@@ -1,0 +1,25 @@
+import time
+
+out_dir = "out-tiny_stories_Neon"
+eval_interval = 5
+eval_iters = 40
+wandb_log = False  # feel free to turn on
+wandb_project = "tiny_stories"
+wandb_run_name = "ft-" + str(time.time())
+
+dataset = "tiny_stories"
+init_from = "gpt2-large"  # this is the largest GPT-2 model
+
+# only save checkpoints if the validation loss improves
+always_save_checkpoint = False
+
+# the number of examples per iter:
+# 1 batch_size * 32 grad_accum * 1024 tokens = 32,768 tokens/iter
+# shakespeare has 301,966 tokens, so 1 epoch ~= 9.2 iters
+batch_size = 1
+gradient_accumulation_steps = 2
+max_iters = 200
+
+# finetune at constant LR
+learning_rate = 3e-05
+decay_lr = False

### 2025-04-26 Nano_gpt_exp/model_Neon.py

@@ -1,0 +1,423 @@
+"""
+Full definition of a GPT Language Model, all of it in this single file.
+References:
+1) the official GPT-2 TensorFlow implementation released by OpenAI:
+https://github.com/openai/gpt-2/blob/master/src/model.py
+2) huggingface/transformers PyTorch implementation:
+https://github.com/huggingface/transformers/blob/main/src/transformers/models/gpt2/modeling_gpt2.py
+"""
+
+import math
+import inspect
+from dataclasses import dataclass
+
+import torch
+import torch.nn as nn
+from torch.nn import functional as F
+import torch.distributed as dist
+import os
+from Neon import Neon
+
+
+class LayerNorm(nn.Module):
+    """LayerNorm but with an optional bias. PyTorch doesn't support simply bias=False"""
+
+    def __init__(self, ndim, bias):
+        super().__init__()
+        self.weight = nn.Parameter(torch.ones(ndim))
+        self.bias = nn.Parameter(torch.zeros(ndim)) if bias else None
+
+    def forward(self, input):
+        return F.layer_norm(input, self.weight.shape, self.weight, self.bias, 1e-5)
+
+
+class CausalSelfAttention(nn.Module):
+
+    def __init__(self, config):
+        super().__init__()
+        assert config.n_embd % config.n_head == 0
+        # key, query, value projections for all heads, but in a batch
+        self.c_attn = nn.Linear(config.n_embd, 3 * config.n_embd, bias=config.bias)
+        # output projection
+        self.c_proj = nn.Linear(config.n_embd, config.n_embd, bias=config.bias)
+        # regularization
+        self.attn_dropout = nn.Dropout(config.dropout)
+        self.resid_dropout = nn.Dropout(config.dropout)
+        self.n_head = config.n_head
+        self.n_embd = config.n_embd
+        self.dropout = config.dropout
+        # flash attention make GPU go brrrrr but support is only in PyTorch >= 2.0
+        self.flash = hasattr(torch.nn.functional, "scaled_dot_product_attention")
+        if not self.flash:
+            print(
+                "WARNING: using slow attention. Flash Attention requires PyTorch >= 2.0"
+            )
+            # causal mask to ensure that attention is only applied to the left in the input sequence
+            self.register_buffer(
+                "bias",
+                torch.tril(torch.ones(config.block_size, config.block_size)).view(
+                    1, 1, config.block_size, config.block_size
+                ),
+            )
+
+    def forward(self, x):
[…]

### 2025-04-26 Nano_gpt_exp/train_Neon.py

@@ -1,0 +1,409 @@
+"""
+This training script can be run both on a single gpu in debug mode,
+and also in a larger training run with distributed data parallel (ddp).
+
+To run on a single GPU, example:
+$ python train.py --batch_size=32 --compile=False
+
+To run with DDP on 4 gpus on 1 node, example:
+$ torchrun --standalone --nproc_per_node=4 train.py
+
+To run with DDP on 4 gpus across 2 nodes, example:
+- Run on the first (master) node with example IP 123.456.123.456:
+$ torchrun --nproc_per_node=8 --nnodes=2 --node_rank=0 --master_addr=123.456.123.456 --master_port=1234 train.py
+- Run on the worker node:
+$ torchrun --nproc_per_node=8 --nnodes=2 --node_rank=1 --master_addr=123.456.123.456 --master_port=1234 train.py
+(If your cluster does not have Infiniband interconnect prepend NCCL_IB_DISABLE=1)
+"""
+
+import os
+import time
+import math
+import pickle
+from contextlib import nullcontext
+
+import numpy as np
+import torch
+from torch.nn.parallel import DistributedDataParallel as DDP
+from torch.distributed import init_process_group, destroy_process_group
+import torch.distributed as dist
+
+from model_Neon import GPTConfig, GPT
+from torch.utils.tensorboard import SummaryWriter
+
+writer = SummaryWriter(log_dir="Neon")
+
+# -----------------------------------------------------------------------------
+# default config values designed to train a gpt2 (124M) on OpenWebText
+# I/O
+out_dir = "out"
+eval_interval = 2000
+log_interval = 1
+eval_iters = 200
+eval_only = False  # if True, script exits right after the first eval
+always_save_checkpoint = True  # if True, always save a checkpoint after each eval
+init_from = "scratch"  # 'scratch' or 'resume' or 'gpt2*'
+# wandb logging
+wandb_log = False  # disabled by default
+wandb_project = "owt"
+wandb_run_name = "gpt2"  # 'run' + str(time.time())
+# data
+dataset = "openwebtext"
+gradient_accumulation_steps = 5 * 8  # used to simulate larger batch sizes
+batch_size = 12  # if gradient_accumulation_steps > 1, this is the micro-batch size
+block_size = 1024
+# model
+n_layer = 12
+n_head = 12
+n_embd = 768
+dropout = 0.0  # for pretraining 0 is good, for finetuning try 0.1+
[…]

### 2025-04-26 article/muon.tex
Citation does not work
finished paragraphs about muon complexity and complexity comparison

--- 2025-04-25 article/muon.tex
@@
 By setting this norm to the RMS-to-RMS norm (a scaled version of the spectral norm), we recover the Muon optimizer \cite{bernstein2025deriving, bernstein2024oldoptimizernewnorm} with an update step defined by:
-\begin{equation}
+\begin{equation}\label{eqn:update_muon}
 \delta W = - \frac{1}{\lambda}\sqrt{\frac{n}{m}}UV^T\,,
@@
 \section{Literature review}
-TO-DO.
+Our review focuses mainly on Muon and Shampoo optimizers, since our algorith extends ideas used to derive those.
+We highlight advantages and disadvantages of those methods, unique effects they introduce, and compare them to Neon.
+
+In the previous section we described the theory behind weights update step in Muon optimizer, but we have not discussed how to obtain required matrices on practice.
+The update step is defined by~\eqref{eqn:update_muon}, which requires $UV^T$ with $U$ and $V^T$ from singular value decomposition of gradient-like matrix $G$.
+Naive solution would be to compute SVD of $G$, and construct required expression.
+However, developers of Muon optimizer developed a workaround which uses Newton-Shulz iterations~\cite{jordan2024muon}.
+Newton-Shulz iterations from original article~\cite{jordan2024muon} require $10$ matrix-matrix multiplication to achieve desired accuracy.
+Asimptotic complexity of such operation is identical to those of SVD and equals $O(nm \min\{m, n\})$ for $m \times n$ matrix, but matrix multiplication on modern GPUs can be performed much more efficiently.
+
+While uniqie advantages and effects introduced by Neon are yet to be discovered, we can already say that our new optimizer introduces additional overhead of $???$ FLOPS per $m \times n$ weight matrix, which is much better then $O(m n \min\{m, n\})$ for Muon optimizer and $O(n^3 + m^3)$ for Shampoo.
+
 \section{Quality metrics}

### 2025-04-26 article/muon.tex
wrote another couple paragraphs on Muon optimizer

@@
 Naive solution would be to compute SVD of $G$, and construct required expression.
-However, developers of Muon optimizer developed a workaround which uses Newton-Shulz iterations~\cite{jordan2024muon}.
-Newton-Shulz iterations from original article~\cite{jordan2024muon} require $10$ matrix-matrix multiplication to achieve desired accuracy.
-Asimptotic complexity of such operation is identical to those of SVD and equals $O(nm \min\{m, n\})$ for $m \times n$ matrix, but matrix multiplication on modern GPUs can be performed much more efficiently.
+However, developers of Muon optimizer came up with a workaround which uses Newton-Shulz iterations~\cite{jordan2024muon}.
+Newton-Shulz iterations from original article~\cite{jordan2024muon} require $10$ matrix-matrix multiplications to achieve desired accuracy.
+Asimptotic complexity of such operation is identical to those of SVD and equals $O(mn \min\{m, n\})$ for $m \times n$ matrix, but matrix multiplication on modern GPUs can be performed much more efficiently.
+
+The performance of Muon on training large language models was tested~\cite{liu2025muon} against those of AdamW.
+The testing demonstrated excellent performance of Muon~--- it was $\sim 2$ times more efficient in terms of FLOPs required to reach certain loss value.
+It is even more remarkable, if cost of one iteration is taken into account: Muon requires additional $O(mn \min\{m, n\})$ FLOPs per $m \times n$ matrix, while AdamW needs only $O(mn)$.
+
+Another interesting discovery about Muon optimizer is that it accelerates grokking\cite{tveit2025muonoptimizeracceleratesgrokking}.
+At the test problem Muon achieved grokking significantly faster then AdamW in terms of passed epochs, mean grokking epoch for Muon was $102.89$, while for AdamW it wes $153.09$.
+Authors suggest that this may be due to the fact that Muon stimulates broder exploration by orthogonalizing gradient matrix, and thus avoids memorization.
+
+Recently certain thoretical guarantees for Muon convergence have been derived~\cite{li2025noteconvergencemuon}. In particular, in $L$-smooth convex case it achieves $1/T^{\frac{1}{4}}$
 

### 2025-04-26 article/muon.tex
finished literature review

--- 2025-04-26 article/muon.tex
@@
 \section{Literature review}
-Our review focuses mainly on Muon and Shampoo optimizers, since our algorith extends ideas used to derive those.
-We highlight advantages and disadvantages of those methods, unique effects they introduce, and compare them to Neon.
+Our review primarily focuses on the Muon and Shampoo optimizers, as our algorithm extends the ideas used to derive these methods.
+We highlight the advantages and disadvantages of these approaches, the unique effects they introduce, and compare them to Neon.
 
-In the previous section we described the theory behind weights update step in Muon optimizer, but we have not discussed how to obtain required matrices on practice.
-The update step is defined by~\eqref{eqn:update_muon}, which requires $UV^T$ with $U$ and $V^T$ from singular value decomposition of gradient-like matrix $G$.
-Naive solution would be to compute SVD of $G$, and construct required expression.
-However, developers of Muon optimizer developed a workaround which uses Newton-Shulz iterations~\cite{jordan2024muon}.
-Newton-Shulz iterations from original article~\cite{jordan2024muon} require $10$ matrix-matrix multiplication to achieve desired accuracy.
-Asimptotic complexity of such operation is identical to those of SVD and equals $O(nm \min\{m, n\})$ for $m \times n$ matrix, but matrix multiplication on modern GPUs can be performed much more efficiently.
+\subsection{Muon optimizer}
 
-While uniqie advantages and effects introduced by Neon are yet to be discovered, we can already say that our new optimizer introduces additional overhead of $???$ FLOPS per $m \times n$ weight matrix, which is much better then $O(m n \min\{m, n\})$ for Muon optimizer and $O(n^3 + m^3)$ for Shampoo.
+In the previous section, we described the theoretical foundation behind the weight update step in the Muon optimizer, but we did not discuss how to obtain the required matrices in practice.
+The update step is defined by \eqref{eqn:update_muon}, which requires $UV^T$ with $U$ and $V^T$ from the singular value decomposition of the gradient-like matrix $G$.
+A naive solution would involve computing the SVD of $G$ and constructing the required expression.
+However, the developers of the Muon optimizer introduced a workaround using Newton-Schulz iterations \cite{jordan2024muon}.
+The Newton-Schulz iterations from the original article \cite{jordan2024muon} require 10 matrix-matrix multiplications to achieve the desired accuracy.
+The asymptotic complexity of such an operation is identical to that of SVD and equals $O(mn \min\{m, n\})$ for an $m \times n$ matrix.
+Nevertheless, matrix multiplication on modern GPUs can be performed much more efficiently.
+
+The performance of Muon in training large language models was tested \cite{liu2025muon} against AdamW.
+The testing demonstrated excellent performance by Muon, which was approximately 2 times more efficient in terms of FLOPs required to reach a certain loss value.
+This is even more remarkable when considering the cost of one iteration: Muon requires an additional $O(mn \min\{m, n\})$ FLOPs per $m \times n$ matrix, while AdamW needs only $O(mn)$.
+
+Another interesting discovery about the Muon optimizer is that it accelerates grokking \cite{tveit2025muonoptimizeracceleratesgrokking}.
+In the test problem, Muon achieved grokking significantly faster than AdamW in terms of passed epochs, with a mean grokking epoch of 102.89 for Muon and 153.09 for AdamW.
+The authors suggest that this may be due to the fact that Muon stimulates broader exploration by orthogonalizing the gradient matrix, thus avoiding memorization.
+
+Recently, theoretical guarantees for Muon convergence have been derived \cite{li2025noteconvergencemuon}.
+In particular, in the $L$-smooth convex case, it achieves $O(1/T^{\frac{1}{2}})$ (with full gradient) and $O(1/T^{\frac{1}{4}})$ (with stochastic gradient) bounds on the Frobenius norm of the gradient or the mathematical expectation of the gradient norm, respectively, where $T$ is the number of iterations.
+
+\subsection{Shampoo optimizer}
+
+Another optimizer that exploits the matrix (and even tensor) structure of weights in neural networks is the Shampoo optimizer.
+We avoid a detailed description of this method here and refer the reader to the original article \cite{gupta2018shampoopreconditionedstochastictensor}, but we outline the key properties of Shampoo and its relation to the Muon optimizer.
+
+The Shampoo optimizer uses left and right preconditioning for the gradient-like matrix, leveling its spectrum.
+The preconditioners are computed from the exponentially averaged gradients, and their computation requires $O(n^3 + m^3)$ per $m \times n$ matrix.
+This exponential averaging is a key feature that provides several distinct interpretations for the preconditioners.
+They can be viewed as an approximation of the Gauss-Newton component of the Hessian or the first step of the power iteration algorithm for computing the optimal Kronecker product approximation \cite{morwani2025a}.
+With exponential averaging turned off, the update step of Shampoo can be simplified and becomes identical to that of Muon \cite{jordan2024muon}.
+
+Convergence analysis is presented in the original article \cite{gupta2018shampoopreconditionedstochastictensor}.
+Shampoo achieves $O(1/T^{\frac{1}{2}})$ convergence for the loss function value in the $L$-smooth convex case.
+
+\subsection{Synthesis and Neon's position}
+
+Recent developments in optimization techniques show that utilizing the matrix structure of weights in neural networks can be very beneficial.
+Optimizers following this path converge faster in terms of iterations or epochs and often even FLOPs, but have a high iteration cost.
+Neon seeks to decrease the iteration cost while preserving fast convergence.
+
+While the uniqie advantages and effects introduced by Neon are yet to be discovered, we can already say that our new optimizer introduces additional overhead of $???$[Depends on variant of Neon we choose] FLOPS per $m \times n$ weight matrix, which is much better then $O(m n \min\{m, n\})$ for Muon optimizer and $O(n^3 + m^3)$ for Shampoo.
 

### 2025-04-27 article/muon.tex

@@
     \item Update rule methods are tested: power iteration vs Lanzos (see playground)
-    \item NanoGPT is tested on Muon and Adam (add fig). The results are strange, because the methods do not converge at the expected rate (Muon is slower)
-    \item Neon, Muon, AdamW and SGD are compared on MLP and CNN (add fig). All methods work correctly, but again there is the problem with which one is the fastest (for now, it's SGD).
+    \item Recorded the distribution of singular values of gradients during NanoGPT training (add fig).
+    \item NanoGPT is tested on Muon and Adam (add fig). The results are strange, because the methods do not converge at the expected rate (Muon is slower). Besides, Neon doe not converge (add fig).
+    \item Neon, Muon, AdamW and SGD are compared on MLP and CNN (see Figures~\ref{fig:mlp_epochs}, \ref{fig:cnn_epochs}, \ref{fig:mlp_time}, and~\ref{fig:cnn_time}).
+All methods work correctly, but again there is the problem with which one is the fastest (for now, it's SGD).
+    \begin{figure}[h!]
+        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/accuracy_vs_epochs_simple.png}}
+        \caption{MLP: self.linear1 = nn.Linear(32*32*3, 512), self.linear2 = nn.Linear(512, 10), self.activ = nn.GELU()}
+        \label{fig:mlp_epochs}
+    \end{figure}
+    \begin{figure}[h!]
+        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/accuracy_vs_epochs_moderate.png}}
+        \caption{CNN: 2 convolutional blocks, 2 fully connected layers, activation + dropout}
+        \label{fig:cnn_epochs}
+    \end{figure}
+    \begin{figure}[h!]
+        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/accuracy_vs_time_simple.png}}
+        \caption{MLP: wallclock time measurements}
+        \label{fig:mlp_time}
+    \end{figure}
+    \begin{figure}[h!]
+        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/accuracy_vs_time_moderate.png}}
+        \caption{CNN: wallclock time measurements}
+        \label{fig:cnn_time}
+    \end{figure}
 \end{enumerate}

### 2025-04-27 article/muon.tex
added missing minus
modified paper a little, to make it obvious that we still need to figure out which version of Neon we accept

@@
 \begin{equation}\label{eqn:update_star}
-  \delta W = \frac{1}{\lambda} u_1 \sigma_1 v_1^T\,,
+  \delta W = -\frac{1}{\lambda} u_1 \sigma_1 v_1^T\,,
 \end{equation}
@@
 \begin{equation}\label{eqn:update_F_star}
-\delta W = \frac{1}{\lambda}UDV^T
+\delta W = -\frac{1}{\lambda}UDV^T
 \end{equation} 
@@
 
-In this article we show how one can quickly compute weight updates defined by \eqref{eqn:update_star} and \eqref{eqn:update_F_star}.
-Then we finilize the methods by adding momentum and test their performance against those of Muon ant training multilayer perceptron and transformer.
-The results will be fast algorithms, which we will convert into a new optimizer classes for PyTorch, as was done with Muon.
+In this article we show how one can quickly compute weight updates defined by \eqref{eqn:update_star} or \eqref{eqn:update_F_star} (for now we need more tests to accept one variant or both of them).
+Then we finilize the method by adding momentum and test their performance against those of Muon at training multilayer perceptron and transformer.
+The results will be fast algorithm, which we will convert into a new optimizer classes for PyTorch, as was done with Muon.
 
@@
     \item Numerical experiments comparing performance of training with Neon, Muon, SVD and Adam on several architectures: MLP, CNN, and NanoGPT.
-    \item Participation in CIFAR-10 challenge with Neon (if the alrothims turns out to be competitive)
+    \item Participation in CIFAR-10 challenge with Neon (if the alrothim turns out to be competitive)
     \item Article describing all the results. If they are satisfactory, we intend to be in time to apply for NeurIPS 2025. It is a challenge, but I think we must give it a try.
@@
 
-While the uniqie advantages and effects introduced by Neon are yet to be discovered, we can already say that our new optimizer introduces additional overhead of $???$[Depends on variant of Neon we choose] FLOPS per $m \times n$ weight matrix, which is much better then $O(m n \min\{m, n\})$ for Muon optimizer and $O(n^3 + m^3)$ for Shampoo.
+While the uniqie advantages and effects introduced by Neon are yet to be discovered, we can already say that our new optimizer introduces additional overhead of $O(mn)$ FLOPS on average per $m \times n$ weight matrix, which is much better then $O(m n \min\{m, n\})$ for Muon optimizer and $O(n^3 + m^3)$ for Shampoo.
 
@@
     \item Recorded the distribution of singular values of gradients during NanoGPT training (add fig).
-    \item NanoGPT is tested on Muon and Adam (add fig). The results are strange, because the methods do not converge at the expected rate (Muon is slower). Besides, Neon doe not converge (add fig).
-    \item Neon, Muon, AdamW and SGD are compared on MLP and CNN (see Figures~\ref{fig:mlp_epochs}, \ref{fig:cnn_epochs}, \ref{fig:mlp_time}, and~\ref{fig:cnn_time}).
+    \item NanoGPT is tested on Muon and Adam (add fig).
+The results are strange, because the methods do not converge at the expected rate (Muon is slower).
+Besides, Neon (rank-1 version) does not converge (add fig).
+    \item Neon (rank-1 version), Muon, AdamW and SGD are compared on MLP and CNN (see Figures~\ref{fig:mlp_epochs}, \ref{fig:cnn_epochs}, \ref{fig:mlp_time}, and~\ref{fig:cnn_time}).
 All methods work correctly, but again there is the problem with which one is the fastest (for now, it's SGD).

### 2025-04-27 article/muon.tex
It's Lanczos, not Lanzos

@@
 Changing spectral norm, which was used to derive Muon, to nuclear norm, we pose a new optimization problem for an update matrix, solution of which defines a novel algorithm we name Neon.
-To make it feasible, we use Lanzos algorithm to find a required step.
+To make it feasible, we use Lanczos algorithm to find a required step.
 After providing theoretical guarantees of Neon convergence, we compare performances of Neon, Muon, and Adam on training multilayer perceptron, convolutional neural network and NanoGPT.
@@
     \item Update rule is derived, see idea
-    \item Update rule methods are tested: power iteration vs Lanzos (see playground)
+    \item Update rule methods are tested: power iteration vs Lanczos (see playground)
     \item Recorded the distribution of singular values of gradients during NanoGPT training (add fig).

### 2025-04-27 article/muon.tex
fixed small inconvinience in Shampoo optimizer subsection
added testing F* norm version of neon to plans

@@
 Convergence analysis is presented in the original article \cite{gupta2018shampoopreconditionedstochastictensor}.
-Shampoo achieves $O(1/T^{\frac{1}{2}})$ convergence for the loss function value in the $L$-smooth convex case.
+Shampoo achieves $O(1/T^{\frac{1}{2}})$ convergence for the loss function value in the $L$-smooth convex case, where $T$ is the number of iterations.
 
@@
 Register at NeurIPS site.
+Test version 2 of Neon (with $F*$ norm)
 
@@
 
-    \item For Alexander: reproduce results of Jordan on NanoGPT and ResNet (CIFAR-10), learn to train both models with Neon.
+    \item For Alexander: reproduce results of Jordan on NanoGPT and ResNet (CIFAR-10), learn to train both models with Neon. Test version 2 of Neon (with $F*$ norm)
 \end{itemize}
@@
 The results are strange, because the methods do not converge at the expected rate (Muon is slower).
-Besides, Neon (rank-1 version) does not converge (add fig).
-    \item Neon (rank-1 version), Muon, AdamW and SGD are compared on MLP and CNN (see Figures~\ref{fig:mlp_epochs}, \ref{fig:cnn_epochs}, \ref{fig:mlp_time}, and~\ref{fig:cnn_time}).
+Besides, Neon (kernel norm version) does not converge (add fig).
+    \item Neon (kernel norm version), Muon, AdamW and SGD are compared on MLP and CNN (see Figures~\ref{fig:mlp_epochs}, \ref{fig:cnn_epochs}, \ref{fig:mlp_time}, and~\ref{fig:cnn_time}).
 All methods work correctly, but again there is the problem with which one is the fastest (for now, it's SGD).

### 2025-04-27 article/muon.tex
Added svd and nano gpt

--- 2025-04-27 article/muon.tex
@@
 
+\subsection{Optimization Strategies for Efficient Large Language Model Training}
+The unprecedented scale of modern Large Language Models (LLMs) has pushed traditional optimizers like AdamW \cite{Loshchilov2017FixingWD} to their limits in terms of computational efficiency and convergence speed \cite{liu2025muon,chen2025cosmoshybridadaptive}.
+This challenge has catalyzed research into more sophisticated optimization approaches that can maintain or improve performance while reducing training costs.
+
+The Muon optimizer  has emerged as a promising alternative based on matrix orthogonalization principles.
+Scaling Muon to billion-parameter LLMs required two key adaptations: the integration of L2 weight decay for stability and the implementation of per-parameter update scaling to handle diverse parameter distributions efficiently .
+Empirical evaluations demonstrate that Muon can match or exceed AdamW's model quality while requiring only about 52 of the training FLOPs.
+The successful training of the Moonlight model series, including a 16B-parameter Mixture-of-Experts model, validates Muon's practicality for production-scale applications \cite{liu2025muon}.
+
+Building on this foundation, hybrid approaches like COSMOS \cite{chen2025cosmoshybridadaptive} further enhance efficiency by combining optimization techniques based on gradient structure.
+COSMOS applies computationally intensive updates to a low-dimensional "leading eigensubspace" while using memory-efficient methods like Muon for the remaining parameters.
+This approach maintains convergence benefits while substantially reducing memory requirements.
+For distributed training environments, optimizers like Dion \cite{ahn2025dioncommunicationefficientoptimizerlarge} specifically target communication efficiency by minimizing data exchange between workers through distributed orthonormalization techniques.
 \section{Quality metrics}
@@
     \item Update rule methods are tested: power iteration vs Lanczos (see playground)
-    \item Recorded the distribution of singular values of gradients during NanoGPT training (add fig).
+    \item Recorded the distribution of singular values of gradients during NanoGPT training (see Firgures ~\ref{fig:svd_all}, \ref{fig:svd_all}).
+    \begin{figure}[h!]
+        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/svd_evolution_all.png}}
+        \caption{Singular values of random 100 parameters via 200 iterations}
+        \label{fig:svd_all}
+    \end{figure}
+    \begin{figure}[h!]
+        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/svd_evolution.png}}
+        \caption{Singular values of random 100 parameters for 5-th,45-th, 65-th,175-th and 200-th iteration}
+        \label{fig:svd}
+    \end{figure}
+
     \item NanoGPT is tested on Muon and Adam (add fig).
 The results are strange, because the methods do not converge at the expected rate (Muon is slower).
-Besides, Neon (rank-1 version) does not converge (add fig).
+Besides, Neon (rank-1 version) does not converge (see Figures ~\ref{fig:train_loss} \ref{fig:val_loss})
+    \begin{figure}[h!]
+        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/loss_train.png}}
+        \caption{Train loss}
+        \label{fig:train_loss}
+    \end{figure}
+    \begin{figure}[h!]
+        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/loss_val.png}}
+        \caption{Validation loss}
+        \label{fig:val_loss}
+    \end{figure}
+
+    The pictures show the best results achieved so far. The experiments were conducted with two 4090 24GB GPUs for nanotgpt-large on the tiny stories dataset
+
     \item Neon (rank-1 version), Muon, AdamW and SGD are compared on MLP and CNN (see Figures~\ref{fig:mlp_epochs}, \ref{fig:cnn_epochs}, \ref{fig:mlp_time}, and~\ref{fig:cnn_time}).
@@
     \end{figure}
-\end{enumerate}
+\end{enumerate}\\
 

### 2025-04-27 article/muon.tex
Updated figures

@@
     \begin{figure}[h!]
-        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/svd_evolution_all.png}}
-        \caption{Singular values of random 100 parameters via 200 iterations}
+        \center{\includegraphics[width=0.8\linewidth]{figs/mlp24/svd_evolution_all.png}}
+        \caption{Singular values of $50257\times 1280$ layer via 200 iterations}
         \label{fig:svd_all}
@@
     \begin{figure}[h!]
-        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/svd_evolution.png}}
-        \caption{Singular values of random 100 parameters for 5-th,45-th, 65-th,175-th and 200-th iteration}
+        \center{\includegraphics[width=0.8\linewidth]{figs/mlp24/svd_evolution.png}}
+        \caption{Singular values of  of $50257\times 1280$ layer for 5-th,45-th, 65-th,175-th and 200-th iteration}
         \label{fig:svd}
@@
     \begin{figure}[h!]
-        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/loss_train.png}}
+        \center{\includegraphics[width=0.8\linewidth]{figs/mlp24/loss_train.png}}
         \caption{Train loss}
@@
     \begin{figure}[h!]
-        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/loss_val.png}}
+        \center{\includegraphics[width=0.8\linewidth]{figs/mlp24/loss_val.png}}
         \caption{Validation loss}
@@
     \begin{figure}[h!]
-        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/accuracy_vs_epochs_simple.png}}
+        \center{\includegraphics[width=0.8\linewidth]{figs/mlp24/accuracy_vs_epochs_simple.png}}
         \caption{MLP: self.linear1 = nn.Linear(32*32*3, 512), self.linear2 = nn.Linear(512, 10), self.activ = nn.GELU()}
@@
     \begin{figure}[h!]
-        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/accuracy_vs_epochs_moderate.png}}
+        \center{\includegraphics[width=0.8\linewidth]{figs/mlp24/accuracy_vs_epochs_moderate.png}}
         \caption{CNN: 2 convolutional blocks, 2 fully connected layers, activation + dropout}
@@
     \begin{figure}[h!]
-        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/accuracy_vs_time_simple.png}}
+        \center{\includegraphics[width=0.8\linewidth]{figs/mlp24/accuracy_vs_time_simple.png}}
         \caption{MLP: wallclock time measurements}
@@
     \begin{figure}[h!]
-        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/accuracy_vs_time_moderate.png}}
+        \center{\includegraphics[width=0.8\linewidth]{figs/mlp24/accuracy_vs_time_moderate.png}}
         \caption{CNN: wallclock time measurements}

### 2025-04-27 article/muon.tex
k-rank updates comparison

--- 2025-04-27 article/muon.tex
@@
 
+\begin{table}[h!]
+    \centering
+    \begin{tabular}{c|c|c|c}
+    \hline
+    Method                              & rtol        & k & time,s \\
+    \hline 
+    Power Iterations                    & 0.01        & 1 & 7.7\\ 
+    SVDS (thick-restart Lanczos method) & 0.01        & 1 & 0.18\\
+    PCA Low Rank (RSVD)                 & 0.01        & 1 & 1.15\\
+    SVDS (thick-restart Lanczos method) & 0.01        & 10 & 0.47\\
+    PCA Low Rank (RSVD)                 & 0.01        & 10 & 19.4\\
+    SVDS (thick-restart Lanczos method) & 0.01        & 100 & 1.96\\
+    PCA Low Rank (RSVD)                 & 0.01        & 100 & 170\\
+    \end{tabular}
+    \caption{k-rank updated comparison}
+    Comparison of different numerical methods to calculate k-rank update on $5000\times5000$ matrix of real numbers, rtol is an error in Frobenius norm relative to the k-rank approximation of truncated svd.
+During the research it was noted that rsvd can give good and fast approximation for singular values, but the matrix of approximation is far from the one given by truncated svd, while Lanczos method gives good and fast approximation for a matrix, but not so good approximation for singular values.
+    \label{tab:my_label}
+\end{table}
+
 \bibliographystyle{unsrt}

### 2025-04-27 article/muon.tex
table of k-rank updates added, sashas figures paths fixed

--- 2025-04-27 article/muon.tex
@@
         \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/svd_evolution_all.png}}
-        \caption{Singular values of random 100 parameters via 200 iterations}
+        \caption{Singular values of $50257\times 1280$ layer via 200 iterations}
         \label{fig:svd_all}
@@
         \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/svd_evolution.png}}
-        \caption{Singular values of random 100 parameters for 5-th,45-th, 65-th,175-th and 200-th iteration}
+        \caption{Singular values of  of $50257\times 1280$ layer for 5-th,45-th, 65-th,175-th and 200-th iteration}
         \label{fig:svd}
@@
     \end{figure}
-\end{enumerate}\\
+\end{enumerate}
+
+\begin{table}[h!]
+    \centering
+    \begin{tabular}{c|c|c|c}
+    \hline
+    Method                              & rtol        & k & time,s \\
+    \hline 
+    Power Iterations                    & 0.01        & 1 & 7.7\\ 
+    SVDS (thick-restart Lanczos method) & 0.01        & 1 & 0.18\\
+    PCA Low Rank (RSVD)                 & 0.01        & 1 & 1.15\\
+    SVDS (thick-restart Lanczos method) & 0.01        & 10 & 0.47\\
+    PCA Low Rank (RSVD)                 & 0.01        & 10 & 19.4\\
+    SVDS (thick-restart Lanczos method) & 0.01        & 100 & 1.96\\
+    PCA Low Rank (RSVD)                 & 0.01        & 100 & 170\\
+    \end{tabular}
+    \caption{k-rank updated comparison}
+    Comparison of different numerical methods to calculate k-rank update on $5000\times5000$ matrix of real numbers, rtol is an error in Frobenius norm relative to the k-rank approximation of truncated svd.
+During the research it was noted that rsvd can give good and fast approximation for singular values, but the matrix of approximation is far from the one given by truncated svd, while Lanczos method gives good and fast approximation for a matrix, but not so good approximation for singular values.
+    \label{tab:my_label}
+\end{table}
 

### 2025-04-28 article/muon.tex
Rewritten with biber and compiled with the table

@@
 \usepackage{algorithmic} 
+\usepackage[backend=biber]{biblatex}
+\addbibresource{muon.bib}
 
@@
     \item Update rule is derived, see idea
-    \item Update rule methods are tested: power iteration vs Lanczos (see playground)
+    \item Update rule methods are tested: power iteration vs Lanczos (see \ref{tab:matrix_methods})
     \item Recorded the distribution of singular values of gradients during NanoGPT training (see Firgures ~\ref{fig:svd_all}, \ref{fig:svd_all}).
@@
 
-    \item NanoGPT is tested on Muon and Adam (add fig).
-The results are strange, because the methods do not converge at the expected rate (Muon is slower).
-Besides, Neon (rank-1 version) does not converge (see Figures ~\ref{fig:train_loss} \ref{fig:val_loss})
+    \item NanoGPT is tested on Muon and Adam. For now, Neon (rank-1 version) does not converge (see Figures ~\ref{fig:train_loss} \ref{fig:val_loss})
     \begin{figure}[h!]
@@
 
-    The pictures show the best results achieved so far. The experiments were conducted with two 4090 24GB GPUs for nanotgpt-large on the tiny stories dataset
+    The pictures show the best results achieved so far. The experiments were conducted with two 4090 24GB GPUs for nanotgpt-large on the tiny stories dataset.
 
@@
     \end{figure}
-\end{enumerate}
-
     \begin{table}[h!]
@@
 During the research it was noted that rsvd can give good and fast approximation for singular values, but the matrix of approximation is far from the one given by truncated svd, while Lanczos method gives good and fast approximation for a matrix, but not so good approximation for singular values.
-    \label{tab:my_label}
+        \label{tab:matrix_methods}
     \end{table}
+\end{enumerate}
 
-\bibliographystyle{unsrt}
-\bibliography{muon.bib}
+\printbibliography
 

### 2025-05-05 poster/poster.tex
add poster & add synctex to gitignore

% Template file for an a0 landscape poster.
% Written by Graeme, 2001-03 based on Norman's original microlensing
% poster.
%
% See discussion and documentation at
% <http://www.astro.gla.ac.uk/users/norman/docs/posters/> 
%
% $Id: poster-template-landscape.tex,v 1.2 2002/12/03 11:25:46 norman Exp $

% Default mode is landscape, which is what we want, however dvips and
% a0poster do not quite do the right thing, so we end up with text in
% landscape style (wide and short) down a portrait page (narrow and
% long). Printing this onto the a0 printer chops the right hand edge.
% However, 'psnup' can save the day, reorienting the text so that the
% poster prints lengthways down an a0 portrait bounding box.
%
% 'psnup -w85cm -h119cm -f poster_from_dvips.ps poster_in_landscape.ps'

\documentclass[a0]{a0poster}
% You might find the 'draft' option to a0 poster useful if you have
% lots of graphics, because they can take some time to process and
% display. (\documentclass[a0,draft]{a0poster})
\input defs
\pagestyle{empty}
\setcounter{secnumdepth}{0}
\renewcommand{\familydefault}{\sfdefault}
\newcommand{\QED}{~~\rule[-1pt]{8pt}{8pt}}\def\qed{\QED}

\renewcommand{\reals}{{\mbox{\bf R}}}

% The textpos package is necessary to position textblocks at arbitary 
% places on the page.
\usepackage[absolute]{textpos}

\usepackage{fleqn,psfrag,wrapfig,tikz}

\usepackage[papersize={38in,28in}]{geometry}

% Graphics to include graphics. Times is nice on posters, but you
% might want to switch it off and go for CMR fonts.
\usepackage{graphics}

% we are running pdflatex, so convert .eps files to .pdf
%\usepackage[pdftex]{graphicx}
%\usepackage{epstopdf}

% These colours are tried and tested for titles and headers. Don't
% over use color!
\usepackage{color}
\definecolor{Red}{rgb}{0.9,0.0,0.1}

\definecolor{bluegray}{rgb}{0.15,0.20,0.40}
\definecolor{bluegraylight}{rgb}{0.35,0.40,0.60}
\definecolor{gray}{rgb}{0.3,0.3,0.3}
\definecolor{lightgray}{rgb}{0.7,0.7,0.7}
\definecolor{darkblue}{rgb}{0.2,0.2,1.0}
\definecolor{darkgreen}{rgb}{0.0,0.5,0.3}

\renewcommand{\labelitemi}{\textcolor{bluegray}\textbullet}
\renewcommand{\labelitemii}{\textcolor{bluegray}{--}}

\setlength{\labelsep}{0.5em}

% see documentation for a0poster class for the size options here
\let\Textsize\normalsize
%\def\Head#1{\noindent\hbox to \hsize{\hfil{\LARGE\color{bluegray} #1}}\bigskip}
\def\Head#1{\noindent{\LARGE\color{bluegray} #1}\bigskip}
\def\LHead#1{\noindent{\LARGE\color{bluegray} #1}\bigskip}
\def\Subhead#1{\noindent{\large\color{bluegray} #1}\bigskip}
\def\Title#1{\noindent{\VeryHuge\color{Red} #1}}

% Set up the grid
%
% Note that [40mm,40mm] is the margin round the edge of the page --
% it is _not_ the grid size. That is always defined as 
% PAGE_WIDTH/HGRID and PAGE_HEIGHT/VGRID. In this case we use
% 23 x 12. This gives us three columns of width 7 boxes, with a gap of
% width 1 in between them. 12 vertical boxes is a good number to work
% with.
%
% Note however that texblocks can be positioned fractionally as well,
% so really any convenient grid size can be used.
%
\TPGrid[40mm,40mm]{23}{12}      % 3 cols of width 7, plus 2 gaps width 1

\parindent=0pt
\parskip=0.2\baselineskip

\begin{document}

% Understanding textblocks is the key to being able to do a poster in
% LaTeX. In
%
%    \begin{textblock}{wid}(x,y)
%    ...
%    \end{textblock}
%
% the first argument gives the block width in units of the grid
% cells specified above in \TPGrid; the second gives the (x,y)
% position on the grid, with the y axis pointing down.

% You will have to do a lot of previewing to get everything in the 
% right place.

% This gives good title positioning for a portrait poster.
% Watch out for hyphenation in titles - LaTeX will do it
% but it looks awful.
\begin{textblock}{23}(0,0)
\Title{The Title of the Project}
\end{textblock}

\begin{textblock}{23}(0,0.6)
{
\LARGE
Firstname Lastname
}

{
\Large
\color{bluegray}
\emph{Optimization Class Project. MIPT}
}
\end{textblock}

% Uni logo in the top right corner. A&A in the bottom left. Gives a
% good visual balance, but you may want to change this depending upon
% the graphics that are in your poster.
%\begin{textblock}{2}(0,10)
%Your logo here
%%\includegraphics{/usr/local/share/images/AandA.epsf}
%\end{textblock}

%\begin{textblock}{2}(21.2,0)
%Another logo here
%%\resizebox{2\TPHorizModule}{!}{\includegraphics{/usr/local/share/images/GUVIu/GUVIu.eps}}
%\end{textblock}

\begin{textblock}{7.0}(0,1.5)

\hrule\medskip
\Head{Introduction}\\
Many approaches to solving regression problems with hierarchical structures 
(\eg, Yuan \& Lin; Meier, van de Geer, B\"uhlmann; Jacob, Obozinski, Vert)
Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\medskip
\hrule\medskip
\Head{Group lasso}\\
One approach is the \emph{group lasso}:
\[
\begin{array}{ll}
\mbox{minimize} & f(x) + \lambda \sum_{i=1}^N \|x_i\|_2
\end{array}
\]
\ie, like lasso, but require groups of variables to be zero or not
\begin{itemize}\itemsep=12pt
\item also called $\ell_{1,2}$ mixed norm regularization
\end{itemize}

\medskip
\hrule\medskip
\Head{Structured group lasso}\\
Another approach is the \emph{structured group lasso}:
\[
\begin{array}{ll}
\mbox{minimize} & f(x) + \sum_{i=1}^N \lambda_i \|x_{g_i}\|_2
\end{array}
\]
where $g_i \subseteq [n]$ and $\mathcal G = \{g_1, \dots, g_N\}$
\begin{itemize}\itemsep=12pt
\item like group lasso, but the groups can overlap arbitrarily
\item particular choices of groups can impose `structured' sparsity
\item \eg, topic models, selecting interaction terms for (graphical) models,
    tree structure of gene networks, fMRI data
\item generalizes to the \textbf{composite absolute penalties family}:
\[
r(x) = \|(\|x_{g_1}\|_{p_1}, \dots, \|x_{g_N}\|_{p_N})\|_{p_0}
\]
\end{itemize}

\medskip
\hrule\medskip
\Head{Hierarchical selection}\\
\begin{center}
\begin{tikzpicture}
[dot/.style={rectangle,draw=black,fill=white,inner sep=5pt,minimum size=5pt}]
\node[dot,draw=orange,thick] at (0,10) (n1) {1};
\node[dot] at (-3,8) (n2) {2};
\node[dot,draw=orange,thick] at (3,8) (n3) {3};
\node[dot] at (-3,6) (n4) {4};
\node[dot,draw=orange,thick] at (1.5,6) (n5) {5};
\node[dot] at (4.5,6) (n6) {6};
\draw[->] (n1) -- (n2);
\draw[->] (n1) -- (n3);
\draw[->] (n2) -- (n4);
\draw[->] (n3) -- (n5);
\draw[->] (n3) -- (n6);
\end{tikzpicture}
\end{center}

\begin{itemize}\itemsep=8pt
    \item $\mathcal G = \{ \{4\}, \textcolor{orange}{\{5\}}, \{6\}, \{2,4\}, 
        \textcolor{orange}{\{3,5,6\}}, \textcolor{orange}{\{1,2,3,4,5,6\} \}}$
\item nonzero variables form a rooted and connected subtree
    \begin{itemize}
        \item if node is selected, so are its ancestors
        \item if node is not selected, neither are its descendants
    \end{itemize}
\end{itemize}

\end{textblock}

\begin{textblock}{7.0}(8,1.5)
\hrule\medskip
\Head{Algorithm}\\
We solve this problem using an ADMM lasso implementation:
\begin{verbatim}
prox_f = @(v,rho) (rho/(1 + rho))*(v - b) + b;
prox_g = @(v,rho) (max(0, v - 1/rho) - max(0, -v - 1/rho));

AA = A*A';
L  = chol(eye(m) + AA);

for iter = 1:MAX_ITER
    xx = prox_g(xz - xt, rho);
    yx = prox_f(yz - yt, rho);

    yz = L \ (L' \ (A*(xx + xt) + AA*(yx + yt)));
    xz = xx + xt + A'*(yx + yt - yz);

    xt = xt + xx - xz;
    yt = yt + yx - yz;
end
\end{verbatim}

\medskip
\hrule\medskip
\Head{Line search}\\
If $L$ is not known (usually the case), can use the following line search:

\noindent\rule[-5pt]{.8\textwidth}{0.4pt}
{\footnotesize
\begin{tabbing}
    {\bf given} $x^k$, $\lambda^{k-1}$, and parameter $\beta \in (0,1)$. \\*[\smallskipamount]
    Let $\lambda := \lambda^{k-1}$. \\*[\smallskipamount]
    {\bf repeat} \\
    \qquad \= 1.\ Let $z := \prox_{\lambda g}(x^{k} - \lambda \nabla f(x^{k}))$. \\
    \> 2.\ {\bf break if} $f(z) \leq \hat{f}_{\lambda}(z, x^{k})$. \\
    \> 3.\ Update $\lambda := \beta \lambda$. \\*[\smallskipamount]
    {\bf return} $\lambda^{k} := \lambda$, $x^{k+1}:=z$.
\end{tabbing}}
\noindent\rule[10pt]{.8\textwidth}{0.4pt}

typical value of $\beta$ is $1/2$, and 
\[
\hat{f}_\lambda(x,y) = f(y) + \nabla f(y)^T (x - y) + 
(1/2\lambda)\|x - y\|_2^2
\]

\medskip
\hrule\medskip
\Head{Convergence proof}\\
Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\end{textblock}

\begin{textblock}{7.0}(16,1.5)

\hrule\medskip
\Head{Numerical example}\\
Consider a numerical example with $f(x) = \|Ax - b\|_2^2$
with $A \in \reals^{10 \times 100}$ and $b \in \reals^{10}$.
Entries of $A$ and $b$ are generated as independent samples from
a standard normal distribution.
Here, we have chosen $\lambda$ using cross validation.

\medskip
\hrule\medskip
\Head{Results}\\
On this numerical example, the ADMM method converges quickly.
We give two realizations corresponding to different parameters $A$ and $b$.
\begin{center}
\psfrag{k}[t][b]{$k$}
\psfrag{fbest - fmin}[b][t]{$f_\mathrm{best}^{(k)} - f^\star$}
\psfrag{noise-free realize}{noise-free case}
\psfrag{realize1}{realization 1}
\psfrag{realize2}{realization 2}
\includegraphics[width=0.9\textwidth]{figures/pwl_error_fbest_realize.eps}
\end{center}

\medskip
\hrule\medskip
\Head{Conclusion}\\
Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\medskip
\hrule\medskip
\Head{Acknowledgements}\\
This material is based upon work supported by the
X Fellowship and my mom.

\end{textblock}

\end{document}

### 2025-05-05 poster/poster.tex
wrote 2 sections in poster

@@
 \usepackage{fleqn,psfrag,wrapfig,tikz}
+\usepackage{enumitem}
+\usepackage{amsmath}
 
@@
 \begin{textblock}{23}(0,0)
-\Title{The Title of the Project}
+\Title{Neon: Nuclear Norm to Beat Muon}
 \end{textblock}
@@
 \LARGE
-Firstname Lastname
+Alexey Kravatskiy, Ivan Kozyrev, Nikolay Kozlov, Alexander Vinogradov
 }
@@
 \Head{Introduction}\\
-Many approaches to solving regression problems with hierarchical structures 
-(\eg, Yuan \& Lin; Meier, van de Geer, B\"uhlmann; Jacob, Obozinski, Vert)
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+Recent advances in optimization techniques have highlighted the benefits of leveraging the matrix structure of neural network weights during training.
+Optimizers like Muon (Jordan et al.) and Shampoo (Gupta, Koren and Singer) have shown promise, but at a significantly higher computational cost than traditional methods like Adam.
+To bridge this gap, we propose Neon, a new optimizer that builds upon the framework of Bernstein and Newhouse.
+By using alternative norms, such as kernel norm (Kernel-Neon) or custom $F*$ norm ($F*$-Neon), we induce low-rank update matrices that enable more efficient computation.
+We evaluate the performance of Neon, Muon, and Adam on multilayer perceptrons, CIFAR10, and NanoGPT, and demonstrate [resting results here].
 
@@
 \hrule\medskip
-\Head{Group lasso}\\
-One approach is the \emph{group lasso}:
-\[
-\begin{array}{ll}
-\mbox{minimize} & f(x) + \lambda \sum_{i=1}^N \|x_i\|_2
-\end{array}
-\]
-\ie, like lasso, but require groups of variables to be zero or not
-\begin{itemize}\itemsep=12pt
-\item also called $\ell_{1,2}$ mixed norm regularization
-\end{itemize}
+\Head{Neon's update rule}\\
+Bernstein and Newhouse suggest obtaining the update step for a weight matrix $W$ as a solution to the optimization problem:
+\begin{equation}
+    \langle G, \delta W \rangle + \frac{\lambda}{2} \norm{\delta W}^2 \to \min_{\delta W}\,,
+\end{equation}
+where $G$ is a gradient-like matrix obtained via backpropagation. Setting norm to RMS-to-RMS norm (scaled version of Spectral norm) produces Muon. We consider two different choices instead:
+\begin{enumerate}
+    \item Choosing kernel norm, $\norm{\cdot}_*$, produces rank-1 update, defined by 
+    \begin{equation}
+        \delta W = -\frac{1}{\lambda} u_1 \sigma_1 v_1^T\,,
+    \end{equation}\label{eqn:update_star}
+    where $\sigma_1$ is largest singular value of $G$, and $u_1$, $v_1$ are corresponding singular values.
 
-\medskip
-\hrule\medskip
-\Head{Structured group lasso}\\
-Another approach is the \emph{structured group lasso}:
-\[
-\begin{array}{ll}
-\mbox{minimize} & f(x) + \sum_{i=1}^N \lambda_i \|x_{g_i}\|_2
-\end{array}
-\]
-where $g_i \subseteq [n]$ and $\mathcal G = \{g_1, \dots, g_N\}$
-\begin{itemize}\itemsep=12pt
-\item like group lasso, but the groups can overlap arbitrarily
-\item particular choices of groups can impose `structured' sparsity
-\item \eg, topic models, selecting interaction terms for (graphical) models,
-    tree structure of gene networks, fMRI data
-\item generalizes to the \textbf{composite absolute penalties family}:
-\[
-r(x) = \|(\|x_{g_1}\|_{p_1}, \dots, \|x_{g_N}\|_{p_N})\|_{p_0}
-\]
-\end{itemize}
+    \item Choosing $F*$ norm, defined by $\norm{\cdot}_{F*} = (\norm{\cdot}_{*} + \norm{\cdot}_{F})/2$, produces a relatively small-rank update, defined by
+    \begin{equation}\label{eqn:update_F_star}
+    \delta W = -\frac{1}{\lambda}UDV^T
+    \end{equation} 
+    with $D = \diag(d_i)$, where $d_i = [\sigma_i - \tau]_+$ and $\tau$ is given by
+    \begin{equation*}
+        \sum_{i=1}^n [\sigma_i - \tau]_+ = \tau\,.
+    \end{equation*}
+\end{enumerate}
 

### 2025-05-06 poster/poster.tex
added algorithms to poster

@@
 \usepackage{amsmath}
+
+\usepackage{caption}
 
@@
 By using alternative norms, such as kernel norm (Kernel-Neon) or custom $F*$ norm ($F*$-Neon), we induce low-rank update matrices that enable more efficient computation.
-We evaluate the performance of Neon, Muon, and Adam on multilayer perceptrons, CIFAR10, and NanoGPT, and demonstrate [resting results here].
+We evaluate the performance of Neon, Muon, and Adam on MLP, CIFAR10, and NanoGPT, and demonstrate [resting results here].
 
@@
     \item Choosing kernel norm, $\norm{\cdot}_*$, produces rank-1 update, defined by 
-    \begin{equation}
+    \begin{equation}\label{eqn:update_star}
         \delta W = -\frac{1}{\lambda} u_1 \sigma_1 v_1^T\,,
-    \end{equation}\label{eqn:update_star}
+    \end{equation}
     where $\sigma_1$ is largest singular value of $G$, and $u_1$, $v_1$ are corresponding singular values.
@@
 \hrule\medskip
-\Head{Hierarchical selection}\\
-\begin{center}
-\begin{tikzpicture}
-[dot/.style={rectangle,draw=black,fill=white,inner sep=5pt,minimum size=5pt}]
-\node[dot,draw=orange,thick] at (0,10) (n1) {1};
-\node[dot] at (-3,8) (n2) {2};
-\node[dot,draw=orange,thick] at (3,8) (n3) {3};
-\node[dot] at (-3,6) (n4) {4};
-\node[dot,draw=orange,thick] at (1.5,6) (n5) {5};
-\node[dot] at (4.5,6) (n6) {6};
-\draw[->] (n1) -- (n2);
-\draw[->] (n1) -- (n3);
-\draw[->] (n2) -- (n4);
-\draw[->] (n3) -- (n5);
-\draw[->] (n3) -- (n6);
-\end{tikzpicture}
-\end{center}
-
-\begin{itemize}\itemsep=8pt
-    \item $\mathcal G = \{ \{4\}, \textcolor{orange}{\{5\}}, \{6\}, \{2,4\}, 
-        \textcolor{orange}{\{3,5,6\}}, \textcolor{orange}{\{1,2,3,4,5,6\} \}}$
-\item nonzero variables form a rooted and connected subtree
-    \begin{itemize}
-        \item if node is selected, so are its ancestors
-        \item if node is not selected, neither are its descendants
-    \end{itemize}
-\end{itemize}
+\Head{Efficient update computation}\\
+A couple of sentences about Lanczos algorithm and reference computation times and required number of iterations for tipical value of $\tau$.
 
@@
 \hrule\medskip
-\Head{Algorithm}\\
-We solve this problem using an ADMM lasso implementation:
-\begin{verbatim}
-prox_f = @(v,rho) (rho/(1 + rho))*(v - b) + b;
-prox_g = @(v,rho) (max(0, v - 1/rho) - max(0, -v - 1/rho));
+\Head{Algorithms}\\
+Now we can write down pseudocode for both versions of Neon.
 
-AA = A*A';
-L  = chol(eye(m) + AA);
-
-for iter = 1:MAX_ITER
-    xx = prox_g(xz - xt, rho);
-    yx = prox_f(yz - yt, rho);
-
-    yz = L \ (L' \ (A*(xx + xt) + AA*(yx + yt)));
-    xz = xx + xt + A'*(yx + yt - yz);
-
-    xt = xt + xx - xz;
-    yt = yt + yx - yz;
-end
-\end{verbatim}
-
-\medskip
-\hrule\medskip
-\Head{Line search}\\
-If $L$ is not known (usually the case), can use the following line search:
-
-\noindent\rule[-5pt]{.8\textwidth}{0.4pt}
-{\footnotesize
-\begin{tabbing}
-    {\bf given} $x^k$, $\lambda^{k-1}$, and parameter $\beta \in (0,1)$. \\*[\smallskipamount]
-    Let $\lambda := \lambda^{k-1}$. \\*[\smallskipamount]
-    {\bf repeat} \\
-    \qquad \= 1.\ Let $z := \prox_{\lambda g}(x^{k} - \lambda \nabla f(x^{k}))$. \\
-    \> 2.\ {\bf break if} $f(z) \leq \hat{f}_{\lambda}(z, x^{k})$. \\
-    \> 3.\ Update $\lambda := \beta \lambda$. \\*[\smallskipamount]
-    {\bf return} $\lambda^{k} := \lambda$, $x^{k+1}:=z$.
-\end{tabbing}}
+\noindent\rule[-5pt]{.8\textwidth}{0.4pt}\\
+{\bf Algorithm 1} Kernel-Neon update step for linear layer\\
+\noindent\rule[10pt]{.8\textwidth}{0.4pt}
+\vspace{-10pt}
+\noindent\begin{tabbing}
+    {\bf Input:} $\lambda$, gradient-like matrix $G$. \\
+    {\bf Output:} Weight matrix update $\delta W$.\\*[\smallskipamount]
+    \qquad \= 1.\ $U, \Sigma, V := \textrm{Lanczos}(G, 1)$. \\
+    {\bf Return} $-\frac{1}{\lambda}U \Sigma V^T$.
+\end{tabbing}
+\vspace{-8pt}
 \noindent\rule[10pt]{.8\textwidth}{0.4pt}
 
-typical value of $\beta$ is $1/2$, and 
-\[
-\hat{f}_\lambda(x,y) = f(y) + \nabla f(y)^T (x - y) + 
-(1/2\lambda)\|x - y\|_2^2
-\]
+For $F*$-Neon it is a bit trickier~\eqref{eqn:update_F_star}, we need to compute $\tau$ and know number of singular values larger then $\tau$, which we denote by $r$.
+Assuming that singular values spectrum of $G$ changes little between iterations (which was true in our experiments) we propose the following algorithm.
 
-\medskip
+\noindent\rule[-5pt]{.8\textwidth}{0.4pt}\\
+{\bf Algorithm 2} $F*$-Neon update step for linear layer\\
+\noindent\rule[10pt]{.8\textwidth}{0.4pt}
+\vspace{-10pt}
+\noindent\begin{tabbing}
+    \qquad\=\qquad\=\qquad\=\kill % Set three tab stops (4em total)
+    {\bf Input:} $\lambda$, $r$, gradient-like matrix $G$. \\
+    {\bf Output:} Weight matrix update $\delta W$.\\*[\smallskipamount]
+    \> 1. $U, \Sigma, V := \textrm{Lanczos}(G, r + 1)$. \\
+    \> 2. $s := \sum_{i=1}^{r}\sigma_r$. \\
+    \> 3.\ {\bf If} $(r + 1)\sigma_{r + 1} > s${\bf:}\\
+    \>\> 4. $r := r + 1$. 5. $\tau := \frac{s + \sigma_{r}}{r + 1}$.\\
+    \> 6.\ {\bf Else if} $(r + 1)\sigma_{r - 1} < s${\bf:}\\
+    \>\> 7. $r := r - 1$. 8. $\tau := \frac{s - \sigma_{r}}{r + 1}$.\\
+    \> 9.\ {\bf Else:}\\
+    \>\> 10. $\tau := \frac{s}{r + 1}$\\
+    \> 11. $D = [\Sigma - \tau I]_+$. \\*[\smallskipamount]
+    {\bf Save} $r$ for the next iteration.\\
+    {\bf Return} $-\frac{1}{\lambda}U D V^T$.
+\end{tabbing}
+\vspace{10pt}
+%\noindent\rule[10pt]{.8\textwidth}{0.4pt}
+
 \hrule\medskip
-\Head{Convergence proof}\\
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+\Head{MLP tests}\\
+Couple of sentences and a vector(!) picture (reference picture for now).
+\begin{figure}[h]
+    \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/accuracy_vs_epochs_simple.png}}
+    \caption{MLP validation loss}
+    \label{fig:mlp_loss}
+\end{figure}
 
@@
 \hrule\medskip
-\Head{Numerical example}\\
-Consider a numerical example with $f(x) = \|Ax - b\|_2^2$
-with $A \in \reals^{10 \times 100}$ and $b \in \reals^{10}$.
-Entries of $A$ and $b$ are generated as independent samples from
-a standard normal distribution.
-Here, we have chosen $\lambda$ using cross validation.
+\Head{CIFAR10 tests}\\
+Couple of sentences and a vector(!) picture (reference picture for now).
+\begin{figure}[h]
+    \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/accuracy_vs_epochs_moderate.png}}
+    \caption{CIFAR10 validation loss}
+    \label{fig:cifar10_loss}
+\end{figure}
 
@@
 \hrule\medskip
-\Head{Results}\\
-On this numerical example, the ADMM method converges quickly.
-We give two realizations corresponding to different parameters $A$ and $b$.
-\begin{center}
-\psfrag{k}[t][b]{$k$}
-\psfrag{fbest - fmin}[b][t]{$f_\mathrm{best}^{(k)} - f^\star$}
-\psfrag{noise-free realize}{noise-free case}
-\psfrag{realize1}{realization 1}
-\psfrag{realize2}{realization 2}
-\includegraphics[width=0.9\textwidth]{figures/pwl_error_fbest_realize.eps}
-\end{center}
+\Head{NanoGPT tests}\\
+Couple of sentences and a vector(!) picture (reference picture for now).
+\begin{figure}[h]
+    \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/loss_val.png}}
+    \caption{NanoGPT validation loss}
+    \label{fig:nanogpt_loss}
+\end{figure}
 

### 2025-05-06 code/config.py
Code has been modified, but cannot be tested (the device is unavailable)

@@ -71,7 +71,7 @@ OPTIMIZER_CONFIGS = {
      'neon_acc': {
         'type': 'neon',
         'label': 'Neon (Accurate)',
-        'num_epochs': 30,
+        'num_epochs': 20,
         'lr': 0.1,
         'momentum': 0.3,
         'nesterov': True,
@@ -112,7 +112,7 @@ OPTIMIZER_CONFIGS = {
 # Training configuration
 TRAIN_CONFIG = {
     'batch_size': batch_size,
-    'model_type': 'simple',
+    'model_type': 'moderate',
     'optimizers': list(OPTIMIZER_CONFIGS.keys())  # List of optimizer keys to use
 }
 
@@ -124,7 +124,7 @@ AUG_CONFIG = {
 
 # Plotting configuration
 PLOT_CONFIG = {
-    'save_dir': 'figures',
+    'save_dir': 'fig/06mlp_no_schedule',
     'figsize': (10, 5),
     'dpi': 100
 } 

### 2025-05-06 code/matrix_functions.py

@@ -27,6 +27,28 @@ def find_tau(S, tau_start, eps=1e-2):
             high = mid
     return (low + high) / 2
 
+def find_tau_torch(S, tau_start, eps=1e-2):
+    smax = torch.max(S)
+    f_mid = torch.sum(torch.clamp(S - tau_start, min=0))
+    if abs(f_mid - tau_start) < eps * smax:
+        # print("Already ok")
+        return tau_start
+    # print("Not ok")
+    if f_mid > tau_start:
+        low = tau_start
+    else:
+        high = tau_start
+ 
+    low, high = torch.tensor(0.0, device=S.device), smax
+    while high - low > eps * smax:
+        mid = (low + high) / 2
+        f_mid = torch.sum(torch.clamp(S - mid, min=0))
+        if f_mid > mid:
+            low = mid
+        else:
+            high = mid
+    return (low + high) / 2
+
 def k_sv_svds_approximation_dlpack(W_torch, k, tau, num_iter=30):
     """SVD approximation using the top k singular values and corresponding vectors."""
     W = cp.from_dlpack(thd.to_dlpack(W_torch)).astype(cp.float32)
@@ -45,6 +67,25 @@ def k_sv_svds_approximation_dlpack(W_torch, k, tau, num_iter=30):
     approx = U @ cp.diag(S_thresholded) @ Vt
     approx_torch = thd.from_dlpack(approx.toDlpack()) 
     return approx_torch, opt_tau, k
+
+
+def svd_full_approximation(A, tau=0):
+    k = 0
+    if (k == 0):
+        k = min(A.shape[0],A.shape[1])
+    # A = A.double() # uncomment this if you need higher precision
+    U, S, Vh = torch.linalg.svd(A, full_matrices=False)
+    opt_tau = find_tau_torch(S, tau)
+    S_thresholded = torch.clamp(S - opt_tau, min=0)
+    # print(sum(S_thresholded), opt_tau)
+    approx = U @ torch.diag(S_thresholded) @ Vh
+    return approx, opt_tau, k
+
+    # S_thresholded = torch.clamp(S - tau, min=0) 
+    # A_reconstructed = U[:,:k] @ torch.diag(S[:k]) @ Vh[:k,:]
+    # print(f"condition {S[0]/S[-1]:.4e}")
+    return A_reconstructed
+
 
 def one_sv_svds_approximation(W_torch, num_iter=30):
     """SVD approximation using the top k singular values and corresponding vectors."""

### 2025-05-06 code/neon_cnn_tester.py

@@ -48,42 +48,44 @@ def create_optimizer(model, total_steps, optimizer_type, **kwargs):
             layers.append((name, module))
     
     for name, layer in layers:
-        if optimizer_type == 'neon':
-            # For Neon, create a new optimizer with random tau for each layer
-            import random
-            tau = random.uniform(0.01, 1)  # Random tau between 0.1 and 1.0
-            layer_kwargs = kwargs.copy()
-            layer_kwargs['tau'] = tau
-            optimizer = Neon(layer.parameters(), **layer_kwargs)
-            # Create scheduler for Neon
-            scheduler = torch.optim.lr_scheduler.LambdaLR(optimizer, lr_lambda=lambda t: lr_lambda(t, total_steps))
-            '''
-            scheduler = torch.optim.lr_scheduler.StepLR(
-                optimizer,
-                step_size=kwargs.get('lr_double_after_epoch', 3),
-                gamma=2.0  # Double the learning rate
-            )'''
-        elif optimizer_type == 'muon':
-            optimizer = Muon(layer.parameters(), **kwargs)
-            # Create scheduler for Muon
-            scheduler = torch.optim.lr_scheduler.StepLR(
-                optimizer,
-                step_size=5,  # Adjust every 5 epochs
-                gamma=0.5  # Halve the learning rate
-            )
-        elif optimizer_type == 'sgd':
-            optimizer = torch.optim.SGD(layer.parameters(), **kwargs)
-            # Create scheduler for SGD
-            scheduler = torch.optim.lr_scheduler.StepLR(
-                optimizer,
-                step_size=5,  # Adjust every 5 epochs
-                gamma=0.5  # Halve the learning rate
-            )
-        else:
-            raise ValueError(f"Unknown optimizer type: {optimizer_type}")
+        # Separate parameters based on their shapes
+        matrix_params = []
+        non_matrix_params = []
         
-        optimizers.append(optimizer)
-        schedulers.append(scheduler)
+        for p in layer.parameters():
+            if p.requires_grad:
+                if len(p.shape) == 2:  # Matrix parameters
+                    matrix_params.append(p)
+                else:  # Non-matrix parameters (bias, conv weights, etc.)
+                    non_matrix_params.append(p)
+        
[…]

### 2025-05-06 code/optimizers.py

@@ -2,7 +2,7 @@ import math
 import torch
 import torch.nn.functional as F
 from torch import nn
-from matrix_functions import k_sv_svds_approximation_dlpack, one_sv_svds_approximation
+from matrix_functions import k_sv_svds_approximation_dlpack, one_sv_svds_approximation, svd_full_approximation
 
 def zeropower_via_newtonschulz5(G, steps=3, eps=1e-7):
     """Simplified Newton-Schulz iteration for whitening"""
@@ -149,4 +149,5 @@ class Neon(torch.optim.Optimizer):
                         update = one_sv_svds_approximation(g_resh, self.lanczos_iter_num)
                     elif self.type == 'accurate':
                         update, self.tau, self.k = k_sv_svds_approximation_dlpack(g_resh, self.k, self.tau, self.lanczos_iter_num)
+                        # update, self.tau, self.k = svd_full_approximation(g_resh, self.tau) -- too slow
                 p.data.add_(update.view(g.shape), alpha=-lr * math.sqrt(n / m)) 

### 2025-05-06 poster/neon_poster.tex
Updated the poster

--- 2025-05-06 poster/poster.tex
@@
 
+%\usepackage{svg}
 \usepackage[papersize={38in,28in}]{geometry}
@@
 \Head{MLP tests}\\
-Couple of sentences and a vector(!) picture (reference picture for now).
+We test Muon, SGD, and Neon on a simple MLP (2 linear layers) that solves CIFAR-10 classification problem. Muon converges faster than Neon and reaches higher accuracy.
 \begin{figure}[h]
-    \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/accuracy_vs_epochs_simple.png}}
+    \center{\includegraphics[width=0.8\linewidth]{../figs/mlp0605/loss_vs_epochs_mlp.pdf}}
     \caption{MLP validation loss}
@@
 \begin{textblock}{7.0}(16,1.5)
+    \begin{figure}[h]
+        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp0605/accuracy_vs_time_mlp.pdf}}
+        \caption{MLP Accuracy vs wallclock time}
+        \label{fig:mlp_loss}
+    \end{figure}
 
 \hrule\medskip
-\Head{CIFAR10 tests}\\
-Couple of sentences and a vector(!) picture (reference picture for now).
-\begin{figure}[h]
-    \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/accuracy_vs_epochs_moderate.png}}
-    \caption{CIFAR10 validation loss}
-    \label{fig:cifar10_loss}
-\end{figure}
+\Head{CIFAR-10 tests}\\
+Now we compare Muon and Neon on ResNet from Keller's cifar10-aribench at GitHub.
+On RTX-4050, Muon is approximately twice as fast as Neon.
+Moreover, Neon has a strange dependency on batch size that has not been observed for Muon: smaller batch size 500 leads to accuracy 84\% in eight epochs, while batch size 2000 results in only 51\%.
+
+TO-DO: we conduct experiments like those with MLP, and add plots here.
 

### 2025-05-06 poster/neon_poster.tex
added nanogpt validation loss (and train too but it does not fit on a poster, just uncomment in code if needed)
added description of Lanczos svd

@@
 \Head{Efficient update computation}\\
-A couple of sentences about Lanczos algorithm and reference computation times and required number of iterations for tipical value of $\tau$.
-
+% A couple of sentences about Lanczos algorithm and reference computation times and required number of iterations for tipical value of $\tau$.
+We use cupy's svds routine to obtain gradients' matrices' SVD approximation formed by largest singular values and corresponding vectors.
+By applying the Lanczos process to either $A^TA$ or $AA^T$, it generates a sequence of orthogonal vectors (Lanczos vectors) that capture 
+the dominant spectral properties of the original matrix. The singular values and vectors are then extracted from the tridiagonal matrix using 
+standard eigenvalue techniques like QR iteration.
 \end{textblock}
@@
 Couple of sentences and a vector(!) picture (reference picture for now).
+% \begin{figure}[h]
+%     \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/loss_val.png}}
+%     \caption{NanoGPT validation loss}
+%     \label{fig:nanogpt_loss}
+% \end{figure} 
+% \begin{figure}[h]
+%     \center{\includegraphics[width=0.8\linewidth]{../figs/nanogpt/train_loss.pdf}}
+%     \caption{NanoGPT train loss}
+%     \label{fig:nanogpt_train_loss}
+% \end{figure}
 \begin{figure}[h]
-    \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/loss_val.png}}
+    \center{\includegraphics[width=0.8\linewidth]{../figs/nanogpt/val_loss.pdf}}
     \caption{NanoGPT validation loss}
-    \label{fig:nanogpt_loss}
+    \label{fig:nanogpt_val_loss}
 \end{figure}
-
 \medskip

### 2025-05-07 article/muon.tex
Added mentors

@@
   \texttt{vinogradov.am@phystech.edu}
+  \and
+  Daniil Merkulov \\
+  \texttt{daniil.merkulov@phystech.edu}
+  \and
+  Ivan Oseledets \\
+  \texttt{i.oseledets@scholtech.ru}
 }
@@
     \begin{figure}[h!]
-        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/accuracy_vs_epochs_simple.png}}
+        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp0605/loss_vs_epochs_mlp.pdf}}
         \caption{MLP: self.linear1 = nn.Linear(32*32*3, 512), self.linear2 = nn.Linear(512, 10), self.activ = nn.GELU()}
@@
     \begin{figure}[h!]
-        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/accuracy_vs_time_simple.png}}
+        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp0605/accuracy_vs_time_mlp.pdf}}
         \caption{MLP: wallclock time measurements}

### 2025-05-13 poster/neon_poster.tex
Added mentors
changed kernel norm to nuclear norm in poster. There is no kernel norm, I am idiot

@@
 \LARGE
-Alexey Kravatskiy, Ivan Kozyrev, Nikolay Kozlov, Alexander Vinogradov
+Alexey Kravatskiy, Ivan Kozyrev, Nikolay Kozlov, Alexander Vinogradov, Daniil Merkulov, Ivan Oseledets
 }
@@
 To bridge this gap, we propose Neon, a new optimizer that builds upon the framework of Bernstein and Newhouse.
-By using alternative norms, such as kernel norm (Kernel-Neon) or custom $F*$ norm ($F*$-Neon), we induce low-rank update matrices that enable more efficient computation.
+By using alternative norms, such as nuclear norm (Nuclear-Neon) or custom $F*$ norm ($F*$-Neon), we induce low-rank update matrices that enable more efficient computation.
 We evaluate the performance of Neon, Muon, and Adam on MLP, CIFAR10, and NanoGPT, and demonstrate [resting results here].
@@
 \begin{enumerate}
-    \item Choosing kernel norm, $\norm{\cdot}_*$, produces rank-1 update, defined by 
+    \item Choosing nuclear norm, $\norm{\cdot}_*$, produces rank-1 update, defined by 
     \begin{equation}\label{eqn:update_star}
@@
 \noindent\rule[-5pt]{.8\textwidth}{0.4pt}\\
-{\bf Algorithm 1} Kernel-Neon update step for linear layer\\
+{\bf Algorithm 1} Nuclear-Neon update step for linear layer\\
 \noindent\rule[10pt]{.8\textwidth}{0.4pt}

### 2025-05-13 poster/neon_poster.tex
added lost 1/2 to poster
made algorithms smaller to leave more space for plots
New pictures from Nanogpt training

@@
     \begin{equation}\label{eqn:update_star}
-        \delta W = -\frac{1}{\lambda} u_1 \sigma_1 v_1^T\,,
+        \delta W = -\frac{1}{2\lambda} u_1 \sigma_1 v_1^T\,,
     \end{equation}
@@
 
+{\footnotesize{}
 \noindent\rule[-5pt]{.8\textwidth}{0.4pt}\\
 {\bf Algorithm 1} Nuclear-Neon update step for linear layer\\
-\noindent\rule[10pt]{.8\textwidth}{0.4pt}
-\vspace{-10pt}
+\noindent\rule[5pt]{.8\textwidth}{0.4pt}
+\vspace{-15pt}
 \noindent\begin{tabbing}
@@
     \qquad \= 1.\ $U, \Sigma, V := \textrm{Lanczos}(G, 1)$. \\
-    {\bf Return} $-\frac{1}{\lambda}U \Sigma V^T$.
+    {\bf Return} $-\frac{1}{2\lambda}U \Sigma V^T$.
 \end{tabbing}
@@
 \noindent\rule[10pt]{.8\textwidth}{0.4pt}
+}
 
@@
 
+{\footnotesize{}
 \noindent\rule[-5pt]{.8\textwidth}{0.4pt}\\
 {\bf Algorithm 2} $F*$-Neon update step for linear layer\\
-\noindent\rule[10pt]{.8\textwidth}{0.4pt}
-\vspace{-10pt}
+\noindent\rule[5pt]{.8\textwidth}{0.4pt}
+\vspace{-15pt}
 \noindent\begin{tabbing}
@@
 \vspace{10pt}
+}
 %\noindent\rule[10pt]{.8\textwidth}{0.4pt}
@@
 \Head{NanoGPT tests}\\
-Couple of sentences and a vector(!) picture (reference picture for now).
+We finetuned NanoGPT by Alex Karpathy on two NVIDIA RTX 4090 GPUs (24 GB each) on tiny stories dataset via 200 iterations.
+Both Neon and Muon showed better convergence then AdamW.
+Neon showed similiar convergence rate to Muon
+
 % \begin{figure}[h]
-%     \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/loss_val.png}}
+%     \center{\includegraphics[width=0.8\linewidth]{../../figs/mlp24/loss_val.png}}
 %     \caption{NanoGPT validation loss}
@@
 % \begin{figure}[h]
-%     \center{\includegraphics[width=0.8\linewidth]{../figs/nanogpt/train_loss.pdf}}
+%     \center{\includegraphics[width=0.8\linewidth]{../../figs/nanogpt/train_loss.pdf}}
 %     \caption{NanoGPT train loss}

### 2025-05-13 code/config.py
Neon vs Muon

@@ -124,7 +124,7 @@ AUG_CONFIG = {
 
 # Plotting configuration
 PLOT_CONFIG = {
-    'save_dir': 'fig/06mlp_no_schedule',
+    'save_dir': 'fig/13mlp_no_schedule',
     'figsize': (10, 5),
     'dpi': 100
 } 

### 2025-05-13 code/muon_airbench.py

@@ -147,11 +147,11 @@ def main():
     whiten_bias_train_steps = ceil(3 * len(train_loader))
 
     # Create optimizers and learning rate schedulers
-    filter_params = [p for p in model.parameters() if len(p.shape) == 4 and p.requires_grad]
+    filter_params = [p for p in model.parameters() if len(p.shape) == 4 and p.requires_grad] # for convolution
     norm_biases = [p for n, p in model.named_parameters() if "norm" in n and p.requires_grad]
-    param_configs = [dict(params=[model.whiten.bias], lr=bias_lr, weight_decay=wd/bias_lr),
-                     dict(params=norm_biases,         lr=bias_lr, weight_decay=wd/bias_lr),
-                     dict(params=[model.head.weight], lr=head_lr, weight_decay=wd/head_lr)]
+    param_configs = [dict(params=[model.whiten.bias], lr=bias_lr, weight_decay=wd/bias_lr), # whitening or normalization layer
+                     dict(params=norm_biases,         lr=bias_lr, weight_decay=wd/bias_lr), # batch normalization
+                     dict(params=[model.head.weight], lr=head_lr, weight_decay=wd/head_lr)] # output
     optimizer1 = torch.optim.SGD(param_configs, momentum=0.85, nesterov=True) #, fused=True)
     # optimizer2 = Muon(filter_params, lr=0.24, momentum=0.6, nesterov=True)
     optimizer2 = Neon(filter_params, lr=0.05, momentum=0.6, nesterov=True, neon_mode='accurate',

### 2025-05-13 code/muon_vs_neon.py

@@ -1,0 +1,414 @@
+# Testing Neon on the benchmark, the code works for Muon around 3s indeed on a powerful GPU
+# Neon is far worse. Somehow its accuracy more depends on the batchsize.
+from math import ceil
+import torch
+torch.backends.cudnn.benchmark = True
+from torch import nn
+import torch.nn.functional as F
+import airbench
+import matplotlib.pyplot as plt
+import time
+import os
+import numpy as np
+
+from optimizers import Neon, Muon
+
+code_version = '13may_muon_neon'
+
+# Neon learning rate schedule by epoch (0-indexed)
+# Modify these values to change the learning rate at each epoch
+NEON_LR_SCHEDULE = {
+    0: 0.001,    # Initial learning rate
+    1: 0.001,    # Epoch 1
+    2: 0.001,    # Epoch 2
+    3: 0.002,    # Epoch 3
+    4: 0.002,    # Epoch 4
+    5: 0.002,   # Epoch 5
+    6: 0.002,   # Epoch 6
+    7: 0.001,  # Epoch 7
+    8: 0.001,
+    9: 0.001,
+    10: 0.001,
+}
+
+# @torch.compile
+def zeropower_via_newtonschulz5(G, steps=3, eps=1e-7):
+    assert len(G.shape) == 2
+    a, b, c = (3.4445, -4.7750,  2.0315)
+    X = G.bfloat16()
+    X /= (X.norm() + eps) # ensure top singular value <= 1
+    if G.size(0) > G.size(1):
+        X = X.T
+    for _ in range(steps):
+        A = X @ X.T
+        B = b * A + c * A @ A
+        X = a * X + B @ X
+    if G.size(0) > G.size(1):
+        X = X.T
+    return X
+
+class Muon(torch.optim.Optimizer):
+    def __init__(self, params, lr=1e-3, momentum=0, nesterov=False):
+        defaults = dict(lr=lr, momentum=momentum, nesterov=nesterov)
+        super().__init__(params, defaults)
+
+    def step(self):
+        for group in self.param_groups:
+            lr = group["lr"]
+            momentum = group["momentum"]
+            for p in group["params"]:
+                g = p.grad
+                if g is None:
+                    continue
+                state = self.state[p]
+
+                if "momentum_buffer" not in state.keys():
+                    state["momentum_buffer"] = torch.zeros_like(g)
+                buf = state["momentum_buffer"]
+                buf.mul_(momentum).add_(g)
[…]

### 2025-05-13 code/neon_cnn_tester.py

@@ -114,208 +114,4 @@ def run_training(model, optimizers, schedulers, train_loader, test_loader, total
                 optimizer.zero_grad()
             
             outputs = model(inputs)
-            loss = criterion(outputs, targets)
-            loss.backward()
-            
-            # Step all optimizers
-            if epoch > 0:  # we want to measure the situation before everything
-                for optimizer in optimizers:
-                    optimizer.step()
-        
-            # Step all schedulers after each epoch
-            if epoch > 0:
-                for scheduler in schedulers:
-                    scheduler.step()
-        
-        train_loss += loss.item()
-        _, predicted = outputs.max(1)
-        train_total += targets.size(0)
-        train_correct += predicted.eq(targets).sum().item()
-        
-        train_acc = 100. * train_correct / train_total
-        train_losses.append(train_loss/len(train_loader))
-        train_accs.append(train_acc)
-        
-        # Test
-        model.eval()
-        test_loss = 0
-        test_correct = 0
-        test_total = 0
-        
-        with torch.no_grad():
-            for inputs, targets in test_loader:
-                outputs = model(inputs)
-                loss = criterion(outputs, targets)
-                
-                test_loss += loss.item()
-                _, predicted = outputs.max(1)
-                test_total += targets.size(0)
-                test_correct += predicted.eq(targets).sum().item()
-        
-        test_acc = 100. * test_correct / test_total
-        test_losses.append(test_loss/len(test_loader))
-        test_accs.append(test_acc)
-        times.append(time.time() - start_time)
-        
-        # Print current learning rates
-        if epoch > 0:
-            lrs = [param_group['lr'] for optimizer in optimizers for param_group in optimizer.param_groups]
-            print(f'Epoch: {epoch}/{total_epochs} | '
-                  f'Train Loss: {train_losses[-1]:.3f} | '
-                  f'Train Acc: {train_acc:.2f}% | '
-                  f'Test Loss: {test_losses[-1]:.3f} | '
-                  f'Test Acc: {test_acc:.2f}% | '
-                  f'LR: {min(lrs):.6f} | '
-                  f'Optimizer: {optimizer_label}')
[…]

### 2025-05-13 poster/neon_poster.tex
Neon vs Muon

@@
 
-\begin{textblock}{7.0}(16,1.5)
-    \begin{figure}[h]
-        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp0605/accuracy_vs_time_mlp.pdf}}
-        \caption{MLP Accuracy vs wallclock time}
-        \label{fig:mlp_loss}
-    \end{figure}
+%\begin{textblock}{7.0}(16,1.5)
+%    \begin{figure}[h]
+%        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp0605/accuracy_vs_time_mlp.pdf}}
+%        \caption{MLP Accuracy vs wallclock time}
+%        \label{fig:mlp_loss}
+%    \end{figure}
 
@@
 On RTX-4050, Muon is approximately twice as fast as Neon.
-Moreover, Neon has a strange dependency on batch size that has not been observed for Muon: smaller batch size 500 leads to accuracy 84\% in eight epochs, while batch size 2000 results in only 51\%.
+That is likely caused by ineffective implementation of Neon's update, because in theory updates of Neon are more light and simple, especially for Nuclear Neon.
+Interestingly, Neon has a ceiling of convergence about 70\% which is not observed for Muon.
+Moreover, Neon requires far smaller learning rates to converge.
+All these aspects should be considered when we improve the algorithm.
 
-TO-DO: we conduct experiments like those with MLP, and add plots here.
+\begin{textblock}{7.0}(16,1.5)
+    \begin{figure}[h]
+        \center{\includegraphics[width=0.8\linewidth]{../figures/13may_muon_neon/test_acc_vs_epoch.pdf}}
+        \caption{Accuracy vs Epoch for Muon vs Neon}
+        \label{fig:mlp_loss}
+    \end{figure}
 

### 2025-05-13 code/models.py

@@ -241,7 +241,7 @@ class Conv(nn.Conv2d):
 class ConvGroup(nn.Module):
     def __init__(self, channels_in, channels_out):
         super().__init__()
-        self.conv1 = Conv(channels_in,  channels_out)
+        self.conv1 = Conv(channels_in, channels_out)
         self.pool = nn.MaxPool2d(2)
         self.norm1 = BatchNorm(channels_out)
         self.conv2 = Conv(channels_out, channels_out)

### 2025-05-13 poster/neon_poster.tex
Poster fixed

@@
 % but it looks awful.
-\begin{textblock}{23}(0,0)
+\begin{textblock}{23}(0,-0.25)
 \Title{Neon: Nuclear Norm to Beat Muon}
@@
 
-\begin{textblock}{23}(0,0.6)
+\begin{textblock}{23}(0,0.35)
 {
@@
 
-\begin{textblock}{7.0}(0,1.5)
+\begin{textblock}{7.0}(0,1.25)
 
@@
 
-\begin{textblock}{7.0}(8,1.5)
+\begin{textblock}{7.0}(8,1.0)
 \hrule\medskip
@@
 \begin{figure}[h]
-    \center{\includegraphics[width=0.8\linewidth]{../figs/mlp0605/loss_vs_epochs_mlp.pdf}}
+    \center{\includegraphics[width=0.7\linewidth]{../figs/mlp0605/loss_vs_epochs_mlp.pdf}}
     \caption{MLP validation loss}
+    \label{fig:mlp_loss}
+\end{figure}
+
+\begin{figure}[h]
+    \center{\includegraphics[width=0.7\linewidth]{../figs/mlp0605/accuracy_vs_time_mlp.pdf}}
+    \caption{MLP Accuracy vs wallclock time}
     \label{fig:mlp_loss}
@@
 %\begin{textblock}{7.0}(16,1.5)
-%    \begin{figure}[h]
-%        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp0605/accuracy_vs_time_mlp.pdf}}
-%        \caption{MLP Accuracy vs wallclock time}
-%        \label{fig:mlp_loss}
-%    \end{figure}
+
+\begin{textblock}{7.0}(16,0.75)
 
@@
 
-\begin{textblock}{7.0}(16,1.5)
     \begin{figure}[h]
-        \center{\includegraphics[width=0.8\linewidth]{../figures/13may_muon_neon/test_acc_vs_epoch.pdf}}
+        \center{\includegraphics[width=0.75\linewidth]{../figures/13may_muon_neon/test_acc_vs_epoch.pdf}}
         \caption{Accuracy vs Epoch for Muon vs Neon}
@@
 \Head{Conclusion}\\
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+A new norm-motivated algorithm proved to be not so effective as Muon, at least from the first sight.
+However, its updates seem to be much faster to compute, and more so for a large matrices which has not tested yet.
+We believe that our algorithm is not an alternative to Muon, but probably a more effective tool for solving some other optimization problem.
+A deep inspection of Nuclear Norm and its properties might give us an insight into Neon's future.
 
 \medskip
-\hrule\medskip
-\Head{Acknowledgements}\\
-This material is based upon work supported by the
-X Fellowship and my mom.
+%\hrule\medskip
+%\Head{Acknowledgements}\\
+%We thank Jeremy Bernstein for his idea to use norms to derive a optimizer.
 

### 2025-08-16 code/optimizer_comparison.py
Alpha version of L-smooth

@@ -759,7 +759,7 @@ if __name__ == "__main__":
         m=1000,           # Matrix rows
         n=100,          # Matrix columns
         num_nodes=20,
-        num_iterations=50000,  # Number of optimization iterations
+        num_iterations=200000,  # Number of optimization iterations
         plot_interval=100,     # Plot data every 50 iterations
     )
     

### 2025-08-16 conf_article/neon.tex
New abstract and preparation for an article

--- 2025-04-28 article/muon.tex
@@
-%% ============================================
-%% ================ Preambule =================
-%% ============================================
-\documentclass[]{scrartcl}
-\usepackage[margin = 0.5in]{geometry}
+\documentclass{article} % For LaTeX2e
+\usepackage{icomp2024_conference,times}
 
-\usepackage[pdftex,unicode, 
-colorlinks=true,
-linkcolor = blue]{hyperref}	% нумерование страниц, ссылки!!!!ИМЕННО В ТАКОМ ПОРЯДКЕ СО СЛЕДУЮЩИМ ПАКЕТОМ
-%\usepackage[warn]{mathtext}				% Поддержка русского текста в формулах
-\usepackage[T1, T2A]{fontenc}			% Пакет выбора кодировки и шрифтов
-\usepackage[utf8]{inputenc} 			% любая желаемая кодировка
-\usepackage[english]{babel}		% поддержка русского языка
-\usepackage{wrapfig}					% Плавающие картинки
-\usepackage{amssymb, amsmath}			% стилевой пакет для формул
+% Optional math command from https://github.com/goodfeli/dlbook_notation.
+\input{math_commands.tex}
+
+\usepackage{hyperref}
+\usepackage{url}
+\usepackage{amsmath,amssymb}
+\usepackage{amsthm}
 \usepackage{algorithm}
 \usepackage{algorithmic}
-\usepackage[backend=biber]{biblatex}
-\addbibresource{muon.bib}
+\usepackage{graphicx}
+\usepackage{subcaption}
+\newtheorem{lemma}{Lemma}
+\newcommand{\norm}[1]{\lVert #1\rVert}
 
-\ifpdf
-\usepackage{cmap} 				% чтобы работал поиск по PDF
-\usepackage[pdftex]{graphicx}
-%\usepackage{pgfplotstable}		% Для вставки таблиц.
-\pdfcompresslevel=9 			% сжимать PDF
-\else
-\usepackage{graphicx}
-\fi
+\title{From Muon to Neon: Introducing Nuclear Norm to Large Matrices}
 
-\graphicspath{{./figures/}}
-\usepackage{subcaption}
+% Authors must not appear in the submitted version. They should be hidden
+% as long as the \icompfinalcopy macro remains commented out below.
+% Non-anonymous submissions will be rejected without review.
 
-%custom commands
-\newcommand{\norm}[1]{\Vert{#1}\Vert}
-
-%% ============================================
-%% ================ Info =================
-%% ============================================
-\title{Neon: Nuclear Norm to Beat Muon}
-
-\author{
-  Alexey Kravatskiy\\
-  \texttt{kravtskii.aiu@phystech.edu}
-  \and
-  Ivan Kozyrev\\
-  \texttt{kozyrev.in@phystech.edu}
-  \and
-  Nikolay Kozlov \\
-  \texttt{kozlov.na@phystech.edu}
-  \and
+\author{Alexey Kravatskiy \\
+Moscow Institute of Physics and Technology (MIPT) \\
+\texttt{kravtskii.aiu@phystech.edu} \\
+\And
+Ivan Kozyrev \\
+Moscow Institute of Physics and Technology (MIPT) \\
+\texttt{kozyrev.in@phystech.edu} \\
+\And
+Nikolai Kozlov \\
+Moscow Institute of Physics and Technology (MIPT) \\
+\texttt{kozlov.na@phystech.edu} \\
+\And
 Alexander Vinogradov \\
-  \texttt{vinogradov.am@phystech.edu}
+Moscow Institute of Physics and Technology (MIPT) \\
+\texttt{vinogradov.am@phystech.edu} \\
+\And
+Daniil Merkulov \\
+Moscow Institute of Physics and Technology (MIPT), Skoltech, HSE, Sber \\
+\texttt{daniil.merkulov@phystech.edu} \\
+\And
+Ivan Oseledets \\
+AIRI, Skoltech \\
+\texttt{i.oseledets@skoltech.ru}
 }
 
-%\author{\begin{tabular}{c c}
-%	  	 Alexey Kravatsky & Nikolay Kozlov & Ivan Kozyrev \\
-%		 \texttt{kravatskii.aiu@phystech.edu} & \texttt{kozlov.n?@phystech.edu} & \texttt{kozyrev.i?@phystech.edu}  
-%		\end{tabular}}
-\date{\today}
+% The \author macro works with any number of authors. There are two commands
+% used to separate the names and addresses of multiple authors: \And and \AND.
+%
+% Using \And between authors leaves it to \LaTeX{} to determine where to break
+% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
+% puts 3 of 4 authors names on the first line, and the last on the second
+% line, try using \AND instead of \And before the third author name.
 
+\newcommand{\fix}{\marginpar{FIX}}
+\newcommand{\new}{\marginpar{NEW}}
+
+%\icompfinalcopy % Uncomment for camera-ready version, but NOT for submission.
 \begin{document}
@@
 \begin{abstract}
-In this paper, we develop a new algorithm for optimization of functions of weight matrices, which are typical for training large language models.
-Changing spectral norm, which was used to derive Muon, to nuclear norm, we pose a new optimization problem for an update matrix, solution of which defines a novel algorithm we name Neon.
-To make it feasible, we use Lanczos algorithm to find a required step.
-After providing theoretical guarantees of Neon convergence, we compare performances of Neon, Muon, and Adam on training multilayer perceptron, convolutional neural network and NanoGPT.
+In this article, we develop a new algorithm for optimizing functions of weight matrices that arise in training large language models.
+By replacing the spectral-norm regularization used in Muon with a nuclear-norm formulation, we obtain a new optimization problem for the update matrix; its solution defines a novel algorithm we call Neon.
+We derive the optimal solution to this problem and propose a numerically efficient method to compute it via the Lanczos algorithm.
+We analyze Neon from a theoretical perspective and provide an extensive empirical study across diverse relevant tasks and settings.
 
-\end{abstract}
+New variant:
+\iffalse This article investigates utilization of matrix norms besides the spectral norm in linear optimization oracle (LMO) framework.
+Making use of the nuclear norm, the mean between the nuclear norm and the Frobenius norm we name the F*-norm, and the dual to F*-norm, we develop 3 new algorithms for optimizing functions of weight matrices that arise in training large language models.\fi
+
+In this article, we develop a family of 3 new algorithms for optimizing functions of weight matrices that arise in training large language models.
+By replacing the spectral norm, which is used when deriving Muon within the linear optimization oracle framework (LMO), with the nuclear norm, the mean between the nuclear norm and the Frobenius norm we name the F*-norm, and the dual to F*-norm we name F2-norm, we obtain 3 new optimization problems for the update matrix; their solutions define 3 novel algorithms we name Neon, F*-Neon, and F2-Muon respectively.
+We derive the optimal solutions to these problems and propose a numerically efficient method to compute updates for Neon and F*-Neon via the Lanczos algorithm (while an update of F2-Muon turns out to be a mean between the Muon update and the Normalized SGD update).
+We analyze all 3 algorithms from a theoretical perspective and provide an extensive empirical study across diverse relevant tasks and settings.
 
 \section{Idea}
+\iffalse
 The goal of the project is to make variations on Muon to speed it up.
-Recently, authors of \cite{bernstein2024oldoptimizernewnorm} have proposed to look at different optimizers as the solution of the optimization problem for an update.
+Recently, authors of \cite{bernstein2024oldoptimizernewnorm} have proposed to derive update step for an optimizer as the solution of the certain optimization problem.
 This approach can be utilized to derive Muon \cite{jordan2024muon}, a novel algorithm for fast training of neural networks.
-Instead of using spectral norm in the problem, we use nuclear norm to produce a new problem.
-%Then, we add momentum to update.
-After finalizing the algorithm, we test it on MLP, and, if results are satisfactory, on transformers (probably something from huggingface).
-The result will be a fast algorithm, which we will convert into a new optimizer class for PyTorch, as was done with Muon.%
-
+Instead of using spectral norm, used to derive Muon, we utilize nuclear norm to produce a new optimization algorithm.
+\fi
 \subsection{Problem (Project description)}
@@
 
-Motivated by the recent achievements of the Muon optimizer (e.g., \cite{liu2025muon}), we consider alternative choices of norms, specifically the kernel norm $\norm{\cdot}_*$ and a custom $F*$ norm, given by $\norm{X}_{F*}^2 = (\norm{X}_F + \norm{X}_*)/2$, where $\norm{\cdot}_F$ denotes the Frobenius norm.
+Motivated by the recent achievements of the Muon optimizer (e.g., \cite{liu2025muon}), we consider alternative choices of norms, specifically the nuclear norm $\norm{\cdot}_*$ and a custom $F*$ norm, given by
+\begin{equation}\label{eqn:F_star}
+    \norm{X}_{F*}^2 = \frac{\norm{X}_F + \norm{X}_*}{2}\,,
+\end{equation}
+where $\norm{\cdot}_F$ denotes the Frobenius norm.
 
-Using the kernel norm in \eqref{eqn:opt_problem_mat} leads to a rank-one update of the weight matrices:
+Using the nuclear norm in \eqref{eqn:opt_problem_mat} leads to a rank-one update of the weight matrices:
 \begin{equation}\label{eqn:update_star}
-  \delta W = -\frac{1}{\lambda} u_1 \sigma_1 v_1^T\,,
+  \delta W = -\frac{1}{2\lambda} u_1 \sigma_1 v_1^T\,,
 \end{equation}
@@
 \end{equation} 
-with $D = \text{diag}(d_i)$, where $d_i = [\sigma_i - \tau]_+$ and $\tau$ is given by:
+with $D = \text{diag}(d_i)$, where $d_i = [\sigma_i - \tau]_+$, and $\tau$ is given by
 \begin{equation}
-    \sum_{i=1}^n [\sigma_i - \tau]_+ = \tau\,.
+    \sum_{i} [\sigma_i - \tau]_+ = \tau\,.
 \end{equation}
@@
 
-In this article we show how one can quickly compute weight updates defined by \eqref{eqn:update_star} or \eqref{eqn:update_F_star} (for now we need more tests to accept one variant or both of them).
+In this article we show how one can quickly compute weight updates defined by \eqref{eqn:update_star} or \eqref{eqn:update_F_star}.
 Then we finilize the method by adding momentum and test their performance against those of Muon at training multilayer perceptron and transformer.
@@
 
-\section{Outcomes}
-The main results of the project are expected to be:
-\begin{enumerate}
-    \item Deriving of Neon and argumentation of the numerical methods which will be used in the algorithm
-    \item Code implementation of Neon as PyTorch optimizer class
-    \item Numerical experiments comparing performance of training with Neon, Muon, SVD and Adam on several architectures: MLP, CNN, and NanoGPT.
-    \item Participation in CIFAR-10 challenge with Neon (if the alrothim turns out to be competitive)
-    \item Article describing all the results. If they are satisfactory, we intend to be in time to apply for NeurIPS 2025. It is a challenge, but I think we must give it a try.
-\end{enumerate}
-
-\section{Literature review}
+% Removed template instructions sections
+\section{Related Work}
 Our review primarily focuses on the Muon and Shampoo optimizers, as our algorithm extends the ideas used to derive these methods.
@@
 For distributed training environments, optimizers like Dion \cite{ahn2025dioncommunicationefficientoptimizerlarge} specifically target communication efficiency by minimizing data exchange between workers through distributed orthonormalization techniques.
+
+\section{Derivation of update rules}
+
+\begin{lemma}\label{lem:opt_star}
+Let $G \in \mathbb{R}^{m \times n}$ and $\lambda > 0$. Then, the following optimization problem
+\begin{equation*} 
+  f(\delta W) = \langle G, \delta W \rangle + \lambda \norm{\delta W}_*^2 \to \min_{\delta W}
+\end{equation*}
+has solution 
+\begin{equation*}
+  \delta W = -\frac{1}{2\lambda} u_1 \sigma_1 v_1^T\,,
+\end{equation*}
+where $\sigma_1$ is the largest singular value of $G$, and $u_1$ and $v_1$ are the corresponding singular vectors.
+\end{lemma}
+
+\begin{proof}
+Let us denote $r = \min\{m, n\}$. Then by Von Neumann's trace inequality,
+\begin{equation*}
+    \vert\langle G, \delta W \rangle\vert \leq \sum_{i=1}^r \sigma_i(G)\sigma_i(\delta W)
+    \Rightarrow
+    \langle G, \delta W \rangle \geq -\sum_{i=1}^r \sigma_i(G)\sigma_i(\delta W)\,.
+\end{equation*}
+Thus, expressing nuclear norm through singular values, we can write down
+\begin{equation*}
+    f(\delta W) \geq -\sum_{i=1}^r \sigma_i(G)\sigma_i(\delta W) + \lambda\left(\sum_{i=1}^{r}\sigma_i(\delta W)\right)^2 \geq \min_{d_1,\dots,d_r \geq 0}-\sum_{i=1}^r \sigma_i(G)d_i + \lambda\left(\sum_{i=1}^{r}d_i\right)^2.
+\end{equation*}
+By Karush-Kuhn-Tucker theorem, necessary conditions of minimum are
+\begin{equation*}
+    \left(d_i \geq 0\ \text{and}\ -\sigma_i(G) + 2\lambda\sum_{j=1}^{r}d_j = 0\right)\ \text{or}\ \left(d_i = 0\ \text{and}\ -\sigma_i(G) + 2\lambda\sum_{j=1}^{r}d_j \geq 0\right) \,\, i = 1,\dots,r\,.
+\end{equation*}
+These conditions simplify to 
+\begin{equation*}
+    \sum_{i \in S}d_i = \sigma_1(G)\,,\ \begin{cases}
+        d_i \geq 0 & \text{if}\ \sigma_i(G) = \sigma_1(G)\,, \\
+        d_i = 0 & \text{otherwise}\,.
+    \end{cases}
+\end{equation*}
+All points satisfying those conditions deliver minimum, and
+\begin{equation*}
+    f(\delta W) \geq -\frac{\sigma_1^2(G)}{4\lambda}\,.
+\end{equation*}
+
+Now let
+\begin{equation*}
+    \delta W^* = -\frac{1}{2\lambda} u_1 \sigma_1(G) v_1^T\,.
+\end{equation*}
+Inserting it to $f(\delta W)$ gives
+\begin{equation*}
+    f\left(\delta W^*\right) = -\frac{\sigma_1(G)^2}{2\lambda} + \frac{\sigma_1(G)^2}{4\lambda} = -\frac{\sigma_1(G)^2}{4\lambda}
+\end{equation*}
+This matches the derived lower bound. Thus, $\delta W^*$ minimizes $f(\delta W)$.
+\end{proof}
+
+\begin{lemma}\label{lem:opt_F_star}
+Let $G \in \mathbb{R}^{m \times n}$, $r = \min\{m. n\}$ and $\lambda > 0$. Then, the following optimization problem
+\begin{equation*} 
+  f(\delta W) = \langle G, \delta W \rangle + \lambda \norm{\delta W}_{F*}^2 \to \min_{\delta W}\,,
+\end{equation*}
+where $\norm{\cdot}_{F*}$ is defined in \eqref{eqn:F_star} has solution 
+\begin{equation}
+\delta W = -\frac{1}{\lambda}UDV^T
+\end{equation} 
+with $D = \text{diag}(d_i)$, where $d_i = [\sigma_i - \tau]_+$, and $\tau$ is given by
+\begin{equation}
+    \sum_{i=1}^r [\sigma_i - \tau]_+ = \tau\,.
+\end{equation}
+\end{lemma}
+
+\begin{proof}
+Analogously to the proof of Lemma~\ref{lem:opt_star}, we can use Von Neumann's trace inequality to write down
+\begin{equation*}
+    f(\delta W) \geq -\sum_{i=1}^r \sigma_i(G)\sigma_i(\delta W) + \frac{\lambda}{2}\left(\sum_{i=1}^{r}\sigma_i(\delta W)\right)^2 + \frac{\lambda}{2}\sum_{i=1}^{r}\sigma_i^2(\delta W)\,,
+\end{equation*}
+\begin{equation}\label{eqn:f_min_F_star}
+    f(\delta W) \geq \frac{1}{\lambda}\min_{d_1,\dots,d_r \geq 0}-\sum_{i=1}^r \sigma_i(G)d_i + \frac{1}{2}\left(\sum_{i=1}^{r}d_i\right)^2 + \frac{1}{2}\left(\sum_{i=1}^{r}d_i^2\right)\,,
+\end{equation}
+By Karush-Kuhn-Tucker theorem, neccesary conditions of minimum are
+\begin{equation*}
+    \left(d_i \geq 0\ \text{and}\ -\sigma_i(G) + \sum_{j=1}^{r}d_j +  d_i = 0\right)\ \text{or}\ \left(d_i = 0\ \text{and}\ -\sigma_i(G) + \sum_{j=1}^{r}d_j + d_i \geq 0\right) \,\, i = 1,\dots,r\,.
+\end{equation*}
+Denoting $\tau = \sum_{i = 1}^r d_i$ gives $d_i = [\sigma_i(G) - \tau]_+$, where $\tau$ satisfies
+\begin{equation}
+    \sum_{i=1}^n [\sigma_i(G) - \tau]_+ = \tau\,.
+\end{equation}
+Inserting found minimum point into~\eqref{eqn:f_min_F_star} yields
+\begin{equation*}
+    f(\delta W) \geq -\sum_{i = 1}^r d_i(\tau + d_i) + \frac{\tau^2}{2\lambda} + \frac{1}{2\lambda}\sum_{i=1}^{r}d_i^2 = -\frac{1}{2 \lambda} \left( \tau^2 + \sum_{i=1}^r d_i^2 \right)\,.
+\end{equation*}
+
+Now let
+\begin{equation}
+\delta W^* = -\frac{1}{\lambda}UDV^T
+\end{equation} 
+with $D = \text{diag}(d_i)$. Inserting it to $f(\delta W)$ gives
+\begin{equation*}
+    f(\delta W^*) = -\sum_{i = 1}^r d_i(\tau + d_i) + \frac{\tau^2}{2\lambda} + \frac{1}{2\lambda}\sum_{i=1}^{r}d_i^2 = -\frac{1}{2 \lambda} \left( \tau^2 + \sum_{i=1}^r d_i^2 \right)\,.
+\end{equation*}
+This matches the derived lower bound. Thus, $\delta W^*$ minimizes $f(\delta W)$.
+\end{proof}
+
 \section{Quality metrics}
@@
     \begin{figure}[h!]
-        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/svd_evolution_all.png}}
+        \center{\includegraphics[width=0.8\linewidth]{figs/mlp24/svd_evolution_all.png}}
         \caption{Singular values of $50257\times 1280$ layer via 200 iterations}
@@
     \begin{figure}[h!]
-        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/svd_evolution.png}}
+        \center{\includegraphics[width=0.8\linewidth]{figs/mlp24/svd_evolution.png}}
         \caption{Singular values of  of $50257\times 1280$ layer for 5-th,45-th, 65-th,175-th and 200-th iteration}
@@
     \begin{figure}[h!]
-        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/accuracy_vs_epochs_simple.png}}
+        \center{\includegraphics[width=0.8\linewidth]{figs/mlp0605/loss_vs_epochs_mlp.pdf}}
         \caption{MLP: self.linear1 = nn.Linear(32*32*3, 512), self.linear2 = nn.Linear(512, 10), self.activ = nn.GELU()}
@@
     \begin{figure}[h!]
-        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/accuracy_vs_time_simple.png}}
+        \center{\includegraphics[width=0.8\linewidth]{figs/mlp0605/accuracy_vs_time_mlp.pdf}}
         \caption{MLP: wallclock time measurements}
@@
 
-\printbibliography
+\bibliography{icomp2024_conference}
+\bibliographystyle{icomp2024_conference}
+
+\appendix
+\section{Appendix}
+You may include other additional sections here.
 

### 2025-08-16 conf_article/neon_old.tex
New abstract and preparation for an article

--- 2025-08-16 conf_article/neon.tex
@@
 \begin{abstract}
-In this article, we develop a new algorithm for optimizing functions of weight matrices that arise in training large language models.
-By replacing the spectral-norm regularization used in Muon with a nuclear-norm formulation, we obtain a new optimization problem for the update matrix; its solution defines a novel algorithm we call Neon.
-We derive the optimal solution to this problem and propose a numerically efficient method to compute it via the Lanczos algorithm.
-We analyze Neon from a theoretical perspective and provide an extensive empirical study across diverse relevant tasks and settings.
-
-New variant:
-\iffalse This article investigates utilization of matrix norms besides the spectral norm in linear optimization oracle (LMO) framework.
-Making use of the nuclear norm, the mean between the nuclear norm and the Frobenius norm we name the F*-norm, and the dual to F*-norm, we develop 3 new algorithms for optimizing functions of weight matrices that arise in training large language models.\fi
-
-In this article, we develop a family of 3 new algorithms for optimizing functions of weight matrices that arise in training large language models.
-By replacing the spectral norm, which is used when deriving Muon within the linear optimization oracle framework (LMO), with the nuclear norm, the mean between the nuclear norm and the Frobenius norm we name the F*-norm, and the dual to F*-norm we name F2-norm, we obtain 3 new optimization problems for the update matrix; their solutions define 3 novel algorithms we name Neon, F*-Neon, and F2-Muon respectively.
-We derive the optimal solutions to these problems and propose a numerically efficient method to compute updates for Neon and F*-Neon via the Lanczos algorithm (while an update of F2-Muon turns out to be a mean between the Muon update and the Normalized SGD update).
-We analyze all 3 algorithms from a theoretical perspective and provide an extensive empirical study across diverse relevant tasks and settings.
+In this paper, we develop a new algorithm for optimization of functions of weight matrices, which are typical for training large language models.
+Changing spectral norm, which was used to derive Muon, to nuclear norm, we pose a new optimization problem for an update matrix, solution of which defines a novel algorithm we name Neon.
+To make it feasible, we use Lanczos algorithm to find a required step.
+After providing theoretical guarantees of Neon convergence, we compare performances of Neon, Muon, and Adam on training multilayer perceptron, convolutional neural network and NanoGPT.
+\end{abstract}
 
 \section{Idea}
-\iffalse
 The goal of the project is to make variations on Muon to speed it up.
@@
 Instead of using spectral norm, used to derive Muon, we utilize nuclear norm to produce a new optimization algorithm.
-\fi
+
 \subsection{Problem (Project description)}
@@
 % Removed template instructions sections
-\section{Related Work}
-Our review primarily focuses on the Muon and Shampoo optimizers, as our algorithm extends the ideas used to derive these methods.
-We highlight the advantages and disadvantages of these approaches, the unique effects they introduce, and compare them to Neon.
-
-\subsection{Muon optimizer}
-
-In the previous section, we described the theoretical foundation behind the weight update step in the Muon optimizer, but we did not discuss how to obtain the required matrices in practice.
-The update step is defined by \eqref{eqn:update_muon}, which requires $UV^T$ with $U$ and $V^T$ from the singular value decomposition of the gradient-like matrix $G$.
-A naive solution would involve computing the SVD of $G$ and constructing the required expression.
-However, the developers of the Muon optimizer introduced a workaround using Newton-Schulz iterations \cite{jordan2024muon}.
-The Newton-Schulz iterations from the original article \cite{jordan2024muon} require 10 matrix-matrix multiplications to achieve the desired accuracy.
-The asymptotic complexity of such an operation is identical to that of SVD and equals $O(mn \min\{m, n\})$ for an $m \times n$ matrix.
-Nevertheless, matrix multiplication on modern GPUs can be performed much more efficiently.
-
-The performance of Muon in training large language models was tested \cite{liu2025muon} against AdamW.
-The testing demonstrated excellent performance by Muon, which was approximately 2 times more efficient in terms of FLOPs required to reach a certain loss value.
-This is even more remarkable when considering the cost of one iteration: Muon requires an additional $O(mn \min\{m, n\})$ FLOPs per $m \times n$ matrix, while AdamW needs only $O(mn)$.
-
-Another interesting discovery about the Muon optimizer is that it accelerates grokking \cite{tveit2025muonoptimizeracceleratesgrokking}.
-In the test problem, Muon achieved grokking significantly faster than AdamW in terms of passed epochs, with a mean grokking epoch of 102.89 for Muon and 153.09 for AdamW.
-The authors suggest that this may be due to the fact that Muon stimulates broader exploration by orthogonalizing the gradient matrix, thus avoiding memorization.
-
-Recently, theoretical guarantees for Muon convergence have been derived \cite{li2025noteconvergencemuon}.
-In particular, in the $L$-smooth convex case, it achieves $O(1/T^{\frac{1}{2}})$ (with full gradient) and $O(1/T^{\frac{1}{4}})$ (with stochastic gradient) bounds on the Frobenius norm of the gradient or the mathematical expectation of the gradient norm, respectively, where $T$ is the number of iterations.
-
-\subsection{Shampoo optimizer}
-
-Another optimizer that exploits the matrix (and even tensor) structure of weights in neural networks is the Shampoo optimizer.
-We avoid a detailed description of this method here and refer the reader to the original article \cite{gupta2018shampoopreconditionedstochastictensor}, but we outline the key properties of Shampoo and its relation to the Muon optimizer.
-
-The Shampoo optimizer uses left and right preconditioning for the gradient-like matrix, leveling its spectrum.
-The preconditioners are computed from the exponentially averaged gradients, and their computation requires $O(n^3 + m^3)$ per $m \times n$ matrix.
-This exponential averaging is a key feature that provides several distinct interpretations for the preconditioners.
-They can be viewed as an approximation of the Gauss-Newton component of the Hessian or the first step of the power iteration algorithm for computing the optimal Kronecker product approximation \cite{morwani2025a}.
-With exponential averaging turned off, the update step of Shampoo can be simplified and becomes identical to that of Muon \cite{jordan2024muon}.
-
-Convergence analysis is presented in the original article \cite{gupta2018shampoopreconditionedstochastictensor}.
-Shampoo achieves $O(1/T^{\frac{1}{2}})$ convergence for the loss function value in the $L$-smooth convex case, where $T$ is the number of iterations.
-
-\subsection{Synthesis and Neon's position}
-
-Recent developments in optimization techniques show that utilizing the matrix structure of weights in neural networks can be very beneficial.
-Optimizers following this path converge faster in terms of iterations or epochs and often even FLOPs, but have a high iteration cost.
-Neon seeks to decrease the iteration cost while preserving fast convergence.
-
-While the uniqie advantages and effects introduced by Neon are yet to be discovered, we can already say that our new optimizer introduces additional overhead of $O(mn)$ FLOPS on average per $m \times n$ weight matrix, which is much better then $O(m n \min\{m, n\})$ for Muon optimizer and $O(n^3 + m^3)$ for Shampoo.
-
-\subsection{Optimization Strategies for Efficient Large Language Model Training}
-The unprecedented scale of modern Large Language Models (LLMs) has pushed traditional optimizers like AdamW \cite{Loshchilov2017FixingWD} to their limits in terms of computational efficiency and convergence speed \cite{liu2025muon,chen2025cosmoshybridadaptive}.
-This challenge has catalyzed research into more sophisticated optimization approaches that can maintain or improve performance while reducing training costs.
-
-The Muon optimizer  has emerged as a promising alternative based on matrix orthogonalization principles.
-Scaling Muon to billion-parameter LLMs required two key adaptations: the integration of L2 weight decay for stability and the implementation of per-parameter update scaling to handle diverse parameter distributions efficiently .
-Empirical evaluations demonstrate that Muon can match or exceed AdamW's model quality while requiring only about 52 of the training FLOPs.
-The successful training of the Moonlight model series, including a 16B-parameter Mixture-of-Experts model, validates Muon's practicality for production-scale applications \cite{liu2025muon}.
-
-Building on this foundation, hybrid approaches like COSMOS \cite{chen2025cosmoshybridadaptive} further enhance efficiency by combining optimization techniques based on gradient structure.
-COSMOS applies computationally intensive updates to a low-dimensional "leading eigensubspace" while using memory-efficient methods like Muon for the remaining parameters.
-This approach maintains convergence benefits while substantially reducing memory requirements.
-For distributed training environments, optimizers like Dion \cite{ahn2025dioncommunicationefficientoptimizerlarge} specifically target communication efficiency by minimizing data exchange between workers through distributed orthonormalization techniques.
 

### 2025-08-16 old_report/muon.tex
New abstract and preparation for an article

--- 2025-05-07 article/muon.tex
@@
 \usepackage{amssymb, amsmath}			% стилевой пакет для формул
+\usepackage{amsthm}                     % поддержка теорем
 \usepackage{algorithm}
@@
 \usepackage{subcaption}
+
+\newtheorem{lemma}{Lemma}
 
@@
 The goal of the project is to make variations on Muon to speed it up.
-Recently, authors of \cite{bernstein2024oldoptimizernewnorm} have proposed to look at different optimizers as the solution of the optimization problem for an update.
+Recently, authors of \cite{bernstein2024oldoptimizernewnorm} have proposed to derive update step for an optimizer as the solution of the certain optimization problem.
 This approach can be utilized to derive Muon \cite{jordan2024muon}, a novel algorithm for fast training of neural networks.
-Instead of using spectral norm in the problem, we use nuclear norm to produce a new problem.
+Instead of using spectral norm, used to derive Muon, we utilize nuclear norm to produce a new optimization algorithm.
 %Then, we add momentum to update.
@@
 For distributed training environments, optimizers like Dion \cite{ahn2025dioncommunicationefficientoptimizerlarge} specifically target communication efficiency by minimizing data exchange between workers through distributed orthonormalization techniques.
+
+\section{Derivation of update rules}
+
+\begin{lemma}\label{lem:opt_star}
+Let $G \in \mathbb{R}^{m \times n}$ and $\lambda > 0$. Then, the following optimization problem
+\begin{equation*} 
+  f(\delta W) = \langle G, \delta W \rangle + \lambda \norm{\delta W}_*^2 \to \min_{\delta W}
+\end{equation*}
+has solution 
+\begin{equation*}
+  \delta W = -\frac{1}{2\lambda} u_1 \sigma_1 v_1^T\,,
+\end{equation*}
+where $\sigma_1$ is the largest singular value of $G$, and $u_1$ and $v_1$ are the corresponding singular vectors.
+\end{lemma}
+
+\begin{proof}
+Let us denote $r = \min\{m, n\}$. Then by Von Neumann's trace inequality,
+\begin{equation*}
+    \vert\langle G, \delta W \rangle\vert \leq \sum_{i=1}^r \sigma_i(G)\sigma_i(\delta W)
+    \Rightarrow
+    \langle G, \delta W \rangle \geq -\sum_{i=1}^r \sigma_i(G)\sigma_i(\delta W)\,.
+\end{equation*}
+Thus, expressing nuclear norm through singular values, we can write down
+\begin{equation*}
+    f(\delta W) \geq -\sum_{i=1}^r \sigma_i(G)\sigma_i(\delta W) + \lambda\left(\sum_{i=1}^{r}\sigma_i(\delta W)\right)^2 \geq \min_{d_1,\dots,d_r \geq 0}-\sum_{i=1}^r \sigma_i(G)d_i + \lambda\left(\sum_{i=1}^{r}d_i\right)^2.
+\end{equation*}
+By Karush-Kuhn-Tucker theorem, necessary conditions of minimum are
+\begin{equation*}
+    \left(d_i \geq 0\ \text{and}\ -\sigma_i(G) + 2\lambda\sum_{j=1}^{r}d_j = 0\right)\ \text{or}\ \left(d_i = 0\ \text{and}\ -\sigma_i(G) + 2\lambda\sum_{j=1}^{r}d_j \geq 0\right) \,,\ i = 1,\dots,r\,.
+\end{equation*}
+These conditions simplify to 
+\begin{equation*}
+    \sum_{i \in S}d_i = \sigma_1(G)\,,\ \begin{cases}
+        d_i \geq 0 & \text{if}\ \sigma_i(G) = \sigma_1(G)\,, \\
+        d_i = 0 & \text{otherwise}\,.
+    \end{cases}
+\end{equation*}
+All points satisfying those conditions deliver minimum, and
+\begin{equation*}
+    f(\delta W) \geq -\frac{\sigma_1^2(G)}{4\lambda}\,.
+\end{equation*}
+
+Now let
+\begin{equation*}
+    \delta W^* = -\frac{1}{2\lambda} u_1 \sigma_1(G) v_1^T\,.
+\end{equation*}
+Inserting it to $f(\delta W)$ gives
+\begin{equation*}
+    f\left(\delta W^*\right) = -\frac{\sigma_1(G)^2}{2\lambda} + \frac{\sigma_1(G)^2}{4\lambda} = -\frac{\sigma_1(G)^2}{4\lambda}
+\end{equation*}
+This matches the derived lower bound. Thus, $\delta W^*$ minimizes $f(\delta W)$.
+\end{proof}
+
+\begin{lemma}\label{lem:opt_F_star}
+Let $G \in \mathbb{R}^{m \times n}$, $r = \min\{m. n\}$ and $\lambda > 0$. Then, the following optimization problem
+\begin{equation*} 
+  f(\delta W) = \langle G, \delta W \rangle + \lambda \norm{\delta W}_{F*}^2 \to \min_{\delta W}\,,
+\end{equation*}
+where $\norm{\cdot}_{F*}$ is defined in \eqref{eqn:F_star} has solution 
+\begin{equation}
+\delta W = -\frac{1}{\lambda}UDV^T
+\end{equation} 
+with $D = \text{diag}(d_i)$, where $d_i = [\sigma_i - \tau]_+$, and $\tau$ is given by
+\begin{equation}
+    \sum_{i=1}^r [\sigma_i - \tau]_+ = \tau\,.
+\end{equation}
+\end{lemma}
+
+\begin{proof}
+Analogously to the proof of Lemma~\ref{lem:opt_star}, we can use Von Neumann's trace inequality to write down
+\begin{equation*}
+    f(\delta W) \geq -\sum_{i=1}^r \sigma_i(G)\sigma_i(\delta W) + \frac{\lambda}{2}\left(\sum_{i=1}^{r}\sigma_i(\delta W)\right)^2 + \frac{\lambda}{2}\sum_{i=1}^{r}\sigma_i^2(\delta W)\,,
+\end{equation*}
+\begin{equation}\label{eqn:f_min_F_star}
+    f(\delta W) \geq \frac{1}{\lambda}\min_{d_1,\dots,d_r \geq 0}-\sum_{i=1}^r \sigma_i(G)d_i + \frac{1}{2}\left(\sum_{i=1}^{r}d_i\right)^2 + \frac{1}{2}\left(\sum_{i=1}^{r}d_i^2\right)\,.
+\end{equation}
+By Karush-Kuhn-Tucker theorem, neccesary conditions of minimum are
+\begin{equation*}
+    \left(d_i \geq 0\ \text{and}\ -\sigma_i(G) + \sum_{j=1}^{r}d_j +  d_i = 0\right)\ \text{or}\ \left(d_i = 0\ \text{and}\ -\sigma_i(G) + \sum_{j=1}^{r}d_j + d_i \geq 0\right) \,,\ i = 1,\dots,r\,.
+\end{equation*}
+Denoting $\tau = \sum_{i = 1}^r d_i$ gives $d_i = [\sigma_i(G) - \tau]_+$, where $\tau$ satisfies
+\begin{equation}
+    \sum_{i=1}^n [\sigma_i(G) - \tau]_+ = \tau\,.
+\end{equation}
+Inserting found minimum point into~\eqref{eqn:f_min_F_star} yields
+\begin{equation*}
+    f(\delta W) \geq -\sum_{i = 1}^r d_i(\tau + d_i) + \frac{\tau^2}{2\lambda} + \frac{1}{2\lambda}\sum_{i=1}^{r}d_i^2 = -\frac{1}{2 \lambda} \left( \tau^2 + \sum_{i=1}^r d_i^2 \right)\,.
+\end{equation*}
+
+Now let
+\begin{equation}
+\delta W^* = -\frac{1}{\lambda}UDV^T
+\end{equation} 
+with $D = \text{diag}(d_i)$. Inserting it to $f(\delta W)$ gives
+\begin{equation*}
+    f(\delta W^*) = -\sum_{i = 1}^r d_i(\tau + d_i) + \frac{\tau^2}{2\lambda} + \frac{1}{2\lambda}\sum_{i=1}^{r}d_i^2 = -\frac{1}{2 \lambda} \left( \tau^2 + \sum_{i=1}^r d_i^2 \right)\,.
+\end{equation*}
+This matches the derived lower bound. Thus, $\delta W^*$ minimizes $f(\delta W)$.
+\end{proof}
+
 \section{Quality metrics}

### 2025-08-16 old_report/proposal_template.tex
New abstract and preparation for an article

%% ============================================
%% ================ Preambule =================
%% ============================================
\documentclass[]{scrartcl}
\usepackage[margin = 0.5in]{geometry}

\usepackage[pdftex,unicode, 
colorlinks=true,
linkcolor = blue]{hyperref}	% нумерование страниц, ссылки!!!!ИМЕННО В ТАКОМ ПОРЯДКЕ СО СЛЕДУЮЩИМ ПАКЕТОМ
%\usepackage[warn]{mathtext}				% Поддержка русского текста в формулах
\usepackage[T1, T2A]{fontenc}			% Пакет выбора кодировки и шрифтов
\usepackage[utf8]{inputenc} 			% любая желаемая кодировка
\usepackage[english]{babel}		% поддержка русского языка
\usepackage{wrapfig}					% Плавающие картинки
\usepackage{amssymb, amsmath}			% стилевой пакет для формул
\usepackage{algorithm}
\usepackage{algorithmic}

\ifpdf
\usepackage{cmap} 				% чтобы работал поиск по PDF
\usepackage[pdftex]{graphicx}
%\usepackage{pgfplotstable}		% Для вставки таблиц.
\pdfcompresslevel=9 			% сжимать PDF
\else
\usepackage{graphicx}
\fi

\graphicspath{{./figures/}}
\usepackage{subcaption}
%% ============================================
%% ================ Info =================
%% ============================================
\title{Название проекта}
\author{\begin{tabular}{c c}
	  	 Добрыня Никитич & Daniil Merkulov \\
		 \texttt{voda@baikal.ru} & \texttt{daniil.merkulov@skoltech.ru} 
		\end{tabular}}
\date{\today}

\begin{document}

\maketitle

\begin{abstract}
Емкое описание предполагаемого проекта в один абзац. Емкое описание предполагаемого проекта в один абзац. Ëмкое описание предполагаемого проекта в один абзац. Это абстракт.
\end{abstract}

\section{Идея}

Обратите внимание на конкретность постановки задачи и её реалистичность. В этом разделе надо вкратце обозначить тему проекта - это повторение/вариация на тему уже готовой статьи, собственная идея, что-нибудь другое. Здесь можно привести немного контекста - например, обуславливающего актуальность проекта.

\subsection{Problem}

В этом разделе необходимо очень конкретно написать решаемую задачу. Понять постановку задачи должен человек с хорошим математическим бэкграундом, но без знаний по узкой теме проекта. В качестве валидации рекомендую дать одному или нескольким знакомым однокурсникам, друзьям и спросить - понятна ли постановка задачи при условии, что они прочитают её так как написано без ваших дополнительных комментариев?

Напишем и процитируем \cite{kaczmarz1951theorie}:  какую нибудь формулу 
\begin{equation}
x_{k+1} = x_k  + \dfrac{b_{r(i)} - \langle a_{r(i)}, x_k\rangle}{\|a_{r(i)}\|_2^2}a_{r(i)}
\end{equation}

Напишем какой нибудь алгоритм

\begin{algorithm}[h!]
	\caption{Randomized Kaczmarz Algorithm}
	\label{RKalg}
	\begin{algorithmic}[1]
		\STATE Initialize $k \leftarrow 0$
		\FOR {$k=0,1,...$}
		\STATE Select row $j$ from $\{1,2,...m\}$ with probability $\frac{||a_j||^2}{||A||_F^2}$
		\STATE Project $x_{k+1} = x_k + \frac{(b_j - a_j^T x_k)}{||a_j||^2} a_j$
		\STATE Update $k \leftarrow k+1$
		\ENDFOR
	\end{algorithmic}
\end{algorithm}

\section{Outcomes}

Опишите, что конкретно будет выходом Вашего проекта (код, теорема, численные эксперименты, телеграм бот, веб сайт, приложение, рассказ).

\begin{figure}[h!]
	\center{\includegraphics[width=0.3\linewidth]{test.pdf}}
	\caption{Картинка, иллюстрирующая идею / выход проекта}
\end{figure}

\section{Литературный обзор}
Краткий обзор релевантных источников по теме, со ссылками на них - минимальное число источников - 7.

В работе \cite{strohmer2009randomized} исследовалось то-то, и даже есть формула:
\begin{equation}
\mathbb{E} \|x_k - x\|_2^2 \leq \left(1 - \kappa (A) ^{-2}\right)^k \|x_0 - x\|_2^2
\end{equation}

А вот в этих работах другой очень интересный взгляд \cite{gower2015randomized}, \cite{dai2014randomized}

\section{Метрики качества}
Приведите формальные и измеряемые показатели, по которым можно оценивать Ваше решение\ проект - это могут быть конкретные метрики качества алгоритмов, соц. опрос, логическое доказательство и т.д. Наша задача в этом разделе - суметь легко ответить на вопрос работает ли данная идея в конкретных задачах или нет (отрицательный результат тоже результат)

\section{Примерный план}
\begin{itemize}
	\item Сначала это к такому то сроку
	\item Потом вот это
	\item И это - программа минимум
	\item А вот это  - если успею программу минимум
\end{itemize}

\section{Отчёт о фазе прототипирования}
Мы попробовали провести эксперименты с существующим кодом - у нас получились такие результаты, попробовали немного изменить постановку, всё сломалось. Вот скриншоты. Вот потребление памяти. Вот это место, в котором неравенство не верно.

\bibliographystyle{unsrt}
\bibliography{biblio}

\end{document}

### 2025-08-16 conf_article/neon.tex
Rough plan

@@
 In this article, we develop a family of 3 new algorithms for optimizing functions of weight matrices that arise in training large language models.
-By replacing the spectral norm, which is used when deriving Muon within the linear optimization oracle framework (LMO), with the nuclear norm, the mean between the nuclear norm and the Frobenius norm we name the F*-norm, and the dual to F*-norm we name F2-norm, we obtain 3 new optimization problems for the update matrix; their solutions define 3 novel algorithms we name Neon, F*-Neon, and F2-Muon respectively.
+By replacing the spectral norm, which was used to derive Muon within the linear minimization oracle (LMO) framework, with the nuclear norm, the mean between the nuclear norm and the Frobenius norm we name the F*-norm, and the dual to F*-norm we name F2-norm, we obtain 3 new optimization problems for the update matrix; their solutions define 3 novel algorithms we name Neon, F*-Neon, and F2-Muon respectively.
 We derive the optimal solutions to these problems and propose a numerically efficient method to compute updates for Neon and F*-Neon via the Lanczos algorithm (while an update of F2-Muon turns out to be a mean between the Muon update and the Normalized SGD update).
 We analyze all 3 algorithms from a theoretical perspective and provide an extensive empirical study across diverse relevant tasks and settings.
+\end{abstract}
+
+\section{Introduction}
+* Hype of Muon and others: 
+- Importance of the functions of a matrix weight. Special structure.
+- Write the equation
+- Memory efficiency and hyperparameter transferability
+- Proven efficient computing (cite benchmarks and Dion and Cosmos)
+* Comparison:
+- Shampoo (very compute intensive for large layers, we need asymptotics!), second order
+- LMO has produced several effective algorithms: Muon, Dion, Scion, Gluon (citations)
+- generalization is better for Muon and others (citations!). Possibly nuSAM here (definitely)
+- Scion and Gluon for multilayer
+- Gluon for (L0, L1)-smooth problems, which is more realistic (low L0, faster convergence - no impractically small stepsizes)
+* Analyis:
+- Several generalizations have been already presented by the community: Old Optimizer and New Norm (idea and evidence that it explains famous algorithms); Pethick (LMO and different layers); Kovalev (as trust region and arbitrary norm); Gluon (generalized smoothness model).
+- However, to the best of our knowledge, no attempt has been made to try other norms. Here we come in.
+- No attempt to mitigate high computation cost of matrix computation has been made. Here again we come in.
+- It has been shown that $O(1/esp^4)$ bounds for Muon, Scion, Gluon and even Adam for L-smooth functions (citations).
+Some bounds depend on the dimensionality of space (cite Kovalev), which for spectral norm gives much more iterations (cite Kovalev)
+- This motivates us to try new norms
+
+\section{Linear Minimization Oracle and Nuclear norm}
+\subsection{Problem statement}
+Norm is not necessarily induced by the inner product.
+Bernstein vs LMO (choose LMO as more correct)
+\subsection{Nuclear Neon}
+Problem and update
+\subsection{F*-Neon}
+Motivation: we need updates of higher rank. Hence, F* (why? regularization?)
+\subsection{F2-Muon}
+Motivation: if the spectral norm (Muon) is dual to the nuclear norm (Neon), then what is dual to F*? F2-Muon! Derivation.
+
+\section{Matrix speedup}
+\subsection{Theory}
+Not only formulas, but also citations!
+Solutions to efficiently find some parts of SVD:
+- Basic SVD (complexity asymptotic!)
+- Newton-Schulz (complexity asymptotic!)
+- Our job: Power iterations, Randomized SVD and Lanczos
+\subsection{Experiments}
+Nikolay's task: SVD, RSVD, Lanczos, Newton-Schulz. We use torch and cupy to test it.
+Options: all possible on torch, or additionally to compare all on cupy. The goal: to find how much do we lose due to cupy
+
+\section{"Smooth" theory via Trust Region bounds}
+* Probably statement of the trust region problem
+* Assumption on unbiasedness and limited variance of stoch. grad.
+* Assumption on L-smoothness as in the Kovalev's article.
+* Assumption on rho. To be more precise, the statement of lemmas on rho for all 4 norms.
+* We utilize the Kovalev's corollary to derive convergence
+* Bonus: convert all the norms into a common sense (L2?) norm to get the true dependency on dimensionality
+
+\section{Experiments}
+\subsection{Synthetic quadratics}
+Description of how to construct L-smooth functions in a given norm, and how to estimate variance of the gradient. Then check of the 
+\subsection{MLP}
+Deep learning?
+\subsection{Benchmarks}
+CNN benchmark
+NanoGPT benchmark
+
+\section{Conclusion}
+* Future work: (L0, L1)-smoothness, probably extrapolation (to make dependency on eps milder)
+\section{Appendix}
+\subsection{Proofs}
+\subsection{Norm derivations}
+* Dual to F* norm
+* Bounds on rho for all 4 norms (with their tighness if possible)
+\subsection{Updates derivations}
+%yep, many KKT and so on
+- Neon
+- F*-Neon
+- F2-Muon
+* F* choice from KKT
+\subsection{}
+~ Note: nuSAM and their update. They had no Lanczos
+Technical details on experiments
 
@@
 \appendix
-\section{Appendix}
+\section{Old Appendix}
 You may include other additional sections here.

### 2025-08-19 conf_article/neon.tex
Introduction and problem statement drafts

--- 2025-08-16 conf_article/neon.tex
@@
 
-\title{From Muon to Neon: Introducing Nuclear Norm to Large Matrices}
+%\title{From Muon to Neon: Introducing Nuclear Norm to Large Matrices}
+\title{The Nuclear Norm and Beyond: Dual Norms and Combinations for Matrix Optimization}
 
@@
 %\icompfinalcopy % Uncomment for camera-ready version, but NOT for submission.
+
+%------------------------------------------------------------------------------------
+\newcommand{\Rmn}{\R^{m\times n}}
+\newcommand{\cD}{\mathcal{D}}
+\newcommand{\cS}{\mathcal{S}}
+\newcommand{\cB}{\mathcal{B}}
+\newcommand{\Ed}[2]{\mathbb{E}_{#1}\left[#2\right]}
+\usepackage{mathtools}
+\DeclarePairedDelimiter{\sqne}{\|}{\|_2^2}
+\DeclarePairedDelimiter{\norme}{\|}{\|_2}
+\DeclarePairedDelimiter{\normf}{\|}{\|_\mathrm{F}}
+\DeclarePairedDelimiter{\sqns}{\|}{\|_{\mathrm{op}}^2}
+\DeclarePairedDelimiter{\norms}{\|}{\|_{\mathrm{op}}}
+\DeclarePairedDelimiter{\sqnn}{\|}{\|_{\mathrm{nuc}}^2}
+\DeclarePairedDelimiter{\sqnf}{\|}{\|_{\mathrm{F}}^2}
+\DeclarePairedDelimiter{\normn}{\|}{\|_{\mathrm{nuc}}}
+\def\<#1,#2>{\langle #1,#2\rangle}
+\DeclarePairedDelimiter{\dotprod}{\langle}{\rangle}
+
+\usepackage{cleveref}
+\usepackage{environ}
+\newcounter{aequation}
+\NewEnviron{aequation}{\refstepcounter{aequation}$$\BODY\eqno{\text{(A\theaequation)}}$$}
+\crefname{aequation}{assumption}{assumptions}
+\creflabelformat{aequation}{(#2A#1#3)}
+
+\usepackage{amsmath}
+\DeclareMathOperator{\tr}{tr}
+
+%------------------------------------------------------------------------------------
+
 \begin{document}
@@
 \begin{abstract}
-In this article, we develop a new algorithm for optimizing functions of weight matrices that arise in training large language models.
-By replacing the spectral-norm regularization used in Muon with a nuclear-norm formulation, we obtain a new optimization problem for the update matrix; its solution defines a novel algorithm we call Neon.
-We derive the optimal solution to this problem and propose a numerically efficient method to compute it via the Lanczos algorithm.
-We analyze Neon from a theoretical perspective and provide an extensive empirical study across diverse relevant tasks and settings.
+In this article, we explore the use of matrix norms for optimizing functions of weight matrices, a crucial problem in training large language models.
+Moving beyond the spectral norm that underlies the Muon update, we leverage the nuclear norm, its affine combinations with other norms, and their corresponding duals to develop a new family of Muon-like algorithms.
+We complement our theoretical analysis with an extensive empirical study of the algorithms across a wide range of tasks and settings.
+\end{abstract}
 
-New variant:
-\iffalse This article investigates utilization of matrix norms besides the spectral norm in linear optimization oracle (LMO) framework.
-Making use of the nuclear norm, the mean between the nuclear norm and the Frobenius norm we name the F*-norm, and the dual to F*-norm, we develop 3 new algorithms for optimizing functions of weight matrices that arise in training large language models.\fi
+\section{Introduction}
+ Minimizing loss functions in unprecedentedly high-dimensional spaces has recently become an integral and crucial part in training large language models.
+Hence, new scalable, time- and memory-efficient algorithms have been demanded.
+Besides well-known Adam and AdamW (citations), recently proposed Muon has shown promising results on different benchmarks (CIFAR, NanoGPT, citations).
+Its key difference from earlier algorthims is that it has been constructed specifically for optimizing functions of weight matrices.
+Such functions are common in modern machine learning (citations), so it does not significantly restrict Muon's applicability to modern problems.
 
-In this article, we develop a family of 3 new algorithms for optimizing functions of weight matrices that arise in training large language models.
-By replacing the spectral norm, which is used when deriving Muon within the linear optimization oracle framework (LMO), with the nuclear norm, the mean between the nuclear norm and the Frobenius norm we name the F*-norm, and the dual to F*-norm we name F2-norm, we obtain 3 new optimization problems for the update matrix; their solutions define 3 novel algorithms we name Neon, F*-Neon, and F2-Muon respectively.
-We derive the optimal solutions to these problems and propose a numerically efficient method to compute updates for Neon and F*-Neon via the Lanczos algorithm (while an update of F2-Muon turns out to be a mean between the Muon update and the Normalized SGD update).
-We analyze all 3 algorithms from a theoretical perspective and provide an extensive empirical study across diverse relevant tasks and settings.
+ That is what can be said from a practical point of view.
+From the perspective of theory, Muon's main innovation was an intentional usage of matrix norms, i.e.
+the spectral norm, to derive the algorithm's update (cite Deriving Muon).
+Several other attempts have been since made to construct new algorithms, mainly generalizations of Muon's paradigm (cite Scion and Gluon).
+
+ Based upon recent theoretical advances that explain some theory behind Muon, Scion and Gluon (cite), we explore application of other matrix norms to optimization of functions of matrices.
+As it has been done with Muon, we stipulate that our algorithms' updates be fast to compute.
+\iffalse
+\begin{itemize}
+    \item Hype of Muon and others:
+    \begin{itemize}
+        \item Importance of the functions of a matrix weight. Special structure.
+        \item Write the equation
+        \item Memory efficiency and hyperparameter transferability
+        \item Proven efficient computing (cite benchmarks and Dion and Cosmos)
+    \end{itemize}
+    \item Comparison:
+    \begin{itemize}
+        \item Shampoo (very compute intensive for large layers, we need asymptotics!), second order
+        \item LMO has produced several effective algorithms: Muon, Dion, Scion, Gluon (citations)
+        \item generalization is better for Muon and others (citations!). Possibly nuSAM here (definitely)
+        \item Scion and Gluon for multilayer
+        \item Gluon for (L0, L1)-smooth problems, which is more realistic (low L0, faster convergence - no impractically small stepsizes)
+    \end{itemize}
+    \item Analyis:
+    \begin{itemize}
+        \item Several generalizations have been already presented by the community: Old Optimizer and New Norm (idea and evidence that it explains famous algorithms); Pethick (LMO and different layers); Kovalev (as trust region and arbitrary norm); Gluon (generalized smoothness model).
+        \item However, to the best of our knowledge, no attempt has been made to try other norms. Here we come in.
+        \item No attempt to mitigate high computation cost of matrix computation has been made. Here again we come in.
+        \item It has been shown that $O(1/esp^4)$ bounds for Muon, Scion, Gluon and even Adam for L-smooth functions (citations).
+Some bounds depend on the dimensionality of space (cite Kovalev), which for spectral norm gives much more iterations (cite Kovalev)
+        \item This motivates us to try new norms
+    \end{itemize}
+\end{itemize}
+\fi
+
+\section{Problem Statement}
+\subsection{Function of a matrix and Assumptions}
+Warning: the text was copied, while formulas were adapted. We need to rephrase the subsection! 
+  We consider the problem of minimizing a differentiable matrix function $F(\cdot)\colon \Rmn \to \R$
+  \begin{equation}\label{eq:mat}
+    \min_{\mX \in \Rmn} f(\mX).
+  \end{equation}
+
+To make the theoretical analysis possible in the future, we make three reasonable non-restrictive assumptions.
+Idea: move gradient assumptions and smoothness to the L-smooth part. Here we need only custom norm and their relation to the Frobenius norm.
+
+{\bf Stochastic gradient estimator.}
+We assume access to a stochastic estimator $g(\cdot\,;\xi)\colon \Rmn \to \Rmn$ of the gradient $\nabla f(\cdot)$, where $\xi \sim \cD$ is a random variable sampled from a probability distribution $\cD$.
+We assume that the stochastic gradient estimator $g(\cdot;\xi)$ is unbiased and has bounded variance, that is, the following relations hold:
+\begin{aequation}\label{eq:variance}
+  \Ed{\xi \sim \cD}{g(\mX;\xi)} = \nabla f(\mX)
+  \quad\text{and}\quad
+  \Ed{\xi \sim \cD}{\sqnf{g(\mX;\xi) - \nabla f(\mX)}} \leq \sigma^2
+  \quad\text{for all}\;
+  \mX \in \Rmn,
+\end{aequation}
+where $\sigma > 0$ is a positive variance parameter, and $\normf{\cdot}$ is the standard Euclidean, i.e.
+Frobenius, norm induced by the inner product $\<\cdot,\cdot>$, i.e., $\normf{\mX} = \sqrt{\<\mX,\mX>}=\sqrt{\tr(\mX^\top \mX)}$.
+These assumptions have been widely adopted for the analysis of many stochastic gradient optimization algorithms \citep{ghadimi2013stochastic,ghadimi2016accelerated,cutkosky2020momentum,sun2023momentum,horvath2023stochastic,gorbunov2020linearly}.
+
+{\bf Non-Euclidean norm setting and Lipschitz continuous gradient.}
+We assume that matrix space $\Rmn$ is equipped with a norm $\norm{\cdot}\colon \Rmn \to \R_+$, which possibly does not coincide with the Frobenius, norm $\normf{\cdot}$.
+In addition, we assume that the gradient $\nabla f(\cdot)$ is Lipschitz continuous with respect to the norm $\norm{\cdot}$, that is, the following inequality holds:
+\begin{aequation}\label{eq:L}
+  \norm{\nabla f(\mX) - \nabla f(\mX')}_* \leq L \norm{\mX-\mX'}
+  \quad\text{for all}\;
+  \mX, \mX' \in \Rmn,
+\end{aequation}
+where $L > 0$ is the gradient Lipschitz constant, and $\norm{\cdot}^\dagger\colon \Rmn \to \R_+$ is the dual norm associated with $\norm{\cdot}$, i.e., $\norm{\mX}^\dagger = \sup_{\norm{\mX'}\leq 1} \<\mX,\mX'>$ for all $\mX \in \Rmn$.
+The assumption of gradient Lipschitz continuity is also widespread in the analysis of first-order optimization methods \citep{ghadimi2013stochastic,gower2019sgd,cutkosky2020momentum,horvath2023stochastic,gorbunov2020linearly}.
+It is important to highlight that while \Cref{eq:L} uses the dual norm $\norm{\cdot}^\dagger$ to measure the difference between the gradients, the variance in \Cref{eq:variance} is measured with respect to the Frobenius norm $\sqnf{\cdot}$, which is necessary to properly utilize the unbiasedness property of the stochastic gradient estimator $g(\cdot;\xi)$.
+Therefore, we need to provide a connection between these norms using the following inequality:
+\begin{aequation}\label{eq:norm}
+  \norm{\mX}^\dagger \leq \rho\cdot\normf{\mX}
+  \quad\text{for all}\;
+  \mX \in \Rmn,
+\end{aequation}
+where $\rho > 0$ is a positive constant.
+Note that such a constant always exists due to the norm equivalence theorem, which always holds in the finite-dimensional space $\Rmn$.
+We recount $\rho$ for different norms $\norm{\cdot}$ in the appendix.
+
+\subsection{Linear Minimization Oracle}
+Let us look at the problem from the perspective of linear minimization oracle (lmo) and unconstrained stochastic conditional gradient descent (uSCG) (\cite{pethick2025training}). lmo is defined as:
+\begin{equation}\label{eq:lmo}
+    \mathrm{lmo}(\mS) \in \arg\min_{\mX \in \cS} \<\mS, \mX>,
+\end{equation}
+where $\cS$ is some set. We are interested in the case when $\cS$ is a ball in our $\norm{\cdot}$ norm:
+\begin{equation}\label{eq:ball}
+    \cS := \cB_\eta :=\{ \mX \mid \norm{\mX} \leq \eta \}.
+\end{equation}
+
+uSCG update is defined as:
+$\mX^{k+1} = \mX^{k} + \gamma_k lmo(\mM^k),$ where $\mM^{k+1} = (1-\alpha_{k+1})\mM^{k} + \alpha_{k+1}g(\mX^k, \xi_k)$ is a momentum.
+
+It can be easily shown that the formula is equvalent to
+\begin{equation}\label{eq:trust_update}
+    \mX^{k+1} = \mX^{k} - \gamma_k\eta \arg\max_{\mX \in \cB_1} \<\mS, \mX> = \mX^{k} - \gamma_k\eta \{\Delta \in \cB_1 \mid \<\mM^k, \Delta> = \norm{\mM^k}^\dagger\}.
+\end{equation}
+
+Let us set $\gamma_k = 1$.
+This transforms algorithm defined by \cref{eq:trust_update} into Algorithm 1 from \cite{kovalev2025understanding}.
+Therefore, we can view \cref{eq:trust_update} both as an lmo-based algorithm and as a trust-region algorithm.
+
+\subsection{Nuclear Neon}
+Problem and update
+\subsection{F*-Neon}
+Motivation: we need updates of higher rank. Hence, F* (why? regularization?)
+\subsection{F2-Muon}
+Motivation: if the spectral norm (Muon) is dual to the nuclear norm (Neon), then what is dual to F*? F2-Muon! Derivation.
+
+\section{Matrix side of updates}
+\subsection{Theory}
+Not only formulas, but also citations!
+Solutions to efficiently find some parts of SVD:
+\begin{itemize}
+    \item Basic SVD (complexity asymptotic!)
+    \item Newton-Schulz (complexity asymptotic!)
+    \item Our job: Power iterations, Randomized SVD and Lanczos
+\end{itemize}
+\subsection{Experiments}
+Nikolay's task: SVD, RSVD, Lanczos, Newton-Schulz. We use torch and cupy to test it.
+Options: all possible on torch, or additionally to compare all on cupy. The goal: to find how much do we lose due to cupy
+
+\section{"Smooth" theory via Trust Region bounds}
+\begin{itemize}
+    \item Probably statement of the trust region problem
+    \item Assumption on unbiasedness and limited variance of stoch. grad.
+    \item Assumption on L-smoothness as in the Kovalev's article.
+    \item Assumption on rho. To be more precise, the statement of lemmas on rho for all 4 norms.
+    \item We utilize the Kovalev's corollary to derive convergence
+    \item Bonus: convert all the norms into a common sense (L2?) norm to get the true dependency on dimensionality
+\end{itemize}
+
+\section{Experiments}
+\subsection{Synthetic quadratics}
+Description of how to construct L-smooth functions in a given norm, and how to estimate variance of the gradient. Then check of the theory.
+It must be the only experiment where we might beat vanilla Muon.
+\subsection{MLP}
+Deep learning?
+\subsection{Benchmarks}
+CNN benchmark
+NanoGPT benchmark
+
+%related work
+
+    \section{Conclusion}
+\begin{itemize}
+    \item Future work: (L0, L1)-smoothness, probably extrapolation (to make dependency on eps milder, but we don't to spend memory here so never mind), hyperparameter-free (Bernstein2023)
+\end{itemize}
+\section{Appendix}
+\subsection{Proofs}
+\subsection{Norm derivations}
+\begin{itemize}
+    \item Dual to F* norm
+    \item Bounds on rho for all 4 norms (with their tighness if possible)
+\end{itemize}
+\subsection{Updates derivations}
+%yep, many KKT and so on
+\begin{itemize}
+    \item Neon
+    \item F*-Neon
+    \item F2-Muon
+    \item F* choice from KKT
+\end{itemize}
+\subsection{}
+~ Note: nuSAM and their update. They had no Lanczos
+Technical details on experiments
 
@@
 \begin{equation}\label{eqn:F_star}
-    \norm{X}_{F*}^2 = \frac{\norm{X}_F + \norm{X}_*}{2}\,,
+    \norm{X}_{F*}^2 = \frac{\norm{X}_F^2 + \norm{X}_*^2}{2}\,,
 \end{equation}

### 2025-08-19 conf_article/neon.tex
Updates included

@@
 \DeclarePairedDelimiter{\sqne}{\|}{\|_2^2}
-\DeclarePairedDelimiter{\norme}{\|}{\|_2}
+%\DeclarePairedDelimiter{\norme}{\|}{\|_2}
 \DeclarePairedDelimiter{\normf}{\|}{\|_\mathrm{F}}
+\DeclarePairedDelimiter{\normfstar}{\|}{\|_\mathrm{F*}}
+\DeclarePairedDelimiter{\normftwo}{\|}{\|_\mathrm{F2}}
 \DeclarePairedDelimiter{\sqns}{\|}{\|_{\mathrm{op}}^2}
@@
 \DeclareMathOperator{\tr}{tr}
+\DeclareMathOperator{\diag}{diag}
 
@@
 
-\subsection{Linear Minimization Oracle}
+\subsection{Linear Minimization Oracle and Trust Region}
 Let us look at the problem from the perspective of linear minimization oracle (lmo) and unconstrained stochastic conditional gradient descent (uSCG) (\cite{pethick2025training}). lmo is defined as:
@@
 It can be easily shown that the formula is equvalent to
-\begin{equation}\label{eq:trust_update}
+\begin{equation}\label{eq:almost_trust_update}
     \mX^{k+1} = \mX^{k} - \gamma_k\eta \arg\max_{\mX \in \cB_1} \<\mS, \mX> = \mX^{k} - \gamma_k\eta \{\Delta \in \cB_1 \mid \<\mM^k, \Delta> = \norm{\mM^k}^\dagger\}.
@@
 Let us set $\gamma_k = 1$.
-This transforms algorithm defined by \cref{eq:trust_update} into Algorithm 1 from \cite{kovalev2025understanding}.
-Therefore, we can view \cref{eq:trust_update} both as an lmo-based algorithm and as a trust-region algorithm.
+This transforms algorithm defined by \cref{eq:almost_trust_update} into Algorithm 1 from \cite{kovalev2025understanding}.
+Therefore, we can view \cref{eq:almost_trust_update} both as an lmo-based algorithm and as a trust-region algorithm.
+
+\section{Different Norms $\norm{\cdot}$ imply different updates}
+% As we have stated in Introduction, we are interested foremost in matrix norms, so we'll start right with them, skipping RMS-RMS norm and others.
+
+Based on different norms $\norm{\cdot}$, we simplify the update defined by the aforementioned equation:
+
+\begin{equation}\label{eq:our_update}
+    \mX^{k+1} = \mX^{k} - \eta \{\Delta \in \cB_1 \mid \<\mM^k, \Delta> = \norm{\mM^k}^\dagger\}
+\end{equation}
+
+In all the work, we define $\mU$, $\mSigma$, $\mV^\top$ as components of the singular value decomposition of $\mM^k$: $\mM^k = \mU \mSigma \mV^\top$.
+We use common notations: $\mU = [u_1, u_2, \dots, u_r]$, $\mSigma = \diag(\sigma_1, \sigma_2, \dots, \sigma_r)$, and  $\mV = [v_1, v_2, \dots, v_r]$.
+
+{\bf $\normf{\mM^{k}}$ and NSGD}
+
+Lemma. When $\norm{\cdot} = \normf{\cdot}$, \cref{eq:our_update} turns into:
+
+\begin{equation}\label{eq:nsgd_update}
+    \mX^{k+1} = \mX^{k} - \eta \frac{\mM^k}{\normf{\mM^k}}
+\end{equation}
+
+It is an interesting observation, because in other works (\cite{pethick2025training}), $\normf{\cdot}$ was used to recover SGD. The difference is in how one states the problem.
+
+{\bf $\norms{\mM^{k}}$ and Muon}
+
+Lemma. When $\norm{\cdot} = \norms{\cdot}$, \cref{eq:our_update} turns into:
+
+\begin{equation}\label{eq:muon_update}
+    \mX^{k+1} = \mX^{k} - \eta U V^\top
+\end{equation}
+
+{\bf $\normn{\mM^{k}}$ and Neon}
+
+Lemma. When $\norm{\cdot} = \normn{\cdot}$, \cref{eq:our_update} turns into:
+
+\begin{equation}\label{eq:neon_update}
+    \mX^{k+1} = \mX^{k} - \eta u_1 v_1^\top
+\end{equation}
+
+We name the derived algorithm \emph{Neon}.
+
+{\bf $\normfstar{\mM^{k}}^\dagger$ and F2-Muon}
+We define $\normfstar{\cdot}$ as a convex combination of $\normn{\cdot}$ and $\normf{\cdot}$:
+
+\begin{equation}\label{eq:fstar_norm}
+    \normfstar{\mX} = \alpha \normn{\mX} + (1-\alpha)\normf{\mX},
+\end{equation}
+
+where $\alpha \in [0, 1]$ defines a specific norm of $F*$-family.
+
+Lemma. When $\norm{\cdot} = \normfstar{\cdot}^\dagger$, \cref{eq:our_update} turns into:
+
+\begin{equation}\label{eq:nsgd_muon_update}
+    \mX^{k+1} = \mX^{k} - \eta (\alpha U V^\top + (1-\alpha)\frac{\mM^k}{\normf{\mM^k}})
+\end{equation}
+
+We name the derived algorithm \emph{F2-Muon}.
+It turns out that F2-Muon is a convex combination of Normalized SGD and Muon, which is curious.
+The implications are significant and discussed in the following sections.
+
+{\bf $\normftwo{\mM^{k}}^\dagger$ and F2-Neon}
+We define $\normftwo{\cdot}$ as a convex combination of $\norms{\cdot}$ and $\normf{\cdot}$:
+
+\begin{equation}\label{eq:ftwo_norm}
+    \normftwo{\mX} = \alpha \norms{\mX} + (1-\alpha)\normf{\mX},
+\end{equation}
+
+where $\alpha \in [0, 1]$ defines a specific norm of $F2$-family.
+
+Lemma. When $\norm{\cdot} = \normftwo{\cdot}^\dagger$, \cref{eq:our_update} turns into:
+
+\begin{equation}\label{eq:nsgd_neon_update}
+    \mX^{k+1} = \mX^{k} - \eta (\alpha u_1 v_1^\top + (1-\alpha)\frac{\mM^k}{\normf{\mM^k}})
+\end{equation}
+
+We name the derived algorithm \emph{F2-Neon}.
+It turns out that F2-Neon is a convex combination of Normalized SGD and Neon, which is curious.
+The implications are significant and discussed in the following sections.
+
+TO-DO: F*-neon? (it would be hard to do it)
 

### 2025-08-19 conf_article/neon.tex
Added proofs

@@
 \subsection{Linear Minimization Oracle and Trust Region}
-Let us look at the problem from the perspective of linear minimization oracle (lmo) and unconstrained stochastic conditional gradient descent (uSCG) (\cite{pethick2025training}). lmo is defined as:
+    Let us look at the problem from the perspective of linear minimization oracle (lmo) and unconstrained stochastic conditional gradient descent (uSCG) (\cite{pethick2025training}).
+lmo is defined as:
     \begin{equation}\label{eq:lmo}
@@
 
-We name the derived algorithm \emph{Neon}.
+        We name the derived algorithm \emph{Neon}. In the section Matrix side of updates, we will discuss how to compute an update efficiently.
 
-{\bf $\normfstar{\mM^{k}}^\dagger$ and F2-Muon}
+    {\bf $\normfstar{\mM^{k}}^\dagger$ and F-Muon}
         We define $\normfstar{\cdot}$ as a convex combination of $\normn{\cdot}$ and $\normf{\cdot}$:
@@
 
-We name the derived algorithm \emph{F2-Muon}.
-It turns out that F2-Muon is a convex combination of Normalized SGD and Muon, which is curious.
+        We name the derived algorithm \emph{F-Muon}.
+It turns out that F-Muon is a convex combination of Normalized SGD and Muon, which is curious.
 The implications are significant and discussed in the following sections.
 
-{\bf $\normftwo{\mM^{k}}^\dagger$ and F2-Neon}
+    {\bf $\normftwo{\mM^{k}}^\dagger$ and F-Neon}
         We define $\normftwo{\cdot}$ as a convex combination of $\norms{\cdot}$ and $\normf{\cdot}$:
@@
 
-where $\alpha \in [0, 1]$ defines a specific norm of $F2$-family.
+        where $\alpha \in [0, 1]$ defines a specific norm of F2-family.
 
@@
 
-We name the derived algorithm \emph{F2-Neon}.
-It turns out that F2-Neon is a convex combination of Normalized SGD and Neon, which is curious.
+        We name the derived algorithm \emph{F-Neon}.
+It turns out that F-Neon is a convex combination of Normalized SGD and Neon, which is curious.
 The implications are significant and discussed in the following sections.
@@
 
-\subsection{Nuclear Neon}
-Problem and update
 \subsection{F*-Neon}
 Motivation: we need updates of higher rank. Hence, F* (why? regularization?)
-\subsection{F2-Muon}
-Motivation: if the spectral norm (Muon) is dual to the nuclear norm (Neon), then what is dual to F*? F2-Muon! Derivation.
+%\subsection{F-Muon}
+%Motivation: if the spectral norm (Muon) is dual to the nuclear norm (Neon), then what is dual to F*? F-Muon! Derivation.
 
@@
 \section{Appendix}
-\subsection{Proofs}
 \subsection{Norm derivations}
-\begin{itemize}
-    \item Dual to F* norm
-    \item Bounds on rho for all 4 norms (with their tighness if possible)
-\end{itemize}
+% \begin{itemize}
+%    \item Dual to F* norm
+%    \item Bounds on rho for all 4 norms (with their tighness if possible)
+% \end{itemize}
 \subsection{Updates derivations}
-%yep, many KKT and so on
-\begin{itemize}
-    \item Neon
-    \item F*-Neon
-    \item F2-Muon
-    \item F* choice from KKT
-\end{itemize}
+
+Derivation of \cref{eq:muon_update} follows from \cref{eq:nsgd_muon_update} with $\alpha=1$. Indeed, $\norms{\cdot}^\dagger=1\cdot\normn{\cdot} + 0 \cdot \normf{\cdot} $.
+
+Derivation of \cref{eq:nsgd_muon_update}:
+Since $\norm{\cdot}^\dagger = \normfstar{\cdot}^{\dagger\dagger} = \normfstar{\cdot}$, the goal is to reach $\normfstar{\mM^k} = \alpha \tr{\mSigma} + (1-\alpha) \normf{\mM^k}$.
+
+Let us note that $\Delta = \alpha (\mU \mV^\top) + (1-\alpha) \frac{\mM^k}{\normf{\mM^k}}$ delivers this value.
+Indeed, by the trace property, $\<\mM^k, \Delta> =\<\mU \mSigma \mV^\top, \alpha \mU \mV^\top + (1-\alpha) \frac{\mU \mSigma \mV^\top}{\normf{\mM^k}}> = \alpha \tr \mSigma + (1-\alpha) \normf{\mM^k} = \normfstar{\mM^k}$, which completes the proof.
+
+Derivation of \cref{eq:neon_update} follows from \cref{eq:nsgd_neon_update} with $\alpha=1$. Indeed, $\normn{\cdot}^\dagger=1\cdot\norms{\cdot} + 0 \cdot \normf{\cdot} $.
+
+Derivation of \cref{eq:nsgd_neon_update}:
+Since $\norm{\cdot}^\dagger = \normftwo{\cdot}^{\dagger\dagger} = \normftwo{\cdot}$, the goal is to reach $\normftwo{\mM^k} = \alpha \sigma_1 + (1-\alpha) \normf{\mM^k}$.
+
+Let us note that $\Delta = \alpha (u_1 v_1^\top) + (1-\alpha) \frac{\mM^k}{\normf{\mM^k}}$ delivers this value.
+Indeed, by the trace property and singular vectors orthogonality, $\<\mM^k, \Delta> =\<\mU \mSigma \mV^\top, \alpha u_1 v_1^\top + (1-\alpha) \frac{\mU \mSigma \mV^\top}{\normf{\mM^k}}> = \alpha \tr \diag({\sigma_1, 0, \dots, 0}), + (1-\alpha) \normf{\mM^k} = \normftwo{\mM^k}$, which completes the proof.
+
 \subsection{}

### 2025-08-20 conf_article/neon.tex
Added L-smooth part

@@
 %------------------------------------------------------------------------------------
+% \epsilon is predefined; override to use \varepsilon
+\renewcommand{\epsilon}{\varepsilon}
 \newcommand{\Rmn}{\R^{m\times n}}
+\newcommand{\cB}{\mathcal{B}}
 \newcommand{\cD}{\mathcal{D}}
+\newcommand{\cO}{\mathcal{O}}
 \newcommand{\cS}{\mathcal{S}}
-\newcommand{\cB}{\mathcal{B}}
 \newcommand{\Ed}[2]{\mathbb{E}_{#1}\left[#2\right]}
@@
 Hence, new scalable, time- and memory-efficient algorithms have been demanded.
-Besides well-known Adam and AdamW (citations), recently proposed Muon has shown promising results on different benchmarks (CIFAR, NanoGPT, citations).
+Besides well-known Adam and AdamW \cite{kingma2014adam}, \cite{loshchilov2017decoupled}, recently proposed Muon has shown promising results on different benchmarks (CIFAR, NanoGPT, citations).
 Its key difference from earlier algorthims is that it has been constructed specifically for optimizing functions of weight matrices.
@@
 the spectral norm, to derive the algorithm's update (cite Deriving Muon).
-Several other attempts have been since made to construct new algorithms, mainly generalizations of Muon's paradigm (cite Scion and Gluon).
+Several other attempts have been since made to construct new algorithms, mainly generalizations of Muon's paradigm (\cite{pethick2025training} and \cite{riabinin2025gluon}).
 
- Based upon recent theoretical advances that explain some theory behind Muon, Scion and Gluon (cite), we explore application of other matrix norms to optimization of functions of matrices.
+ Based upon recent theoretical advances that explain some theory behind Muon, Scion and Gluon (\cite{bernstein2025deriving,kovalev2025understanding,pethick2025training,riabinin2025gluon}), we explore application of other matrix norms to optimization of functions of matrices.
 As it has been done with Muon, we stipulate that our algorithms' updates be fast to compute.
@@
     \begin{aequation}\label{eq:L}
-    \norm{\nabla f(\mX) - \nabla f(\mX')}_* \leq L \norm{\mX-\mX'}
+    \norm{\nabla f(\mX) - \nabla f(\mX')}^\dagger \leq L \norm{\mX-\mX'}
     \quad\text{for all}\;
@@
 
-        Lemma. When $\norm{\cdot} = \normf{\cdot}$, \cref{eq:our_update} turns into:
+        \begin{lemma}\label{lemma:nsgd_update}
+            When $\norm{\cdot} = \normf{\cdot}$, \cref{eq:our_update} turns into:
 
@@
 
+        \end{lemma}
         It is an interesting observation, because in other works (\cite{pethick2025training}), $\normf{\cdot}$ was used to recover SGD. The difference is in how one states the problem.
@@
     {\bf $\norms{\mM^{k}}$ and Muon}
-
-        Lemma. When $\norm{\cdot} = \norms{\cdot}$, \cref{eq:our_update} turns into:
+    \begin{lemma}\label{lemma:muon_update}
+        When $\norm{\cdot} = \norms{\cdot}$, \cref{eq:our_update} turns into:
 
@@
         \end{equation}
+    \end{lemma}
 
     {\bf $\normn{\mM^{k}}$ and Neon}
-
-        Lemma. When $\norm{\cdot} = \normn{\cdot}$, \cref{eq:our_update} turns into:
+    \begin{lemma}\label{lemma:neon_update}
+        When $\norm{\cdot} = \normn{\cdot}$, \cref{eq:our_update} turns into:
 
@@
         \end{equation}
-
+    \end{lemma}
         We name the derived algorithm \emph{Neon}. In the section Matrix side of updates, we will discuss how to compute an update efficiently.
@@
 
-        Lemma. When $\norm{\cdot} = \normfstar{\cdot}^\dagger$, \cref{eq:our_update} turns into:
+    \begin{lemma}\label{lemma:nsgd_muon_update}
+        When $\norm{\cdot} = \normfstar{\cdot}^\dagger$, \cref{eq:our_update} turns into:
 
@@
         \end{equation}
+    \end{lemma}
 
@@
 
-        Lemma. When $\norm{\cdot} = \normftwo{\cdot}^\dagger$, \cref{eq:our_update} turns into:
+    \begin{lemma}\label{lemma:nsgd_neon_update}    
+        When $\norm{\cdot} = \normftwo{\cdot}^\dagger$, \cref{eq:our_update} turns into:
 
@@
 
-\section{"Smooth" theory via Trust Region bounds}
-\begin{itemize}
-    \item Probably statement of the trust region problem
-    \item Assumption on unbiasedness and limited variance of stoch. grad.
-    \item Assumption on L-smoothness as in the Kovalev's article.
-    \item Assumption on rho. To be more precise, the statement of lemmas on rho for all 4 norms.
-    \item We utilize the Kovalev's corollary to derive convergence
-    \item Bonus: convert all the norms into a common sense (L2?) norm to get the true dependency on dimensionality
-\end{itemize}
+\section{Trust Region Bounds for L-smooth functions}
+    First, we analyze the problem in the unstochastic case.
+From Corollary 1 of \cite{kovalev2025understanding}, we directly get the following result that matches lower bounds, as was noted in \cite{kovalev2025understanding}.
+
+    \begin{lemma}\label{lemma:no_noise_tr}
+    To reach the precision $\min_{k=1\ldots K} \norm{\nabla f(\mX_k)}^\dagger \leq \epsilon$ by the iterations~\eqref{eq:our_update} under the conditions of \Cref{eq:L}, it is sufficient to choose the stepsize $\eta$ and the number of iterations $K$ as follows:
+    \begin{equation}
+      \eta = \cO\left(\frac{\epsilon}{L}\right),\qquad K = \cO\left(\frac{L\Delta_0}{\epsilon^2}\right).
+      % \eta = \frac{\epsilon}{3L},\qquad K = \ceil*{\frac{6L\Delta_0}{\epsilon^2}}.
+    \end{equation}
+    \end{lemma}
+
+    In the stochastic case, from Corollary of 2 of \cite{kovalev2025understanding}, we directly get the following result that matches lower bounds, as was noted in \cite{kovalev2025understanding}:
+
+    \begin{lemma}\label{lemma:stoch_tr}
+    To reach the precision $\E{\min_{k=1\ldots K} \norm{\nabla f(\mX_k)}^\dagger} \leq \epsilon$ by ~\eqref{eq:our_update} under the assumptions \Cref{eq:variance}, \Cref{eq:L}, \Cref{eq:norm}, it is sufficient to choose the parameters as follows:
+    \begin{align}
+        \eta &= \cO\left(\min\left\{\frac{\epsilon}{L}, \frac{\epsilon^3}{\rho^2\sigma^2L}\right\}\right),
+        \qquad
+        \alpha = \cO\left(\min\left\{1, \frac{\epsilon^2}{\rho^2\sigma^2}\right\}\right),
+        \\
+        \label{eq:str_K_nonconvex}
+        K &= \cO\left(\max\left\{
+            \frac{\rho\sigma}{\epsilon},
+            \frac{\rho^3\sigma^3}{\epsilon^3},
+            \frac{L\Delta_0}{\epsilon^2},
+            \frac{L\Delta_0\rho^2\sigma^2}{\epsilon^4}
+        \right\}\right).
+    \end{align}
+    \end{lemma}
+
+    As the norms $\normf{\cdot}$, $\normn{\cdot}$, $\normf{\cdot}$ are almost proportional to each other when $m, n \rightarrow \infty$ (with high probability for random matrices), the expected convergence guarantees in terms of $\normf{\cdot}$ are the same (it can be easily shown by noting that $\norm{\mX}\sim\alpha\normf{\mX}$, $\norm{\nabla f(\mX)}^\dagger \sim \frac{1}{\alpha} \normf{\nabla f(\mX)}$, and expressing $L$-constant via $L_F$-constant for the Frobenius norm).
 
@@
 Technical details on experiments
+\iffalse
 
@@
 
-\bibliography{icomp2024_conference}
-\bibliographystyle{icomp2024_conference}
-
 \appendix
@@
 
+\fi
+\bibliography{icomp2024_conference}
+\bibliographystyle{icomp2024_conference}
+
 \end{document}

### 2025-08-21 conf_article/neon.tex
LLS description half done

@@
 \newcommand{\cD}{\mathcal{D}}
+\newcommand{\cN}{\mathcal{N}}
 \newcommand{\cO}{\mathcal{O}}
@@
 
-        TO-DO: F*-neon? (it would be hard to do it)
-
-\subsection{F*-Neon}
-Motivation: we need updates of higher rank. Hence, F* (why? regularization?)
-%\subsection{F-Muon}
-%Motivation: if the spectral norm (Muon) is dual to the nuclear norm (Neon), then what is dual to F*? F-Muon! Derivation.
+\subsection{Algorithms for Matrices $\leftrightarrow$ Algorithms for Vectors}
+Table that compares Muons and Neons to vector algorithms (NSGD).
 
@@
     To reach the precision $\min_{k=1\ldots K} \norm{\nabla f(\mX_k)}^\dagger \leq \epsilon$ by the iterations~\eqref{eq:our_update} under the conditions of \Cref{eq:L}, it is sufficient to choose the stepsize $\eta$ and the number of iterations $K$ as follows:
-    \begin{equation}
+    \begin{equation}\label{eq:unstoch_tr}
       \eta = \cO\left(\frac{\epsilon}{L}\right),\qquad K = \cO\left(\frac{L\Delta_0}{\epsilon^2}\right).
@@
 
+    From the theory of random martices and the Marchenko-Pastur law, we get that random $\mM \in \Rmn: \mM \sim \cN(0, \sigma^2 \Rmn)$ has the following asymptotics of its norms:
+
+    Nuclear: $\sigma n \sqrt{m}$
+
+    Frobenius: $\sigma \sqrt{m n}$
+
+    Spectral: $\sigma(\sqrt{m} + \sqrt{n})$
+
+    This means that for square random matrices $n \times n$  the following asymptotics take place: $\normf{\cdot} \sim \frac{\sqrt{n}}{2}\norms{\cdot}$ and $\normn{\cdot} \sim \frac{n}{2}\norms{\cdot}$.
+
 \section{Experiments}
-    \subsection{Synthetic quadratics}
-    Description of how to construct L-smooth functions in a given norm, and how to estimate variance of the gradient. Then check of the theory.
-    It must be the only experiment where we might beat vanilla Muon.
-    \subsection{MLP}
-    Deep learning?
+    \subsection{Randomized Linear Least Squared}
+    Since the provided by other authors \cite{kovalev2025understanding,riabinin2025gluon} theoretical guarantees are almost norm-independent, we have to test them in practice.
+
+    To test the bounds from \cite{kovalev2025understanding} in practice, we construct the following L-smooth problem:
+
+    \begin{equation}\label{eq:lls}
+        F(\mX) = \frac{1}{2} \<(\mX-\mS), \mM (\mX - \mS) \mN>
+    \end{equation}
+
+    where $\mX \in \Rmn$, $m=10$, $n=10$, $S \in \Rmn$, $\mM \in \mathbb{S}^m_{+}$ and $\mN \in \mathbb{S}^n_{+}$ are positive-semidefinite matrices.
+Spectra of $\mM$ and $\mN$ are uniformly distributed in (0, 1) interval.
+
+    It it easy to derive the gradient
+    \begin{equation}\label{eq:grad_lls}
+        \nabla F(\mX) = \mM (\mX - \mS) \mN,
+    \end{equation}
+
+    Let us define $\gamma$ as $\norm{\cdot} \sim \gamma \normf{\cdot}$, which is asymptotics from the previous section.
+Then $\norm{\cdot}^\dagger \sim \frac{1}{\gamma}\normf{\cdot}$, as $\normf{\cdot}^\dagger = \normf{\cdot}$.
+Hence, $\norm{\cdot} \sim \gamma^2 \norm{\cdot}^\dagger$
+
+    Then $\norm{\nabla f(\mX) - \nabla f(\mY)}^\dagger = \norm{\mM (\mX - \mY) \mN}^\dagger \leq \norm{\mM}^\dagger \norm{\mN}^\dagger \norm{\mX - \mY}^\dagger \sim \norm{\mM}^\dagger \norm{\mN}^\dagger \gamma^2 \norm{\mX - \mY}$, and $L \sim \gamma^2 \norm{\mM}^\dagger \norm{\mN}^\dagger$.
+
+    Now for the known norms:
+
+    Frobenius (NSGD): $L = \normf{M} \normf{N}$
+
+    Spectral (Muon): $\gamma \sim \frac{2}{\sqrt{n}} \implies L \sim \frac{2}{n}\normn{M}\normn{N}$
+
+    Nuclear (Neon): $\gamma \sim \sqrt{n} \implies L \sim n\norms{M}\norms{N}$
+
+    We take learning rate $\eta$ and iteration number $K$ from \cref{eq:unstoch_tr}.
+
+    We are interested to get $\epsilon_2=$ 1e-1 precision in $\norms{\cdot}$.
+Hence, we target $\epsilon = \epsilon_2$ for the nuclear norm (Neon), $\epsilon = \frac{n}{2}$ for the spectral norm (Muon), and $\epsilon = \frac{\sqrt{n}}{2}$ for the Frobenius norm (NSGD).
+
+    {\bf An independent experiment}
+    To make the problem more real-world, we do not use known information about smoothness.
+We run NormalizedSGD, Muon, F-Muon with $\alpha=1/2$, Neon, and F-Neon with $\alpha=1/2$ for 100 000 iterations with learning rate = 1e-3.
+
+    \subsection{Logistic Regression}
     \subsection{Benchmarks}
@@
 Let us note that $\Delta = \alpha (u_1 v_1^\top) + (1-\alpha) \frac{\mM^k}{\normf{\mM^k}}$ delivers this value.
-Indeed, by the trace property and singular vectors orthogonality, $\<\mM^k, \Delta> =\<\mU \mSigma \mV^\top, \alpha u_1 v_1^\top + (1-\alpha) \frac{\mU \mSigma \mV^\top}{\normf{\mM^k}}> = \alpha \tr \diag({\sigma_1, 0, \dots, 0}), + (1-\alpha) \normf{\mM^k} = \normftwo{\mM^k}$, which completes the proof.
+Indeed, by the trace property and singular vectors orthogonality, $\<\mM^k, \Delta> =\<\mU \mSigma \mV^\top, \alpha u_1 v_1^\top + (1-\alpha) \frac{\mU \mSigma \mV^\top}{\normf{\mM^k}}> = \alpha \tr \diag({\sigma_1, 0, \dots, 0}) + (1-\alpha) \normf{\mM^k} = \normftwo{\mM^k}$, which completes the proof.
 

### 2025-08-21 code/benchmark_plotter.py
Working on LLS and Lipschitz

@@ -109,16 +109,16 @@ def build_default_panels(experiments: Dict[str, Dict[str, Any]]) -> List[Dict[st
             "yscale": "log",
         },
         {
-            "title": "Gradient Frobenius Norm vs Iteration",
+            "title": "Gradient Spectral Norm vs Iteration",
             "x_label": "Iteration",
-            "y_label": "||∇f(X)||_F",
+            "y_label": "||∇f(X)||_2",
             "series": collect_series("iterations", "grad_frobenius_norms"),
             "yscale": "log",
         },
         {
-            "title": "Gradient Frobenius Norm vs Time",
+            "title": "Gradient Spectral Norm vs Time",
             "x_label": "Time (s)",
-            "y_label": "||∇f(X)||_F",
+            "y_label": "||∇f(X)||_2",
             "series": collect_series("cumulative_time", "grad_frobenius_norms"),
             "yscale": "log",
         },

### 2025-08-21 code/benchmark_runner.py

@@ -21,13 +21,14 @@ class MatrixProblem:
 
 
 def run_optimizer_on_problem(
+    name: str,
     optimizer_class: Callable[..., torch.optim.Optimizer],
     optimizer_kwargs: Dict[str, Any],
     problem: MatrixProblem,
     X_init: torch.Tensor,
     num_iterations: int = 500,
     record_interval: int = 1,
-    verbose: bool = False,
+    verbose: bool = False
 ) -> Dict[str, Any]:
     """
     Run an optimizer on a matrix problem and record metrics.
@@ -38,7 +39,7 @@ def run_optimizer_on_problem(
       - losses: list[float]
       - grad_frobenius_norms: list[float]  (||∇f(X)||_F)
     """
-
+    print(f"Testing {name}")
     X = X_init.clone()
 
     optimizer = optimizer_class([X], **optimizer_kwargs)
@@ -46,7 +47,7 @@ def run_optimizer_on_problem(
     iterations: list[int] = []
     cumulative_time: list[float] = []
     losses: list[float] = []
-    grad_fro_norms: list[float] = []
+    grad_fro_norms: list[float] = [] # now spectral norms
 
     start_time = time.time()
     
@@ -62,14 +63,14 @@ def run_optimizer_on_problem(
                 cumulative_time.append(cur_time)
                 losses.append(problem.objective(X).item())
                 grad_now = problem.gradient(X)
-                grad_fro_norms.append(torch.linalg.norm(grad_now, ord="fro").item())
+                grad_fro_norms.append(torch.linalg.norm(grad_now, ord=2).item())
 
         # Optional progress
         if verbose and (t % max(1, (num_iterations // 10)) == 0):
             with torch.no_grad():
                 loss_val = problem.objective(X).item()
-                grad_norm_val = torch.linalg.norm(grad, ord="fro").item()
-                print(f"Iter {t}: loss={loss_val:.6f}, ||∇f(X)||_F={grad_norm_val:.6f}")
+                grad_norm_val = torch.linalg.norm(grad, ord=2).item()
+                print(f"Iter {t}: loss={loss_val:.6f}, ||∇f(X)||_2={grad_norm_val:.6f}")
 
         # Optimizer step
         optimizer.step()

### 2025-08-21 conf_article/neon.tex
Expanded introduction

@@
 Hence, new scalable, time- and memory-efficient algorithms have been demanded.
-Besides well-known Adam and AdamW \cite{kingma2014adam}, \cite{loshchilov2017decoupled}, recently proposed Muon has shown promising results on different benchmarks (CIFAR, NanoGPT, citations).
-Its key difference from earlier algorthims is that it has been constructed specifically for optimizing functions of weight matrices.
-Such functions are common in modern machine learning (citations), so it does not significantly restrict Muon's applicability to modern problems.
+Besides well-known Adam and AdamW \cite{kingma2014adam}, \cite{loshchilov2017decoupled}, recently proposed Muon has shown promising results on training very large models \cite{jordan2024muon,ahn2025dioncommunicationefficientoptimizerlarge}.
+Its key difference from earlier algorthims is that it has been constructed specifically for optimizing functions of weight matrices, which are common in deep learning.
 
@@
 From the perspective of theory, Muon's main innovation was an intentional usage of matrix norms, i.e.
-the spectral norm, to derive the algorithm's update (cite Deriving Muon).
-Several other attempts have been since made to construct new algorithms, mainly generalizations of Muon's paradigm (\cite{pethick2025training} and \cite{riabinin2025gluon}).
+the spectral norm, to derive the algorithm's update \cite{bernstein2025deriving}.
+Several other attempts have been since made to construct new algorithms, mainly generalizations of Muon's paradigm, like Scion (\cite{pethick2025training} and Gluon \cite{riabinin2025gluon}).
 
@@
 As it has been done with Muon, we stipulate that our algorithms' updates be fast to compute.
+
+ In this article, we focus on the two most common matrix norms akin to the spectral, namely, the Nuclear Norm and the Frobenius norm.
+Working in the linear minimization oracle (lmo) approach, which is equivalent to a factor to the trust region and the steepest descent under norm constraint approaches, we derive Neon, our algorithm based on the nuclear norm.
+In the section {\it Matrix side of the updates}, we explain how Neon updates can be computed asymptotically faster than Muon updates by the Newton-Schulz iterations.
+
+ Noticing that Neon and Muon are diametrical in terms of the rank of the update matrix, we bridge the space by "regularizing" them by NormalizedSGD, which is derived in lmo with the Frobenius norm.
+We do this in the same lmo approach by considering a norm that is dual to the convex combination of the Frobenius norm and the spectral or the nuclear norms respectively.
+So we derive the algorithms we name F-Neon and F-Muon respectively.
+
+ Table that compares methods, updates, and sends to the proofs?
+
+ Having faced the array of different Muon-like optimizers, according to the upper bounds from \cite{kovalev2025understanding,riabinin2025gluon}, with similar convergence behavior, we painstakingly compare them on a synthetic linear least squares problem with known Lipschitz constant.
+The efforts results in comparison of the algorithms by their convergence not in terms of the dual norms of their gradient, but in terms of the common spectral norm, which may strongly differ, especially in large matrices, from the initial dual norms.
+Thus, we compare the algorithms in a unified fashion.
+
+ Finally, we test Muon, Neon, NSGD, F-Muon and F-Neon on deep-learning tasks: training convolutional network on CIFAR-10 and fine-tuning NanoGPT.
+The results support the supremacy of Muon, but the most striking result of the tests is that F-Muon, only half of which is Muon, surpasses Muon's accuracy on the CIFAR tasks by a margin.
+The case of F-Muon answers in the affirmative to the question of feasibility of constructing a mixture of optimization algorithms to increase robustness of the composite algorithm.
+
 \iffalse
@@
 \section{Experiments}
-    \subsection{Randomized Linear Least Squared}
+    \subsection{Randomized Linear Least Squares}
     Since the provided by other authors \cite{kovalev2025understanding,riabinin2025gluon} theoretical guarantees are almost norm-independent, we have to test them in practice.
@@
 
-%related work
+\section{Related Work}
+What to include: Muon and its upgrades (say that the same can be done with Neon), alternatives to Newton-Schulz, nuSAM and so on.
 
@@
 \end{itemize}
+
+\bibliography{icomp2024_conference}
+\bibliographystyle{icomp2024_conference}
+
 \section{Appendix}
-\subsection{Norm derivations}
-% \begin{itemize}
-%    \item Dual to F* norm
-%    \item Bounds on rho for all 4 norms (with their tighness if possible)
-% \end{itemize}
+\subsection{Norms $\normfstar{\cdot}^\dagger$ and $\normftwo{\cdot}^\dagger$}
+Here we provide the derivation of these norms and plot them in the case of $2\times2$ matrices.
 \subsection{Updates derivations}
@@
 
-\subsection{}
+\subsection{Technical details of the experiments}
 ~ Note: nuSAM and their update. They had no Lanczos
-Technical details on experiments
+
 \iffalse
@@
 \fi
-\bibliography{icomp2024_conference}
-\bibliographystyle{icomp2024_conference}
 

### 2025-08-21 conf_article/neon.tex
Moved stochastic into the appendix

@@
 
-{\bf Stochastic gradient estimator.}
-    We assume access to a stochastic estimator $g(\cdot\,;\xi)\colon \Rmn \to \Rmn$ of the gradient $\nabla f(\cdot)$, where $\xi \sim \cD$ is a random variable sampled from a probability distribution $\cD$.
-We assume that the stochastic gradient estimator $g(\cdot;\xi)$ is unbiased and has bounded variance, that is, the following relations hold:
-    \begin{aequation}\label{eq:variance}
-    \Ed{\xi \sim \cD}{g(\mX;\xi)} = \nabla f(\mX)
-    \quad\text{and}\quad
-    \Ed{\xi \sim \cD}{\sqnf{g(\mX;\xi) - \nabla f(\mX)}} \leq \sigma^2
-    \quad\text{for all}\;
-    \mX \in \Rmn,
-    \end{aequation}
-    where $\sigma > 0$ is a positive variance parameter, and $\normf{\cdot}$ is the standard Euclidean, i.e.
-Frobenius, norm induced by the inner product $\<\cdot,\cdot>$, i.e., $\normf{\mX} = \sqrt{\<\mX,\mX>}=\sqrt{\tr(\mX^\top \mX)}$.
-These assumptions have been widely adopted for the analysis of many stochastic gradient optimization algorithms \citep{ghadimi2013stochastic,ghadimi2016accelerated,cutkosky2020momentum,sun2023momentum,horvath2023stochastic,gorbunov2020linearly}.
-
 {\bf Non-Euclidean norm setting and Lipschitz continuous gradient.}
@@
     where $L > 0$ is the gradient Lipschitz constant, and $\norm{\cdot}^\dagger\colon \Rmn \to \R_+$ is the dual norm associated with $\norm{\cdot}$, i.e., $\norm{\mX}^\dagger = \sup_{\norm{\mX'}\leq 1} \<\mX,\mX'>$ for all $\mX \in \Rmn$.
-The assumption of gradient Lipschitz continuity is also widespread in the analysis of first-order optimization methods \citep{ghadimi2013stochastic,gower2019sgd,cutkosky2020momentum,horvath2023stochastic,gorbunov2020linearly}.
-It is important to highlight that while \Cref{eq:L} uses the dual norm $\norm{\cdot}^\dagger$ to measure the difference between the gradients, the variance in \Cref{eq:variance} is measured with respect to the Frobenius norm $\sqnf{\cdot}$, which is necessary to properly utilize the unbiasedness property of the stochastic gradient estimator $g(\cdot;\xi)$.
-Therefore, we need to provide a connection between these norms using the following inequality:
-    \begin{aequation}\label{eq:norm}
-    \norm{\mX}^\dagger \leq \rho\cdot\normf{\mX}
-    \quad\text{for all}\;
-    \mX \in \Rmn,
-    \end{aequation}
-    where $\rho > 0$ is a positive constant.
-Note that such a constant always exists due to the norm equivalence theorem, which always holds in the finite-dimensional space $\Rmn$.
-We recount $\rho$ for different norms $\norm{\cdot}$ in the appendix.
 
@@
 \section{Appendix}
+\subsection{Formal assumptions}
+{\bf Stochastic gradient estimator.}
+    We assume access to a stochastic estimator $g(\cdot\,;\xi)\colon \Rmn \to \Rmn$ of the gradient $\nabla f(\cdot)$, where $\xi \sim \cD$ is a random variable sampled from a probability distribution $\cD$.
+We assume that the stochastic gradient estimator $g(\cdot;\xi)$ is unbiased and has bounded variance, that is, the following relations hold:
+    \begin{aequation}\label{eq:variance}
+    \Ed{\xi \sim \cD}{g(\mX;\xi)} = \nabla f(\mX)
+    \quad\text{and}\quad
+    \Ed{\xi \sim \cD}{\sqnf{g(\mX;\xi) - \nabla f(\mX)}} \leq \sigma^2
+    \quad\text{for all}\;
+    \mX \in \Rmn,
+    \end{aequation}
+    where $\sigma > 0$ is a positive variance parameter, and $\normf{\cdot}$ is the standard Euclidean, i.e.
+Frobenius, norm induced by the inner product $\<\cdot,\cdot>$, i.e., $\normf{\mX} = \sqrt{\<\mX,\mX>}=\sqrt{\tr(\mX^\top \mX)}$.
+These assumptions have been widely adopted for the analysis of many stochastic gradient optimization algorithms \citep{ghadimi2013stochastic,ghadimi2016accelerated,cutkosky2020momentum,sun2023momentum,horvath2023stochastic,gorbunov2020linearly}.
+
+    It is important to highlight that while \Cref{eq:L} uses the dual norm $\norm{\cdot}^\dagger$ to measure the difference between the gradients, the variance in \Cref{eq:variance} is measured with respect to the Frobenius norm $\sqnf{\cdot}$, which is necessary to properly utilize the unbiasedness property of the stochastic gradient estimator $g(\cdot;\xi)$.
+Therefore, we need to provide a connection between these norms using the following inequality:
+    \begin{aequation}\label{eq:norm}
+    \norm{\mX}^\dagger \leq \rho\cdot\normf{\mX}
+    \quad\text{for all}\;
+    \mX \in \Rmn,
+    \end{aequation}
+    where $\rho > 0$ is a positive constant.
+Note that such a constant always exists due to the norm equivalence theorem, which always holds in the finite-dimensional space $\Rmn$.
+We recount $\rho$ for different norms $\norm{\cdot}$ in the appendix.
+
 \subsection{Norms $\normfstar{\cdot}^\dagger$ and $\normftwo{\cdot}^\dagger$}

### 2025-08-22 conf_article/neon.tex
Ky Fan k-norm introduced

@@
 \DeclarePairedDelimiter{\normf}{\|}{\|_\mathrm{F}}
+\DeclarePairedDelimiter{\normkfk}{\|}{\|_\mathrm{KF-k}}
 \DeclarePairedDelimiter{\normfstar}{\|}{\|_\mathrm{F*}}
@@
 Hence, new scalable, time- and memory-efficient algorithms have been demanded.
-Besides well-known Adam and AdamW \cite{kingma2014adam}, \cite{loshchilov2017decoupled}, recently proposed Muon has shown promising results on training very large models \cite{jordan2024muon,ahn2025dioncommunicationefficientoptimizerlarge}.
+Besides well-known Adam and AdamW \cite{kingma2014adam}, \cite{loshchilov2017decoupled}, recently proposed Muon has shown promising results on training very large models \cite{jordan2024muon,liu2025muon}.
 Its key difference from earlier algorthims is that it has been constructed specifically for optimizing functions of weight matrices, which are common in deep learning.
@@
 
+    {\bf $\normkfk{\mM^k}^\dagger$ and Muon, Neon, and Dion without error feedback}
+        We remind the reader that the Ky Fan k-norm \cite{fan1951maximum}, which we denote as $\normkfk{\cdot}$, is the sum of the k largest singular values of the matrix.
+It can be proven that $\normkfk{\cdot}^\dagger = \max\{\frac{1}{k} \normn{\cdot}, \norms{\cdot}\}$ (see Appendix).
+Special cases of Ky Fan k-norm are the Ky Fan 1-norm, which is the sprectral norm, and the Ky Fan $\min\{m, n\}$-norm, which is the nuclear norm.
+
+        \begin{lemma}\label{lemma:ky_fan_update}    
+            When $\norm{\cdot} = \normkfk{\mM^k}^\dagger$, \cref{eq:our_update} turns into:
+
+            \begin{equation}\label{eq:ky_fan_update}
+                \mX^{k+1} = \mX^{k} - \eta \sum_{i=1}^{k}u_i v_i^\top
+            \end{equation}
+        \end{lemma}
+
+        This lemma recovers the updates of Muon, when $k=\min\{m, n\}$, of Neon, when $k=1$, and of Dion \cite{ahn2025dioncommunicationefficientoptimizerlarge} without the error feedback.
+
+        Moreover, one can consider F-KF-k-norm = $\alpha \normkfk{\cdot} + (1-\alpha)\normf{\cdot}$, the dual to which will produce algorithms like F-Dion without the error feedback.
+
 \subsection{Algorithms for Matrices $\leftrightarrow$ Algorithms for Vectors}
@@
 \section{Related Work}
-What to include: Muon and its upgrades (say that the same can be done with Neon), alternatives to Newton-Schulz, nuSAM and so on.
+As Muon \cite{jordan2024muon} is a very successful and popular optimizer for functions of weight matrices, a lot of reseach has been put into explaining its success and further improving its performance.
+
+Regarding the second point, in less than a year, a large number of applications and improvements of Muon has been proposed.
+\cite{liu2025muon} adapted the algorthm for training language models larger than NanoGPT.
+\cite{shah2025practical} organized efficient hyperparameter transfer by combining Muon with maximal update parametrization.
+\cite{chen2025cosmoshybridadaptive} applied computationally intensive updates of SOAP optimizer to a low-dimensional "leading eigensubspace" while using memory-efficient methods like Muon for the remaining parameters.
+\cite{amsel2025polar} proposed a more efficient alternative to Newton-Schulz operations.
+\cite{si2025adamuon} introduced AdaMuon which combines element-wise adaptivity with orthogonal updates.
+
+Regarding the first point, simplistic derivation of \cite{bernstein2025deriving} which is based on \cite{bernstein2024oldoptimizernewnorm} excluded, there has been a gap in theory behind Muon, which, as it seems to us, is not even now completely closed.
+\cite{kovalev2025understanding} has provided convergence guarantees of Muon in various settings, from which, however, Muon's supremacy cannot be recovered.
+Indeed, although the obtained bounds depend on the norm choice, the asymptotics of the convergence remain the same as for NSGD and other optimizers, $K = \cO(\epsilon^{-4})$ in a L-smooth stochastic case.
+
+Similar drawback has a recent article \cite{riabinin2025gluon}, where L-smoothness assumption is replaced with a more practical $(L_0, L_1)$-smoothness.
+The authors derived from their theorems optimal stepsizes for Muon and Scion that match fine-tuned stepsizes reported by \cite{pethick2025training}.
+But still, they did not explain why, for instance, NSGD is inferior to Muon in training large-language models.
 
@@
 
-\subsection{Norms $\normfstar{\cdot}^\dagger$ and $\normftwo{\cdot}^\dagger$}
+\subsection{Norms $\normfstar{\cdot}^\dagger$, $\normftwo{\cdot}^\dagger$, and $\normkfk{\cdot}^\dagger$}
 Here we provide the derivation of these norms and plot them in the case of $2\times2$ matrices.
+
 \subsection{Updates derivations}
 
-Derivation of \cref{eq:muon_update} follows from \cref{eq:nsgd_muon_update} with $\alpha=1$. Indeed, $\norms{\cdot}^\dagger=1\cdot\normn{\cdot} + 0 \cdot \normf{\cdot} $.
+Proof of \cref{lemma:muon_update} follows from \cref{eq:nsgd_muon_update} with $\alpha=1$. Indeed, $\norms{\cdot}^\dagger=1\cdot\normn{\cdot} + 0 \cdot \normf{\cdot} $.
 
-Derivation of \cref{eq:nsgd_muon_update}:
+Proof of \cref{lemma:nsgd_muon_update}:
 Since $\norm{\cdot}^\dagger = \normfstar{\cdot}^{\dagger\dagger} = \normfstar{\cdot}$, the goal is to reach $\normfstar{\mM^k} = \alpha \tr{\mSigma} + (1-\alpha) \normf{\mM^k}$.
 
-Let us note that $\Delta = \alpha (\mU \mV^\top) + (1-\alpha) \frac{\mM^k}{\normf{\mM^k}}$ delivers this value.
+Let us note that $\Delta = \alpha \mU \mV^\top + (1-\alpha) \frac{\mM^k}{\normf{\mM^k}}$ delivers this value.
 Indeed, by the trace property, $\<\mM^k, \Delta> =\<\mU \mSigma \mV^\top, \alpha \mU \mV^\top + (1-\alpha) \frac{\mU \mSigma \mV^\top}{\normf{\mM^k}}> = \alpha \tr \mSigma + (1-\alpha) \normf{\mM^k} = \normfstar{\mM^k}$, which completes the proof.
 
-Derivation of \cref{eq:neon_update} follows from \cref{eq:nsgd_neon_update} with $\alpha=1$. Indeed, $\normn{\cdot}^\dagger=1\cdot\norms{\cdot} + 0 \cdot \normf{\cdot} $.
+Proof of \cref{lemma:neon_update} follows from \cref{eq:nsgd_neon_update} with $\alpha=1$. Indeed, $\normn{\cdot}^\dagger=1\cdot\norms{\cdot} + 0 \cdot \normf{\cdot} $.
 
-Derivation of \cref{eq:nsgd_neon_update}:
+Proof of \cref{lemma:nsgd_neon_update}:
 Since $\norm{\cdot}^\dagger = \normftwo{\cdot}^{\dagger\dagger} = \normftwo{\cdot}$, the goal is to reach $\normftwo{\mM^k} = \alpha \sigma_1 + (1-\alpha) \normf{\mM^k}$.
@@
 Indeed, by the trace property and singular vectors orthogonality, $\<\mM^k, \Delta> =\<\mU \mSigma \mV^\top, \alpha u_1 v_1^\top + (1-\alpha) \frac{\mU \mSigma \mV^\top}{\normf{\mM^k}}> = \alpha \tr \diag({\sigma_1, 0, \dots, 0}) + (1-\alpha) \normf{\mM^k} = \normftwo{\mM^k}$, which completes the proof.
+
+Proof of \cref{lemma:ky_fan_update}:
+Since $\norm{\cdot}^\dagger = \normkfk{\cdot}^{\dagger\dagger} = \normkfk{\cdot}$, the goal to reach $\normkfk{\mM^k}$.
+
+Let us note that $\Delta = \sum_{i=1}^{k} u_i v_i^\top$ delivers the value.
+Indeed, $\<\mM^k, \Delta> =\<\mU \mSigma \mV^\top, \sum_{i=1}^{k} u_i v_i^\top> = \sum_{i,j=1}^{r, k}\<u_i \sigma_i v_i^\top, u_j v_j^\top> = \sum_{i=1}^{k}\sigma_i = \normkfk{\mM^k}$, which completes the proof.
 

### 2025-08-22 code/base_airbench.py
Dion implementation

@@ -22,6 +22,7 @@ import torch.nn.functional as F
 import torchvision
 import torchvision.transforms as T
 from matrix_functions import one_sv_svds_approximation
+from optimizers import Dion
 
 torch.backends.cudnn.benchmark = True
 
@@ -479,7 +480,8 @@ def main(run, model):
                      dict(params=[model.head.weight], lr=head_lr, weight_decay=wd/head_lr)]
     optimizer1 = torch.optim.SGD(param_configs, momentum=0.85, nesterov=True, fused=True)
     # optimizer2 = NormalizedMuon(filter_params, lr=0.24, momentum=0.5, nesterov=True, sgd_coeff=-0.15) # important
-    optimizer2 = SignSGDMuon(filter_params, lr=0.45, momentum=0.7, nesterov=True, sgd_coeff=0.5)
+    optimizer2 = Dion(filter_params, lr=0.24, momentum=0.6, rank=5, momentum_decay=0.6)
+    # optimizer2 = SignSGDMuon(filter_params, lr=0.45, momentum=0.7, nesterov=True, sgd_coeff=0.5)
     # optimizer2 = Muon(filter_params, lr=0.24, momentum=0.6, nesterov=True)
     optimizers = [optimizer1, optimizer2]
     for opt in optimizers:

### 2025-08-22 code/matrix_functions.py

@@ -69,6 +69,8 @@ def k_sv_svds_approximation_dlpack(W_torch, k, tau, num_iter=30):
     # print(opt_tau)
     return approx_torch, opt_tau, k
 
+    
+
 
 def svd_full_approximation(A, tau=0):
     k = 0
@@ -251,22 +253,32 @@ def randomized_svd_torch(G, n_components=1, n_oversamples=5, n_iter=2):
 
 def several_sv_svds_approximation(W_torch, k, num_iter=50):
     """SVD approximation using the top k singular values and corresponding vectors."""
+    # Store original device and dtype
+    original_device = W_torch.device
+    original_dtype = W_torch.dtype
+    
     W = cp.from_dlpack(thd.to_dlpack(W_torch)).astype(cp.float32)
     U, S, Vt = cupyx_svds(W, k=min([k, W.shape[0] - 1, W.shape[1] - 1]), maxiter=num_iter, which='LM')
 
-    # approx = U @ Vt #cp.diag(S) - we don't need this!
-    approx_torch_U = thd.from_dlpack(U.toDlpack()) 
-    approx_torch_Vt = thd.from_dlpack(Vt.toDlpack())
-    return approx_torch_U, thd.from_dlpack(S.toDlpack()), approx_torch_Vt
+    # Convert back to torch tensors and ensure they're on the correct device
+    approx_torch_U = thd.from_dlpack(U.toDlpack()).to(device=original_device, dtype=original_dtype)
+    approx_torch_S = thd.from_dlpack(S.toDlpack()).to(device=original_device, dtype=original_dtype)
+    approx_torch_Vt = thd.from_dlpack(Vt.toDlpack()).to(device=original_device, dtype=original_dtype)
+    
+    return approx_torch_U, approx_torch_S, approx_torch_Vt
 
 def one_sv_svds_approximation(W_torch, num_iter=30):
     """SVD approximation using the top k singular values and corresponding vectors."""
     k = 1
+    # Store original device and dtype
+    original_device = W_torch.device
+    original_dtype = W_torch.dtype
+    
     W = cp.from_dlpack(thd.to_dlpack(W_torch)).astype(cp.float32)
     U, S, Vt = cupyx_svds(W, k=min([k, W.shape[0] - 1, W.shape[1] - 1]), maxiter=num_iter, which='LM')
 
     approx = U @ Vt #cp.diag(S) - we don't need this!
-    approx_torch = thd.from_dlpack(approx.toDlpack()) 
[…]

### 2025-08-22 code/optimizers.py

@@ -136,7 +136,6 @@ class Neon(torch.optim.Optimizer):
                 if g is None:
                     continue
                 state = self.state[p]
-
                 if 'momentum_buffer' not in state.keys():
                     state['momentum_buffer'] = torch.zeros_like(g)
                 buf = state['momentum_buffer']
@@ -172,7 +171,11 @@ class Neon(torch.optim.Optimizer):
                         # s[1] = 0
                         update = u @ torch.diag(s) @ vt
                         '''
-                        # update, self.tau, self.k = svd_full_approximation(g_resh, self.tau) -- too slow
+
+                    elif self.type == 'kyfan': # no EF here, so it won't work!
+                        u, s, vt = several_sv_svds_approximation(g_resh, self.k)
+                        update = u @ vt
+                        error = u @ torch.diag(s) @ vt
                 update2 = (1-self.sgd_coeff) * update + self.sgd_coeff * g_resh / (g_resh.norm() + 1e-12)
                 '''
                 b = float(g_resh.norm())
@@ -215,3 +218,122 @@ class NormalizedMuon(torch.optim.Optimizer):
                 else:
                     update = self.sgd_coeff * g_normalized
                 p.data.add_(update, alpha=-lr) # take a step
+
+
+class Dion(torch.optim.Optimizer):
+    def __init__(self, params, lr=1e-3, momentum=0, nesterov=False, rank=1, momentum_decay=0.9):
+        if lr < 0.0:
+            raise ValueError(f"Invalid learning rate: {lr}")
+        if momentum < 0.0:
+            raise ValueError(f"Invalid momentum value: {momentum}")
+        if nesterov and momentum <= 0:
+            raise ValueError("Nesterov momentum requires a momentum")
+        if momentum_decay < 0.0 or momentum_decay > 1.0:
+            raise ValueError(f"Invalid momentum decay value: {momentum_decay}")
+        if rank < 1:
+            raise ValueError(f"Invalid rank value: {rank}")
+        
+        defaults = dict(lr=lr, momentum=momentum, nesterov=nesterov)
+        super().__init__(params, defaults)
+        self.rank = rank
+        self.momentum_decay = momentum_decay
+    '''
+    def power_iter1(self, B, Q):
+        """
+        Single power iteration (from Q)
+        B: gradient matrix (m x n)
+        Q: right factor from previous iteration (n x r)
+        Returns: P (m x r), R (n x r) where P is orthonormal
+        """
[…]

### 2025-08-22 test_error_feedback.py
We don't need EF

@@ -1,0 +1,84 @@
+#!/usr/bin/env python3
+"""
+Test script for Error Feedback Muon optimizers
+Demonstrates the error feedback mechanism and compares with standard optimizers
+"""
+
+import torch
+import torch.nn as nn
+import torch.nn.functional as F
+from code.base_airbench import ErrorFeedbackMuon, ErrorFeedbackQuantizedMuon, NormalizedMuon
+
+# Simple test model
+class SimpleModel(nn.Module):
+    def __init__(self, input_size=10, hidden_size=20, output_size=1):
+        super().__init__()
+        self.fc1 = nn.Linear(input_size, hidden_size)
+        self.fc2 = nn.Linear(hidden_size, hidden_size)
+        self.fc3 = nn.Linear(hidden_size, output_size)
+        
+    def forward(self, x):
+        x = F.relu(self.fc1(x))
+        x = F.relu(self.fc2(x))
+        x = self.fc3(x)
+        return x
+
+def test_optimizer(optimizer_class, optimizer_kwargs, name):
+    """Test an optimizer on a simple optimization problem"""
+    print(f"\n=== Testing {name} ===")
+    
+    # Create model and data
+    model = SimpleModel()
+    x = torch.randn(100, 10)
+    y = torch.randn(100, 1)
+    
+    # Create optimizer
+    optimizer = optimizer_class(model.parameters(), **optimizer_kwargs)
+    
+    # Training loop
+    losses = []
+    for epoch in range(50):
+        optimizer.zero_grad()
+        output = model(x)
+        loss = F.mse_loss(output, y)
+        loss.backward()
+        optimizer.step()
+        losses.append(loss.item())
+        
+        if epoch % 10 == 0:
+            print(f"Epoch {epoch}: Loss = {loss.item():.6f}")
+    
+    print(f"Final loss: {losses[-1]:.6f}")
+    print(f"Loss improvement: {losses[0] - losses[-1]:.6f}")
+    return losses
+
+def main():
+    print("Error Feedback Muon Optimizer Test")
+    print("=" * 50)
+    
+    # Test different optimizers
+    optimizers_to_test = [
+        (NormalizedMuon, {"lr": 0.01, "momentum": 0.9, "sgd_coeff": 0.5}, "NormalizedMuon (Baseline)"),
[…]

### 2025-08-22 code/base_airbench.py
We don't need EF

@@ -3,6 +3,12 @@ airbench94_muon.py
 Runs in 2.59 seconds on a 400W NVIDIA A100
 Attains 94.004 mean accuracy (n=200 trials)
 Descends from https://github.com/tysam-code/hlb-CIFAR10/blob/main/main.py
+
+New Error Feedback Optimizers:
+- ErrorFeedbackMuon: Implements error feedback mechanism to accumulate and compensate for 
+  optimization errors, improving convergence stability
+- ErrorFeedbackQuantizedMuon: Extends error feedback with actual gradient quantization,
+  useful for distributed training with compressed gradients
 """
 
 #############################################
@@ -22,7 +28,7 @@ import torch.nn.functional as F
 import torchvision
 import torchvision.transforms as T
 from matrix_functions import one_sv_svds_approximation
-from optimizers import Dion
+from optimizers import Dion, Neon
 
 torch.backends.cudnn.benchmark = True
 
@@ -30,7 +36,7 @@ torch.backends.cudnn.benchmark = True
 #               Muon optimizer              #
 #############################################
 
-@torch.compile
+# @torch.compile
 def zeropower_via_newtonschulz5(G, steps=3, eps=1e-7):
     """
     Newton-Schulz iteration to compute the zeroth power / orthogonalization of G. We opt to use a
@@ -203,6 +209,134 @@ class RankOneNeon(torch.optim.Optimizer):
                 update = one_sv_svds_approximation(g_resh, 50)
                 # p.data.add_(update, alpha=-lr) # take a step
                 p.data.add_(update.view(g.shape), alpha=-lr)
+
+class ErrorFeedbackMuon(torch.optim.Optimizer):
+    def __init__(self, params, lr=1e-3, momentum=0, nesterov=False, sgd_coeff=0, error_feedback_decay=0.9):
+        defaults = dict(lr=lr, momentum=momentum, nesterov=nesterov)
+        super().__init__(params, defaults)
+        self.sgd_coeff = sgd_coeff
+        self.error_feedback_decay = error_feedback_decay
+
+    def step(self):
+        for group in self.param_groups:
+            lr = group["lr"]
+            momentum = group["momentum"]
+            for p in group["params"]:
+                g = p.grad
+                if g is None:
+                    continue
+                state = self.state[p]
+
+                # Initialize error feedback buffer if not exists
+                if "error_feedback_buffer" not in state.keys():
+                    state["error_feedback_buffer"] = torch.zeros_like(g)
+                
+                # Initialize momentum buffer if not exists
[…]

### 2025-08-22 code/optimizers.py

@@ -221,7 +221,7 @@ class NormalizedMuon(torch.optim.Optimizer):
 
 
 class Dion(torch.optim.Optimizer):
-    def __init__(self, params, lr=1e-3, momentum=0, nesterov=False, rank=1, momentum_decay=0.9):
+    def __init__(self, params, lr=1e-3, momentum=0, nesterov=False, rank=1, momentum_decay=0.9, sgd_coeff=0):
         if lr < 0.0:
             raise ValueError(f"Invalid learning rate: {lr}")
         if momentum < 0.0:
@@ -237,6 +237,7 @@ class Dion(torch.optim.Optimizer):
         super().__init__(params, defaults)
         self.rank = rank
         self.momentum_decay = momentum_decay
+        self.sgd_coeff = sgd_coeff
     '''
     def power_iter1(self, B, Q):
         """
@@ -331,9 +332,9 @@ class Dion(torch.optim.Optimizer):
                 # Step 8: Xt ← Xt−1 − η√(m/n) PtQ⊤t (scaled orthonormal update)
                 update = u @ vt # Pt @ Qt.T
                 # scaled_update = update * math.sqrt(m / n)
-                
+                update2 = (1-self.sgd_coeff) * update + self.sgd_coeff * g_reshaped / (g_reshaped.norm() + 1e-12)
                 # Update the parameter
-                p.data.add_(update.view(g.shape), alpha=-lr)
+                p.data.add_(update2.view(g.shape), alpha=-lr)
                 
                 # Store Q for next iteration
                 # state['Q'] = Qt

### 2025-08-22 conf_article/neon.tex
Refactored related work and added conclusion

@@
 %\title{From Muon to Neon: Introducing Nuclear Norm to Large Matrices}
-\title{The Nuclear Norm and Beyond: Dual Norms and Combinations for Matrix Optimization}
+% \title{The Nuclear Norm and Beyond: Dual Norms and Combinations for Matrix Optimization}
+
+\title{The Ky Fan Norms and Beyond: Dual Norms and Combinations for Matrix Optimization}
 
@@
 \begin{abstract}
+\iffalse
     In this article, we explore the use of matrix norms for optimizing functions of weight matrices, a crucial problem in training large language models.
 Moving beyond the spectral norm that underlies the Muon update, we leverage the nuclear norm, its affine combinations with other norms, and their corresponding duals to develop a new family of Muon-like algorithms.
+We complement our theoretical analysis with an extensive empirical study of the algorithms across a wide range of tasks and settings.
+\fi
+In this article, we explore the use of matrix norms for optimizing functions of weight matrices, a crucial problem in training large language models.
+Moving beyond the spectral norm that underlies the Muon update, we leverage the Ky Fan k-norm, its affine combinations with other norms, and their corresponding duals to develop a new family of Muon-like algorithms.
 We complement our theoretical analysis with an extensive empirical study of the algorithms across a wide range of tasks and settings.
@@
 Besides well-known Adam and AdamW \cite{kingma2014adam}, \cite{loshchilov2017decoupled}, recently proposed Muon has shown promising results on training very large models \cite{jordan2024muon,liu2025muon}.
-Its key difference from earlier algorthims is that it has been constructed specifically for optimizing functions of weight matrices, which are common in deep learning.
+Its key difference from Adam and AdamW is that it has been constructed specifically for optimizing functions of weight matrices, which are common in deep learning.
 
@@
 the spectral norm, to derive the algorithm's update \cite{bernstein2025deriving}.
-Several other attempts have been since made to construct new algorithms, mainly generalizations of Muon's paradigm, like Scion (\cite{pethick2025training} and Gluon \cite{riabinin2025gluon}).
-
 Based upon recent theoretical advances that explain some theory behind Muon, Scion and Gluon (\cite{bernstein2025deriving,kovalev2025understanding,pethick2025training,riabinin2025gluon}), we explore application of other matrix norms to optimization of functions of matrices.
@@
 
- In this article, we focus on the two most common matrix norms akin to the spectral, namely, the Nuclear Norm and the Frobenius norm.
-Working in the linear minimization oracle (lmo) approach, which is equivalent to a factor to the trust region and the steepest descent under norm constraint approaches, we derive Neon, our algorithm based on the nuclear norm.
+ In this article, we focus on the two most common matrix norms akin to the spectral, namely, the nuclear Norm and the Frobenius norm.
+Working in the linear minimization oracle (lmo) framework, which is equivalent to a factor to the trust region approach and the steepest descent under norm constraint, we derive Neon, our algorithm based on the nuclear norm.
 In the section {\it Matrix side of the updates}, we explain how Neon updates can be computed asymptotically faster than Muon updates by the Newton-Schulz iterations.
@@
 \section{Different Norms $\norm{\cdot}$ imply different updates}
-    % As we have stated in Introduction, we are interested foremost in matrix norms, so we'll start right with them, skipping RMS-RMS norm and others.
 
@@
     First, we analyze the problem in the unstochastic case.
-From Corollary 1 of \cite{kovalev2025understanding}, we directly get the following result that matches lower bounds, as was noted in \cite{kovalev2025understanding}.
+From Corollary 1 of \cite{kovalev2025understanding}, we directly get the following result that matches lower bounds, as was noted in that article.
 
@@
 
-    In the stochastic case, from Corollary of 2 of \cite{kovalev2025understanding}, we directly get the following result that matches lower bounds, as was noted in \cite{kovalev2025understanding}:
+    In the stochastic case, from Corollary of 2 of \cite{kovalev2025understanding}, we directly get the following result that pnce more matches lower bounds:
 
@@
 \section{Related Work}
-As Muon \cite{jordan2024muon} is a very successful and popular optimizer for functions of weight matrices, a lot of reseach has been put into explaining its success and further improving its performance.
+As Muon \cite{jordan2024muon} is a very successful and popular optimizer for functions of weight matrices, a lot of reseach has been put into, first, further improving its performance, and, second, in explaining its success.
 
-Regarding the second point, in less than a year, a large number of applications and improvements of Muon has been proposed.
+{\bf Improvements of Muon.} Regarding the first point, in less than a year, a large number of applications and improvements of Muon has been proposed.
 \cite{liu2025muon} adapted the algorthm for training language models larger than NanoGPT.
 \cite{shah2025practical} organized efficient hyperparameter transfer by combining Muon with maximal update parametrization.
-\cite{chen2025cosmoshybridadaptive} applied computationally intensive updates of SOAP optimizer to a low-dimensional "leading eigensubspace" while using memory-efficient methods like Muon for the remaining parameters.
+To construct their COSMOS optimizer, \cite{chen2025cosmoshybridadaptive} applied computationally intensive updates of SOAP optimizer to a low-dimensional "leading eigensubspace" while using memory-efficient methods like Muon for the remaining parameters.
 \cite{amsel2025polar} proposed a more efficient alternative to Newton-Schulz operations.
 \cite{si2025adamuon} introduced AdaMuon which combines element-wise adaptivity with orthogonal updates.
+We suppose that the described above or similar techniques can be applied to our optimizers as well.
+For example, F-Muon also benefits from faster alternatives to Newton-Schulz iterations, and Neon may be a great substitute to Muon in COSMOS, because, as we have shown in {\it the Matrix side of the updates}, Lanczos algorithms is much faster than Newton-Schulz iterations on large matrices.
 
-Regarding the first point, simplistic derivation of \cite{bernstein2025deriving} which is based on \cite{bernstein2024oldoptimizernewnorm} excluded, there has been a gap in theory behind Muon, which, as it seems to us, is not even now completely closed.
-\cite{kovalev2025understanding} has provided convergence guarantees of Muon in various settings, from which, however, Muon's supremacy cannot be recovered.
+{\bf Theory behind Muon.} Regarding the second point, there has been a prolonged gap in theory behind Muon, simplistic derivation of \cite{bernstein2025deriving} based on \cite{bernstein2024oldoptimizernewnorm} excluded.
+This gap, as it seems to us, is not even now completely closed.
+For example, \cite{kovalev2025understanding} has provided convergence guarantees of Muon in various settings, from which, however, Muon's supremacy cannot be recovered.
 Indeed, although the obtained bounds depend on the norm choice, the asymptotics of the convergence remain the same as for NSGD and other optimizers, $K = \cO(\epsilon^{-4})$ in a L-smooth stochastic case.
@@
 
+We suppose that the reason for the recorded by us discrepancy between Neon and Muon performance, both of which are described by Scion or Gluon frameworks, lies in the structure of the norm ball, which must be an object of further research.
+
+{\bf The nuclear norm in lmo.} As we found out only when writing this article, the nuclear norm has been already explored in the context of the linear minimization oracle.
+\cite{pethick2025sam} applied it to create $\nu$SAM, a new sharpness-aware minimization technique.
+The update from their lemma 3.1 coincides with the update of Neon, but is used for completely different purposes.
+In addition, \cite{pethick2025sam} use power iterations to find $u_1$ and $v_1^\top$, while we suggest utilizing much more efficient and precise Lanczos algorithm.
+
+{\bf The Ky Fan Norm and Dion.} Rank-$k$ Centralized Dion, Algorithm 1 from \cite{ahn2025dioncommunicationefficientoptimizerlarge}, without an error feedback and scaling of the update, turns out to be an lmo-based algorithm under the $\normkfk{\cdot}^\dagger$ norm, which we described in {\it Different norms imply different updates}.
+Reported by the authors necessity of using error feedback to obtain satisfactory accuracy may take place in the cases of our algorithms as well, for example, F-Neon.
+This we leave to future research.
+
 \section{Conclusion}
-\begin{itemize}
-    \item Future work: (L0, L1)-smoothness, probably extrapolation (to make dependency on eps milder, but we don't to spend memory here so never mind), hyperparameter-free (Bernstein2023)
-\end{itemize}
+In this article, we have generalized several successful algorithms, like Muon and Dion, to the lmo-based algorithms in the $\normkfk{\cdot}^\dagger$ norm.
+Also we have proposed the technique of ``regularizing'' the updates with NSGD, a trick to increase the robustness of the algorithms and motivated by the consideration of the $\normfstar{\cdot}$ and $\normftwo{\cdot}$ norms.
+Generalizations of well-known norms and subsequent combination of them may further improve performance of lmo-based algorithms.
+If a theory is developed that explains the discrepancy between performance of different algorithms based on the matrix norms, one will probably be able to apply it to the generalizations and combinations of the norms as well, which leaves an open space to construct norms specific for given architectures and probably even their parameters.
 

### 2025-08-22 conf_article/neon.tex
Minor revisions in Updates Section

@@
 %\title{From Muon to Neon: Introducing Nuclear Norm to Large Matrices}
-% \title{The Nuclear Norm and Beyond: Dual Norms and Combinations for Matrix Optimization}
 
@@
 \begin{abstract}
-\iffalse
 In this article, we explore the use of matrix norms for optimizing functions of weight matrices, a crucial problem in training large language models.
-Moving beyond the spectral norm that underlies the Muon update, we leverage the nuclear norm, its affine combinations with other norms, and their corresponding duals to develop a new family of Muon-like algorithms.
-We complement our theoretical analysis with an extensive empirical study of the algorithms across a wide range of tasks and settings.
-\fi
-In this article, we explore the use of matrix norms for optimizing functions of weight matrices, a crucial problem in training large language models.
-Moving beyond the spectral norm that underlies the Muon update, we leverage the Ky Fan k-norm, its affine combinations with other norms, and their corresponding duals to develop a new family of Muon-like algorithms.
+Moving beyond the spectral norm that underlies the Muon update, we leverage the Ky Fan k-norms, their affine combinations with the Frobenius norms, and corresponding duals to develop a new family of Muon-like algorithms.
 We complement our theoretical analysis with an extensive empirical study of the algorithms across a wide range of tasks and settings.
@@
 
-\iffalse
-\begin{itemize}
-    \item Hype of Muon and others:
-    \begin{itemize}
-        \item Importance of the functions of a matrix weight. Special structure.
-        \item Write the equation
-        \item Memory efficiency and hyperparameter transferability
-        \item Proven efficient computing (cite benchmarks and Dion and Cosmos)
-    \end{itemize}
-    \item Comparison:
-    \begin{itemize}
-        \item Shampoo (very compute intensive for large layers, we need asymptotics!), second order
-        \item LMO has produced several effective algorithms: Muon, Dion, Scion, Gluon (citations)
-        \item generalization is better for Muon and others (citations!). Possibly nuSAM here (definitely)
-        \item Scion and Gluon for multilayer
-        \item Gluon for (L0, L1)-smooth problems, which is more realistic (low L0, faster convergence - no impractically small stepsizes)
-    \end{itemize}
-    \item Analyis:
-    \begin{itemize}
-        \item Several generalizations have been already presented by the community: Old Optimizer and New Norm (idea and evidence that it explains famous algorithms); Pethick (LMO and different layers); Kovalev (as trust region and arbitrary norm); Gluon (generalized smoothness model).
-        \item However, to the best of our knowledge, no attempt has been made to try other norms. Here we come in.
-        \item No attempt to mitigate high computation cost of matrix computation has been made. Here again we come in.
-        \item It has been shown that $O(1/esp^4)$ bounds for Muon, Scion, Gluon and even Adam for L-smooth functions (citations).
-Some bounds depend on the dimensionality of space (cite Kovalev), which for spectral norm gives much more iterations (cite Kovalev)
-        \item This motivates us to try new norms
-    \end{itemize}
-\end{itemize}
-\fi
-
 \section{Problem Statement}
-\subsection{Function of a matrix and Assumptions}
-    Warning: the text was copied, while formulas were adapted. We need to rephrase the subsection! 
+\subsection{Function of a matrix}
     We consider the problem of minimizing a differentiable matrix function $F(\cdot)\colon \Rmn \to \R$
@@
 
-    To make the theoretical analysis possible in the future, we make three reasonable non-restrictive assumptions.
-    Idea: move gradient assumptions and smoothness to the L-smooth part. Here we need only custom norm and their relation to the Frobenius norm.
-
-{\bf Non-Euclidean norm setting and Lipschitz continuous gradient.}
-    We assume that matrix space $\Rmn$ is equipped with a norm $\norm{\cdot}\colon \Rmn \to \R_+$, which possibly does not coincide with the Frobenius, norm $\normf{\cdot}$.
-In addition, we assume that the gradient $\nabla f(\cdot)$ is Lipschitz continuous with respect to the norm $\norm{\cdot}$, that is, the following inequality holds:
-    \begin{aequation}\label{eq:L}
-    \norm{\nabla f(\mX) - \nabla f(\mX')}^\dagger \leq L \norm{\mX-\mX'}
-    \quad\text{for all}\;
-    \mX, \mX' \in \Rmn,
-    \end{aequation}
-    where $L > 0$ is the gradient Lipschitz constant, and $\norm{\cdot}^\dagger\colon \Rmn \to \R_+$ is the dual norm associated with $\norm{\cdot}$, i.e., $\norm{\mX}^\dagger = \sup_{\norm{\mX'}\leq 1} \<\mX,\mX'>$ for all $\mX \in \Rmn$.
+    We equip the matrix space $\Rmn$ with a norm $\norm{\cdot}\colon \Rmn \to \R_+$, which possibly does not coincide with the Frobenius norm $\normf{\cdot}$.
+$\norm{\cdot}^\dagger\colon \Rmn \to \R_+$ is the dual norm associated with $\norm{\cdot}$, i.e., $\norm{\mX}^\dagger = \sup_{\norm{\mX'}\leq 1} \<\mX,\mX'>$ for all $\mX \in \Rmn$.
 
 \subsection{Linear Minimization Oracle and Trust Region}
-    Let us look at the problem from the perspective of linear minimization oracle (lmo) and unconstrained stochastic conditional gradient descent (uSCG) (\cite{pethick2025training}).
-lmo is defined as:
+    We analyze the problem from the perspective of linear minimization oracle (lmo) and unconstrained stochastic conditional gradient descent (uSCG) (\cite{pethick2025training}). lmo is defined as:
     \begin{equation}\label{eq:lmo}
@@
 
-    Based on different norms $\norm{\cdot}$, we simplify the update defined by the aforementioned equation:
+    Based on different norms $\norm{\cdot}$, we will simplify the update defined by the aforementioned equation:
 
@@
 
-    In all the work, we define $\mU$, $\mSigma$, $\mV^\top$ as components of the singular value decomposition of $\mM^k$: $\mM^k = \mU \mSigma \mV^\top$.
-We use common notations: $\mU = [u_1, u_2, \dots, u_r]$, $\mSigma = \diag(\sigma_1, \sigma_2, \dots, \sigma_r)$, and  $\mV = [v_1, v_2, \dots, v_r]$.
+    Throughout this work, we denote the singular value decomposition (SVD) of $\mM^k$ as $\mM^k = \mU \mSigma \mV^\top$, where $\mU = [u_1, u_2, \dots, u_r]$, $\mSigma = \diag(\sigma_1, \sigma_2, \dots, \sigma_r)$, and  $\mV = [v_1, v_2, \dots, v_r]$.
 
@@
         \end{lemma}
-        It is an interesting observation, because in other works (\cite{pethick2025training}), $\normf{\cdot}$ was used to recover SGD. The difference is in how one states the problem.
+
+        The result is the same as in (Table 1, \cite{pethick2025training}), but here we use the Euclidean norm as $\normf{\mM^k} = \sqrt{\sum_{i=1}^{\min\{m,n\}}\sigma_i^2}$, which clearly is a matrix norm.
 
@@
 
+    Though the update is well-known, we provide a much simpler proof in the appendix, when compared to \cite{bernstein2024oldoptimizernewnorm}.
+
     {\bf $\normn{\mM^{k}}$ and Neon}
@@
     \end{lemma}
-        We name the derived algorithm \emph{Neon}. In the section Matrix side of updates, we will discuss how to compute an update efficiently.
+        We name the derived algorithm \emph{Neon}. In the section {\it The Matrix side of updates}, we will discuss how to compute the update efficiently.
 
@@
         We name the derived algorithm \emph{F-Muon}.
-It turns out that F-Muon is a convex combination of Normalized SGD and Muon, which is curious.
+It turns out that F-Muon is a convex combination of Normalized SGD and Muon.
 The implications are significant and discussed in the following sections.
@@
         We name the derived algorithm \emph{F-Neon}.
-It turns out that F-Neon is a convex combination of Normalized SGD and Neon, which is curious.
+It turns out that F-Neon is a convex combination of Normalized SGD and Neon.
 The implications are significant and discussed in the following sections.
 
-    {\bf $\normkfk{\mM^k}^\dagger$ and Muon, Neon, and Dion without error feedback}
-        We remind the reader that the Ky Fan k-norm \cite{fan1951maximum}, which we denote as $\normkfk{\cdot}$, is the sum of the k largest singular values of the matrix.
-It can be proven that $\normkfk{\cdot}^\dagger = \max\{\frac{1}{k} \normn{\cdot}, \norms{\cdot}\}$ (see Appendix).
-Special cases of Ky Fan k-norm are the Ky Fan 1-norm, which is the sprectral norm, and the Ky Fan $\min\{m, n\}$-norm, which is the nuclear norm.
+    {\bf $\normkfk{\mM^k}^\dagger$ and Muon, Neon, and Centralized Dion without error feedback}
+        We remind the reader that the Ky Fan k-norm (\cite{bhatia2013matrix}, p.
+92), which we denote as $\normkfk{\cdot}$, is the sum of the k largest singular values of the matrix.
+It can be proved that $\normkfk{\cdot}^\dagger = \max\{\frac{1}{k} \normn{\cdot}, \norms{\cdot}\}$ (see \cite{bhatia2013matrix}, p.
+96).
+Special cases of the Ky Fan k-norm are the Ky Fan 1-norm, which is the sprectral norm, and the Ky Fan $\min\{m, n\}$-norm, which is the nuclear norm.
 
@@
 \subsection{Formal assumptions}
+Warning: the text was copied, while formulas were adapted. We need to rephrase the subsection and probably contract it.
+
+{\bf Non-Euclidean norm setting and Lipschitz continuous gradient.}
+    In addition, we assume that the gradient $\nabla f(\cdot)$ is Lipschitz continuous with respect to the norm $\norm{\cdot}$, that is, the following inequality holds:
+    \begin{aequation}\label{eq:L}
+    \norm{\nabla f(\mX) - \nabla f(\mX')}^\dagger \leq L \norm{\mX-\mX'}
+    \quad\text{for all}\;
+    \mX, \mX' \in \Rmn,
+    \end{aequation}
+    where $L > 0$ is the gradient Lipschitz constant.
+
 {\bf Stochastic gradient estimator.}
@@
 
-\subsection{Norms $\normfstar{\cdot}^\dagger$, $\normftwo{\cdot}^\dagger$, and $\normkfk{\cdot}^\dagger$}
+\subsection{Norms $\normfstar{\cdot}^\dagger$ and $\normftwo{\cdot}^\dagger$}
 Here we provide the derivation of these norms and plot them in the case of $2\times2$ matrices.
+
+\subsection{Visualization of different matrix norms}
+
+TO-DO: Vibe-code F2, F* norm plots and their duals. Add 3D plot of Ky Fan and its dual.
 
@@
 \subsection{Technical details of the experiments}
-~ Note: nuSAM and their update. They had no Lanczos
 

### 2025-08-22 conf_article/neon.tex
Table for lp and Sp algorithms

@@
 
+\usepackage{arydshln} % dashed lines
+
 %------------------------------------------------------------------------------------
@@
 So we derive the algorithms we name F-Neon and F-Muon respectively.
-
- Table that compares methods, updates, and sends to the proofs?
 
@@
 \subsection{Algorithms for Matrices $\leftrightarrow$ Algorithms for Vectors}
-Table that compares Muons and Neons to vector algorithms (NSGD).
+
+lmo optimizers in Schatten $S_p$ norms and in $l_p$ norms with common $p$ may be analogous to each other, as is illustrated by Table \ref{tbl:mat_vs_vec_lmo}.
+The analogies go beyond similarities in the updates.
+SignSGD is very close to Adam, as noted in \cite{bernstein2024oldoptimizernewnorm}, and both Adam and Muon perform well in training large models.
+NSGD is the same for both matrix and vector cases.
+Greedy Coordinate Descent methods are not applied to high-dimensional problems, from this perspective, it is not surprising that Neon underperforms on such problems.
+
+Despite such similarities, no theory has been yet proposed that would reduce the matrix case to the optimization of a function of a singular numbers vector.
+If such a theory is developed, analysis of matrix algorithms like Muon will be greatly simplified.
+
+\begin{table*}[t]
+    \caption{lmo optimizers in Schatten $S_p$ norms and in $l_p$ norms. $g$ is the gradient. When it is a matrix, $g = \mU \mSigma \mV^\top$}
+    \label{tbl:mat_vs_vec_lmo}
+    \bgroup
+    \def\arraystretch{1.2}
+    \resizebox{\textwidth}{!}{
+    \begin{tabular}{|l|c|c|c|}
+    \hline
+    Method & lmo constraint set $\mathcal D$ & lmo & Reference \\
+    \hline
+    \hline
+    Normalized SGD & $l_2$-ball, $S_2$-ball & $-\eta \tfrac{g}{\norm{g}_2} = -\eta \tfrac{g}{\norm{g}_F}$ & \citep{hazan2015beyond} \\
+    Momentum Normalized SGD & Ball in $l_2$, or Ball in $S_2$ & $-\eta \tfrac{g}{\norm{g}_2} = -\eta\tfrac{g}{\norm{g}_F}$ & \citep{cutkosky2020momentum}\\
+    \hline
+    SignSGD & Ball in Max-norm $l_\infty$ & $-\eta \sign(g)$ & \citep[Thm. 1]{bernstein2018signsgd} \\
+    Signum  & Ball in Max-norm $l_\infty$ & $-\eta \sign(g)$ & \citep[Thm. 3]{bernstein2018signsgd} \\
+    \hdashline
+    Muon & Ball in Spectral $S_\infty$ & $-\eta UV^\top$ & \citep{jordan2024muon} \\
+    \hline
+    Gauss-Southwell Coordinate Descent & Ball in $l_1$ & $-\eta \{i: g_i \geq g_k \forall k\}$ & \citep[p. 19]{shi2016primer}\\
+    \hdashline 
+    Neon & Ball in Nuclear $S_1$ & $-\eta u_1 v_1^\top$ & This work\\
+    \hline
+    \end{tabular}
+    }
+    \egroup
+    \end{table*}
 

### 2025-08-23 code/gpt_kyfan_dual_plotter.py
Added Ky Fan and its Dual plots

@@ -1,0 +1,82 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from mpl_toolkits.mplot3d.art3d import Poly3DCollection
+
+# Rebuild after kernel reset
+
+# Define the 12 vertices (all permutations of (±1, ±1, 0))
+verts = []
+base = [(1,1,0), (1,-1,0), (-1,1,0), (-1,-1,0)]
+permute = [(0,1,2), (0,2,1), (1,0,2), (1,2,0), (2,0,1), (2,1,0)]
+unique = set()
+for x,y,z in base:
+    for p in permute:
+        v = ( (x,y,z)[p[0]], (x,y,z)[p[1]], (x,y,z)[p[2]] )
+        unique.add(tuple(v))
+verts = np.array(sorted(list(unique)))
+
+# 6 square faces on planes x=±1, y=±1, z=±1
+def sq_face(plane, sgn):
+    idx = {'x':0,'y':1,'z':2}[plane]
+    face = [v for v in verts if np.isclose(v[idx], sgn)]
+    other = [i for i in [0,1,2] if i != idx]
+    center = np.mean(face, axis=0)
+    angles = []
+    for v in face:
+        vec = np.array([v[other[0]]-center[other[0]], v[other[1]]-center[other[1]]])
+        angle = np.arctan2(vec[1], vec[0])
+        angles.append(angle)
+    order = np.argsort(angles)
+    return [face[i] for i in order]
+
+square_faces = []
+for plane in ['x','y','z']:
+    for sgn in [-1, 1]:
+        square_faces.append(sq_face(plane, sgn))
+
+# 8 triangular faces in planes |x|+|y|+|z|=2 (one per octant)
+tri_faces = []
+for sx in [-1,1]:
+    for sy in [-1,1]:
+        for sz in [-1,1]:
+            tri_faces.append([(sx*1, sy*1, 0), (sx*1, 0, sz*1), (0, sy*1, sz*1)])
+
+faces = square_faces + tri_faces
+
+# Plot solid polyhedron
+fig = plt.figure(figsize=(8,8))
+ax = fig.add_subplot(111, projection='3d')
+poly = Poly3DCollection(faces, alpha=0.95, edgecolor='darkblue', linewidths=1.5)
+ax.add_collection3d(poly)
+ax.scatter(verts[:,0], verts[:,1], verts[:,2], s=20, c='darkblue')
+
+res = 41
[…]

### 2025-08-23 code/gpt_kyfan_plotter.py

@@ -1,0 +1,53 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from mpl_toolkits.mplot3d.art3d import Poly3DCollection
+
+# Generate all vertices that satisfy the constraints
+verts = set()
+for x in [-1, -0.5, 0, 0.5, 1]:
+    for y in [-1, -0.5, 0, 0.5, 1]:
+        for z in [-1, -0.5, 0, 0.5, 1]:
+            if abs(x) + abs(y) <= 1 + 1e-9 and abs(x) + abs(z) <= 1 + 1e-9 and abs(y) + abs(z) <= 1 + 1e-9:
+                verts.add((x, y, z))
+
+verts = [
+    ( 1.0, 0.0, 0.0), (-1.0, 0.0, 0.0),
+    ( 0.0, 1.0, 0.0), ( 0.0,-1.0, 0.0),
+    ( 0.0, 0.0, 1.0), ( 0.0, 0.0,-1.0),
+]
+for sx in (+0.5, -0.5):
+    for sy in (+0.5, -0.5):
+        for sz in (+0.5, -0.5):
+            verts.append((sx, sy, sz))
+verts = np.array(sorted(list(verts)))
+# Convex hull of points
+from scipy.spatial import ConvexHull
+points = np.array(verts)
+hull = ConvexHull(points)
+
+# Plot
+fig = plt.figure(figsize=(8, 8))
+
+ax = fig.add_subplot(111, projection='3d')
+for simplex in hull.simplices:
+    tri = [points[i] for i in simplex]
+    ax.add_collection3d(Poly3DCollection([tri], alpha=0.95, edgecolor='darkblue', linewidths=1.5))
+ax.scatter(verts[:,0], verts[:,1], verts[:,2], s=20, c='darkblue')
+
+ax.set_box_aspect([1,1,1])
+
+ax.set_xlabel('X')
+ax.set_ylabel('Y')
+ax.set_zlabel('Z')
+ax.set_xlim([-1, 1])
+ax.set_ylim([-1, 1])
+ax.set_zlim([-1, 1])
+
+
+ax.set_xticks(np.arange(-1, 1.1, 0.5))
+ax.set_yticks(np.arange(-1, 1.1, 0.5))
+ax.set_zticks(np.arange(-1, 1.1, 0.5))
+plt.show()
+
+fig.savefig('KyFan.pdf', dpi=300, bbox_inches='tight')
+

### 2025-08-23 conf_article/neon.tex
Added Ky Fan and its Dual plots

@@
 \newcommand{\norm}[1]{\lVert #1\rVert}
+\newcommand{\abs}[1]{\lvert #1 \rvert}
 
@@
 \section{Matrix side of updates}
+To compute the algorithms' updates, we use thick-restarted Lanczos method for singular value problem (TRLan) on $A^TA$ or $AA^T$ matrices (the one with less size is picked), implemented in cupy lSibrary [https://docs.cupy.dev/en/stable/reference/generated/cupyx.scipy.sparse.linalg.svds.html] and described in \cite{simonz1998thick}.
+
+This method is designed for efficiently approximating the largest singular values and vectors of large matrices.
+Its thick-restart strategy retains the most informative Ritz vectors at each cycle, which accelerates convergence while avoiding excessive memory growth.
+We are specifically interested in this algorithm because it allows us to extract several largest singular values and related singular vectors of the matrix to make a Neon step.
+Moreover, TRLan is stable GPU-friendly algorithm because it mainly operates with matrix-vector multiplications (MVs), which are higly-parallel, and does not require full reorthogonalization against the whole Krylov basis by managing  short recurrent formulas and incorparating thick restarting.
+
+Per-cycle complexity is $O(mn^2 + n^2k + nk^2)$, where m and n are dimensions of target matrix and n is the smallest one, k is retained subspace's size.
+
 \subsection{Theory}
@@
 \subsection{Norms $\normfstar{\cdot}^\dagger$ and $\normftwo{\cdot}^\dagger$}
+From \cite[p.
+2]{yu2012arithmetic} it follows that the norm ball in $\normfstar{\cdot}^\dagger$ is the Minkowski sum of the norm ball in $\alpha\normn{\cdot}$ and $(1-\alpha)\normf{\cdot}$.
+The norm could be reconstructed from this geometric observation.
+
+Similarly, the norm ball in $\normftwo{\cdot}^\dagger$ is the Minkowski sum of the norm ball in $\alpha\norms{\cdot}$ and $(1-\alpha)\normf{\cdot}$.
+
 Here we provide the derivation of these norms and plot them in the case of $2\times2$ matrices.
@@
 
-TO-DO: Vibe-code F2, F* norm plots and their duals. Add 3D plot of Ky Fan and its dual.
+TO-DO: Vibe-code F2, F* norm plots and their duals.
+
+1-balls in $l_\infty$, $l_1$ and $l_2$ norms are well-known from textbooks. But what about Ky Fan $k$-norm? How can it be represented?
+
+To showcase the complex structure of the Ky Fan $k$-norm and its dual, we suggest the illustrations \ref{fig:kyfan_combined} with the ball in the Ky Fan $2$-norm in \ref{fig:kyfan} and its dual in \ref{fig:kyfandual}.
+On x-, y-, and z-axes, there are singular values $\sigma_1$, $\sigma_2$, and $\sigma_3$ respectively of a matrix from $\Rmn$ with $\min\{m,n\}=3$.
+In this particular case, we do not sort the singular values.
+In the proposed representation, we actually plot balls Top-$2$ norm $\max\{\abs{x}+\abs{y}, \abs{x}+\abs{z}, \abs{y}+\abs{z}\}$ and its dual norm $\max\{\max(\abs(x),\abs{y},\abs{z}), \frac{1}{2} (\abs{x}+\abs{y}+\abs{z})\}$.
+The resulting balls are much more complex than balls in $l_\infty$, $l_1$ and $l_2$ norms.
+
+In fact, those balls can be described easier if we use the results from \cite{yu2012arithmetic}.
+The Ky Fan $2$-norm ball is an intersection of three $l_1$ balls in $(x,y)$, $(x, z)$, and $(y,z)$ spaces.
+The 1-ball in the dual Ky Fan $2$-norm is an intersection of 1-ball the in $l_\infty$ norm and $\frac{1}{2}$-ball in the $l_1$ norm.
+
+\begin{figure}[h!]
+    \centering
+    \begin{subfigure}[b]{0.45\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{KyFan.pdf}
+        \caption{Ky Fan 2-norm}
+        \label{fig:kyfan}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[b]{0.45\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{KyFanDual.pdf}
+        \caption{Dual to Ky Fan 2-norm}
+        \label{fig:kyfandual}
+    \end{subfigure}
+    \caption{Ky Fan 2-norm and its dual}
+    \label{fig:kyfan_combined}
+\end{figure}
 

### 2025-08-23 code/kyfan_plotter.py

@@ -1,0 +1,451 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from mpl_toolkits.mplot3d import Axes3D
+from mpl_toolkits.mplot3d.art3d import Poly3DCollection
+import matplotlib.patches as patches
+from matplotlib.patches import FancyBboxPatch
+import matplotlib.patches as mpatches
+from scipy.spatial import ConvexHull
+from scipy.spatial.distance import cdist
+
+def generate_ball_points(norm_func, num_points=1000, bounds=(-2, 2)):
+    """
+    Generate points on the surface of the ball defined by the given norm function.
+    
+    Args:
+        norm_func: Function that computes the norm of a point
+        num_points: Number of points to generate
+        bounds: Bounds for the search space
+    
+    Returns:
+        points: Array of points on the surface
+    """
+    points = []
+    
+    # Generate random points and filter those with norm close to 1
+    for _ in range(num_points * 10):  # Generate more points to account for filtering
+        if len(points) >= num_points:
+            break
+            
+        # Random point in the search space
+        point = np.random.uniform(bounds[0], bounds[1], 3)
+        
+        # Compute norm
+        norm = norm_func(point)
+        
+        # Check if point is on the surface (within tolerance)
+        if abs(norm - 1.0) < 0.05:
+            points.append(point)
+    
+    return np.array(points)
+
+def max_norm_inf_l1(point, alpha=0.5):
+    """
+    Compute the norm max{||y||_∞, α||y||_1}
+    
+    Args:
+        point: 3D point (x, y, z)
+        alpha: Weight for L1 norm (default 0.5)
+    
+    Returns:
+        norm value
+    """
+    linf_norm = np.max(np.abs(point))
+    l1_norm = np.sum(np.abs(point))
+    return max(linf_norm, alpha * l1_norm)
+
+def plot_3d_ball(norm_func, title="1-Ball in R³", num_points=2000, alpha=0.5):
+    """
+    Plot the 1-ball in R³ as a solid 3D body.
+    
+    Args:
+        norm_func: Function that computes the norm
+        title: Plot title
+        num_points: Number of points to generate
+        alpha: Weight for L1 norm
+    """
[…]

### 2025-08-23 conf_article/neon.tex
Defined Fanions

@@
 In this article, we explore the use of matrix norms for optimizing functions of weight matrices, a crucial problem in training large language models.
-Moving beyond the spectral norm that underlies the Muon update, we leverage the Ky Fan k-norms, their affine combinations with the Frobenius norms, and corresponding duals to develop a new family of Muon-like algorithms.
+Moving beyond the spectral norm that underlies the Muon update, we leverage the Ky Fan k-norms, their affine combinations with the Frobenius norms, and corresponding duals to develop a new family of Muon-like algorithms we name {\it Fanions}.
 We complement our theoretical analysis with an extensive empirical study of the algorithms across a wide range of tasks and settings.
@@
 
-    {\bf $\normf{\mM^{k}}$ and NSGD}
+    \subsection{\texorpdfstring{$\normf{\mM^{k}}$ and Normalized SGD}{NSGD}}
 
@@
 
-    {\bf $\norms{\mM^{k}}$ and Muon}
+    \subsection{\texorpdfstring{$\norms{\mM^{k}}$ and Muon}{Muon}}
     \begin{lemma}\label{lemma:muon_update}
@@
 
-    {\bf $\normn{\mM^{k}}$ and Neon}
+    \subsection{\texorpdfstring{$\normn{\mM^{k}}$ and Neon}{Neon}}
     \begin{lemma}\label{lemma:neon_update}
@@
 
-    {\bf $\normfstar{\mM^{k}}^\dagger$ and F-Muon}
+    \subsection{\texorpdfstring{$\normfstar{\mM^{k}}^\dagger$ and F-Muon}{F-Muon}}
         We define $\normfstar{\cdot}$ as a convex combination of $\normn{\cdot}$ and $\normf{\cdot}$:
@@
 
-        where $\alpha \in [0, 1]$ defines a specific norm of $F*$-family.
+        where $\alpha \in [0, 1]$ defines a specific norm from the $F*$-family.
 
@@
 
-    {\bf $\normftwo{\mM^{k}}^\dagger$ and F-Neon}
+    \subsection{\texorpdfstring{$\normftwo{\mM^{k}}^\dagger$ and F-Neon}{F-Neon}}
         We define $\normftwo{\cdot}$ as a convex combination of $\norms{\cdot}$ and $\normf{\cdot}$:
@@
 
-        where $\alpha \in [0, 1]$ defines a specific norm of F2-family.
+        where $\alpha \in [0, 1]$ defines a specific norm from the F2-family.
 
@@
 
-    {\bf $\normkfk{\mM^k}^\dagger$ and Muon, Neon, and Centralized Dion without error feedback}
+    \subsection{\texorpdfstring{$\normkfk{\mM^k}^\dagger$ and Muon, Neon, and Centralized Dion without error feedback}{Muon, Neon, and Centralized Dion without error feedback}}
         We remind the reader that the Ky Fan k-norm (\cite{bhatia2013matrix}, p.
@@
         \end{lemma}
+        In the section {\it The Matrix side of updates}, we will discuss how to compute the updates efficiently.
 
-        This lemma recovers the updates of Muon, when $k=\min\{m, n\}$, of Neon, when $k=1$, and of Dion \cite{ahn2025dioncommunicationefficientoptimizerlarge} without the error feedback.
+        We introduce the family of {\it Fanions}, which consists of {\it $k$-Fanions}, lmo-based algorithms under the $\normkfk{\mM^k}^\dagger$ norms.
+By this terminology and the lemma, Muon is a $\min\{n,m\}$-Fanion, while Neon is an 1-Fanion.
+Moreover, the centralized version of rank-$r$ Dion \cite{ahn2025dioncommunicationefficientoptimizerlarge} without the error feedback and without scaling of the update, from the perspective of lmo, is actually a $r$-Fanion.
 
-        Moreover, one can consider F-KF-k-norm = $\alpha \normkfk{\cdot} + (1-\alpha)\normf{\cdot}$, the dual to which will produce algorithms like F-Dion without the error feedback.
+        In addition, one can consider F-KF-k-norm = $\alpha \normkfk{\cdot} + (1-\alpha)\normf{\cdot}$, the dual to which will produce algorithms like F-Dion without the error feedback.
+The resulting family can be named {\it F-Fanions}.
 
@@
 
-Here we provide the derivation of these norms and plot them in the case of $2\times2$ matrices.
+\subsection{Visualization of different matrix norms}
+\subsubsection{Duals to F* and F2 norms}
+In \ref{fig:fduals} we plot these norms. On x-axis and y-axis, there are singular values $\sigma_1$, $\sigma_2$ respectively of a matrix from $\Rmn$ with $\min\{m,n\}=2$.
 
-\subsection{Visualization of different matrix norms}
+\begin{figure}[h!]
+    \centering
+    \begin{subfigure}[b]{0.48\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{fstardualball.pdf}
+        \caption{The dual to F* norm for different $\alpha$}
+        \label{fig:fstardual}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[b]{0.48\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{ftwodualball.pdf}
+        \caption{The dual to F2 norm for different $\alpha$}
+        \label{fig:ftwodual}
+    \end{subfigure}
+    \caption{The duals to F* and F2 norms for different $\alpha$}
+    \label{fig:fduals}
+\end{figure}
 
-TO-DO: Vibe-code F2, F* norm plots and their duals.
-
-1-balls in $l_\infty$, $l_1$ and $l_2$ norms are well-known from textbooks. But what about Ky Fan $k$-norm? How can it be represented?
+\subsubsection{The Ky Fan norm and its dual}
+1-balls in $l_\infty$, $l_1$ and $l_2$ norms are well-known from textbooks. But what about the Ky Fan $k$-norm? How can it be represented?
 

### 2025-08-23 code/norms/fstar_dual_plotter.py

@@ -1,0 +1,66 @@
+import numpy as np
+import matplotlib.cm as cm
+import matplotlib.pyplot as plt
+
+def minkowski_sum_boundary(alpha, num_angles=2000):
+    r1 = alpha            # L1-ball radius (diamond)
+    r2 = 1.0 - alpha      # L2-ball radius (circle)
+    thetas = np.linspace(0, 2*np.pi, num_angles, endpoint=False)
+    pts = []
+    for th in thetas:
+        u = np.array([np.cos(th), np.sin(th)])  # direction
+        
+        # Support point on r2 * L2 ball is simply r2 * u
+        s2 = r2 * u
+        
+        # Support point on r1 * L1 ball:
+        # pick the axis with larger |component|, place full mass there with the sign
+        if abs(u[0]) >= abs(u[1]):
+            s1 = np.array([r1 * np.sign(u[0]), 0.0])
+        else:
+            s1 = np.array([0.0, r1 * np.sign(u[1])])
+        
+        pts.append(s1 + s2)
+    return np.array(pts)
+
+def l1_ball_boundary(radius, num=800):
+    t = np.linspace(0, 2*np.pi, num, endpoint=False)
+    x = np.cos(t)
+    y = np.sin(t)
+    denom = np.maximum(np.abs(x) + np.abs(y), 1e-12)
+    return radius * x / denom, radius * y / denom
+
+def l2_ball_boundary(radius, num=800):
+    t = np.linspace(0, 2*np.pi, num, endpoint=False)
+    return radius * np.cos(t), radius * np.sin(t)
+
+# List of alpha values to test
+alphas = [0.25, 0.5, 0.75]
+colormap = cm.get_cmap('viridis')
+colors = [colormap(i / (len(alphas) - 1)) for i in range(len(alphas) - 1, -1, -1)]
+# Plot
+plt.figure(figsize=(8, 8))
+
+# Plot Minkowski sums for each alpha
+for alpha, color in zip(alphas, colors):
+    sum_pts = minkowski_sum_boundary(alpha)
+    plt.plot(sum_pts[:, 0], sum_pts[:, 1], label=f"F*$^\dagger$-ball (α={alpha})", 
+             linewidth=2, color=color)
+
+# Plot reference L1 and L2 balls
+x1, y1 = l1_ball_boundary(1)
+x2, y2 = l2_ball_boundary(1)
[…]

### 2025-08-23 code/norms/ftwo_dual_plotter.py

@@ -1,0 +1,68 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.cm as cm
+
+def minkowski_sum_boundary(alpha, num_angles=2000):
+    r1 = alpha            # L∞-ball radius (square)
+    r2 = 1.0 - alpha      # L2-ball radius (circle)
+    thetas = np.linspace(0, 2*np.pi, num_angles, endpoint=False)
+    pts = []
+    for th in thetas:
+        u = np.array([np.cos(th), np.sin(th)])  # direction
+        
+        # Support point on r2 * L2 ball is simply r2 * u
+        s2 = r2 * u
+        
+        # Support point on r1 * L∞ ball:
+        # For L∞ ball, the support point in direction u is r1 * sign(u)
+        s1 = r1 * np.sign(u)
+        
+        pts.append(s1 + s2)
+    return np.array(pts)
+
+def linfty_ball_boundary(radius, num=800):
+    # Parametrize the L∞ "circle" (square) using angle and normalization
+    t = np.linspace(0, 2*np.pi, num, endpoint=False)
+    x = np.cos(t)
+    y = np.sin(t)
+    # Compute the maximum of |x| and |y| element-wise, then ensure no division by zero
+    denom = np.maximum(np.abs(x), np.abs(y))
+    denom = np.maximum(denom, 1e-12)  # Avoid division by zero
+    return radius * x / denom, radius * y / denom
+
+def l2_ball_boundary(radius, num=800):
+    t = np.linspace(0, 2*np.pi, num, endpoint=False)
+    return radius * np.cos(t), radius * np.sin(t)
+
+# List of alpha values to test
+alphas = [0.25, 0.5, 0.75]
+# Use viridis colormap for nicer color gradient
+colormap = cm.get_cmap('viridis')
+colors = [colormap(i / (len(alphas) - 1)) for i in range(len(alphas) - 1, -1, -1)]
+
+# Plot
+plt.figure(figsize=(8, 8))
+
+# Plot Minkowski sums for each alpha
+for alpha, color in zip(alphas, colors):
+    sum_pts = minkowski_sum_boundary(alpha)
+    plt.plot(sum_pts[:, 0], sum_pts[:, 1], label=f"F2$^\dagger$-ball (α={alpha})", 
+             linewidth=2, color=color)
+
[…]

### 2025-08-23 conf_article/neon.tex
Added expressions for f2 and f* duals

@@
         where $\alpha \in [0, 1]$ defines a specific norm from the $F*$-family.
+$\normfstar{\cdot}^\dagger$ can be expressed as in \ref{eq:normfstardual}.
+However, its norm ball is a simple Minkowski sum of $\alpha \normn{\cdot}$ and $(1-\alpha)\normf{\cdot}$ balls \ref{fig:fstardual}.
 
@@
 
-        where $\alpha \in [0, 1]$ defines a specific norm from the F2-family.
+        where $\alpha \in [0, 1]$ defines a specific norm from the F2-family.$\normftwo{\cdot}^\dagger$ can be expressed as in \ref{eq:normftwodual}.
+However, its norm ball is a simple Minkowski sum of $\alpha \norms{\cdot}$ and $(1-\alpha)\normf{\cdot}$ balls \ref{fig:ftwodual}.
 
@@
 
-Despite such similarities, no theory has been yet proposed that would reduce the matrix case to the optimization of a function of a singular numbers vector.
+Despite such similarities, no theory has been yet proposed that would reduce the matrix case to the optimization of a function of a singular values vector.
 If such a theory is developed, analysis of matrix algorithms like Muon will be greatly simplified.
@@
 \subsection{Norms $\normfstar{\cdot}^\dagger$ and $\normftwo{\cdot}^\dagger$}
-From \cite[p.
-2]{yu2012arithmetic} it follows that the norm ball in $\normfstar{\cdot}^\dagger$ is the Minkowski sum of the norm ball in $\alpha\normn{\cdot}$ and $(1-\alpha)\normf{\cdot}$.
-The norm could be reconstructed from this geometric observation.
+First, we need a well-known fact mentioned, for example, in \cite[Table 1]{yu2012arithmetic}. For the sake of completeness, we provide a proof of the fact.
 
-Similarly, the norm ball in $\normftwo{\cdot}^\dagger$ is the Minkowski sum of the norm ball in $\alpha\norms{\cdot}$ and $(1-\alpha)\normf{\cdot}$.
+\begin{lemma}\label{lem:dual_to_conv_comb}
+    Let $\norm{\cdot}_{(1)}$ and $\norm{\cdot}_{(2)}$ be norms on a finite-dimensional Euclidean space, and let $\alpha,\beta \geq 0$. Define 
+    $$
+    \norm{x} := \alpha \norm{x}_{(1)} + \beta \norm{x}_{(2)}.
+    $$
+    Then the dual unit ball of $\norm{\cdot}$ satisfies
+    $$
+    B_{\norm{\cdot}^\dagger} 
+    = \alpha B_{\norm{\cdot}_{(1)}^\dagger} + \beta B_{\norm{\cdot}_{(2)}^\dagger},
+    $$
+    where $+$ denotes the Minkowski sum and $B_{\norm{\cdot}_{(i)}^\dagger}$ is the unit ball of the dual norm $\norm{\cdot}_{(i)}^\dagger$.
+\end{lemma}
+
+    \begin{proof}
+    Write $f(x) = \alpha \norm{x}_{(1)}$ and $g(x) = \beta \norm{x}_{(2)}$, so 
+    $$
+    \norm{x} = f(x)+g(x).
+    $$
+    Recall two standard facts:
+    \begin{enumerate}
+        \item For any norm $\norm{\cdot}$ and $\lambda>0$, 
+        $$
+        (\lambda \norm{\cdot})^*(y) = 
+        \sup_{x}\bigl( \langle y,x \rangle - \lambda \norm{x}\bigr)
+        = \delta_{\lambda B_{\norm{\cdot}^\dagger}}(y),
+        $$
+        i.e., the indicator function of the scaled dual ball.
+        \item The Fenchel conjugate of a sum satisfies
+        $$
+        (f+g)^*(y) = \inf_{u+v=y} \bigl(f^*(u) + g^*(v)\bigr).
+        $$
+    \end{enumerate}
+    Applying these to $f$ and $g$, we have 
+    $$
+    f^*(u) = \delta_{\alpha B_{\norm{\cdot}_{(1)}^\dagger}}(u), 
+    \quad 
+    g^*(v) = \delta_{\beta B_{\norm{\cdot}_{(2)}^\dagger}}(v).
+    $$
+    Thus
+    $$
+    \norm{\cdot}^*(y) 
+    = (f+g)^*(y)
+    = \inf_{u+v=y}
+    \bigl(
+    \delta_{\alpha B_{\norm{\cdot}_{(1)}^\dagger}}(u)
+    +
+    \delta_{\beta B_{\norm{\cdot}_{(2)}^\dagger}}(v)
+    \bigr)
+    = 
+    \delta_{\alpha B_{\norm{\cdot}_{(1)}^\dagger}+\beta B_{\norm{\cdot}_{(2)}^\dagger}}(y).
+    $$
+    But by definition, the conjugate of a norm is exactly the indicator of its dual unit ball:
+    $$
+    \norm{\cdot}^*(y) = \delta_{B_{\norm{\cdot}^\dagger}}(y).
+    $$
+    Therefore,
+    $$
+    B_{\norm{\cdot}^\dagger} = \alpha B_{\norm{\cdot}_{(1)}^\dagger} + \beta B_{\norm{\cdot}_{(2)}^\dagger}.
+    $$
+    \end{proof}
+
+Consequently,
+\begin{equation}\label{eq:dual_to_conv_comb}
+\norm{y}^\dagger
+= \inf\bigl\{ t \ge 0 : y \in t\bigl(\alpha B_{\norm{\cdot}_{(1)}^\dagger} + \beta B_{\norm{\cdot}_{(2)}^\dagger}\bigr) \bigr\}
+= \inf_{z,t}\bigl\{t \ge 0: z \in t \alpha B_{\norm{\cdot}_{(1)}^\dagger}, y - z \in t \beta B_{\norm{\cdot}_{(1)}^\dagger}\bigr\}
+\end{equation}
+
+Thus, we immediately find $\normfstar{\cdot}^\dagger$, which is related to F-Muon update.
+Indeed, after setting $\beta=1-\alpha$ and remembering that for smooth and bounded cases we can use $\min$ instead of $\inf$, we get
+\begin{equation}\label{eq:normfstardual}
+\normfstar{\mY}^\dagger = \min_{\mZ} \min_t\{t, s.t. \norms{\mZ}\leq \alpha t, \normf{\mY-\mZ}\leq (1-\alpha) t\}
+\end{equation}
+
+If $\alpha = 1$, then $\mZ = \mY$, and we get $\normfstar{\mY}^\dagger = \norms{\mY}$. If $\alpha = 0$, then $\mZ = 0$, and we get $\normfstar{\mY}^\dagger = \normf{\mY}$.
+
+Similarly, we find $\normftwo{\cdot}^\dagger$, which is related to F-Neon update:
+\begin{equation}\label{eq:normftwodual}
+\normftwo{\mY}^\dagger = \min_{\mZ} \min_t\{t, s.t. \normn{\mZ}\leq \alpha t, \normf{\mY-\mZ}\leq (1-\alpha) t\}
+\end{equation}
+
+If $\alpha = 1$, then $\mZ = \mY$, and we get $\normftwo{\mY}^\dagger = \normn{\mY}$. If $\alpha = 0$, then $\mZ = 0$, and we get $\normftwo{\mY}^\dagger = \normf{\mY}$.
 
@@
 \subsubsection{Duals to F* and F2 norms}
+
+ It follows from \ref{lem:dual_to_conv_comb} that the norm ball in $\normfstar{\cdot}^\dagger$ is the Minkowski sum of the norm ball in $\alpha\normn{\cdot}$ and $(1-\alpha)\normf{\cdot}$ and the norm ball in $\normftwo{\cdot}^\dagger$ is the Minkowski sum of the norm ball in $\alpha\norms{\cdot}$ and $(1-\alpha)\normf{\cdot}$.
+
 In \ref{fig:fduals} we plot these norms. On x-axis and y-axis, there are singular values $\sigma_1$, $\sigma_2$ respectively of a matrix from $\Rmn$ with $\min\{m,n\}=2$.
@@
         \includegraphics[width=\linewidth]{fstardualball.pdf}
-        \caption{The dual to F* norm for different $\alpha$}
+        \caption{lmo balls for F-Muon for different $\alpha$}
         \label{fig:fstardual}
@@
         \includegraphics[width=\linewidth]{ftwodualball.pdf}
-        \caption{The dual to F2 norm for different $\alpha$}
+        \caption{lmo ball for F-Neon for different $\alpha$}
         \label{fig:ftwodual}
     \end{subfigure}
-    \caption{The duals to F* and F2 norms for different $\alpha$}
+    \caption{Balls in the duals to F* and F2 norms for different $\alpha$}
     \label{fig:fduals}
@@
 In this particular case, we do not sort the singular values.
-In the proposed representation, we actually plot balls Top-$2$ norm $\max\{\abs{x}+\abs{y}, \abs{x}+\abs{z}, \abs{y}+\abs{z}\}$ and its dual norm $\max\{\max(\abs(x),\abs{y},\abs{z}), \frac{1}{2} (\abs{x}+\abs{y}+\abs{z})\}$.
+In the proposed representation, we actually plot balls in the Top-$2$ norm $\max\{\abs{x}+\abs{y}, \abs{x}+\abs{z}, \abs{y}+\abs{z}\}$ and its dual norm $\max\{\max(\abs(x),\abs{y},\abs{z}), \frac{1}{2} (\abs{x}+\abs{y}+\abs{z})\}$.
 The resulting balls are much more complex than balls in $l_\infty$, $l_1$ and $l_2$ norms.

### 2025-08-24 code/airbench_muon.py
Tuned to test CIFAR

@@ -1,0 +1,449 @@
+"""
+airbench94_muon.py
+Runs in 2.59 seconds on a 400W NVIDIA A100 using torch==2.4.1
+Attains 94.01 mean accuracy (n=200 trials)
+Descends from https://github.com/tysam-code/hlb-CIFAR10/blob/main/main.py
+"""
+
+#############################################
+#                  Setup                    #
+#############################################
+
+import os
+import sys
+with open(sys.argv[0]) as f:
+    code = f.read()
+import uuid
+from math import ceil
+
+import torch
+from torch import nn
+import torch.nn.functional as F
+import torchvision
+import torchvision.transforms as T
+
+torch.backends.cudnn.benchmark = True
+
+#############################################
+#               Muon optimizer              #
+#############################################
+
+@torch.compile
+def zeropower_via_newtonschulz5(G, steps=3, eps=1e-7):
+    """
+    Newton-Schulz iteration to compute the zeroth power / orthogonalization of G. We opt to use a
+    quintic iteration whose coefficients are selected to maximize the slope at zero. For the purpose
+    of minimizing steps, it turns out to be empirically effective to keep increasing the slope at
+    zero even beyond the point where the iteration no longer converges all the way to one everywhere
+    on the interval. This iteration therefore does not produce UV^T but rather something like US'V^T
+    where S' is diagonal with S_{ii}' \sim Uniform(0.5, 1.5), which turns out not to hurt model
+    performance at all relative to UV^T, where USV^T = G is the SVD.
+    """
+    assert len(G.shape) == 2
+    a, b, c = (3.4445, -4.7750,  2.0315)
+    X = G.bfloat16()
+    X /= (X.norm() + eps) # ensure top singular value <= 1
+    if G.size(0) > G.size(1):
+        X = X.T
+    for _ in range(steps):
+        A = X @ X.T
+        B = b * A + c * A @ A
+        X = a * X + B @ X
+    if G.size(0) > G.size(1):
+        X = X.T
+    return X
+
+class Muon(torch.optim.Optimizer):
+    def __init__(self, params, lr=1e-3, momentum=0, nesterov=False):
+        if lr < 0.0:
+            raise ValueError(f"Invalid learning rate: {lr}")
+        if momentum < 0.0:
+            raise ValueError(f"Invalid momentum value: {momentum}")
+        if nesterov and momentum <= 0:
+            raise ValueError("Nesterov momentum requires a momentum")
[…]

### 2025-08-24 code/base_airbench.py
Tuned to test CIFAR

@@ -645,7 +645,8 @@ def main(run, model):
                      dict(params=norm_biases, lr=bias_lr, weight_decay=wd/bias_lr),
                      dict(params=[model.head.weight], lr=head_lr, weight_decay=wd/head_lr)]
     optimizer1 = torch.optim.SGD(param_configs, momentum=0.85, nesterov=True)#, fused=True)
-    optimizer2 = NormalizedMuon(filter_params, lr=0.5, momentum=0.6, sgd_coeff=0.5, nesterov=True) # important
+    optimizer2 = NormalizedMuon(filter_params, lr=0.4, momentum=0.65, sgd_coeff=0.5, nesterov=True) # important
+    # optimizer2 = NormalizedMuon(filter_params, lr=0.24, momentum=0.6, sgd_coeff=0.6, nesterov=True) # important
     # optimizer2 = DistrMuonNormalized(filter_params, lr=0.05, momentum=0.95, sgd_coeff=0.5) # important
     
     # optimizer2 = Neon(filter_params, neon_mode='kyfan', lr=0.45, momentum=0.65, nesterov=True, sgd_coeff=0.6)
@@ -658,6 +659,7 @@ def main(run, model):
     for opt in optimizers:
         for group in opt.param_groups:
             group["initial_lr"] = group["lr"]
+            group["target_momentum"] = group.get("momentum", 0)  # default to 0 if not set
 
     # For accurately timing GPU code
     starter = torch.cuda.Event(enable_timing=True)
@@ -704,6 +706,12 @@ def main(run, model):
                 group["lr"] = group["initial_lr"] * (1 - step / whiten_bias_train_steps)
             for group in optimizer1.param_groups[1:]+optimizer2.param_groups:
                 group["lr"] = group["initial_lr"] * (1 - step / total_train_steps)
+            for group in optimizer2.param_groups:
+                eta = (step / total_train_steps) # percentage of training
+                # group["lr"] = (0.45 * eta + 0.3 * (1 - eta)) * (1-eta)
+            for group in optimizer2.param_groups:
+                eta = (step / total_train_steps) # percentage of training
+                group["momentum"] = (group["target_momentum"] - 0.1) * eta + group["target_momentum"] * (1 - eta)
             for opt in optimizers:
                 opt.step()
             model.zero_grad(set_to_none=True)
@@ -742,7 +750,7 @@ if __name__ == "__main__":
 
     print_columns(logging_columns_list, is_head=True)
[…]

### 2025-08-24 code/airbench_muon.py
Preparing F-Muon

@@ -84,6 +84,39 @@ class Muon(torch.optim.Optimizer):
                 update = zeropower_via_newtonschulz5(g.reshape(len(g), -1)).view(g.shape) # whiten the update
                 p.data.add_(update, alpha=-lr) # take a step
 
+class NormalizedMuon(torch.optim.Optimizer):
+    def __init__(self, params, lr=1e-3, momentum=0, nesterov=False, sgd_coeff=0):
+        defaults = dict(lr=lr, momentum=momentum, nesterov=nesterov)
+        super().__init__(params, defaults)
+        self.sgd_coeff = sgd_coeff
+
+    def step(self):
+        for group in self.param_groups:
+            lr = group["lr"]
+            momentum = group["momentum"]
+            for p in group["params"]:
+                g = p.grad
+                if g is None:
+                    continue
+                state = self.state[p]
+
+                if "momentum_buffer" not in state.keys():
+                    state["momentum_buffer"] = torch.zeros_like(g)
+                buf = state["momentum_buffer"]
+                buf.mul_(momentum).add_(g)
+                g = g.add(buf, alpha=momentum) if group["nesterov"] else buf
+
+                p.data.mul_(len(p.data)**0.5 / p.data.norm()) # normalize the weight
+                # update = zeropower_via_newtonschulz5(g.reshape(len(g), -1)).view(g.shape) # whiten the update
+                eps = 1e-12
+                g_normalized = g / (g.norm() + eps)
+                if self.sgd_coeff != 1:
+                    update_part = zeropower_via_newtonschulz5(g.reshape(len(g), -1)).view(g.shape)
+                    update = (1-self.sgd_coeff) * update_part + self.sgd_coeff * g_normalized
+                else:
+                    update = self.sgd_coeff * g_normalized
+                p.data.add_(update, alpha=-lr) # take a step
+
 #############################################
 #                DataLoader                 #
 #############################################
@@ -359,11 +392,12 @@ def main(run, model):
                      dict(params=norm_biases,         lr=bias_lr, weight_decay=wd/bias_lr),
                      dict(params=[model.head.weight], lr=head_lr, weight_decay=wd/head_lr)]
     optimizer1 = torch.optim.SGD(param_configs, momentum=0.85, nesterov=True, fused=True)
-    optimizer2 = Muon(filter_params, lr=0.24, momentum=0.6, nesterov=True)
[…]

### 2025-08-24 code/norms/muon_fmuon_cifar.py
Added Matrix methods, contributions, and a Cifar draft

@@ -1,0 +1,60 @@
+import numpy as np
+import matplotlib.cm as cm
+import matplotlib.pyplot as plt
+
+def minkowski_sum_boundary(alpha, num_angles=2000):
+    r1 = alpha            # L1-ball radius (diamond)
+    r2 = 1.0 - alpha      # L2-ball radius (circle)
+    thetas = np.linspace(0, 2*np.pi, num_angles, endpoint=False)
+    pts = []
+    for th in thetas:
+        u = np.array([np.cos(th), np.sin(th)])  # direction
+        
+        # Support point on r2 * L2 ball is simply r2 * u
+        s2 = r2 * u
+        
+        # Support point on r1 * L1 ball:
+        # pick the axis with larger |component|, place full mass there with the sign
+        if abs(u[0]) >= abs(u[1]):
+            s1 = np.array([r1 * np.sign(u[0]), 0.0])
+        else:
+            s1 = np.array([0.0, r1 * np.sign(u[1])])
+        
+        pts.append(s1 + s2)
+    return np.array(pts)
+
+def l1_ball_boundary(radius, num=800):
+    t = np.linspace(0, 2*np.pi, num, endpoint=False)
+    x = np.cos(t)
+    y = np.sin(t)
+    denom = np.maximum(np.abs(x) + np.abs(y), 1e-12)
+    return radius * x / denom, radius * y / denom
+
+# List of alpha values to test
+alphas = [0.5]
+colormap = cm.get_cmap('viridis')
+colors = [colormap(i / (len(alphas))) for i in range(len(alphas) - 1, -1, -1)]
+# Plot
+plt.figure(figsize=(8, 8))
+
+# Plot Minkowski sums for each alpha
+for alpha, color in zip(alphas, colors):
+    sum_pts = 0.4 * minkowski_sum_boundary(alpha)
+    plt.plot(sum_pts[:, 0], sum_pts[:, 1], label=f"F-Muon ball (α={alpha}, lr=0.4)", 
+             linewidth=2, color=color)
+
+# Plot reference L1 and L2 balls
+x1, y1 = l1_ball_boundary(0.24)
+plt.plot(x1, y1, linestyle="--", label="Muon ball (lr=0.24)", color="black", alpha=0.9)
+
+# Configure plot
+plt.gca().set_aspect("equal", adjustable="box")
[…]

### 2025-08-24 conf_article/neon.tex
Added Matrix methods, contributions, and a Cifar draft

@@
 \begin{abstract}
-In this article, we explore the use of matrix norms for optimizing functions of weight matrices, a crucial problem in training large language models.
+% In this article, we explore the use of matrix norms for optimizing functions of weight matrices, a crucial problem in training large language models.
 Moving beyond the spectral norm that underlies the Muon update, we leverage the Ky Fan k-norms, their affine combinations with the Frobenius norms, and corresponding duals to develop a new family of Muon-like algorithms we name {\it Fanions}.
 We complement our theoretical analysis with an extensive empirical study of the algorithms across a wide range of tasks and settings.
+
+In this article, we explore the use of matrix norms for optimizing functions of weight matrices, a crucial problem in training large language models.
+Moving beyond the spectral norm that underlies the Muon update, we leverage the Ky Fan k-norms to introduce a new family of Muon-like algorithms we name {\it Fanions}.
+Then, we consider their convex combinations with the Frobenius norm and corresponding duals to develop a new family of algorithms we name {\it F-Fanions}, one of them being {\it F-Muon}.
+We complement our theoretical analysis with an extensive empirical study of the algorithms across a wide range of tasks and settings, from which it follows that properly-tuned F-Muon is on par with Muon, which raises the question about Muon's optimality.
+
 \end{abstract}
@@
 Hence, new scalable, time- and memory-efficient algorithms have been demanded.
-Besides well-known Adam and AdamW \cite{kingma2014adam}, \cite{loshchilov2017decoupled}, recently proposed Muon has shown promising results on training very large models \cite{jordan2024muon,liu2025muon}.
+Besides well-known Adam \citep{kingma2014adam} and AdamW \citep{loshchilov2017decoupled}, recently proposed Muon has shown promising results on training very large models \citep{jordan2024muon,liu2025muon}.
 Its key difference from Adam and AdamW is that it has been constructed specifically for optimizing functions of weight matrices, which are common in deep learning.
@@
 From the perspective of theory, Muon's main innovation was an intentional usage of matrix norms, i.e.
-the spectral norm, to derive the algorithm's update \cite{bernstein2025deriving}.
-Based upon recent theoretical advances that explain some theory behind Muon, Scion and Gluon (\cite{bernstein2025deriving,kovalev2025understanding,pethick2025training,riabinin2025gluon}), we explore application of other matrix norms to optimization of functions of matrices.
+the spectral norm, to derive the algorithm's update \citep{bernstein2025deriving}.
+Based upon recent theoretical advances that explain some theory behind Muon, Scion and Gluon \citep{bernstein2025deriving,kovalev2025understanding,pethick2025training,riabinin2025gluon}, we explore application of other matrix norms to optimization of functions of matrices.
 As it has been done with Muon, we stipulate that our algorithms' updates be fast to compute.
@@
 
- Having faced the array of different Muon-like optimizers, according to the upper bounds from \cite{kovalev2025understanding,riabinin2025gluon}, with similar convergence behavior, we painstakingly compare them on a synthetic linear least squares problem with known Lipschitz constant.
+ Having faced the array of different Muon-like optimizers, according to the upper bounds from \citet{kovalev2025understanding,riabinin2025gluon}, with similar convergence behavior, we painstakingly compare them on a synthetic linear least squares problem with known Lipschitz constant.
 The efforts results in comparison of the algorithms by their convergence not in terms of the dual norms of their gradient, but in terms of the common spectral norm, which may strongly differ, especially in large matrices, from the initial dual norms.
@@
 \subsection{Linear Minimization Oracle and Trust Region}
-    We analyze the problem from the perspective of linear minimization oracle (lmo) and unconstrained stochastic conditional gradient descent (uSCG) (\cite{pethick2025training}). lmo is defined as:
+    We analyze the problem from the perspective of linear minimization oracle (lmo) and unconstrained stochastic conditional gradient descent (uSCG) \citep{pethick2025training}. lmo is defined as:
     \begin{equation}\label{eq:lmo}
@@
     Let us set $\gamma_k = 1$.
-This transforms algorithm defined by \cref{eq:almost_trust_update} into Algorithm 1 from \cite{kovalev2025understanding}.
+This transforms algorithm defined by \cref{eq:almost_trust_update} into Algorithm 1 from \citet{kovalev2025understanding}.
 Therefore, we can view \cref{eq:almost_trust_update} both as an lmo-based algorithm and as a trust-region algorithm.
@@
 
-        The result is the same as in (Table 1, \cite{pethick2025training}), but here we use the Euclidean norm as $\normf{\mM^k} = \sqrt{\sum_{i=1}^{\min\{m,n\}}\sigma_i^2}$, which clearly is a matrix norm.
+        The result is the same as in (Table 1, \citet{pethick2025training}), but here we use the Euclidean norm as $\normf{\mM^k} = \sqrt{\sum_{i=1}^{\min\{m,n\}}\sigma_i^2}$, which clearly is a matrix norm.
 
@@
 
-    Though the update is well-known, we provide a much simpler proof in the appendix, when compared to \cite{bernstein2024oldoptimizernewnorm}.
+    Though the update is well-known, we provide a much simpler proof in the appendix, when compared to \citet{bernstein2024oldoptimizernewnorm}.
 
@@
         where $\alpha \in [0, 1]$ defines a specific norm from the $F*$-family.
-$\normfstar{\cdot}^\dagger$ can be expressed as in \ref{eq:normfstardual}.
-However, its norm ball is a simple Minkowski sum of $\alpha \normn{\cdot}$ and $(1-\alpha)\normf{\cdot}$ balls \ref{fig:fstardual}.
+$\normfstar{\cdot}^\dagger$ can be expressed as in \cref{eq:normfstardual}.
+However, its norm ball is a simple Minkowski sum of $\alpha \normn{\cdot}$ and $(1-\alpha)\normf{\cdot}$ balls \cref{fig:fstardual}.
 
@@
 
-        where $\alpha \in [0, 1]$ defines a specific norm from the F2-family.$\normftwo{\cdot}^\dagger$ can be expressed as in \ref{eq:normftwodual}.
-However, its norm ball is a simple Minkowski sum of $\alpha \norms{\cdot}$ and $(1-\alpha)\normf{\cdot}$ balls \ref{fig:ftwodual}.
+        where $\alpha \in [0, 1]$ defines a specific norm from the F2-family.$\normftwo{\cdot}^\dagger$ can be expressed as in \cref{eq:normftwodual}.
+However, its norm ball is a simple Minkowski sum of $\alpha \norms{\cdot}$ and $(1-\alpha)\normf{\cdot}$ balls \cref{fig:ftwodual}.
 
@@
     \subsection{\texorpdfstring{$\normkfk{\mM^k}^\dagger$ and Muon, Neon, and Centralized Dion without error feedback}{Muon, Neon, and Centralized Dion without error feedback}}
-        We remind the reader that the Ky Fan k-norm (\cite{bhatia2013matrix}, p.
+        We remind the reader that the Ky Fan k-norm (\citet{bhatia2013matrix}, p.
 92), which we denote as $\normkfk{\cdot}$, is the sum of the k largest singular values of the matrix.
-It can be proved that $\normkfk{\cdot}^\dagger = \max\{\frac{1}{k} \normn{\cdot}, \norms{\cdot}\}$ (see \cite{bhatia2013matrix}, p.
+It can be proved that $\normkfk{\cdot}^\dagger = \max\{\frac{1}{k} \normn{\cdot}, \norms{\cdot}\}$ (see \citet{bhatia2013matrix}, p.
 96).
@@
 By this terminology and the lemma, Muon is a $\min\{n,m\}$-Fanion, while Neon is an 1-Fanion.
-Moreover, the centralized version of rank-$r$ Dion \cite{ahn2025dioncommunicationefficientoptimizerlarge} without the error feedback and without scaling of the update, from the perspective of lmo, is actually a $r$-Fanion.
+Moreover, the centralized version of rank-$r$ Dion \citep{ahn2025dioncommunicationefficientoptimizerlarge} without the error feedback and without scaling of the update, from the perspective of lmo, is actually a $r$-Fanion.
 
@@
 
-lmo optimizers in Schatten $S_p$ norms and in $l_p$ norms with common $p$ may be analogous to each other, as is illustrated by Table \ref{tbl:mat_vs_vec_lmo}.
+lmo optimizers in Schatten $S_p$ norms and in $l_p$ norms with common $p$ may be analogous to each other, as is illustrated by \cref{tbl:mat_vs_vec_lmo}.
 The analogies go beyond similarities in the updates.
-SignSGD is very close to Adam, as noted in \cite{bernstein2024oldoptimizernewnorm}, and both Adam and Muon perform well in training large models.
+SignSGD is very close to Adam, as noted in \citet{bernstein2024oldoptimizernewnorm}, and both Adam and Muon perform well in training large models.
 NSGD is the same for both matrix and vector cases.
@@
 \section{Matrix side of updates}
-To compute the algorithms' updates, we use thick-restarted Lanczos method for singular value problem (TRLan) on $A^TA$ or $AA^T$ matrices (the one with less size is picked), implemented in cupy lSibrary [https://docs.cupy.dev/en/stable/reference/generated/cupyx.scipy.sparse.linalg.svds.html] and described in \cite{simonz1998thick}.
+To compute the algorithms' updates, we use thick-restarted Lanczos method for singular value problem (TRLan) on $\mM^{k\top}\mM^k$ or $\mM^k \mM^{k\top}$ matrices (the one with less size is picked), implemented in CuPy lSibrary \citep{cupy_svds_ref} and described in \citet{simonz1998thick}.
 
@@
 
-\subsection{Theory}
-        Not only formulas, but also citations!
-        Solutions to efficiently find some parts of SVD:
-        \begin{itemize}
-            \item Basic SVD (complexity asymptotic!)
-            \item Newton-Schulz (complexity asymptotic!)
-            \item Our job: Power iterations, Randomized SVD and Lanczos
-        \end{itemize}
-    \subsection{Experiments}
-        Nikolay's task: SVD, RSVD, Lanczos, Newton-Schulz. We use torch and cupy to test it.
-        Options: all possible on torch, or additionally to compare all on cupy. The goal: to find how much do we lose due to cupy
-
+In \cref{tbl:matrix_methods}, we compare performance of TRLan, RSVD, and power iterations on calculating k-rank update, which is used in k-Fanion.
+The results highlight that TrLan is much faster that its competitors.
+
+During the research it was noted that RSVD can give good and fast approximation for singular values, but the matrix of approximation is far from the one given by truncated SVD, while TRLan gives good and fast approximation of a matrix, but not so good approximation for singular values.
+That means that TRLan may be not a perfect choice for algorithms like Dion, where $\sigma_i$ are required for error feedback.
+\begin{table}[h!]
+    \centering
+    \begin{tabular}{c|c|c|c}
+    \hline
+    Method                              & rtol        & k & time,s \\
+    \hline 
+    Power Iterations                    & 0.01        & 1 & 7.7\\ 
+    SVDS (thick-restart Lanczos method) & 0.01        & 1 & 0.18\\
+    PCA Low Rank (RSVD)                 & 0.01        & 1 & 1.15\\
+    SVDS (thick-restart Lanczos method) & 0.01        & 10 & 0.47\\
+    PCA Low Rank (RSVD)                 & 0.01        & 10 & 19.4\\
+    SVDS (thick-restart Lanczos method) & 0.01        & 100 & 1.96\\
+    PCA Low Rank (RSVD)                 & 0.01        & 100 & 170\\
+    \end{tabular}
+    \caption{k-rank updated comparison}
+    Comparison of different numerical methods to calculate k-rank update on $5000\times5000$ matrix of real numbers, rtol is an error in Frobenius norm relative to the k-rank approximation. 
+    \label{tbl:matrix_methods}
+\end{table}
 \section{Trust Region Bounds for L-smooth functions}
     First, we analyze the problem in the unstochastic case.
-From Corollary 1 of \cite{kovalev2025understanding}, we directly get the following result that matches lower bounds, as was noted in that article.
+From Corollary 1 of \citet{kovalev2025understanding}, we directly get the following result that matches lower bounds, as was noted in that article.
 
@@
 
-    In the stochastic case, from Corollary of 2 of \cite{kovalev2025understanding}, we directly get the following result that pnce more matches lower bounds:
+    In the stochastic case, from Corollary of 2 of \citet{kovalev2025understanding}, we directly get the following result that once more matches lower bounds:
 
@@
 \section{Experiments}
+    \iffalse
     \subsection{Randomized Linear Least Squares}
-    Since the provided by other authors \cite{kovalev2025understanding,riabinin2025gluon} theoretical guarantees are almost norm-independent, we have to test them in practice.
-
-    To test the bounds from \cite{kovalev2025understanding} in practice, we construct the following L-smooth problem:
+    Since the provided by other authors \citep{kovalev2025understanding,riabinin2025gluon} theoretical guarantees are almost norm-independent, we have to test them in practice.
+
+    To test the bounds from \citet{kovalev2025understanding} in practice, we construct the following L-smooth problem:
 
@@
     \subsection{Logistic Regression}
-    \subsection{Benchmarks}
-    CNN benchmark
-    NanoGPT benchmark
+    \fi
+    \subsection{CIFAR-10 airbench}
+    We adapt Keller Jordan's code to test F-Muon, Neon, and F-Neon on the CIFAR-10 airbench \citep{cifar2023airbench}.
+    First, we run F-Muon for different $\alpha$ with the same {\tt lr=0.24(1 - step/total\_steps), momentum=0.6, nesterov=True}, as have been finetuned by Jordan.
+We record the accuracy after 8 epochs of training.
+
+    % add figure
+
+    Then, we tune F-Muon with $\alpha = 0.5$.
+Tuned parameters are {\tt lr=0.4(1 - step/total\_steps), momentum=0.65 - 0.1(1 - step/total\_steps), nesterov = True}.
+While Muon reaches $94.00 \pm 0.13$\% accuracy after 8 epochs, tuned F-Muon reaches $93.97 \pm 0.14\%$ after 8 epochs.
+
+    Finally, we take the set of tuned parameters and test on different $\alpha$. We notice that even when $\alpha=0.1$, the accuracy is much higher than in case of the pure NSGD.
+
+    The results are curious and could be represented by \cref{fig:cifar_ball}: lmo ball, which we plotted in a 2D space of singular values, has drastically changed, but the observed convergence after tuning has not degraded.
+These observations raise the question of how much lmo-based algorithms are sensitive to the constraint area, i.e.
+what will happen if the ball is corrupted.
+In this particular example, we have had a blurred ball, which proved as robust as the original ball.
+
+    \begin{figure}[h]
+        \centering
+        \includegraphics[width=0.6\textwidth]{figs/fstardual_cifar.pdf} % adjust width as needed
+        \caption{Visualization of lmo balls for Muon and F-Muon.}
+        \label{fig:cifar_ball}
+    \end{figure}
+
+    \subsection{Modded NanoGPT speedrun}
+    We test F-Muon on NanoGPT speedrun \citep{modded_nanogpt_2024}.
 
 \section{Related Work}
-As Muon \cite{jordan2024muon} is a very successful and popular optimizer for functions of weight matrices, a lot of reseach has been put into, first, further improving its performance, and, second, in explaining its success.
+As Muon \citep{jordan2024muon} is a very successful and popular optimizer for functions of weight matrices, a lot of reseach has been put into, first, further improving its performance, and, second, in explaining its success.
 
 {\bf Improvements of Muon.} Regarding the first point, in less than a year, a large number of applications and improvements of Muon has been proposed.
-\cite{liu2025muon} adapted the algorthm for training language models larger than NanoGPT.
-\cite{shah2025practical} organized efficient hyperparameter transfer by combining Muon with maximal update parametrization.
-To construct their COSMOS optimizer, \cite{chen2025cosmoshybridadaptive} applied computationally intensive updates of SOAP optimizer to a low-dimensional "leading eigensubspace" while using memory-efficient methods like Muon for the remaining parameters.
-\cite{amsel2025polar} proposed a more efficient alternative to Newton-Schulz operations.
-\cite{si2025adamuon} introduced AdaMuon which combines element-wise adaptivity with orthogonal updates.
+\citet{liu2025muon} adapted the algorthm for training language models larger than NanoGPT.
+\citet{shah2025practical} organized efficient hyperparameter transfer by combining Muon with maximal update parametrization.
+To construct their COSMOS optimizer, \citet{chen2025cosmoshybridadaptive} applied computationally intensive updates of SOAP optimizer to a low-dimensional "leading eigensubspace" while using memory-efficient methods like Muon for the remaining parameters.
+\citet{amsel2025polar} proposed a more efficient alternative to Newton-Schulz operations.
+\citet{si2025adamuon} introduced AdaMuon which combines element-wise adaptivity with orthogonal updates.
 We suppose that the described above or similar techniques can be applied to our optimizers as well.
@@
 
-{\bf Theory behind Muon.} Regarding the second point, there has been a prolonged gap in theory behind Muon, simplistic derivation of \cite{bernstein2025deriving} based on \cite{bernstein2024oldoptimizernewnorm} excluded.
+{\bf Theory behind Muon.} Regarding the second point, there has been a prolonged gap in theory behind Muon, simplistic derivation of \citet{bernstein2025deriving} based on \citet{bernstein2024oldoptimizernewnorm} excluded.
 This gap, as it seems to us, is not even now completely closed.
-For example, \cite{kovalev2025understanding} has provided convergence guarantees of Muon in various settings, from which, however, Muon's supremacy cannot be recovered.
+For example, \citet{kovalev2025understanding} has provided convergence guarantees of Muon in various settings, from which, however, Muon's supremacy cannot be recovered.
 Indeed, although the obtained bounds depend on the norm choice, the asymptotics of the convergence remain the same as for NSGD and other optimizers, $K = \cO(\epsilon^{-4})$ in a L-smooth stochastic case.
 
-Similar drawback has a recent article \cite{riabinin2025gluon}, where L-smoothness assumption is replaced with a more practical $(L_0, L_1)$-smoothness.
-The authors derived from their theorems optimal stepsizes for Muon and Scion that match fine-tuned stepsizes reported by \cite{pethick2025training}.
+Similar drawback has a recent article \citet{riabinin2025gluon}, where L-smoothness assumption is replaced with a more practical $(L_0, L_1)$-smoothness.
+The authors derived from their theorems optimal stepsizes for Muon and Scion that match fine-tuned stepsizes reported by \citet{pethick2025training}.
 But still, they did not explain why, for instance, NSGD is inferior to Muon in training large-language models.
@@
 {\bf The nuclear norm in lmo.} As we found out only when writing this article, the nuclear norm has been already explored in the context of the linear minimization oracle.
-\cite{pethick2025sam} applied it to create $\nu$SAM, a new sharpness-aware minimization technique.
-The update from their lemma 3.1 coincides with the update of Neon, but is used for completely different purposes.
-In addition, \cite{pethick2025sam} use power iterations to find $u_1$ and $v_1^\top$, while we suggest utilizing much more efficient and precise Lanczos algorithm.
-
-{\bf The Ky Fan Norm and Dion.} Rank-$k$ Centralized Dion, Algorithm 1 from \cite{ahn2025dioncommunicationefficientoptimizerlarge}, without an error feedback and scaling of the update, turns out to be an lmo-based algorithm under the $\normkfk{\cdot}^\dagger$ norm, which we described in {\it Different norms imply different updates}.
+\citet{pethick2025sam} applied it to create $\nu$SAM, a new sharpness-aware minimization technique.
+The update from their Lemma 3.1 coincides with the update of Neon, but is used for completely different purposes.
+In addition, \citet{pethick2025sam} use power iterations to find $u_1$ and $v_1^\top$, while we suggest utilizing much more efficient and precise Lanczos algorithm.
+
+{\bf The Ky Fan Norm and Dion.} Rank-$k$ Centralized Dion, Algorithm 1 from \citet{ahn2025dioncommunicationefficientoptimizerlarge}, without an error feedback and scaling of the update, turns out to be an lmo-based algorithm under the $\normkfk{\cdot}^\dagger$ norm, which we described in {\it Different norms imply different updates}.
 Reported by the authors necessity of using error feedback to obtain satisfactory accuracy may take place in the cases of our algorithms as well, for example, F-Neon.
@@
 
+\section{Author Contributions}
+IO suggested using the nuclear norm in the \citet{bernstein2024oldoptimizernewnorm} framework.
+DM supervised the project and helped with editing the article.
+IK, NK, and AV participated mainly on the first stage of research, when it has been a project in the optimization course at MIPT.
+IK suggested using composite norms (though not F2 and F*) and KKT conditions to find the resulting updates.
+NK suggested Lanczos algorithm as the fastest tool to compute Neon's updates and conducted experiments to prove it.
+AV tested Neon on the finetuning of NanoGPT.
+All other work was done by AK.
+
 \bibliography{icomp2024_conference}
@@
 
-\section{Appendix}
-\subsection{Formal assumptions}
-Warning: the text was copied, while formulas were adapted. We need to rephrase the subsection and probably contract it.
-
-{\bf Non-Euclidean norm setting and Lipschitz continuous gradient.}
-    In addition, we assume that the gradient $\nabla f(\cdot)$ is Lipschitz continuous with respect to the norm $\norm{\cdot}$, that is, the following inequality holds:
+\appendix
+
+\section{Formal assumptions}
+Here we restate assumptions from \citet{kovalev2025understanding} that are used in the theorems from the main part of the article.
+{\bf Lipschitz continuous gradient.}
+    We assume that the gradient $\nabla f(\cdot)$ is Lipschitz continuous with respect to the norm $\norm{\cdot}$, that is, the following inequality holds:
     \begin{aequation}\label{eq:L}
@@
 Frobenius, norm induced by the inner product $\<\cdot,\cdot>$, i.e., $\normf{\mX} = \sqrt{\<\mX,\mX>}=\sqrt{\tr(\mX^\top \mX)}$.
-These assumptions have been widely adopted for the analysis of many stochastic gradient optimization algorithms \citep{ghadimi2013stochastic,ghadimi2016accelerated,cutkosky2020momentum,sun2023momentum,horvath2023stochastic,gorbunov2020linearly}.
 
@@
     where $\rho > 0$ is a positive constant.
-Note that such a constant always exists due to the norm equivalence theorem, which always holds in the finite-dimensional space $\Rmn$.
-We recount $\rho$ for different norms $\norm{\cdot}$ in the appendix.
-
-\subsection{Norms $\normfstar{\cdot}^\dagger$ and $\normftwo{\cdot}^\dagger$}
-First, we need a well-known fact mentioned, for example, in \cite[Table 1]{yu2012arithmetic}. For the sake of completeness, we provide a proof of the fact.
-
-\begin{lemma}\label{lem:dual_to_conv_comb}
+
+\section{Norms $\normfstar{\cdot}^\dagger$ and $\normftwo{\cdot}^\dagger$}
+First, we need a well-known fact mentioned, for example, in \citet[Table 1]{yu2012arithmetic}. For the sake of completeness, we provide a proof of the fact.
+
+\begin{lemma}\label{lemma:dual_to_conv_comb}
     Let $\norm{\cdot}_{(1)}$ and $\norm{\cdot}_{(2)}$ be norms on a finite-dimensional Euclidean space, and let $\alpha,\beta \geq 0$. Define 
@@
 
-\subsection{Visualization of different matrix norms}
-\subsubsection{Duals to F* and F2 norms}
-
- It follows from \ref{lem:dual_to_conv_comb} that the norm ball in $\normfstar{\cdot}^\dagger$ is the Minkowski sum of the norm ball in $\alpha\normn{\cdot}$ and $(1-\alpha)\normf{\cdot}$ and the norm ball in $\normftwo{\cdot}^\dagger$ is the Minkowski sum of the norm ball in $\alpha\norms{\cdot}$ and $(1-\alpha)\normf{\cdot}$.
-
-In \ref{fig:fduals} we plot these norms. On x-axis and y-axis, there are singular values $\sigma_1$, $\sigma_2$ respectively of a matrix from $\Rmn$ with $\min\{m,n\}=2$.
+\section{Visualization of different matrix norms}
+\subsection{Duals to F* and F2 norms}
+
+ It follows from \cref{lemma:dual_to_conv_comb} that the norm ball in $\normfstar{\cdot}^\dagger$ is the Minkowski sum of the norm ball in $\alpha\normn{\cdot}$ and $(1-\alpha)\normf{\cdot}$ and the norm ball in $\normftwo{\cdot}^\dagger$ is the Minkowski sum of the norm ball in $\alpha\norms{\cdot}$ and $(1-\alpha)\normf{\cdot}$.
+
+In Fig. \cref{fig:fduals} we plot these norms. On x-axis and y-axis, there are singular values $\sigma_1$, $\sigma_2$ respectively of a matrix from $\Rmn$ with $\min\{m,n\}=2$.
 
@@
 
-\subsubsection{The Ky Fan norm and its dual}
+\subsection{The Ky Fan norm and its dual}
 1-balls in $l_\infty$, $l_1$ and $l_2$ norms are well-known from textbooks. But what about the Ky Fan $k$-norm? How can it be represented?
 
-To showcase the complex structure of the Ky Fan $k$-norm and its dual, we suggest the illustrations \ref{fig:kyfan_combined} with the ball in the Ky Fan $2$-norm in \ref{fig:kyfan} and its dual in \ref{fig:kyfandual}.
+To showcase the complex structure of the Ky Fan $k$-norm and its dual, we suggest the illustrations \cref{fig:kyfan_combined} with the ball in the Ky Fan $2$-norm in \cref{fig:kyfan} and its dual in \cref{fig:kyfandual}.
 On x-, y-, and z-axes, there are singular values $\sigma_1$, $\sigma_2$, and $\sigma_3$ respectively of a matrix from $\Rmn$ with $\min\{m,n\}=3$.
@@
 
-In fact, those balls can be described easier if we use the results from \cite{yu2012arithmetic}.
+In fact, those balls can be described easier if we use the results from \citet{yu2012arithmetic}.
 The Ky Fan $2$-norm ball is an intersection of three $l_1$ balls in $(x,y)$, $(x, z)$, and $(y,z)$ spaces.
@@
 
-\subsection{Updates derivations}
+\section{Updates derivations}
 
@@
 
-\subsection{Technical details of the experiments}
+\section{Technical details of the experiments}
+We compared TRLand with other methods in Google Colab. We used RTX A4000 for CIFAR airbench.
 
@@
 During the research it was noted that rsvd can give good and fast approximation for singular values, but the matrix of approximation is far from the one given by truncated svd, while Lanczos method gives good and fast approximation for a matrix, but not so good approximation for singular values.
-        \label{tab:matrix_methods}
+        \label{tab:matrix_methods_old}
     \end{table}

### 2025-08-24 conf_article/neon.tex
Added logs from NanoGPT speedrun

@@
     Then, we tune F-Muon with $\alpha = 0.5$.
-Tuned parameters are {\tt lr=0.4(1 - step/total\_steps), momentum=0.65 - 0.1(1 - step/total\_steps), nesterov = True}.
-While Muon reaches $94.00 \pm 0.13$\% accuracy after 8 epochs, tuned F-Muon reaches $93.97 \pm 0.14\%$ after 8 epochs.
+Tuned parameters are {\tt lr=0.4(1 - step/total\_steps), momentum=0.65, nesterov = True}.
+While Muon reaches $94.00 \pm 0.13$\% accuracy after 8 epochs, tuned F-Muon reaches $94.01 \pm 0.14\%$ after 8 epochs.
 
@@
     We test F-Muon on NanoGPT speedrun \citep{modded_nanogpt_2024}.
+For $\alpha = 0.5$, the optimal parameters are {\tt lr=0.07, momentum=0.95}, while for Muon they were  {\tt lr=0.05, momentum=0.95}.
+After testing for 1750 steps, as has been done on the speedruns, we get 3.281 cross-entropy loss, while on Muon, it falls below the target threshold 3.28 reaching 3.279.
+
+    TO-DO: more discussions of experiments.
 

### 2025-08-24 code/optimizers.py

@@ -496,3 +496,191 @@ class DistrMuonNormalized(torch.optim.Optimizer):
                 )
 
         torch.futures.collect_all(all_reduce_futures).wait()
+
+# does not work
+class FastNormMuon(torch.optim.Optimizer):
+    """
+    Muon - MomentUm Orthogonalized by Newton-schulz
+
+    https://kellerjordan.github.io/posts/muon/
+
+    Muon internally runs standard SGD-momentum, and then performs an orthogonalization post-
+    processing step, in which each 2D parameter's update is replaced with the nearest orthogonal
+    matrix. To efficiently orthogonalize each update, we use a Newton-Schulz iteration, which has
+    the advantage that it can be stably run in bfloat16 on the GPU.
+
+    Warning: This optimizer should not be used for the embedding layer, the final fully connected layer,
+    or any {0,1}-D parameters; those should all be optimized by a standard method (e.g., AdamW).
+    """
+    def __init__(self, params, lr=0.02, weight_decay=0.01, momentum=0.95, sgd_coeff=0):
+        defaults = dict(lr=lr, weight_decay=weight_decay, momentum=momentum, sgd_coeff=sgd_coeff)
+        params = list(params)
+        sizes = {p.shape for p in params}
+        # create one buffer per unique parameter-size
+        param_groups = []
+        for size in sizes:
+            group_params = [p for p in params if p.shape == size]
+            param_groups.append(dict(params=group_params))
+        super().__init__(param_groups, defaults)
+        self.sgd_coeff = sgd_coeff
+
+    @torch.no_grad()
+    def step(self):
+        # Efficient systems-wise implementation of step developed by @YouJiacheng,
+        # @KonstantinWilleke, @alexrgilbert, @adricarda, @tuttyfrutyee, @vdlad,
+        # @ryanyang0, and @vagrawal.
+        rank = dist.get_rank()
+        world_size = dist.get_world_size()
+        reduce_scatter_futures: list[torch.Future] = []
+        all_reduce_futures: list[torch.Future] = []
+        for group in self.param_groups:
+            params: list[Tensor] = group["params"]
+            grad = torch.empty_like(params[-1])
+            grad_pad = [param.grad for param in params] + [torch.zeros_like(params[-1])] * world_size
+            for base_i in range(0, len(params), world_size):
+                if base_i + rank < len(params):
+                    grad = params[base_i + rank].grad
+                # This gives strange dynamo warnings
[…]

### 2025-08-24 code/plot_results.py
Working on CIFAR plots

@@ -1,0 +1,107 @@
+"""
+Standalone script to plot SGD coefficient experiment results.
+Run this script after the experiment to visualize results from saved .pt files.
+"""
+
+import torch
+import matplotlib.pyplot as plt
+import numpy as np
+
+def plot_results(sgd_coeffs, mean_accs, std_accs):
+    """Create a scatter plot with error bars showing the results."""
+    plt.figure(figsize=(12, 8))
+    # Create scatter plot with error bars
+    ymin = 0.85
+    sgd_coeffs = np.array(sgd_coeffs)
+    mean_accs = np.array(mean_accs)
+    std_accs = np.array(std_accs)
+
+    # Create a mask for points above ymin
+    mask = mean_accs > ymin
+
+    # Plot only the filtered points
+    plt.errorbar(
+        1 - sgd_coeffs[mask],       # x values
+        mean_accs[mask],            # y values
+        yerr=std_accs[mask],        # errors
+        fmt='o', markersize=8, capsize=5, capthick=2,
+        color='#440154', ecolor='#440154', alpha=0.8, linewidth=2 # deep purple
+    )
+    # Customize the plot
+    plt.xlabel('α', fontsize=16)
+    plt.ylabel('Accuracy', fontsize=16)
+    plt.grid(True, alpha=0.3)
+    
+    # Set y-axis limits
+    plt.ylim(ymin, 0.96)
+    
+    # Add horizontal line at 94.01 (the original reported accuracy)
+    plt.axhline(y=0.9401, color='gray', linestyle='--', alpha=0.7, 
+                label='Top Muon: 94.01%')
+    plt.legend(fontsize=12)
+    
+    # Add value labels above points
+    for coeff, mean, std in zip(sgd_coeffs, mean_accs, std_accs):
+        plt.text(1 - coeff, max(ymin, mean + std), f'{mean:.4f}', 
+                ha='center', va='bottom', fontweight='bold', fontsize=10)
+    plt.xticks(np.arange(-0.2, 1.3, 0.2))
+    plt.tight_layout()
+    plt.savefig('sgd_coeff_results.pdf', dpi=300, bbox_inches='tight')
+    plt.show()
+
+def plot_results_from_file(filename='sgd_coeff_experiment_results.pt'):
+    """Load and plot results from a saved .pt file."""
+    try:
+        # Load the results
+        results = torch.load(filename)
+        
+        sgd_coeffs = results['sgd_coeffs']
[…]

### 2025-08-24 code/test_sgd_coeff.py

@@ -1,0 +1,36 @@
+"""
+Simple test script to verify the setup works before running the full experiment.
+dull code
+"""
+
+import torch
+from airbench_muon import CifarNet, NormalizedMuon
+
+def test_sgd_coeff():
+    """Test that we can create a model and optimizer with different sgd_coeff values."""
+    
+    print("Testing SGD coefficient setup...")
+    
+    # Create a simple model
+    model = CifarNet().cuda().to(memory_format=torch.channels_last)
+    
+    # Test different sgd_coeff values
+    test_coeffs = [-0.1, 0.0, 0.5, 1.0, 1.1]
+    
+    for coeff in test_coeffs:
+        try:
+            # Get filter parameters
+            filter_params = [p for p in model.parameters() if len(p.shape) == 4 and p.requires_grad]
+            
+            # Create optimizer with this coefficient
+            optimizer = NormalizedMuon(filter_params, lr=0.4, momentum=0.65, nesterov=True, sgd_coeff=coeff)
+            
+            print(f"  ✓ Successfully created NormalizedMuon with sgd_coeff = {coeff}")
+            
+        except Exception as e:
+            print(f"  ✗ Failed to create NormalizedMuon with sgd_coeff = {coeff}: {e}")
+    
+    print("\nTest completed!")
+
+if __name__ == "__main__":
+    test_sgd_coeff()

### 2025-08-24 code/airbench_muon.py
Working on CIFAR plots

@@ -21,7 +21,7 @@ from torch import nn
 import torch.nn.functional as F
 import torchvision
 import torchvision.transforms as T
-
+import math
 torch.backends.cudnn.benchmark = True
 
 #############################################
@@ -115,6 +115,38 @@ class NormalizedMuon(torch.optim.Optimizer):
                     update = (1-self.sgd_coeff) * update_part + self.sgd_coeff * g_normalized
                 else:
                     update = self.sgd_coeff * g_normalized
+                p.data.add_(update, alpha=-lr) # take a step
+
+class SignedMuon(torch.optim.Optimizer):
+    def __init__(self, params, lr=1e-3, momentum=0, nesterov=False, sgd_coeff=0):
+        defaults = dict(lr=lr, momentum=momentum, nesterov=nesterov)
+        super().__init__(params, defaults)
+        self.sgd_coeff = sgd_coeff
+
+    def step(self):
+        for group in self.param_groups:
+            lr = group["lr"]
+            momentum = group["momentum"]
+            for p in group["params"]:
+                g = p.grad
+                if g is None:
+                    continue
+                state = self.state[p]
+
+                if "momentum_buffer" not in state.keys():
+                    state["momentum_buffer"] = torch.zeros_like(g)
+                buf = state["momentum_buffer"]
+                buf.mul_(momentum).add_(g)
+                g = g.add(buf, alpha=momentum) if group["nesterov"] else buf
+
+                p.data.mul_(len(p.data)**0.5 / p.data.norm()) # normalize the weight
+                # update = zeropower_via_newtonschulz5(g.reshape(len(g), -1)).view(g.shape) # whiten the update
+                eps = 1e-12
+                if self.sgd_coeff != 1:
+                    update_part = zeropower_via_newtonschulz5(g.reshape(len(g), -1)).view(g.shape)
+                    update = (1-self.sgd_coeff) * update_part + self.sgd_coeff * torch.sign(g) / math.sqrt(g.shape[0] * g.shape[1])
+                else:
+                    update = self.sgd_coeff * torch.sign(g) / math.sqrt(g.shape[0] * g.shape[1])
                 p.data.add_(update, alpha=-lr) # take a step
 
 #############################################
@@ -392,7 +424,7 @@ def main(run, model):
                      dict(params=norm_biases,         lr=bias_lr, weight_decay=wd/bias_lr),
[…]

### 2025-08-24 code/quick_test.py

@@ -1,0 +1,92 @@
+"""
+Quick test script to run a small experiment with fewer runs.
+Useful for testing the setup before running the full experiment.
+"""
+
+import torch
+import matplotlib.pyplot as plt
+import numpy as np
+from run_sgd_coeff_experiment import create_modified_main_function
+from plot_results import plot_results, plot_results_from_file
+
+def quick_test():
+    """Run a quick test with fewer coefficient values and fewer runs."""
+    
+    print("Running Quick Test - SGD Coefficient Experiment")
+    print("=" * 50)
+    
+    # Test fewer coefficient values for quick testing
+    # sgd_coeffs = [-0.1, 0.0, 0.5, 1.0, 1.1]
+    sgd_coeffs = np.arange(-0.2, 1.3, 0.1)
+    print(f"Testing {len(sgd_coeffs)} different sgd_coeff values: {sgd_coeffs}")
+    print()
+    
+    # Store results
+    mean_accs = []
+    std_accs = []
+    
+    # Run experiments for each sgd_coeff value
+    for sgd_coeff in sgd_coeffs:
+        print(f"Running experiment with sgd_coeff = {sgd_coeff}")
+        
+        # Create and compile the model
+        from airbench_muon import CifarNet
+        model = CifarNet().cuda().to(memory_format=torch.channels_last)
+        model.compile(mode="max-autotune")
+        
+        # Get the modified main function
+        modified_main = create_modified_main_function(sgd_coeff)
+        
+        # Warmup run
+        print("  Running warmup...")
+        modified_main("warmup", model)
+        
+        # Actual runs (fewer for quick testing)
+        num_runs = 5
+        print(f"  Running {num_runs} experiments...")
+        accs = torch.tensor([modified_main(run, model) for run in range(num_runs)])
+        
+        mean_acc = accs.mean().item()
+        std_acc = accs.std().item()
+        
+        print(f"  Results: Mean = {mean_acc:.4f}, Std = {std_acc:.4f}")
+        print()
+        
+        mean_accs.append(mean_acc)
+        std_accs.append(std_acc)
+    
+    # Plot the results
+    print("Creating visualization...")
+    plot_results(sgd_coeffs, mean_accs, std_accs)
+    
+    # Print summary
+    print("\n" + "=" * 50)
+    print("QUICK TEST SUMMARY")
+    print("=" * 50)
[…]

### 2025-08-24 code/run_sgd_coeff_experiment.py

@@ -1,0 +1,298 @@
+"""
+Script to run airbench_muon.py with different sgd_coeff values and plot the results.
+"""
+
+import os
+import sys
+import torch
+import matplotlib.pyplot as plt
+import numpy as np
+from math import ceil
+from airbench_muon import CifarNet, CifarLoader, evaluate, BatchNorm, Conv, ConvGroup
+
+def create_modified_main_function(sgd_coeff):
+    """
+    Create a modified version of the main function with the specified sgd_coeff.
+    """
+    def modified_main(run, model):
+        batch_size = 2000
+        bias_lr = 0.053
+        head_lr = 0.67
+        wd = 2e-6 * batch_size
+
+        test_loader = CifarLoader("cifar10", train=False, batch_size=2000)
+        train_loader = CifarLoader("cifar10", train=True, batch_size=batch_size, aug=dict(flip=True, translate=2))
+        if run == "warmup":
+            # The only purpose of the first run is to warmup the compiled model, so we can use dummy data
+            train_loader.labels = torch.randint(0, 10, size=(len(train_loader.labels),), device=train_loader.labels.device)
+        total_train_steps = torch.ceil(torch.tensor(8 * len(train_loader)))
+        whiten_bias_train_steps = torch.ceil(torch.tensor(3 * len(train_loader)))
+
+        # Create optimizers and learning rate schedulers
+        filter_params = [p for p in model.parameters() if len(p.shape) == 4 and p.requires_grad]
+        norm_biases = [p for n, p in model.named_parameters() if "norm" in n and p.requires_grad]
+        param_configs = [dict(params=[model.whiten.bias], lr=bias_lr, weight_decay=wd/bias_lr),
+                         dict(params=norm_biases,         lr=bias_lr, weight_decay=wd/bias_lr),
+                         dict(params=[model.head.weight], lr=head_lr, weight_decay=wd/head_lr)]
+        optimizer1 = torch.optim.SGD(param_configs, momentum=0.85, nesterov=True, fused=True)
+        
+        # Import NormalizedMuon here to avoid circular imports
+        from airbench_muon import NormalizedMuon
+        optimizer2 = NormalizedMuon(filter_params, lr=0.4, momentum=0.65, nesterov=True, sgd_coeff=sgd_coeff)
+        optimizers = [optimizer1, optimizer2]
+        for opt in optimizers:
+            for group in opt.param_groups:
+                group["initial_lr"] = group["lr"]
[…]

### 2025-08-24 code/plot_results.py
F-muon alphas compared

@@ -24,12 +24,12 @@ def plot_results(sgd_coeffs, mean_accs, std_accs):
         1 - sgd_coeffs[mask],       # x values
         mean_accs[mask],            # y values
         yerr=std_accs[mask],        # errors
-        fmt='o', markersize=8, capsize=5, capthick=2,
+        fmt='o', markersize=6, capsize=5, capthick=2,
         color='#440154', ecolor='#440154', alpha=0.8, linewidth=2 # deep purple
     )
     # Customize the plot
-    plt.xlabel('α', fontsize=16)
-    plt.ylabel('Accuracy', fontsize=16)
+    plt.xlabel('α', fontsize=18)
+    plt.ylabel('Accuracy, %', fontsize=18)
     plt.grid(True, alpha=0.3)
     
     # Set y-axis limits
@@ -38,12 +38,13 @@ def plot_results(sgd_coeffs, mean_accs, std_accs):
     # Add horizontal line at 94.01 (the original reported accuracy)
     plt.axhline(y=0.9401, color='gray', linestyle='--', alpha=0.7, 
                 label='Top Muon: 94.01%')
-    plt.legend(fontsize=12)
+    plt.legend(fontsize=16)
     
     # Add value labels above points
-    for coeff, mean, std in zip(sgd_coeffs, mean_accs, std_accs):
-        plt.text(1 - coeff, max(ymin, mean + std), f'{mean:.4f}', 
-                ha='center', va='bottom', fontweight='bold', fontsize=10)
+    
+    #for coeff, mean, std in zip(sgd_coeffs, mean_accs, std_accs):
+    #    plt.text(1 - coeff, max(ymin, mean + std), f'{100 * mean:.2f}', 
+    #            ha='center', va='bottom', fontweight='bold', fontsize=15)
     plt.xticks(np.arange(-0.2, 1.3, 0.2))
     plt.tight_layout()
     plt.savefig('sgd_coeff_results.pdf', dpi=300, bbox_inches='tight')
@@ -52,8 +53,8 @@ def plot_results(sgd_coeffs, mean_accs, std_accs):
 def plot_results_from_file(filename='sgd_coeff_experiment_results.pt'):
     """Load and plot results from a saved .pt file."""
     try:
-        # Load the results
-        results = torch.load(filename)
+        # Load the results with weights_only=False to handle numpy arrays
+        results = torch.load(filename, weights_only=False)
         
         sgd_coeffs = results['sgd_coeffs']
         mean_accs = results['mean_accs']
[…]

### 2025-08-24 code/quick_test.py

@@ -85,8 +85,15 @@ def quick_test():
         'note': 'Quick test with only 5 runs per coefficient'
     }
     
-    torch.save(results, 'quick_test_results.pt')
+    # Save in a more compatible format
+    torch.save(results, 'quick_test_results.pt', _use_new_zipfile_serialization=False)
     print(f"\nQuick test results saved to 'quick_test_results.pt'")
+    
+    # Also save as a pickle file for better compatibility
+    import pickle
+    with open('quick_test_results.pkl', 'wb') as f:
+        pickle.dump(results, f)
+    print(f"Results also saved to 'quick_test_results.pkl'")
 
 if __name__ == "__main__":
     quick_test()

### 2025-08-24 code/run_sgd_coeff_experiment.py
F-muon alphas compared

@@ -38,7 +38,7 @@ def create_modified_main_function(sgd_coeff):
         
         # Import NormalizedMuon here to avoid circular imports
         from airbench_muon import NormalizedMuon
-        optimizer2 = NormalizedMuon(filter_params, lr=0.4, momentum=0.65, nesterov=True, sgd_coeff=sgd_coeff)
+        optimizer2 = NormalizedMuon(filter_params, lr=0.24, momentum=0.6, nesterov=True, sgd_coeff=sgd_coeff)
         optimizers = [optimizer1, optimizer2]
         for opt in optimizers:
             for group in opt.param_groups:
@@ -291,8 +291,15 @@ def main_experiment():
         'best_std': best_std
     }
     
-    torch.save(results, 'full_sgd_coeff_experiment_results.pt')
+    # Save in a more compatible format
+    torch.save(results, 'full_sgd_coeff_experiment_results.pt', _use_new_zipfile_serialization=False)
     print(f"\nResults saved to 'full_sgd_coeff_experiment_results.pt'")
+    
+    # Also save as a pickle file for better compatibility
+    import pickle
+    with open('full_sgd_coeff_experiment_results.pkl', 'wb') as f:
+        pickle.dump(results, f)
+    print(f"Results also saved to 'full_sgd_coeff_experiment_results.pkl'")
 
 if __name__ == "__main__":
     main_experiment()

### 2025-08-24 code/quick_test.py
Improved plot

@@ -17,7 +17,7 @@ def quick_test():
     
     # Test fewer coefficient values for quick testing
     # sgd_coeffs = [-0.1, 0.0, 0.5, 1.0, 1.1]
-    sgd_coeffs = np.arange(-0.2, 1.3, 0.1)
+    sgd_coeffs = np.arange(-0.5, 1.3, 0.1)
     print(f"Testing {len(sgd_coeffs)} different sgd_coeff values: {sgd_coeffs}")
     print()
     
@@ -42,7 +42,7 @@ def quick_test():
         modified_main("warmup", model)
         
         # Actual runs (fewer for quick testing)
-        num_runs = 5
+        num_runs = 10
         print(f"  Running {num_runs} experiments...")
         accs = torch.tensor([modified_main(run, model) for run in range(num_runs)])
         

### 2025-08-24 code/plot_results.py
Exps on CIFAR done

@@ -45,7 +45,7 @@ def plot_results(sgd_coeffs, mean_accs, std_accs):
     #for coeff, mean, std in zip(sgd_coeffs, mean_accs, std_accs):
     #    plt.text(1 - coeff, max(ymin, mean + std), f'{100 * mean:.2f}', 
     #            ha='center', va='bottom', fontweight='bold', fontsize=15)
-    plt.xticks(np.arange(-0.2, 1.3, 0.2))
+    plt.xticks(np.arange(0, 1.7, 0.2))
     plt.tight_layout()
     plt.savefig('sgd_coeff_results.pdf', dpi=300, bbox_inches='tight')
     plt.show()
@@ -140,7 +140,7 @@ if __name__ == "__main__":
     
     if available_files:
         # Try to plot the main results file first
-        if 'sgd_coeff_experiment_results.pt' in available_files:
+        if 'muon_alphas.pt' in available_files:
             print("\nPlotting main experiment results...")
             plot_results_from_file('sgd_coeff_experiment_results.pt')
         elif 'quick_test_results.pt' in available_files:

### 2025-08-24 code/optimizers.py
Exps on quadratics

@@ -25,7 +25,7 @@ def zeropower_via_newtonschulz5(G, steps=3, eps=1e-7):
         X = X.T
     return X
 '''
-@torch.compile
+# @torch.compile
 def zeropower_via_newtonschulz5(G: Tensor, steps: int=5) -> Tensor:
     """
     Newton-Schulz iteration to compute the zeroth power / orthogonalization of G. We opt to use a

### 2025-08-24 code/run_benchmarks_simpler.py

@@ -49,14 +49,14 @@ def main() -> None:
     lr_nuc = grad_op_eps / beta_nuc / lipsch_const_nuc
     lr_fro = grad_op_eps / beta_fro / lipsch_const_fro
     print(f"Muon lr: {lr_op: .2E}, Neon iter: {lr_nuc: .2e}, Frobenius iter: {lr_fro:.2e}")
+    lr_op = lr_nuc = lr_fro = 0.01
 
     iter_num_op = int(lipsch_const_op * delta / grad_op_eps / grad_op_eps / beta_op / beta_op)
     iter_num_nuc = int(lipsch_const_nuc * delta / grad_op_eps / grad_op_eps / beta_nuc / beta_nuc)
     iter_num_fro = int(lipsch_const_fro * delta / grad_op_eps / grad_op_eps / beta_fro / beta_fro)
 
-    print(lr_op, lr_nuc, lr_fro)
     print(f"Muon iter: {iter_num_op}, Neon iter: {iter_num_nuc}, Frobenius iter: {iter_num_fro}")
-    iter_num_nuc = iter_num_op = iter_num_fro = 10000
+    iter_num_nuc = iter_num_op = iter_num_fro = 5000
     # iter_num = 100000
     X0 = 0.1 * torch.randn(m, n, dtype=torch.float32, device=device)
     # Define optimizers and settings
@@ -143,7 +143,7 @@ def main() -> None:
         verbose=True,
         name=sgd_name
     )
-    experiments[sgd_name] = sgd_results
+    # experiments[sgd_name] = sgd_results
     # experiments[sgd_name] = sgd_results
     # Build panels and plot
     panels = build_default_panels(experiments)

### 2025-08-24 code/benchmark_plotter.py
Exps on quadratics

@@ -1,4 +1,5 @@
 import math
+import os
 from typing import Any, Dict, List, Optional, Tuple
 
 import numpy as np
@@ -54,12 +55,10 @@ def plot_from_descriptions(
         ax.set_xlabel(panel.get("x_label", ""))
         ax.set_ylabel(panel.get("y_label", ""))
         base_title = panel.get("title", "")
-        ax.set_title(base_title + (f" ({title_suffix})" if title_suffix else ""))
+        # ax.set_title(base_title + (f" ({title_suffix})" if title_suffix else ""))
         yscale = panel.get("yscale")
         if yscale:
             ax.set_yscale(yscale)
-            #if yscale == "log":
-            #    ax.set_ylim(bottom=10**deg)
         ax.grid(True, alpha=0.3)
         ax.legend()
 
@@ -69,8 +68,80 @@ def plot_from_descriptions(
 
     fig.tight_layout()
     if save_path is not None:
-        plt.savefig(save_path, dpi=300, bbox_inches="tight")
+        plt.savefig(save_path, dpi=300, bbox_inches="tight", format='pdf')
     plt.show()
+
+
+def plot_and_save_individual_panels(
+    panel_descriptions: List[Dict[str, Any]],
+    base_save_dir: str = "simple_lls",
+    title_suffix: str = "",
+    figsize: Tuple[int, int] = (10, 8),
+) -> None:
+    """
+    Plot and save each panel individually as a separate PDF file in the specified directory.
+    
+    Args:
+        panel_descriptions: List of panel descriptions (same format as plot_from_descriptions)
+        base_save_dir: Directory to save the PDF files (default: "simple_lls")
+        title_suffix: Optional suffix to add to plot titles
+        figsize: Figure size for individual plots
+    """
+    
+    # Create the save directory if it doesn't exist
+    os.makedirs(base_save_dir, exist_ok=True)
+    
+    for idx, panel in enumerate(panel_descriptions):
+        # Create a new figure for each panel
+        fig, ax = plt.subplots(figsize=figsize)
+        
+        yscale = panel.get("yscale")
+        deg = -3
+        
+        # Plot all series for this panel
+        for s in panel.get("series", []):
+            y_values = s["y"]
+            if yscale == "log":
+                y_array = np.asarray(y_values, dtype=float)
+                y_values = np.maximum(y_array, 10**deg)
+            
+            ax.plot(
+                s["x"],
+                y_values,
+                label=s.get("name", "series"),
[…]

### 2025-08-24 conf_article/neon.tex
Article has been done

@@
 Then, we consider their convex combinations with the Frobenius norm and corresponding duals to develop a new family of algorithms we name {\it F-Fanions}, one of them being {\it F-Muon}.
-We complement our theoretical analysis with an extensive empirical study of the algorithms across a wide range of tasks and settings, from which it follows that properly-tuned F-Muon is on par with Muon, which raises the question about Muon's optimality.
+We complement our theoretical analysis with an extensive empirical study of the algorithms across a wide range of tasks and settings, from which it follows that properly-tuned F-Muon is almost on par with Muon, which questions the necessity of finding the exact solution to the problem in linear optimization oracle framework.
 
@@
 Hence, new scalable, time- and memory-efficient algorithms have been demanded.
-Besides well-known Adam \citep{kingma2014adam} and AdamW \citep{loshchilov2017decoupled}, recently proposed Muon has shown promising results on training very large models \citep{jordan2024muon,liu2025muon}.
+Besides well-known Adam \citep{kingma2014adam} and AdamW \citep{loshchilov2017decoupled}, recently proposed Muon \citep{jordan2024muon} has shown promising results on training very large models \citep{liu2025muon}.
 Its key difference from Adam and AdamW is that it has been constructed specifically for optimizing functions of weight matrices, which are common in deep learning.
@@
 
- In this article, we focus on the two most common matrix norms akin to the spectral, namely, the nuclear Norm and the Frobenius norm.
+ In this article, we focus on the two most common matrix norms akin to the spectral, namely, the nuclear norm and the Frobenius norm.
 Working in the linear minimization oracle (lmo) framework, which is equivalent to a factor to the trust region approach and the steepest descent under norm constraint, we derive Neon, our algorithm based on the nuclear norm.
-In the section {\it Matrix side of the updates}, we explain how Neon updates can be computed asymptotically faster than Muon updates by the Newton-Schulz iterations.
-
- Noticing that Neon and Muon are diametrical in terms of the rank of the update matrix, we bridge the space by "regularizing" them by NormalizedSGD, which is derived in lmo with the Frobenius norm.
-We do this in the same lmo approach by considering a norm that is dual to the convex combination of the Frobenius norm and the spectral or the nuclear norms respectively.
+In the section {\it Matrix side of the updates}, we explain how Neon updates can be computed asymptotically with the same $\cO(m n \min\{m, n\})$ number of FLOPs as Muon updates are computed via Newton-Schulz iterations.
+
+ Noticing that Neon and Muon are diametrical in terms of the rank of the update matrix, we bridge the space by "regularizing" them by Normalized SGD, which is derived in lmo with the Frobenius norm.
+We do this in the same lmo approach by considering a norm that is dual to the convex combination of the Frobenius norm and the spectral or the nuclear norm respectively.
 So we derive the algorithms we name F-Neon and F-Muon respectively.
@@
 The efforts results in comparison of the algorithms by their convergence not in terms of the dual norms of their gradient, but in terms of the common spectral norm, which may strongly differ, especially in large matrices, from the initial dual norms.
-Thus, we compare the algorithms in a unified fashion.
-
- Finally, we test Muon, Neon, NSGD, F-Muon and F-Neon on deep-learning tasks: training convolutional network on CIFAR-10 and fine-tuning NanoGPT.
+Thus, we compare the algorithms in a unified fashion and find that if an algorithm converges fast by the loss, it does not mean it converges fast by the norm of the gradient, and vice versa.
+
+ Finally, we test Muon and F-Muon on two deep-learning tasks, training convolutional network on CIFAR-10 \citep{cifar2023airbench} and pre-training NanoGPT on FineWeb \citep{modded_nanogpt_2024}.
 The results support the supremacy of Muon, but the most striking result of the tests is that F-Muon, only half of which is Muon, surpasses Muon's accuracy on the CIFAR tasks by a margin.
@@
 
-Per-cycle complexity is $O(mn^2 + n^2k + nk^2)$, where m and n are dimensions of target matrix and n is the smallest one, k is retained subspace's size.
+Per-cycle complexity is $O(mn^2 + n^2k + nk^2)$, where m and n are dimensions of the target matrix and n is the smaller one, k is retained subspace's size.
 
@@
 That means that TRLan may be not a perfect choice for algorithms like Dion, where $\sigma_i$ are required for error feedback.
+
+The practical drawback of the TRLAn is the absence of its implementation in PyTorch.
 \begin{table}[h!]
@@
 \section{Experiments}
-    \iffalse
     \subsection{Randomized Linear Least Squares}
@@
 
-    To test the bounds from \citet{kovalev2025understanding} in practice, we construct the following L-smooth problem:
+    To test the bounds from \citet{kovalev2025understanding}, we construct the following L-smooth problem:
 
@@
 
-    We are interested to get $\epsilon_2=$ 1e-1 precision in $\norms{\cdot}$.
+    We are interested to get $\epsilon_2=$ 1e0 precision in $\norms{\cdot}$.
 Hence, we target $\epsilon = \epsilon_2$ for the nuclear norm (Neon), $\epsilon = \frac{n}{2}$ for the spectral norm (Muon), and $\epsilon = \frac{\sqrt{n}}{2}$ for the Frobenius norm (NSGD).
-
-    {\bf An independent experiment}
+    With that, we get the following lrs: Muon lr: 2.33E-04, Neon lr: 2.10e-02, Frobenius (NSGD) lr: 1.75e-02, and the following number of iterations: Muon iter: 68, Neon iter: 11919, Frobenius iter: 323.
+L-constants are 171.4, 47.7, and 16.2 respectively.
+As we will see, these results do not relate well to the practice.
+
     To make the problem more real-world, we do not use known information about smoothness.
-We run NormalizedSGD, Muon, F-Muon with $\alpha=1/2$, Neon, and F-Neon with $\alpha=1/2$ for 100 000 iterations with learning rate = 1e-3.
-
-    \subsection{Logistic Regression}
-    \fi
+We run NormalizedSGD, Muon, F-Muon with $\alpha=1/2$, Neon, and F-Neon with $\alpha=1/2$ for 5 000 iterations with learning rate = 1e-2, which has been close to computed theoretically above.
+The results are in \cref{fig:lls}.
+
+    Though Neon displays the worst convergence in terms of loss, it is a leader in terms of the norm of the gradient.
+This might signify that the known bounds are not sufficient even in the case of a convex optimization: lmo-based algorithms may converge well by these bounds, but be impractical like Neon.
+
+    This experiment may also show that F-Muon may be more robust under general conditions than Muon.
+
+    \begin{figure}[h!]
+        \centering
+        \begin{subfigure}[t]{0.48\linewidth}
+            \centering
+            \includegraphics[width=\linewidth]{figs/simple_lls/loss_vs_iteration_50x50.pdf}
+            \caption{Loss}
+            \label{fig:lls_loss}
+        \end{subfigure}
+        \hfill
+        \begin{subfigure}[t]{0.48\linewidth}
+            \centering
+            \includegraphics[width=\linewidth]{figs/simple_lls/gradient_spectral_norm_vs_iteration_50x50.pdf}
+            \caption{Spectral norm of the gradient}
+            \label{fig:lls_grad_norm}
+        \end{subfigure}
+        \caption{Linear least squares problem for a 50 x 50 matrix}
+        \label{fig:lls}
+
+    \end{figure}
+
     \subsection{CIFAR-10 airbench}
     We adapt Keller Jordan's code to test F-Muon, Neon, and F-Neon on the CIFAR-10 airbench \citep{cifar2023airbench}.
-    First, we run F-Muon for different $\alpha$ with the same {\tt lr=0.24(1 - step/total\_steps), momentum=0.6, nesterov=True}, as have been finetuned by Jordan.
+    First, we run F-Muon for different $\alpha$ with the same {\tt lr=0.24(1 - step/total\_steps), momentum=0.6, nesterov=True}, as have been finetuned by Jordan, 10 repetitions for each $\alpha$.
 We record the accuracy after 8 epochs of training.
-
-    % add figure
-
-    Then, we tune F-Muon with $\alpha = 0.5$.
+The results are in \cref{fig:muon_alphas}.
+
+    Then we tune F-Muon with $\alpha = 0.5$.
 Tuned parameters are {\tt lr=0.4(1 - step/total\_steps), momentum=0.65, nesterov = True}.
-While Muon reaches $94.00 \pm 0.13$\% accuracy after 8 epochs, tuned F-Muon reaches $94.01 \pm 0.14\%$ after 8 epochs.
-
-    Finally, we take the set of tuned parameters and test on different $\alpha$. We notice that even when $\alpha=0.1$, the accuracy is much higher than in case of the pure NSGD.
+While Muon reaches $94.00 \pm 0.13$\% accuracy after 8 epochs, tuned F-Muon reaches $94.01 \pm 0.14\%$ after 8 epochs (average was done by 200 iterations).
+
+    Finally, we take this set of tuned parameters and test on different $\alpha$, 5 times for each $\alpha$.
+The results are in \cref{fig:fmuon_alphas}.
+We notice that even when $\alpha=0.1$, the accuracy is much higher than in case of the pure NSGD.
+
+    \begin{figure}[h!]
+        \centering
+        \begin{subfigure}[b]{0.48\linewidth}
+            \centering
+            \includegraphics[width=\linewidth]{figs/muon_tuned_diff_alpha.pdf}
+            \caption{With parameters tuned for Muon}
+            \label{fig:muon_alphas}
+        \end{subfigure}
+        \hfill
+        \begin{subfigure}[b]{0.48\linewidth}
+            \centering
+            \includegraphics[width=\linewidth]{figs/fmuon_tuned_diff_alpha.pdf}
+            \caption{With parameters tuned for F-Muon}
+            \label{fig:fmuon_alphas}
+        \end{subfigure}
+        \caption{Mean accuracies for different $\alpha$ of F-Muon.}
+        \label{fig:diff_alphas}
+    \end{figure}
 
@@
 
+    The most pathological case is $\alpha < 0$, which corresponds to the lmo with an area that is not a norm ball!
+Despite this violation, the mixture of algorithms reaches almost the same accuracy as vanilla Muon.
+
     \subsection{Modded NanoGPT speedrun}
@@
 For $\alpha = 0.5$, the optimal parameters are {\tt lr=0.07, momentum=0.95}, while for Muon they were  {\tt lr=0.05, momentum=0.95}.
-After testing for 1750 steps, as has been done on the speedruns, we get 3.281 cross-entropy loss, while on Muon, it falls below the target threshold 3.28 reaching 3.279.
-
-    TO-DO: more discussions of experiments.
+After testing for 1750 steps, as it has been done on the speedruns, we get 3.281 cross-entropy loss, while on Muon, it falls below the target threshold 3.28 reaching 3.279.
+
+    Again, as on the CIFAR airbench, if we set $\alpha=-0.1$, F* to which is not a norm, we get a 3.2818 loss, which is not higher than for $\alpha=0.3$, for example.
 
@@
 Here we restate assumptions from \citet{kovalev2025understanding} that are used in the theorems from the main part of the article.
+
 {\bf Lipschitz continuous gradient.}

### 2025-08-24 code/norms/fmuon_negative_nsgd.py
Article has been done

@@ -1,0 +1,60 @@
+import numpy as np
+import matplotlib.cm as cm
+import matplotlib.pyplot as plt
+
+def minkowski_sum_boundary(alpha, num_angles=2000):
+    r1 = alpha            # L1-ball radius (diamond)
+    r2 = 1.0 - alpha      # L2-ball radius (circle)
+    thetas = np.linspace(0, 2*np.pi, num_angles, endpoint=False)
+    pts = []
+    for th in thetas:
+        u = np.array([np.cos(th), np.sin(th)])  # direction
+        
+        # Support point on r2 * L2 ball is simply r2 * u
+        s2 = r2 * u
+        
+        # Support point on r1 * L1 ball:
+        # pick the axis with larger |component|, place full mass there with the sign
+        if abs(u[0]) >= abs(u[1]):
+            s1 = np.array([r1 * np.sign(u[0]), 0.0])
+        else:
+            s1 = np.array([0.0, r1 * np.sign(u[1])])
+        
+        pts.append(s1 + s2)
+    return np.array(pts)
+
+def l1_ball_boundary(radius, num=800):
+    t = np.linspace(0, 2*np.pi, num, endpoint=False)
+    x = np.cos(t)
+    y = np.sin(t)
+    denom = np.maximum(np.abs(x) + np.abs(y), 1e-12)
+    return radius * x / denom, radius * y / denom
+
+# List of alpha values to test
+alphas = [1.5]
+colormap = cm.get_cmap('viridis')
+colors = [colormap(i / (len(alphas))) for i in range(len(alphas) - 1, -1, -1)]
+# Plot
+plt.figure(figsize=(8, 8))
+
+# Plot Minkowski sums for each alpha
+for alpha, color in zip(alphas, colors):
+    sum_pts = 0.24 * minkowski_sum_boundary(alpha)
+    plt.plot(sum_pts[:, 0], sum_pts[:, 1], label=f"F-Muon ball (α={alpha}, lr=0.4)", 
+             linewidth=2, color=color)
+
+# Plot reference L1 and L2 balls
+x1, y1 = l1_ball_boundary(0.24)
+plt.plot(x1, y1, linestyle="--", label="Muon ball (lr=0.24)", color="black", alpha=0.9)
+
+# Configure plot
+plt.gca().set_aspect("equal", adjustable="box")
[…]

### 2025-08-25 code/norms/fmuon_negative_nsgd.py
Cleared in the code/norm folder

@@ -1,3 +1,5 @@
+# we are not sure that it is done correct here in terms of lmo: we do not understand such a Minkowski sum
+# and it does not work so, because duals work the way we wrote only for norms 
 import numpy as np
 import matplotlib.cm as cm
 import matplotlib.pyplot as plt

### 2025-08-25 code/norms/fstar_dual_plotter.py
Cleared in the code/norm folder

@@ -1,3 +1,4 @@
+# for different alphas lmos for F-Muon
 import numpy as np
 import matplotlib.cm as cm
 import matplotlib.pyplot as plt

### 2025-08-25 code/norms/ftwo_dual_plotter.py

@@ -1,3 +1,4 @@
+# different alphas lmos for F-Neon
 import numpy as np
 import matplotlib.pyplot as plt
 import matplotlib.cm as cm

### 2025-08-25 code/norms/kyfan2_dual_plotter.py

@@ -1,3 +1,4 @@
+# plots lmo norm for a 2-Fanion (=2-Dion without an error feedback)
 import numpy as np
 import matplotlib.pyplot as plt
 from mpl_toolkits.mplot3d.art3d import Poly3DCollection

### 2025-08-25 code/norms/kyfan2_plotter.py

@@ -1,3 +1,4 @@
+# plots Ky Fan rank-2 norm
 import numpy as np
 import matplotlib.pyplot as plt
 from mpl_toolkits.mplot3d.art3d import Poly3DCollection

### 2025-08-25 code/norms/muon_fmuon_cifar.py

@@ -1,3 +1,4 @@
+# the code plots F-Muon and Muon optimal lmo balls for CIFAR airbench
 import numpy as np
 import matplotlib.cm as cm
 import matplotlib.pyplot as plt
@@ -39,12 +40,12 @@ plt.figure(figsize=(8, 8))
 
 # Plot Minkowski sums for each alpha
 for alpha, color in zip(alphas, colors):
-    sum_pts = 0.4 * minkowski_sum_boundary(alpha)
+    sum_pts = 0.4 * minkowski_sum_boundary(alpha) # 0.4 is for F-Muon lr, which is \eta in lmo
     plt.plot(sum_pts[:, 0], sum_pts[:, 1], label=f"F-Muon ball (α={alpha}, lr=0.4)", 
              linewidth=2, color=color)
 
 # Plot reference L1 and L2 balls
-x1, y1 = l1_ball_boundary(0.24)
+x1, y1 = l1_ball_boundary(0.24) # 0.24 is for Muon lr
 plt.plot(x1, y1, linestyle="--", label="Muon ball (lr=0.24)", color="black", alpha=0.9)
 
 # Configure plot

### 2025-08-25 code/scrap/scrap_norms/kyfan_plotter.py

@@ -1,3 +1,4 @@
+# rubbish from Cursor
 import numpy as np
 import matplotlib.pyplot as plt
 from mpl_toolkits.mplot3d import Axes3D

### 2025-08-25 code/cifar_bench/quick_test.py
Clearing the code: CIFAR and others

@@ -6,7 +6,7 @@ Useful for testing the setup before running the full experiment.
 import torch
 import matplotlib.pyplot as plt
 import numpy as np
-from run_sgd_coeff_experiment import create_modified_main_function
+from create_fmuon_testing import create_modified_main_function
 from plot_results import plot_results, plot_results_from_file
 
 def quick_test():
@@ -30,7 +30,7 @@ def quick_test():
         print(f"Running experiment with sgd_coeff = {sgd_coeff}")
         
         # Create and compile the model
-        from airbench_muon import CifarNet
+        from code.cifar_bench.airbench_muon import CifarNet
         model = CifarNet().cuda().to(memory_format=torch.channels_last)
         model.compile(mode="max-autotune")
         

### 2025-08-25 code/cifar_loader.py

@@ -1,87 +1,0 @@
-#############################################
-#                DataLoader                 #
-#############################################
-import torch
-import torchvision
-import torch.nn.functional as F
-import torchvision.transforms as T
-from config import device
-
-from math import ceil
-import os
-
-CIFAR_MEAN = torch.tensor((0.4914, 0.4822, 0.4465), device=device)
-CIFAR_STD = torch.tensor((0.2470, 0.2435, 0.2616), device=device)
-
-#@torch.compile(mode='max-autotune')
-def batch_flip_lr(inputs):
-    flip_mask = (torch.rand(len(inputs), device=inputs.device) < 0.5).view(-1, 1, 1, 1)
-    return torch.where(flip_mask, inputs.flip(-1), inputs)
-
-#@torch.compile(mode='max-autotune')
-def batch_crop(images, crop_size):
-    r = (images.size(-1) - crop_size)//2
-    shifts = torch.randint(-r, r+1, size=(len(images), 2), device=images.device)
-    images_out = torch.empty((len(images), 3, crop_size, crop_size), device=images.device, dtype=images.dtype)
-    if r <= 2:
-        for sy in range(-r, r+1):
-            for sx in range(-r, r+1):
-                mask = (shifts[:, 0] == sy) & (shifts[:, 1] == sx)
-                images_out[mask] = images[mask, :, r+sy:r+sy+crop_size, r+sx:r+sx+crop_size]
-    else:
-        images_tmp = torch.empty((len(images), 3, crop_size, crop_size+2*r), device=images.device, dtype=images.dtype)
-        for s in range(-r, r+1):
-            mask = (shifts[:, 0] == s)
-            images_tmp[mask] = images[mask, :, r+s:r+s+crop_size, :]
-        for s in range(-r, r+1):
-            mask = (shifts[:, 1] == s)
-            images_out[mask] = images_tmp[mask, :, :, r+s:r+s+crop_size]
-    return images_out
-
-class CifarLoader:
-    def __init__(self, path, train=True, batch_size=500, aug=None):
-        data_path = os.path.join(path, 'train.pt' if train else 'test.pt')
-        if not os.path.exists(data_path):
-            dset = torchvision.datasets.CIFAR10(path, download=True, train=train)
-            images = torch.tensor(dset.data)
-            labels = torch.tensor(dset.targets)
-            torch.save({'images': images, 'labels': labels, 'classes': dset.classes}, data_path)
-
-        data = torch.load(data_path, map_location='cpu')
[…]

### 2025-08-25 code/airbench.py
Clearing the code: CIFAR and others

@@ -1,367 +1,0 @@
-import os
-from math import ceil
-
-import torch
-from torch import nn
-import torch.nn.functional as F
-import torchvision
-import torchvision.transforms as T
-
-
-def infer(model, loader, tta_level=0):
-    
-    def infer_basic(inputs, net):
-        return net(inputs).clone()
-
-    def infer_mirror(inputs, net):
-        return 0.5 * net(inputs) + 0.5 * net(inputs.flip(-1))
-
-    def infer_mirror_translate(inputs, net):
-        logits = infer_mirror(inputs, net)
-        pad = 1 
-        padded_inputs = F.pad(inputs, (pad,)*4, 'reflect')
-        inputs_translate_list = [ 
-            padded_inputs[:, :, 0:32, 0:32],
-            padded_inputs[:, :, 2:34, 2:34],
-        ]   
-        logits_translate_list = [infer_mirror(inputs_translate, net)
-                                 for inputs_translate in inputs_translate_list]
-        logits_translate = torch.stack(logits_translate_list).mean(0)
-        return 0.5 * logits + 0.5 * logits_translate
-
-    model.eval()
-    test_images = loader.normalize(loader.images)
-    infer_fn = [infer_basic, infer_mirror, infer_mirror_translate][tta_level]
-    with torch.no_grad():
-        return torch.cat([infer_fn(inputs, model) for inputs in test_images.split(2000)])
-
-def evaluate(model, loader, tta_level=0):
-    logits = infer(model, loader, tta_level)
-    return (logits.argmax(1) == loader.labels).float().mean().item()
-
-#############################################
-#                DataLoader                 #
-#############################################
-
-CIFAR_MEAN = torch.tensor((0.4914, 0.4822, 0.4465))
-CIFAR_STD = torch.tensor((0.2470, 0.2435, 0.2616))
-
-def batch_flip_lr(inputs):
-    flip_mask = (torch.rand(len(inputs), device=inputs.device) < 0.5).view(-1, 1, 1, 1)
-    return torch.where(flip_mask, inputs.flip(-1), inputs)
-
-def batch_crop(images, crop_size):
-    r = (images.size(-1) - crop_size)//2
-    shifts = torch.randint(-r, r+1, size=(len(images), 2), device=images.device)
-    images_out = torch.empty((len(images), 3, crop_size, crop_size), device=images.device, dtype=images.dtype)
[…]

### 2025-08-25 code/cifar_bench/airbench_muon.py

@@ -1,10 +1,11 @@
 """
+This file reproduces 94.01% accuracy of training with F-Muon
 airbench94_muon.py
 Runs in 2.59 seconds on a 400W NVIDIA A100 using torch==2.4.1
 Attains 94.01 mean accuracy (n=200 trials)
 Descends from https://github.com/tysam-code/hlb-CIFAR10/blob/main/main.py
 """
-
+# we added: tuned NormalizedMuon and possibility of a momentum slowdown, which, however, is not useful
 #############################################
 #                  Setup                    #
 #############################################

### 2025-08-25 code/cifar_bench/create_fmuon_testing.py

@@ -1,0 +1,118 @@
+"""
+Script to run airbench_muon.py with different sgd_coeff values and plot the results.
+"""
+
+import os
+import sys
+import torch
+import matplotlib.pyplot as plt
+import numpy as np
+from math import ceil
+from airbench_muon import CifarNet, CifarLoader, evaluate, BatchNorm, Conv, ConvGroup
+from airbench_muon import NormalizedMuon
+
+def create_modified_main_function(sgd_coeff):
+    """
+    Create a modified version of the main function with the specified sgd_coeff.
+    """
+    def modified_main(run, model):
+        batch_size = 2000
+        bias_lr = 0.053
+        head_lr = 0.67
+        wd = 2e-6 * batch_size
+
+        test_loader = CifarLoader("cifar10", train=False, batch_size=2000)
+        train_loader = CifarLoader("cifar10", train=True, batch_size=batch_size, aug=dict(flip=True, translate=2))
+        if run == "warmup":
+            # The only purpose of the first run is to warmup the compiled model, so we can use dummy data
+            train_loader.labels = torch.randint(0, 10, size=(len(train_loader.labels),), device=train_loader.labels.device)
+        total_train_steps = torch.ceil(torch.tensor(8 * len(train_loader)))
+        whiten_bias_train_steps = torch.ceil(torch.tensor(3 * len(train_loader)))
+
+        # Create optimizers and learning rate schedulers
+        filter_params = [p for p in model.parameters() if len(p.shape) == 4 and p.requires_grad]
+        norm_biases = [p for n, p in model.named_parameters() if "norm" in n and p.requires_grad]
+        param_configs = [dict(params=[model.whiten.bias], lr=bias_lr, weight_decay=wd/bias_lr),
+                         dict(params=norm_biases,         lr=bias_lr, weight_decay=wd/bias_lr),
+                         dict(params=[model.head.weight], lr=head_lr, weight_decay=wd/head_lr)]
+        optimizer1 = torch.optim.SGD(param_configs, momentum=0.85, nesterov=True, fused=True)
+        
+        # Import NormalizedMuon here to avoid circular imports
+        optimizer2 = NormalizedMuon(filter_params, lr=0.24, momentum=0.6, nesterov=True, sgd_coeff=sgd_coeff)
+        optimizers = [optimizer1, optimizer2]
+        for opt in optimizers:
+            for group in opt.param_groups:
+                group["initial_lr"] = group["lr"]
[…]

### 2025-08-25 code/run_sgd_coeff_experiment.py

@@ -1,305 +1,0 @@
-"""
-Script to run airbench_muon.py with different sgd_coeff values and plot the results.
-"""
-
-import os
-import sys
-import torch
-import matplotlib.pyplot as plt
-import numpy as np
-from math import ceil
-from airbench_muon import CifarNet, CifarLoader, evaluate, BatchNorm, Conv, ConvGroup
-
-def create_modified_main_function(sgd_coeff):
-    """
-    Create a modified version of the main function with the specified sgd_coeff.
-    """
-    def modified_main(run, model):
-        batch_size = 2000
-        bias_lr = 0.053
-        head_lr = 0.67
-        wd = 2e-6 * batch_size
-
-        test_loader = CifarLoader("cifar10", train=False, batch_size=2000)
-        train_loader = CifarLoader("cifar10", train=True, batch_size=batch_size, aug=dict(flip=True, translate=2))
-        if run == "warmup":
-            # The only purpose of the first run is to warmup the compiled model, so we can use dummy data
-            train_loader.labels = torch.randint(0, 10, size=(len(train_loader.labels),), device=train_loader.labels.device)
-        total_train_steps = torch.ceil(torch.tensor(8 * len(train_loader)))
-        whiten_bias_train_steps = torch.ceil(torch.tensor(3 * len(train_loader)))
-
-        # Create optimizers and learning rate schedulers
-        filter_params = [p for p in model.parameters() if len(p.shape) == 4 and p.requires_grad]
-        norm_biases = [p for n, p in model.named_parameters() if "norm" in n and p.requires_grad]
-        param_configs = [dict(params=[model.whiten.bias], lr=bias_lr, weight_decay=wd/bias_lr),
-                         dict(params=norm_biases,         lr=bias_lr, weight_decay=wd/bias_lr),
-                         dict(params=[model.head.weight], lr=head_lr, weight_decay=wd/head_lr)]
-        optimizer1 = torch.optim.SGD(param_configs, momentum=0.85, nesterov=True, fused=True)
-        
-        # Import NormalizedMuon here to avoid circular imports
-        from airbench_muon import NormalizedMuon
-        optimizer2 = NormalizedMuon(filter_params, lr=0.24, momentum=0.6, nesterov=True, sgd_coeff=sgd_coeff)
-        optimizers = [optimizer1, optimizer2]
-        for opt in optimizers:
-            for group in opt.param_groups:
-                group["initial_lr"] = group["lr"]
[…]

### 2025-08-25 code/test_sgd_coeff.py

@@ -1,36 +1,0 @@
-"""
-Simple test script to verify the setup works before running the full experiment.
-dull code
-"""
-
-import torch
-from airbench_muon import CifarNet, NormalizedMuon
-
-def test_sgd_coeff():
-    """Test that we can create a model and optimizer with different sgd_coeff values."""
-    
-    print("Testing SGD coefficient setup...")
-    
-    # Create a simple model
-    model = CifarNet().cuda().to(memory_format=torch.channels_last)
-    
-    # Test different sgd_coeff values
-    test_coeffs = [-0.1, 0.0, 0.5, 1.0, 1.1]
-    
-    for coeff in test_coeffs:
-        try:
-            # Get filter parameters
-            filter_params = [p for p in model.parameters() if len(p.shape) == 4 and p.requires_grad]
-            
-            # Create optimizer with this coefficient
-            optimizer = NormalizedMuon(filter_params, lr=0.4, momentum=0.65, nesterov=True, sgd_coeff=coeff)
-            
-            print(f"  ✓ Successfully created NormalizedMuon with sgd_coeff = {coeff}")
-            
-        except Exception as e:
-            print(f"  ✗ Failed to create NormalizedMuon with sgd_coeff = {coeff}: {e}")
-    
-    print("\nTest completed!")
-
-if __name__ == "__main__":
-    test_sgd_coeff()

### 2025-08-25 code/L_smooth_tests/problems.py
Clearing L-smooth part

@@ -1,9 +1,10 @@
+# we use RandomQuadraticPSDProblem
 from math import sqrt
 from typing import Optional, Tuple
 
 import torch
 
-from benchmark_runner import MatrixProblem
+from optimizer_runner import MatrixProblem
 
 
 def _rand_orthogonal(n: int, device: torch.device, dtype: torch.dtype = torch.float32) -> torch.Tensor:

### 2025-08-25 code/cifar_bench/create_fmuon_testing.py

@@ -37,7 +37,6 @@ def create_modified_main_function(sgd_coeff):
                          dict(params=[model.head.weight], lr=head_lr, weight_decay=wd/head_lr)]
         optimizer1 = torch.optim.SGD(param_configs, momentum=0.85, nesterov=True, fused=True)
         
-        # Import NormalizedMuon here to avoid circular imports
         optimizer2 = NormalizedMuon(filter_params, lr=0.24, momentum=0.6, nesterov=True, sgd_coeff=sgd_coeff)
         optimizers = [optimizer1, optimizer2]
         for opt in optimizers:

### 2025-08-25 code/L_smooth_tests/run_benchmarks.py
Clearing L-smooth part

@@ -4,7 +4,7 @@ from math import sqrt
 import torch
 
 from optimizers import Muon, Neon, NormalizedMuon
-from benchmark_runner import MatrixProblem, run_optimizer_on_problem
+from optimizer_runner import MatrixProblem, run_optimizer_on_problem
 from benchmark_plotter import build_default_panels, plot_from_descriptions, plot_and_save_default_panels
 from problems import RandomQuadraticPSDProblem as SimpleQuadratic, LogisticRegressionProblem
 
@@ -49,14 +49,14 @@ def main() -> None:
     lr_nuc = grad_op_eps / beta_nuc / lipsch_const_nuc
     lr_fro = grad_op_eps / beta_fro / lipsch_const_fro
     print(f"Muon lr: {lr_op: .2E}, Neon iter: {lr_nuc: .2e}, Frobenius iter: {lr_fro:.2e}")
-    lr_op = lr_nuc = lr_fro = 0.01
 
     iter_num_op = int(lipsch_const_op * delta / grad_op_eps / grad_op_eps / beta_op / beta_op)
     iter_num_nuc = int(lipsch_const_nuc * delta / grad_op_eps / grad_op_eps / beta_nuc / beta_nuc)
     iter_num_fro = int(lipsch_const_fro * delta / grad_op_eps / grad_op_eps / beta_fro / beta_fro)
 
+    print(lr_op, lr_nuc, lr_fro)
     print(f"Muon iter: {iter_num_op}, Neon iter: {iter_num_nuc}, Frobenius iter: {iter_num_fro}")
-    iter_num_nuc = iter_num_op = iter_num_fro = 5000
+    iter_num_nuc = iter_num_op = iter_num_fro = 10000
     # iter_num = 100000
     X0 = 0.1 * torch.randn(m, n, dtype=torch.float32, device=device)
     # Define optimizers and settings
@@ -143,7 +143,7 @@ def main() -> None:
         verbose=True,
         name=sgd_name
     )
-    # experiments[sgd_name] = sgd_results
+    experiments[sgd_name] = sgd_results
     # experiments[sgd_name] = sgd_results
     # Build panels and plot
     panels = build_default_panels(experiments)

### 2025-08-25 code/cifar_bench/compare_sgd_coeffs.py

@@ -6,8 +6,9 @@ Useful for testing the setup before running the full experiment.
 import torch
 import matplotlib.pyplot as plt
 import numpy as np
+from airbench_muon import CifarNet
 from create_fmuon_testing import create_modified_main_function
-from plot_results import plot_results, plot_results_from_file
+from plot_sgd_coeffs_results import plot_results, plot_results_from_file
 
 def quick_test():
     """Run a quick test with fewer coefficient values and fewer runs."""
@@ -30,7 +31,6 @@ def quick_test():
         print(f"Running experiment with sgd_coeff = {sgd_coeff}")
         
         # Create and compile the model
-        from code.cifar_bench.airbench_muon import CifarNet
         model = CifarNet().cuda().to(memory_format=torch.channels_last)
         model.compile(mode="max-autotune")
         

### 2025-08-25 code/run_benchmarks.py

@@ -1,161 +1,0 @@
-from typing import Any, Dict, Optional, Tuple
-from math import sqrt
-
-import torch
-
-from optimizers import Muon, Neon, NormalizedMuon
-from benchmark_runner import MatrixProblem, run_optimizer_on_problem
-from benchmark_plotter import build_default_panels, plot_from_descriptions, plot_and_save_default_panels
-from problems import RandomQuadraticPSDProblem as SimpleQuadratic, LogisticRegressionProblem
-
-class SinFrobeniusNormProblem(MatrixProblem):
-    """
-    Problem: f(X) = sin(||X||_F)
-    ∂f/∂X = cos(r) * X / r, r = ||X||_F, and 0 if r = 0.
-    """
-
-    def objective(self, X: torch.Tensor) -> torch.Tensor:
-        r = torch.linalg.norm(X, ord="fro")
-        return torch.sin(r)
-
-    def gradient(self, X: torch.Tensor) -> torch.Tensor:
-        r = torch.linalg.norm(X, ord="fro")
-        if r.item() == 0.0:
-            return torch.zeros_like(X)
-        return torch.cos(r) * X / r
-
-
-def main() -> None:
-    device = torch.device("cuda" if torch.cuda.is_available() else "cpu")
-
-    # Define problem
-    
-    # Initial point
-    m, n = 50, 50
-    # problem = LogisticRegressionProblem(m, n) 
-    problem = SimpleQuadratic(m, n, device=device, seed=42, eig_range_M=(0,1), eig_range_N=(0,1))
-    delta = 250 # let it be
-    grad_op_eps = 1
-    lipsch_const_op = problem.lipschitz_constant(norm=2)
-    lipsch_const_nuc = problem.lipschitz_constant(norm='nuc')
-    lipsch_const_fro = problem.lipschitz_constant(norm='fro')
-    print(lipsch_const_op, lipsch_const_nuc, lipsch_const_fro)
-
-    beta_op = n / 2
-    beta_nuc = 1
-    beta_fro = sqrt(n)/2
-
-    lr_op = grad_op_eps / beta_op / lipsch_const_op
-    lr_nuc = grad_op_eps / beta_nuc / lipsch_const_nuc
-    lr_fro = grad_op_eps / beta_fro / lipsch_const_fro
-    print(f"Muon lr: {lr_op: .2E}, Neon iter: {lr_nuc: .2e}, Frobenius iter: {lr_fro:.2e}")
-
[…]

### 2025-08-25 code/matrix_functions.py
Preparing the code

@@ -69,8 +69,6 @@ def k_sv_svds_approximation_dlpack(W_torch, k, tau, num_iter=30):
     # print(opt_tau)
     return approx_torch, opt_tau, k
 
-    
-
 
 def svd_full_approximation(A, tau=0):
     k = 0
@@ -89,8 +87,6 @@ def svd_full_approximation(A, tau=0):
     # print(f"condition {S[0]/S[-1]:.4e}")
     return A_reconstructed
 
-
-import torch
 
 def randomized_svd_torch_full(G, n_components=1, n_oversamples=5, n_iter=2):
     """
@@ -251,6 +247,7 @@ def randomized_svd_torch(G, n_components=1, n_oversamples=5, n_iter=2):
     
     return u1, sigma1, v1
 
+# used in k-Neon
 def several_sv_svds_approximation(W_torch, k, num_iter=50):
     """SVD approximation using the top k singular values and corresponding vectors."""
     # Store original device and dtype
@@ -267,6 +264,7 @@ def several_sv_svds_approximation(W_torch, k, num_iter=50):
     
     return approx_torch_U, approx_torch_S, approx_torch_Vt
 
+# used in Neon
 def one_sv_svds_approximation(W_torch, num_iter=30):
     """SVD approximation using the top k singular values and corresponding vectors."""
     k = 1
@@ -304,6 +302,7 @@ def one_sv_svds_approximation(W_torch, num_iter=30):
     '''
     # return torch.outer(u1, v1), sigma1 # * - we do not need this! Muon doesn't have this 
 
+# similar to several_sv_svds, but returns only a product
 def lanczos_svdt(W_torch, k, num_iter=50):
     """SVD approximation using the top k singular values and corresponding vectors."""
     

### 2025-08-25 code/muon_airbench.py

@@ -1,194 +1,0 @@
-# Testing Neon on the benchmark, the code works for Muon around 3s indeed on a powerful GPU
-# Neon is far worse. Somehow its accuracy more depends on the batchsize.
-from math import ceil
-import torch
-torch.backends.cudnn.benchmark = True
-from torch import nn
-import torch.nn.functional as F
-import airbench
-
-from optimizers import Neon
-
-# @torch.compile
-def zeropower_via_newtonschulz5(G, steps=3, eps=1e-7):
-    assert len(G.shape) == 2
-    a, b, c = (3.4445, -4.7750,  2.0315)
-    X = G.bfloat16()
-    X /= (X.norm() + eps) # ensure top singular value <= 1
-    if G.size(0) > G.size(1):
-        X = X.T
-    for _ in range(steps):
-        A = X @ X.T
-        B = b * A + c * A @ A
-        X = a * X + B @ X
-    if G.size(0) > G.size(1):
-        X = X.T
-    return X
-
-class Muon(torch.optim.Optimizer):
-    def __init__(self, params, lr=1e-3, momentum=0, nesterov=False):
-        defaults = dict(lr=lr, momentum=momentum, nesterov=nesterov)
-        super().__init__(params, defaults)
-
-    def step(self):
-        for group in self.param_groups:
-            lr = group["lr"]
-            momentum = group["momentum"]
-            for p in group["params"]:
-                g = p.grad
-                if g is None:
-                    continue
-                state = self.state[p]
-
-                if "momentum_buffer" not in state.keys():
-                    state["momentum_buffer"] = torch.zeros_like(g)
-                buf = state["momentum_buffer"]
-                buf.mul_(momentum).add_(g)
-                g = g.add(buf, alpha=momentum) if group["nesterov"] else buf
-
-                p.data.mul_(len(p.data)**0.5 / p.data.norm()) # normalize the weight
-                update = zeropower_via_newtonschulz5(g.reshape(len(g), -1)).view(g.shape) # whiten the update
-                p.data.add_(update, alpha=-lr) # take a step
-
-# note the use of low BatchNorm stats momentum
-class BatchNorm(nn.BatchNorm2d):
-    def __init__(self, num_features, momentum=0.6, eps=1e-12):
-        super().__init__(num_features, eps=eps, momentum=1-momentum)
-        self.weight.requires_grad = False
[…]

### 2025-08-25 code/cifar_bench/airbench_muon.py
Revised comparison of matrix methods

@@ -425,7 +425,7 @@ def main(run, model):
                      dict(params=norm_biases,         lr=bias_lr, weight_decay=wd/bias_lr),
                      dict(params=[model.head.weight], lr=head_lr, weight_decay=wd/head_lr)]
     optimizer1 = torch.optim.SGD(param_configs, momentum=0.85, nesterov=True, fused=True)
-    optimizer2 = NormalizedMuon(filter_params, lr=0.4, momentum=0.65, nesterov=True, sgd_coeff=1)
+    optimizer2 = NormalizedMuon(filter_params, lr=0.4, momentum=0.65, nesterov=True, sgd_coeff=0.5)
     optimizers = [optimizer1, optimizer2]
     for opt in optimizers:
         for group in opt.param_groups:

### 2025-09-17 conf_article/neon.tex
Cleared in the code/norm folder
Minor revision in the article

@@
  That is what can be said from a practical point of view.
-From the perspective of theory, Muon's main innovation was an intentional usage of matrix norms, i.e.
+From the perspective of theory, the main innovation of Muon has been an intentional usage of matrix norms, i.e.
 the spectral norm, to derive the algorithm's update \citep{bernstein2025deriving}.
@@
 
- In this article, we focus on the two most common matrix norms akin to the spectral, namely, the nuclear norm and the Frobenius norm.
+ In this article, we focus on the two most common matrix norms akin to the spectral norm, namely the nuclear norm and the Frobenius norm.
 Working in the linear minimization oracle (lmo) framework, which is equivalent to a factor to the trust region approach and the steepest descent under norm constraint, we derive Neon, our algorithm based on the nuclear norm.
@@
 
- Noticing that Neon and Muon are diametrical in terms of the rank of the update matrix, we bridge the space by "regularizing" them by Normalized SGD, which is derived in lmo with the Frobenius norm.
-We do this in the same lmo approach by considering a norm that is dual to the convex combination of the Frobenius norm and the spectral or the nuclear norm respectively.
-So we derive the algorithms we name F-Neon and F-Muon respectively.
-
- Having faced the array of different Muon-like optimizers, according to the upper bounds from \citet{kovalev2025understanding,riabinin2025gluon}, with similar convergence behavior, we painstakingly compare them on a synthetic linear least squares problem with known Lipschitz constant.
-The efforts results in comparison of the algorithms by their convergence not in terms of the dual norms of their gradient, but in terms of the common spectral norm, which may strongly differ, especially in large matrices, from the initial dual norms.
-Thus, we compare the algorithms in a unified fashion and find that if an algorithm converges fast by the loss, it does not mean it converges fast by the norm of the gradient, and vice versa.
-
- Finally, we test Muon and F-Muon on two deep-learning tasks, training convolutional network on CIFAR-10 \citep{cifar2023airbench} and pre-training NanoGPT on FineWeb \citep{modded_nanogpt_2024}.
-The results support the supremacy of Muon, but the most striking result of the tests is that F-Muon, only half of which is Muon, surpasses Muon's accuracy on the CIFAR tasks by a margin.
-The case of F-Muon answers in the affirmative to the question of feasibility of constructing a mixture of optimization algorithms to increase robustness of the composite algorithm.
+ Noticing that Neon and Muon are diametrical in terms of the rank of the update matrix, we bridge the space by considering the dual to the Ky Fan $k$-norm, the primal norm being the sum of the $k$ largest singular values of the matrix.
+The resulting update turns out to be Dion \citet{ahn2025dioncommunicationefficientoptimizerlarge} without error feedback.
+
+ To make Neon and Dion without error feedback more robust, we "regularize" them by Normalized SGD (NSGD), which is derived in lmo with the Frobenius norm.
+We do this in the same lmo approach by considering a norm that is dual to the convex combination of the Frobenius norm and the Ky Fan k-norm.
+So we derive the algorithms we name F-Neon and F-Dion respectively.
+The same procedure we follow with Muon to obtain F-Muon.
+
+ Having faced the array of different Muon-like optimizers, which, according to the upper bounds from \citet{kovalev2025understanding,riabinin2025gluon}, must converge similarly, we painstakingly compare them on a synthetic linear least squares problem with a known Lipschitz constant.
+The efforts result in comparison of the algorithms by their convergence not in terms of the dual norms of the gradient, but in terms of the common spectral norm, which may strongly differ, especially in large matrices, from the initial dual norms.
+Thus, we compare the algorithms in a unified fashion and find that sometimes if an algorithm converges fast by the loss, it does not imply its fast convergence by the norm of the gradient, and vice versa.
+
+ Finally, observing the poor performace of Neon and Dion without error feedback, which negatively impacts F-Neon and F-Dion as well, we test Muon and F-Muon on two deep-learning tasks, training convolutional network on CIFAR-10 \citep{cifar2023airbench} and pre-training NanoGPT on FineWeb \citep{modded_nanogpt_2024}.
+The results support the supremacy of Muon, but the most striking outcome of the tests is that F-Muon, only half of which is Muon, is on par with Muon on training the convolutional network and only slighter worse that Muon on NanoGPT pre-training, while NSGD is far from that.
+The case of F-Muon answers in the affirmative to the question of possibility to perturb the norm that underpins the algorithm without sacrificing its performace.
+We envision that this effect will be used in the future to constuct easier to compute approximations of empirically optimal algorithms like Muon.
 
@@
         \begin{lemma}\label{lemma:ky_fan_update}    
-            When $\norm{\cdot} = \normkfk{\mM^k}^\dagger$, \cref{eq:our_update} turns into:
+            When $\norm{\cdot} = \normkfk{\mM^t}^\dagger$, \cref{eq:our_update} turns into:
 
             \begin{equation}\label{eq:ky_fan_update}
-                \mX^{k+1} = \mX^{k} - \eta \sum_{i=1}^{k}u_i v_i^\top
+                \mX^{t+1} = \mX^{t} - \eta \sum_{i=1}^{k}u_i v_i^\top
             \end{equation}
@@
 
-        We introduce the family of {\it Fanions}, which consists of {\it $k$-Fanions}, lmo-based algorithms under the $\normkfk{\mM^k}^\dagger$ norms.
+        We introduce the family of {\it Fanions}, which consists of {\it $k$-Fanions}, lmo-based algorithms under the $\normkfk{\mM^t}^\dagger$ norms.
 By this terminology and the lemma, Muon is a $\min\{n,m\}$-Fanion, while Neon is an 1-Fanion.
@@
 
-        In addition, one can consider F-KF-k-norm = $\alpha \normkfk{\cdot} + (1-\alpha)\normf{\cdot}$, the dual to which will produce algorithms like F-Dion without the error feedback.
+        In addition, one can consider F-KF-k-norm: $\norm{\cdot}_{\mathrm{F-KF-k}}$ = $\alpha \normkfk{\cdot} + (1-\alpha)\normf{\cdot}$, the dual to which will produce algorithms like F-Dion without the error feedback.
 The resulting family can be named {\it F-Fanions}.
@@
 
-    As the norms $\normf{\cdot}$, $\normn{\cdot}$, $\normf{\cdot}$ are almost proportional to each other when $m, n \rightarrow \infty$ (with high probability for random matrices), the expected convergence guarantees in terms of $\normf{\cdot}$ are the same (it can be easily shown by noting that $\norm{\mX}\sim\alpha\normf{\mX}$, $\norm{\nabla f(\mX)}^\dagger \sim \frac{1}{\alpha} \normf{\nabla f(\mX)}$, and expressing $L$-constant via $L_F$-constant for the Frobenius norm).
+    As the norms $\normf{\cdot}$, $\normn{\cdot}$, $\norms{\cdot}$ are almost proportional to each other when $m, n \rightarrow \infty$ (with high probability for random matrices), the expected convergence guarantees in terms of $\normf{\cdot}$ are the same (it can be easily shown by noting that $\norm{\mX}\sim\alpha\normf{\mX}$, $\norm{\nabla f(\mX)}^\dagger \sim \frac{1}{\alpha} \normf{\nabla f(\mX)}$, and expressing $L$-constant via $L_F$-constant for the Frobenius norm).
 
@@
 In this article, we have generalized several successful algorithms, like Muon and Dion, to the lmo-based algorithms in the $\normkfk{\cdot}^\dagger$ norm.
-Also we have proposed the technique of ``regularizing'' the updates with NSGD, a trick to increase the robustness of the algorithms and motivated by the consideration of the $\normfstar{\cdot}$ and $\normftwo{\cdot}$ norms.
-Generalizations of well-known norms and subsequent combination of them may further improve performance of lmo-based algorithms.
-If a theory is developed that explains the discrepancy between performance of different algorithms based on the matrix norms, one will probably be able to apply it to the generalizations and combinations of the norms as well, which leaves an open space to construct norms specific for given architectures and probably even their parameters.
+Also we have proposed the technique of ``regularizing'' the updates with NSGD, a trick to increase the robustness of the algorithms, which is motivated by the consideration of the $\normfstar{\cdot}$, $\normftwo{\cdot}$, and general $\norm{\cdot}_{\mathrm{F-KF-k}}$ norms.
+Generalizations of well-known norms and subsequent combinations of them may further improve performance of lmo-based algorithms.
+If a theory is developed that explains the discrepancy between performance of different algorithms based on the matrix norms, one will probably be able to intentionally construct norms optimal to given architectures and probably even their parameters.
 
@@
         \centering
-        \includegraphics[width=\linewidth]{fstardualball.pdf}
+        \includegraphics[width=\linewidth]{figs/fstardualball.pdf}
         \caption{lmo balls for F-Muon for different $\alpha$}
@@
         \centering
-        \includegraphics[width=\linewidth]{ftwodualball.pdf}
+        \includegraphics[width=\linewidth]{figs/ftwodualball.pdf}
         \caption{lmo ball for F-Neon for different $\alpha$}
@@
         \centering
-        \includegraphics[width=\linewidth]{KyFan.pdf}
+        \includegraphics[width=\linewidth]{figs/KyFan.pdf}
         \caption{Ky Fan 2-norm}
@@
         \centering
-        \includegraphics[width=\linewidth]{KyFanDual.pdf}
+        \includegraphics[width=\linewidth]{figs/KyFanDual.pdf}
         \caption{Dual to Ky Fan 2-norm}
@@
 Proof of \cref{lemma:ky_fan_update}:
-Since $\norm{\cdot}^\dagger = \normkfk{\cdot}^{\dagger\dagger} = \normkfk{\cdot}$, the goal to reach $\normkfk{\mM^k}$.
+Since $\norm{\cdot}^\dagger = \normkfk{\cdot}^{\dagger\dagger} = \normkfk{\cdot}$, the goal to reach $\normkfk{\mM^t}$.
 
 Let us note that $\Delta = \sum_{i=1}^{k} u_i v_i^\top$ delivers the value.
-Indeed, $\<\mM^k, \Delta> =\<\mU \mSigma \mV^\top, \sum_{i=1}^{k} u_i v_i^\top> = \sum_{i,j=1}^{r, k}\<u_i \sigma_i v_i^\top, u_j v_j^\top> = \sum_{i=1}^{k}\sigma_i = \normkfk{\mM^k}$, which completes the proof.
+Indeed, $\<\mM^t, \Delta> =\<\mU \mSigma \mV^\top, \sum_{i=1}^{k} u_i v_i^\top> = \sum_{i,j=1}^{r, k}\<u_i \sigma_i v_i^\top, u_j v_j^\top> = \sum_{i=1}^{k}\sigma_i = \normkfk{\mM^t}$, which completes the proof.
 

### 2025-09-24 conf_article/neon.tex
prepared the article to be commented on

@@
 \usepackage{subcaption}
+\usepackage{xcolor}
+\usepackage{changes}
+\definechangesauthor[name={Ivan Kozyrev}, color=orange]{IK}
 \newtheorem{lemma}{Lemma}
@@
 
-%\icompfinalcopy % Uncomment for camera-ready version, but NOT for submission.
+\icompfinalcopy % Uncomment for camera-ready version, but NOT for submission.
 
@@
 \usepackage{arydshln} % dashed lines
+\usepackage[right=5cm, marginparwidth=4.6cm, marginparsep=0.2cm]{geometry} % For comments!
 
@@
 \section{Introduction}
+This is the original text. \added[id=IK, comment={I think we should mention this.}]{This is a new sentence that has been added.}
+
  Minimizing loss functions in unprecedentedly high-dimensional spaces has recently become an integral and crucial part in training large language models.

### 2025-09-24 conf_article/neon.tex
did some editing, mainly in introduction

@@
 In this article, we explore the use of matrix norms for optimizing functions of weight matrices, a crucial problem in training large language models.
-Moving beyond the spectral norm that underlies the Muon update, we leverage the Ky Fan k-norms to introduce a new family of Muon-like algorithms we name {\it Fanions}.
-Then, we consider their convex combinations with the Frobenius norm and corresponding duals to develop a new family of algorithms we name {\it F-Fanions}, one of them being {\it F-Muon}.
-We complement our theoretical analysis with an extensive empirical study of the algorithms across a wide range of tasks and settings, from which it follows that properly-tuned F-Muon is almost on par with Muon, which questions the necessity of finding the exact solution to the problem in linear optimization oracle framework.
+Moving beyond the spectral norm that underlies the Muon update, we leverage the duals to the Ky Fan k-norms to introduce a \deleted[id=IK, comment={Redundant}]{new} family of Muon-like algorithms we name {\it Fanions}, which happen to be similar to Dion.
+\replaced[id=IK, comment={This rewrite clarifies the sentence structure, making it clear that the norms (not the algorithms) are being combined.}]{Subsequently, we construct a second family of algorithms, dubbed {\it F-Fanions}, based on the duals of convex combinations of Ky Fan k-norms and the Frobenius norm.
+One prominent member of this family is {\it F-Muon}.}{Then, we consider their convex combinations with the Frobenius norm and corresponding duals to develop a new family of algorithms we name {\it F-Fanions}, one of them being {\it F-Muon}.} We complement our theoretical analysis with an extensive empirical study of the algorithms across a wide range of tasks and settings, from which it follows that properly-tuned F-Muon is almost on par with Muon, \replaced[id=IK, comment={Alex, clarify this please.
+It's a very non-obvious and non-clear conclusion, especially when knowing that Muon itself only provides an approximate answer to the lmo.}]{which questions the necessity of finding the exact solution to the problem in linear optimization oracle framework.}
 
@@
 \section{Introduction}
-This is the original text. \added[id=IK, comment={I think we should mention this.}]{This is a new sentence that has been added.}
 
@@
 That is what can be said from a practical point of view.
-From the perspective of theory, the main innovation of Muon has been an intentional usage of matrix norms, i.e.
-the spectral norm, to derive the algorithm's update \citep{bernstein2025deriving}.
+\replaced[id=IK, comment={Rewrote this to be more precise.
+The innovation wasn't just 'using' a norm, but deriving the update from a principled, constrained optimization problem.
+I think this better reflects the contribution of the Muon paper.}]{From a theoretical perspective, a key innovation of Muon was its principled derivation of the update rule, which emerges as the solution to an optimization problem constrained by the spectral norm \citep{bernstein2025deriving}}{From the perspective of theory, the main innovation of Muon has been an intentional usage of matrix norms, i.e.
+the spectral norm, to derive the algorithm's update}.
 Based upon recent theoretical advances that explain some theory behind Muon, Scion and Gluon \citep{bernstein2025deriving,kovalev2025understanding,pethick2025training,riabinin2025gluon}, we explore application of other matrix norms to optimization of functions of matrices.
@@
 
- In this article, we focus on the two most common matrix norms akin to the spectral norm, namely the nuclear norm and the Frobenius norm.
+In this article, \replaced[id=IK, comment={I think this is much clearer to introduce all algorithms like that, the paragraphs where one of every two words is a ***on are confusing.}]{we generalize the spectral-norm-based approach by first leveraging the family of Ky Fan k-norms to derive a new class of algorithms, which we call Fanions.
+Subsequently, we create a second, hybrid family named F-Fanions, by constructing convex combinations of the Ky Fan norms with the Frobenius norm and taking dual of that composite norm.
+Working within the linear minimization oracle (lmo) framework~--- which is equivalent (up to a factor) to the trust region approach~--- we derive the explicit update formulas for both algorithm families.
+Later in \cref{sec:matrix-side} we explain how their updates can be approximated with the same $\cO(m n \min\{m,n\})$ complexity as Muon updates are computed via Newton-Schulz iterations.}{we focus on the two most common matrix norms akin to the spectral norm, namely the nuclear norm and the Frobenius norm.
 Working in the linear minimization oracle (lmo) framework, which is equivalent to a factor to the trust region approach and the steepest descent under norm constraint, we derive Neon, our algorithm based on the nuclear norm.
-In the section {\it Matrix side of the updates}, we explain how Neon updates can be computed asymptotically with the same $\cO(m n \min\{m, n\})$ number of FLOPs as Muon updates are computed via Newton-Schulz iterations.
+In the section {\it Matrix side of the updates}, we explain how Neon updates can be computed asymptotically with the same $\cO(m n \min\{m, n\})$ number of FLOPs as Muon updates are computed via Newton-Schulz iterations.}
 
@@
 
-\section{Matrix side of updates}
+\section{Matrix side of updates}\label{sec:matrix-side}
 To compute the algorithms' updates, we use thick-restarted Lanczos method for singular value problem (TRLan) on $\mM^{k\top}\mM^k$ or $\mM^k \mM^{k\top}$ matrices (the one with less size is picked), implemented in CuPy lSibrary \citep{cupy_svds_ref} and described in \citet{simonz1998thick}.
@@
 
-    The most pathological case is $\alpha < 0$, which corresponds to the lmo with an area that is not a norm ball!
+    The most pathological case is \replaced[id=IK, comment={In the pictures there are no experiments with $\alpha < 0$, but there is a one with $\alpha $}]{$\alpha > 1$}{$\alpha < 0$}, which corresponds to the lmo with an area that is not a norm ball!
 Despite this violation, the mixture of algorithms reaches almost the same accuracy as vanilla Muon.

### 2025-10-09 slides/neon_slides.tex
Draft of the slides done

\documentclass[aspectratio=169]{beamer}

% Theme and basics
\usetheme{Madrid}
\usecolortheme{default}
\usefonttheme{professionalfonts}
\setbeamertemplate{navigation symbols}{}

% Custom footline: short author and short title on all non-title slides
\setbeamertemplate{footline}{%
  \leavevmode
  \hbox{%
    \begin{beamercolorbox}[wd=\paperwidth,ht=2.5ex,dp=1ex,center]{author in head/foot}
      \usebeamerfont{author in head/foot}\insertshortauthor\,\textemdash\,\insertshorttitle
    \end{beamercolorbox}%
  }%
}

% Math and utilities
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{arydshln}
\usepackage{array}
\usepackage{mathtools}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{fontawesome5} % for question mark
%\usepackage[authoryear]{natbib}

% Citations use the standard natbib formatting (as in the article)
\usepackage[style=authoryear, maxcitenames=1, maxbibnames=3, terseinits=true, uniquelist=false, natbib=true,]{biblatex}
\addbibresource{../conf_article/icomp2024_conference.bib}
% Reuse article macros
\input{../conf_article/math_commands.tex}

\newcommand{\norm}[1]{\lVert #1\rVert}
\newcommand{\abs}[1]{\lvert #1 \rvert}

% \epsilon is predefined; override to use \varepsilon
\renewcommand{\epsilon}{\varepsilon}
\newcommand{\Rmn}{\R^{m\times n}}
\newcommand{\cB}{\mathcal{B}}
\newcommand{\cD}{\mathcal{D}}
\newcommand{\cN}{\mathcal{N}}
\newcommand{\cO}{\mathcal{O}}
\newcommand{\cS}{\mathcal{S}}
\newcommand{\Ed}[2]{\mathbb{E}_{#1}\left[#2\right]}
\usepackage{mathtools}
\DeclarePairedDelimiter{\sqne}{\|}{\|_2^2}
%\DeclarePairedDelimiter{\norme}{\|}{\|_2}
\DeclarePairedDelimiter{\normf}{\|}{\|_\mathrm{F}}
\DeclarePairedDelimiter{\normkfk}{\|}{\|_\mathrm{KF-k}}
\DeclarePairedDelimiter{\normfstar}{\|}{\|_\mathrm{F*}}
\DeclarePairedDelimiter{\normftwo}{\|}{\|_\mathrm{F2}}
\DeclarePairedDelimiter{\sqns}{\|}{\|_{\mathrm{op}}^2}
\DeclarePairedDelimiter{\norms}{\|}{\|_{\mathrm{op}}}
\DeclarePairedDelimiter{\sqnn}{\|}{\|_{\mathrm{nuc}}^2}
\DeclarePairedDelimiter{\sqnf}{\|}{\|_{\mathrm{F}}^2}
\DeclarePairedDelimiter{\normn}{\|}{\|_{\mathrm{nuc}}}
\DeclarePairedDelimiter{\normfkfk}{\|}{\|_{\mathrm{F-KF-k}}}
\def\<#1,#2>{\langle #1,#2\rangle}
\DeclarePairedDelimiter{\dotprod}{\langle}{\rangle}

\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator{\diag}{diag}

% Graphics from the article repo
\graphicspath{{../conf_article/figs/}{../figures/13may_muon_neon/}}

% Title
\title[The Ky Fan Norms and Beyond]{The Ky Fan Norms and Beyond: Dual Norms and Combinations for Matrix Optimization}
\author[A. Kravatskiy et al.]{Alexey Kravatskiy\inst{1} \and Ivan Kozyrev\inst{1} \and Nikolai Kozlov\inst{1} \and Alexander Vinogradov\inst{1} \and Daniil Merkulov\inst{1,2,3,4} \and Ivan Oseledets\inst{5,2}}
\institute[]{%
\inst{1} MIPT \and
\inst{2} Skoltech \and
\inst{3} HSE \and
\inst{4} SberAI \and
\inst{5} AIRI
}
\date{\today}

\begin{document}

%-------------------------------------------------------------------------------------
\begin{frame}[plain]
  \titlepage
\end{frame}
%-------------------------------------------------------------------------------------
\begin{frame}{Moving beyond the spectral norm and Muon}
    \begin{block}{Objective:  $\min_{\mX\in\R^{m\times n}} f(\mX)$}
    \faQuestionCircle \space What will happen if we change the spectral norm in the derivation of Muon?
    \end{block}
    \begin{block}{F-Fanions: Muon, Neon, NSGD, Dion without EF, and so much more}
        From $\norms{\cdot}$ and Muon:
        $$\mX^{t+1} = \mX^{t} - \eta U V^\top$$
        To $\normkfk{\cdot}^\dagger$ and a general F-Fanion:
        $$\mX^{t+1} = \mX^{t} - \eta \left(\alpha \sum_{i=1}^k{u_i v_i^\top} + (1-\alpha)\frac{\mM^t}{\normf{\mM^t}}\right), \alpha \in [0,1]$$
    \end{block}

\end{frame}
%-------------------------------------------------------------------------------------
\begin{frame}{Linear Minimization Oracle (LMO) and Trust Region}
% after omitting gamma_k, we obtain:
Let us equip $\Rmn$ with a norm $\norm{\cdot}$. Its dual is $\norm{\mX}^\dagger = \sup_{\norm{\mX'}\leq 1} \<\mX,\mX'>$. $\<\cdot, \cdot>$ is a Frobenius product.
\vspace{0.4em}

Both LMO and Trust Region lead to the update
$$\mX^{t+1} = \mX^{t} - \eta \arg\max_{\mX \in \cB_1} \<\mM^t, \mX> = \mX^{t} - \eta \{\Delta \in \cB_1 \mid \<\mM^t, \Delta> = \norm{\mM^t}^\dagger\}$$

\vspace{0.4em}

\textbf{Recipe}: It means we seek $\Delta$ from the 1-norm ball that delivers $\<\mM^t, \Delta> = \norm{\mM^t}^\dagger$. We will often return to the SVD $\mM^t = \mU \Sigma \mV^\top$.
\end{frame}
%-------------------------------------------------------------------------------------
\begin{frame}{Frobenius $\normf{\mM^{k}}$ and Normalized SGD}

\begin{block}{Deriving NSGD by Recipe} $\normf{\mM^t}^\dagger = \normf{\mM^t}$, and $\Delta = \frac{\mM^t}{\normf{\mM^t}}$ with $\normf{\Delta} = 1$ delivers it. Hence,
$$\mX^{t+1} = \mX^{t} - \eta \frac{\mM^t}{\normf{\mM^t}}$$
\end{block}
\end{frame}
%-------------------------------------------------------------------------------------
\begin{frame}{Spectral $\norms{\mM^{k}}$ and Muon. Nuclear $\normn{\mM^k}$ and Neon.}
\begin{block}{Deriving Muon by Recipe} $\norms{\mM^t}^\dagger = \normn{\mM^t}$, and $\Delta = \mU\mV^\top$ delivers it: $\norms{\Delta}$ = 1 and

$$\<\mU\Sigma\mV^\top, \mU\mV^\top> = \tr(\mV \Sigma \mU^\top \mU \mV^\top) = \tr\Sigma = \normn{\mM^t}$$

Hence,
$$\mX^{t+1} = \mX^{t} - \eta \mU \mV^\top$$

\end{block}

\begin{block}{Deriving Neon by Recipe} $\normn{\mM^t}^\dagger = \norms{\mM^t}$, and $\Delta = u_1 v_1^\top$ delivers it: $\normn{\Delta}$ = 1 and

$$\<\mU\Sigma\mV^\top, u_1 v_1^\top> = \tr(\mV \Sigma \mU^\top u_1 v_1^\top) = \tr\diag({\sigma_1, 0, \dots, 0}) = \sigma_1 = \norms{\mM^t}$$

Hence,
$$\mX^{t+1} = \mX^{t} - \eta u_1 v_1^\top$$
\end{block}
\end{frame}

%-------------------------------------------------------------------------------------
\begin{frame}{Of Matrix and Vector Algorithms}
    %\centering
    %\def\arraystretch{1.1}
    \footnotesize
    \begin{table}
    \caption{lmo optimizers in Schatten $S_p$ norms and in $l_p$ norms.}
    \label{tbl:mat_vs_vec_lmo}
    \begin{tabularx}{\linewidth}{|>{\raggedright\arraybackslash}X|c|c|c|}
    \hline
    Method & lmo constraint set $\mathcal D$ & lmo & Reference \\
    \hline\hline
    Normalized SGD & $l_2$-ball, $S_2$-ball & $-\eta \tfrac{g}{\norm{g}_2} = -\eta \tfrac{g}{\norm{g}_F}$ & \citep{hazan2015beyond} \\
    Momentum Normalized SGD & Ball in $l_2$, or Ball in $S_2$ & $-\eta \tfrac{g}{\norm{g}_2} = -\eta\tfrac{g}{\norm{g}_F}$ & \citep{cutkosky2020momentum}\\
    \hline
    SignSGD & Ball in Max-norm $l_\infty$ & $-\eta \sign(g)$ & \citep[Thm. 1]{bernstein2018signsgd} \\
    Signum  & Ball in Max-norm $l_\infty$ & $-\eta \sign(g)$ & \citep[Thm. 3]{bernstein2018signsgd} \\
    \hdashline
    Muon & Ball in Spectral $S_\infty$ & $-\eta UV^\top$ & \citep{jordan2024muon} \\
    \hline
    Gauss-Southwell Coordinate Descent & Ball in $l_1$ & $-\eta \{i: g_i \geq g_k \forall k\}$ & \citep[p.19]{shi2016primer}\\
    \hdashline 
    Neon & Ball in Nuclear $S_1$ & $-\eta u_1 v_1^\top$ & This work\\
    \hline
    \end{tabularx}
    \end{table}
\end{frame}
%------------------------------------------------------------------------------------------------------------
\begin{frame}{Understanding Dion by Thomas Pethick}
    Without momentum, Dion is simplified to
    $$
    \begin{aligned}
    \Delta &\leftarrow g + e  \\
    e &\leftarrow \Delta - \sum_{i=1}^r \sigma_i u_i v_i^\top \\
    x &\leftarrow x - \gamma \sum_{i=1}^r u_i v_i^\top
    \end{aligned}
    $$
    where $\gamma > 0$ and $\sum_{i=1}^r \sigma_i u_iv^\top_i$ is the rank-$r$ truncated SVD of $\Delta \in \Rmn$.
    \vspace{0.4em}

    \faQuestionCircle \space What will we get if we try $\normkfk{\mM}:=\sum_{i=1}^r{\sigma_i}$?
    \end{frame}
%-------------------------------------------------------------------------------------
\begin{frame}{Ky Fan $k$-rank $\normkfk{\mM^k}^\dagger$ and Fanions}

\begin{block}{Deriving Fanions by Recipe} $\normkfk{\mM^t}^{\dagger\dagger} = \normkfk{\mM^t}$, and $\Delta = \sum_{i=1}^k{u_i v_i^\top}$ with $\normkfk{\Delta}^\dagger = \max\{\frac{1}{k} \normn{\Delta}, \norms{\Delta}\} = \max\{\frac{1}{k} k, 1\} = 1$ delivers it:
    $$\<\mM^t, \Delta> =\<\mU \mSigma \mV^\top, \sum_{i=1}^{k} u_i v_i^\top> = \sum_{i,j=1}^{r, k}\<u_i \sigma_i v_i^\top, u_j v_j^\top> = \sum_{i=1}^{k}\sigma_i = \normkfk{\mM^t}$$

    Hence,
    $$\mX^{t+1} = \mX^{t} - \eta \sum_{i=1}^{k} u_i v_i^\top$$
    \end{block}
\end{frame}
%-----------------------------------------------------------------------------------------------------------
\begin{frame}{Computing the $k$-rank update: Lanczos method}
    \begin{center}
    \small
    \begin{tabular}{lccc}
      \toprule
      Method & rtol & $k$ & time (s) \\
      \midrule
      Power Iterations & 0.01 & 1 & 7.7 \\
      SVDS (TRLan) & 0.01 & 1 & \textbf{0.18} \\
      PCA Low Rank (RSVD) & 0.01 & 1 & 1.15 \\
      SVDS (TRLan) & 0.01 & 10 & \textbf{0.47} \\
      PCA Low Rank (RSVD) & 0.01 & 10 & 19.4 \\
      SVDS (TRLan) & 0.01 & 100 & \textbf{1.96} \\
      PCA Low Rank (RSVD) & 0.01 & 100 & 170 \\
      \bottomrule
    \end{tabular}
    \end{center}
    \vspace{0.4em}
    \centering
    \footnotesize Comparison on a \(5000\times5000\) matrix; rtol is the relative Frobenius error of $\sum_{i=1}^{k}u_i \sigma_i v_i^\top$
  \end{frame}
%--------------------------------------------------------------------------------------------------------------
\begin{frame}{Frobeniusize the norms!}
\faExclamationCircle \space $\normf{\cdot}^\dagger = \normf{\cdot}$ is not a Ky Fan norm, so NSGD is not a Fanion.
\vspace{1em}

Let us consider the convex combination of the Ky Fan rank-$k$ norm and the Frobenius norm, which we call F-KF-k-norm: $$\norm{\cdot}_{\mathrm{F-KF-k}} = \alpha \normkfk{\cdot} + (1-\alpha)\normf{\cdot}, \alpha \in [0,1]$$

\end{frame}
%--------------------------------------------------------------------------------------------------------------

\begin{frame}{Balls of Duals to Convex Combinations of Norms}
    \begin{lemma}\label{lemma:dual_to_conv_comb}
        Let $\norm{\cdot}_{(1)}$ and $\norm{\cdot}_{(2)}$ be norms on a finite-dimensional Euclidean space, and let $\alpha,\beta \geq 0$. Define 
        $$
        \norm{x} := \alpha \norm{x}_{(1)} + \beta \norm{x}_{(2)}.
        $$
        Then the dual unit ball of $\norm{\cdot}$ satisfies
        $$
        B_{\norm{\cdot}^\dagger} 
        = \alpha B_{\norm{\cdot}_{(1)}^\dagger} + \beta B_{\norm{\cdot}_{(2)}^\dagger},
        $$
        where $+$ denotes the Minkowski sum and $B_{\norm{\cdot}_{(i)}^\dagger}$ is the unit ball of the dual norm $\norm{\cdot}_{(i)}^\dagger$.
    \end{lemma}

\end{frame}
%-----------------------------------------------------------------------------------------------------------
\begin{frame}{$\normfkfk{\mM^k}^\dagger$ and F-Fanions}

    \begin{block}{Deriving F-Fanions by Recipe}
    $\normfkfk{\mM^t}^{\dagger\dagger} = \normfkfk{\mM^t}$, and $\Delta = \alpha\sum_{i=1}^k{u_i v_i^\top} + (1-\alpha)\frac{\mM^k}{\normf{\mM^k}}$ delivers it:
\begin{enumerate}
    \item $\normfkfk{\Delta}^\dagger \leq 1$ because $\sum_{i=1}^k{u_i v_i^\top}$ lies in $B_{\normkfk{\cdot}}$ and $\frac{\mM^k}{\normf{\mM^k}}$ lies in $B_{\normf{\cdot}}$, so $\Delta$ lies in the Minkowski sum
    \item $\<\mM^t, \Delta> =\alpha \<\mU \mSigma \mV^\top, \sum_{i=1}^{k} u_i v_i^\top> + (1-\alpha)\<\mM^t, \frac{\mM^t}{\normf{\mM^t}}> = \alpha \sum_{i=1}^{k}\sigma_i + (1-\alpha) \normf{\mM^t}= \normfkfk{\mM^t}$
\end{enumerate}
\vspace{1em}

        Hence,
        $$\mX^{t+1} = \mX^{t} - \eta \Big(\alpha\sum_{i=1}^{k} u_i v_i^\top + (1-\alpha)\frac{\mM^t}{\normf{\mM^t}}\Big)$$
        \end{block}
\end{frame}
%------------------------------------------------------------------------------------------------------------
% The next slides were generated by Cursor
\begin{frame}{F-Muon and F-Neon}
  Convex combinations with Frobenius norm:
  \[
    \|\mX\|_{F*}=\alpha\|\mX\|_*+(1-\alpha)\|\mX\|_F,\qquad
    \|\mX\|_{F2}=\alpha\|\mX\|_{op}+(1-\alpha)\|\mX\|_F.
  \]
  Dual-induced updates:
  \[
    \text{F-Muon: }\mX^{k+1}=\mX^k-\eta\Big(\alpha\,\mU\mV^\top+(1-\alpha)\tfrac{\mM^k}{\|\mM^k\|_F}\Big),
  \]
  \[
    \text{F-Neon: }\mX^{k+1}=\mX^k-\eta\Big(\alpha\,u_1 v_1^\top+(1-\alpha)\tfrac{\mM^k}{\|\mM^k\|_F}\Big).
  \]
\end{frame}
%-----------------------------------------------------------------------------------------------
\begin{frame}{Geometric intuition: Dual balls}
    \centering
    \begin{minipage}{0.48\textwidth}
        \includegraphics[height=0.75\textheight,keepaspectratio]{fstardualball.pdf}
    \end{minipage}\hfill
    \begin{minipage}{0.48\textwidth}
        \includegraphics[height=0.75\textheight,keepaspectratio]{ftwodualball.pdf}
    \end{minipage}

    \vspace{0.3em}
    \footnotesize Dual balls for F-Muon and F-Neon across \(\alpha\) in a 2D singular-value space
    \end{frame}
%---------------------------------------
\begin{frame}{F-Muon on CIFAR-10 airbench}
    \begin{columns}[T,totalwidth=\textwidth]
      \begin{column}{0.48\textwidth}
        \includegraphics[width=\linewidth]{muon_tuned_diff_alpha.pdf}
        \centering
        \scriptsize Params tuned for Muon \citep{cifar2023airbench}
      \end{column}
      \begin{column}{0.48\textwidth}
        \includegraphics[width=\linewidth]{fmuon_tuned_diff_alpha.pdf}
        \centering
        \scriptsize Params tuned for F-Muon
      \end{column}
    \end{columns}
    \vspace{0.6em}
    \centering

    \footnotesize F-Muon with \(\alpha\approx 0.5\) matches Muon after tuning
  \end{frame}

%----------------------------------------
\begin{frame}{LMO balls on CIFAR-10}
  \begin{figure}[t]
    \includegraphics[height=0.7\textheight]{fstardual_cifar.pdf}
    \centering

    \footnotesize LMO balls for learning rates from the experiment

    \faQuestionCircle \space How should we interpret cases with $\alpha > 1$?
  \end{figure}
\end{frame}
%-----------------------------------------------------------------------------------------------
\begin{frame}{NanoGPT speedrun}
    \begin{itemize}
      \item Setup from \citet{modded_nanogpt_2024}: 1750 iterations and cross-entropy loss lower than 3.28
      \item Muon: {\tt lr=0.05, momentum=0.95}, 3.279
      \item F-Muon with \(\alpha=0.5\): {\tt lr=0.07, momentum=0.95}, 3.281
    \end{itemize}
\end{frame}
%----------------------------------------
\begin{frame}{Trust-region bounds (L-smooth)}
  \begin{itemize}
    \item Deterministic: choose \(\eta=\cO(\epsilon/L)\), \(K=\cO(L\Delta_0/\epsilon^2)\) \citep{kovalev2025understanding}.
    \item Stochastic (with variance \(\sigma^2\), link between norms \(\rho\)): detailed schedules for \(\eta,\alpha,K\).
    \item Asymptotically, norms are proportional for large random matrices (Marchenko–Pastur), aligning guarantees across choices.
  \end{itemize}
\end{frame}

%----------------------------------------
\begin{frame}{Experiment: Random linear least squares}
  \begin{columns}[T,totalwidth=\textwidth]
    \begin{column}{0.48\textwidth}
      \includegraphics[width=\linewidth]{simple_lls/loss_vs_iteration_50x50.pdf}
      \centering
      \scriptsize Loss vs iteration
    \end{column}
    \begin{column}{0.48\textwidth}
      \includegraphics[width=\linewidth]{simple_lls/gradient_spectral_norm_vs_iteration_50x50.pdf}
      \centering
      \scriptsize Spectral norm of gradient vs iteration
    \end{column}
  \end{columns}
  \vspace{0.3em}
  \footnotesize Setup informed by \citet{kovalev2025understanding}.
\end{frame}

%----------------------------------------
\begin{frame}{Ky Fan geometry}
  \begin{figure}[t]
    \centering
    \includegraphics[width=0.45\linewidth]{KyFan.pdf}\hfill
    \includegraphics[width=0.45\linewidth]{KyFanDual.pdf}
    \caption{Ky Fan 2-norm ball and its dual (singular-value space).}
  \end{figure}
\end{frame}

%----------------------------------------
\begin{frame}{Conclusions}
  \begin{itemize}
    \item Fanions unify Muon/Neon via Ky Fan norms; F-Fanions add robustness via Frobenius mixing.
    \item TRLan enables efficient k-rank updates; practical speed benefits over power/RSVD.
    \item Empirically, F-Muon remains competitive on CIFAR-10 and NanoGPT.
  \end{itemize}
\end{frame}

%----------------------------------------
\begin{frame}[allowframebreaks]{References}
  % Match article's bibliography style and entries
  \printbibliography
  %\bibliographystyle{../conf_article/icomp2024_conference}
  %\bibliography{../conf_article/icomp2024_conference}
\end{frame}

\end{document}

### 2025-10-10 slides/neon_slides.tex
longer presentation done

@@
 %\usepackage[authoryear]{natbib}
+
+\usepackage{environ}
+\newcounter{aequation}
+\NewEnviron{aequation}{\refstepcounter{aequation}$$\BODY\eqno{\text{(A\theaequation)}}$$}
+\crefname{aequation}{assumption}{assumptions}
+\creflabelformat{aequation}{(#2A#1#3)}
 
@@
 \title[The Ky Fan Norms and Beyond]{The Ky Fan Norms and Beyond: Dual Norms and Combinations for Matrix Optimization}
-\author[A.
-Kravatskiy et al.]{Alexey Kravatskiy\inst{1} \and Ivan Kozyrev\inst{1} \and Nikolai Kozlov\inst{1} \and Alexander Vinogradov\inst{1} \and Daniil Merkulov\inst{1,2,3,4} \and Ivan Oseledets\inst{5,2}}
+\author[A. Kravatskiy et al.]{%
+  Alexey Kravatskiy\inst{1} \and
+  Ivan Kozyrev\inst{1} \and
+  Nikolai Kozlov\inst{1} \and
+  Alexander Vinogradov\inst{1} \and
+  Daniil Merkulov\inst{1,2,3,4} \and
+  Ivan Oseledets\inst{5,2}\\[0.3em]
+}
 \institute[]{%
-\inst{1} MIPT \and
-\inst{2} Skoltech \and
-\inst{3} HSE \and
-\inst{4} SberAI \and
+\inst{1} MIPT \hspace{1em}%
+\inst{2} Skoltech \hspace{1em}%
+\inst{3} HSE \hspace{1em}%
+\inst{4} AI4Science \hspace{1em}%
 \inst{5} AIRI
@@
         \centering
-        \scriptsize Params tuned for Muon \citep{cifar2023airbench}
+        \scriptsize With params tuned for Muon, as in \citep{cifar2023airbench}
       \end{column}
@@
         \centering
-        \scriptsize Params tuned for F-Muon
+        \scriptsize With params tuned for F-Muon
       \end{column}
@@
 
-    \footnotesize F-Muon with \(\alpha\approx 0.5\) matches Muon after tuning
+    \footnotesize F-Muon with \(\alpha = 0.5\) matches Muon after tuning
+
+    \faQuestionCircle \space How should we interpret cases with $\alpha > 1$?
   \end{frame}
@@
 
-    \faQuestionCircle \space How should we interpret cases with $\alpha > 1$?
   \end{figure}
@@
 \begin{frame}{NanoGPT speedrun}
+  Setup from \citet{modded_nanogpt_2024}: 1750 iterations and cross-entropy loss lower than 3.28
     \begin{itemize}
-      \item Setup from \citet{modded_nanogpt_2024}: 1750 iterations and cross-entropy loss lower than 3.28
-      \item Muon: {\tt lr=0.05, momentum=0.95}, 3.279
-      \item F-Muon with \(\alpha=0.5\): {\tt lr=0.07, momentum=0.95}, 3.281
+      \item Muon: {\tt lr=0.05, momentum=0.95}, final loss = 3.279
+      \item F-Muon with \(\alpha=0.5\): {\tt lr=0.07, momentum=0.95}, final loss = 3.281
+      \item NSGD: {\tt lr=0.07, momentum=0.96}, final loss = 3.4651
+      \item F-Muon with $\alpha=0.5$: {\tt lr=0.07, momentum=0.96}, final loss = 3.2824!
     \end{itemize}
 \end{frame}
+%----------------------------------------------------------------------------------------------
+\begin{frame}{Formal assumptions}
+  L-smoothness:
+  \begin{aequation}\label{eq:L}
+    \norm{\nabla f(\mX) - \nabla f(\mX')}^\dagger \leq L \norm{\mX-\mX'}
+    \quad\text{for all}\;
+    \mX, \mX' \in \Rmn,
+    \end{aequation}
+
+  Bounded variance:
+    \begin{aequation}\label{eq:variance}
+      \Ed{\xi \sim \cD}{g(\mX;\xi)} = \nabla f(\mX)
+      \quad\text{and}\quad
+      \Ed{\xi \sim \cD}{\sqnf{g(\mX;\xi) - \nabla f(\mX)}} \leq \sigma^2
+      \quad\text{for all}\;
+      \mX \in \Rmn,
+      \end{aequation}
+
+  Bounded norm:
+      \begin{aequation}\label{eq:norm}
+        \norm{\mX}^\dagger \leq \rho\cdot\normf{\mX}
+        \quad\text{for all}\;
+        \mX \in \Rmn,
+        \end{aequation}
+
+  From Marchenko-Pastur law, $\normf{\cdot} \sim \frac{\sqrt{n}}{2}\norms{\cdot}$ and $\normn{\cdot} \sim \frac{n}{2}\norms{\cdot}$ for large $n$.
+
+\end{frame}
+\begin{frame}{Bounds from \citep{kovalev2025understanding}}
+  \begin{lemma}\label{lemma:stoch_tr}
+    To reach the precision $\E{\min_{k=1\ldots K} \norm{\nabla f(\mX_k)}^\dagger} \leq \epsilon$ under the assumptions \Cref{eq:L}, \Cref{eq:variance}, \Cref{eq:norm}, it is sufficient to choose the parameters as follows:
+    \begin{align}
+        \eta &= \cO\left(\min\left\{\frac{\epsilon}{L}, \frac{\epsilon^3}{\rho^2\sigma^2L}\right\}\right),
+        \qquad
+        \alpha = \cO\left(\min\left\{1, \frac{\epsilon^2}{\rho^2\sigma^2}\right\}\right),
+        \\
+        \label{eq:str_K_nonconvex}
+        K &= \cO\left(\max\left\{
+            \frac{\rho\sigma}{\epsilon},
+            \frac{\rho^3\sigma^3}{\epsilon^3},
+            \frac{L\Delta_0}{\epsilon^2},
+            \frac{L\Delta_0\rho^2\sigma^2}{\epsilon^4}
+        \right\}\right).
+    \end{align}
+    \end{lemma}
+
+\end{frame}
 %----------------------------------------
-\begin{frame}{Trust-region bounds (L-smooth)}
-  \begin{itemize}
-    \item Deterministic: choose \(\eta=\cO(\epsilon/L)\), \(K=\cO(L\Delta_0/\epsilon^2)\) \citep{kovalev2025understanding}.
-    \item Stochastic (with variance \(\sigma^2\), link between norms \(\rho\)): detailed schedules for \(\eta,\alpha,K\).
-    \item Asymptotically, norms are proportional for large random matrices (Marchenko–Pastur), aligning guarantees across choices.
-  \end{itemize}
-\end{frame}
-
-%----------------------------------------
-\begin{frame}{Experiment: Random linear least squares}
+\begin{frame}{Random linear least squares: Loss and the Spectral Norm of the gradient}
   \begin{columns}[T,totalwidth=\textwidth]
@@
   \vspace{0.3em}
-  \footnotesize Setup informed by \citet{kovalev2025understanding}.
+  \centering
+  \footnotesize Linear least squares problem for a 50 x 50 matrix. Common {\tt lr=0.01}.
 \end{frame}
@@
 %----------------------------------------
-\begin{frame}{Ky Fan geometry}
-  \begin{figure}[t]
+\begin{frame}{Geometric intuition: Ky Fan norms}
   \centering
-    \includegraphics[width=0.45\linewidth]{KyFan.pdf}\hfill
-    \includegraphics[width=0.45\linewidth]{KyFanDual.pdf}
-    \caption{Ky Fan 2-norm ball and its dual (singular-value space).}
-  \end{figure}
+  \begin{minipage}{0.46\textwidth}
+      \includegraphics[height=0.75\textheight,keepaspectratio]{KyFan.pdf}
+  \end{minipage}\hfill
+  \begin{minipage}{0.46\textwidth}
+      \includegraphics[height=0.75\textheight,keepaspectratio]{KyFanDual.pdf}
+  \end{minipage}
+
+  \vspace{0.3em}
+  \footnotesize Ky Fan rank-2 ball and its dual in a 3D singular-value space
   \end{frame}
-
 %----------------------------------------
@@
   \begin{itemize}
-    \item Fanions unify Muon/Neon via Ky Fan norms; F-Fanions add robustness via Frobenius mixing.
-    \item TRLan enables efficient k-rank updates; practical speed benefits over power/RSVD.
-    \item Empirically, F-Muon remains competitive on CIFAR-10 and NanoGPT.
+    \item Muon and NSGD are special cases of F-Fanions
+    \item Mixture of norm-based updates is a norm-based update!
+    \item No existing bounds describe the superiority of the spectral norm
   \end{itemize}
+
+  \vspace{2em}
+  \textbf{For future experiments}: Exploration of matrix and vector algorithms correspondence
 \end{frame}

### 2025-10-11 conf_article/neon.tex
continued with updates in introductiuon

@@
 the spectral norm, to derive the algorithm's update}.
+
+Motivated by the success of Muon, many generalizations and variations of it were proposed.
+Among the notable ones are Scion \citep{}, Deon \citep{} and Gluon \citep{}.
+One question still persists however: Why
+
 Based upon recent theoretical advances that explain some theory behind Muon, Scion and Gluon \citep{bernstein2025deriving,kovalev2025understanding,pethick2025training,riabinin2025gluon}, we explore application of other matrix norms to optimization of functions of matrices.
@@
 
-In this article, \replaced[id=IK, comment={I think this is much clearer to introduce all algorithms like that, the paragraphs where one of every two words is a ***on are confusing.}]{we generalize the spectral-norm-based approach by first leveraging the family of Ky Fan k-norms to derive a new class of algorithms, which we call Fanions.
+\replaced[id = IK, comment={Subsequent paragraphs in the introduction are almost completely new.}]{}{}
+
+%In this article, \replaced[id=IK, comment={I think this is much clearer to introduce all algorithms like that, the paragraphs where one of every two words is a ***on are confusing.}]{we generalize the spectral-norm-based approach by first leveraging the family of Ky Fan k-norms to derive a new class of algorithms, which we call Fanions.
 Subsequently, we create a second, hybrid family named F-Fanions, by constructing convex combinations of the Ky Fan norms with the Frobenius norm and taking dual of that composite norm.
@@
 
-Noticing that Neon and Muon are diametrical in terms of the rank of the update matrix, we bridge the space by considering the dual to the Ky Fan $k$-norm, the primal norm being the sum of the $k$ largest singular values of the matrix.
+%Noticing that Neon and Muon are diametrical in terms of the rank of the update matrix, we bridge the space by considering the dual to the Ky Fan $k$-norm, the primal norm being the sum of the $k$ largest singular values of the matrix.
 The resulting update turns out to be Dion \citet{ahn2025dioncommunicationefficientoptimizerlarge} without error feedback.
 
-To make Neon and Dion without error feedback more robust, we "regularize" them by Normalized SGD (NSGD), which is derived in lmo with the Frobenius norm.
+%To make Neon and Dion without error feedback more robust, we "regularize" them by Normalized SGD (NSGD), which is derived in lmo with the Frobenius norm.
 We do this in the same lmo approach by considering a norm that is dual to the convex combination of the Frobenius norm and the Ky Fan k-norm.
@@
 
-Having faced the array of different Muon-like optimizers, which, according to the upper bounds from \citet{kovalev2025understanding,riabinin2025gluon}, must converge similarly, we painstakingly compare them on a synthetic linear least squares problem with a known Lipschitz constant.
+%Having faced the array of different Muon-like optimizers, which, according to the upper bounds from \citet{kovalev2025understanding,riabinin2025gluon}, must converge similarly, we painstakingly compare them on a synthetic linear least squares problem with a known Lipschitz constant.
 The efforts result in comparison of the algorithms by their convergence not in terms of the dual norms of the gradient, but in terms of the common spectral norm, which may strongly differ, especially in large matrices, from the initial dual norms.
@@
 
-Finally, observing the poor performace of Neon and Dion without error feedback, which negatively impacts F-Neon and F-Dion as well, we test Muon and F-Muon on two deep-learning tasks, training convolutional network on CIFAR-10 \citep{cifar2023airbench} and pre-training NanoGPT on FineWeb \citep{modded_nanogpt_2024}.
+%Finally, observing the poor performace of Neon and Dion without error feedback, which negatively impacts F-Neon and F-Dion as well, we test Muon and F-Muon on two deep-learning tasks, training convolutional network on CIFAR-10 \citep{cifar2023airbench} and pre-training NanoGPT on FineWeb \citep{modded_nanogpt_2024}.
 The results support the supremacy of Muon, but the most striking outcome of the tests is that F-Muon, only half of which is Muon, is on par with Muon on training the convolutional network and only slighter worse that Muon on NanoGPT pre-training, while NSGD is far from that.

### 2025-10-13 poster/old/neon_poster_airi.tex
finished poster

--- 2025-05-13 poster/neon_poster.tex
@@
 
+% Russian language support
+\usepackage[utf8]{inputenc}
+\usepackage[T2A]{fontenc}
+\usepackage[russian]{babel}
+
 % These colours are tried and tested for titles and headers. Don't
@@
 %\def\Head#1{\noindent\hbox to \hsize{\hfil{\LARGE\color{bluegray} #1}}\bigskip}
-\def\Head#1{\noindent{\LARGE\color{bluegray} #1}\bigskip}
+\def\Head#1{\noindent{\LARGE\color{black} #1}\bigskip}
 \def\LHead#1{\noindent{\LARGE\color{bluegray} #1}\bigskip}
 \def\Subhead#1{\noindent{\large\color{bluegray} #1}\bigskip}
-\def\Title#1{\noindent{\VeryHuge\color{Red} #1}}
+\def\Title#1{\noindent\hbox to \hsize{\hfil{\fontsize{72}{86}\selectfont\color{Red} #1}}}
 
@@
 
+% AIRIxTGU logo in the top left corner
+\begin{textblock}{2}(0,0)
+\includegraphics[width=4\TPHorizModule]{AIRIxTGU_logos.pdf}
+\end{textblock}
+
+% "Лето с AIRI" in the top right corner
+\begin{textblock}{5}(19.5, 0)
+    \begin{center}
+    {\Large \textcolor{bluegray}{\textbf{Лето с AIRI}}}
+    \end{center}
+    \end{textblock}
+
 % This gives good title positioning for a portrait poster.
@@
 % but it looks awful.
-\begin{textblock}{23}(0,-0.25)
-\Title{Neon: Nuclear Norm to Beat Muon}
+\begin{textblock}{23}(0,-0.1)
+    \begin{center}
+    {\fontsize{96}{110}\selectfont\color{Red} \textbf{Neon: Nuclear Norm to Beat Muon}}
+    \end{center}
     \end{textblock}
 
-\begin{textblock}{23}(0,0.35)
-{
-\LARGE
-Alexey Kravatskiy, Ivan Kozyrev, Nikolay Kozlov, Alexander Vinogradov, Daniil Merkulov, Ivan Oseledets
-}
+\begin{textblock}{23}(0,0.2)
+    \begin{center}
+    {\large Alexey Kravatskiy, Ivan Kozyrev, Nikolay Kozlov, Alexander Vinogradov, Daniil Merkulov, Ivan Oseledets}
 
-{
-\Large
-\color{bluegray}
-\emph{Optimization Class Project. MIPT}
-}
+    \vspace{0.5em}
+
+    {\normalsize \color{bluegray} \emph{Optimization Class Project, MIPT}}
+    \end{center}
     \end{textblock}
 
-% Uni logo in the top right corner. A&A in the bottom left. Gives a
-% good visual balance, but you may want to change this depending upon
-% the graphics that are in your poster.
-%\begin{textblock}{2}(0,10)
-%Your logo here
-%%\includegraphics{/usr/local/share/images/AandA.epsf}
-%\end{textblock}
+% Vertical lines to separate sections - very thin and grey
+\begin{textblock}{0.02}(7.5,1.0)
+\textcolor{lightgray}{\rule{0.02\TPHorizModule}{10\TPVertModule}}
+\end{textblock}
 
-%\begin{textblock}{2}(21.2,0)
-%Another logo here
-%%\resizebox{2\TPHorizModule}{!}{\includegraphics{/usr/local/share/images/GUVIu/GUVIu.eps}}
-%\end{textblock}
+\begin{textblock}{0.02}(15.5,1.0)
+\textcolor{lightgray}{\rule{0.02\TPHorizModule}{10\TPVertModule}}
+\end{textblock}
 
@@
 
-\hrule\medskip
 \Head{Introduction}\\
@@
 To bridge this gap, we propose Neon, a new optimizer that builds upon the framework of Bernstein and Newhouse.
-By using alternative norms, such as nuclear norm (Nuclear-Neon) or custom $F*$ norm ($F*$-Neon), we induce low-rank update matrices that enable more efficient computation.
-We evaluate the performance of Neon, Muon, and Adam on MLP, CIFAR10, and NanoGPT, and demonstrate [resting results here].
+By using alternative norms, such as nuclear norm (Nuclear-Neon) or custom $F*$ norm ($F*$-Neon), we obtain low-rank update matrices that enable more efficient computation.
+We evaluate the performance of Neon, Muon, and Adam on training MLP and CNN on CIFAR-10 and on finetuning NanoGPT.
 
 \medskip
-\hrule\medskip
 \Head{Neon's update rule}\\
@@
 \medskip
-\hrule\medskip
 \Head{Efficient update computation}\\
@@
 \begin{textblock}{7.0}(8,1.0)
-\hrule\medskip
 \Head{Algorithms}\\
@@
 
-\hrule\medskip
+\medskip
 \Head{MLP tests}\\
@@
 
-%\begin{textblock}{7.0}(16,1.5)
-
 \begin{textblock}{7.0}(16,0.75)
-
-\hrule\medskip
 \Head{CIFAR-10 tests}\\
@@
 \medskip
-\hrule\medskip
 \Head{NanoGPT tests}\\
@@
 \medskip
-\hrule\medskip
 \Head{Conclusion}\\
 A new norm-motivated algorithm proved to be not so effective as Muon, at least from the first sight.
-However, its updates seem to be much faster to compute, and more so for a large matrices which has not tested yet.
-We believe that our algorithm is not an alternative to Muon, but probably a more effective tool for solving some other optimization problem.
-A deep inspection of Nuclear Norm and its properties might give us an insight into Neon's future.
+However, its updates seem to be much faster to compute, and more so for large matrices that have not been tested yet.
+We believe our algorithm is not an alternative to Muon, but probably a more effective tool for solving some other optimization problems.
+A deep inspection of nuclear norm and its properties might give us an insight into Neon's future.
 
 \medskip
-%\hrule\medskip
 %\Head{Acknowledgements}\\
@@
 
+\begin{textblock}{6}(-0,11.3)
+    \footnotesize
+    \raggedright
+    \textcolor{gray}{\emph{Jeremy Bernstein and Laker Newhouse. Old optimizer, new norm: An anthology (2024)}}
+    \end{textblock}
+
+    \begin{textblock}{6}(-0,11.5)
+    \footnotesize
+    \raggedright
+    \textcolor{gray}{\emph{Jeremy Bernstein. Deriving Muon (2025)}}
+    \end{textblock}
+
+    \begin{textblock}{6}(-0,11.7)
+    \footnotesize
+    \raggedright
+    \textcolor{gray}{\emph{Keller Jordan et al. Muon: An optimizer for hidden layers in neural networks (2024)}}
+    \end{textblock}
+
+% QR code in the lower right corner
+\begin{textblock}{2}(21.2,9.6)
+\begin{center}
+\includegraphics[width=0.7\TPHorizModule]{neon_repo_qr.png}
+\end{center}
+\end{textblock}
+
 \end{document}

### 2025-10-13 poster/old/neon_poster.tex

--- 2025-05-13 poster/neon_poster.tex
@@
 
-\documentclass[a0]{a0poster}
+\documentclass[a1]{a1poster}
 % You might find the 'draft' option to a0 poster useful if you have
@@
 To bridge this gap, we propose Neon, a new optimizer that builds upon the framework of Bernstein and Newhouse.
-By using alternative norms, such as nuclear norm (Nuclear-Neon) or custom $F*$ norm ($F*$-Neon), we induce low-rank update matrices that enable more efficient computation.
-We evaluate the performance of Neon, Muon, and Adam on MLP, CIFAR10, and NanoGPT, and demonstrate [resting results here].
+By using alternative norms, such as nuclear norm (Nuclear-Neon) or custom $F*$ norm ($F*$-Neon), we obtain low-rank update matrices that enable more efficient computation.
+We evaluate the performance of Neon, Muon, and Adam on training MLP and CNN on CIFAR-10 and on finetuning NanoGPT.
 
@@
 A new norm-motivated algorithm proved to be not so effective as Muon, at least from the first sight.
-However, its updates seem to be much faster to compute, and more so for a large matrices which has not tested yet.
-We believe that our algorithm is not an alternative to Muon, but probably a more effective tool for solving some other optimization problem.
-A deep inspection of Nuclear Norm and its properties might give us an insight into Neon's future.
+However, its updates seem to be much faster to compute, and more so for large matrices that have not been tested yet.
+We believe our algorithm is not an alternative to Muon, but probably a more effective tool for solving some other optimization problems.
+A deep inspection of nuclear norm and its properties might give us an insight into Neon's future.
 

### 2025-10-13 poster/main.tex

\documentclass[fontsize=10pt]{beamer}
%% Possible paper sizes: a0, a0b, a1, a2, a3, a4.
%% Possible orientations: portrait, landscape
%% Font sizes can be changed using the scale option.
\usepackage[size=custom, width=120, height=90, orientation=landscape, scale=1]{beamerposter}
\usetheme{LLT-poster}
\usecolortheme{ComingClean}
%\usecolortheme{Entrepreneur}
%\usecolortheme{ConspiciousCreep}  %% VERY garish.

%\usepackage[utf8]{inputenc}
\usepackage{libertine}
\usepackage[scaled=1]{inconsolata}
\usepackage{svg}
\usepackage{tabularx}
\usepackage{calc}
\usepackage{makecell}
\usepackage{csquotes}
\usepackage{caption}
\usepackage{mathtools}
\usepackage{arydshln}  % For \hdashline in tables
\usepackage{booktabs}
\usepackage{multirow}

% 1. Load the unicode-math package. This is the modern engine for math fonts.
\usepackage{unicode-math}
\usepackage{mwe}

% Custom operatore and contractions
\input{defs.tex}

% Graphics from the article repo
\graphicspath{{../conf_article/figs/}{../figures/13may_muon_neon/}}

% Define eqspace command (for text mode spacing)
\newcommand{\eqspace}{\par\vspace{0.2em}}

\defineauthorblock{%
  \renewcommand\cellalign{l}
  % We use tabular* spanning slightly less than full width to leave room for margins
  % Then center it to distribute margins equally
  \hspace*{\fill}%
  \begin{tabular*}{\dimexpr\linewidth-4em\relax}{
    @{}
    l
    @{\extracolsep{\fill}}
    l
    l
    l
    l
    l
    l
    @{}
  }
    % --- Column 1: First Author ---
    \makecell{
    \textbf{\Large Alexey Kravatskiy$^{1}$} \\[2ex]
    {\Large kravtskii.aiu@phystech.edu}
    } &

    % --- Column 2: Second Author ---
    \makecell{
    \textbf{\Large Ivan Kozyrev$^{1}$} \\[2ex]
    {\Large kozyrev.in@phystech.edu}
    } &

    % --- Column 3: Third Author ---
    \makecell{
    \textbf{\Large Nikolai Kozlov$^{1}$} \\[2ex]
    {\Large kozlov.na@phystech.edu}
    } &

    % --- Column 4: Fourth Author ---
    \makecell{
    \textbf{\Large Alexander Vinogradov$^{1}$} \\[2ex]
    {\Large vinogradov.am@phystech.edu}
    } &

    % --- Column 5: Fifth Author ---
    \makecell{
    \textbf{\Large Daniil Merkulov$^{1, 2, 3, 4}$} \\[2ex]
    {\Large daniil.merkulov@phystech.edu}
    } &

    % --- Column 6: Sixth Author ---
    \makecell{
    \textbf{\Large Ivan Oseledets$^{5, 2}$} \\[2ex]
    {\Large i.oseledets@skoltech.ru}
    } &

    % --- Column 7: All Institutions ---
    \makecell{
    \Large $^{1}$Moscow Institute of Physics and Technology\\
    \Large $^{2}$Skolkovo Institute of Science and Technology\\
    \Large $^{3}$HSE\; \Large $^{4}$AI4Science\; \Large $^{5}$AIRI
    }
  \end{tabular*}%
  \hspace*{\fill}%
}

\title{The Ky Fan Norms and Beyond: Dual Norms and Combinations for Matrix Optimization}

\begin{document}
\begin{frame}[fragile]
\begin{columns}[T]
\hspace{0.02\textwidth}% Left margin
%%%% First Column
\begin{column}{0.30\textwidth}
% Apply paragraph spacing inside column
\setlength{\parskip}{0.7em}
\setlength{\parindent}{0pt}
\Large

\textbf{\Huge\color{Zen}Introduction}\\[0.3em]

Training large language models requires optimizers more scalable and efficient than Adam/AdamW. The recent Muon~[2] optimizer shows great promise, sometimes doubling AdamW's efficiency, due to its unique design:
\begin{itemize}
  \item \textbf{\color{HazySummerEve}Matrix-Aware Design:} It is constructed specifically for weight matrices, not generic vectors.
  \item \textbf{\color{HazySummerEve}Principled Update Rule:} Muon's update is not merely heuristic. It is derived as the solution to a linear minimization problem constrained by the operator (spectral) norm, giving it a solid mathematical foundation.
\end{itemize}

Muon's success has inspired variants like Dion~[1] algorithm and frameworks of Scion and Gluon~[4]. We build on this work by asking a fundamental question:
\begin{attention}
In deriving Muon's update step, why constrain by the spectral norm? How would alternative norms affect performance and computational cost?
\end{attention}
To investigate this, we first analyze the link between Muon-like algorithms and their defining norms. We then propose \textbf{\color{HazySummerEve}F-Fanions}, a novel family of optimizers derived from exotic, mixed norms. Finally, we present an empirical comparison between Muon and F-Fanions across various numerical experiments.

\vspace{0.5em}
\textbf{\Huge\color{Zen}How Norms Shape the Update Step}\\[0.3em]

Many optimizers are defined by a norm-constrained Linear Minimization Oracle (LMO)~[3], which computes the update $\Delta\mX^t$ based on the gradient (possibly stochastic / with momentum) $\mG^t$:
\begin{equation*}
    \Delta X^t = \mX^{t+1} - \mX^t \in \eta \arg\min_{\norm{\mX} \leq 1} \inner{\mG^t}{\mX}\,.
\end{equation*}
The choice of norm dictates the algorithm. For matrix methods, the update often uses the Singular Value Decomposition (SVD) of the gradient, $\mG^t = \mU \mSigma \mV^\top$.

\centering
% Using 'tabular' now for better width control based on content
\begin{tabular}{clcc}
\toprule
\textbf{Case} & \textbf{Method} & \textbf{Constraining Norm} & \textbf{Update Formula} \\
\midrule
\multirow{3}{*}{\rotatebox[origin=c]{90}{\parbox{1.5cm}{\centering Vec.}}}
& Normalized SGD    & $\ell_2$          & $-\eta g / \norm{g}_2$ \\
& SignSGD           & $\ell_\infty$     & $-\eta\sign(g)$ \\
& Coordinate Descent    & $\ell_1$          & $-\eta e_{\mathrm{argmax}|g_i|}$ \\
\midrule
\multirow{3}{*}{\rotatebox[origin=c]{90}{\parbox{1.5cm}{\centering Mat.}}}
& Normalized SGD    & $\normf{\cdot}$   & $-\eta\mG^t / \normf{\mG^t}$ \\
& Muon     & $\norms{\cdot}$   & $-\eta\mU \mV^\top$ \\
& Dion (no feedback)   & $\norm{\cdot}_{KF-k}^\dag$ & $-\eta\mU_k \mV_k^\top$ \\
\bottomrule
\end{tabular}

\end{column}
\hspace{0.02\textwidth}% Space between columns
%%%% Second Column
\begin{column}{0.30\textwidth}
\setlength{\parskip}{0.7em}
\setlength{\parindent}{0pt}
\Large

We introduce a generalized norm that unifies and interpolates between these methods:
\begin{equation*}
    \norm{\mX}_{\text{gen}} = \left( \alpha \normkfk{\mX} + (1 - \alpha) \normf{\mX} \right)^\dag\,.
\end{equation*}

This defines the \textbf{\color{HazySummerEve}F-Fanions}, a new family of optimizers parameterized by $\alpha \in [0, 1]$ and $k$, with the update rule:
\begin{equation*}
    \Delta \mX^t = \eta \alpha \mU_k \mV_k^\top + \eta(1 - \alpha) \frac{\mG^t}{\normf{\mG^t}}\,.
\end{equation*}

This framework recovers existing algorithms as special cases:
\begin{itemize}
    \item \textbf{\color{HazySummerEve}Normalized SGD:} $\alpha = 0$.
    \item \textbf{\color{HazySummerEve}Dion} (no feedback): $\alpha = 1$, $k < \min\{m, n\}$.
    \item \textbf{\color{HazySummerEve}Muon:} $\alpha = 1$, $k = \min\{m, n\}$.
\end{itemize}
We also identify and analyze \textbf{\color{HazySummerEve}F-Neon}, a computationally efficient member of this family where $k=1$.

\vspace{0.5em}
\textbf{\Huge\color{Zen}Efficiently Computing the Update}\\[0.3em]

Computing the F-Fanion update is efficient at the extremes of the rank parameter $k$. We propose using the GPU-friendly \textbf{\color{HazySummerEve}Thick-Restarted Lanczos (TRLan)} method in two distinct ways:
\begin{itemize}
    \item \textbf{\color{HazySummerEve}For small $k$ (e.g., F-Neon):} TRLan is used directly to find the few leading singular vectors.
    \item \textbf{\color{HazySummerEve}For large $k$ (e.g., Muon):} TRLan is used to compute trailing singular values, they are substracted from a full-rank term, obtained through Newton-Shulz iterations.
\end{itemize}

The table below demonstrates the speed of the direct approach for finding $k$ leading singular vectors, from a $5000 \times 5000$ matrix.

{
\centering
\begin{tabular}{lccc}
\toprule
\textbf{Method} & \textbf{k=1} & \textbf{k=10} & \textbf{k=100} \\
\midrule
TRLan (our choice) & 0.18s & 0.47s & 1.96s \\
Randomized SVD (RSVD)       & 1.15s          & 19.4s          & 170.0s \\
Power Iterations            & 7.70s          & ---            & --- \\
\bottomrule
\end{tabular}
\par
\vspace{0.5em}
}

TRLan's dramatic speed advantage in this direct computation makes low-rank F-Fanions, like F-Neon, highly efficient.

\vspace{0.5em}
\textbf{\Huge\color{Zen}Modded NanoGPT Speedrun}\\[0.3em]

We benchmarked our method against Muon on the \texttt{modded\_nanogpt\_2024} speedrun, a time-constrained training challenge. Below are the results after 1750 steps,

{
\centering
\begin{tabular}{lccc}
\toprule
\textbf{Method} & \textbf{LR} & \textbf{Momentum} & \textbf{Final Loss} \\
\midrule
Muon (Baseline) & 0.05 & 0.95 & 3.279 (Pass) \\
F-Muon ($\alpha=0.5$) & 0.07 & 0.95 & 3.281 (Near Miss) \\
\bottomrule
\end{tabular}
\par
}

\end{column}
\hspace{0.02\textwidth}% Space between columns

%%%% Third column
\begin{column}{0.30\textwidth}
\setlength{\parskip}{0.7em}
\setlength{\parindent}{0pt}
\Large

\textbf{\Huge\color{Zen}CIFAR-10 Airbench Experiments}\\[0.3em]

We evaluate F-Muon against Muon by training a Convolutional Neural Network (CNN) on the CIFAR-10 Airbench. All results are averaged over multiple runs after 8 epochs.

\begin{columns}[T,totalwidth=\textwidth]
      \begin{column}{0.48\textwidth}
        \includegraphics[width=\linewidth]{muon_tuned_diff_alpha.pdf}

        {\centering With parameters from tuned version of Muon\par}
      \end{column}
      \begin{column}{0.48\textwidth}
        \includegraphics[width=\linewidth]{fmuon_tuned_diff_alpha.pdf}

        {\centering With parameters tuned for F-Muon, $\alpha = 0.5$\par}
      \end{column}
    \end{columns}\vspace{1em}

\begin{columns}[T,totalwidth=\textwidth]
  \begin{column}{0.47\textwidth}
    \setlength{\parskip}{0.7em}
\setlength{\parindent}{0pt}
\Large
    We can see that Properly tuned F-Muon can perform on par with Muon, achieving the same 94\% accuracy.

    Another Interesting observation here is that tuned F-Muon has much larger trust region then Muon, and thus makes larger step sizes (figure to the right).
  \end{column}
  \begin{column}{0.49\textwidth}
    \includegraphics[width=\linewidth]{fstardual_cifar.pdf}
    \centering
  \end{column}
\end{columns}

\vspace{0.5em}
\textbf{\Huge\color{Zen}Conclusion \& Outlook}\\[0.6em]

\begin{itemize}
    \item We introduced the \textbf{\color{HazySummerEve}F-Fanion framework}, unifying LMO-based optimizers like Muon and Dion via mixed norms.

    \item The choice of norm is a flexible design parameter, not a fixed rule. Our results show alternatives to the spectral norm can be equally effective.

    \item This work opens the core question of how to theoretically guide norm selection. Future work could also explore adapting the norm (via $\alpha$ and $k$) dynamically during training.
\end{itemize}

\vspace{0.5em}
\textbf{\Huge\color{Zen}Key References}\\[0.6em]

\begin{itemize}
   \large
    \setlength{\itemsep}{0pt} % Remove spacing between items
    \item[{[1]}] Ahn, K. \& Xu, B. *Dion*. arXiv:2504.05295, 2025.
    \item[{[2]}] Bernstein, J. *Deriving muon*. 2025. URL: jeremybernste.in/writing/deriving-muon
    \item[{[3]}] Bernstein, J. \& Newhouse, L. *Old optimizer, new norm*. arXiv:2409.20325, 2024.
    \item[{[4]}] Riabinin, A. et al. *Gluon: Making muon \& scion great again!* arXiv:2505.13416, 2025.
\end{itemize}

\end{column}
\hspace*{0.02\textwidth}% Right margin (using \hspace* to prevent it from being dropped)
\end{columns}

\end{frame}
\end{document}

### 2025-10-13 poster/main.tex

@@
     \makecell{
-    \Large $^{1}$Moscow Institute of Physics and Technology\\
-    \Large $^{2}$Skolkovo Institute of Science and Technology\\
-    \Large $^{3}$HSE\; \Large $^{4}$AI4Science\; \Large $^{5}$AIRI
+    \Large $^{1}$MIPT\; \Large $^{2}$Skoltech\\
+    \Large $^{3}$HSE\; \Large $^{4}$AI4Science\\
+    \Large $^{5}$AIRI
     }
@@
 \begin{itemize}
-  \item \textbf{\color{HazySummerEve}Matrix-Aware Design:} It is constructed specifically for weight matrices, not generic vectors.
+  \item \textbf{\color{HazySummerEve}Matrix-Aware Design:} It is constructed specifically for weight matrices rather than generic vectors.
   \item \textbf{\color{HazySummerEve}Principled Update Rule:} Muon's update is not merely heuristic.
@@
 We then propose \textbf{\color{HazySummerEve}F-Fanions}, a novel family of optimizers derived from exotic, mixed norms.
-Finally, we present an empirical comparison between Muon and F-Fanions across various numerical experiments.
+Finally, we present an empirical comparison between Muon and F-Fanions.
 
@@
 
+{
 \centering
@@
 \toprule
-\textbf{Case} & \textbf{Method} & \textbf{Constraining Norm} & \textbf{Update Formula} \\
+\textbf{Case} & \textbf{Method} & \textbf{Norm} & \textbf{Update Formula} \\
 \midrule
 \multirow{3}{*}{\rotatebox[origin=c]{90}{\parbox{1.5cm}{\centering Vec.}}}
-& Normalized SGD    & $\ell_2$          & $-\eta g / \norm{g}_2$ \\
-& SignSGD           & $\ell_\infty$     & $-\eta\sign(g)$ \\
-& Coordinate Descent    & $\ell_1$          & $-\eta e_{\mathrm{argmax}|g_i|}$ \\
+& Normalized SGD    & $\ell_2$          & $-\eta g^t / \norm{g^t}_2$ \\
+& SignSGD           & $\ell_\infty$     & $-\eta\sign(g^t)$ \\
+& Gauss-Southwell CD    & $\ell_1$      & $-\eta \sum_{i \in \argmax|g_i^t|} \sign(g_i^t)e_i$ \\
 \midrule
@@
 & Muon     & $\norms{\cdot}$   & $-\eta\mU \mV^\top$ \\
-& Dion (no feedback)   & $\norm{\cdot}_{KF-k}^\dag$ & $-\eta\mU_k \mV_k^\top$ \\
+& Dion (without EF)   & $\norm{\cdot}_{KF-k}^\dag$ & $-\eta\mU_k \mV_k^\top$ \\
 \bottomrule
-\end{tabular}
+\end{tabular}\eqspace
+}
+Here $\norm{\cdot}_{KF-k}$ denotes Ky Fan $k$ norm, $\norm{\cdot}^\dag$ is the dual norm.
 
@@
 \begin{equation*}
-    \Delta \mX^t = \eta \alpha \mU_k \mV_k^\top + \eta(1 - \alpha) \frac{\mG^t}{\normf{\mG^t}}\,.
+    \Delta \mX^t = -\eta \alpha \mU_k \mV_k^\top - \eta(1 - \alpha) \frac{\mG^t}{\normf{\mG^t}}\,.
 \end{equation*}
@@
     \item \textbf{\color{HazySummerEve}Normalized SGD:} $\alpha = 0$.
-    \item \textbf{\color{HazySummerEve}Dion} (no feedback): $\alpha = 1$, $k < \min\{m, n\}$.
+    \item \textbf{\color{HazySummerEve}Dion} (without error feedback): $\alpha = 1$, $k < \min\{m, n\}$.
     \item \textbf{\color{HazySummerEve}Muon:} $\alpha = 1$, $k = \min\{m, n\}$.
@@
 
-We benchmarked our method against Muon on the \texttt{modded\_nanogpt\_2024} speedrun, a time-constrained training challenge. Below are the results after 1750 steps,
+We benchmarked our method against Muon on the \texttt{modded nanogpt 2024} speedrun, a pretraining of a real LLM. Below are the results after 1750 steps,
 
@@
 
-    Another Interesting observation here is that tuned F-Muon has much larger trust region then Muon, and thus makes larger step sizes (figure to the right).
+    Another Interesting observation here is that tuned F-Muon has much larger trust region than Muon, and thus makes larger step sizes (figure to the right).
   \end{column}
@@
 \begin{itemize}
-    \item We introduced the \textbf{\color{HazySummerEve}F-Fanion framework}, unifying LMO-based optimizers like Muon and Dion via mixed norms.
+    \item We introduced the \textbf{\color{HazySummerEve}F-Fanion framework}, unifying LMO-based optimizers like Muon and Normalized SGD via mixed norms.
 
@@
     \setlength{\itemsep}{0pt} % Remove spacing between items
-    \item[{[1]}] Ahn, K. \& Xu, B. *Dion*. arXiv:2504.05295, 2025.
-    \item[{[2]}] Bernstein, J. *Deriving muon*. 2025. URL: jeremybernste.in/writing/deriving-muon
-    \item[{[3]}] Bernstein, J. \& Newhouse, L. *Old optimizer, new norm*. arXiv:2409.20325, 2024.
-    \item[{[4]}] Riabinin, A. et al. *Gluon: Making muon \& scion great again!* arXiv:2505.13416, 2025.
+    \item[{[1]}] Ahn, K. \& Xu, B. \textit{Dion: A Communication-Efficient Optimizer for Large Models}. arXiv:2504.05295, 2025.
+    \item[{[2]}] Bernstein, J. \textit{Deriving Muon}. 2025. URL: jeremybernste.in/writing/deriving-muon
+    \item[{[3]}] Bernstein, J. \& Newhouse, L. \textit{Old Optimizer, New Norm: An Anthology}. arXiv:2409.20325, 2024.
+    \item[{[4]}] Riabinin, A. et al. \textit{Gluon: Making Muon \& Scion Great Again!} arXiv:2505.13416, 2025.
 \end{itemize}

### 2025-10-17 poster/main.tex
Added short slides

@@
 
-Training large language models requires optimizers more scalable and efficient than Adam/AdamW.
-The recent Muon~[2] optimizer shows great promise, sometimes doubling AdamW's efficiency, due to its unique design:
+LLM training requires more scalable and efficient optimizers. The recent Muon~[2] sometimes doubles AdamW's efficiency due to Muon's unique design :
 \begin{itemize}
-  \item \textbf{\color{HazySummerEve}Matrix-Aware Design:} It is constructed specifically for weight matrices rather than generic vectors.
-  \item \textbf{\color{HazySummerEve}Principled Update Rule:} Muon's update is not merely heuristic.
+  \item \textbf{\color{HazySummerEve}Matrix-Aware Design:} Muon is constructed specifically for weight matrices rather than for generic vectors.
+  \item \textbf{\color{HazySummerEve}Theory-Based Update Rule:} Muon's update is not a mere heuristic.
 It is derived as the solution to a linear minimization problem constrained by the operator (spectral) norm, giving it a solid mathematical foundation.
@@
 \begin{attention}
-In deriving Muon's update step, why constrain by the spectral norm? How would alternative norms affect performance and computational cost?
+Why should the choice of a matrix norm in the Muon’s update derivation be restricted to the spectral norm? How might other matrix norms influence the algorithm’s performance and computational cost?
 \end{attention}
@@
 \vspace{0.5em}
-\textbf{\Huge\color{Zen}How Norms Shape the Update Step}\\[0.3em]
+\textbf{\Huge\color{Zen}How Norms Shape the Update}\\[0.3em]
 
@@
 \midrule
-\multirow{3}{*}{\rotatebox[origin=c]{90}{\parbox{1.5cm}{\centering Mat.}}}
+\multirow{4}{*}{\rotatebox[origin=c]{90}{\parbox{1.5cm}{\centering Mat.}}}
 & Normalized SGD    & $\normf{\cdot}$   & $-\eta\mG^t / \normf{\mG^t}$ \\
 & Muon     & $\norms{\cdot}$   & $-\eta\mU \mV^\top$ \\
+& Neon (ours)   & $\norm{\cdot}_{nuc}$ $ & $-\eta u_1 v_1^\top$ \\
 & Dion (without EF)   & $\norm{\cdot}_{KF-k}^\dag$ & $-\eta\mU_k \mV_k^\top$ \\
+
 \bottomrule
@@
 \end{itemize}
-We also identify and analyze \textbf{\color{HazySummerEve}F-Neon}, a computationally efficient member of this family where $k=1$.
+We also identify and analyze \textbf{\color{HazySummerEve}F-Neon}, a computationally efficient member of this family with $k=1$.
 
@@
 
-We benchmarked our method against Muon on the \texttt{modded nanogpt 2024} speedrun, a pretraining of a real LLM. Below are the results after 1750 steps,
+We benchmarked our method against Muon and NSGD on the \texttt{modded nanogpt 2024} speedrun, which is pretraining of a real LLM.
+Below are the results after 1,750 steps.
+The objective was to achieve a loss lower than 3.28.
 
@@
 \midrule
-Muon (Baseline) & 0.05 & 0.95 & 3.279 (Pass) \\
-F-Muon ($\alpha=0.5$) & 0.07 & 0.95 & 3.281 (Near Miss) \\
+Muon (Baseline)         & 0.05 & 0.95 & 3.279  \\
+F-Muon ($\alpha=0.5$)   & 0.07 & 0.95 & 3.281 \\
+NSGD                    & 0.07 & 0.96 & 3.4651 \\
+F-Muon ($\alpha=0.5$)   & 0.07 & 0.96 & 3.2824 \\
 \bottomrule
@@
 
-We evaluate F-Muon against Muon by training a Convolutional Neural Network (CNN) on the CIFAR-10 Airbench. All results are averaged over multiple runs after 8 epochs.
+We evaluate F-Muon against Muon by training a Convolutional Neural Network (CNN) on the CIFAR-10 Airbench.
+All results are averaged over multiple runs and reported after 8 epochs with batch size = 2,000.
 
@@
 \Large
-    We can see that Properly tuned F-Muon can perform on par with Muon, achieving the same 94\% accuracy.
+    We observe that a properly tuned F-Muon performs on par with Muon and achieves the same 94.01\% accuracy.
 
-    Another Interesting observation here is that tuned F-Muon has much larger trust region than Muon, and thus makes larger step sizes (figure to the right).
+    Moreover, the tuned F-Muon has a significantly larger trust region than Muon, and the learning rate is so high that Muon yields lower test accuracy when trained with it (see the figure on the right).
   \end{column}
@@
 \begin{itemize}
-    \item We introduced the \textbf{\color{HazySummerEve}F-Fanion framework}, unifying LMO-based optimizers like Muon and Normalized SGD via mixed norms.
+    \item We introduced the \textbf{\color{HazySummerEve}F-Fanions} that unify LMO-based optimizers like Muon and Normalized SGD via mixed norms.
 
-    \item The choice of norm is a flexible design parameter, not a fixed rule. Our results show alternatives to the spectral norm can be equally effective.
+    \item The choice of a norm is a flexible design parameter rather than a fixed rule. Our results show that other norms may be as efficient as the spectral norm.
 
-    \item This work opens the core question of how to theoretically guide norm selection. Future work could also explore adapting the norm (via $\alpha$ and $k$) dynamically during training.
+    \item Our work poses a crucial question of how to theoretically guide norm selection. Future work could also explore adapting the norm (via $\alpha$ and $k$) dynamically during training.
 \end{itemize}

### 2025-10-17 slides/neon_short_slides.tex
Added short slides

--- 2025-10-10 slides/neon_slides.tex
@@
 %\usepackage[authoryear]{natbib}
+
+% Absolute positioning for overlaying content on the title slide
+\usepackage[absolute,overlay]{textpos}
 
@@
   \titlepage
+  % Right-side figure on the title slide
+  \begin{textblock*}{0.2\paperwidth}(0.75\paperwidth,0.6\paperheight)
+    \includegraphics[width=\linewidth]{fstardual_cifar.pdf}
+  \end{textblock*}
+  \begin{textblock*}{0.22\paperwidth}(0.05\paperwidth,0.55\paperheight)
+    \includegraphics[width=\linewidth]{KyFanDual.pdf}
+  \end{textblock*}
 \end{frame}
@@
 %-------------------------------------------------------------------------------------
-\begin{frame}{Linear Minimization Oracle (LMO) and Trust Region}
+\begin{frame}{Linear Minimization Oracle (LMO)}
 % after omitting gamma_k, we obtain:
-Let us equip $\Rmn$ with a norm $\norm{\cdot}$. Its dual is $\norm{\mX}^\dagger = \sup_{\norm{\mX'}\leq 1} \<\mX,\mX'>$. $\<\cdot, \cdot>$ is a Frobenius product.
+Let us equip $\Rmn$ with a norm $\norm{\cdot}$. Its dual is $\norm{\mX}^\dagger = \sup_{\norm{\mX'}\leq 1} \<\mX,\mX'>$. $\<\cdot, \cdot>$ is the Frobenius product.
 \vspace{0.4em}
 
-Both LMO and Trust Region lead to the update
+LMO leads to the update
 $$\mX^{t+1} = \mX^{t} - \eta \arg\max_{\mX \in \cB_1} \<\mM^t, \mX> = \mX^{t} - \eta \{\Delta \in \cB_1 \mid \<\mM^t, \Delta> = \norm{\mM^t}^\dagger\}$$
@@
 
-\textbf{Recipe}: It means we seek $\Delta$ from the 1-norm ball that delivers $\<\mM^t, \Delta> = \norm{\mM^t}^\dagger$. We will often return to the SVD $\mM^t = \mU \Sigma \mV^\top$.
+\textbf{Recipe}: It means we seek $\Delta$ from the 1-norm ball that delivers $\<\mM^t, \Delta> = \norm{\mM^t}^\dagger$
+\vspace{0.4em}
+
+We will often utilize the SVD of $\mM^t$: $\mM^t = \mU \Sigma \mV^\top$.
 \end{frame}
-%-------------------------------------------------------------------------------------
-\begin{frame}{Frobenius $\normf{\mM^{k}}$ and Normalized SGD}
-
-\begin{block}{Deriving NSGD by Recipe} $\normf{\mM^t}^\dagger = \normf{\mM^t}$, and $\Delta = \frac{\mM^t}{\normf{\mM^t}}$ with $\normf{\Delta} = 1$ delivers it. Hence,
-$$\mX^{t+1} = \mX^{t} - \eta \frac{\mM^t}{\normf{\mM^t}}$$
-\end{block}
-\end{frame}
-%-------------------------------------------------------------------------------------
-\begin{frame}{Spectral $\norms{\mM^{k}}$ and Muon. Nuclear $\normn{\mM^k}$ and Neon.}
-\begin{block}{Deriving Muon by Recipe} $\norms{\mM^t}^\dagger = \normn{\mM^t}$, and $\Delta = \mU\mV^\top$ delivers it: $\norms{\Delta}$ = 1 and
-
-$$\<\mU\Sigma\mV^\top, \mU\mV^\top> = \tr(\mV \Sigma \mU^\top \mU \mV^\top) = \tr\Sigma = \normn{\mM^t}$$
-
-Hence,
-$$\mX^{t+1} = \mX^{t} - \eta \mU \mV^\top$$
-
-\end{block}
-
-\begin{block}{Deriving Neon by Recipe} $\normn{\mM^t}^\dagger = \norms{\mM^t}$, and $\Delta = u_1 v_1^\top$ delivers it: $\normn{\Delta}$ = 1 and
-
-$$\<\mU\Sigma\mV^\top, u_1 v_1^\top> = \tr(\mV \Sigma \mU^\top u_1 v_1^\top) = \tr\diag({\sigma_1, 0, \dots, 0}) = \sigma_1 = \norms{\mM^t}$$
-
-Hence,
-$$\mX^{t+1} = \mX^{t} - \eta u_1 v_1^\top$$
-\end{block}
-\end{frame}
-
-%-------------------------------------------------------------------------------------
 \begin{frame}{Of Matrix and Vector Algorithms}
@@
     \hline
-    Gauss-Southwell Coordinate Descent & Ball in $l_1$ & $-\eta \{i: g_i \geq g_k \forall k\}$ & \citep[p.19]{shi2016primer}\\
+    Gauss-Southwell Coordinate Descent & Ball in $l_1$ & $-\eta \sum_{i \in \argmax|g_i^t|} \sign(g_i^t)e_i$ & \citep[p.19]{shi2016primer}\\
     \hdashline 
@@
 
-    \faQuestionCircle \space What will we get if we try $\normkfk{\mM}:=\sum_{i=1}^r{\sigma_i}$?
+    \faQuestionCircle \space What will we obtain if we try $\normkfk{\mM}:=\sum_{i=1}^r{\sigma_i}$?
     \end{frame}
@@
     \end{block}
+\end{frame}
+%----------------------------------------
+\begin{frame}{Geometric intuition: Ky Fan norms}
+  \centering
+  \begin{minipage}{0.46\textwidth}
+      \includegraphics[height=0.75\textheight,keepaspectratio]{KyFan.pdf}
+  \end{minipage}\hfill
+  \begin{minipage}{0.46\textwidth}
+      \includegraphics[height=0.75\textheight,keepaspectratio]{KyFanDual.pdf}
+  \end{minipage}
+
+  \vspace{0.3em}
+  \footnotesize Ky Fan rank-2 ball and its dual in a 3D singular-value space
   \end{frame}
@@
 
-Let us consider the convex combination of the Ky Fan rank-$k$ norm and the Frobenius norm, which we call F-KF-k-norm: $$\norm{\cdot}_{\mathrm{F-KF-k}} = \alpha \normkfk{\cdot} + (1-\alpha)\normf{\cdot}, \alpha \in [0,1]$$
+Let us consider a convex combination of the Ky Fan rank-$k$ norm and the Frobenius norm, which we call the F-KF-k-norm: $$\norm{\cdot}_{\mathrm{F-KF-k}} = \alpha \normkfk{\cdot} + (1-\alpha)\normf{\cdot}, \alpha \in [0,1]$$
 
@@
 \begin{frame}{F-Muon and F-Neon}
-  Convex combinations with Frobenius norm:
+  Convex combinations with the Frobenius norm:
   \[
@@
 \end{frame}
-%-----------------------------------------------------------------------------------------------
-\begin{frame}{Geometric intuition: Dual balls}
-    \centering
-    \begin{minipage}{0.48\textwidth}
-        \includegraphics[height=0.75\textheight,keepaspectratio]{fstardualball.pdf}
-    \end{minipage}\hfill
-    \begin{minipage}{0.48\textwidth}
-        \includegraphics[height=0.75\textheight,keepaspectratio]{ftwodualball.pdf}
-    \end{minipage}
-
-    \vspace{0.3em}
-    \footnotesize Dual balls for F-Muon and F-Neon across \(\alpha\) in a 2D singular-value space
-    \end{frame}
 %---------------------------------------
@@
         \centering
-        \scriptsize With params tuned for F-Muon
+        \scriptsize With params tuned for $\alpha=0.5$ F-Muon
       \end{column}
@@
 \end{frame}
-%----------------------------------------------------------------------------------------------
-\begin{frame}{Formal assumptions}
-  L-smoothness:
-  \begin{aequation}\label{eq:L}
-    \norm{\nabla f(\mX) - \nabla f(\mX')}^\dagger \leq L \norm{\mX-\mX'}
-    \quad\text{for all}\;
-    \mX, \mX' \in \Rmn,
-    \end{aequation}
-
-  Bounded variance:
-    \begin{aequation}\label{eq:variance}
-      \Ed{\xi \sim \cD}{g(\mX;\xi)} = \nabla f(\mX)
-      \quad\text{and}\quad
-      \Ed{\xi \sim \cD}{\sqnf{g(\mX;\xi) - \nabla f(\mX)}} \leq \sigma^2
-      \quad\text{for all}\;
-      \mX \in \Rmn,
-      \end{aequation}
-
-  Bounded norm:
-      \begin{aequation}\label{eq:norm}
-        \norm{\mX}^\dagger \leq \rho\cdot\normf{\mX}
-        \quad\text{for all}\;
-        \mX \in \Rmn,
-        \end{aequation}
-
-  From Marchenko-Pastur law, $\normf{\cdot} \sim \frac{\sqrt{n}}{2}\norms{\cdot}$ and $\normn{\cdot} \sim \frac{n}{2}\norms{\cdot}$ for large $n$.
-
-\end{frame}
-\begin{frame}{Bounds from \citep{kovalev2025understanding}}
-  \begin{lemma}\label{lemma:stoch_tr}
-    To reach the precision $\E{\min_{k=1\ldots K} \norm{\nabla f(\mX_k)}^\dagger} \leq \epsilon$ under the assumptions \Cref{eq:L}, \Cref{eq:variance}, \Cref{eq:norm}, it is sufficient to choose the parameters as follows:
-    \begin{align}
-        \eta &= \cO\left(\min\left\{\frac{\epsilon}{L}, \frac{\epsilon^3}{\rho^2\sigma^2L}\right\}\right),
-        \qquad
-        \alpha = \cO\left(\min\left\{1, \frac{\epsilon^2}{\rho^2\sigma^2}\right\}\right),
-        \\
-        \label{eq:str_K_nonconvex}
-        K &= \cO\left(\max\left\{
-            \frac{\rho\sigma}{\epsilon},
-            \frac{\rho^3\sigma^3}{\epsilon^3},
-            \frac{L\Delta_0}{\epsilon^2},
-            \frac{L\Delta_0\rho^2\sigma^2}{\epsilon^4}
-        \right\}\right).
-    \end{align}
-    \end{lemma}
-
-\end{frame}
 %----------------------------------------
-\begin{frame}{Random linear least squares: Loss and the Spectral Norm of the gradient}
+\begin{frame}{Random linear least squares: Loss and the Spectral Norm of the Gradient}
   \begin{columns}[T,totalwidth=\textwidth]
@@
 \end{frame}
-
 %----------------------------------------
-\begin{frame}{Geometric intuition: Ky Fan norms}
-  \centering
-  \begin{minipage}{0.46\textwidth}
-      \includegraphics[height=0.75\textheight,keepaspectratio]{KyFan.pdf}
-  \end{minipage}\hfill
-  \begin{minipage}{0.46\textwidth}
-      \includegraphics[height=0.75\textheight,keepaspectratio]{KyFanDual.pdf}
-  \end{minipage}
-
-  \vspace{0.3em}
-  \footnotesize Ky Fan rank-2 ball and its dual in a 3D singular-value space
-  \end{frame}
-%----------------------------------------
-\begin{frame}{Conclusions}
+\begin{frame}{Conclusion}
   \begin{itemize}
     \item Muon and NSGD are special cases of F-Fanions
-    \item Mixture of norm-based updates is a norm-based update!
-    \item No existing bounds describe the superiority of the spectral norm
+    \item A combination of norm-based updates is a norm-based update
+    \item No existing bound describes the superiority of the spectral norm
   \end{itemize}
@@
   \vspace{2em}
-  \textbf{For future experiments}: Exploration of matrix and vector algorithms correspondence
+  \textbf{For future experiments}: Exploration of the matrix and vector algorithms correspondence
 \end{frame}

### 2025-11-02 conf_article/neon.tex
Finished the introduction rewrite

@@
 \replaced[id=IK, comment={This rewrite clarifies the sentence structure, making it clear that the norms (not the algorithms) are being combined.}]{Subsequently, we construct a second family of algorithms, dubbed {\it F-Fanions}, based on the duals of convex combinations of Ky Fan k-norms and the Frobenius norm.
-One prominent member of this family is {\it F-Muon}.}{Then, we consider their convex combinations with the Frobenius norm and corresponding duals to develop a new family of algorithms we name {\it F-Fanions}, one of them being {\it F-Muon}.} We complement our theoretical analysis with an extensive empirical study of the algorithms across a wide range of tasks and settings, from which it follows that properly-tuned F-Muon is almost on par with Muon, \replaced[id=IK, comment={Alex, clarify this please.
-It's a very non-obvious and non-clear conclusion, especially when knowing that Muon itself only provides an approximate answer to the lmo.}]{which questions the necessity of finding the exact solution to the problem in linear optimization oracle framework.}
+One prominent member of this family is {\it F-Muason}.}{Then, we consider their convex combinations with the Frobenius norm and corresponding duals to develop a new family of algorithms we name {\it F-Fanions}, one of them being {\it F-Muon}.} We complement our theoretical analysis with an extensive empirical study of the algorithms across a wide range of tasks and settings, from which it follows that properly-tuned F-Muon is almost on par with Muon, \replaced[id=IK, comment={Alex, clarify this please.
+It's a very non-obvious and non-clear conclusion, especially when knowing that Muon itself only provides an approximate answer to the LMO.
+I provided an alternative phrase.}]{which questions the necessity of finding the exact solution to the problem in linear optimization oracle framework.}{which shows the flexibility in the choice of the norm and opens possibilities for easier-to-compute variants.}
 
@@
 The innovation wasn't just 'using' a norm, but deriving the update from a principled, constrained optimization problem.
-I think this better reflects the contribution of the Muon paper.}]{From a theoretical perspective, a key innovation of Muon was its principled derivation of the update rule, which emerges as the solution to an optimization problem constrained by the spectral norm \citep{bernstein2025deriving}}{From the perspective of theory, the main innovation of Muon has been an intentional usage of matrix norms, i.e.
+I think this better reflects the contribution of the Muon paper.}]{From a theoretical perspective, a key innovation of Muon was its principled derivation of the update rule, which emerges as the solution to an optimization problem constrained by the RMS-to-RMS norm (scaled version of spectral norm) \citep{bernstein2025deriving}}{From the perspective of theory, the main innovation of Muon has been an intentional usage of matrix norms, i.e.
 the spectral norm, to derive the algorithm's update}.
 
+\replaced[id = IK, comment={Subsequent paragraphs in the introduction are very heavily modified.}]{}{}
+
 Motivated by the success of Muon, many generalizations and variations of it were proposed.
-Among the notable ones are Scion \citep{}, Deon \citep{} and Gluon \citep{}.
-One question still persists however: Why
-
-Based upon recent theoretical advances that explain some theory behind Muon, Scion and Gluon \citep{bernstein2025deriving,kovalev2025understanding,pethick2025training,riabinin2025gluon}, we explore application of other matrix norms to optimization of functions of matrices.
-As it has been done with Muon, we stipulate that our algorithms' updates be fast to compute.
-
-\replaced[id = IK, comment={Subsequent paragraphs in the introduction are almost completely new.}]{}{}
+Among the notable ones are Scion \citep{pethick2025training}, Dion \citep{ahn2025dioncommunicationefficientoptimizerlarge} and Gluon \citep{riabinin2025gluon}.
+Those researches target Muon's efficiency and establish convergence bounds.
+One central question, however, remains unanswered:
+
+\textit{In deriving Muon's update step, why constrain by the spectral norm? How would alternative norms affect performance and computational cost?}
+
+In this article, we tackle this question by first showing the connection between matrix norms and corresponding existing algorithms and discuss the theoretic bounds derived for those algorithms.
+We then leverage the family of norms dual to Ky Fan k-norms to derive a new class of algorithms with low-rank updates, which we call Fanions.
+Subsequently, we create a second, hybrid family named F-Fanions by constructing convex combinations of the Ky Fan norms with the Frobenius norm and taking dual of that composite norm.
+Working within the linear minimization oracle (\replaced[id = IK, comment={I think it's better this way.}]{LMO}{lmo}) framework we derive the explicit update formulas for both algorithm families.
+As it was done with Muon, we stipulate our algorithms to be fast to approximate, for which we utilize the Lanczos algorithm as described in \cref{sec:matrix-side}.
+
+We then compare the performance of the algorithm families across various benchmarks (\cref{sec:experiments}):
+\begin{itemize}
+    \item Synthetic least squares experiment,
+    \item CIFAR-10 airbench,
+    \item Pre-training NanoGPT on FineWeb dataset.
+\end{itemize}
+
+Our experiments reveal important insights about the role of matrix norms in optimization.
+On a synthetic least squares problem, we observe a striking discrepancy: while some algorithms converge slowly in terms of loss, they may converge quickly in terms of gradient norm, and vice versa.
+This suggests that existing theoretical guarantees may not fully explain practical algorithm performance.
+
+Most notably, our experiments on real-world tasks demonstrate that the choice of underlying matrix norm is remarkably flexible.
+On CIFAR-10 airbench, properly-tuned F-Muon achieves $94.01 \pm 0.14\%$ accuracy, essentially matching Muon's $94.00 \pm 0.13\%$ performance.
+On NanoGPT pre-training, F-Muon achieves 3.281 cross-entropy loss, only marginally worse than Muon's 3.279.
+These results show that Muon-like algorithms can maintain competitive performance even when the underlying norm constraint is significantly modified, answering affirmatively the central question posed above.
+This flexibility suggests potential for developing easier-to-compute variants of successful algorithms like Muon.
 
@@
 \section{Problem Statement}
+
 \subsection{Function of a matrix}
@@
 
-\section{Experiments}
+\section{Experiments}\label{sec:experiments}
     \subsection{Randomized Linear Least Squares}

### 2025-11-02 conf_article/neon.tex
revised the first section

@@
 
-%In this article, \replaced[id=IK, comment={I think this is much clearer to introduce all algorithms like that, the paragraphs where one of every two words is a ***on are confusing.}]{we generalize the spectral-norm-based approach by first leveraging the family of Ky Fan k-norms to derive a new class of algorithms, which we call Fanions.
-Subsequently, we create a second, hybrid family named F-Fanions, by constructing convex combinations of the Ky Fan norms with the Frobenius norm and taking dual of that composite norm.
-Working within the linear minimization oracle (lmo) framework~--- which is equivalent (up to a factor) to the trust region approach~--- we derive the explicit update formulas for both algorithm families.
-Later in \cref{sec:matrix-side} we explain how their updates can be approximated with the same $\cO(m n \min\{m,n\})$ complexity as Muon updates are computed via Newton-Schulz iterations.}{we focus on the two most common matrix norms akin to the spectral norm, namely the nuclear norm and the Frobenius norm.
-Working in the linear minimization oracle (lmo) framework, which is equivalent to a factor to the trust region approach and the steepest descent under norm constraint, we derive Neon, our algorithm based on the nuclear norm.
-In the section {\it Matrix side of the updates}, we explain how Neon updates can be computed asymptotically with the same $\cO(m n \min\{m, n\})$ number of FLOPs as Muon updates are computed via Newton-Schulz iterations.}
-
-%Noticing that Neon and Muon are diametrical in terms of the rank of the update matrix, we bridge the space by considering the dual to the Ky Fan $k$-norm, the primal norm being the sum of the $k$ largest singular values of the matrix.
-The resulting update turns out to be Dion \citet{ahn2025dioncommunicationefficientoptimizerlarge} without error feedback.
-
-%To make Neon and Dion without error feedback more robust, we "regularize" them by Normalized SGD (NSGD), which is derived in lmo with the Frobenius norm.
-We do this in the same lmo approach by considering a norm that is dual to the convex combination of the Frobenius norm and the Ky Fan k-norm.
-So we derive the algorithms we name F-Neon and F-Dion respectively.
-The same procedure we follow with Muon to obtain F-Muon.
-
-%Having faced the array of different Muon-like optimizers, which, according to the upper bounds from \citet{kovalev2025understanding,riabinin2025gluon}, must converge similarly, we painstakingly compare them on a synthetic linear least squares problem with a known Lipschitz constant.
-The efforts result in comparison of the algorithms by their convergence not in terms of the dual norms of the gradient, but in terms of the common spectral norm, which may strongly differ, especially in large matrices, from the initial dual norms.
-Thus, we compare the algorithms in a unified fashion and find that sometimes if an algorithm converges fast by the loss, it does not imply its fast convergence by the norm of the gradient, and vice versa.
-
-%Finally, observing the poor performace of Neon and Dion without error feedback, which negatively impacts F-Neon and F-Dion as well, we test Muon and F-Muon on two deep-learning tasks, training convolutional network on CIFAR-10 \citep{cifar2023airbench} and pre-training NanoGPT on FineWeb \citep{modded_nanogpt_2024}.
-The results support the supremacy of Muon, but the most striking outcome of the tests is that F-Muon, only half of which is Muon, is on par with Muon on training the convolutional network and only slighter worse that Muon on NanoGPT pre-training, while NSGD is far from that.
-The case of F-Muon answers in the affirmative to the question of possibility to perturb the norm that underpins the algorithm without sacrificing its performace.
-We envision that this effect will be used in the future to constuct easier to compute approximations of empirically optimal algorithms like Muon.
-
-\section{Problem Statement}
-
-\subsection{Function of a matrix}
-    We consider the problem of minimizing a differentiable matrix function $F(\cdot)\colon \Rmn \to \R$
+\section{\replaced[id = IK, comment={I think the old one was a bit misleading.
+Merged what previously had been 2 subsections into this one and improved the notation.
+Also removed the thing about trust region, it'll be referenced in a succeeding sections.}]{Update Step as Constrained Optimization}{Problem Statement}}
+
+Training a neural network involves optimizing a function of weight matrices. This naturally leads us to consider the problem of minimizing a differentiable matrix function: 
 \begin{equation}\label{eq:mat}
-        \min_{\mX \in \Rmn} f(\mX).
-    \end{equation}
-
-    We equip the matrix space $\Rmn$ with a norm $\norm{\cdot}\colon \Rmn \to \R_+$, which possibly does not coincide with the Frobenius norm $\normf{\cdot}$.
-$\norm{\cdot}^\dagger\colon \Rmn \to \R_+$ is the dual norm associated with $\norm{\cdot}$, i.e., $\norm{\mX}^\dagger = \sup_{\norm{\mX'}\leq 1} \<\mX,\mX'>$ for all $\mX \in \Rmn$.
-
-\subsection{Linear Minimization Oracle and Trust Region}
-    We analyze the problem from the perspective of linear minimization oracle (lmo) and unconstrained stochastic conditional gradient descent (uSCG) \citep{pethick2025training}. lmo is defined as:
+    F(\cdot)\colon \Rmn \to \R\,,\qquad F(\mX) \to \min_{\mX \in \Rmn}\,.
+\end{equation}
+We equip the matrix space $\Rmn$ with a standard dot product $\<\cdot, \cdot> \to \R$ and a norm $\norm{\cdot}\colon \Rmn \to \R_+$, which possibly does not coincide with the Frobenius norm $\normf{\cdot}$.
+The dual norm $\norm{\cdot}^\dagger\colon \Rmn \to \R_+$ associated with $\norm{\cdot}$ is defined as
+\begin{equation}\label{eq:dual_norm}
+    \norm{\mG}^\dagger = \sup_{\mD \in \Rmn\,,\norm{\mD}\leq 1} \<\mG,\mD>\,.
+\end{equation}
+
+One approach to optimizing such functions is to employ an iterative algorithm based on Linear Minimization Oracle (LMO):
 \begin{equation}\label{eq:lmo}
-        \mathrm{lmo}(\mS) \in \arg\min_{\mX \in \cS} \<\mS, \mX>,
-    \end{equation}
-    where $\cS$ is some set. We are interested in the case when $\cS$ is a ball in our $\norm{\cdot}$ norm:
+    \mathrm{LMO}(\mG) \in \argmin_{\mD \in \cS} \<\mG, \mD>\,,
+\end{equation}
+where $\mG$ is typically a gradient (possibly with momentum) of $F$ and $\cS \subset \Rmn$ is some set.
+The update rule used in Muon optimizer is uSCG \citep{pethick2025training}.
+In the most general case, which involves momentum and stochastic gradient $g(\mX^k, \xi^k)$, it can be written as
+\begin{equation}\label{eq:usgd_update}
+    \begin{aligned}
+    \mM^{k} &= \alpha_{k} g(\mX^k, \xi^k) + (1 - \alpha_k)\mM^{k - 1}\,,\\
+    \mX^{k + 1} &= \mX^k + \gamma_k\mathrm{LMO}(\mM^{k})\,.
+    \end{aligned}
+\end{equation}
+
+Here we are particularly interested in the case when $\cS$ is a ball in the $\norm{\cdot}$ norm:
 \begin{equation}\label{eq:ball}
-        \cS := \cB_\eta :=\{ \mX \mid \norm{\mX} \leq \eta \}.
-    \end{equation}
-
-    uSCG update is defined as:
-    $\mX^{k+1} = \mX^{k} + \gamma_k lmo(\mM^k),$ where $\mM^{k+1} = (1-\alpha_{k+1})\mM^{k} + \alpha_{k+1}g(\mX^k, \xi_k)$ is a momentum.
-
-    It can be easily shown that the formula is equvalent to
+    \cS = \cB_\eta = \{ \mD \in \Rmn \mid \norm{\mD} \leq \eta \}\,.
+\end{equation}
+In this case it can be easily shown that
+\begin{equation} \label{eq:lmo_in_norm_ball}
+    \argmin_{\mD \in \cS} \<\mG, \mD> = - \eta \{ \mD \in \cB_1 \mid \<\mG, \mD> = \norm{\mG}^\dagger\}\,.
+\end{equation}
+And update for $\mX$ in \cref{eq:usgd_update} becomes
 \begin{equation}\label{eq:almost_trust_update}
-        \mX^{k+1} = \mX^{k} - \gamma_k\eta \arg\max_{\mX \in \cB_1} \<\mS, \mX> = \mX^{k} - \gamma_k\eta \{\Delta \in \cB_1 \mid \<\mM^k, \Delta> = \norm{\mM^k}^\dagger\}.
-    \end{equation}
-
-    Let us set $\gamma_k = 1$.
-This transforms algorithm defined by \cref{eq:almost_trust_update} into Algorithm 1 from \citet{kovalev2025understanding}.
-Therefore, we can view \cref{eq:almost_trust_update} both as an lmo-based algorithm and as a trust-region algorithm.
+    \mX^{k+1} = \mX^{k} - \gamma_k\eta \{\mD \in \cB_1 \mid \<\mM^k, \mD> = \norm{\mM^k}^\dagger\}\,.
+\end{equation}
+This formula will later allow us to easily compute updates for various norms.
 

### 2025-11-02 conf_article/neon.tex
edited a bit more and applied latex format

@@
 
-\section{\replaced[id = IK, comment={I think the old one was a bit misleading.
-Merged what previously had been 2 subsections into this one and improved the notation.
-Also removed the thing about trust region, it'll be referenced in a succeeding sections.}]{Update Step as Constrained Optimization}{Problem Statement}}
-
-Training a neural network involves optimizing a function of weight matrices. This naturally leads us to consider the problem of minimizing a differentiable matrix function: 
+\section{\added[id = IK, comment={I think it goes well into one big section.}]{Update Step as Constrained Optimization}}
+
+\subsection{\replaced[id = IK, comment={Merged what previously had been section with two subsections into this one and improved the notation.
+Also removed the thing about trust region, it'll be referenced in succeeding sections.}]{Linear Minimization Oracle}{Problem Statement}}
+
+Training a neural network involves optimizing a function of weight matrices. For simplicity, we consider the problem of minimizing a differentiable function of a single matrix:
 \begin{equation}\label{eq:mat}
@@
 \end{equation}
+In practice, neural networks have multiple weight matrices $\mX_1, \mX_2, \ldots$, and the objective function depends on all of them.
+The framework we consider applies to each weight matrix independently, where the gradient $\mG_i = \nabla_{\mX_i} F$ for each matrix is computed via backpropagation, and the update rule is applied separately to each matrix.
+
 We equip the matrix space $\Rmn$ with a standard dot product $\<\cdot, \cdot> \to \R$ and a norm $\norm{\cdot}\colon \Rmn \to \R_+$, which possibly does not coincide with the Frobenius norm $\normf{\cdot}$.
@@
 	\begin{aligned}
-    \mM^{k} &= \alpha_{k} g(\mX^k, \xi^k) + (1 - \alpha_k)\mM^{k - 1}\,,\\
-    \mX^{k + 1} &= \mX^k + \gamma_k\mathrm{LMO}(\mM^{k})\,.
+		\mM^{k}     & = \alpha_{k} g(\mX^k, \xi^k) + (1 - \alpha_k)\mM^{k - 1}\,, \\
+		\mX^{k + 1} & = \mX^k + \gamma_k\mathrm{LMO}(\mM^{k})\,.
 	\end{aligned}
@@
 And update for $\mX$ in \cref{eq:usgd_update} becomes
-\begin{equation}\label{eq:almost_trust_update}
+\begin{equation}\label{eq:our_update}
 	\mX^{k+1} = \mX^{k} - \gamma_k\eta \{\mD \in \cB_1 \mid \<\mM^k, \mD> = \norm{\mM^k}^\dagger\}\,.
@@
 
-\section{Different Norms $\norm{\cdot}$ imply different updates}
-
-    Based on different norms $\norm{\cdot}$, we will simplify the update defined by the aforementioned equation:
-
-    \begin{equation}\label{eq:our_update}
-        \mX^{k+1} = \mX^{k} - \eta \{\Delta \in \cB_1 \mid \<\mM^k, \Delta> = \norm{\mM^k}^\dagger\}
-    \end{equation}
-
-    Throughout this work, we denote the singular value decomposition (SVD) of $\mM^k$ as $\mM^k = \mU \mSigma \mV^\top$, where $\mU = [u_1, u_2, \dots, u_r]$, $\mSigma = \diag(\sigma_1, \sigma_2, \dots, \sigma_r)$, and  $\mV = [v_1, v_2, \dots, v_r]$.
+\subsection{Different Norms $\norm{\cdot}$ imply different updates}
+
+In this subsection we will derive and discuss updates induced by the choice of the norm in \cref{eq:our_update}.
+To describe those updates, the singular value decomposition (SVD) of $\mM^k$ is required.
+We denote SVD of $\mM^k$ as $\mM^k = \mU \mSigma \mV^\top$, where $\mU = [u_1, u_2, \dots, u_r]$, $\mSigma = \diag(\sigma_1, \sigma_2, \dots, \sigma_r)$, and  $\mV = [v_1, v_2, \dots, v_r]$.
 
@@
 
-        We name the derived algorithm \emph{F-Muon}.
-It turns out that F-Muon is a convex combination of Normalized SGD and Muon.
-The implications are significant and discussed in the following sections.
+We name the derived algorithm \emph{F-Muon}. It turns out that F-Muon is a convex combination of Normalized SGD and Muon. The implications are significant and discussed in the following sections.
 
@@
 
-        We name the derived algorithm \emph{F-Neon}.
-It turns out that F-Neon is a convex combination of Normalized SGD and Neon.
-The implications are significant and discussed in the following sections.
+We name the derived algorithm \emph{F-Neon}. It turns out that F-Neon is a convex combination of Normalized SGD and Neon. The implications are significant and discussed in the following sections.
 
@@
 
-Despite such similarities, no theory has been yet proposed that would reduce the matrix case to the optimization of a function of a singular values vector.
-If such a theory is developed, analysis of matrix algorithms like Muon will be greatly simplified.
+\deleted[id = IK, comment={This is essentially like saying uh oh the optimization in multy-dimensional space is so hard, why there is no algorithm to reduce this problem to 1-d case.}]{Despite such similarities, no theory has been yet proposed that would reduce the matrix case to the optimization of a function of a singular values vector.
+If such a theory is developed, analysis of matrix algorithms like Muon will be greatly simplified.}
 
@@
 			Normalized SGD                     & $l_2$-ball, $S_2$-ball          & $-\eta \tfrac{g}{\norm{g}_2} = -\eta \tfrac{g}{\norm{g}_F}$ & \citep{hazan2015beyond}              \\
-    Momentum Normalized SGD & Ball in $l_2$, or Ball in $S_2$ & $-\eta \tfrac{g}{\norm{g}_2} = -\eta\tfrac{g}{\norm{g}_F}$ & \citep{cutkosky2020momentum}\\
+			Momentum Normalized SGD            & Ball in $l_2$, or Ball in $S_2$ & $-\eta \tfrac{g}{\norm{g}_2} = -\eta\tfrac{g}{\norm{g}_F}$  & \citep{cutkosky2020momentum}         \\
 			\hline
@@
 			\hline
-    Gauss-Southwell Coordinate Descent & Ball in $l_1$ & $-\eta \{i: g_i \geq g_k \forall k\}$ & \citep[p. 19]{shi2016primer}\\
+			Gauss-Southwell Coordinate Descent & Ball in $l_1$                   & $-\eta \{i: g_i \geq g_k \forall k\}$                       & \citep[p. 19]{shi2016primer}         \\
 			\hdashline
-    Neon & Ball in Nuclear $S_1$ & $-\eta u_1 v_1^\top$ & This work\\
+			Neon                               & Ball in Nuclear $S_1$           & $-\eta u_1 v_1^\top$                                        & This work                            \\
 			\hline
@@
 		\hline
-    Power Iterations                    & 0.01        & 1 & 7.7\\ 
-    SVDS (thick-restart Lanczos method) & 0.01        & 1 & 0.18\\
-    PCA Low Rank (RSVD)                 & 0.01        & 1 & 1.15\\
-    SVDS (thick-restart Lanczos method) & 0.01        & 10 & 0.47\\
-    PCA Low Rank (RSVD)                 & 0.01        & 10 & 19.4\\
-    SVDS (thick-restart Lanczos method) & 0.01        & 100 & 1.96\\
-    PCA Low Rank (RSVD)                 & 0.01        & 100 & 170\\
+		Power Iterations                    & 0.01 & 1   & 7.7    \\
+		SVDS (thick-restart Lanczos method) & 0.01 & 1   & 0.18   \\
+		PCA Low Rank (RSVD)                 & 0.01 & 1   & 1.15   \\
+		SVDS (thick-restart Lanczos method) & 0.01 & 10  & 0.47   \\
+		PCA Low Rank (RSVD)                 & 0.01 & 10  & 19.4   \\
+		SVDS (thick-restart Lanczos method) & 0.01 & 100 & 1.96   \\
+		PCA Low Rank (RSVD)                 & 0.01 & 100 & 170    \\
 	\end{tabular}
@@
 	\begin{align}
-        \eta &= \cO\left(\min\left\{\frac{\epsilon}{L}, \frac{\epsilon^3}{\rho^2\sigma^2L}\right\}\right),
+		\eta & = \cO\left(\min\left\{\frac{\epsilon}{L}, \frac{\epsilon^3}{\rho^2\sigma^2L}\right\}\right),
 		\qquad
@@
 		\label{eq:str_K_nonconvex}
-        K &= \cO\left(\max\left\{
+		K    & = \cO\left(\max\left\{
 		\frac{\rho\sigma}{\epsilon},
@@
 \iffalse
-
 	\section{Idea}
@@
 				      \hline
-        Power Iterations                    & 0.01        & 1 & 7.7\\ 
-        SVDS (thick-restart Lanczos method) & 0.01        & 1 & 0.18\\
-        PCA Low Rank (RSVD)                 & 0.01        & 1 & 1.15\\
-        SVDS (thick-restart Lanczos method) & 0.01        & 10 & 0.47\\
-        PCA Low Rank (RSVD)                 & 0.01        & 10 & 19.4\\
-        SVDS (thick-restart Lanczos method) & 0.01        & 100 & 1.96\\
-        PCA Low Rank (RSVD)                 & 0.01        & 100 & 170\\
+				      Power Iterations                    & 0.01 & 1   & 7.7    \\
+				      SVDS (thick-restart Lanczos method) & 0.01 & 1   & 0.18   \\
+				      PCA Low Rank (RSVD)                 & 0.01 & 1   & 1.15   \\
+				      SVDS (thick-restart Lanczos method) & 0.01 & 10  & 0.47   \\
+				      PCA Low Rank (RSVD)                 & 0.01 & 10  & 19.4   \\
+				      SVDS (thick-restart Lanczos method) & 0.01 & 100 & 1.96   \\
+				      PCA Low Rank (RSVD)                 & 0.01 & 100 & 170    \\
 			      \end{tabular}

### 2025-11-02 conf_article/neon.tex
small changes for clarity

@@
 \subsection{\replaced[id = IK, comment={Merged what previously had been section with two subsections into this one and improved the notation.
-Also removed the thing about trust region, it'll be referenced in succeeding sections.}]{Linear Minimization Oracle}{Problem Statement}}
+Also removed the thing about trust region, it'll be referenced in succeeding sections.}]{Linear Minimization Oracle Framework}{Problem Statement}}
 
@@
 \end{equation}
-In practice, neural networks have multiple weight matrices $\mX_1, \mX_2, \ldots$, and the objective function depends on all of them.
-The framework we consider applies to each weight matrix independently, where the gradient $\mG_i = \nabla_{\mX_i} F$ for each matrix is computed via backpropagation, and the update rule is applied separately to each matrix.
-
 We equip the matrix space $\Rmn$ with a standard dot product $\<\cdot, \cdot> \to \R$ and a norm $\norm{\cdot}\colon \Rmn \to \R_+$, which possibly does not coincide with the Frobenius norm $\normf{\cdot}$.
@@
 \end{equation}
-where $\mG$ is typically a gradient (possibly with momentum) of $F$ and $\cS \subset \Rmn$ is some set.
-The update rule used in Muon optimizer is uSCG \citep{pethick2025training}.
-In the most general case, which involves momentum and stochastic gradient $g(\mX^k, \xi^k)$, it can be written as
+where $\mG$ is typically a gradient (possibly with momentum) of $F$ and $\cS \subset \Rmn$ is some set. The simple update formula then can look like
+\begin{equation}\label{eq:simple_update}
+	\mX^{k + 1} = \mX^k + \mathrm{LMO}\left(\nabla F\left(\mX^k\right)\right)\,.
+\end{equation}
+
+In a typical neural network, the objective function $F$ depends on a set of weight matrices $\{ \mX_1, \mX_2, \ldots \}$.
+The optimization framework we have described is applied in a layer-wise fashion.
+At each iteration $k$, a stochastic gradient $g(\mX^k, \xi^k)$ is computed using a mini-batch of data $\xi^k$ via backpropagation.
+This yields a separate gradient component, $\mG_i^k$, for each matrix $\mX_i$.
+The LMO-based update rule is then applied to each matrix $\mX_i$ using its corresponding gradient component $\mG_i^k$.
+
+The update rule used in Muon optimizer is uSCG \citep{pethick2025training}. In the most general case, which involves momentum, it can be written as
 \begin{equation}\label{eq:usgd_update}

### 2025-11-06 conf_article/neon.tex
Expanded the experiments

@@
 \replaced[id=IK, comment={This rewrite clarifies the sentence structure, making it clear that the norms (not the algorithms) are being combined.}]{Subsequently, we construct a second family of algorithms, dubbed {\it F-Fanions}, based on the duals of convex combinations of Ky Fan k-norms and the Frobenius norm.
-One prominent member of this family is {\it F-Muason}.}{Then, we consider their convex combinations with the Frobenius norm and corresponding duals to develop a new family of algorithms we name {\it F-Fanions}, one of them being {\it F-Muon}.} We complement our theoretical analysis with an extensive empirical study of the algorithms across a wide range of tasks and settings, from which it follows that properly-tuned F-Muon is almost on par with Muon, \replaced[id=IK, comment={Alex, clarify this please.
+One prominent member of this family is {\it F-Muon}.}{Then, we consider their convex combinations with the Frobenius norm and corresponding duals to develop a new family of algorithms we name {\it F-Fanions}, one of them being {\it F-Muon}.} We complement our theoretical analysis with an extensive empirical study of the algorithms across a wide range of tasks and settings, from which it follows that properly-tuned F-Muon is almost on par with Muon, \replaced[id=IK, comment={Alex, clarify this please.
 It's a very non-obvious and non-clear conclusion, especially when knowing that Muon itself only provides an approximate answer to the LMO.
@@
 
-We introduce the family of {\it Fanions}, which consists of {\it $k$-Fanions}, lmo-based algorithms under the $\normkfk{\mM^t}^\dagger$ norms.
-By this terminology and the lemma, Muon is a $\min\{n,m\}$-Fanion, while Neon is an 1-Fanion.
-Moreover, the centralized version of rank-$r$ Dion \citep{ahn2025dioncommunicationefficientoptimizerlarge} without the error feedback and without scaling of the update, from the perspective of lmo, is actually a $r$-Fanion.
+We introduce the family of {\it Fanions}, which consists of {\it Fanion-$k$}, lmo-based algorithms under the $\normkfk{\mM^t}^\dagger$ norms.
+By this terminology and the lemma, Muon is a Fanion-$\min\{n,m\}$, while Neon is an Fanion-1.
+Moreover, the centralized version of rank-$r$ Dion \citep{ahn2025dioncommunicationefficientoptimizerlarge} without the error feedback and without scaling of the update, from the perspective of lmo, is actually a Fanion-$r$.
 
@@
 			\hline
-			Gauss-Southwell Coordinate Descent & Ball in $l_1$                   & $-\eta \{i: g_i \geq g_k \forall k\}$                       & \citep[p. 19]{shi2016primer}         \\
+			Gauss-Southwell Coordinate Descent & Ball in $l_1$                   & $-\eta \sum_{i \in \argmax|g_i^t|} \sign(g_i^t)e_i$                       & \citep[p. 19]{shi2016primer}         \\
 			\hdashline
@@
 \end{table}
-\section{Trust Region Bounds for L-smooth functions}
-First, we analyze the problem in the unstochastic case.
-From Corollary 1 of \citet{kovalev2025understanding}, we directly get the following result that matches lower bounds, as was noted in that article.
-
-\begin{lemma}\label{lemma:no_noise_tr}
-	To reach the precision $\min_{k=1\ldots K} \norm{\nabla f(\mX_k)}^\dagger \leq \epsilon$ by the iterations~\eqref{eq:our_update} under the conditions of \Cref{eq:L}, it is sufficient to choose the stepsize $\eta$ and the number of iterations $K$ as follows:
-	\begin{equation}\label{eq:unstoch_tr}
-		\eta = \cO\left(\frac{\epsilon}{L}\right),\qquad K = \cO\left(\frac{L\Delta_0}{\epsilon^2}\right).
-		% \eta = \frac{\epsilon}{3L},\qquad K = \ceil*{\frac{6L\Delta_0}{\epsilon^2}}.
-	\end{equation}
-\end{lemma}
-
-In the stochastic case, from Corollary of 2 of \citet{kovalev2025understanding}, we directly get the following result that once more matches lower bounds:
-
-\begin{lemma}\label{lemma:stoch_tr}
-	To reach the precision $\E{\min_{k=1\ldots K} \norm{\nabla f(\mX_k)}^\dagger} \leq \epsilon$ by ~\eqref{eq:our_update} under the assumptions \Cref{eq:variance}, \Cref{eq:L}, \Cref{eq:norm}, it is sufficient to choose the parameters as follows:
-	\begin{align}
-		\eta & = \cO\left(\min\left\{\frac{\epsilon}{L}, \frac{\epsilon^3}{\rho^2\sigma^2L}\right\}\right),
-		\qquad
-		\alpha = \cO\left(\min\left\{1, \frac{\epsilon^2}{\rho^2\sigma^2}\right\}\right),
-		\\
-		\label{eq:str_K_nonconvex}
-		K    & = \cO\left(\max\left\{
-		\frac{\rho\sigma}{\epsilon},
-		\frac{\rho^3\sigma^3}{\epsilon^3},
-		\frac{L\Delta_0}{\epsilon^2},
-		\frac{L\Delta_0\rho^2\sigma^2}{\epsilon^4}
-		\right\}\right).
-	\end{align}
-\end{lemma}
-
-As the norms $\normf{\cdot}$, $\normn{\cdot}$, $\norms{\cdot}$ are almost proportional to each other when $m, n \rightarrow \infty$ (with high probability for random matrices), the expected convergence guarantees in terms of $\normf{\cdot}$ are the same (it can be easily shown by noting that $\norm{\mX}\sim\alpha\normf{\mX}$, $\norm{\nabla f(\mX)}^\dagger \sim \frac{1}{\alpha} \normf{\nabla f(\mX)}$, and expressing $L$-constant via $L_F$-constant for the Frobenius norm).
-
-From the theory of random martices and the Marchenko-Pastur law, we get that random $\mM \in \Rmn: \mM \sim \cN(0, \sigma^2 \Rmn)$ has the following asymptotics of its norms:
-
-Nuclear: $\sigma n \sqrt{m}$
-
-Frobenius: $\sigma \sqrt{m n}$
-
-Spectral: $\sigma(\sqrt{m} + \sqrt{n})$
-
-This means that for square random matrices $n \times n$  the following asymptotics take place: $\normf{\cdot} \sim \frac{\sqrt{n}}{2}\norms{\cdot}$ and $\normn{\cdot} \sim \frac{n}{2}\norms{\cdot}$.
 
@@
 \subsection{Randomized Linear Least Squares}
-Since the provided by other authors \citep{kovalev2025understanding,riabinin2025gluon} theoretical guarantees are almost norm-independent, we have to test them in practice.
-
-To test the bounds from \citet{kovalev2025understanding}, we construct the following L-smooth problem:
+We first compare F-Fanions on the following L-smooth problem:
 
 \begin{equation}\label{eq:lls}
-	F(\mX) = \frac{1}{2} \<(\mX-\mS), \mM (\mX - \mS) \mN>
+	F(\mX) = \frac{1}{2} \<(\mX-\mS), \mM (\mX - \mS) \mN> \to \min_{\mX \in \Rmn}\
 \end{equation}
 
-where $\mX \in \Rmn$, $m=10$, $n=10$, $S \in \Rmn$, $\mM \in \mathbb{S}^m_{+}$ and $\mN \in \mathbb{S}^n_{+}$ are positive-semidefinite matrices.
-Spectra of $\mM$ and $\mN$ are uniformly distributed in (0, 1) interval.
+where $\mX \in \Rmn$, $m=500$ and $n=500$ are typical dimensions of a neural network weight matrix, $S \in \Rmn$, $\mM \in \mathbb{S}^m_{+}$ and $\mN \in \mathbb{S}^n_{+}$ are positive-semidefinite matrices.
+The spectra of $\mM$ and $\mN$ are uniformly distributed in the (0, 1) interval.
+We set $\mS=0$, and $\mX^0 \sim 0.1 \gN(0, \mI)$.
 
-It it easy to derive the gradient
-\begin{equation}\label{eq:grad_lls}
-	\nabla F(\mX) = \mM (\mX - \mS) \mN,
-\end{equation}
+We run different Fanions and their respective F-Fanions with $\alpha=1/2$: Neon (Fanion-1), Fanion-2, Fanion-5, Fanion-100, and Muon (Fanion-500).
+We test them against Normalized SGD, which is also a F-Fanion with $\alpha = 0$ and an arbitrary $k$.
 
-Let us define $\gamma$ as $\norm{\cdot} \sim \gamma \normf{\cdot}$, which is asymptotics from the previous section.
-Then $\norm{\cdot}^\dagger \sim \frac{1}{\gamma}\normf{\cdot}$, as $\normf{\cdot}^\dagger = \normf{\cdot}$.
-Hence, $\norm{\cdot} \sim \gamma^2 \norm{\cdot}^\dagger$
+Since theoretical bounds \citep{kovalev2025understanding, riabinin2025gluon} rely on a very loose norm bound $\norm{\cdot}\le \rho \normf{\cdot}$, we do not derive learning rate from smoothness.
+Rather, we set the learning rate of $1/\sqrt{T}$ for each optimizer and the momentum of 0.6.
+12 000 iterations for each algorithm ensure that the rate at some point gets low enough for each algorithm to converge.
+The results are presented in \cref{fig:lls} and ~\cref{sec:lls_plots_section}.
 
-Then $\norm{\nabla f(\mX) - \nabla f(\mY)}^\dagger = \norm{\mM (\mX - \mY) \mN}^\dagger \leq \norm{\mM}^\dagger \norm{\mN}^\dagger \norm{\mX - \mY}^\dagger \sim \norm{\mM}^\dagger \norm{\mN}^\dagger \gamma^2 \norm{\mX - \mY}$, and $L \sim \gamma^2 \norm{\mM}^\dagger \norm{\mN}^\dagger$.
-
-Now for the known norms:
-
-Frobenius (NSGD): $L = \normf{M} \normf{N}$
-
-Spectral (Muon): $\gamma \sim \frac{2}{\sqrt{n}} \implies L \sim \frac{2}{n}\normn{M}\normn{N}$
-
-Nuclear (Neon): $\gamma \sim \sqrt{n} \implies L \sim n\norms{M}\norms{N}$
-
-We take learning rate $\eta$ and iteration number $K$ from \cref{eq:unstoch_tr}.
-
-We are interested to get $\epsilon_2=$ 1e0 precision in $\norms{\cdot}$.
-Hence, we target $\epsilon = \epsilon_2$ for the nuclear norm (Neon), $\epsilon = \frac{n}{2}$ for the spectral norm (Muon), and $\epsilon = \frac{\sqrt{n}}{2}$ for the Frobenius norm (NSGD).
-With that, we get the following lrs: Muon lr: 2.33E-04, Neon lr: 2.10e-02, Frobenius (NSGD) lr: 1.75e-02, and the following number of iterations: Muon iter: 68, Neon iter: 11919, Frobenius iter: 323.
-L-constants are 171.4, 47.7, and 16.2 respectively.
-As we will see, these results do not relate well to the practice.
-
-To make the problem more real-world, we do not use known information about smoothness.
-We run NormalizedSGD, Muon, F-Muon with $\alpha=1/2$, Neon, and F-Neon with $\alpha=1/2$ for 5 000 iterations with learning rate = 1e-2, which has been close to computed theoretically above.
-The results are in \cref{fig:lls}.
-
-Though Neon displays the worst convergence in terms of loss, it is a leader in terms of the norm of the gradient.
-This might signify that the known bounds are not sufficient even in the case of a convex optimization: lmo-based algorithms may converge well by these bounds, but be impractical like Neon.
-
-This experiment may also show that F-Muon may be more robust under general conditions than Muon.
+In terms of the loss minimization, F-Muon had the fastest convergence.
+In terms of the Frobenius norm of the gradient, the lowest was observed for NSGD.
+Then go F-Neon (F-Fanion-1), F-Fanion-2, F-Fanion-5, F-Fanion-100, and F-Muon (Fanion-500).
+It is noticeable that all F-Fanions have lower Frobenius norms of the gradients than their respective Fanioins.
+The same behavior will be recorded in the next section.
 
@@
     \centering
-	\begin{subfigure}[t]{0.48\linewidth}
+    \begin{subfigure}[t]{0.8\linewidth}
         \centering
-		\includegraphics[width=\linewidth]{figs/simple_lls/loss_vs_iteration_50x50.pdf}
-		\caption{Loss}
+        \includegraphics[width=\linewidth]{figs/simple_lls/loss_vs_iteration_500x500.pdf}
+        \caption{The loss}
         \label{fig:lls_loss}
     \end{subfigure}
-	\hfill
-	\begin{subfigure}[t]{0.48\linewidth}
+    \vspace{1em} % optional: add vertical spacing between subfigures
+    \begin{subfigure}[t]{0.8\linewidth}
         \centering
-		\includegraphics[width=\linewidth]{figs/simple_lls/gradient_spectral_norm_vs_iteration_50x50.pdf}
-		\caption{Spectral norm of the gradient}
-		\label{fig:lls_grad_norm}
+        \includegraphics[width=\linewidth]{figs/simple_lls/gradient_frobenius_norm_vs_iteration_500x500.pdf}
+        \caption{The Frobenius norm of the gradient}
+        \label{fig:lls_fro_grad_norm}
     \end{subfigure}
-	\caption{Linear least squares problem for a 50 x 50 matrix}
+    \caption{Linear least squares problem for a 500x500 matrix}
     \label{fig:lls}
-
 \end{figure}
@@
 	\centering
-	\includegraphics[width=0.6\textwidth]{figs/fstardual_cifar.pdf} % adjust width as needed
+	\includegraphics[width=0.6\textwidth]{figs/norms/fstardual_cifar.pdf} % adjust width as needed
 	\caption{Visualization of lmo balls for Muon and F-Muon.}
@@
 
-\subsection{Modded NanoGPT speedrun}
+\subsection{NanoGPT speedrun}
 We test F-Muon on NanoGPT speedrun \citep{modded_nanogpt_2024}.
@@
 After testing for 1750 steps, as it has been done on the speedruns, we get 3.281 cross-entropy loss, while on Muon, it falls below the target threshold 3.28 reaching 3.279.
+However, this difference is negligible, if one looks at \cref{fig:gpt_loss}.
+It is even more striking, considering the fact that F-Muon is an average of Muon and NSGD, and the later performed quite poorly.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.7\textwidth]{figs/nanogpt/val_loss_vs_step.pdf}
+	\caption{The validation loss for NanoGPT}
+	\label{fig:gpt_loss}
+\end{figure}
 
@@
 \appendix
-
-\section{Formal assumptions}
-Here we restate assumptions from \citet{kovalev2025understanding} that are used in the theorems from the main part of the article.
-
-	{\bf Lipschitz continuous gradient.}
-We assume that the gradient $\nabla f(\cdot)$ is Lipschitz continuous with respect to the norm $\norm{\cdot}$, that is, the following inequality holds:
-\begin{aequation}\label{eq:L}
-	\norm{\nabla f(\mX) - \nabla f(\mX')}^\dagger \leq L \norm{\mX-\mX'}
-	\quad\text{for all}\;
-	\mX, \mX' \in \Rmn,
-\end{aequation}
-where $L > 0$ is the gradient Lipschitz constant.
-
-	{\bf Stochastic gradient estimator.}
-We assume access to a stochastic estimator $g(\cdot\,;\xi)\colon \Rmn \to \Rmn$ of the gradient $\nabla f(\cdot)$, where $\xi \sim \cD$ is a random variable sampled from a probability distribution $\cD$.
-We assume that the stochastic gradient estimator $g(\cdot;\xi)$ is unbiased and has bounded variance, that is, the following relations hold:
-\begin{aequation}\label{eq:variance}
-	\Ed{\xi \sim \cD}{g(\mX;\xi)} = \nabla f(\mX)
-	\quad\text{and}\quad
-	\Ed{\xi \sim \cD}{\sqnf{g(\mX;\xi) - \nabla f(\mX)}} \leq \sigma^2
-	\quad\text{for all}\;
-	\mX \in \Rmn,
-\end{aequation}
-where $\sigma > 0$ is a positive variance parameter, and $\normf{\cdot}$ is the standard Euclidean, i.e.
-Frobenius, norm induced by the inner product $\<\cdot,\cdot>$, i.e., $\normf{\mX} = \sqrt{\<\mX,\mX>}=\sqrt{\tr(\mX^\top \mX)}$.
-
-It is important to highlight that while \Cref{eq:L} uses the dual norm $\norm{\cdot}^\dagger$ to measure the difference between the gradients, the variance in \Cref{eq:variance} is measured with respect to the Frobenius norm $\sqnf{\cdot}$, which is necessary to properly utilize the unbiasedness property of the stochastic gradient estimator $g(\cdot;\xi)$.
-Therefore, we need to provide a connection between these norms using the following inequality:
-\begin{aequation}\label{eq:norm}
-	\norm{\mX}^\dagger \leq \rho\cdot\normf{\mX}
-	\quad\text{for all}\;
-	\mX \in \Rmn,
-\end{aequation}
-where $\rho > 0$ is a positive constant.
 
@@
 		\centering
-		\includegraphics[width=\linewidth]{figs/fstardualball.pdf}
+		\includegraphics[width=\linewidth]{figs/norms/fstardualball.pdf}
 		\caption{lmo balls for F-Muon for different $\alpha$}
@@
 		\centering
-		\includegraphics[width=\linewidth]{figs/ftwodualball.pdf}
+		\includegraphics[width=\linewidth]{figs/norms/ftwodualball.pdf}
 		\caption{lmo ball for F-Neon for different $\alpha$}
@@
 		\centering
-		\includegraphics[width=\linewidth]{figs/KyFan.pdf}
+		\includegraphics[width=\linewidth]{figs/norms/KyFan.pdf}
 		\caption{Ky Fan 2-norm}
@@
 		\centering
-		\includegraphics[width=\linewidth]{figs/KyFanDual.pdf}
+		\includegraphics[width=\linewidth]{figs/norms/KyFanDual.pdf}
 		\caption{Dual to Ky Fan 2-norm}
@@
 
+\section{More plots for Linear Least Squares}
+\label{sec:lls_plots_section}
+
+\begin{figure}[h!]
+    \centering
+    \begin{subfigure}[t]{0.6\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{figs/simple_lls/gradient_spectral_norm_vs_iteration_500x500.pdf}
+        \caption{The spectral norm of the gradient}
+        \label{fig:lls_op_grad_norm}
+    \end{subfigure}
+    \vspace{1em} % optional: add vertical spacing between subfigures
+    \begin{subfigure}[t]{0.6\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{figs/simple_lls/gradient_nuclear_norm_vs_iteration_500x500.pdf}
+        \caption{The nuclear norm of the gradient}
+        \label{fig:lls_nuclear_grad_norm}
+    \end{subfigure}
+    \vspace{1em} % optional: add vertical spacing between subfigures
+    \begin{subfigure}[t]{0.6\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{figs/simple_lls/loss_vs_time_500x500.pdf}
+        \caption{The loss over time}
+        \label{fig:lls_loss_time}
+    \end{subfigure}
+    \caption{More images for Linear least squares problem for a 500x500 matrix}
+    \label{fig:app_lls}
+\end{figure}
+
 \section{Technical details of the experiments}
@@
 
-\iffalse
-	\section{Idea}
-	\iffalse
-		The goal of the project is to make variations on Muon to speed it up.
-Recently, authors of \cite{bernstein2024oldoptimizernewnorm} have proposed to derive update step for an optimizer as the solution of the certain optimization problem.
-This approach can be utilized to derive Muon \cite{jordan2024muon}, a novel algorithm for fast training of neural networks.
-Instead of using spectral norm, used to derive Muon, we utilize nuclear norm to produce a new optimization algorithm.
-	\fi
-	\subsection{Problem (Project description)}
-
-	In this subsection, we provide a more detailed description of our idea and formulate it as a mathematical problem.
-The authors of \cite{bernstein2024oldoptimizernewnorm} suggest obtaining the update step as a solution to the optimization problem:
-	\begin{equation}
-		\langle g, \delta w \rangle + \lambda \norm{\delta w}^2 \to \min_{\delta w}\,,
-	\end{equation}
-	where $w$ is the weight vector, $g$ is a gradient-like vector (e.g., obtained via momentum SGD), and $\norm{\cdot}$ represents a certain norm.
-Many popular optimizers, such as Adam (with exponential moving average disabled) and vanilla SGD, can be cast within this framework \cite{bernstein2024oldoptimizernewnorm}.
-
-	In large language models, most weights are structured as matrices, which offers additional opportunities for optimization.
-Let $W$ be the weight matrix of a linear layer, and $G$ be a gradient-like matrix.
-Then, the update step $\delta W$ can be obtained as a solution to the optimization problem:
-	\begin{equation}\label{eqn:opt_problem_mat}
-		\langle G, \delta W \rangle + \lambda \norm{\delta W}^2 \to \min_{\delta W}\,,
-	\end{equation}
-	where $\norm{\cdot}$ denotes a certain matrix norm.
-By setting this norm to the RMS-to-RMS norm (a scaled version of the spectral norm), we recover the Muon optimizer \cite{bernstein2025deriving, bernstein2024oldoptimizernewnorm} with an update step defined by:
-	\begin{equation}\label{eqn:update_muon}
-		\delta W = - \frac{1}{\lambda}\sqrt{\frac{n}{m}}UV^T\,,
-	\end{equation}
-	where $m$ is the input dimension of the layer, $n$ is the output dimension, and $U$ and $V$ are obtained from the singular value decomposition of the gradient matrix $G = U \Sigma V$.
-
-	Motivated by the recent achievements of the Muon optimizer (e.g., \cite{liu2025muon}), we consider alternative choices of norms, specifically the nuclear norm $\norm{\cdot}_*$ and a custom $F*$ norm, given by
-	\begin{equation}\label{eqn:F_star}
-		\norm{X}_{F*}^2 = \frac{\norm{X}_F^2 + \norm{X}_*^2}{2}\,,
-	\end{equation}
-	where $\norm{\cdot}_F$ denotes the Frobenius norm.
-
-	Using the nuclear norm in \eqref{eqn:opt_problem_mat} leads to a rank-one update of the weight matrices:
-	\begin{equation}\label{eqn:update_star}
-		\delta W = -\frac{1}{2\lambda} u_1 \sigma_1 v_1^T\,,
-	\end{equation}
-	where $\sigma_1$ is the largest singular value, and $u_1$ and $v_1$ are the corresponding singular vectors.
-We expect one iteration of this method to be significantly faster than one iteration of Muon.
-
-	Another choice is the $F*$ norm. With this choice, \eqref{eqn:opt_problem_mat} yields
-	\begin{equation}\label{eqn:update_F_star}
-		\delta W = -\frac{1}{\lambda}UDV^T
-	\end{equation}
-	with $D = \text{diag}(d_i)$, where $d_i = [\sigma_i - \tau]_+$, and $\tau$ is given by
-	\begin{equation}
-		\sum_{i} [\sigma_i - \tau]_+ = \tau\,.
-	\end{equation}
-	We anticipate that the method with this update step will perform well with large batch sizes.
-
-	In this article we show how one can quickly compute weight updates defined by \eqref{eqn:update_star} or \eqref{eqn:update_F_star}.
-Then we finilize the method by adding momentum and test their performance against those of Muon at training multilayer perceptron and transformer.
-The results will be fast algorithm, which we will convert into a new optimizer classes for PyTorch, as was done with Muon.
-
-	% Removed template instructions sections
-	\section{Related Work}
-	Our review primarily focuses on the Muon and Shampoo optimizers, as our algorithm extends the ideas used to derive these methods.
-We highlight the advantages and disadvantages of these approaches, the unique effects they introduce, and compare them to Neon.
-
-	\subsection{Muon optimizer}
-
-	In the previous section, we described the theoretical foundation behind the weight update step in the Muon optimizer, but we did not discuss how to obtain the required matrices in practice.
-The update step is defined by \eqref{eqn:update_muon}, which requires $UV^T$ with $U$ and $V^T$ from the singular value decomposition of the gradient-like matrix $G$.
-A naive solution would involve computing the SVD of $G$ and constructing the required expression.
-However, the developers of the Muon optimizer introduced a workaround using Newton-Schulz iterations \cite{jordan2024muon}.
-The Newton-Schulz iterations from the original article \cite{jordan2024muon} require 10 matrix-matrix multiplications to achieve the desired accuracy.
-The asymptotic complexity of such an operation is identical to that of SVD and equals $O(mn \min\{m, n\})$ for an $m \times n$ matrix.
-Nevertheless, matrix multiplication on modern GPUs can be performed much more efficiently.
-
-	The performance of Muon in training large language models was tested \cite{liu2025muon} against AdamW.
-The testing demonstrated excellent performance by Muon, which was approximately 2 times more efficient in terms of FLOPs required to reach a certain loss value.
-This is even more remarkable when considering the cost of one iteration: Muon requires an additional $O(mn \min\{m, n\})$ FLOPs per $m \times n$ matrix, while AdamW needs only $O(mn)$.
-
-	Another interesting discovery about the Muon optimizer is that it accelerates grokking \cite{tveit2025muonoptimizeracceleratesgrokking}.
-In the test problem, Muon achieved grokking significantly faster than AdamW in terms of passed epochs, with a mean grokking epoch of 102.89 for Muon and 153.09 for AdamW.
-The authors suggest that this may be due to the fact that Muon stimulates broader exploration by orthogonalizing the gradient matrix, thus avoiding memorization.
-
-	Recently, theoretical guarantees for Muon convergence have been derived \cite{li2025noteconvergencemuon}.
-In particular, in the $L$-smooth convex case, it achieves $O(1/T^{\frac{1}{2}})$ (with full gradient) and $O(1/T^{\frac{1}{4}})$ (with stochastic gradient) bounds on the Frobenius norm of the gradient or the mathematical expectation of the gradient norm, respectively, where $T$ is the number of iterations.
-
-	\subsection{Shampoo optimizer}
-
-	Another optimizer that exploits the matrix (and even tensor) structure of weights in neural networks is the Shampoo optimizer.
-We avoid a detailed description of this method here and refer the reader to the original article \cite{gupta2018shampoopreconditionedstochastictensor}, but we outline the key properties of Shampoo and its relation to the Muon optimizer.
-
-	The Shampoo optimizer uses left and right preconditioning for the gradient-like matrix, leveling its spectrum.
-The preconditioners are computed from the exponentially averaged gradients, and their computation requires $O(n^3 + m^3)$ per $m \times n$ matrix.
-This exponential averaging is a key feature that provides several distinct interpretations for the preconditioners.
-They can be viewed as an approximation of the Gauss-Newton component of the Hessian or the first step of the power iteration algorithm for computing the optimal Kronecker product approximation \cite{morwani2025a}.
-With exponential averaging turned off, the update step of Shampoo can be simplified and becomes identical to that of Muon \cite{jordan2024muon}.
-
-	Convergence analysis is presented in the original article \cite{gupta2018shampoopreconditionedstochastictensor}.
-Shampoo achieves $O(1/T^{\frac{1}{2}})$ convergence for the loss function value in the $L$-smooth convex case, where $T$ is the number of iterations.
-
-	\subsection{Synthesis and Neon's position}
-
-	Recent developments in optimization techniques show that utilizing the matrix structure of weights in neural networks can be very beneficial.
-Optimizers following this path converge faster in terms of iterations or epochs and often even FLOPs, but have a high iteration cost.
-Neon seeks to decrease the iteration cost while preserving fast convergence.
-
-	While the uniqie advantages and effects introduced by Neon are yet to be discovered, we can already say that our new optimizer introduces additional overhead of $O(mn)$ FLOPS on average per $m \times n$ weight matrix, which is much better then $O(m n \min\{m, n\})$ for Muon optimizer and $O(n^3 + m^3)$ for Shampoo.
-
-	\subsection{Optimization Strategies for Efficient Large Language Model Training}
-	The unprecedented scale of modern Large Language Models (LLMs) has pushed traditional optimizers like AdamW \cite{Loshchilov2017FixingWD} to their limits in terms of computational efficiency and convergence speed \cite{liu2025muon,chen2025cosmoshybridadaptive}.
-This challenge has catalyzed research into more sophisticated optimization approaches that can maintain or improve performance while reducing training costs.
-
-	The Muon optimizer  has emerged as a promising alternative based on matrix orthogonalization principles.
-Scaling Muon to billion-parameter LLMs required two key adaptations: the integration of L2 weight decay for stability and the implementation of per-parameter update scaling to handle diverse parameter distributions efficiently .
-Empirical evaluations demonstrate that Muon can match or exceed AdamW's model quality while requiring only about 52 of the training FLOPs.
-The successful training of the Moonlight model series, including a 16B-parameter Mixture-of-Experts model, validates Muon's practicality for production-scale applications \cite{liu2025muon}.
-
-	Building on this foundation, hybrid approaches like COSMOS \cite{chen2025cosmoshybridadaptive} further enhance efficiency by combining optimization techniques based on gradient structure.
-COSMOS applies computationally intensive updates to a low-dimensional "leading eigensubspace" while using memory-efficient methods like Muon for the remaining parameters.
-This approach maintains convergence benefits while substantially reducing memory requirements.
-For distributed training environments, optimizers like Dion \cite{ahn2025dioncommunicationefficientoptimizerlarge} specifically target communication efficiency by minimizing data exchange between workers through distributed orthonormalization techniques.
-
-	\section{Derivation of update rules}
-
-	\begin{lemma}\label{lem:opt_star}
-		Let $G \in \mathbb{R}^{m \times n}$ and $\lambda > 0$. Then, the following optimization problem
-		\begin{equation*}
-			f(\delta W) = \langle G, \delta W \rangle + \lambda \norm{\delta W}_*^2 \to \min_{\delta W}
-		\end{equation*}
-		has solution
-		\begin{equation*}
-			\delta W = -\frac{1}{2\lambda} u_1 \sigma_1 v_1^T\,,
-		\end{equation*}
-		where $\sigma_1$ is the largest singular value of $G$, and $u_1$ and $v_1$ are the corresponding singular vectors.
-	\end{lemma}
-
-	\begin{proof}
-		Let us denote $r = \min\{m, n\}$. Then by Von Neumann's trace inequality,
-		\begin{equation*}
-			\vert\langle G, \delta W \rangle\vert \leq \sum_{i=1}^r \sigma_i(G)\sigma_i(\delta W)
-			\Rightarrow
-			\langle G, \delta W \rangle \geq -\sum_{i=1}^r \sigma_i(G)\sigma_i(\delta W)\,.
-		\end{equation*}
-		Thus, expressing nuclear norm through singular values, we can write down
-		\begin{equation*}
-			f(\delta W) \geq -\sum_{i=1}^r \sigma_i(G)\sigma_i(\delta W) + \lambda\left(\sum_{i=1}^{r}\sigma_i(\delta W)\right)^2 \geq \min_{d_1,\dots,d_r \geq 0}-\sum_{i=1}^r \sigma_i(G)d_i + \lambda\left(\sum_{i=1}^{r}d_i\right)^2.
-		\end{equation*}
-		By Karush-Kuhn-Tucker theorem, necessary conditions of minimum are
-		\begin{equation*}
-			\left(d_i \geq 0\ \text{and}\ -\sigma_i(G) + 2\lambda\sum_{j=1}^{r}d_j = 0\right)\ \text{or}\ \left(d_i = 0\ \text{and}\ -\sigma_i(G) + 2\lambda\sum_{j=1}^{r}d_j \geq 0\right) \,\, i = 1,\dots,r\,.
-		\end{equation*}
-		These conditions simplify to
-		\begin{equation*}
-			\sum_{i \in S}d_i = \sigma_1(G)\,,\ \begin{cases}
-				d_i \geq 0 & \text{if}\ \sigma_i(G) = \sigma_1(G)\,, \\
-				d_i = 0    & \text{otherwise}\,.
-			\end{cases}
-		\end{equation*}
-		All points satisfying those conditions deliver minimum, and
-		\begin{equation*}
-			f(\delta W) \geq -\frac{\sigma_1^2(G)}{4\lambda}\,.
-		\end{equation*}
-
-		Now let
-		\begin{equation*}
-			\delta W^* = -\frac{1}{2\lambda} u_1 \sigma_1(G) v_1^T\,.
-		\end{equation*}
-		Inserting it to $f(\delta W)$ gives
-		\begin{equation*}
-			f\left(\delta W^*\right) = -\frac{\sigma_1(G)^2}{2\lambda} + \frac{\sigma_1(G)^2}{4\lambda} = -\frac{\sigma_1(G)^2}{4\lambda}
-		\end{equation*}
-		This matches the derived lower bound. Thus, $\delta W^*$ minimizes $f(\delta W)$.
-	\end{proof}
-
-	\begin{lemma}\label{lem:opt_F_star}
-		Let $G \in \mathbb{R}^{m \times n}$, $r = \min\{m. n\}$ and $\lambda > 0$. Then, the following optimization problem
-		\begin{equation*}
-			f(\delta W) = \langle G, \delta W \rangle + \lambda \norm{\delta W}_{F*}^2 \to \min_{\delta W}\,,
-		\end{equation*}
-		where $\norm{\cdot}_{F*}$ is defined in \eqref{eqn:F_star} has solution
-		\begin{equation}
-			\delta W = -\frac{1}{\lambda}UDV^T
-		\end{equation}
-		with $D = \text{diag}(d_i)$, where $d_i = [\sigma_i - \tau]_+$, and $\tau$ is given by
-		\begin{equation}
-			\sum_{i=1}^r [\sigma_i - \tau]_+ = \tau\,.
-		\end{equation}
-	\end{lemma}
-
-	\begin{proof}
-		Analogously to the proof of Lemma~\ref{lem:opt_star}, we can use Von Neumann's trace inequality to write down
-		\begin{equation*}
-			f(\delta W) \geq -\sum_{i=1}^r \sigma_i(G)\sigma_i(\delta W) + \frac{\lambda}{2}\left(\sum_{i=1}^{r}\sigma_i(\delta W)\right)^2 + \frac{\lambda}{2}\sum_{i=1}^{r}\sigma_i^2(\delta W)\,,
-		\end{equation*}
-		\begin{equation}\label{eqn:f_min_F_star}
-			f(\delta W) \geq \frac{1}{\lambda}\min_{d_1,\dots,d_r \geq 0}-\sum_{i=1}^r \sigma_i(G)d_i + \frac{1}{2}\left(\sum_{i=1}^{r}d_i\right)^2 + \frac{1}{2}\left(\sum_{i=1}^{r}d_i^2\right)\,,
-		\end{equation}
-		By Karush-Kuhn-Tucker theorem, neccesary conditions of minimum are
-		\begin{equation*}
-			\left(d_i \geq 0\ \text{and}\ -\sigma_i(G) + \sum_{j=1}^{r}d_j +  d_i = 0\right)\ \text{or}\ \left(d_i = 0\ \text{and}\ -\sigma_i(G) + \sum_{j=1}^{r}d_j + d_i \geq 0\right) \,\, i = 1,\dots,r\,.
-		\end{equation*}
-		Denoting $\tau = \sum_{i = 1}^r d_i$ gives $d_i = [\sigma_i(G) - \tau]_+$, where $\tau$ satisfies
-		\begin{equation}
-			\sum_{i=1}^n [\sigma_i(G) - \tau]_+ = \tau\,.
-		\end{equation}
-		Inserting found minimum point into~\eqref{eqn:f_min_F_star} yields
-		\begin{equation*}
-			f(\delta W) \geq -\sum_{i = 1}^r d_i(\tau + d_i) + \frac{\tau^2}{2\lambda} + \frac{1}{2\lambda}\sum_{i=1}^{r}d_i^2 = -\frac{1}{2 \lambda} \left( \tau^2 + \sum_{i=1}^r d_i^2 \right)\,.
-		\end{equation*}
-
-		Now let
-		\begin{equation}
-			\delta W^* = -\frac{1}{\lambda}UDV^T
-		\end{equation}
-		with $D = \text{diag}(d_i)$. Inserting it to $f(\delta W)$ gives
-		\begin{equation*}
-			f(\delta W^*) = -\sum_{i = 1}^r d_i(\tau + d_i) + \frac{\tau^2}{2\lambda} + \frac{1}{2\lambda}\sum_{i=1}^{r}d_i^2 = -\frac{1}{2 \lambda} \left( \tau^2 + \sum_{i=1}^r d_i^2 \right)\,.
-		\end{equation*}
-		This matches the derived lower bound. Thus, $\delta W^*$ minimizes $f(\delta W)$.
-	\end{proof}
-
-	\section{Quality metrics}
-	\begin{enumerate}
-		\item The derivation is theoretically solid
-		\item The numerical procedure used to compute a step is grounded and has estimated time overhead (say, in FLOPS)
-		\item The code with Neon trains MLP and CNN (and NanoGPT, but it's a bonus) less than 3 times slower than Adam
-		\item Instruction of setting the parameters of the algorithm are presented and justified
-		\item The announced article has full structure (Abstract, Introduction, Theory, Experiments, Conclusion, Appendix)
-		\item If results are positive, it is written with NeurIPS template.
-	\end{enumerate}
-
-	\section{Preliminary plan}
-	\paragraph{Week April 28 - May 4}
-	\begin{itemize}
-		\item For Alexey: solve how to tune the algorithms for MLP and CNN, try formulating theory (and an appropriate model of the problem) why Muon and Neon are so successful, and create the drafts of the proofs.
-Register at NeurIPS site.
-
-		\item For Ivan: write the theory for an update from the algebra point of view (as for an article)
-
-		\item For Nikolay: write the theory for computing an update, and implement the method, if required
-
-		\item For Alexander: reproduce results of Jordan on NanoGPT and ResNet (CIFAR-10), learn to train both models with Neon.
-	\end{itemize}
-
-	\paragraph{Week May 5 - May 11}
-	\begin{itemize}
-		\item For Alexey: finalize the proofs. Verify them via small experiments on MLP and CNN. Write with Alexander Experiments for the article.
-
-		\item For Ivan: join Nikolay to finalize algebra part of the article. Estimate FLOPS, memory and other overheads (produce O bounds)
-
-		\item For Nikolay: write a draft of the poster (before May 6), and work with Ivan
-
-		\item For Alexander: agressively test algorithms, prove that Neon outperforms competitors and prepare the results for the article.
-
-		\item For everybody: write and edit the article
-
-		\item May 11: submit an abstract to NeurIPS.
-
-		\item May 12-14: the article is being polished.
-
-		\item May 15: the article must be sent.
-	\end{itemize}
-
-	\section{Prototyping phase report}
-	\begin{enumerate}
-		\item Update rule is derived, see idea
-		\item Update rule methods are tested: power iteration vs Lanczos (see \ref{tab:matrix_methods})
-		\item Recorded the distribution of singular values of gradients during NanoGPT training (see Firgures ~\ref{fig:svd_all}, \ref{fig:svd_all}).
-		      \begin{figure}[h!]
-			      \center{\includegraphics[width=0.8\linewidth]{figs/mlp24/svd_evolution_all.png}}
-			      \caption{Singular values of $50257\times 1280$ layer via 200 iterations}
-			      \label{fig:svd_all}
-		      \end{figure}
-		      \begin{figure}[h!]
-			      \center{\includegraphics[width=0.8\linewidth]{figs/mlp24/svd_evolution.png}}
-			      \caption{Singular values of  of $50257\times 1280$ layer for 5-th,45-th, 65-th,175-th and 200-th iteration}
-			      \label{fig:svd}
-		      \end{figure}
-
-		\item NanoGPT is tested on Muon and Adam. For now, Neon (rank-1 version) does not converge (see Figures ~\ref{fig:train_loss} \ref{fig:val_loss})
-		      \begin{figure}[h!]
-			      \center{\includegraphics[width=0.8\linewidth]{figs/mlp24/loss_train.png}}
-			      \caption{Train loss}
-			      \label{fig:train_loss}
-		      \end{figure}
-		      \begin{figure}[h!]
-			      \center{\includegraphics[width=0.8\linewidth]{figs/mlp24/loss_val.png}}
-			      \caption{Validation loss}
-			      \label{fig:val_loss}
-		      \end{figure}
-
-		      The pictures show the best results achieved so far. The experiments were conducted with two 4090 24GB GPUs for nanotgpt-large on the tiny stories dataset.
-
-		\item Neon (rank-1 version), Muon, AdamW and SGD are compared on MLP and CNN (see Figures~\ref{fig:mlp_epochs}, \ref{fig:cnn_epochs}, \ref{fig:mlp_time}, and~\ref{fig:cnn_time}).
-All methods work correctly, but again there is the problem with which one is the fastest (for now, it's SGD).
-		      \begin{figure}[h!]
-			      \center{\includegraphics[width=0.8\linewidth]{figs/mlp0605/loss_vs_epochs_mlp.pdf}}
-			      \caption{MLP: self.linear1 = nn.Linear(32*32*3, 512), self.linear2 = nn.Linear(512, 10), self.activ = nn.GELU()}
-			      \label{fig:mlp_epochs}
-		      \end{figure}
-		      \begin{figure}[h!]
-			      \center{\includegraphics[width=0.8\linewidth]{figs/mlp24/accuracy_vs_epochs_moderate.png}}
-			      \caption{CNN: 2 convolutional blocks, 2 fully connected layers, activation + dropout}
-			      \label{fig:cnn_epochs}
-		      \end{figure}
-		      \begin{figure}[h!]
-			      \center{\includegraphics[width=0.8\linewidth]{figs/mlp0605/accuracy_vs_time_mlp.pdf}}
-			      \caption{MLP: wallclock time measurements}
-			      \label{fig:mlp_time}
-		      \end{figure}
-		      \begin{figure}[h!]
-			      \center{\includegraphics[width=0.8\linewidth]{figs/mlp24/accuracy_vs_time_moderate.png}}
-			      \caption{CNN: wallclock time measurements}
-			      \label{fig:cnn_time}
-		      \end{figure}
-		      \begin{table}[h!]
-			      \centering
-			      \begin{tabular}{c|c|c|c}
-				      \hline
-				      Method                              & rtol & k   & time,s \\
-				      \hline
-				      Power Iterations                    & 0.01 & 1   & 7.7    \\
-				      SVDS (thick-restart Lanczos method) & 0.01 & 1   & 0.18   \\
-				      PCA Low Rank (RSVD)                 & 0.01 & 1   & 1.15   \\
-				      SVDS (thick-restart Lanczos method) & 0.01 & 10  & 0.47   \\
-				      PCA Low Rank (RSVD)                 & 0.01 & 10  & 19.4   \\
-				      SVDS (thick-restart Lanczos method) & 0.01 & 100 & 1.96   \\
-				      PCA Low Rank (RSVD)                 & 0.01 & 100 & 170    \\
-			      \end{tabular}
-			      \caption{k-rank updated comparison}
-			      Comparison of different numerical methods to calculate k-rank update on $5000\times5000$ matrix of real numbers, rtol is an error in Frobenius norm relative to the k-rank approximation of truncated svd.
-During the research it was noted that rsvd can give good and fast approximation for singular values, but the matrix of approximation is far from the one given by truncated svd, while Lanczos method gives good and fast approximation for a matrix, but not so good approximation for singular values.
-			      \label{tab:matrix_methods_old}
-		      \end{table}
-	\end{enumerate}
-
-	\appendix
-	\section{Old Appendix}
-	You may include other additional sections here.
-
-\fi
-
 \end{document}

### 2025-11-06 code/cifar_bench/base_airbench.py
Added norm logger for CIFAR

@@ -358,7 +358,7 @@ def main(run, model):
     optimizer1 = torch.optim.SGD(param_configs, momentum=0.85, nesterov=True)#, fused=True)
     # random mix, 93.3%, 11.26 s on bs 2000 with lr=0.4, mom=0.65
     # optimizer2 = RandomNormalizedMuon(filter_params, lr=0.24, momentum=0.6, sgd_coeff=0.5, nesterov=True) # and 92.9% for bs 200
-    # optimizer2 = NormalizedMuon(filter_params, lr=0.4, momentum=0.65, sgd_coeff=0.5, nesterov=True) # the best tuned F-Muon, 94.0%
+    optimizer2 = NormalizedMuon(filter_params, lr=0.4, momentum=0.65, sgd_coeff=0.5, nesterov=True) # the best tuned F-Muon, 94.0%
     # optimizer2 = MuonCringeMomentum(filter_params, lr=0.24, momentum=0.6, nesterov=True) # it's not bad, but without Nesterov=true it does not work
     # optimizer2 = Muon(filter_params, lr=0.24, momentum=0.6, nesterov=True) # base Muon, 94.01% 11.4 s
     switch_to_muon = True
@@ -388,7 +388,7 @@ def main(run, model):
     # optimizer2 = Dion(filter_params, lr=0.45, momentum=0.65, rank=20, momentum_decay=0.9, sgd_coeff=0) # 89.3%, with 0.2% variance
     
 
-    optimizer2 = SignSGDMuon(filter_params, lr=0.4, momentum=0.65, nesterov=True, sgd_coeff=0.5) # 94%, on par with Muon
+    # optimizer2 = SignSGDMuon(filter_params, lr=0.4, momentum=0.65, nesterov=True, sgd_coeff=0.5) # 94%, on par with Muon
     # optimizer2 = SGDMuon(filter_params, lr=0.24, momentum=0.6, nesterov=True, sgd_coeff=0.1) # 87% - does not work well, because it's not an LMO algorithm
     
     # optimizer2 = ErrorFeedbackMuon(filter_params, lr=0.24, momentum=0.6, nesterov=True, sgd_coeff=0, error_feedback_decay=0.9) # 89.6%, unfeasible
@@ -421,16 +421,18 @@ def main(run, model):
     stop_timer()
 
     for epoch in range(ceil(total_train_steps / len(train_loader))):
-
         ####################
         #     Training     #
         # ####################
-        # if epoch == 8 and switch_to_muon: # to go to Muon
+        # no gain:
[…]

### 2025-11-06 code/cifar_bench/norm_logger_airbench.py

@@ -10,13 +10,16 @@ import sys
 with open(sys.argv[0]) as f:
     code = f.read()
 import uuid
+import random
 from math import ceil
+from datetime import datetime
 
 import torch
 from torch import nn
 import torch.nn.functional as F
 import torchvision
 import torchvision.transforms as T
+import pandas as pd
 from optimizers import Dion, Muon, Neon, NormalizedMuon, SGDMuon, SignSGDMuon, zeropower_via_newtonschulz5, RandomNormalizedMuon
 
 
@@ -331,36 +334,176 @@ def evaluate(model, loader, tta_level=0):
     return (logits.argmax(1) == loader.labels).float().mean().item()
 
 ############################################
+#            Seed Setting                  #
+############################################
+
+def set_seed(seed=42):
+    """Set random seeds for reproducibility."""
+    random.seed(seed)
+    torch.manual_seed(seed)
+    torch.cuda.manual_seed(seed)
+    torch.cuda.manual_seed_all(seed)
+    # Set deterministic behavior for CUDA operations
+    torch.backends.cudnn.deterministic = True
+    torch.backends.cudnn.benchmark = False
+    # Set seed for numpy if available
+    try:
+        import numpy as np
+        np.random.seed(seed)
+    except ImportError:
+        pass
+
+############################################
 #                Training                  #
 ############################################
-# TODO: create a pandas dataframe out of these values:
 
-def log_grad_frobenius_norms(model, step=None, logger=None, silent=False):
-    norms = {}
-    total_sq = 0.0
+def log_grad_frobenius_norms(model, step=None, epoch=None, loss=None, outputs=None, labels=None, logger=None, silent=False, norm_data_list=None):
+    """
+    Compute and log Frobenius, spectral, and nuclear norms of gradients.
+    
+    Args:
+        model: The model to compute gradients for
+        step: Current training step
+        epoch: Current epoch
+        loss: Current loss value (optional)
+        outputs: Model outputs for computing train accuracy (optional)
+        labels: Ground truth labels for computing train accuracy (optional)
+        logger: Optional logger (e.g., tensorboard)
+        silent: If True, don't print
[…]

### 2025-11-06 code/cifar_norms/cifar_norm_plotter.py

@@ -1,0 +1,206 @@
+import pandas as pd
+import matplotlib.pyplot as plt
+import os
+from pathlib import Path
+from typing import Dict, List, Union
+
+
+def plot_norm_comparison(
+    log_files: Dict[str, str],
+    metrics: List[str],
+    x_axis: str = "step",
+    output_dir: str = None,
+    figsize: tuple = (10, 6),
+    dpi: int = 300
+):
+    """
+    Plot and compare different norm logs.
+    
+    Args:
+        log_files: Dictionary mapping title/name to CSV file path
+                   e.g., {"Muon": "grad_norms_Muon_run0.csv", "NormalizedMuon": "grad_norms_NormalizedMuon_run0.csv"}
+        metrics: List of metrics to plot. Can be:
+                 - Exact column names (e.g., "total_frobenius", "layers.1.conv1.weight_nuclear")
+                 - Partial matches (e.g., "nuclear norm for conv1" will match columns containing "conv1" and "nuclear")
+                 - Pattern matches (e.g., "total_*" for all total metrics)
+        x_axis: Column to use for x-axis (default: "step", can also use "epoch")
+        output_dir: Directory to save PDF files (default: same directory as first log file)
+        figsize: Figure size tuple (width, height)
+        dpi: Resolution for saved figures
+    """
+    # Load all dataframes
+    dataframes = {}
+    for title, filepath in log_files.items():
+        if not os.path.exists(filepath):
+            raise FileNotFoundError(f"Log file not found: {filepath}")
+        df = pd.read_csv(filepath)
+        # Remove empty rows
+        df = df.dropna(how='all')
+        dataframes[title] = df
+    
+    # Determine output directory
+    if output_dir is None:
+        first_file = list(log_files.values())[0]
+        output_dir = os.path.dirname(os.path.abspath(first_file))
+    os.makedirs(output_dir, exist_ok=True)
+    
+    # Get all available columns (excluding x_axis and loss)
+    # Note: train_acc and val_acc are included and will be plotted separately
+    all_columns = set()
+    for df in dataframes.values():
+        all_columns.update(df.columns)
+    all_columns.discard(x_axis)
+    all_columns.discard('loss')
+    # Keep epoch, train_acc, and val_acc as they are valid metrics to plot
+    
+    # Resolve metrics to actual column names
+    resolved_metrics = []
+    for metric in metrics:
+        matched_columns = _resolve_metric(metric, all_columns)
+        if not matched_columns:
[…]

### 2025-11-06 code/optimizers.py

@@ -1,7 +1,7 @@
 import math
 import torch
 import torch.nn.functional as F
-from torch import nn, Tensor
+from torch import nn, Tensor, nuclear_norm
 import torch.distributed as dist
 from matrix_methods.matrix_functions import k_sv_svds_approximation_dlpack, one_sv_svds_approximation, svd_full_approximation, several_sv_svds_approximation
 
@@ -27,6 +27,27 @@ def zeropower_via_newtonschulz5(G, steps=3, eps=1e-12):
     if G.size(0) > G.size(1):
         X = X.T
     return X
+
+def zeropower_via_newtonschulz5_with_fro_norm(G, steps=3, eps=1e-12):
+    """Simplified Newton-Schulz iteration for whitening"""
+    assert len(G.shape) == 2
+    a, b, c = (3.4445, -4.7750, 2.0315)
+    X = G.bfloat16()
+    # Add numerical stability
+    norm = X.norm() + eps
+    if norm < eps:
+        print("Norm lower than eps")
+        return torch.zeros_like(X)
+    X /= norm
+    if G.size(0) > G.size(1):
+        X = X.T
+    for _ in range(steps):
+        A = X @ X.T
+        B = b * A + c * A @ A
+        X = a * X + B @ X
+    if G.size(0) > G.size(1):
+        X = X.T
+    return X, norm
 
 
 def newtonschulz5_with_nuclear(G, steps=3, eps=1e-12):
@@ -259,13 +280,27 @@ class NormalizedMuon(torch.optim.Optimizer):
                     norm = 1e-10
                 p.data.mul_(len(p.data)**0.5 / norm) # normalize the weight
                 # update = zeropower_via_newtonschulz5(g.reshape(len(g), -1)).view(g.shape) # whiten the update
-                g_normalized = g / (g.norm() + eps)
                 if self.sgd_coeff != 1:
-                    update_part = zeropower_via_newtonschulz5(g.reshape(len(g), -1)).view(g.shape)
-                    update = (1-self.sgd_coeff) * update_part + self.sgd_coeff * g_normalized
+                    # zeropower expects 2D input: reshape and restore
+                    update_part, fro_norm = zeropower_via_newtonschulz5_with_fro_norm(
+                        g.reshape(g.size(0), -1)
+                    )
+                    update_part = update_part.view_as(g)
+
+                    # normalize g by fro_norm in-place (safe after update_part computed)
+                    # If you prefer not to mutate g, use g_div = g / fro_norm instead.
+                    g.div_(fro_norm)
[…]

### 2025-11-06 code/cifar_bench/base_airbench.py
CIFAR Norm Logger stores and plots all the data

@@ -424,7 +424,6 @@ def main(run, model):
         ####################
         #     Training     #
         # ####################
-        # no gain:
         # if epoch == 5 and switch_to_muon: # to go to Muon
         #     # optimizers[1] = NormalizedMuon(filter_params, lr=0.4, momentum=0.65, nesterov=True, sgd_coeff=0.5)
         #     # optimizers[1].sgd_coeff = 0.5

### 2025-11-06 code/cifar_bench/norm_logger_airbench.py

@@ -238,7 +238,7 @@ class CifarNet(nn.Module):
                 mod.half()
 
     def reset(self):
-        for m in model.modules():
+        for m in self.modules():
             if type(m) in (nn.Conv2d, Conv, BatchNorm, nn.Linear):
                 m.reset_parameters()
         w = self.head.weight.data
@@ -459,7 +459,7 @@ def log_grad_frobenius_norms(model, step=None, epoch=None, loss=None, outputs=No
     
     return (frobenius_norms, spectral_norms, nuclear_norms), (total_frobenius, total_spectral, total_nuclear)
 
-def save_norm_dataframe(norm_data_list, run=None, optimizer_name=None, epoch_val_acc=None):
+def save_norm_dataframe(norm_data_list, run=None, optimizer_name=None, combo_name=None, epoch_val_acc=None):
     """
     Convert norm data list to DataFrame and save to cifar_norms/ folder.
     
@@ -467,6 +467,7 @@ def save_norm_dataframe(norm_data_list, run=None, optimizer_name=None, epoch_val
         norm_data_list: List of dictionaries containing norm data
         run: Run identifier (optional)
         optimizer_name: Optimizer name for filename (optional)
+        combo_name: Combo name for filename (optional, takes precedence over optimizer_name)
         epoch_val_acc: Dictionary mapping epoch -> val_acc to backfill (optional)
     
     Returns:
@@ -487,10 +488,12 @@ def save_norm_dataframe(norm_data_list, run=None, optimizer_name=None, epoch_val
     output_dir = 'cifar_norms'
     os.makedirs(output_dir, exist_ok=True)
     
-    # Create meaningful filename
+    # Create meaningful filename - use combo_name if provided, otherwise optimizer_name
     timestamp = datetime.now().strftime("%Y%m%d_%H%M%S")
     filename_parts = ['grad_norms']
-    if optimizer_name:
+    if combo_name:
+        filename_parts.append(combo_name)
+    elif optimizer_name:
         filename_parts.append(optimizer_name)
     if run is not None:
         filename_parts.append(f'run{run}')
@@ -505,7 +508,19 @@ def save_norm_dataframe(norm_data_list, run=None, optimizer_name=None, epoch_val
     
     return filepath
 
[…]

### 2025-11-06 code/cifar_norms/cifar_norm_plotter.py

@@ -180,16 +180,24 @@ def _sanitize_filename(filename: str) -> str:
     filename = filename.replace('.', '_').replace(' ', '_')
     return filename
 
+import re
+
+def collect_grad_norm_files(folder_path):
+    pattern = re.compile(r"grad_norms_(.*?)_run\d+_\d{8}_\d{6}\.csv$")
+    result = {}
+
+    for filename in os.listdir(folder_path):
+        if filename.endswith(".csv"):
+            match = pattern.match(filename)
+            if match:
+                key = match.group(1)
+                result[key] = filename
+
+    return result
 
 if __name__ == "__main__":
     # Example usage
-    log_files = {
-        "Muon": "grad_norms_Muon_run0_20251106_222819.csv",
-        "NormalizedMuon": "grad_norms_NormalizedMuon_run0_20251106_223129.csv",
-        "Neon 0.4": "grad_norms_Neon_run_lr_0_4.csv",
-        "Neon 0.04": "grad_norms_Neon_run_lr_004.csv",
-        "Neon 4": "grad_norms_Neon_run_lr_4.csv"
-    }
+    log_files = collect_grad_norm_files(".")
     
     metrics = [
         "total_frobenius",

### 2025-11-09 conf_article/neon.tex
Article updated

@@
 Daniil Merkulov \\
-Moscow Institute of Physics and Technology (MIPT), Skoltech, HSE, Sber \\
+Moscow Institute of Physics and Technology (MIPT), Skoltech, HSE, AI4Science \\
 \texttt{daniil.merkulov@phystech.edu} \\
@@
 \begin{abstract}
-	% In this article, we explore the use of matrix norms for optimizing functions of weight matrices, a crucial problem in training large language models.
-Moving beyond the spectral norm that underlies the Muon update, we leverage the Ky Fan k-norms, their affine combinations with the Frobenius norms, and corresponding duals to develop a new family of Muon-like algorithms we name {\it Fanions}.
-We complement our theoretical analysis with an extensive empirical study of the algorithms across a wide range of tasks and settings.
-
-	In this article, we explore the use of matrix norms for optimizing functions of weight matrices, a crucial problem in training large language models.
-Moving beyond the spectral norm that underlies the Muon update, we leverage the duals to the Ky Fan k-norms to introduce a \deleted[id=IK, comment={Redundant}]{new} family of Muon-like algorithms we name {\it Fanions}, which happen to be similar to Dion.
-\replaced[id=IK, comment={This rewrite clarifies the sentence structure, making it clear that the norms (not the algorithms) are being combined.}]{Subsequently, we construct a second family of algorithms, dubbed {\it F-Fanions}, based on the duals of convex combinations of Ky Fan k-norms and the Frobenius norm.
-One prominent member of this family is {\it F-Muon}.}{Then, we consider their convex combinations with the Frobenius norm and corresponding duals to develop a new family of algorithms we name {\it F-Fanions}, one of them being {\it F-Muon}.} We complement our theoretical analysis with an extensive empirical study of the algorithms across a wide range of tasks and settings, from which it follows that properly-tuned F-Muon is almost on par with Muon, \replaced[id=IK, comment={Alex, clarify this please.
-It's a very non-obvious and non-clear conclusion, especially when knowing that Muon itself only provides an approximate answer to the LMO.
-I provided an alternative phrase.}]{which questions the necessity of finding the exact solution to the problem in linear optimization oracle framework.}{which shows the flexibility in the choice of the norm and opens possibilities for easier-to-compute variants.}
+	In this article, we explore the use of various matrix norms for optimizing functions of weight matrices, a crucial problem in training large language models.
+Moving beyond the spectral norm that underlies the Muon update, we leverage the duals to the Ky Fan k-norms to introduce a family of Muon-like algorithms we name {\it Fanions}, which happen to be similar to Dion.
+Subsequently, we construct a second family of {\it F-Fanions}, which are based on the duals of convex combinations of Ky Fan k-norms and the Frobenius norm.
+One prominent member of this family is {\it F-Muon}.
+We complement our theoretical analysis with an extensive empirical study of the algorithms across a wide range of tasks and settings, from which it follows that F-Muon is on par with Muon, which questions the exclusivity of a spectral norm.
 
@@
 That is what can be said from a practical point of view.
-\replaced[id=IK, comment={Rewrote this to be more precise.
-The innovation wasn't just 'using' a norm, but deriving the update from a principled, constrained optimization problem.
-I think this better reflects the contribution of the Muon paper.}]{From a theoretical perspective, a key innovation of Muon was its principled derivation of the update rule, which emerges as the solution to an optimization problem constrained by the RMS-to-RMS norm (scaled version of spectral norm) \citep{bernstein2025deriving}}{From the perspective of theory, the main innovation of Muon has been an intentional usage of matrix norms, i.e.
-the spectral norm, to derive the algorithm's update}.
-
-\replaced[id = IK, comment={Subsequent paragraphs in the introduction are very heavily modified.}]{}{}
+From a theoretical perspective, a key innovation of Muon was its principled derivation of the update rule, which emerges as the solution to an optimization problem constrained by the RMS-to-RMS norm (scaled version of spectral norm) \citep{bernstein2025deriving}
 
@@
 Subsequently, we create a second, hybrid family named F-Fanions by constructing convex combinations of the Ky Fan norms with the Frobenius norm and taking dual of that composite norm.
-Working within the linear minimization oracle (\replaced[id = IK, comment={I think it's better this way.}]{LMO}{lmo}) framework we derive the explicit update formulas for both algorithm families.
+Working within the linear minimization oracle (LMO) framework we derive the explicit update formulas for both algorithm families.
 As it was done with Muon, we stipulate our algorithms to be fast to approximate, for which we utilize the Lanczos algorithm as described in \cref{sec:matrix-side}.
@@
 
-\section{\added[id = IK, comment={I think it goes well into one big section.}]{Update Step as Constrained Optimization}}
+\section{Update Step as Constrained Optimization}
 
-\subsection{\replaced[id = IK, comment={Merged what previously had been section with two subsections into this one and improved the notation.
-Also removed the thing about trust region, it'll be referenced in succeeding sections.}]{Linear Minimization Oracle Framework}{Problem Statement}}
+\subsection{Linear Minimization Oracle Framework}
 
-Training a neural network involves optimizing a function of weight matrices. For simplicity, we consider the problem of minimizing a differentiable function of a single matrix:
-\begin{equation}\label{eq:mat}
-	F(\cdot)\colon \Rmn \to \R\,,\qquad F(\mX) \to \min_{\mX \in \Rmn}\,.
+Training a neural network is essentially an optimization of a function of several weight matrices.
+Let us start by considering the problem of minimizing a differentiable function of a {\it single} matrix:
+\begin{equation}\label{eq:matrix_problem}
+	F(\cdot)\colon \Rmn \to \R\,,\qquad F(\mX) \to \min_{\mX \in \Rmn}\,
 \end{equation}
-We equip the matrix space $\Rmn$ with a standard dot product $\<\cdot, \cdot> \to \R$ and a norm $\norm{\cdot}\colon \Rmn \to \R_+$, which possibly does not coincide with the Frobenius norm $\normf{\cdot}$.
+We equip the matrix space $\Rmn$ with a standard dot product $\<\cdot, \cdot> \to \R$ and a norm $\norm{\cdot}\colon \Rmn \to \R_+$, which does not have to coincide with the Frobenius norm $\normf{\cdot}$.
 The dual norm $\norm{\cdot}^\dagger\colon \Rmn \to \R_+$ associated with $\norm{\cdot}$ is defined as
 \begin{equation}\label{eq:dual_norm}
-	\norm{\mG}^\dagger = \sup_{\mD \in \Rmn\,,\norm{\mD}\leq 1} \<\mG,\mD>\,.
+	\norm{\mG}^\dagger = \sup_{\mD \in \Rmn\,:\norm{\mD}\leq 1} \<\mG,\mD>\,.
 \end{equation}
 
-One approach to optimizing such functions is to employ an iterative algorithm based on Linear Minimization Oracle (LMO):
+Such problems can be solved with an iterative algorithm based on the Linear Minimization Oracle (LMO):
 \begin{equation}\label{eq:lmo}
@@
 \begin{equation}\label{eq:simple_update}
-	\mX^{k + 1} = \mX^k + \mathrm{LMO}\left(\nabla F\left(\mX^k\right)\right)\,.
+	\mX^{k + 1} = \mX^k + \gamma_k\mathrm{LMO}\left(\nabla F\left(\mX^k\right)\right)\,.
 \end{equation}
 
-In a typical neural network, the objective function $F$ depends on a set of weight matrices $\{ \mX_1, \mX_2, \ldots \}$.
+In a typical neural network, the objective function $F$ depends on a set of weight matrices $\{ \mW_1, \mW_2, \ldots \}$.
 The optimization framework we have described is applied in a layer-wise fashion.
-At each iteration $k$, a stochastic gradient $g(\mX^k, \xi^k)$ is computed using a mini-batch of data $\xi^k$ via backpropagation.
+At each iteration $k$, a stochastic gradient $g(\mX^k, \xi^k)$ is computed using a mini-batch of data with the $\xi^k$ noise via backpropagation.
 This yields a separate gradient component, $\mG_i^k$, for each matrix $\mX_i$.
@@
 Greedy Coordinate Descent methods are not applied to high-dimensional problems, from this perspective, it is not surprising that Neon underperforms on such problems.
-
-\deleted[id = IK, comment={This is essentially like saying uh oh the optimization in multy-dimensional space is so hard, why there is no algorithm to reduce this problem to 1-d case.}]{Despite such similarities, no theory has been yet proposed that would reduce the matrix case to the optimization of a function of a singular values vector.
-If such a theory is developed, analysis of matrix algorithms like Muon will be greatly simplified.}
 
@@
 
-The results are curious and could be represented by \cref{fig:cifar_ball}: lmo ball, which we plotted in a 2D space of singular values, has drastically changed, but the observed convergence after tuning has not degraded.
+The results are curious and could be represented by \cref{fig:cifar_ball}: the LMO ball that we plotted in a 2D space of singular values has drastically changed, but the observed convergence after tuning has not degraded.
 These observations raise the question of how much lmo-based algorithms are sensitive to the constraint area, i.e.
@@
 	\includegraphics[width=0.6\textwidth]{figs/norms/fstardual_cifar.pdf} % adjust width as needed
-	\caption{Visualization of lmo balls for Muon and F-Muon.}
+	\caption{Visualization of the LMO balls for Muon and F-Muon.}
 	\label{fig:cifar_ball}
@@
 
-The most pathological case is \replaced[id=IK, comment={In the pictures there are no experiments with $\alpha < 0$, but there is a one with $\alpha $}]{$\alpha > 1$}{$\alpha < 0$}, which corresponds to the lmo with an area that is not a norm ball!
+The most pathological case is $\alpha > 1$, which corresponds to the lmo with an area that is not a norm ball!
 Despite this violation, the mixture of algorithms reaches almost the same accuracy as vanilla Muon.

### 2025-11-10 conf_article/neon.tex
Major revision of the article, now raw but better structured.

@@
 Moving beyond the spectral norm that underlies the Muon update, we leverage the duals to the Ky Fan k-norms to introduce a family of Muon-like algorithms we name {\it Fanions}, which happen to be similar to Dion.
-Subsequently, we construct a second family of {\it F-Fanions}, which are based on the duals of convex combinations of Ky Fan k-norms and the Frobenius norm.
+Subsequently, we construct a second family of {\it F-Fanions}, which are based on the duals of convex combinations of the Ky Fan k-norms and the Frobenius norm.
 One prominent member of this family is {\it F-Muon}.
@@
 Working within the linear minimization oracle (LMO) framework we derive the explicit update formulas for both algorithm families.
-As it was done with Muon, we stipulate our algorithms to be fast to approximate, for which we utilize the Lanczos algorithm as described in \cref{sec:matrix-side}.
+As it was done with Muon, we stipulate our algorithms to be fast to approximate, for which we utilize the Lanczos algorithm as described in ~\cref{sec:matrix-side}.
 
@@
 
-\section{Update Step as Constrained Optimization}
-
-\subsection{Linear Minimization Oracle Framework}
+\section{Preliminaries: Linear Minimization Oracle Framework}
 
@@
 
-In a typical neural network, the objective function $F$ depends on a set of weight matrices $\{ \mW_1, \mW_2, \ldots \}$.
-The optimization framework we have described is applied in a layer-wise fashion.
-At each iteration $k$, a stochastic gradient $g(\mX^k, \xi^k)$ is computed using a mini-batch of data with the $\xi^k$ noise via backpropagation.
-This yields a separate gradient component, $\mG_i^k$, for each matrix $\mX_i$.
-The LMO-based update rule is then applied to each matrix $\mX_i$ using its corresponding gradient component $\mG_i^k$.
-
-The update rule used in Muon optimizer is uSCG \citep{pethick2025training}. In the most general case, which involves momentum, it can be written as
-\begin{equation}\label{eq:usgd_update}
-	\begin{aligned}
-		\mM^{k}     & = \alpha_{k} g(\mX^k, \xi^k) + (1 - \alpha_k)\mM^{k - 1}\,, \\
-		\mX^{k + 1} & = \mX^k + \gamma_k\mathrm{LMO}(\mM^{k})\,.
-	\end{aligned}
-\end{equation}
-
 Here we are particularly interested in the case when $\cS$ is a ball in the $\norm{\cdot}$ norm:
@@
 
-\subsection{Different Norms $\norm{\cdot}$ imply different updates}
-
-In this subsection we will derive and discuss updates induced by the choice of the norm in \cref{eq:our_update}.
+In the next subsection we will derive and discuss updates induced by the choice of the norm in \cref{eq:our_update}.
 To describe those updates, the singular value decomposition (SVD) of $\mM^k$ is required.
@@
 
-\subsection{\texorpdfstring{$\normf{\mM^{k}}$ and Normalized SGD}{NSGD}}
+When $\norm{\cdot} = \norms{\cdot}$, \cref{eq:our_update} turns into:
 
-\begin{lemma}\label{lemma:nsgd_update}
-	When $\norm{\cdot} = \normf{\cdot}$, \cref{eq:our_update} turns into:
+\begin{equation}\label{eq:muon_update}
+	\mX^{k+1} = \mX^{k} - \eta U V^\top
+\end{equation}.
+
+And when $\norm{\cdot} = \normf{\cdot}$, \cref{eq:our_update} turns into:
 
@@
 
-\end{lemma}
-
-The result is the same as in (Table 1, \citet{pethick2025training}), but here we use the Euclidean norm as $\normf{\mM^k} = \sqrt{\sum_{i=1}^{\min\{m,n\}}\sigma_i^2}$, which clearly is a matrix norm.
-
-\subsection{\texorpdfstring{$\norms{\mM^{k}}$ and Muon}{Muon}}
-\begin{lemma}\label{lemma:muon_update}
-	When $\norm{\cdot} = \norms{\cdot}$, \cref{eq:our_update} turns into:
-
-	\begin{equation}\label{eq:muon_update}
-		\mX^{k+1} = \mX^{k} - \eta U V^\top
-	\end{equation}
-\end{lemma}
-
-Though the update is well-known, we provide a much simpler proof in the appendix, when compared to \citet{bernstein2024oldoptimizernewnorm}.
+\section{Duals to Ky Fan norms instead of the spectral norm}
+%\subsection{Non-operator Norms $\norm{\cdot}$ lead to new updates}
 
@@
 \end{lemma}
-We name the derived algorithm \emph{Neon}. In the section {\it The Matrix side of updates}, we will discuss how to compute the update efficiently.
-
-\subsection{\texorpdfstring{$\normfstar{\mM^{k}}^\dagger$ and F-Muon}{F-Muon}}
-We define $\normfstar{\cdot}$ as a convex combination of $\normn{\cdot}$ and $\normf{\cdot}$:
-
-\begin{equation}\label{eq:fstar_norm}
-	\normfstar{\mX} = \alpha \normn{\mX} + (1-\alpha)\normf{\mX},
-\end{equation}
-
-where $\alpha \in [0, 1]$ defines a specific norm from the $F*$-family.
-$\normfstar{\cdot}^\dagger$ can be expressed as in \cref{eq:normfstardual}.
-However, its norm ball is a simple Minkowski sum of $\alpha \normn{\cdot}$ and $(1-\alpha)\normf{\cdot}$ balls \cref{fig:fstardual}.
-
-\begin{lemma}\label{lemma:nsgd_muon_update}
-	When $\norm{\cdot} = \normfstar{\cdot}^\dagger$, \cref{eq:our_update} turns into:
-
-	\begin{equation}\label{eq:nsgd_muon_update}
-		\mX^{k+1} = \mX^{k} - \eta (\alpha U V^\top + (1-\alpha)\frac{\mM^k}{\normf{\mM^k}})
-	\end{equation}
-\end{lemma}
-
-We name the derived algorithm \emph{F-Muon}. It turns out that F-Muon is a convex combination of Normalized SGD and Muon. The implications are significant and discussed in the following sections.
-
-\subsection{\texorpdfstring{$\normftwo{\mM^{k}}^\dagger$ and F-Neon}{F-Neon}}
-We define $\normftwo{\cdot}$ as a convex combination of $\norms{\cdot}$ and $\normf{\cdot}$:
-
-\begin{equation}\label{eq:ftwo_norm}
-	\normftwo{\mX} = \alpha \norms{\mX} + (1-\alpha)\normf{\mX},
-\end{equation}
-
-where $\alpha \in [0, 1]$ defines a specific norm from the F2-family.$\normftwo{\cdot}^\dagger$ can be expressed as in \cref{eq:normftwodual}.
-However, its norm ball is a simple Minkowski sum of $\alpha \norms{\cdot}$ and $(1-\alpha)\normf{\cdot}$ balls \cref{fig:ftwodual}.
-
-\begin{lemma}\label{lemma:nsgd_neon_update}
-	When $\norm{\cdot} = \normftwo{\cdot}^\dagger$, \cref{eq:our_update} turns into:
-
-	\begin{equation}\label{eq:nsgd_neon_update}
-		\mX^{k+1} = \mX^{k} - \eta (\alpha u_1 v_1^\top + (1-\alpha)\frac{\mM^k}{\normf{\mM^k}})
-	\end{equation}
-\end{lemma}
-
-We name the derived algorithm \emph{F-Neon}. It turns out that F-Neon is a convex combination of Normalized SGD and Neon. The implications are significant and discussed in the following sections.
+We name the derived algorithm \emph{Neon}. In the section ~\cref{sec:matrix-side}, we will discuss how to compute the update efficiently.
 
 \subsection{\texorpdfstring{$\normkfk{\mM^k}^\dagger$ and Muon, Neon, and Centralized Dion without error feedback}{Muon, Neon, and Centralized Dion without error feedback}}
+
+Now: what will we get if we try to generalize?
+
 We remind the reader that the Ky Fan k-norm (\citet{bhatia2013matrix}, p.
@@
 \end{lemma}
-In the section {\it The Matrix side of updates}, we will discuss how to compute the updates efficiently.
+In the section ~\cref{sec:matrix-side}, we will discuss how to compute the updates efficiently.
 
@@
 
-In addition, one can consider F-KF-k-norm: $\norm{\cdot}_{\mathrm{F-KF-k}}$ = $\alpha \normkfk{\cdot} + (1-\alpha)\normf{\cdot}$, the dual to which will produce algorithms like F-Dion without the error feedback.
-The resulting family can be named {\it F-Fanions}.
+Add more about Dion, and what its difference from us. I don't understand why their algorithm works well. I don't believe that the problem is with the error feedback.
 
-\subsection{Algorithms for Matrices $\leftrightarrow$ Algorithms for Vectors}
-
+\section{Algorithms for Matrices vs Algorithms for Vectors}
 lmo optimizers in Schatten $S_p$ norms and in $l_p$ norms with common $p$ may be analogous to each other, as is illustrated by \cref{tbl:mat_vs_vec_lmo}.
@@
 
-\section{Matrix side of updates}\label{sec:matrix-side}
+\section{Conic combination of LMO-algorithms is an LMO-algorithm}
+
+NSGD, Muon, and SignSGD work fair enough separately. Could we boost the quality if we combined them? To answer this question, we have to build a theoretical base for that.
+
+\subsection{Duals to Convex Combination of Norm}
+First, we need a well-known fact mentioned, for example, in \citet[Table 1]{yu2012arithmetic}. For the sake of completeness, we provide a proof of the fact.
+
+\begin{lemma}\label{lemma:dual_to_conv_comb}
+	Let $\norm{\cdot}_{(1)}$ and $\norm{\cdot}_{(2)}$ be norms on a finite-dimensional Euclidean space, and let $\alpha,\beta \geq 0$. Define
+	$$
+		\norm{x} := \alpha \norm{x}_{(1)} + \beta \norm{x}_{(2)}.
+	$$
+	Then the dual unit ball of $\norm{\cdot}$ satisfies
+	$$
+		B_{\norm{\cdot}^\dagger}
+		= \alpha B_{\norm{\cdot}_{(1)}^\dagger} + \beta B_{\norm{\cdot}_{(2)}^\dagger},
+	$$
+	where $+$ denotes the Minkowski sum and $B_{\norm{\cdot}_{(i)}^\dagger}$ is the unit ball of the dual norm $\norm{\cdot}_{(i)}^\dagger$.
+\end{lemma}
+
+\begin{proof}
+	Write $f(x) = \alpha \norm{x}_{(1)}$ and $g(x) = \beta \norm{x}_{(2)}$, so
+	$$
+		\norm{x} = f(x)+g(x).
+	$$
+	Recall two standard facts:
+	\begin{enumerate}
+		\item For any norm $\norm{\cdot}$ and $\lambda>0$,
+		      $$
+			      (\lambda \norm{\cdot})^*(y) =
+			      \sup_{x}\bigl( \langle y,x \rangle - \lambda \norm{x}\bigr)
+			      = \delta_{\lambda B_{\norm{\cdot}^\dagger}}(y),
+		      $$
+		      i.e., the indicator function of the scaled dual ball.
+		\item The Fenchel conjugate of a sum satisfies
+		      $$
+			      (f+g)^*(y) = \inf_{u+v=y} \bigl(f^*(u) + g^*(v)\bigr).
+		      $$
+	\end{enumerate}
+	Applying these to $f$ and $g$, we have
+	$$
+		f^*(u) = \delta_{\alpha B_{\norm{\cdot}_{(1)}^\dagger}}(u),
+		\quad
+		g^*(v) = \delta_{\beta B_{\norm{\cdot}_{(2)}^\dagger}}(v).
+	$$
+	Thus
+	$$
+		\norm{\cdot}^*(y)
+		= (f+g)^*(y)
+		= \inf_{u+v=y}
+		\bigl(
+		\delta_{\alpha B_{\norm{\cdot}_{(1)}^\dagger}}(u)
+		+
+		\delta_{\beta B_{\norm{\cdot}_{(2)}^\dagger}}(v)
+		\bigr)
+		=
+		\delta_{\alpha B_{\norm{\cdot}_{(1)}^\dagger}+\beta B_{\norm{\cdot}_{(2)}^\dagger}}(y).
+	$$
+	But by definition, the conjugate of a norm is exactly the indicator of its dual unit ball:
+	$$
+		\norm{\cdot}^*(y) = \delta_{B_{\norm{\cdot}^\dagger}}(y).
+	$$
+	Therefore,
+	$$
+		B_{\norm{\cdot}^\dagger} = \alpha B_{\norm{\cdot}_{(1)}^\dagger} + \beta B_{\norm{\cdot}_{(2)}^\dagger}.
+	$$
+\end{proof}
+
+Consequently,
+\begin{equation}\label{eq:dual_to_conv_comb}
+	\norm{y}^\dagger
+	= \inf\bigl\{ t \ge 0 : y \in t\bigl(\alpha B_{\norm{\cdot}_{(1)}^\dagger} + \beta B_{\norm{\cdot}_{(2)}^\dagger}\bigr) \bigr\}
+	= \inf_{z,t}\bigl\{t \ge 0: z \in t \alpha B_{\norm{\cdot}_{(1)}^\dagger}, y - z \in t \beta B_{\norm{\cdot}_{(1)}^\dagger}\bigr\}
+\end{equation}
+
+Now we can work with different norms.
+
+\subsection{General case}
+We have norm \#1 with update \#1 and norm \#2 with update \#2. We prove the lemma that says that the update of their conic combination is a conic combination of updates \#1 and \#2.
+
+\subsection{Frobeniusize them: F-Muon and F-Neon}
+Let us consider a mixture of Fanions with another excellent LMO-algorithm, NSGD.
+
+We define $\normfstar{\cdot}$ as a convex combination of $\normn{\cdot}$ and $\normf{\cdot}$:
+
+\begin{equation}\label{eq:fstar_norm}
+	\normfstar{\mX} = \alpha \normn{\mX} + (1-\alpha)\normf{\mX},
+\end{equation}
+
+where $\alpha \in [0, 1]$ defines a specific norm from the $F*$-family.
+$\normfstar{\cdot}^\dagger$ can be expressed as in \cref{eq:normfstardual}.
+However, its norm ball is a simple Minkowski sum of $\alpha \normn{\cdot}$ and $(1-\alpha)\normf{\cdot}$ balls \cref{fig:fstardual}.
+
+\begin{lemma}\label{lemma:nsgd_muon_update}
+	When $\norm{\cdot} = \normfstar{\cdot}^\dagger$, \cref{eq:our_update} turns into:
+
+	\begin{equation}\label{eq:nsgd_muon_update}
+		\mX^{k+1} = \mX^{k} - \eta (\alpha U V^\top + (1-\alpha)\frac{\mM^k}{\normf{\mM^k}})
+	\end{equation}
+\end{lemma}
+
+We name the derived algorithm \emph{F-Muon}.
+
+We define $\normftwo{\cdot}$ as a convex combination of $\norms{\cdot}$ and $\normf{\cdot}$:
+
+\begin{equation}\label{eq:ftwo_norm}
+	\normftwo{\mX} = \alpha \norms{\mX} + (1-\alpha)\normf{\mX},
+\end{equation}
+
+where $\alpha \in [0, 1]$ defines a specific norm from the F2-family.$\normftwo{\cdot}^\dagger$ can be expressed as in \cref{eq:normftwodual}.
+However, its norm ball is a simple Minkowski sum of $\alpha \norms{\cdot}$ and $(1-\alpha)\normf{\cdot}$ balls \cref{fig:ftwodual}.
+
+\begin{lemma}\label{lemma:nsgd_neon_update}
+	When $\norm{\cdot} = \normftwo{\cdot}^\dagger$, \cref{eq:our_update} turns into:
+
+	\begin{equation}\label{eq:nsgd_neon_update}
+		\mX^{k+1} = \mX^{k} - \eta (\alpha u_1 v_1^\top + (1-\alpha)\frac{\mM^k}{\normf{\mM^k}})
+	\end{equation}
+\end{lemma}
+
+We name the derived algorithm \emph{F-Neon}. F-Neon is a convex combination of Normalized SGD and Neon.
+
+In general, one can consider F-KF-k-norm: $\norm{\cdot}_{\mathrm{F-KF-k}}$ = $\alpha \normkfk{\cdot} + (1-\alpha)\normf{\cdot}$, the dual to which will produce algorithms like F-Dion without the error feedback.
+The resulting family can be named {\it F-Fanions}.
+
+\subsection{Add SignSGD: S-Muon and S-Neon}
+Trivial formulas, but a need for more experiments.
+
+\section{Computing the updates}\label{sec:matrix-side}
 To compute the algorithms' updates, we use thick-restarted Lanczos method for singular value problem (TRLan) on $\mM^{k\top}\mM^k$ or $\mM^k \mM^{k\top}$ matrices (the one with less size is picked), implemented in CuPy lSibrary \citep{cupy_svds_ref} and described in \citet{simonz1998thick}.
@@
 \subsection{Randomized Linear Least Squares}
-We first compare F-Fanions on the following L-smooth problem:
+First, we compare F-Fanions on the following L-smooth problem:
 
@@
 
+	{\bf Orthogonal research} Fanions, F-Fanions, and S-Fanions benefit from the general theoretical description of \citet{riabinin2025gluon}.
+They could be transformed to Drop-Fanions etc.
+(cite Drop-Muon article), they could be viewed as Broximal Point Methods (cite the article), they could be subject to EF (cite the EF for all).
+Of course, they could be clipped (cite ClipScion), and they could be analyzed when the LMO is not calculated accurately (cite Beyond the Ideal).
+
 \section{Conclusion}
@@
 \appendix
+\section{LMO for Neural Networks}
+In a typical neural network, the objective function $F$ depends on a set of weight matrices $\{ \mW_1, \mW_2, \ldots \}$.
+The optimization framework we have described is applied in a layer-wise fashion.
+At each iteration $k$, a stochastic gradient $g(\mX^k, \xi^k)$ is computed using a mini-batch of data with the $\xi^k$ noise via backpropagation.
+This yields a separate gradient component, $\mG_i^k$, for each matrix $\mX_i$.
+The LMO-based update rule is then applied to each matrix $\mX_i$ using its corresponding gradient component $\mG_i^k$.
+
+The update rule used in Muon optimizer is uSCG \citep{pethick2025training}. In the most general case, which involves momentum, it can be written as
+\begin{equation}\label{eq:usgd_update}
+	\begin{aligned}
+		\mM^{k}     & = \alpha_{k} g(\mX^k, \xi^k) + (1 - \alpha_k)\mM^{k - 1}\,, \\
+		\mX^{k + 1} & = \mX^k + \gamma_k\mathrm{LMO}(\mM^{k})\,.
+	\end{aligned}
+\end{equation}
 
 \section{Norms $\normfstar{\cdot}^\dagger$ and $\normftwo{\cdot}^\dagger$}
-First, we need a well-known fact mentioned, for example, in \citet[Table 1]{yu2012arithmetic}. For the sake of completeness, we provide a proof of the fact.
 
-\begin{lemma}\label{lemma:dual_to_conv_comb}
-	Let $\norm{\cdot}_{(1)}$ and $\norm{\cdot}_{(2)}$ be norms on a finite-dimensional Euclidean space, and let $\alpha,\beta \geq 0$. Define
-	$$
-		\norm{x} := \alpha \norm{x}_{(1)} + \beta \norm{x}_{(2)}.
-	$$
-	Then the dual unit ball of $\norm{\cdot}$ satisfies
-	$$
-		B_{\norm{\cdot}^\dagger}
-		= \alpha B_{\norm{\cdot}_{(1)}^\dagger} + \beta B_{\norm{\cdot}_{(2)}^\dagger},
-	$$
-	where $+$ denotes the Minkowski sum and $B_{\norm{\cdot}_{(i)}^\dagger}$ is the unit ball of the dual norm $\norm{\cdot}_{(i)}^\dagger$.
-\end{lemma}
-
-\begin{proof}
-	Write $f(x) = \alpha \norm{x}_{(1)}$ and $g(x) = \beta \norm{x}_{(2)}$, so
-	$$
-		\norm{x} = f(x)+g(x).
-	$$
-	Recall two standard facts:
-	\begin{enumerate}
-		\item For any norm $\norm{\cdot}$ and $\lambda>0$,
-		      $$
-			      (\lambda \norm{\cdot})^*(y) =
-			      \sup_{x}\bigl( \langle y,x \rangle - \lambda \norm{x}\bigr)
-			      = \delta_{\lambda B_{\norm{\cdot}^\dagger}}(y),
-		      $$
-		      i.e., the indicator function of the scaled dual ball.
-		\item The Fenchel conjugate of a sum satisfies
-		      $$
-			      (f+g)^*(y) = \inf_{u+v=y} \bigl(f^*(u) + g^*(v)\bigr).
-		      $$
-	\end{enumerate}
-	Applying these to $f$ and $g$, we have
-	$$
-		f^*(u) = \delta_{\alpha B_{\norm{\cdot}_{(1)}^\dagger}}(u),
-		\quad
-		g^*(v) = \delta_{\beta B_{\norm{\cdot}_{(2)}^\dagger}}(v).
-	$$
-	Thus
-	$$
-		\norm{\cdot}^*(y)
-		= (f+g)^*(y)
-		= \inf_{u+v=y}
-		\bigl(
-		\delta_{\alpha B_{\norm{\cdot}_{(1)}^\dagger}}(u)
-		+
-		\delta_{\beta B_{\norm{\cdot}_{(2)}^\dagger}}(v)
-		\bigr)
-		=
-		\delta_{\alpha B_{\norm{\cdot}_{(1)}^\dagger}+\beta B_{\norm{\cdot}_{(2)}^\dagger}}(y).
-	$$
-	But by definition, the conjugate of a norm is exactly the indicator of its dual unit ball:
-	$$
-		\norm{\cdot}^*(y) = \delta_{B_{\norm{\cdot}^\dagger}}(y).
-	$$
-	Therefore,
-	$$
-		B_{\norm{\cdot}^\dagger} = \alpha B_{\norm{\cdot}_{(1)}^\dagger} + \beta B_{\norm{\cdot}_{(2)}^\dagger}.
-	$$
-\end{proof}
-
-Consequently,
-\begin{equation}\label{eq:dual_to_conv_comb}
-	\norm{y}^\dagger
-	= \inf\bigl\{ t \ge 0 : y \in t\bigl(\alpha B_{\norm{\cdot}_{(1)}^\dagger} + \beta B_{\norm{\cdot}_{(2)}^\dagger}\bigr) \bigr\}
-	= \inf_{z,t}\bigl\{t \ge 0: z \in t \alpha B_{\norm{\cdot}_{(1)}^\dagger}, y - z \in t \beta B_{\norm{\cdot}_{(1)}^\dagger}\bigr\}
-\end{equation}
-
-Thus, we immediately find $\normfstar{\cdot}^\dagger$, which is related to F-Muon update.
+Based on the lemma, we immediately find $\normfstar{\cdot}^\dagger$, which is related to F-Muon update.
 Indeed, after setting $\beta=1-\alpha$ and remembering that for smooth and bounded cases we can use $\min$ instead of $\inf$, we get
@@
 If $\alpha = 1$, then $\mZ = \mY$, and we get $\normftwo{\mY}^\dagger = \normn{\mY}$. If $\alpha = 0$, then $\mZ = 0$, and we get $\normftwo{\mY}^\dagger = \normf{\mY}$.
+
+Unfortunately, $\normfstar{\cdot}^\dagger$ and $\normftwo{\cdot}^\dagger$ do not have a closed-form expression.
 
@@
 \section{Updates derivations}
-
-Proof of \cref{lemma:muon_update} follows from \cref{eq:nsgd_muon_update} with $\alpha=1$. Indeed, $\norms{\cdot}^\dagger=1\cdot\normn{\cdot} + 0 \cdot \normf{\cdot} $.
-
 Proof of \cref{lemma:nsgd_muon_update}:

### 2025-11-10 conf_article/neon.tex
Shortened the proofs

@@
 \section{Updates derivations}
-Proof of \cref{lemma:nsgd_muon_update}:
-Since $\norm{\cdot}^\dagger = \normfstar{\cdot}^{\dagger\dagger} = \normfstar{\cdot}$, the goal is to reach $\normfstar{\mM^k} = \alpha \tr{\mSigma} + (1-\alpha) \normf{\mM^k}$.
+% we prove a general lemma instead
+%Proof of \cref{lemma:nsgd_muon_update}:
+%Since $\norm{\cdot}^\dagger = \normfstar{\cdot}^{\dagger\dagger} = \normfstar{\cdot}$, the goal is to reach $\normfstar{\mM^k} = \alpha \tr{\mSigma} + (1-\alpha) \normf{\mM^k}$.
 
-Let us note that $\Delta = \alpha \mU \mV^\top + (1-\alpha) \frac{\mM^k}{\normf{\mM^k}}$ delivers this value.
+%Let us note that $\Delta = \alpha \mU \mV^\top + (1-\alpha) \frac{\mM^k}{\normf{\mM^k}}$ delivers this value.
 Indeed, by the trace property, $\<\mM^k, \Delta> =\<\mU \mSigma \mV^\top, \alpha \mU \mV^\top + (1-\alpha) \frac{\mU \mSigma \mV^\top}{\normf{\mM^k}}> = \alpha \tr \mSigma + (1-\alpha) \normf{\mM^k} = \normfstar{\mM^k}$, which completes the proof.
 
-Proof of \cref{lemma:neon_update} follows from \cref{eq:nsgd_neon_update} with $\alpha=1$. Indeed, $\normn{\cdot}^\dagger=1\cdot\norms{\cdot} + 0 \cdot \normf{\cdot} $.
+Proof of \cref{lemma:neon_update}:
+Since $\norm{\cdot}^\dagger = \norms{\cdot}$, the goal is to reach $\norms{\mM^k} = \sigma_1$.
 
-Proof of \cref{lemma:nsgd_neon_update}:
-Since $\norm{\cdot}^\dagger = \normftwo{\cdot}^{\dagger\dagger} = \normftwo{\cdot}$, the goal is to reach $\normftwo{\mM^k} = \alpha \sigma_1 + (1-\alpha) \normf{\mM^k}$.
+Let us note that $\Delta = u_1 v_1^\top$ delivers this value.
+Indeed, by the trace property and singular vectors orthogonality, $\<\mM^k, \Delta> =\<\mU \mSigma \mV^\top, u_1 v_1^\top> = \tr \diag({\sigma_1, 0, \dots, 0}) = \norms{\mM^k}$, which completes the proof.
 
-Let us note that $\Delta = \alpha (u_1 v_1^\top) + (1-\alpha) \frac{\mM^k}{\normf{\mM^k}}$ delivers this value.
+% we prove a general lemma instead
+%Proof of \cref{lemma:nsgd_neon_update}:
+%Since $\norm{\cdot}^\dagger = \normftwo{\cdot}^{\dagger\dagger} = \normftwo{\cdot}$, the goal is to reach $\normftwo{\mM^k} = \alpha \sigma_1 + (1-\alpha) \normf{\mM^k}$.
+
+% Let us note that $\Delta = \alpha (u_1 v_1^\top) + (1-\alpha) \frac{\mM^k}{\normf{\mM^k}}$ delivers this value.
 Indeed, by the trace property and singular vectors orthogonality, $\<\mM^k, \Delta> =\<\mU \mSigma \mV^\top, \alpha u_1 v_1^\top + (1-\alpha) \frac{\mU \mSigma \mV^\top}{\normf{\mM^k}}> = \alpha \tr \diag({\sigma_1, 0, \dots, 0}) + (1-\alpha) \normf{\mM^k} = \normftwo{\mM^k}$, which completes the proof.

### 2025-11-23 conf_article/neon.tex
Added Schatten, revised preliminaries

@@
 \DeclarePairedDelimiter{\normf}{\|}{\|_\mathrm{F}}
-\DeclarePairedDelimiter{\normkfk}{\|}{\|_\mathrm{KF-k}}
+\DeclarePairedDelimiter{\normkfk}{\|}{\|_{\text{KF-}k}}
 \DeclarePairedDelimiter{\normfstar}{\|}{\|_\mathrm{F*}}
 \DeclarePairedDelimiter{\normftwo}{\|}{\|_\mathrm{F2}}
-\DeclarePairedDelimiter{\sqns}{\|}{\|_{\mathrm{op}}^2}
-\DeclarePairedDelimiter{\norms}{\|}{\|_{\mathrm{op}}}
+\DeclarePairedDelimiter{\sqns}{\|}{\|_{\mathrm{2}}^2}
+\DeclarePairedDelimiter{\norms}{\|}{\|_{\mathrm{2}}}
 \DeclarePairedDelimiter{\sqnn}{\|}{\|_{\mathrm{nuc}}^2}
@@
 
-Training a neural network is essentially an optimization of a function of several weight matrices.
+Training a neural network is essentially an optimization of a function of several weight matrices and a few vectors.
 Let us start by considering the problem of minimizing a differentiable function of a {\it single} matrix:
@@
 We equip the matrix space $\Rmn$ with a standard dot product $\<\cdot, \cdot> \to \R$ and a norm $\norm{\cdot}\colon \Rmn \to \R_+$, which does not have to coincide with the Frobenius norm $\normf{\cdot}$.
-The dual norm $\norm{\cdot}^\dagger\colon \Rmn \to \R_+$ associated with $\norm{\cdot}$ is defined as
+The dual norm $\norm{\cdot}^\dagger\colon \Rmn \to \R_+$ that is associated with $\norm{\cdot}$ is defined as
 \begin{equation}\label{eq:dual_norm}
@@
 \begin{equation}\label{eq:lmo}
-	\mathrm{LMO}(\mG) \in \argmin_{\mD \in \cS} \<\mG, \mD>\,,
+	\mathrm{LMO}(\mM^k) \in \argmin_{\mD \in \cS} \<\mM^k, \mD>\,,
 \end{equation}
-where $\mG$ is typically a gradient (possibly with momentum) of $F$ and $\cS \subset \Rmn$ is some set. The simple update formula then can look like
+where $\mM^k$ is a gradient (or a momentum) of $F$ in $\mX^k$ and $\cS \subset \Rmn$ is some set. The update of the algorithm is defined as follows:
 \begin{equation}\label{eq:simple_update}
-	\mX^{k + 1} = \mX^k + \gamma_k\mathrm{LMO}\left(\nabla F\left(\mX^k\right)\right)\,.
+	\mX^{k + 1} = \mX^k + \gamma_k\mathrm{LMO}\left(\mM\right)\,.
 \end{equation}
 
-Here we are particularly interested in the case when $\cS$ is a ball in the $\norm{\cdot}$ norm:
-\begin{equation}\label{eq:ball}
-	\cS = \cB_\eta = \{ \mD \in \Rmn \mid \norm{\mD} \leq \eta \}\,.
+We are particularly interested in the case when $\cS$ is a ball in the norm $\norm{\cdot}$:
+
+$$\cS = \cB_\eta \in \{ \mD \in \Rmn \mid \norm{\mD} \leq \eta \}.$$
+In this case,
+
+$$\argmin_{\mD \in \cS} \<\mM^k, \mD> = - \eta \{ \mD \in \cB_1 \mid \<\mM^k, \mD> = \norm{\mM^k}^\dagger\}$$
+
+and update for $\mX^k$ in \cref{eq:simple_update} simplifies to
+\begin{equation}\label{eq:our_update}
+	\mX^{k+1} = \mX^{k} - \eta_k \{\mD \in \cB_1 \mid \<\mM^k, \mD> = \norm{\mM^k}^\dagger\}\,
 \end{equation}
-In this case it can be easily shown that
-\begin{equation} \label{eq:lmo_in_norm_ball}
-	\argmin_{\mD \in \cS} \<\mG, \mD> = - \eta \{ \mD \in \cB_1 \mid \<\mG, \mD> = \norm{\mG}^\dagger\}\,.
-\end{equation}
-And update for $\mX$ in \cref{eq:usgd_update} becomes
-\begin{equation}\label{eq:our_update}
-	\mX^{k+1} = \mX^{k} - \gamma_k\eta \{\mD \in \cB_1 \mid \<\mM^k, \mD> = \norm{\mM^k}^\dagger\}\,.
-\end{equation}
-This formula will later allow us to easily compute updates for various norms.
 
-In the next subsection we will derive and discuss updates induced by the choice of the norm in \cref{eq:our_update}.
-To describe those updates, the singular value decomposition (SVD) of $\mM^k$ is required.
-We denote SVD of $\mM^k$ as $\mM^k = \mU \mSigma \mV^\top$, where $\mU = [u_1, u_2, \dots, u_r]$, $\mSigma = \diag(\sigma_1, \sigma_2, \dots, \sigma_r)$, and $\mV = [v_1, v_2, \dots, v_r]$.
+By this formula, it is easy to compute updates for algorithms induced by various norms $\norm{\cdot}$.
+For example, when the norm $\norm{\cdot}$ is the Frobenius norm $\normf{\cdot}$, \cref{eq:our_update} turns into
 
-When $\norm{\cdot} = \norms{\cdot}$, \cref{eq:our_update} turns into:
+\begin{equation}\label{eq:nsgd_update}
+	\mX^{k+1} = \mX^{k} - \eta_k \frac{\mM^k}{\normf{\mM^k}},
+\end{equation} which recovers Normalized SGD (NSGD). When one has the spectral norm $\norms{\cdot}$, one gets
 
 \begin{equation}\label{eq:muon_update}
-	\mX^{k+1} = \mX^{k} - \eta U V^\top
-\end{equation}.
-
-And when $\norm{\cdot} = \normf{\cdot}$, \cref{eq:our_update} turns into:
-
-	\begin{equation}\label{eq:nsgd_update}
-		\mX^{k+1} = \mX^{k} - \eta \frac{\mM^k}{\normf{\mM^k}}
-	\end{equation}
+	\mX^{k+1} = \mX^{k} - \eta_k U V^\top,
+\end{equation} which is Muon without the $\sqrt{m/n}$ factor.
+Here, $\mM^k = \mU \mSigma \mV^\top$ is the Singular Value Decomposition (SVD) of $\mM^k$ ($\mU = [u_1, u_2, \dots, u_r]$, $\mSigma = \diag(\sigma_1, \space \sigma_2, \dots, \sigma_r)$, and $\mV = [v_1, v_2, \dots, v_r]$).
+Muon can be recovered by taking the RMS-to-RMS operator norm $\sqrt{\frac{n}{m}}\norms{\cdot}$.
 
 \section{Duals to Ky Fan norms instead of the spectral norm}
-%\subsection{Non-operator Norms $\norm{\cdot}$ lead to new updates}
 
 \subsection{\texorpdfstring{$\normn{\mM^{k}}$ and Neon}{Neon}}
+After having considered $\normf{\mM^k}$ and $\norms{\mM^k}$, it is natural to look at the nuclear norm $\normn{\mM^k}$.
 \begin{lemma}\label{lemma:neon_update}
@@
 	\begin{equation}\label{eq:neon_update}
-		\mX^{k+1} = \mX^{k} - \eta u_1 v_1^\top
+		\mX^{k+1} = \mX^{k} - \eta_k u_1 v_1^\top
 	\end{equation}
 \end{lemma}
-We name the derived algorithm \emph{Neon}. In the section ~\cref{sec:matrix-side}, we will discuss how to compute the update efficiently.
+\begin{proof}
+Since $\norm{\cdot}^\dagger = \norms{\cdot}$, the goal is to reach $\norms{\mM^k} = \sigma_1$.
+
+Let us note that $\Delta = u_1 v_1^\top$ delivers this value.
+Indeed, $\normn{\Delta}=1$ and by the trace property and orthogonality of the singular vectors, $\<\mM^k, \Delta> =\<\mU \mSigma \mV^\top, u_1 v_1^\top> = \tr \diag({\sigma_1, 0, \dots, 0}) = \norms{\mM^k}$, which completes the proof.
+\end{proof}
+
+We name the derived algorithm \emph{Neon}. In the~\cref{sec:matrix-side}, we will discuss how to compute the Neon's update.
 
@@
 
-Now: what will we get if we try to generalize?
+Neon's and Muon's updates seem to be complete opposites: one has rank one, while the other is full-rank. It would be interesting to derive algorithms with updates of various ranks.
 
-We remind the reader that the Ky Fan k-norm (\citet{bhatia2013matrix}, p.
-92), which we denote as $\normkfk{\cdot}$, is the sum of the k largest singular values of the matrix.
-It can be proved that $\normkfk{\cdot}^\dagger = \max\{\frac{1}{k} \normn{\cdot}, \norms{\cdot}\}$ (see \citet{bhatia2013matrix}, p.
+\textbf{Schatten norms}
+\citet{cesista2025schattenp} considered Schatten-$p$ norms $\norm{\mM^k}_{S_p} = \left(\sum_{i=1}^{\min(m, n)}\sigma_i^p\right)^{1/p}$.
+The update was as follows: $$X^{k+1} = X^k - \eta_k \mU \frac{\diag\left(\sigma_1^{q-1},\dots, \sigma_{\min(m,n)}^{q-1}\right)}{\norm{\mM^k}^{q-1}_{S_q}}\mV^\top$$ for $q$: $\frac{1}{p} + \frac{1}{q} = 1$.
+This formula recovers Neon when $p\rightarrow1$ provided that $\sigma_1 > \sigma_2$, which is true on real data; NSGD when $p=2$, and Muon when $p\rightarrow\infty$.
+
+However, Schatten norms do not fill the gap rank: indeed, when $p > 1$, the rank of the update is full, while when $p=1$ ($p \rightarrow 1$) it is one.
+Moreover, it is not clear how to calculate the update for $p \neq 1, 2, \infty$: it seems one has to know all $\sigma_i$ to compute the update, which makes the problem as hard as the full SVD.
+
+\textbf{Ky Fan norms}
+There is another family of matrix norms, which might help us: Ky Fan norms.
+For $k \in \{1, \dots, \min(m, n)\}$, the Ky Fan $k$-norm $\normkfk{\cdot}$ is $\sum_{i=1}^{k}\sigma_i$, i.e.
+the sum of the $k$ largest singular values of the matrix.
+Special cases of the Ky Fan $k$-norm are the Ky Fan $1$-norm, which is the sprectral norm, and the Ky Fan $\min\{m, n\}$-norm, which is the nuclear norm.
+
+\textbf{Duals to Ky Fan norms}
+While Schatten norms are closed under dualization and $\norm{\mM^k}_{S_p}^\dagger = \norm{\mM^k}_{S_q}$, for Ky Fan norms it is not generally the case: $\normkfk{\cdot}^\dagger = \max\{\frac{1}{k} \normn{\cdot}, \norms{\cdot}\}$ (see \citet{bhatia2013matrix}, p.
 96).
-Special cases of the Ky Fan k-norm are the Ky Fan 1-norm, which is the sprectral norm, and the Ky Fan $\min\{m, n\}$-norm, which is the nuclear norm.
+$k = 1$ and $k=\min(m, n)$ are exceptional: for $k=1$ the dual norm is $\normn{\cdot}$ with $k = \min(m, n)$ and for $k = \min(m, n)$ the dual norm is $\norms{\cdot}$ with $k=1$.
+For each other $k$ $\normkfk{\cdot} \neq \norm{\cdot}_{\text{KF-}k'}$ for any $k'$: $k'=\min(m, n)$ and $k'= 1$ correspond to the already discussed cases, while for other $k'$ one can change $\norm{\mM^k}_{\text{KF-}k'}$ by moving a small value between $\sigma_{k'}$ and $\sigma_{k'+1}$.
+During this change, $\norms{\mM^k}$ and $\normn{\mM^k}$ will remain constant, hence $\normkfk{\mM^k}$ will not change as well.
 
@@
 
+Then prove a lemma for any set of norms \#1, \dots, \#n (by induction)
+
 \subsection{Frobeniusize them: F-Muon and F-Neon}
@@
 \subsection{Add SignSGD: S-Muon and S-Neon}
-Trivial formulas, but a need for more experiments.
+Trivial formulas, but we need more experiments.
 
@@
 
-Proof of \cref{lemma:neon_update}:
-Since $\norm{\cdot}^\dagger = \norms{\cdot}$, the goal is to reach $\norms{\mM^k} = \sigma_1$.
-
-Let us note that $\Delta = u_1 v_1^\top$ delivers this value.
-Indeed, by the trace property and singular vectors orthogonality, $\<\mM^k, \Delta> =\<\mU \mSigma \mV^\top, u_1 v_1^\top> = \tr \diag({\sigma_1, 0, \dots, 0}) = \norms{\mM^k}$, which completes the proof.
-
 % we prove a general lemma instead

### 2025-11-24 conf_article/neon.tex
Improved related work, added Ky Fan analysis

@@
 Motivated by the success of Muon, many generalizations and variations of it were proposed.
-Among the notable ones are Scion \citep{pethick2025training}, Dion \citep{ahn2025dioncommunicationefficientoptimizerlarge} and Gluon \citep{riabinin2025gluon}.
+Among the notable ones are Scion \citep{pethick2025training}, Dion \citep{ahn2025dion} and Gluon \citep{riabinin2025gluon}.
 Those researches target Muon's efficiency and establish convergence bounds.
@@
 
-\textbf{Schatten norms}
-\citet{cesista2025schattenp} considered Schatten-$p$ norms $\norm{\mM^k}_{S_p} = \left(\sum_{i=1}^{\min(m, n)}\sigma_i^p\right)^{1/p}$.
-The update was as follows: $$X^{k+1} = X^k - \eta_k \mU \frac{\diag\left(\sigma_1^{q-1},\dots, \sigma_{\min(m,n)}^{q-1}\right)}{\norm{\mM^k}^{q-1}_{S_q}}\mV^\top$$ for $q$: $\frac{1}{p} + \frac{1}{q} = 1$.
+\textbf{Schatten norms.}
+\citet{cesista2025schattenp} considered Schatten-$p$ norms $\norm{\mM^k}_{S_p} = \left(\sum_{i=1}^{\min(m, n)}\sigma_i^p\right)^{1/p}$, which produce the update:
+$$X^{k+1} = X^k - \eta_k \mU \frac{\diag\left(\sigma_1^{q-1},\dots, \sigma_{\min(m,n)}^{q-1}\right)}{\left(\sum_{i=1}^{\min(m,n)}{\sigma_i^q}\right)^{\frac{q-1}{q}}}\mV^\top$$
+for $q$: $\frac{1}{p} + \frac{1}{q} = 1$.
 This formula recovers Neon when $p\rightarrow1$ provided that $\sigma_1 > \sigma_2$, which is true on real data; NSGD when $p=2$, and Muon when $p\rightarrow\infty$.
@@
 
-\textbf{Ky Fan norms}
+\textbf{Ky Fan norms.}
 There is another family of matrix norms, which might help us: Ky Fan norms.
@@
 
-\textbf{Duals to Ky Fan norms}
-While Schatten norms are closed under dualization and $\norm{\mM^k}_{S_p}^\dagger = \norm{\mM^k}_{S_q}$, for Ky Fan norms it is not generally the case: $\normkfk{\cdot}^\dagger = \max\{\frac{1}{k} \normn{\cdot}, \norms{\cdot}\}$ (see \citet{bhatia2013matrix}, p.
-96).
+Let us derive the update for an arbitrary $k$.
+Since $\normkfk{\cdot}^\dagger = \max\{\frac{1}{k} \normn{\cdot}, \norms{\cdot}\}$ (see \citet{bhatia2013matrix}, p.
+96), the goal is to reach $\max\{\frac{1}{k}\sum_{i=1}^{\min(m,n)}\sigma_i, \sigma_1\}$.
+$\Delta = u_1 v_1^\top$ from Neon delivers $\sigma_1$, while $\Delta = \frac{1}{k}\sum_{i=1}^{\min(m,n)}u_i v_i^\top$ from Muon (with an extra factor of $1/k$) delivers $\frac{1}{k}\sum_{i=1}^{\min(m,n)}\sigma_i$.
+Thus, the update is either
+$$\mX^{t+1} = \mX^{t} - \eta_t u_1 v_1^\top\text{ or }\mX^{t+1} = \mX^{t} - \frac{\eta_t}{k}\mU \mV^\top,$$
+depending on which one better minimizes the function $F(\mX^t) + \mM^t(\mX^{t+1} - \mX^t)$.
+
+The rank gap is not closed by the Ky Fan norms because the obtained update is either one-rank or full-rank.
+
+\textbf{Duals to Ky Fan norms.}
+While Schatten norms are closed under dualization and $\norm{\mM^k}_{S_p}^\dagger = \norm{\mM^k}_{S_q}$, for Ky Fan norms it is not generally the case.
 $k = 1$ and $k=\min(m, n)$ are exceptional: for $k=1$ the dual norm is $\normn{\cdot}$ with $k = \min(m, n)$ and for $k = \min(m, n)$ the dual norm is $\norms{\cdot}$ with $k=1$.
-For each other $k$ $\normkfk{\cdot} \neq \norm{\cdot}_{\text{KF-}k'}$ for any $k'$: $k'=\min(m, n)$ and $k'= 1$ correspond to the already discussed cases, while for other $k'$ one can change $\norm{\mM^k}_{\text{KF-}k'}$ by moving a small value between $\sigma_{k'}$ and $\sigma_{k'+1}$.
+For each other $k$, $\normkfk{\cdot} \neq \norm{\cdot}_{\text{KF-}k'}$ for any $k'$: $k'=\min(m, n)$ and $k'= 1$ correspond to the already discussed cases, while for other $k'$ one can change $\norm{\mM^k}_{\text{KF-}k'}$ by moving a small value between $\sigma_{k'}$ and $\sigma_{k'+1}$.
 During this change, $\norms{\mM^k}$ and $\normn{\mM^k}$ will remain constant, hence $\normkfk{\mM^k}$ will not change as well.
@@
 \end{lemma}
-In the section ~\cref{sec:matrix-side}, we will discuss how to compute the updates efficiently.
 
-We introduce the family of {\it Fanions}, which consists of {\it Fanion-$k$}, lmo-based algorithms under the $\normkfk{\mM^t}^\dagger$ norms.
-By this terminology and the lemma, Muon is a Fanion-$\min\{n,m\}$, while Neon is an Fanion-1.
-Moreover, the centralized version of rank-$r$ Dion \citep{ahn2025dioncommunicationefficientoptimizerlarge} without the error feedback and without scaling of the update, from the perspective of lmo, is actually a Fanion-$r$.
+\begin{proof}
+Since $\norm{\cdot}^\dagger = \normkfk{\cdot}^{\dagger\dagger} = \normkfk{\cdot}$, the goal to reach $\normkfk{\mM^t}$.
 
-Add more about Dion, and what its difference from us. I don't understand why their algorithm works well. I don't believe that the problem is with the error feedback.
-
-\section{Algorithms for Matrices vs Algorithms for Vectors}
-lmo optimizers in Schatten $S_p$ norms and in $l_p$ norms with common $p$ may be analogous to each other, as is illustrated by \cref{tbl:mat_vs_vec_lmo}.
-The analogies go beyond similarities in the updates.
-SignSGD is very close to Adam, as noted in \citet{bernstein2024oldoptimizernewnorm}, and both Adam and Muon perform well in training large models.
-NSGD is the same for both matrix and vector cases.
-Greedy Coordinate Descent methods are not applied to high-dimensional problems, from this perspective, it is not surprising that Neon underperforms on such problems.
-
-\begin{table*}[t]
-	\caption{lmo optimizers in Schatten $S_p$ norms and in $l_p$ norms. $g$ is the gradient. When it is a matrix, $g = \mU \mSigma \mV^\top$}
-	\label{tbl:mat_vs_vec_lmo}
-	\bgroup
-	\def\arraystretch{1.2}
-	\resizebox{\textwidth}{!}{
-		\begin{tabular}{|l|c|c|c|}
-			\hline
-			Method                             & lmo constraint set $\mathcal D$ & lmo                                                         & Reference                            \\
-			\hline
-			\hline
-			Normalized SGD                     & $l_2$-ball, $S_2$-ball          & $-\eta \tfrac{g}{\norm{g}_2} = -\eta \tfrac{g}{\norm{g}_F}$ & \citep{hazan2015beyond}              \\
-			Momentum Normalized SGD            & Ball in $l_2$, or Ball in $S_2$ & $-\eta \tfrac{g}{\norm{g}_2} = -\eta\tfrac{g}{\norm{g}_F}$  & \citep{cutkosky2020momentum}         \\
-			\hline
-			SignSGD                            & Ball in Max-norm $l_\infty$     & $-\eta \sign(g)$                                            & \citep[Thm. 1]{bernstein2018signsgd} \\
-			Signum                             & Ball in Max-norm $l_\infty$     & $-\eta \sign(g)$                                            & \citep[Thm. 3]{bernstein2018signsgd} \\
-			\hdashline
-			Muon                               & Ball in Spectral $S_\infty$     & $-\eta UV^\top$                                             & \citep{jordan2024muon}               \\
-			\hline
-			Gauss-Southwell Coordinate Descent & Ball in $l_1$                   & $-\eta \sum_{i \in \argmax|g_i^t|} \sign(g_i^t)e_i$                       & \citep[p. 19]{shi2016primer}         \\
-			\hdashline
-			Neon                               & Ball in Nuclear $S_1$           & $-\eta u_1 v_1^\top$                                        & This work                            \\
-			\hline
-		\end{tabular}
-	}
-	\egroup
-\end{table*}
+Let us note that $\Delta = \sum_{i=1}^{k} u_i v_i^\top$ delivers the value.
+Indeed, $\<\mM^t, \Delta> =\<\mU \mSigma \mV^\top, \sum_{i=1}^{k} u_i v_i^\top> = \sum_{i,j=1}^{r, k}\<u_i \sigma_i v_i^\top, u_j v_j^\top> = \sum_{i=1}^{k}\sigma_i = \normkfk{\mM^t}$, which completes the proof.
+\end{proof}
+These updates constitute a family {\it Fanions}, LMO-based algorithms under the $\normkfk{\mM^t}^\dagger$ norms.
+The algorithm for a particular $k$ we will call {\it Fanion-$k$}.
+Thus, Muon is a Fanion-$\min\{m,n\}$, while Neon is a Fanion-1.
+Moreover, the unsharded version of the rank-$r$ Dion (Algorithm 1 from \citet{ahn2025dion}) without the error feedback and without scaling of the update is actually a Fanion-$r$ (see \citet{pethick2025understandingdion}, where Dion is written down in a notation more similar to ours).
 
@@
 
-\section{Related Work}
+\section{Algorithms for Vectors $\rightarrow$ Algorithms for Matrices}
+
+One cannot but notice the similarity between the updates LMO optimizers in the vector $l_p$ and the matrix Schatten $S_p$ norms, which is illustrated by \cref{tbl:mat_vs_vec_lmo}.
+However, the parallels are not limited to this similarity and can be traced even in the empirical observations.
+SignSGD is very close to Adam, as noted in \citet{bernstein2024oldoptimizernewnorm}, and both Adam and Muon perform well in training large models.
+NSGD is the same for both matrix and vector cases.
+Greedy Coordinate Descent is not applied to high-dimensional problems, so from this perspective it is not surprising that one-rank Neon underperforms on such problems.
+
+\begin{table*}[t]
+	\caption{lmo optimizers in Schatten $S_p$ norms and in $l_p$ norms. $g$ is the gradient. When it is a matrix, $g = \mU \mSigma \mV^\top$}
+	\label{tbl:mat_vs_vec_lmo}
+	\bgroup
+	\def\arraystretch{1.2}
+	\resizebox{\textwidth}{!}{
+		\begin{tabular}{|l|c|c|c|}
+			\hline
+			Method                             & lmo constraint set $\mathcal D$ & lmo                                                         & Reference                            \\
+			\hline
+			\hline
+			Normalized SGD                     & $l_2$-ball, $S_2$-ball          & $-\eta \tfrac{g}{\norm{g}_2} = -\eta \tfrac{g}{\norm{g}_F}$ & \citep{hazan2015beyond}              \\
+			Momentum Normalized SGD            & Ball in $l_2$, or Ball in $S_2$ & $-\eta \tfrac{g}{\norm{g}_2} = -\eta\tfrac{g}{\norm{g}_F}$  & \citep{cutkosky2020momentum}         \\
+			\hline
+			SignSGD                            & Ball in Max-norm $l_\infty$     & $-\eta \sign(g)$                                            & \citep[Thm. 1]{bernstein2018signsgd} \\
+			Signum                             & Ball in Max-norm $l_\infty$     & $-\eta \sign(g)$                                            & \citep[Thm. 3]{bernstein2018signsgd} \\
+			\hdashline
+			Muon                               & Ball in Spectral $S_\infty$     & $-\eta UV^\top$                                             & \citep{jordan2024muon}               \\
+			\hline
+			Gauss-Southwell Coordinate Descent & Ball in $l_1$                   & $-\eta \sum_{i \in \argmax|g_i^t|} \sign(g_i^t)e_i$                       & \citep[p. 19]{shi2016primer}         \\
+			\hdashline
+			Neon                               & Ball in Nuclear $S_1$           & $-\eta u_1 v_1^\top$                                        & This work                            \\
+			\hline
+		\end{tabular}
+	}
+	\egroup
+\end{table*}
+
+\section{Related Work and Discussion}
 As Muon \citep{jordan2024muon} is a very successful and popular optimizer for functions of weight matrices, a lot of reseach has been put into, first, further improving its performance, and, second, in explaining its success.
@@
 
-	{\bf The nuclear norm in lmo.} As we found out only when writing this article, the nuclear norm has been already explored in the context of the linear minimization oracle.
+	{\bf The LMO and Error Feedback.} We already mentioned that rank-$k$ unsharded Dion without error feedback and scaling of the update is Fanion-$k$.
+Since error feedback for Dion is very important as discovered by the ablation study in \citet{ahn2025dion}, F-Fanions and S-Fanions will benefit from it as well.
+In the context of federated learning, error feedback is effective even for compressed Muon \citep{gruntkowska2025efmuon}.
+The advantage of Fanions and S-Fanions is that less bits are required to transmit the updates: $\sum_{i=1}^k u_i v_i^\top$ is easily transmitted as $\{u_1, v_1, \dots, u_k, v_k\}$ ($(m+n)\times k$ floats), while the sign part is as usual coded in $m \times n$ bits.
+Thus, the compression is "built-in".
+Moreover, there is an intriguing possibility to construct differentially private Fanions and S-Fanions with the specific more optimal non-Gaussian noise, as was done with DP-signSGD \citep{jang2024dplogsign}.
+This we leave for future research.
+
+	{\bf The nuclear norm in the LMO.} As we found out only when transforming our results into an article, the nuclear norm has already been explored in the context of the linear minimization oracle.
 \citet{pethick2025sam} applied it to create $\nu$SAM, a new sharpness-aware minimization technique.
-The update from their Lemma 3.1 coincides with the update of Neon, but is used for completely different purposes.
-In addition, \citet{pethick2025sam} use power iterations to find $u_1$ and $v_1^\top$, while we suggest utilizing much more efficient and precise Lanczos algorithm.
-
-	{\bf The Ky Fan Norm and Dion.} Rank-$k$ Centralized Dion, Algorithm 1 from \citet{ahn2025dioncommunicationefficientoptimizerlarge}, without an error feedback and scaling of the update, turns out to be an lmo-based algorithm under the $\normkfk{\cdot}^\dagger$ norm, which we described in {\it Different norms imply different updates}.
-Reported by the authors necessity of using error feedback to obtain satisfactory accuracy may take place in the cases of our algorithms as well, for example, F-Neon.
-This we leave to future research.
+It would be interesting to substitute $\normn{\cdot}$ with $\normkfk{\cdot}^\dagger$ in their approach: $k > 1$ might increase accuracy boost, while preserving a small memory footprint and a small time overhead, if Dion-style power iterations are used.
 
 	{\bf Orthogonal research} Fanions, F-Fanions, and S-Fanions benefit from the general theoretical description of \citet{riabinin2025gluon}.
-They could be transformed to Drop-Fanions etc.
-(cite Drop-Muon article), they could be viewed as Broximal Point Methods (cite the article), they could be subject to EF (cite the EF for all).
-Of course, they could be clipped (cite ClipScion), and they could be analyzed when the LMO is not calculated accurately (cite Beyond the Ideal).
+They could be transformed to Drop-Fanions by updating only some layers as in \citet{gruntkowska2025dropmuon}.
+They could be viewed as approximations of the Non-Euclidean Broximal Point Method for the corresponding norms \citep{gruntkowska2025noneuclideanbroximal}.
+One can clip them to produce ClippedScion-like algorithms \citep{pethick2025generalizedgradientnormclipping}.
+Finally, the results from \citet{shulgin2025inexactmuon} can be used to theoretically derive faster yet working approximate schemes to calculate the LMO; power iterations with a very limited number of iterations is a good thing to start with.
 
@@
 If a theory is developed that explains the discrepancy between performance of different algorithms based on the matrix norms, one will probably be able to intentionally construct norms optimal to given architectures and probably even their parameters.
+
+We suggest that future works that deal with the non-Euclidean LMO explain in corollaries to their convergence theorems the empirical superiority of Muon to other Fanions.
+Without that, it is hard to believe that the bounds are relevant for practitioners.
 
@@
 
-\section{Updates derivations}
+%\section{Updates derivations}
 % we prove a general lemma instead
@@
 Indeed, by the trace property and singular vectors orthogonality, $\<\mM^k, \Delta> =\<\mU \mSigma \mV^\top, \alpha u_1 v_1^\top + (1-\alpha) \frac{\mU \mSigma \mV^\top}{\normf{\mM^k}}> = \alpha \tr \diag({\sigma_1, 0, \dots, 0}) + (1-\alpha) \normf{\mM^k} = \normftwo{\mM^k}$, which completes the proof.
-
-Proof of \cref{lemma:ky_fan_update}:
-Since $\norm{\cdot}^\dagger = \normkfk{\cdot}^{\dagger\dagger} = \normkfk{\cdot}$, the goal to reach $\normkfk{\mM^t}$.
-
-Let us note that $\Delta = \sum_{i=1}^{k} u_i v_i^\top$ delivers the value.
-Indeed, $\<\mM^t, \Delta> =\<\mU \mSigma \mV^\top, \sum_{i=1}^{k} u_i v_i^\top> = \sum_{i,j=1}^{r, k}\<u_i \sigma_i v_i^\top, u_j v_j^\top> = \sum_{i=1}^{k}\sigma_i = \normkfk{\mM^t}$, which completes the proof.
 

### 2025-11-24 code/L_smooth_tests/finetune_lrs.py
Added tuning on the quadratics

@@ -1,0 +1,417 @@
+from typing import Any, Dict, List, Optional, Tuple
+from math import sqrt
+import matplotlib.cm as cm
+import numpy as np
+import pandas as pd
+import os
+
+import torch
+
+from optimizers import Muon, Neon, NormalizedMuon, RandomNormalizedMuon, NeonMuon
+from mlion import MLion, Lion
+from L_smooth_tests.optimizer_runner import MatrixProblem, run_optimizer_on_problem
+from L_smooth_tests.benchmark_plotter import build_default_panels, plot_from_descriptions, plot_and_save_default_panels, save_experiments_to_csv
+from L_smooth_tests.problems import RandomQuadraticPSDProblem as SimpleQuadratic, LogisticRegressionProblem
+
+
+def grid_search_learning_rates(
+    algorithm_specs: List[Dict[str, Any]],
+    learning_rates: np.ndarray,
+    momentums: Optional[np.ndarray],
+    problem: MatrixProblem,
+    X_init: torch.Tensor,
+    csv_path: str,
+    loss_threshold: float = 0,
+    verbose: bool = False,
+) -> List[Dict[str, Any]]:
+    """
+    Perform grid search over learning rates and momentums for multiple algorithms.
+    Writes results to CSV incrementally as they are computed.
+    
+    Args:
+        algorithm_specs: List of algorithm specifications. Each spec should contain:
+            - name: str algorithm name
+            - optimizer_class: optimizer class
+            - optimizer_kwargs: dict of kwargs (without 'lr' and 'momentum', which will be set dynamically)
+            - num_iterations: int number of iterations
+            - record_interval: int interval to record metrics
+            - requires_cuda: bool (optional)
+            - use_momentum: bool (optional) - if True, momentum will be grid searched; if False/None, momentum from kwargs is used
+        learning_rates: Array of learning rates to try
+        momentums: Array of momentums to try (None if momentum should not be grid searched)
+        problem: MatrixProblem instance
+        X_init: Initial point tensor
+        csv_path: Path to CSV file for incremental writing
+        loss_threshold: Loss threshold for below_thresh_iter
+        verbose: Whether to print progress
+        
+    Returns:
+        List of dicts with keys: algorithm_name, learning_rate, momentum, below_thresh_iter
+    """
+    results: List[Dict[str, Any]] = []
+    device = X_init.device
+    
+    # Calculate total runs
+    if momentums is not None:
+        total_runs = len(algorithm_specs) * len(learning_rates) * len(momentums)
+    else:
[…]

### 2025-11-24 code/L_smooth_tests/optimizer_runner.py

@@ -27,6 +27,7 @@ def run_optimizer_on_problem(
     optimizer_kwargs: Dict[str, Any],
     problem: MatrixProblem,
     X_init: torch.Tensor,
+    loss_threshold: float = 0,
     num_iterations: int = 500,
     record_interval: int = 1,
     verbose: bool = False,
@@ -43,7 +44,7 @@ def run_optimizer_on_problem(
       - grad_spectral_norms: list[float]   (||∇f(X)||_2)
       - grad_nuclear_norms: list[float]    (||∇f(X)||_* )
     """
-    print(f"Testing {name}")
+    print(f"\nTesting {name}")
     X = X_init.clone()
 
     optimizer = optimizer_class([X], **optimizer_kwargs)
@@ -61,6 +62,7 @@ def run_optimizer_on_problem(
     start_time = time.time()
     overhead_time = 0.0  # time spent on costly gradient norm computations (excluded from timing)
     
+    iteration_reach_loss_thresh = -1
     for t in range(num_iterations):
         # Compute gradient and set it manually
         grad = problem.gradient(X)
@@ -71,7 +73,11 @@ def run_optimizer_on_problem(
                 cur_time = time.time() - start_time - overhead_time
                 iterations.append(t)
                 cumulative_time.append(cur_time)
-                losses.append(problem.objective(X).item())
+                loss = problem.objective(X).item()
+                losses.append(loss)
+                if loss < loss_threshold and iteration_reach_loss_thresh < 0:
+                    iteration_reach_loss_thresh = t
+                    print(f"Loss lower than {loss_threshold} on {t}th iteration")
                 grad_now = problem.gradient(X)
                 # Frobenius, spectral, and nuclear norms
                 _oh_start = time.time()
@@ -110,7 +116,7 @@ def run_optimizer_on_problem(
         X.grad = None
 
     print(f"Gradient norm computation overhead (excluded from timing): {overhead_time:.6f}s")
-
+    print(f"Last loss: {losses[-1]}")
     return {
         "iterations": iterations,
         "cumulative_time": cumulative_time,
@@ -119,6 +125,7 @@ def run_optimizer_on_problem(
         "grad_spectral_norms": grad_spectral_norms,
         "grad_nuclear_norms": grad_nuclear_norms,
         "final_X": X,
+        "below_thresh_iter": iteration_reach_loss_thresh,
     }
 
 

### 2025-11-24 code/L_smooth_tests/run_benchmarks_new.py

@@ -5,7 +5,7 @@ import numpy as np
 
 import torch
 
-from optimizers import Muon, Neon, NormalizedMuon, RandomNormalizedMuon
+from optimizers import Muon, Neon, NormalizedMuon, RandomNormalizedMuon, NeonMuon
 from mlion import MLion, Lion
 from L_smooth_tests.optimizer_runner import MatrixProblem, run_optimizer_on_problem
 from L_smooth_tests.benchmark_plotter import build_default_panels, plot_from_descriptions, plot_and_save_default_panels, save_experiments_to_csv
@@ -105,19 +105,21 @@ def main() -> None:
     # Inverse-square-root scheduler: lr(T) = lr0 / sqrt(T)
     def inv_sqrt_lr(T: int, base_lr: float) -> float:
         return base_lr / T**(1/2)
+    def const_lr(T: int, base_lr: float) -> float:
+        return base_lr
 
     # Define experiment specifications
     colors = cm.tab20b(np.linspace(0, 1, 5))[::-1]
     experiment_specs: Dict[str, Dict[str, Any]] = {
-        "NSGD": dict(
-            optimizer_class=NormalizedMuon,
-            optimizer_kwargs=dict(lr=0.1, momentum=0.6, nesterov=True, sgd_coeff=1),
-            num_iterations=iter_num_fro,
-            record_interval=100,
-            verbose=True,
-            lr_scheduler=inv_sqrt_lr,
-            color="gray"
-        ),
+        # "NSGD": dict(
+        #     optimizer_class=NormalizedMuon,
+        #     optimizer_kwargs=dict(lr=0.077, momentum=0.6, nesterov=True, sgd_coeff=1),
+        #     num_iterations=iter_num_fro,
+        #     record_interval=100,
+        #     verbose=True,
+        #     lr_scheduler=const_lr,
+        #     color="gray"
+        # ),
         #  "Random F-Muon": dict(
         #     optimizer_class=RandomNormalizedMuon,
         #     optimizer_kwargs=dict(lr=1, momentum=0.95, nesterov=True, sgd_coeff=0.5),
@@ -126,24 +128,43 @@ def main() -> None:
         #     verbose=True,
         #     lr_scheduler=inv_sqrt_lr,
         # ),
-        
-        # "Neon": dict(
-        #     optimizer_class=Neon,
-        #     optimizer_kwargs=dict(lr=1, nesterov=True, momentum=0.6, neon_mode="kyfan", iter_num=5),
+        # "NeonMuon": dict(
[…]

### 2025-11-24 code/cifar_bench/base_airbench.py

@@ -19,8 +19,8 @@ import torch.nn.functional as F
 import torchvision
 import torchvision.transforms as T
 from optimizers import Dion, Muon, Neon, NormalizedMuon, SGDMuon, SignSGDMuon, zeropower_via_newtonschulz5, RandomNormalizedMuon, NuclearNormalizedMuon, MuonCringeMomentum
-from optimizers import SpectrallyNormalizedNeon
-from mlion import MLion
+from optimizers import SpectrallyNormalizedNeon, MuonOrNSGD, RealFanion, NeonMuon, SingleDeviceNorMuon
+from mlion import MLion, NLion
 
 #############################################
 #               Muon optimizer              #
@@ -346,7 +346,7 @@ def main(run, model):
     if run == 'warmup':
         # The only purpose of the first run is to warmup the compiled model, so we can use dummy data
         train_loader.labels = torch.randint(0, 10, size=(len(train_loader.labels),), device=train_loader.labels.device)
-    total_train_steps = ceil(12 * len(train_loader))
+    total_train_steps = ceil(8 * len(train_loader))
     whiten_bias_train_steps = ceil(3 * len(train_loader))
 
     # Create optimizers and learning rate schedulers
@@ -359,10 +359,18 @@ def main(run, model):
     optimizer1 = torch.optim.SGD(param_configs, momentum=0.85, nesterov=True)#, fused=True)
     # random mix, 93.3%, 11.26 s on bs 2000 with lr=0.4, mom=0.65
     # optimizer2 = RandomNormalizedMuon(filter_params, lr=0.24, momentum=0.6, sgd_coeff=0.5, nesterov=True) # and 92.9% for bs 200
-    # optimizer2 = NormalizedMuon(filter_params, lr=0.4, momentum=0.65, sgd_coeff=0.5, nesterov=True) # the best tuned F-Muon, 94.0%
+    # optimizer2 = NormalizedMuon(filter_params, lr=0.4, momentum=0.65, sgd_coeff=0.5, nesterov=True) # the best tuned F-Muon, 94.0% for bs = 2000; 51.2 +-11 for bs = 10000; 92.56% for bs=200
[…]

### 2025-11-24 code/mlion.py

@@ -2,8 +2,48 @@
 import torch
 from optimizers import zeropower_via_newtonschulz5
 from torch.optim import Optimizer
+from matrix_methods.matrix_functions import several_sv_svds_approximation
 
-class MLion(Optimizer):
+class WrongMLion(Optimizer):
+    def __init__(self, params, lr=1e-4, betas=(0.9, 0.99), weight_decay=0.0):
+        defaults = dict(lr=lr, betas=betas, weight_decay=weight_decay)
+        super().__init__(params, defaults)
+
+    @torch.no_grad()
+    def step(self, closure=None):
+        loss = None
+        if closure is not None:
+            loss = closure()
+
+        for group in self.param_groups:
+            lr = group['lr']
+            beta1, beta2 = group['betas']
+            wd = group['weight_decay']
+
+            for p in group['params']:
+                if p.grad is None:
+                    continue
+
+                g = p.grad
+                if wd != 0:
+                    g = g.add(p, alpha=wd)
+
+                
+                state = self.state[p]
+
+                
+                if len(state) == 0:
+                    state['exp_avg'] = torch.zeros_like(p)
+                exp_avg = state['exp_avg']
+                exp_avg.mul_(beta1).add_(g, alpha=1 - beta1)
+                
+                update = zeropower_via_newtonschulz5(exp_avg.reshape(len(g), -1)).view(exp_avg.shape) # whiten the update
+                # p.add_(torch.sign(exp_avg), alpha=-lr)
+                p.add_(update, alpha=-lr)#  * max(1, p.size(-2) / p.size(-1)) ** 0.5)
+
+        return loss
+
+class NLion(Optimizer): # does not work for airbench
     def __init__(self, params, lr=1e-4, betas=(0.9, 0.99), weight_decay=0.0):
         defaults = dict(lr=lr, betas=betas, weight_decay=weight_decay)
         super().__init__(params, defaults)
@@ -33,11 +73,14 @@ class MLion(Optimizer):
                 exp_avg = state['exp_avg']
                 exp_avg.mul_(beta1).add_(g, alpha=1 - beta1)
                 
-                update = zeropower_via_newtonschulz5(exp_avg.reshape(len(g), -1)).view(exp_avg.shape) # whiten the update
+                u, s, vt = several_sv_svds_approximation(exp_avg.reshape(len(g), -1), 1, 10)
+                update = (u @ vt).view(exp_avg.shape)
                 # p.add_(torch.sign(exp_avg), alpha=-lr)
[…]

### 2025-11-24 code/optimizers.py

@@ -285,6 +285,7 @@ class Neon(torch.optim.Optimizer):
                 p.data.add_(update2, alpha=-lr)
 
 
+
 # the same as F-Muon, the copy is in airbench_muon.py
 class NormalizedMuon(torch.optim.Optimizer):
     def __init__(self, params, lr=1e-3, momentum=0, nesterov=False, sgd_coeff=0):
@@ -434,36 +435,10 @@ class Dion(torch.optim.Optimizer):
         self.rank = rank
         self.momentum_decay = momentum_decay
         self.sgd_coeff = sgd_coeff
-    '''
-    def power_iter1(self, B, Q):
-        """
-        Single power iteration (from Q)
-        B: gradient matrix (m x n)
-        Q: right factor from previous iteration (n x r)
-        Returns: P (m x r), R (n x r) where P is orthonormal
-        """
-        # P ← BQ
-        P = B @ Q
-        
-        # P ← Orthogonalize(P) using QR decomposition
-        P, _ = torch.linalg.qr(P)
-        
-        # R ← B^T P
-        R = B.T @ P
-        
-        return P, R
 
-    def column_normalize(self, R):
-        """Normalize columns of R to unit norm"""
-        norms = torch.norm(R, dim=0, keepdim=True)
-        # Add small epsilon for numerical stability
-        norms = torch.clamp(norms, min=1e-8)
-        return R / norms
-    '''
     def step(self):
         for group in self.param_groups:
             lr = group['lr']
-            momentum = group['momentum']
             for p in group['params']:
                 g = p.grad
                 if g is None:
@@ -474,25 +449,15 @@ class Dion(torch.optim.Optimizer):
                 # Initialize state variables if not present
                 if 'momentum_buffer' not in state:
                     state['momentum_buffer'] = torch.zeros_like(g)
-                '''
-                if 'Q' not in state:
-                    # Initialize Q randomly with correct shape
-                    n = p.numel()
-                    m = g.numel()
-                    if m > n:
-                        state['Q'] = torch.randn(n, self.rank, device=p.device, dtype=p.dtype)
-                    else:
-                        state['Q'] = torch.randn(m, self.rank, device=p.device, dtype=p.dtype)
-                
-                Q = state['Q']
-                '''
                 buf = state['momentum_buffer']
                 
                 # Apply momentum
-                buf.mul_(momentum).add_(g)
[…]

### 2025-11-24 code/simple_network.py

@@ -183,7 +183,7 @@ MODEL_CONFIGS = {
         'input_size': 32 * 32 * 3,
         'hidden_sizes': [2048, 1024, 512],
         'output_size': 10,
-        'dropout': 0.1
+        'dropout': 0
     },
 }
 
@@ -316,11 +316,7 @@ def evaluate(model, loader, tta_level=0):
 
 def main(run, model):
 
-    batch_size = 2000
-    bias_lr = 0.053
-    head_lr = 0.67
-    wd = 2e-6 * batch_size
-
+    batch_size = 10
     test_loader = CifarLoader("cifar10", train=False, batch_size=2000)
     train_loader = CifarLoader("cifar10", train=True, batch_size=batch_size, aug=dict(flip=True, translate=2))
     if run == "warmup":
@@ -330,7 +326,9 @@ def main(run, model):
     whiten_bias_train_steps = ceil(3 * len(train_loader))
 
     # optimizer2 = NormalizedMuon(model.parameters(), lr=0.1, momentum=0.6, nesterov=True, sgd_coeff=0)
-    optimizer2 = Muon(model.parameters(), lr=0.1, momentum=0.6, nesterov=True)
+    opt_mode = 'custom'
+    optimizer2 = Muon(model.parameters(), lr=0.1, momentum=0, nesterov=False)
+
     
     # optimizer2 = torch.optim.Adam(model.parameters(), lr=0.0001)
     optimizers = [optimizer2]
@@ -370,11 +368,78 @@ def main(run, model):
         model.train()
         for inputs, labels in train_loader:
             outputs = model(inputs) # , whiten_bias_grad=(step < whiten_bias_train_steps))
-            F.cross_entropy(outputs, labels, label_smoothing=0.2, reduction="sum").backward()
-            for group in optimizer2.param_groups:
-                group["lr"] = group["initial_lr"] * (1 - step / total_train_steps)
-            for opt in optimizers:
-                opt.step()
+            if opt_mode == 'custom':
+                losses = F.cross_entropy(outputs, labels, label_smoothing=0.2, reduction="none")
+
+                # Prepare storage for accumulated gradients
+                param_sums = [torch.zeros_like(p) for p in model.parameters()]
+                us = [[] for p in model.parameters()]
+                vs = [[] for p in model.parameters()]
+
+                for i in range(len(losses)):
+                    # Compute gradient for single sample
+                    grads = torch.autograd.grad(
+                        losses[i],
+                        model.parameters(),
[…]

### 2025-11-25 conf_article/neon.tex
provided proofs in section 4

@@
 Our experiments reveal important insights about the role of matrix norms in optimization.
-On a synthetic least squares problem, we observe a striking discrepancy: while some algorithms converge slowly in terms of loss, they may converge quickly in terms of gradient norm, and vice versa.
+\textcolor{red}{On a synthetic least squares problem, we observe a striking discrepancy: while some algorithms converge slowly in terms of loss, they may converge quickly in terms of gradient norm, and vice versa}.
 This suggests that existing theoretical guarantees may not fully explain practical algorithm performance.
@@
 
-\section{Duals to Ky Fan norms instead of the spectral norm}
+\section{Duals to Ky Fan Norms Instead of the Spectral Norm}
 
@@
 
-\section{Conic combination of LMO-algorithms is an LMO-algorithm}
+\section{Conic Combination of LMO-algorithms is an LMO-algorithm}
 
-NSGD, Muon, and SignSGD work fair enough separately. Could we boost the quality if we combined them? To answer this question, we have to build a theoretical base for that.
+Simply applying norm dual to the Ky Fank $k$-norm however, is not enough to provide family of algorithms diverse enough for our cause.
+Next we consider linear combinations of algorithms from LMO framework, which happen to be LMO-algorithms too, as we show in succeeding paragraphs.
 
-\subsection{Duals to Convex Combination of Norm}
-First, we need a well-known fact mentioned, for example, in \citet[Table 1]{yu2012arithmetic}. For the sake of completeness, we provide a proof of the fact.
+\subsection{General Case}
+First, we need a well-known fact mentioned, for example, in \citet[Table 1]{yu2012arithmetic}. For the sake of completeness, we provide a proof here.
 
 \begin{lemma}\label{lemma:dual_to_conv_comb}
-	Let $\norm{\cdot}_{(1)}$ and $\norm{\cdot}_{(2)}$ be norms on a finite-dimensional Euclidean space, and let $\alpha,\beta \geq 0$. Define
+	Let $\norm{\cdot}_{(1)},\dots,\norm{\cdot}_{(n)}$ be norms on a finite-dimensional Euclidean space, and let $\alpha_1,\dots,\alpha_n$ be non-negative reals. Define
 	$$
-		\norm{x} := \alpha \norm{x}_{(1)} + \beta \norm{x}_{(2)}.
+		\norm{x} := \sum_{i = 1}^n \alpha_i \norm{x}_{(i)}.
 	$$
@@
 	$$
-		B_{\norm{\cdot}^\dagger}
-		= \alpha B_{\norm{\cdot}_{(1)}^\dagger} + \beta B_{\norm{\cdot}_{(2)}^\dagger},
+		\cB_{\norm{\cdot}^\dagger}
+		= \sum_{i = 1}^n \alpha_i \cB_{\norm{\cdot}_{(i)}^\dagger}
 	$$
-	where $+$ denotes the Minkowski sum and $B_{\norm{\cdot}_{(i)}^\dagger}$ is the unit ball of the dual norm $\norm{\cdot}_{(i)}^\dagger$.
+	where $\sum$ denotes the Minkowski sum and $\cB_{\norm{\cdot}_{(i)}^\dagger}$ is the unit ball of the dual norm $\norm{\cdot}_{(i)}^\dagger$.
 \end{lemma}
@@
 \begin{proof}
-	Write $f(x) = \alpha \norm{x}_{(1)}$ and $g(x) = \beta \norm{x}_{(2)}$, so
-	$$
-		\norm{x} = f(x)+g(x).
-	$$
-	Recall two standard facts:
+	Let us first prove the lemma for the case $n = 2$. Denote $f(x) = \alpha_1 \norm{x}_{(1)}$ and $g(x) = \alpha_2 \norm{x}_{(2)}$, such that $\norm{x} = f(x) + g(x)$. Recall two standard facts:
 	\begin{enumerate}
@@
 			      \sup_{x}\bigl( \langle y,x \rangle - \lambda \norm{x}\bigr)
-			      = \delta_{\lambda B_{\norm{\cdot}^\dagger}}(y),
+			      = \delta_{\lambda \cB_{\norm{\cdot}^\dagger}}(y)\,,
 		      $$
@@
 		      $$
-			      (f+g)^*(y) = \inf_{u+v=y} \bigl(f^*(u) + g^*(v)\bigr).
+			      (f+g)^*(y) = \inf_{u+v=y} \bigl(f^*(u) + g^*(v)\bigr)\,.
 		      $$
@@
 	$$
-		f^*(u) = \delta_{\alpha B_{\norm{\cdot}_{(1)}^\dagger}}(u),
+		f^*(u) = \delta_{\alpha_1 \cB_{\norm{\cdot}_{(1)}^\dagger}}(u)\,,
 		\quad
-		g^*(v) = \delta_{\beta B_{\norm{\cdot}_{(2)}^\dagger}}(v).
+		g^*(v) = \delta_{\alpha_2 \cB_{\norm{\cdot}_{(2)}^\dagger}}(v)\,.
 	$$
-	Thus
+	Thus,
 	$$
@@
 		\bigl(
-		\delta_{\alpha B_{\norm{\cdot}_{(1)}^\dagger}}(u)
+		\delta_{\alpha_1 \cB_{\norm{\cdot}_{(1)}^\dagger}}(u)
 		+
-		\delta_{\beta B_{\norm{\cdot}_{(2)}^\dagger}}(v)
+		\delta_{\alpha_2 \cB_{\norm{\cdot}_{(2)}^\dagger}}(v)
 		\bigr)
 		=
-		\delta_{\alpha B_{\norm{\cdot}_{(1)}^\dagger}+\beta B_{\norm{\cdot}_{(2)}^\dagger}}(y).
+		\delta_{\alpha_1 \cB_{\norm{\cdot}_{(1)}^\dagger}+\alpha_2 \cB_{\norm{\cdot}_{(2)}^\dagger}}(y).
 	$$
-	But by definition, the conjugate of a norm is exactly the indicator of its dual unit ball:
+	By definition, the conjugate of a norm is exactly the indicator of its dual unit ball: 
 	$$
-		\norm{\cdot}^*(y) = \delta_{B_{\norm{\cdot}^\dagger}}(y).
+	\norm{\cdot}^*(y) = \delta_{\cB_{\norm{\cdot}^\dagger}}(y)\,.
 	$$
-	Therefore,
+	Therefore, $\cB_{\norm{\cdot}^\dagger} = \alpha_1 \cB_{\norm{\cdot}_{(1)}^\dagger} + \alpha_2 \cB_{\norm{\cdot}_{(2)}^\dagger}$.
+
+	Now we prove the general case by induction. The base case ($n=2$) is already proven. Suppose that the assumption of the lemma holds for $n = k$. Then, for $n = k + 1$, 
 	$$
-		B_{\norm{\cdot}^\dagger} = \alpha B_{\norm{\cdot}_{(1)}^\dagger} + \beta B_{\norm{\cdot}_{(2)}^\dagger}.
+	\norm{x} = \sum_{i = 1}^k \alpha_i \norm{x}_{(i)} + \alpha_{k + 1}\norm{x}_{(k + 1)} = \norm{x}_{(1:k)} + \alpha_{k + 1}\norm{x}_{(k + 1)}\,.
 	$$
+	Applying the result for $n=2$ combined with the induction assumption, we obtain
+	$$
+	\cB_{\norm{\cdot}^\dagger} = \cB_{\norm{\cdot}_{(1:k)}^\dagger} + \alpha_{k + 1} \cB_{\norm{\cdot}_{(k + 1)}^\dagger} = \sum_{i = 1}^{k + 1} \alpha_i \cB_{\norm{\cdot}_{(i)}^\dagger}\,,
+	$$
+	which proves the lemma.
 \end{proof}
 
-Consequently,
-\begin{equation}\label{eq:dual_to_conv_comb}
-	\norm{y}^\dagger
-	= \inf\bigl\{ t \ge 0 : y \in t\bigl(\alpha B_{\norm{\cdot}_{(1)}^\dagger} + \beta B_{\norm{\cdot}_{(2)}^\dagger}\bigr) \bigr\}
-	= \inf_{z,t}\bigl\{t \ge 0: z \in t \alpha B_{\norm{\cdot}_{(1)}^\dagger}, y - z \in t \beta B_{\norm{\cdot}_{(1)}^\dagger}\bigr\}
-\end{equation}
+\begin{lemma}\label{lemma:lin_colm_alg}
+	Let $\norm{\cdot}_{(1)}, \dots, \norm{\cdot}_{(n)}$ be norms on a finite-dimensional Euclidean space, and let $\alpha_1, \dots, \alpha_n$ be non-negative reals.
+Consider Linear Minimization Oracles $\mathrm{LMO}_{1}, \dots, \mathrm{LMO}_{n}$, corresponding to these norms.
+Then, $\sum_{i = 1}^n \alpha_i \mathrm{LMO}_{i}$ is the LMO corresponding to the norm $\norm{\cdot}^\dagger$ dual to the $\sum_{i = 1}^n \alpha_i \norm{\cdot}_{(i)}^\dagger$.
+\end{lemma}
 
-Now we can work with different norms.
+\begin{proof}
+	From \cref{lemma:dual_to_conv_comb} we know the general form of the unit ball in the $\norm{\cdot}$ norm: $\cB_{\norm{\cdot}} = \sum_{i = 1}^n \alpha_i \cB_{\norm{\cdot}_{(i)}^\dagger}$.
+Thus, the linear function minimization objective on a $\norm{\cdot}$ norm ball can be transformed as follows:
+	$$
+	\argmin_{\mD \in \cB_{\norm{\cdot}}}\<\mM, \mD> = \argmin_{\mD_1 \in \alpha_i \cB_i,\dots,\mD_n \in \alpha_n\cB_n}\<\mM, \sum_{i = 1}^n\mD_i> = \sum_{i = 1}^n \argmin_{D_i \in \cB_i}\<\mM, \mD_i>\,.
+	$$
+	This immediately implies $\sum_{i = 1}^n \alpha_i \mathrm{LMO}_{i} \in \argmin_{\mD \in \cB_{\norm{\cdot}}}\<\mM, \mD>$, which proves the claim of the lemma.
+\end{proof}
 
-\subsection{General case}
-We have norm \#1 with update \#1 and norm \#2 with update \#2. We prove the lemma that says that the update of their conic combination is a conic combination of updates \#1 and \#2.
+\subsection{Frobeniusize Them: F-Muon and F-Neon}
+Let us construct the concrete examples of algorithms given by linear combinations of LMO-algorithms.
+It follows from \cref{lemma:lin_colm_alg} that those algorithms can also be viewed as LMO-algorithms.
 
-Then prove a lemma for any set of norms \#1, \dots, \#n (by induction)
-
-\subsection{Frobeniusize them: F-Muon and F-Neon}
-Let us consider a mixture of Fanions with another excellent LMO-algorithm, NSGD.
-
-We define $\normfstar{\cdot}$ as a convex combination of $\normn{\cdot}$ and $\normf{\cdot}$:
+Consider a mixture of Fanions with another excellent LMO-algorithm, NSGD. We define $\normfstar{\cdot}$ as a convex combination of $\normn{\cdot}$ and $\normf{\cdot}$:
 
@@
 
-\section{Computing the updates}\label{sec:matrix-side}
+\section{Computing the Updates}\label{sec:matrix-side}
 To compute the algorithms' updates, we use thick-restarted Lanczos method for singular value problem (TRLan) on $\mM^{k\top}\mM^k$ or $\mM^k \mM^{k\top}$ matrices (the one with less size is picked), implemented in CuPy lSibrary \citep{cupy_svds_ref} and described in \citet{simonz1998thick}.

### 2025-11-25 conf_article/neon.tex
Added TRLan analysis from Nikolay

@@
 \section{Computing the Updates}\label{sec:matrix-side}
-To compute the algorithms' updates, we use thick-restarted Lanczos method for singular value problem (TRLan) on $\mM^{k\top}\mM^k$ or $\mM^k \mM^{k\top}$ matrices (the one with less size is picked), implemented in CuPy lSibrary \citep{cupy_svds_ref} and described in \citet{simonz1998thick}.
+We employ the thick-restarted Lanczos method for the symmetric eigenvalue problem (thick-restart Lanczos, TRLan) to compute the low-rank updates of Fanions.
+We apply TRLan to either $\mM^{k\top}\mM^k$ or $\mM^k \mM^{k\top}$ (whichever matrix is smaller).
+We use the CuPy implementation of \texttt{cupy.sparse.linalg.svds} \citep{cupy_svds_ref} which internally relies on TRLan \citep{simonz1998thick}.
 
-This method is designed for efficiently approximating the largest singular values and vectors of large matrices.
-Its thick-restart strategy retains the most informative Ritz vectors at each cycle, which accelerates convergence while avoiding excessive memory growth.
-We are specifically interested in this algorithm because it allows us to extract several largest singular values and related singular vectors of the matrix to make a Neon step.
-Moreover, TRLan is stable GPU-friendly algorithm because it mainly operates with matrix-vector multiplications (MVs), which are higly-parallel, and does not require full reorthogonalization against the whole Krylov basis by managing  short recurrent formulas and incorparating thick restarting.
+TRLan is specifically designed for efficiently approximating the largest singular values and corresponding singular vectors of very large matrices.
+The thick-restart strategy retains the most informative Ritz vectors across restarts, which dramatically accelerates convergence while keeping memory consumption moderate.
+TRLan is particularly attractive in our GPU setting because its dominant cost is a modest number of highly parallelizable matrix-vector multiplications (matvecs) and it avoids full reorthogonalization against the entire Krylov basis by using short recurrence relations combined with thick restarting.
 
-Per-cycle complexity is $O(mn^2 + n^2k + nk^2)$, where m and n are dimensions of the target matrix and n is the smaller one, k is retained subspace's size.
+The per-restart complexity is $\mathcal{O}(mn^2 + n^2k + nk^2)$, where $m \geq n$ are the dimensions of the target matrix and $k$ is the size of the retained subspace (typically $k \ll n$).
 
-In \cref{tbl:matrix_methods}, we compare performance of TRLan, RSVD, and power iterations on calculating k-rank update, which is used in k-Fanion.
-The results highlight that TrLan is much faster that its competitors.
+In Table~\ref{tbl:matrix_methods}, we compare TRLan against randomized SVD (RSVD) and simple power iterations when computing the rank-$k$ update used in Fanion-$k$ and related algorithms.
+Experiments are performed on dense random matrices with i.i.d.
+$\mathcal{N}(0,1)$ entries and use CPU implementations for fair comparison.
+We report:
+\begin{itemize}
+  \item $err_1$: relative error in the Frobenius norm of $\sum_i^k u_i v_i^T$,
+  \item $err_2$: relative error in the Frobenius norm of $\sum_i^k \sigma_i u_i v_i^T$.\end{itemize}
 
-During the research it was noted that RSVD can give good and fast approximation for singular values, but the matrix of approximation is far from the one given by truncated SVD, while TRLan gives good and fast approximation of a matrix, but not so good approximation for singular values.
-That means that TRLan may be not a perfect choice for algorithms like Dion, where $\sigma_i$ are required for error feedback.
+On $500\times500$ matrices, TRLan and RSVD require comparable wall-clock time, but TRLan delivers orders-of-magnitude lower error and far fewer matvecs.
+On larger $5000\times5000$ matrices the advantage becomes even more pronounced: TRLan is 3-4 times faster than RSVD while using $\sim$30 times fewer matvecs at comparable or better accuracy.
 
-The practical drawback of the TRLAn is the absence of its implementation in PyTorch.
-\begin{table}[h!]
+An interesting empirical observation is that RSVD tends to approximate the \emph{singular values} themselves reasonably well but the reconstructed low-rank matrix noticeably deviates from the truncated SVD, whereas TRLan provides an excellent approximation to the truncated SVD matrix itself (low $err_2$) at the cost of occasionally less accurate individual singular values.
+This makes TRLan the preferred choice for algorithms like Neon/Fanion-$k$ that only need the low-rank term $\sum \sigma_i u_i v_i^\top$, but less ideal for methods (e.g., Dion) that explicitly require accurate $\sigma_i$ for error feedback or step-size control.
+
+A current practical limitation is the lack of a native PyTorch implementation of thick-restart Lanczos; existing PyTorch-based randomized SVD routines cannot match TRLan's accuracy/efficiency combination for the matrix reconstruction task.
+
+For reference, Table~\ref{tbl:newton_schulz_reference} shows results for the Newton-Schulz polar decomposition iteration on the same matrices, $err_1$ is the relative error of $UV^T$ (29-30 iterations to converge, significantly higher matvec count than TRLan).
+
+\begin{table}[ht]
 	\centering
-	\begin{tabular}{c|c|c|c}
+	\begin{tabular}{|c|c|c|c|c|c|c|c|}
 	  \hline
-		Method                              & rtol & k   & time,s \\
+	  Matrix sizes & $k$  & Method           & Time (s) & Matvecs & Iterations & $err_1$         & $err_2$       \\
 	  \hline
-		Power Iterations                    & 0.01 & 1   & 7.7    \\
-		SVDS (thick-restart Lanczos method) & 0.01 & 1   & 0.18   \\
-		PCA Low Rank (RSVD)                 & 0.01 & 1   & 1.15   \\
-		SVDS (thick-restart Lanczos method) & 0.01 & 10  & 0.47   \\
-		PCA Low Rank (RSVD)                 & 0.01 & 10  & 19.4   \\
-		SVDS (thick-restart Lanczos method) & 0.01 & 100 & 1.96   \\
-		PCA Low Rank (RSVD)                 & 0.01 & 100 & 170    \\
+	  $500 \times 500$      & 5  & Power Iterations           & 0.062   & 2005    & 200        & 9.2$\cdot 10^{-3}$ & 9.1$\cdot10^{-3}$\\
+	  $500 \times 500$      & 5  & RSVD             & 0.017   & 1170    & 38         & 9.8$\cdot 10^{-3}$ & 9.6$\cdot10^{-3}$\\
+	  $500 \times 500$      & 5  & TRLan            & 0.018   & 131     & 65         & 9.6$\cdot 10^{-5}$ & 9.4$\cdot10^{-5}$\\
+	  \hdashline
+	  $500 \times 500$      & 50 & Power Iterations           & 0.44    & 43750   & 437      & 9.9$\cdot 10^{-3}$ & 9.0$\cdot10^{-3}$\\
+	  $500 \times 500$      & 50 & RSVD             & 0.61    & 6120    & 50      & 9.9$\cdot 10^{-3}$ & 9.1$\cdot10^{-3}$\\
+	  $500 \times 500$      & 50 & TRLan            & 0.16    & 462     & 231      & 3.3$\cdot 10^{-7}$ & 3.0$\cdot10^{-7}$\\
+	  \hdashline 
+	  $5000 \times 5000$    & 5  & Power Iterations           & 9.6     & 9065    & 906      & 8.6$\cdot 10^{-3}$ & 8.6$\cdot10^{-3}$\\
+	  $5000 \times 5000$    & 5  & RSVD             & 2.1     & 5640    & 187      & 9.7$\cdot 10^{-3}$ & 9.7$\cdot10^{-3}$\\
+	  $5000 \times 5000$    & 5  & TRLan            & 0.70    & 205     & 102      & 7.7$\cdot 10^{-3}$ & 7.7$\cdot10^{-3}$\\
+	  \hline
 	\end{tabular}
-	\caption{k-rank updated comparison}
-	Comparison of different numerical methods to calculate k-rank update on $5000\times5000$ matrix of real numbers, rtol is an error in Frobenius norm relative to the k-rank approximation.
+  \caption{Comparison of methods for computing rank-$k$ updates on dense random matrices (CPU, double precision). Lower is better in all columns.}
   \label{tbl:matrix_methods}
+  \end{table}
+
+  \begin{table}[ht]
+  \centering
+  \begin{tabular}{|c|c|c|c|c|}
+  \hline
+  Matrix size    & Time (s) & Matvecs & Iterations & $err_1$ \\
+  \hline
+  $500 \times 500$     & 0.041    & 27 000  & 27         & 4.8e-3  \\
+  $5000 \times 5000$   & 26.4     & 290 000 & 29         & 6.5e-3  \\
+  \hline
+  \end{tabular}
+  \caption{Newton–Schulz iteration on the same matrices (for reference).}
+  \label{tbl:newton_schulz_reference}
   \end{table}
@@
 \subsection{Randomized Linear Least Squares}
-First, we compare F-Fanions on the following L-smooth problem:
+First, we compare F-Fanions on the following convex L-smooth problem:
 
@@
 
-where $\mX \in \Rmn$, $m=500$ and $n=500$ are typical dimensions of a neural network weight matrix, $S \in \Rmn$, $\mM \in \mathbb{S}^m_{+}$ and $\mN \in \mathbb{S}^n_{+}$ are positive-semidefinite matrices.
+where $\mX \in \Rmn$, $m=500$ and $n=500$ are typical dimensions of a neural network weight matrix, $\mS \in \Rmn$, $\mM \in \mathbb{S}^m_{+}$ and $\mN \in \mathbb{S}^n_{+}$ are positive-semidefinite matrices.
 The spectra of $\mM$ and $\mN$ are uniformly distributed in the (0, 1) interval.
@@
 
-Since theoretical bounds \citep{kovalev2025understanding, riabinin2025gluon} rely on a very loose norm bound $\norm{\cdot}\le \rho \normf{\cdot}$, we do not derive learning rate from smoothness.
-Rather, we set the learning rate of $1/\sqrt{T}$ for each optimizer and the momentum of 0.6.
-12 000 iterations for each algorithm ensure that the rate at some point gets low enough for each algorithm to converge.
-The results are presented in \cref{fig:lls} and ~\cref{sec:lls_plots_section}.
+Since theoretical bounds \citep{kovalev2025understanding, riabinin2025gluon} rely on a very loose norm bound $\norm{\cdot}\le \rho \normf{\cdot}$, we do not derive learning rate or momentum from the smoothness constants.
+Rather, we find such \texttt{lr, momentum} pair that the loss threshold $0.01$ is reached in a minimal number of iterations.
+This setting is the most realistic, and at the same time aligns with the corollaries of the convergence theorems where constant learning rate and momentum coefficients are proposed.
+
+12 000 iterations for each algorithm ensure that the rate at some point gets low enough for each algorithm to converge. The results are presented in \cref{fig:lls} and ~\cref{sec:lls_plots_section}.
 
@@
 \subsection{NanoGPT speedrun}
-We test F-Muon on NanoGPT speedrun \citep{modded_nanogpt_2024}.
+We test F-Muon on the NanoGPT speedrun \citep{modded_nanogpt_2024}.
 For $\alpha = 0.5$, the optimal parameters are {\tt lr=0.07, momentum=0.95}, while for Muon they were {\tt lr=0.05, momentum=0.95}.
-After testing for 1750 steps, as it has been done on the speedruns, we get 3.281 cross-entropy loss, while on Muon, it falls below the target threshold 3.28 reaching 3.279.
+After testing for 1750 steps we get 3.281 cross-entropy loss.
+For Muon, it falls below the target threshold 3.28 and reaches 3.279.
 However, this difference is negligible, if one looks at \cref{fig:gpt_loss}.
-It is even more striking, considering the fact that F-Muon is an average of Muon and NSGD, and the later performed quite poorly.
+It is even more striking, considering the fact that F-Muon is an average of Muon and NSGD, and the latter performed quite poorly.
 

### 2025-11-25 conf_article/neon.tex
edited section 4 even more

@@
 \newtheorem{lemma}{Lemma}
+\newtheorem{corollary}{Corollary}
 \newcommand{\norm}[1]{\lVert #1\rVert}
@@
 \begin{equation}\label{eq:simple_update}
-	\mX^{k + 1} = \mX^k + \gamma_k\mathrm{LMO}\left(\mM\right)\,.
+	\mX^{k + 1} = \mX^k + \gamma_k\mathrm{LMO}\left(\mM^k\right)\,.
 \end{equation}
 
-We are particularly interested in the case when $\cS$ is a ball in the norm $\norm{\cdot}$:
+We are particularly interested in the case when $\cS$ is a unit ball in the norm $\norm{\cdot}$:
 
-$$\cS = \cB_\eta \in \{ \mD \in \Rmn \mid \norm{\mD} \leq \eta \}.$$
+$$
+\cS = \cB_{\norm{\cdot}} = \{ \mD \in \Rmn \mid \norm{\mD} \leq 1 \}\,.
+$$
 In this case,
-
-$$\argmin_{\mD \in \cS} \<\mM^k, \mD> = - \eta \{ \mD \in \cB_1 \mid \<\mM^k, \mD> = \norm{\mM^k}^\dagger\}$$
-
+$$
+\argmin_{\mD \in \cS} \<\mM^k, \mD> = - \{ \mD \in \cB_1 \mid \<\mM^k, \mD> = \norm{\mM^k}^\dagger\}\,,
+$$
 and update for $\mX^k$ in \cref{eq:simple_update} simplifies to
 \begin{equation}\label{eq:our_update}
-	\mX^{k+1} = \mX^{k} - \eta_k \{\mD \in \cB_1 \mid \<\mM^k, \mD> = \norm{\mM^k}^\dagger\}\,
+	\mX^{k+1} = \mX^{k} - \gamma_k \{\mD \in \cB_1 \mid \<\mM^k, \mD> = \norm{\mM^k}^\dagger\}\,.
 \end{equation}
 
-By this formula, it is easy to compute updates for algorithms induced by various norms $\norm{\cdot}$.
+Using this formula it is easy to compute updates for algorithms induced by various norms $\norm{\cdot}$.
 For example, when the norm $\norm{\cdot}$ is the Frobenius norm $\normf{\cdot}$, \cref{eq:our_update} turns into
-
 \begin{equation}\label{eq:nsgd_update}
-	\mX^{k+1} = \mX^{k} - \eta_k \frac{\mM^k}{\normf{\mM^k}},
-\end{equation} which recovers Normalized SGD (NSGD). When one has the spectral norm $\norms{\cdot}$, one gets
-
+	\mX^{k+1} = \mX^{k} - \eta_k \frac{\mM^k}{\normf{\mM^k}}\,,
+\end{equation} 
+which recovers Normalized SGD (NSGD). When one has the spectral norm $\norms{\cdot}$, one gets
 \begin{equation}\label{eq:muon_update}
-	\mX^{k+1} = \mX^{k} - \eta_k U V^\top,
-\end{equation} which is Muon without the $\sqrt{m/n}$ factor.
+	\mX^{k+1} = \mX^{k} - \eta_k U V^\top\,,
+\end{equation} 
+which is Muon without the $\sqrt{m/n}$ factor.
 Here, $\mM^k = \mU \mSigma \mV^\top$ is the Singular Value Decomposition (SVD) of $\mM^k$ ($\mU = [u_1, u_2, \dots, u_r]$, $\mSigma = \diag(\sigma_1, \space \sigma_2, \dots, \sigma_r)$, and $\mV = [v_1, v_2, \dots, v_r]$).
@@
 \subsection{\texorpdfstring{$\normn{\mM^{k}}$ and Neon}{Neon}}
-After having considered $\normf{\mM^k}$ and $\norms{\mM^k}$, it is natural to look at the nuclear norm $\normn{\mM^k}$.
+After considering $\normf{\mM^k}$ and $\norms{\mM^k}$, it is natural to look at the nuclear norm $\normn{\mM^k}$.
 \begin{lemma}\label{lemma:neon_update}
@@
 Thus, the update is either
-$$\mX^{t+1} = \mX^{t} - \eta_t u_1 v_1^\top\text{ or }\mX^{t+1} = \mX^{t} - \frac{\eta_t}{k}\mU \mV^\top,$$
+$$
+\mX^{t+1} = \mX^{t} - \eta_t u_1 v_1^\top\text{ or }\mX^{t+1} = \mX^{t} - \frac{\eta_t}{k}\mU \mV^\top\,,
+$$
 depending on which one better minimizes the function $F(\mX^t) + \mM^t(\mX^{t+1} - \mX^t)$.
@@
 
-\begin{lemma}\label{lemma:lin_colm_alg}
+\begin{lemma}\label{lemma:lin_comb_lmo}
 	Let $\norm{\cdot}_{(1)}, \dots, \norm{\cdot}_{(n)}$ be norms on a finite-dimensional Euclidean space, and let $\alpha_1, \dots, \alpha_n$ be non-negative reals.
-Consider Linear Minimization Oracles $\mathrm{LMO}_{1}, \dots, \mathrm{LMO}_{n}$, corresponding to these norms.
-Then, $\sum_{i = 1}^n \alpha_i \mathrm{LMO}_{i}$ is the LMO corresponding to the norm $\norm{\cdot}^\dagger$ dual to the $\sum_{i = 1}^n \alpha_i \norm{\cdot}_{(i)}^\dagger$.
+Consider Linear Minimization Oracles $\mathrm{LMO}_{1}, \dots, \mathrm{LMO}_{n}$, corresponding to the unit balls of these norms.
+Then, $\sum_{i = 1}^n \alpha_i \mathrm{LMO}_{i}$ is the LMO corresponding to the norm $\norm{\cdot}$ dual to the $\sum_{i = 1}^n \alpha_i \norm{\cdot}_{(i)}^\dagger$.
 \end{lemma}
@@
 \begin{proof}
-	From \cref{lemma:dual_to_conv_comb} we know the general form of the unit ball in the $\norm{\cdot}$ norm: $\cB_{\norm{\cdot}} = \sum_{i = 1}^n \alpha_i \cB_{\norm{\cdot}_{(i)}^\dagger}$.
+	Using \cref{lemma:dual_to_conv_comb} and the fact $\norm{\cdot}^{\dagger \dagger} = \norm{\cdot}$, we can obtain general form of the unit ball in the $\norm{\cdot}$ norm: $\cB_{\norm{\cdot}} = \sum_{i = 1}^n \alpha_i \cB_{\norm{\cdot}_{(i)}}$.
 Thus, the linear function minimization objective on a $\norm{\cdot}$ norm ball can be transformed as follows:
 	$$
-	\argmin_{\mD \in \cB_{\norm{\cdot}}}\<\mM, \mD> = \argmin_{\mD_1 \in \alpha_i \cB_i,\dots,\mD_n \in \alpha_n\cB_n}\<\mM, \sum_{i = 1}^n\mD_i> = \sum_{i = 1}^n \argmin_{D_i \in \cB_i}\<\mM, \mD_i>\,.
+	\argmin_{\mD \in \cB_{\norm{\cdot}}}\<\mM, \mD> = \argmin_{\mD_1 \in \alpha_i \cB_{\norm{\cdot}_{(1)}},\dots,\mD_n \in \alpha_n\cB_{\norm{\cdot}_{(n)}}}\<\mM, \sum_{i = 1}^n\mD_i> = \sum_{i = 1}^n \argmin_{D_i \in \cB_{\norm{\cdot}_{(i)}}}\<\mM, \mD_i>\,,
 	$$
+	where the last summation denotes the Minkowski sum.
 This immediately implies $\sum_{i = 1}^n \alpha_i \mathrm{LMO}_{i} \in \argmin_{\mD \in \cB_{\norm{\cdot}}}\<\mM, \mD>$, which proves the claim of the lemma.
@@
 
+Applying it to optimization algorithms we obtain the following corollary.
+
+\begin{corollary}\label{corollary:lin_comb_alg}
+	Let there be a finite family of LMO based algorithms indexed by $i = 1,\dots,n$, where the update of the $i$-th algorithm is defined by 
+	$$
+	\mX^{k + 1} - \mX^{k} = \gamma_k \mathrm{LMO}_{i}(\mM^{k})\,,
+	$$
+	and $\mathrm{LMO}_i$ corresponds to the unit ball of norm $\norm{\cdot}_i$. For arbitrary non-negative $\alpha_1,\dots,\alpha_n$, the algorithm with the update given by 
+	$$
+	\mX^{k + 1} - \mX^{k} = \gamma_k \sum_{i = 1}^n \alpha_k \mathrm{LMO}_{\norm{\cdot}_{(i)}}(\mM^k)
+	$$
+	is an LMO-algorithm itself, with LMO corresponding to the unit ball of the norm $\norm{\cdot}$ dual to the norm given by $\sum_{i = 1}^n \alpha_i \norm{\cdot}_{(i)}^\dagger$.
+\end{corollary}
+
 \subsection{Frobeniusize Them: F-Muon and F-Neon}
 Let us construct the concrete examples of algorithms given by linear combinations of LMO-algorithms.
-It follows from \cref{lemma:lin_colm_alg} that those algorithms can also be viewed as LMO-algorithms.
+It follows from \cref{corollary:lin_comb_alg} that those algorithms can also be viewed as LMO-algorithms.
 
-Consider a mixture of Fanions with another excellent LMO-algorithm, NSGD. We define $\normfstar{\cdot}$ as a convex combination of $\normn{\cdot}$ and $\normf{\cdot}$:
+Combining Fanions-$k$ with another excellent LMO-algorithm, NSGD, we obtain the family of algorithms with updates defined by
+\begin{equation}\label{eq:nsgd_muon_update}
+		\mX^{t+1} = \mX^{t} - \gamma \left(\alpha \sum_{i = 1}^k u_i v_i^\top + (1-\alpha)\frac{\mM^k}{\normf{\mM^k}}\right)\,,
+\end{equation}
+where $u_1, \dots, u_k$ ($v_1, \dots, v_k$) are first $k$ left (right) sighular vectors of $\mM^k$.
+By \cref{corollary:lin_comb_alg}, this is an LMO-algorithm corresponding to the unit ball of the norm $\norm{\cdot}_{\mathrm{F-KF-k}}$, given by
+\begin{equation}\label{eq:fstar_norm}
+	\norm{\cdot}_{\mathrm{F-KF-k}}^\dagger = \alpha \normkfk{\cdot}^\dagger + (1-\alpha)\normf{\cdot}^\dagger\,.
+\end{equation}
+We name the derived algorithms \emph{F-Fanions}.
+A bit of information and visualizations related to the $\normfstar{\cdot}$ norm is presented in the appendix (see \cref{eq:normfstardual}, \cref{fig:fstardual}).
 
-\begin{equation}\label{eq:fstar_norm}
-	\normfstar{\mX} = \alpha \normn{\mX} + (1-\alpha)\normf{\mX},
-\end{equation}
-
-where $\alpha \in [0, 1]$ defines a specific norm from the $F*$-family.
-$\normfstar{\cdot}^\dagger$ can be expressed as in \cref{eq:normfstardual}.
-However, its norm ball is a simple Minkowski sum of $\alpha \normn{\cdot}$ and $(1-\alpha)\normf{\cdot}$ balls \cref{fig:fstardual}.
-
-\begin{lemma}\label{lemma:nsgd_muon_update}
-	When $\norm{\cdot} = \normfstar{\cdot}^\dagger$, \cref{eq:our_update} turns into:
-
-	\begin{equation}\label{eq:nsgd_muon_update}
-		\mX^{k+1} = \mX^{k} - \eta (\alpha U V^\top + (1-\alpha)\frac{\mM^k}{\normf{\mM^k}})
-	\end{equation}
-\end{lemma}
-
-We name the derived algorithm \emph{F-Muon}.
-
-We define $\normftwo{\cdot}$ as a convex combination of $\norms{\cdot}$ and $\normf{\cdot}$:
-
-\begin{equation}\label{eq:ftwo_norm}
-	\normftwo{\mX} = \alpha \norms{\mX} + (1-\alpha)\normf{\mX},
-\end{equation}
-
-where $\alpha \in [0, 1]$ defines a specific norm from the F2-family.$\normftwo{\cdot}^\dagger$ can be expressed as in \cref{eq:normftwodual}.
-However, its norm ball is a simple Minkowski sum of $\alpha \norms{\cdot}$ and $(1-\alpha)\normf{\cdot}$ balls \cref{fig:ftwodual}.
-
-\begin{lemma}\label{lemma:nsgd_neon_update}
-	When $\norm{\cdot} = \normftwo{\cdot}^\dagger$, \cref{eq:our_update} turns into:
-
-	\begin{equation}\label{eq:nsgd_neon_update}
-		\mX^{k+1} = \mX^{k} - \eta (\alpha u_1 v_1^\top + (1-\alpha)\frac{\mM^k}{\normf{\mM^k}})
-	\end{equation}
-\end{lemma}
-
-We name the derived algorithm \emph{F-Neon}. F-Neon is a convex combination of Normalized SGD and Neon.
-
-In general, one can consider F-KF-k-norm: $\norm{\cdot}_{\mathrm{F-KF-k}}$ = $\alpha \normkfk{\cdot} + (1-\alpha)\normf{\cdot}$, the dual to which will produce algorithms like F-Dion without the error feedback.
-The resulting family can be named {\it F-Fanions}.
+Two edge members of this family, with $k = 1$ and $k = \min\{m, n\}$ correspondingly, are \emph{F-Neon} and \emph{F-Muon}.
 

### 2025-11-25 conf_article/neon.tex
Almost wrote Related Work

--- 2025-11-25 conf_article/neon.tex
@@
 	In this article, we explore the use of various matrix norms for optimizing functions of weight matrices, a crucial problem in training large language models.
-Moving beyond the spectral norm that underlies the Muon update, we leverage the duals to the Ky Fan k-norms to introduce a family of Muon-like algorithms we name {\it Fanions}, which happen to be similar to Dion.
-Subsequently, we construct a second family of {\it F-Fanions}, which are based on the duals of convex combinations of the Ky Fan k-norms and the Frobenius norm.
-One prominent member of this family is {\it F-Muon}.
-We complement our theoretical analysis with an extensive empirical study of the algorithms across a wide range of tasks and settings, from which it follows that F-Muon is on par with Muon, which questions the exclusivity of a spectral norm.
+Moving beyond the spectral norm that underlies the Muon update, we leverage the duals to the Ky Fan $k$-norms to introduce a family of Muon-like algorithms we name {\it Fanions}, which happen to be similar to Dion.
+Then working with the duals of convex combinations of the Ky Fan $k$-norms and the Frobenius norm or the $l_\infty$ norm, we construct the families of {\it F-Fanions} and {\it S-Fanions} respectively.
+Their most prominent members are {\it F-Muon} and {\it S-Muon}.
+We complement our theoretical analysis with an extensive empirical study of the algorithms across a wide range of tasks and settings, from which it follows that F-Muon and S-Muon are always on par with Muon and S-Muo, while on fine-tuning of NanoGPT and synthetic linear least squares even better than vanilla Muon.
 
@@
 
-The per-restart complexity is $\mathcal{O}(mn^2 + n^2k + nk^2)$, where $m \geq n$ are the dimensions of the target matrix and $k$ is the size of the retained subspace (typically $k \ll n$).
+The per-cycle complexity is $\mathcal{O}(mn^2 + n^2k + nk^2)$, where $m \geq n$ are the dimensions of the target matrix and $k$ is the size of the retained subspace (typically $k \ll n$).
 
@@
 \section{Related Work and Discussion}
-As Muon \citep{jordan2024muon} is a very successful and popular optimizer for functions of weight matrices, a lot of reseach has been put into, first, further improving its performance, and, second, in explaining its success.
+Since Muon \citep{jordan2024muon} is a very successful and popular optimizer for functions of weight matrices, a lot of reseach has been put into, first, further improving its performance and, second, explaining its success.
 
-	{\bf Improvements of Muon.} Regarding the first point, in less than a year, a large number of applications and improvements of Muon has been proposed.
+	{\bf Improvements of Muon.} Regarding the first point, a large number of applications and improvements of Muon has been proposed in less than a year.
 \citet{liu2025muon} adapted the algorthm for training language models larger than NanoGPT.
 \citet{shah2025practical} organized efficient hyperparameter transfer by combining Muon with maximal update parametrization.
-To construct their COSMOS optimizer, \citet{chen2025cosmoshybridadaptive} applied computationally intensive updates of SOAP optimizer to a low-dimensional "leading eigensubspace" while using memory-efficient methods like Muon for the remaining parameters.
-\citet{amsel2025polar} proposed a more efficient alternative to Newton-Schulz operations.
+To construct their COSMOS optimizer, \citet{chen2025cosmoshybridadaptive} applied computationally intensive updates of SOAP optimizer \citep{vyas2025soap} to a low-dimensional "leading eigensubspace" while using memory-efficient methods like Muon for the remaining parameters.
+\citet{amsel2025polar} proposed a more efficient alternative to Newton-Schulz algorithm.
 \citet{si2025adamuon} introduced AdaMuon which combines element-wise adaptivity with orthogonal updates.
 We suppose that the described above or similar techniques can be applied to our optimizers as well.
-For example, F-Muon also benefits from faster alternatives to Newton-Schulz iterations, and Neon may be a great substitute to Muon in COSMOS, because, as we have shown in {\it the Matrix side of the updates}, Lanczos algorithms is much faster than Newton-Schulz iterations on large matrices.
+For example, F-Muon and S-Muon also benefit from faster alternatives to Newton-Schulz iterations, and Fanions may be a great substitute to Muon in COSMOS, because, as we have shown in Section~\ref{sec:matrix-side}, Lanczos algorithm is much faster than Newton-Schulz iterations on very large matrices.
 
@@
 This gap, as it seems to us, is not even now completely closed.
-For example, \citet{kovalev2025understanding} has provided convergence guarantees of Muon in various settings, from which, however, Muon's supremacy cannot be recovered.
-Indeed, although the obtained bounds depend on the norm choice, the asymptotics of the convergence remain the same as for NSGD and other optimizers, $K = \cO(\epsilon^{-4})$ in a L-smooth stochastic case.
+For example, \citet{kovalev2025understanding} has provided convergence guarantees of Muon and in various settings, from which, however, Muon's supremacy cannot be recovered.
+Indeed, although the obtained bounds depend on the norm choice, the asymptotics of the convergence remain the same as for NSGD and other optimizers, $K = \cO(\epsilon^{-4})$ in an L-smooth stochastic case.
 
-Similar drawback has a recent article \citet{riabinin2025gluon}, where L-smoothness assumption is replaced with a more practical $(L_0, L_1)$-smoothness.
-The authors derived from their theorems optimal stepsizes for Muon and Scion that match fine-tuned stepsizes reported by \citet{pethick2025training}.
-But still, they did not explain why, for instance, NSGD is inferior to Muon in training large-language models.
+	A similar drawback has the article \citet{riabinin2025gluon}, where L-smoothness assumption is replaced with a more practical $(L_0, L_1)$-smoothness.
+By estimating smoothness and substituting it into their Theorem 1, the authors recovered the optimal fine-tuned stepsizes reported by \citet{pethick2025training}.
+However, they have not showed the optimality of the spectral or the RMS-to-RMS norm, which is observed in practice, as our comparison with NSGD highlights.
 
-We suppose that the reason for the recorded by us discrepancy between Neon and Muon performance, both of which are described by Scion or Gluon frameworks, lies in the structure of the norm ball, which must be an object of further research.
+	Common important drawback of the analyses is the consideration of the convergence by the norm of the gradient.
+As we showed in our experiments on CIFAR, the gradient norm may decrease by a factor of ten, while the accuracy reaches 100\%.
+
+	We suppose that the reason for the recorded by us stark discrepancy between Neon and Muon performance, both of which are described by Stochastic Conditional Gradient or Gluon frameworks, lies in the structure of the norm ball or in the preconditioner interpretation \citep{pethick2025trainingneuralnetworksscale}, which must be an object of further research.
 
@@
 In the context of federated learning, error feedback is effective even for compressed Muon \citep{gruntkowska2025efmuon}.
-The advantage of Fanions and S-Fanions is that less bits are required to transmit the updates: $\sum_{i=1}^k u_i v_i^\top$ is easily transmitted as $\{u_1, v_1, \dots, u_k, v_k\}$ ($(m+n)\times k$ floats), while the sign part is as usual coded in $m \times n$ bits.
+The advantage of Fanions and S-Fanions is that less bits are required for transmission: $\sum_{i=1}^k u_i v_i^\top$ is easily transmitted as $\{u_1, v_1, \dots, u_k, v_k\}$ ($(m+n)\times k$ floats), while the sign part is as usual coded in $m \times n$ bits.
 Thus, the compression is "built-in".
@@
 \citet{pethick2025sam} applied it to create $\nu$SAM, a new sharpness-aware minimization technique.
-It would be interesting to substitute $\normn{\cdot}$ with $\normkfk{\cdot}^\dagger$ in their approach: $k > 1$ might increase accuracy boost, while preserving a small memory footprint and a small time overhead, if Dion-style power iterations are used.
+It would be interesting to substitute $\normn{\cdot}$ with $\normkfk{\cdot}^\dagger$ in their approach.
+Since the SAM-neighborhood becomes more diverse, $k > 1$ might add to the accuracy boost, while preserving a small memory footprint and a small time overhead, if Dion-style power iterations are used.
 
-	{\bf Orthogonal research} Fanions, F-Fanions, and S-Fanions benefit from the general theoretical description of \citet{riabinin2025gluon}.
+	{\bf Orthogonal research.} Fanions, F-Fanions, and S-Fanions benefit from the general theoretical description of \citet{riabinin2025gluon}, and better learning rates can be predicted by calculating the trajectory smoothness.
 They could be transformed to Drop-Fanions by updating only some layers as in \citet{gruntkowska2025dropmuon}.
@@
 One can clip them to produce ClippedScion-like algorithms \citep{pethick2025generalizedgradientnormclipping}.
-Finally, the results from \citet{shulgin2025inexactmuon} can be used to theoretically derive faster yet working approximate schemes to calculate the LMO; power iterations with a very limited number of iterations is a good thing to start with.
+They could be made more memory-effective by the zero-order techniques that worked for Muon \citep{petrov2025zeromuon}.
+Finally, the results from \citet{shulgin2025inexactmuon} can be used to theoretically derive faster yet working approximate schemes to calculate the LMO; power iterations with a very limited number of iterations is a good aglorithm to consider for the analysis.
 

### 2025-11-26 code/L_smooth_tests/finetune_lrs.py
Logged all conv grad norms

@@ -254,7 +254,7 @@ def main() -> None:
     parser.add_argument("--momentum-start", type=float, default=None, help="Start of momentum range (default: None, no momentum grid search)")
     parser.add_argument("--momentum-end", type=float, default=None, help="End of momentum range (default: None)")
     parser.add_argument("--momentum-num", type=int, default=5, help="Number of momentum points (default: 5)")
-    parser.add_argument("--loss-threshold", type=float, default=0.01, help="Loss threshold for below_thresh_iter (default: 0.01)")
+    parser.add_argument("--loss-threshold", type=float, default=0.001, help="Loss threshold for below_thresh_iter (default: 0.01)") # CHANGED
     parser.add_argument("--log-csv", type=str, default="L_smooth_tests/results/lr_grid_search.csv", help="Path to save grid search results CSV")
     parser.add_argument("--no-plot", action="store_true", help="If set, skip plotting after running experiments")
     parser.add_argument("--title-suffix", type=str, default=None, help="Optional title suffix for saved plots (defaults to size m x n)")
@@ -278,13 +278,15 @@ def main() -> None:
         np.round(np.linspace(0.01, 0.1, 19), 3),
         np.round(np.linspace(0.15, 1, 18), 3)
     ))
-    # learning_rates = np.round(np.linspace(0.15, 1, 18), 3)
+    # learning_rates = np.round(np.linspace(0.01, 0.1, 19), 3) # - for 0.01 loss
+    # learning_rates = learning_rates[learning_rates < 0.06]
+    learning_rates = np.round(np.linspace(0.005, 0.01, 6), 3) # - for 0.001 loss
     print(f"Learning rate grid: {learning_rates}")
     
     # Generate momentum grid if specified
     momentums = None
     if True:# args.momentum_start is not None and args.momentum_end is not None:
-        momentums = [0.9, 0.95]# [0, 0.5, 0.9, 0.95, 0.99]# np.linspace(args.momentum_start, args.momentum_end, args.momentum_num)
+        momentums = [0.5, 0.8, 0.9] # [0.1, 0.5, 0.8, 0.9, 0.95]# [0, 0.5, 0.9, 0.95, 0.99]# np.linspace(args.momentum_start, args.momentum_end, args.momentum_num)
         momentums = np.round(momentums, 3)
         print(f"Momentum grid: {momentums}")
         print(f"Loss threshold: {args.loss_threshold}")
[…]

### 2025-11-26 code/L_smooth_tests/run_benchmarks_new.py

@@ -251,7 +251,7 @@ def main() -> None:
         # ),
         "S-Muon": dict(
             optimizer_class=SignSGDMuon,
-            optimizer_kwargs=dict(lr=0.035, momentum=0.9, nesterov=True, sign_lr_mult=0.01, sgd_coeff=0.5),
+            optimizer_kwargs=dict(lr=0.01, momentum=0.9, nesterov=True, sign_lr_mult=0.01, sgd_coeff=0.5), # lr=0.035, momentum=0.8 for 0.01 loss
             num_iterations=iter_num_fro,
             record_interval=record_period,
             verbose=True,
@@ -269,7 +269,7 @@ def main() -> None:
         ),
         "Muon": dict(
             optimizer_class=NormalizedMuon,
-            optimizer_kwargs=dict(lr=0.02, momentum=0.9, nesterov=True),
+            optimizer_kwargs=dict(lr=0.007, momentum=0.5, nesterov=True), # 0.025, 0.5 for 0.01 loss
             num_iterations=iter_num_fro,
             record_interval=record_period,
             verbose=True,
@@ -278,7 +278,7 @@ def main() -> None:
         ),
         "F-Muon": dict(
             optimizer_class=NormalizedMuon,
-            optimizer_kwargs=dict(lr=0.04, momentum=0.6, nesterov=True, sgd_coeff=0.5),
+            optimizer_kwargs=dict(lr=0.014, momentum=0.5, nesterov=True, sgd_coeff=0.5), # lr=0.035, momentum=0.5 for 0.01 loss
             num_iterations=iter_num_fro,
             record_interval=record_period,
             verbose=True,

### 2025-11-26 code/cifar_bench/norm_logger_airbench.py

@@ -721,7 +721,7 @@ def run_optimizer_experiments(num_runs=1, warmup=True):
 
     # Define optimizer configurations with names
     optimizer_configs = {
-        'Muon_lr024': lambda filter_params: Muon(
+        'Muon': lambda filter_params: Muon(
             filter_params, lr=0.24, momentum=0.6, nesterov=True, norm_weight=False
         ),
         # 'Neon_kyfan_lr04': lambda filter_params: Neon(
@@ -743,11 +743,11 @@ def run_optimizer_experiments(num_runs=1, warmup=True):
         #     filter_params, neon_mode='kyfan', lr=0.45, k=10, momentum=0.65, nesterov=True, sgd_coeff=0.5
         # ),
         
-        'FMuon_lr04': lambda filter_params: NormalizedMuon(
+        'F-Muon_lr04': lambda filter_params: NormalizedMuon(
             filter_params, lr=0.4, momentum=0.65, sgd_coeff=0.5, nesterov=True, norm_weight=False
         ),
-        'SignSGDMuon_lr04': lambda filter_params: SignSGDMuon(
-            filter_params, lr=0.4, momentum=0.65, nesterov=True, sgd_coeff=0.5, sign_lr_mult=0.002, norm_weight=False 
+        'S-Muon': lambda filter_params: SignSGDMuon(
+            filter_params, lr=0.42, momentum=0.65, nesterov=True, sgd_coeff=0.5, sign_lr_mult=0.003, norm_weight=False 
         ),
         'NSGD_lr_1': lambda filter_params: NormalizedMuon(filter_params, lr=1, momentum=0.6, sgd_coeff=1, nesterov=True, norm_weight=False)
     }

### 2025-11-26 code/cifar_norms/cifar_norm_plotter.py

@@ -186,7 +186,8 @@ import os
 import re
 from datetime import datetime
 
-def collect_grad_norm_files(folder_path, after_dt=datetime(2025, 11, 25, 13, 50, 0)):
+
+def collect_grad_norm_files(folder_path, after_dt=datetime(2025, 11, 26, 10, 50, 0)):
     """
     Collect grad_norm CSV logs whose timestamp is strictly after `after_dt`.
 
@@ -237,8 +238,13 @@ if __name__ == "__main__":
         "total_spectral",
         "total_nuclear",
         "nuclear norm for conv1",  # Will match layers.*.conv1.weight_nuclear
-        "layers.1.conv1.weight_frobenius",
-        "layers.2.conv2.weight_spectral",
+        "nuclear norm for conv2",
+        "Frobenius norm for conv1",
+        "Frobenius norm for conv2",
+
+        # "Frobenius norm for norm2",
+        "spectral norm for conv1",
+        "spectral norm for conv2",
         "train_acc",  # Train accuracy (plotted separately)
         "val_acc"     # Validation accuracy (plotted separately)
     ]

### 2025-11-26 conf_article/icomp2025camera_ready/full-paper-template.tex
Improved intro and exps

%\documentclass[wcp,gray]{jmlr} % test grayscale version
 %\documentclass[wcp]{jmlr}% former name JMLR W\&CP
\documentclass[pmlr]{jmlr}% new name PMLR (Proceedings of Machine Learning)

\RequirePackage{graphicx}
 % The following packages will be automatically loaded:
 % amsmath, amssymb, natbib, graphicx, url, algorithm2e
 \usepackage{booktabs}
 %\usepackage{rotating}% for sideways figures and tables
\usepackage{longtable}% for long tables
 % The booktabs package is used by this sample document
 % (it provides \toprule, \midrule and \bottomrule).
 % 
 % book quality tables

 % The siunitx package is used by this sample document
 % to align numbers in a column by their decimal point.
 % Remove the next line if you don't require it.
\makeatletter
\def\set@curr@file#1{\def\@curr@file{#1}} %temp workaround for 2019 latex release
\makeatother
\usepackage[load-configurations=version-1]{siunitx} % newer version

 % The following command is just for this sample document:
\newcommand{\cs}[1]{\texttt{\char`\\#1}}

 % Define an unnumbered theorem just for this sample document:
\theorembodyfont{\upshape}
\theoremheaderfont{\scshape}
\theorempostheader{:}
\theoremsep{\newline}
\newtheorem*{note}{Note}

 % change the arguments, as appropriate, in the following:
\jmlrvolume{tbd}
\jmlryear{2025}
\jmlrworkshop{International Conference on Computational Optimization}

% H: Not sure if we need this:
% Short headings should be running head and authors last names
% \ShortHeadings{A Really Awesome MLHC Article}{Lastname, PhD and Lastname, MD}
% \firstpageno{1}

\title[Short Title]{Title of Your ICOMP Article}

\author{\Name{Firstname Lastname}
       \Email{name@email.edu}\\ 
       \addr Department\\
       University\\
       City, State, Country 
       \AND
       \Name{Firstname Lastname}
       \Email{name@email.edu}\\ 
       \addr Department\\
       University\\
       City, State, Country}

\begin{document}

\maketitle

\begin{abstract}
  Summary of the article.  Be sure to highlight how the work
  contributes to our understanding of machine learning and healthcare.
\end{abstract}

\section{Introduction}
\label{sec:intro}

This is a sample article that uses the \textsf{jmlr} class with
the \texttt{pmlr} class option.  Please follow the guidelines in
this sample document as it can help to reduce complications when
combining the articles into a book. Please avoid using obsolete
commands, such as \verb|\rm|, and obsolete packages, such as
\textsf{epsfig}.\footnote{See
\url{http://www.ctan.org/pkg/l2tabu}}

Please also ensure that your document will compile with PDF\LaTeX.
If you have an error message that's puzzling you, first check for it
at the UK TUG FAQ
\url{https://texfaq.org/FAQ-man-latex}.  If
that doesn't help, create a minimal working example (see
\url{https://www.dickimaw-books.com/latex/minexample}) and post
to somewhere like TeX on StackExchange
(\url{https://tex.stackexchange.com/}) or the LaTeX Community Forum
(\url{https://latex.org/forum/}).

\begin{note}
This is an numbered theorem-like environment that was defined in
this document's preamble.
\end{note}

\subsection{Sub-sections}

Sub-sections are produced using \verb|\subsection|.

\subsubsection{Sub-sub-sections}

Sub-sub-sections are produced using \verb|\subsubsection|.

\paragraph{Sub-sub-sub-sections}

Sub-sub-sub-sections are produced using \verb|\paragraph|.
These are unnumbered with a running head.

\subparagraph{Sub-sub-sub-sub-sections}

Sub-sub-sub-sub-sections are produced using \verb|\subparagraph|.
These are unnumbered with a running head.

\section{Cross-Referencing}

Always use \verb|\label| and \verb|\ref| (or one of the commands
described below) when cross-referencing.  For example, the next
section is Section~\ref{sec:math}. The \textsf{jmlr} class
provides some convenient cross-referencing commands:
\verb|\sectionref|, \verb|\equationref|, \verb|\tableref|,
\verb|\figureref|, \verb|\algorithmref|, \verb|\theoremref|,
\verb|\lemmaref|, \verb|\remarkref|, \verb|\corollaryref|,
\verb|\definitionref|, \verb|\conjectureref|, \verb|\axiomref|,
\verb|\exampleref| and \verb|\appendixref|. The argument of these
commands may either be a single label or a comma-separated list
of labels. Examples:

Referencing sections: \sectionref{sec:math} or
\sectionref{sec:intro,sec:math} or
\sectionref{sec:intro,sec:math,sec:tables,sec:figures}.

Referencing equations: \equationref{eq:trigrule} or
\equationref{eq:trigrule,eq:df} or
\equationref{eq:trigrule,eq:f,eq:df,eq:y}.

Referencing tables: \tableref{tab:operatornames} or
\tableref{tab:operatornames,tab:example} or
\tableref{tab:operatornames,tab:example,tab:example-booktabs}.

Referencing figures: \figureref{fig:image} or
\figureref{fig:image,fig:teximage} or
\figureref{fig:image,fig:teximage,fig:subfigex} or
\figureref{fig:image-a,fig:image-b}.

Referencing algorithms: \algorithmref{alg:gauss} or
\algorithmref{alg:gauss,alg:moore} or
\algorithmref{alg:gauss,alg:moore,alg:net}.

Referencing theorem-like environments: \theoremref{thm:eigenpow},
\lemmaref{lem:sample}, \remarkref{rem:sample}, 
\corollaryref{cor:sample}, \definitionref{def:sample},
\conjectureref{con:sample}, \axiomref{ax:sample} and
\exampleref{ex:sample}.

Referencing appendices: \appendixref{apd:first} or
\appendixref{apd:first,apd:second}.

\section{Equations}
\label{sec:math}

The \textsf{jmlr} class loads the \textsf{amsmath} package, so
you can use any of the commands and environments defined there.
(See the \textsf{amsmath} documentation for further
details.\footnote{Either \texttt{texdoc amsmath} or
\url{http://www.ctan.org/pkg/amsmath}})

Unnumbered single-lined equations should be displayed using
\verb|\[| and \verb|\]|. For example:
\[E = m c^2\]
Numbered single-line equations should be displayed using the
\texttt{equation} environment. For example:
\begin{equation}\label{eq:trigrule}
\cos^2\theta + \sin^2\theta \equiv 1
\end{equation}
This can be referenced using \verb|\label| and \verb|\equationref|.
For example, \equationref{eq:trigrule}.

Multi-lined numbered equations should be displayed using the
\texttt{align} environment.\footnote{For reasons why you 
shouldn't use the obsolete \texttt{eqnarray} environment, see
Lars Madsen, \emph{Avoid eqnarray!} TUGboat 33(1):21--25, 2012.} For example:
\begin{align}
f(x) &= x^2 + x\label{eq:f}\\
f'(x) &= 2x + 1\label{eq:df}
\end{align}
Unnumbered multi-lined equations should be displayed using the
\texttt{align*} environment. For example:
\begin{align*}
f(x) &= (x+1)(x-1)\\
&= x^2 - 1
\end{align*}
If you want to mix numbered with unnumbered lines use the
align environment and suppress unwanted line numbers with
\verb|\nonumber|. For example:
\begin{align}
y &= x^2 + 3x - 2x + 1\nonumber\\
&= x^2 + x + 1\label{eq:y}
\end{align}
An equation that is too long to fit on a single line can be
displayed using the \texttt{split} environment. 
Text can be embedded in an equation using \verb|\text| or
\verb|\intertext| (as used in \theoremref{thm:eigenpow}).
See the \textsf{amsmath} documentation for further details.

\subsection{Operator Names}
\label{sec:op}

Predefined operator names are listed in \tableref{tab:operatornames}.
For additional operators, either use \verb|\operatorname|,
for example $\operatorname{var}(X)$ or declare it with
\verb|\DeclareMathOperator|, for example
\begin{verbatim}
\DeclareMathOperator{\var}{var}
\end{verbatim}
and then use this new command. If you want limits that go above and
below the operator (like \verb|\sum|) use the starred versions
(\verb|\operatorname*| or \verb|\DeclareMathOperator*|).

\begin{table}[htbp]
\floatconts
  {tab:operatornames}%
  {\caption{Predefined Operator Names (taken from 
   \textsf{amsmath} documentation)}}%
  {%
\begin{tabular}{rlrlrlrl}
\cs{arccos} & $\arccos$ &  \cs{deg} & $\deg$ &  \cs{lg} & $\lg$ &  \cs{projlim} & $\projlim$ \\
\cs{arcsin} & $\arcsin$ &  \cs{det} & $\det$ &  \cs{lim} & $\lim$ &  \cs{sec} & $\sec$ \\
\cs{arctan} & $\arctan$ &  \cs{dim} & $\dim$ &  \cs{liminf} & $\liminf$ &  \cs{sin} & $\sin$ \\
\cs{arg} & $\arg$ &  \cs{exp} & $\exp$ &  \cs{limsup} & $\limsup$ &  \cs{sinh} & $\sinh$ \\
\cs{cos} & $\cos$ &  \cs{gcd} & $\gcd$ &  \cs{ln} & $\ln$ &  \cs{sup} & $\sup$ \\
\cs{cosh} & $\cosh$ &  \cs{hom} & $\hom$ &  \cs{log} & $\log$ &  \cs{tan} & $\tan$ \\
\cs{cot} & $\cot$ &  \cs{inf} & $\inf$ &  \cs{max} & $\max$ &  \cs{tanh} & $\tanh$ \\
\cs{coth} & $\coth$ &  \cs{injlim} & $\injlim$ &  \cs{min} & $\min$ \\
\cs{csc} & $\csc$ &  \cs{ker} & $\ker$ &  \cs{Pr} & $\Pr$
\end{tabular}\par
\begin{tabular}{rlrl}
\cs{varlimsup} & $\varlimsup$ 
& \cs{varinjlim} & $\varinjlim$\\
\cs{varliminf} & $\varliminf$ 
& \cs{varprojlim} & $\varprojlim$
\end{tabular}
}
\end{table}

\section{Vectors and Sets}
\label{sec:vec}

Vectors should be typeset using \cs{vec}. For example $\vec{x}$.
The \textsf{jmlr} class also provides \cs{set} to typeset a
set. For example $\set{S}$.

\section{Floats}
\label{sec:floats}

Floats, such as figures, tables and algorithms, are moving
objects and are supposed to float to the nearest convenient
location. Please don't force them to go in a particular place. In
general it's best to use the \texttt{htbp} specifier and don't
put the figure or table in the middle of a paragraph (that is
make sure there's a paragraph break above and below the float).
Floats are supposed to have a little extra space above and below
them to make them stand out from the rest of the text. This extra
spacing is put in automatically and shouldn't need modifying.

To ensure consistency, please \emph{don't} try changing the format of the caption by doing
something like:
\begin{verbatim}
\caption{\textit{A Sample Caption.}}
\end{verbatim}
or
\begin{verbatim}
\caption{\em A Sample Caption.}
\end{verbatim}
You can, of course, change the font for individual words or 
phrases, for example:
\begin{verbatim}
\caption{A Sample Caption With Some \emph{Emphasized Words}.}
\end{verbatim}

\subsection{Tables}
\label{sec:tables}

Tables should go in the \texttt{table} environment. Within this
environment use \verb|\floatconts| (defined by \textsf{jmlr})
to set the caption correctly and center the table contents.

\begin{table}[htbp]
 % The first argument is the label.
 % The caption goes in the second argument, and the table contents
 % go in the third argument.
\floatconts
  {tab:example}%
  {\caption{An Example Table}}%
  {\begin{tabular}{ll}
  \bfseries Dataset & \bfseries Result\\
  Data1 & 0.12345\\
  Data2 & 0.67890\\
  Data3 & 0.54321\\
  Data4 & 0.09876
  \end{tabular}}
\end{table}

If you want horizontal rules you can use the \textsf{booktabs}
package which provides the commands \verb|\toprule|, 
\verb|\midrule| and \verb|\bottomrule|. For example, see
\tableref{tab:example-booktabs}.

\begin{table}[hbtp]
\floatconts
  {tab:example-booktabs}
  {\caption{A Table With Horizontal Lines}}
  {\begin{tabular}{ll}
  \toprule
  \bfseries Dataset & \bfseries Result\\
  \midrule
  Data1 & 0.12345\\
  Data2 & 0.67890\\
  Data3 & 0.54321\\
  Data4 & 0.09876\\
  \bottomrule
  \end{tabular}}
\end{table}

If you want vertical lines as well, you can't use the
\textsf{booktabs} commands as there'll be some unwanted gaps.
Instead you can use \LaTeX's \verb|\hline|, but the rows may
appear a bit cramped.  You can add extra space above or below a
row using \verb|\abovestrut| and \verb|\belowstrut|. For example,
see \tableref{tab:example-hline}.

\begin{table}[htbp]
\floatconts
  {tab:example-hline}
  {\caption{A Table With Horizontal and Vertical Lines}}%
  {%
    \begin{tabular}{|l|l|}
    \hline
    \abovestrut{2.2ex}\bfseries Dataset & \bfseries Result\\\hline
    \abovestrut{2.2ex}Data1 & 0.12345\\
    Data2 & 0.67890\\
    Data3 & 0.54321\\
    \belowstrut{0.2ex}Data4 & 0.09876\\\hline
    \end{tabular}
  }
\end{table}

If you want to align numbers on their decimal point, you can
use the \textsf{siunitx} package. For example, see
\tableref{tab:example-siunitx}. For further details see the
\textsf{siunitx} documentation\footnote{Either \texttt{texdoc
siunitx} or \url{http://www.ctan.org/pkg/siunitx}}.

\begin{table}[htbp]
\floatconts
  {tab:example-siunitx}
  {\caption{A Table With Numbers Aligned on the Decimal Point}}
  {\begin{tabular}{lS}
  \bfseries Dataset & {\bfseries Result}\\
  Data1 & 0.12345\\
  Data2 & 10.6789\\
  Data3 & 50.543\\
  Data4 & 200.09876
  \end{tabular}}
\end{table}

If the table is too wide, you can adjust the inter-column
spacing by changing the value of \verb|\tabcolsep|. For
example:
\begin{verbatim}
\setlength{\tabcolsep}{3pt}
\end{verbatim}
If the table is very wide but not very long, you can use the
\texttt{sidewaystable} environment defined in the
\textsf{rotating} package (so use \verb|\usepackage{rotating}|).
If the table is too long to fit on a page, you should use the
\texttt{longtable} environment defined in the \textsf{longtable}
package (so use \verb|\usepackage{longtable}|).

\subsection{Figures}
\label{sec:figures}

Figures should go in the \texttt{figure} environment. Within this
environment, use \verb|\floatconts| to correctly position the
caption and center the image. Use \verb|\includegraphics|
for external graphics files but omit the file extension. Do not
use \verb|\epsfig| or \verb|\psfig|. If you want to scale the
image, it's better to use a fraction of the line width rather
than an explicit length. For example, see \figureref{fig:image}.

\begin{figure}[htbp]
 % Caption and label go in the first argument and the figure contents
 % go in the second argument
\floatconts
  {fig:image}
  {\caption{Example Image}}
  {\includegraphics[width=0.5\linewidth]{example-image}}
\end{figure}

If your image is made up of \LaTeX\ code (for example, commands
provided by the \textsf{pgf} package) you can include it using
\cs{includeteximage} (defined by the \textsf{jmlr} class). This
can be scaled and rotated in the same way as \cs{includegraphics}.
For example, see \figureref{fig:teximage}.

\begin{figure}[htbp]
\floatconts
  {fig:teximage}
  {\caption{Image Created Using \LaTeX\ Code}}
  {\includeteximage[angle=45]{example-teximage}}
\end{figure}

If the figure is too wide to fit on the page, you can use the
\texttt{sidewaysfigure} environment defined in the
\textsf{rotating} package.

It's best not to use \verb|\graphicspath| with the \textsf{jmlr} class
as it can cause problems with the production editing process. If the
images are contained in a subdirectory, specify this when you
include the image, for example \verb|\includegraphics{figures/mypic}|.

\subsection{Sub-Figures}
\label{sec:subfigures}

Sub-figures can be created using \verb|\subfigure|, which is
defined by the \textsf{jmlr} class. The optional argument allows
you to provide a subcaption. The label should be placed in the
mandatory argument of \verb|\subfigure|. You can reference the
entire figure, for example \figureref{fig:subfigex}, or you can
reference part of the figure using \verb|\figureref|, for example
\figureref{fig:image-a}. Alternatively you can reference the
subfigure using \verb|\subfigref|, for example
\subfigref{fig:image-a,fig:image-b} in \figureref{fig:subfigex}.

\begin{figure}[htbp]
\floatconts
  {fig:subfigex}
  {\caption{An Example With Sub-Figures.}}
  {%
    \subfigure[Image A]{\label{fig:image-a}%
      \includegraphics[width=0.2\linewidth]{example-image-a}}%
    \qquad
    \subfigure[Image B]{\label{fig:image-b}%
      \includegraphics[width=0.25\linewidth]{example-image-b}}
  }
\end{figure}

By default, the sub-figures are aligned on the baseline.
This can be changed using the second optional argument
of \verb|\subfigure|. This may be \texttt{t} (top), \texttt{c}
(centered) or \texttt{b} (bottom). For example, the subfigures
\subfigref{fig:image-a2,fig:image-b2} in \figureref{fig:subfigex2}
both have \verb|[c]| as the second optional argument.

\begin{figure}[htbp]
\floatconts
  {fig:subfigex2}
  {\caption{Another Example With Sub-Figures (with a Cramped
    Sub-Caption).}}
  {%
    \subfigure[Image A][c]{\label{fig:image-a2}%
      \includegraphics[width=0.1\linewidth]{example-image-a}}%
    \qquad
    \subfigure[Image B][c]{\label{fig:image-b2}%
      \includegraphics[width=0.3\linewidth]{example-image-b}}
  }%
\end{figure}

Note that a very narrow sub-float will leave little space for the
sub-caption, which will likely cause Underfull/Overfull hbox warnings.

\subsection{Sub-Tables}
\label{sec:subtables}
There is an analogous command \verb|\subtable| for sub-tables.
It has the same syntax as \verb|\subfigure| described above.
You can reference the table using \verb|\tableref|, for example
\tableref{tab:subtabex} or you can reference part of the table,
for example \tableref{tab:ab}. Alternatively you can reference the
subtable using \verb|\subtabref|, for example
\subtabref{tab:ab,tab:cd} in \tableref{tab:subtabex}.

\begin{table}[htbp]
\floatconts
 {tab:subtabex}
 {\caption{An Example With Sub-Tables}}
 {%
   \subtable{%
     \label{tab:ab}%
     \begin{tabular}{cc}
     \bfseries A & \bfseries B\\
     1 & 2
     \end{tabular}
   }\qquad
   \subtable{%
     \label{tab:cd}%
     \begin{tabular}{cc}
     \bfseries C & \bfseries D\\
     3 & 4\\
     5 & 6
     \end{tabular}
   }
 }
\end{table}

By default, the sub-tables are aligned on the top.
This can be changed using the second optional argument
of \verb|\subtable|. This may be \texttt{t} (top), \texttt{c}
(centered) or \texttt{b} (bottom). For example, the sub-tables
\subtabref{tab:ab2,tab:cd2} in \tableref{tab:subtabex2}
both have \verb|[c]| as the second optional argument.

\begin{table}[htbp]
\floatconts
 {tab:subtabex2}
 {\caption{Another Example With Sub-Tables}}
 {%
   \subtable[][c]{%
     \label{tab:ab2}%
     \begin{tabular}{cc}
     \bfseries A & \bfseries B\\
     1 & 2
     \end{tabular}
   }\qquad
   \subtable[][c]{%
     \label{tab:cd2}%
     \begin{tabular}{cc}
     \bfseries C & \bfseries D\\
     3 & 4\\
     5 & 6
     \end{tabular}
   }
 }
\end{table}

\subsection{Algorithms}
\label{sec:algorithms}

Enumerated textual algorithms can be displayed using the
\texttt{algorithm} environment. Within this environment, use
use an \texttt{enumerate} or nested \texttt{enumerate} environments.
For example, see \algorithmref{alg:gauss}. Note that algorithms
float like figures and tables.

\begin{algorithm}[htbp]
\floatconts
{alg:gauss}% label
{\caption{The Gauss-Seidel Algorithm}}
{% contents
\begin{enumerate}
  \item For $k=1$ to maximum number of iterations
    \begin{enumerate}
      \item For $i=1$ to $n$
        \begin{enumerate}
        \item $x_i^{(k)} = 
          \frac{b_i - \sum_{j=1}^{i-1}a_{ij}x_j^{(k)}
          - \sum_{j=i+1}^{n}a_{ij}x_j^{(k-1)}}{a_{ii}}$
        \item If $\|\vec{x}^{(k)}-\vec{x}^{(k-1)} < \epsilon\|$,
          where $\epsilon$ is a specified stopping criteria, stop.
      \end{enumerate}
    \end{enumerate}
\end{enumerate}
}
\end{algorithm}

If you'd rather have the same numbering throughout the algorithm
but still want the convenient indentation of nested 
\texttt{enumerate} environments, you can use the
\texttt{enumerate*} environment provided by the \textsf{jmlr}
class. For example, see \algorithmref{alg:moore}.

\begin{algorithm}
\floatconts
{alg:moore}% label
{\caption{Moore's Shortest Path}}%caption
{% main float content
Given a connected graph $G$, where the length of each edge is 1:
\begin{enumerate*}
  \item Set the label of vertex $s$ to 0
  \item Set $i=0$
  \begin{enumerate*}
    \item \label{step:locate}Locate all unlabelled vertices 
          adjacent to a vertex labelled $i$ and label them $i+1$
    \item If vertex $t$ has been labelled,
    \begin{enumerate*}
      \item[] the shortest path can be found by backtracking, and 
      the length is given by the label of $t$.
    \end{enumerate*}
    otherwise
    \begin{enumerate*}
      \item[] increment $i$ and return to step~\ref{step:locate}
    \end{enumerate*}
  \end{enumerate*}
\end{enumerate*}
}
\end{algorithm}

Pseudo code can be displayed using the \texttt{algorithm2e}
environment. This is defined by the \textsf{algorithm2e} package
(which is automatically loaded) so check the \textsf{algorithm2e}
documentation for further details.\footnote{Either \texttt{texdoc
algorithm2e} or \url{http://www.ctan.org/pkg/algorithm2e}}
For an example, see \algorithmref{alg:net}.

\begin{algorithm2e}
\caption{Computing Net Activation}
\label{alg:net}
 % older versions of algorithm2e have \dontprintsemicolon instead
 % of the following:
 %\DontPrintSemicolon
 % older versions of algorithm2e have \linesnumbered instead of the
 % following:
 %\LinesNumbered
\KwIn{$x_1, \ldots, x_n, w_1, \ldots, w_n$}
\KwOut{$y$, the net activation}
$y\leftarrow 0$\;
\For{$i\leftarrow 1$ \KwTo $n$}{
  $y \leftarrow y + w_i*x_i$\;
}
\end{algorithm2e}

\section{Description Lists}

The \textsf{jmlr} class also provides a description-like 
environment called \texttt{altdescription}. This has an
argument that should be the widest label in the list. Compare:
\begin{description}
\item[add] A method that adds two variables.
\item[differentiate] A method that differentiates a function.
\end{description}
with
\begin{altdescription}{differentiate}
\item[add] A method that adds two variables.
\item[differentiate] A method that differentiates a function.
\end{altdescription}

\section{Theorems, Lemmas etc}
\label{sec:theorems}

The following theorem-like environments are predefined by
the \textsf{jmlr} class: \texttt{theorem}, \texttt{example},
\texttt{lemma}, \texttt{proposition}, \texttt{remark}, 
\texttt{corollary}, \texttt{definition}, \texttt{conjecture}
and \texttt{axiom}. You can use the \texttt{proof} environment
to display the proof if need be, as in \theoremref{thm:eigenpow}.

\begin{theorem}[Eigenvalue Powers]\label{thm:eigenpow}
If $\lambda$ is an eigenvalue of $\vec{B}$ with eigenvector
$\vec{\xi}$, then $\lambda^n$ is an eigenvalue of $\vec{B}^n$
with eigenvector $\vec{\xi}$.
\begin{proof}
Let $\lambda$ be an eigenvalue of $\vec{B}$ with eigenvector
$\xi$, then
\begin{align*}
\vec{B}\vec{\xi} &= \lambda\vec{\xi}
\intertext{premultiply by $\vec{B}$:}
\vec{B}\vec{B}\vec{\xi} &= \vec{B}\lambda\vec{\xi}\\
\Rightarrow \vec{B}^2\vec{\xi} &= \lambda\vec{B}\vec{\xi}\\
&= \lambda\lambda\vec{\xi}\qquad
\text{since }\vec{B}\vec{\xi}=\lambda\vec{\xi}\\
&= \lambda^2\vec{\xi}
\end{align*}
Therefore true for $n=2$. Now assume true for $n=k$:
\begin{align*}
\vec{B}^k\vec{\xi} &= \lambda^k\vec{\xi}
\intertext{premultiply by $\vec{B}$:}
\vec{B}\vec{B}^k\vec{\xi} &= \vec{B}\lambda^k\vec{\xi}\\
\Rightarrow \vec{B}^{k+1}\vec{\xi} &= \lambda^k\vec{B}\vec{\xi}\\
&= \lambda^k\lambda\vec{\xi}\qquad
\text{since }\vec{B}\vec{\xi}=\lambda\vec{\xi}\\
&= \lambda^{k+1}\vec{\xi}
\end{align*}
Therefore true for $n=k+1$. Therefore, by induction, true for all
$n$.
\end{proof}
\end{theorem}

\begin{lemma}[A Sample Lemma]\label{lem:sample}
This is a lemma.
\end{lemma}

\begin{remark}[A Sample Remark]\label{rem:sample}
This is a remark.
\end{remark}

\begin{corollary}[A Sample Corollary]\label{cor:sample}
This is a corollary.
\end{corollary}

\begin{definition}[A Sample Definition]\label{def:sample}
This is a definition.
\end{definition}

\begin{conjecture}[A Sample Conjecture]\label{con:sample}
This is a conjecture.
\end{conjecture}

\begin{axiom}[A Sample Axiom]\label{ax:sample}
This is an axiom.
\end{axiom}

\begin{example}[An Example]\label{ex:sample}
This is an example.
\end{example}

\section{Citations and Bibliography}
\label{sec:cite}

The \textsf{jmlr} class automatically loads \textsf{natbib}.
This sample file has the citations defined in the accompanying
BibTeX file \texttt{pmlr-sample.bib}. For a parenthetical
citation use \verb|\citep|. For example
\citep{guyon-elisseeff-03}. For a textual citation use
\verb|\citet|. For example \citet{guyon2007causalreport}.
Both commands may take a comma-separated list, for example
\citet{guyon-elisseeff-03,guyon2007causalreport}.

These commands have optional arguments and have a starred
version. See the \textsf{natbib} documentation for further
details.\footnote{Either \texttt{texdoc natbib} or
\url{http://www.ctan.org/pkg/natbib}}

The bibliography is displayed using \verb|\bibliography|.

\acks{Acknowledgements go here.}

\bibliography{sample}

\appendix

\section{First Appendix}\label{apd:first}

This is the first appendix.

\section{Second Appendix}\label{apd:second}

This is the second appendix.

\end{document}

\end{document}

### 2025-11-26 code/L_smooth_tests/finetune_lrs.py
Measured F-Muon, S-Muon, and Muon on LLS finally

@@ -2,6 +2,7 @@ from typing import Any, Dict, List, Optional, Tuple
 from math import sqrt
 import matplotlib.cm as cm
 import numpy as np
+from numpy.random.mtrand import randint
 import pandas as pd
 import os
 
@@ -280,13 +281,15 @@ def main() -> None:
     ))
     # learning_rates = np.round(np.linspace(0.01, 0.1, 19), 3) # - for 0.01 loss
     # learning_rates = learning_rates[learning_rates < 0.06]
-    learning_rates = np.round(np.linspace(0.005, 0.01, 6), 3) # - for 0.001 loss
+    learning_rates = np.round(np.logspace(np.log10(0.0005), np.logspace(0.1), 15), 3) # - for 0.001 loss
+    # learning_rates = np.round(np.linspace(0.005, 0.020, 16), 3) # - for 0.001 loss
+    
     print(f"Learning rate grid: {learning_rates}")
     
     # Generate momentum grid if specified
     momentums = None
     if True:# args.momentum_start is not None and args.momentum_end is not None:
-        momentums = [0.5, 0.8, 0.9] # [0.1, 0.5, 0.8, 0.9, 0.95]# [0, 0.5, 0.9, 0.95, 0.99]# np.linspace(args.momentum_start, args.momentum_end, args.momentum_num)
+        momentums = [0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 0.95] # [0.1, 0.5, 0.8, 0.9, 0.95]# [0, 0.5, 0.9, 0.95, 0.99]# np.linspace(args.momentum_start, args.momentum_end, args.momentum_num)
         momentums = np.round(momentums, 3)
         print(f"Momentum grid: {momentums}")
         print(f"Loss threshold: {args.loss_threshold}")
@@ -319,38 +322,38 @@ def main() -> None:
         #     num_iterations=iter_num_op,
         #     record_interval=100,
         # ),
-        dict(
-            name="Muon",
-            optimizer_class=NormalizedMuon,
-            optimizer_kwargs=dict(nesterov=True),
-            num_iterations=1500,
-            record_interval=rint,
-            use_momentum=momentums is not None,
-        ),
-        dict(
-            name="F-Muon",
-            optimizer_class=NormalizedMuon,
-            optimizer_kwargs=dict(nesterov=True, sgd_coeff=0.5),
-            num_iterations=1500,
-            record_interval=rint,
[…]

### 2025-11-26 code/L_smooth_tests/run_benchmarks_new.py

@@ -110,6 +110,15 @@ def main() -> None:
     def const_lr(T: int, base_lr: float) -> float:
         return base_lr
 
+
+    muon_lr = 0.007
+    fmuon_lr = 0.015
+    smuon_lr = 0.011
+    
+    muon_mom = 0.5
+    fmuon_mom = 0.7
+    smuon_mom = 0.9
+
     # Define experiment specifications
     colors = cm.tab20b(np.linspace(0, 1, 5))[::-1]
     experiment_specs: Dict[str, Dict[str, Any]] = {
@@ -150,86 +159,86 @@ def main() -> None:
         #     requires_cuda=True,
         #     color=colors[0]
         # ),
-        # "Neon": dict(
-        #     optimizer_class=Neon,
-        #     optimizer_kwargs=dict(lr=0.45, nesterov=True, momentum=0.95, neon_mode="kyfan", iter_num=5),
-        #     num_iterations=iter_num_nuc,
-        #     record_interval=50,
-        #     verbose=True,
-        #     lr_scheduler=const_lr,
-        #     requires_cuda=True,
-        #     color=colors[0]
-        # ),
-        # "F-Neon": dict(
-        #     optimizer_class=Neon,
-        #     optimizer_kwargs=dict(lr=0.25, nesterov=True, momentum=0.9, neon_mode="kyfan", iter_num=5, sgd_coeff=0.5),
-        #     num_iterations=iter_num_nuc,
-        #     record_interval=50,
-        #     verbose=True,
-        #     lr_scheduler=const_lr,
-        #     requires_cuda=True,
-        #     color=colors[0]
-        # ),
-        # "Fanion-2": dict(
-        #     optimizer_class=Neon,
-        #     optimizer_kwargs=dict(lr=1, nesterov=True, momentum=0.6, neon_mode="kyfan", iter_num=5, k=2),
-        #     num_iterations=iter_num_nuc,
-        #     record_interval=100,
-        #     verbose=True,
-        #     lr_scheduler=inv_sqrt_lr,
-        #     requires_cuda=True,
-        #     color=colors[1]
-        # ),
-        # "F-Fanion-2": dict(
-        #     optimizer_class=Neon,
-        #     optimizer_kwargs=dict(lr=1, nesterov=True, momentum=0.6, neon_mode="kyfan", iter_num=5, sgd_coeff=0.5, k = 2),
-        #     num_iterations=iter_num_nuc,
[…]

### 2025-11-26 code/cifar_bench/airbench_muon.py

@@ -427,8 +427,9 @@ def main(run, model):
                      dict(params=norm_biases,         lr=bias_lr, weight_decay=wd/bias_lr),
                      dict(params=[model.head.weight], lr=head_lr, weight_decay=wd/head_lr)]
     optimizer1 = torch.optim.SGD(param_configs, momentum=0.85, nesterov=True, fused=True)
-    # optimizer2 =  NormalizedMuon(filter_params, lr=0.4, momentum=0.65, nesterov=True, sgd_coeff=0.5)
-    optimizer2 =  SignSGDMuon(filter_params, lr=0.42, momentum=0.65, nesterov=True, sgd_coeff=0.5, sign_lr_mult=0.003)
+    optimizer2 =  Muon(filter_params, lr=0.24, momentum=0.63, nesterov=True)
+    # optimizer2 =  NormalizedMuon(filter_params, lr=0.44, momentum=0.63, nesterov=True, sgd_coeff=0.5)
+    # optimizer2 =  SignSGDMuon(filter_params, lr=0.42, momentum=0.63, nesterov=True, sgd_coeff=0.5, sign_lr_mult=0.003)
     # optimizer2 = MuonOrSign(filter_params, lr=0.42, momentum=0.65, nesterov=True, sign_coeff=0.003) # 94.00%, for now a bad idea
 
     optimizers = [optimizer1, optimizer2]
@@ -515,7 +516,7 @@ if __name__ == "__main__":
 
     print_columns(logging_columns_list, is_head=True)
     main("warmup", model)
-    accs = torch.tensor([main(run, model) for run in range(200)])
+    accs = torch.tensor([main(run, model) for run in range(50)])
     print("Mean: %.4f    Std: %.4f" % (accs.mean(), accs.std()))
 
     log_dir = os.path.join("logs", str(uuid.uuid4()))

### 2025-11-26 code/cifar_bench/base_airbench.py

@@ -19,7 +19,7 @@ import torch.nn.functional as F
 import torchvision
 import torchvision.transforms as T
 from optimizers import Dion, Muon, Neon, NormalizedMuon, SGDMuon, SignSGDMuon, zeropower_via_newtonschulz5, RandomNormalizedMuon, NuclearNormalizedMuon, MuonCringeMomentum
-from optimizers import SpectrallyNormalizedNeon, MuonOrNSGD, MuonOrSign, RealFanion, NeonMuon, SingleDeviceNorMuon
+from optimizers import SpectrallyNormalizedNeon, MuonOrNSGD, MuonOrSign, RealFanion, NeonMuon, SingleDeviceNorMuon, MuonSignedUpdate
 from mlion import MLion, NLion
 
 #############################################
@@ -418,11 +418,14 @@ def main(run, model):
     
 
     # optimizer2 = SignSGDMuon(filter_params, lr=0.4, momentum=0.6, nesterov=True, sgd_coeff=1, sign_lr_mult=0.005) # 90.94 +- 0.4%
+    optimizer2 = SignSGDMuon(filter_params, lr=0.003, momentum=0.6, nesterov=True, sgd_coeff=1, sign_lr_mult=1) # 90.94 +- 0.4%
     # optimizer2 = SignSGDMuon(filter_params, lr=0.4, momentum=0.6, nesterov=True, sgd_coeff=1, sign_lr_mult=0.007) # 90.95 +- 0.16%
     # optimizer2 = SignSGDMuon(filter_params, lr=0.4, momentum=0.6, nesterov=True, sgd_coeff=0.5, sign_lr_mult=0.002) # 0.005-> 93.93%, 0.002 -> 94.00%, 0.001-> 93.89%
     # optimizer2 = SignSGDMuon(filter_params, lr=0.4, momentum=0.6, nesterov=True, sgd_coeff=1, sign_lr_mult=0.009) # 89.66 +- 0.33%
     # optimizer2 = SGDMuon(filter_params, lr=0.24, momentum=0.6, nesterov=True, sgd_coeff=0.1) # 87% - does not work well, because it's not an LMO algorithm
-    optimizer2 = MuonOrSign(filter_params, lr=0.42, momentum=0.65, nesterov=True, sign_coeff=0.003) # 94.00% - see airbench_muon.py
[…]

### 2025-11-26 code/optimizers.py

@@ -154,6 +154,43 @@ class Muon(torch.optim.Optimizer):
                 update = zeropower_via_newtonschulz5(g.reshape(len(g), -1)).view(g.shape) # whiten the update
                 p.data.add_(update, alpha=-lr) # take a step
 
+
+class MuonSignedUpdate(torch.optim.Optimizer):
+    def __init__(self, params, lr=1e-3, momentum=0, nesterov=False, norm_weight=True):
+        if lr < 0.0:
+            raise ValueError(f"Invalid learning rate: {lr}")
+        if momentum < 0.0:
+            raise ValueError(f"Invalid momentum value: {momentum}")
+        if nesterov and momentum <= 0:
+            raise ValueError("Nesterov momentum requires a momentum")
+        defaults = dict(lr=lr, momentum=momentum, nesterov=nesterov)
+        super().__init__(params, defaults)
+        self.norm_weight = norm_weight
+
+    def step(self):
+        for group in self.param_groups:
+            lr = group["lr"]
+            momentum = group["momentum"]
+            for p in group["params"]:
+                g = p.grad
+                if g is None:
+                    continue
+                state = self.state[p]
+
+                if "momentum_buffer" not in state.keys():
+                    state["momentum_buffer"] = torch.zeros_like(g)
+                buf = state["momentum_buffer"]
+                buf.mul_(momentum).add_(g)
+                g = g.add(buf, alpha=momentum) if group["nesterov"] else buf
+
+                if self.norm_weight:
+                    norm = p.data.norm()
+                    if norm < 1e-10:
+                        norm = 1e-10
+                    p.data.mul_(len(p.data)**0.5 / norm) # normalize the weight
+                update = zeropower_via_newtonschulz5(g.reshape(len(g), -1)).view(g.shape).sign() # whiten the update
+                p.data.add_(update, alpha=-lr) # take a step
+
 class Muon2(torch.optim.Optimizer): # not ready yet!, we must have access to gradients by sample!
     def __init__(self, params, lr=1e-3, momentum=0, nesterov=False, norm_weight=True):
         if lr < 0.0:

### 2025-11-26 conf_article/neon.tex
Added Chebyshev

@@
 That is what can be said from a practical point of view.
-From a theoretical perspective, a key innovation of Muon was its principled derivation of the update rule, which emerges as the solution to an optimization problem constrained by the RMS-to-RMS norm (scaled version of spectral norm) \citep{bernstein2025deriving}
+From a theoretical perspective, a key innovation of Muon was its principled derivation of the update rule, which emerges as the solution to an optimization problem constrained by the RMS-to-RMS norm (scaled version of the spectral norm) \citep{bernstein2025deriving}
 
@@
 Among the notable ones are Scion \citep{pethick2025training}, Dion \citep{ahn2025dion} and Gluon \citep{riabinin2025gluon}.
-Those researches target Muon's efficiency and establish convergence bounds.
+Those works try to explain Muon's efficiency and establish convergence bounds.
 One central question, however, remains unanswered:
 
-\textit{In deriving Muon's update step, why constrain by the spectral norm? How would alternative norms affect performance and computational cost?}
+\textit{In deriving Muon's update step, why should one constrain by the spectral or any operator norm? How would alternative norms affect performance and computational cost?}
 
-In this article, we tackle this question by first showing the connection between matrix norms and corresponding existing algorithms and discuss the theoretic bounds derived for those algorithms.
-We then leverage the family of norms dual to Ky Fan k-norms to derive a new class of algorithms with low-rank updates, which we call Fanions.
-Subsequently, we create a second, hybrid family named F-Fanions by constructing convex combinations of the Ky Fan norms with the Frobenius norm and taking dual of that composite norm.
-Working within the linear minimization oracle (LMO) framework we derive the explicit update formulas for both algorithm families.
-As it was done with Muon, we stipulate our algorithms to be fast to approximate, for which we utilize the Lanczos algorithm as described in ~\cref{sec:matrix-side}.
+In this article, we tackle this question by actually showing that there are many viable non-operator norms.
+We leverage the family of norms dual to Ky Fan $k$-norms to derive a new family of Fanions, algorithms with low-rank updates.
+This approach theoretically explains the backbone of Dion's update \citep{ahn2025dion} and generalizes the memory-motivated application of the nuclear norm to Sharpness-Aware Minimization \citep{pethick2025sam}.
+As it was done with Muon, we come up with an effective procedure for computing Fanions' updates, Lanczos algorithm as described in ~\cref{sec:matrix-side} is the most effective algorithm, which, however, lacks an effective GPU- and PyTorch-friendly implementation.
+
+Working with duals to conic combinations of dual norms, we consturct the families of F-Fanions and S-Fanions, which are hybrids between Muon and NormalizedSGD and SignSGD, respectively.
 
@@
 	\item Pre-training NanoGPT on FineWeb dataset.
+	\item Fine-tuning NanoGPT on Tiny Stories
 \end{itemize}
 
-Our experiments reveal important insights about the role of matrix norms in optimization.
+Our experiments reveal important insights into the role of matrix norms in optimization.
+First, we show on the example of Neon, a one-rank Fanion, that not every LMO-based algorithm is effective, despite the general bounds of \citet{kovalev2025understanding} and \citet{riabinin2025gluon}.
+
 \textcolor{red}{On a synthetic least squares problem, we observe a striking discrepancy: while some algorithms converge slowly in terms of loss, they may converge quickly in terms of gradient norm, and vice versa}.
@@
 Most notably, our experiments on real-world tasks demonstrate that the choice of underlying matrix norm is remarkably flexible.
-On CIFAR-10 airbench, properly-tuned F-Muon achieves $94.01 \pm 0.14\%$ accuracy, essentially matching Muon's $94.00 \pm 0.13\%$ performance.
-On NanoGPT pre-training, F-Muon achieves 3.281 cross-entropy loss, only marginally worse than Muon's 3.279.
+On CIFAR-10 airbench, properly-tuned F-Muon and S-Muon achieve $94.01 \pm 0.14\%$ and $94.02 \pm 0.13\%$ accuracy, essentially matching Muon's $94.01 \pm 0.13\%$ performance.
+On NanoGPT pre-training, F-Muon achieves 3.281 cross-entropy loss, while fully-tuned Muon achieves 3.279.
+Finally, S-Muon beats both Muon and F-Muon on fine-tuning of NanoGPT on Tiny Stories.
 These results show that Muon-like algorithms can maintain competitive performance even when the underlying norm constraint is significantly modified, answering affirmatively the central question posed above.
@@
 $$
-depending on which one better minimizes the function $F(\mX^t) + \mM^t(\mX^{t+1} - \mX^t)$.
+depending on which one better minimizes the linear function $L(\mX) \equiv F(\mX^t) + \mM^t(\mX - \mX^t)$.
 
@@
 \section{Conic Combination of LMO-algorithms is an LMO-algorithm}
+\label{sec:conic_combo}
 
@@
 	$$
-	\mX^{k + 1} - \mX^{k} = \gamma_k \sum_{i = 1}^n \alpha_k \mathrm{LMO}_{\norm{\cdot}_{(i)}}(\mM^k)
+	\mX^{k + 1} - \mX^{k} = \gamma_k \sum_{i = 1}^n \alpha_i \mathrm{LMO}_{\norm{\cdot}_{(i)}}(\mM^k)
 	$$
@@
 \begin{equation}\label{eq:nsgd_muon_update}
-		\mX^{t+1} = \mX^{t} - \gamma \left(\alpha \sum_{i = 1}^k u_i v_i^\top + (1-\alpha)\frac{\mM^k}{\normf{\mM^k}}\right)\,,
+		\mX^{t+1} = \mX^{t} - \gamma \left(\alpha \sum_{i = 1}^k u_i v_i^\top + (1-\alpha)\frac{\mM^k}{\normf{\mM^k}}\right)\,.
 \end{equation}
-where $u_1, \dots, u_k$ ($v_1, \dots, v_k$) are first $k$ left (right) sighular vectors of $\mM^k$.
 By \cref{corollary:lin_comb_alg}, this is an LMO-algorithm corresponding to the unit ball of the norm $\norm{\cdot}_{\mathrm{F-KF-k}}$, given by
@@
 \subsection{Randomized Linear Least Squares}
-First, we compare F-Fanions on the following convex L-smooth problem:
+First, we compare F-Fanions on the following convex $L$-smooth problem:
 
@@
 We run different Fanions and their respective F-Fanions with $\alpha=1/2$: Neon (Fanion-1), Fanion-2, Fanion-5, Fanion-100, and Muon (Fanion-500).
-We test them against Normalized SGD, which is also a F-Fanion with $\alpha = 0$ and an arbitrary $k$.
+Also, we test S-Fanions with $\alpha=1/2$ and \texttt{sign\_lr\_coeff=0.01}.
+We test them against Normalized SGD and SignSGD, which is are also F-Fanion and S-Fanion respectively with $\alpha = 0$ and an arbitrary $k$.
 
-Since theoretical bounds \citep{kovalev2025understanding, riabinin2025gluon} rely on a very loose norm bound $\norm{\cdot}\le \rho \normf{\cdot}$, we do not derive learning rate or momentum from the smoothness constants.
-Rather, we find such \texttt{lr, momentum} pair that the loss threshold $0.01$ is reached in a minimal number of iterations.
+Since theoretical bounds \citep{kovalev2025understanding, riabinin2025gluon} rely on a very loose norm bound $\norm{\cdot}\le \rho \normf{\cdot}$, we do not derive learning rate or momentum from the smoothness constants that also depend on the norm choice.
+Rather, we find such \texttt{(lr, momentum)} pair that the loss threshold $0.01$ is reached in a minimal number of iterations.
 This setting is the most realistic, and at the same time aligns with the corollaries of the convergence theorems where constant learning rate and momentum coefficients are proposed.
 
-12 000 iterations for each algorithm ensure that the rate at some point gets low enough for each algorithm to converge. The results are presented in \cref{fig:lls} and ~\cref{sec:lls_plots_section}.
+The results are presented in \cref{fig:lls} and ~\cref{sec:lls_plots_section}.
 
-In terms of the loss minimization, F-Muon had the fastest convergence.
-In terms of the Frobenius norm of the gradient, the lowest was observed for NSGD.
-Then go F-Neon (F-Fanion-1), F-Fanion-2, F-Fanion-5, F-Fanion-100, and F-Muon (Fanion-500).
-It is noticeable that all F-Fanions have lower Frobenius norms of the gradients than their respective Fanioins.
-The same behavior will be recorded in the next section.
+Both F-Muon and S-Muon converge faster and to a lower loss and a lower Frobenius norm of the gradient than NSGD, Muon, or SignSGD.
 
@@
 \subsection{CIFAR-10 airbench}
-We adapt Keller Jordan's code to test F-Muon, Neon, and F-Neon on the CIFAR-10 airbench \citep{cifar2023airbench}.
-First, we run F-Muon for different $\alpha$ with the same {\tt lr=0.24(1 - step/total\_steps), momentum=0.6, nesterov=True}, as have been finetuned by Jordan, 10 repetitions for each $\alpha$.
-We record the accuracy after 8 epochs of training.
-The results are in \cref{fig:muon_alphas}.
+
+We compare the algorithms on the CIFAR-10 airbench \citep{cifar2023airbench}.
+First, to test the impact of $\alpha$, we run F-Muon for different $\alpha$ with the same {\tt lr=0.24(1 - step/total\_steps), momentum=0.6, nesterov=True} and weight normalization, as have been finetuned by Keller Jordan, 10 repetitions for each $\alpha$.
+We record the accuracy after 8 epochs of training (\cref{fig:muon_alphas}).
 
@@
 Tuned parameters are {\tt lr=0.4(1 - step/total\_steps), momentum=0.65, nesterov = True}.
-While Muon reaches $94.00 \pm 0.13$\% accuracy after 8 epochs, tuned F-Muon reaches $94.01 \pm 0.14\%$ after 8 epochs (average was done by 200 iterations).
+Tuned F-Muon reaches $94.01 \pm 0.14\%$ after 8 epochs (averaged by 200 iterations), matching Muon's accuracy and variance.
+Once again we measure \texttt{val\_accuracy}($\alpha$) curve (\cref{fig:fmuon_alphas}).
+We notice that even when $\alpha=0.1$, the accuracy is much higher than in the case of vanilla NSGD.
 
-Finally, we take this set of tuned parameters and test on different $\alpha$, 5 times for each $\alpha$.
-The results are in \cref{fig:fmuon_alphas}.
-We notice that even when $\alpha=0.1$, the accuracy is much higher than in case of the pure NSGD.
+We tune S-Muon with $\alpha=0.5$ for the best \texttt{(lr, momentum, sign\_lr\_coeff)} and by setting {\tt lr=0.42(1 - step/total\_steps), momentum=0.65, nesterov = True, sign\_lr\_coeff = 0.003} get $\bf{94.02} \pm 0.13\%$, which is even higher than for vanilla Muon.
 
@@
 	\end{subfigure}
-	\caption{Mean accuracies for different $\alpha$ of F-Muon.}
+	\caption{Mean validation accuracies for F-Muon with different $\alpha$}
 	\label{fig:diff_alphas}
@@
 
-The results are curious and could be represented by \cref{fig:cifar_ball}: the LMO ball that we plotted in a 2D space of singular values has drastically changed, but the observed convergence after tuning has not degraded.
-These observations raise the question of how much lmo-based algorithms are sensitive to the constraint area, i.e.
-what will happen if the ball is corrupted.
-In this particular example, we have had a blurred ball, which proved as robust as the original ball.
+The Muon-level performace is surprising for the contorted geometry of F-Muon and S-Muon.
+F-Muon can be visualized by \cref{fig:cifar_ball}, where the LMO balls of Muon and F-Muon for real \texttt{lr} are plotted in a 2D space of singular values.
+S-Muon cannot be visualized because the $l_\infty$ norm is not a function of singular values, while $\norms{\cdot}$, when plotted for the $\Rmn$ space with $m, n > 2$, requires at least a 4D space.
+However, it is plain that the LMO ball of S-Muon significantly differs from the ball of Muon because the $\tt{lr}(1-\alpha)\tt{sign\_lr\_coeff} = 6.3 \times 10^{-4}$, which is comparable to the usual \texttt{lr} of SignSGD for deep learning problems.
+
+The most pathological case is $\alpha > 1$, which corresponds to the LMO with an area that is not a norm ball!
+Despite this violation, the mixture of algorithms reaches almost the same accuracy as vanilla Muon.
+
+ These observations raise the question of how much LMO-based algorithms are sensitive to the constraint area.
 
@@
 
-The most pathological case is $\alpha > 1$, which corresponds to the lmo with an area that is not a norm ball!
-Despite this violation, the mixture of algorithms reaches almost the same accuracy as vanilla Muon.
-
 \subsection{NanoGPT speedrun}
@@
 For Muon, it falls below the target threshold 3.28 and reaches 3.279.
-However, this difference is negligible, if one looks at \cref{fig:gpt_loss}.
-It is even more striking, considering the fact that F-Muon is an average of Muon and NSGD, and the latter performed quite poorly.
+However, this difference is negligible if one looks at \cref{fig:gpt_loss}.
+It is even more striking, considering the fact that this F-Muon is an average of Muon and NSGD, and the latter performed quite poorly.
 
@@
 
-Again, as on the CIFAR airbench, if we set $\alpha=-0.1$, F* to which is not a norm, we get a 3.2818 loss, which is not higher than for $\alpha=0.3$, for example.
+Again, as on the CIFAR airbench, if we set $\alpha=-0.1$ that corresponds to no ball, we get a 3.2818 loss, which is a small difference from the baseline.
 
@@
 To construct their COSMOS optimizer, \citet{chen2025cosmoshybridadaptive} applied computationally intensive updates of SOAP optimizer \citep{vyas2025soap} to a low-dimensional "leading eigensubspace" while using memory-efficient methods like Muon for the remaining parameters.
-\citet{amsel2025polar} proposed a more efficient alternative to Newton-Schulz algorithm.
+\citet{amsel2025polar, grishina2025chebyshev} proposed more efficient alternatives to Newton-Schulz algorithm.
 \citet{si2025adamuon} introduced AdaMuon which combines element-wise adaptivity with orthogonal updates.
@@
 For example, \citet{kovalev2025understanding} has provided convergence guarantees of Muon and in various settings, from which, however, Muon's supremacy cannot be recovered.
-Indeed, although the obtained bounds depend on the norm choice, the asymptotics of the convergence remain the same as for NSGD and other optimizers, $K = \cO(\epsilon^{-4})$ in an L-smooth stochastic case.
+Indeed, although the obtained bounds depend on the norm choice, the asymptotics of the convergence remain the same as for NSGD and other optimizers, $K = \cO(\epsilon^{-4})$ in an $L$-smooth stochastic case.
 
-	A similar drawback has the article \citet{riabinin2025gluon}, where L-smoothness assumption is replaced with a more practical $(L_0, L_1)$-smoothness.
+	A similar drawback has the article \citet{riabinin2025gluon}, where $L$-smoothness assumption is replaced with a more practical $(L_0, L_1)$-smoothness.
 By estimating smoothness and substituting it into their Theorem 1, the authors recovered the optimal fine-tuned stepsizes reported by \citet{pethick2025training}.
@@
 	Common important drawback of the analyses is the consideration of the convergence by the norm of the gradient.
-As we showed in our experiments on CIFAR, the gradient norm may decrease by a factor of ten, while the accuracy reaches 100\%.
+As we showed in our experiments on CIFAR, the gradient norm may decrease only by a factor of ten, when the accuracy reaches 100\%.
 
-	We suppose that the reason for the recorded by us stark discrepancy between Neon and Muon performance, both of which are described by Stochastic Conditional Gradient or Gluon frameworks, lies in the structure of the norm ball or in the preconditioner interpretation \citep{pethick2025trainingneuralnetworksscale}, which must be an object of further research.
+	We suppose that the reason for the recorded stark discrepancy between Neon and Muon performance, both of which are described by Stochastic Conditional Gradient \citep{pethick2025training} or Gluon frameworks, lies in the structure of the norm ball or in the preconditioner interpretation \citep{pethick2025trainingneuralnetworksscale}, which must be an object of further research.
 
@@
 They could be made more memory-effective by the zero-order techniques that worked for Muon \citep{petrov2025zeromuon}.
-Finally, the results from \citet{shulgin2025inexactmuon} can be used to theoretically derive faster yet working approximate schemes to calculate the LMO; power iterations with a very limited number of iterations is a good aglorithm to consider for the analysis.
+Finally, the results from \citet{shulgin2025inexactmuon} can be used to explain the robustness of Muon to the norm change that occured in our experiments and to theoretically derive faster yet working approximate schemes to calculate the LMO; power iterations with a very limited number of iterations is a good aglorithm to consider for the analysis.
 
@@
 DM supervised the project and helped with editing the article.
-IK, NK, and AV participated mainly on the first stage of research, when it has been a project in the optimization course at MIPT.
-IK suggested using composite norms (though not F2 and F*) and KKT conditions to find the resulting updates.
-NK suggested Lanczos algorithm as the fastest tool to compute Neon's updates and conducted experiments to prove it.
-AV tested Neon on the finetuning of NanoGPT.
+IK suggested using composite norms (though not the ones that induce F-Fanions and S-Fanions) and refactored \cref{sec:conic_combo}.
+NK suggested Lanczos algorithm as the most precise means to compute Fanions' updates, conducted experiments to prove it, and wrote ~\cref{sec:matrix-side}.
+AV tested Neon on the finetuning of NanoGPT on Tiny Stories.
 All other work was done by AK.
@@
 
-\section{More plots for Linear Least Squares}
+\section{More data for Linear Least Squares}
 \label{sec:lls_plots_section}
+
+\begin{table}[ht]
+	\centering
+	\begin{tabular}{|c|c|c|c|}
+	  \hline
+	  Method & lr  & momentum           & Iterations to 0.01 loss \\
+	  \hline
+	  NSGD & 0.25 & 0.9 & 350 \\
+	  \hdashline
+	  SignSGD & 0.055 $\cdot$ 0.01 & 0.95 & 700 \\
+	  \hdashline
+	  Muon & 0.025 & 0.5 & 290 \\
+	  \hdashline
+	  F-Muon & 0.035 & 0.5 & 320 \\
+	  \hdashline
+	  S-Muon & 0.035 & 0.8 & 270 \\
+	  \hline
+	\end{tabular}
+  \caption{Tuning lrs and momentum coefficients}
+  \label{tbl:lls_lrs}
+  \end{table}
 

### 2025-11-26 conf_article/neon.tex
added affilation for me

@@
 Ivan Kozyrev \\
-Moscow Institute of Physics and Technology (MIPT) \\
+Moscow Institute of Physics and Technology (MIPT), \\
+Marchuk Institute of Numerical Mathematics \\
 \texttt{kozyrev.in@phystech.edu} \\
@@
 Their most prominent members are {\it F-Muon} and {\it S-Muon}.
-We complement our theoretical analysis with an extensive empirical study of the algorithms across a wide range of tasks and settings, from which it follows that F-Muon and S-Muon are always on par with Muon and S-Muo, while on fine-tuning of NanoGPT and synthetic linear least squares even better than vanilla Muon.
+We complement our theoretical analysis with an extensive empirical study of the algorithms across a wide range of tasks and settings, from which it follows that F-Muon and S-Muon are always on par with Muon and S-Muon, while on fine-tuning of NanoGPT and synthetic linear least squares even better than vanilla Muon.
 
@@
 This approach theoretically explains the backbone of Dion's update \citep{ahn2025dion} and generalizes the memory-motivated application of the nuclear norm to Sharpness-Aware Minimization \citep{pethick2025sam}.
-As it was done with Muon, we come up with an effective procedure for computing Fanions' updates, Lanczos algorithm as described in ~\cref{sec:matrix-side} is the most effective algorithm, which, however, lacks an effective GPU- and PyTorch-friendly implementation.
+As it was done with Muon, we come up with an effective procedure for computing Fanions' updates, Lanczos algorithm as described in~\cref{sec:matrix-side} is the most effective algorithm, which, however, lacks an effective GPU- and PyTorch-friendly implementation.
 

### 2025-11-26 code/L_smooth_tests/benchmark_plotter.py
LLS done

@@ -61,7 +61,7 @@ def plot_from_descriptions(
         if yscale:
             ax.set_yscale(yscale)
         ax.grid(True, alpha=0.3)
-        ax.legend()
+        ax.legend(ncol=2, frameon=True, fancybox=True, shadow=True)
 
     # Hide any unused subplots
     for j in range(num_panels, len(axes_list)):
@@ -128,7 +128,7 @@ def plot_and_save_individual_panels(
         
         # Add grid and legend
         ax.grid(True, alpha=0.3)
-        ax.legend()
+        ax.legend(ncol=2, frameon=True, fancybox=True, shadow=True)
         
         # Create filename from title (sanitized for filesystem)
         filename = base_title.lower().replace(" ", "_").replace("(", "").replace(")", "").replace("||", "").replace("∇", "grad").replace("f", "f").replace("X", "X").replace(":", "").replace("__", "_")

### 2025-11-26 code/L_smooth_tests/finetune_lrs.py

@@ -279,10 +279,10 @@ def main() -> None:
         np.round(np.linspace(0.01, 0.1, 19), 3),
         np.round(np.linspace(0.15, 1, 18), 3)
     ))
-    # learning_rates = np.round(np.linspace(0.01, 0.1, 19), 3) # - for 0.01 loss
+    learning_rates = np.round(np.linspace(0.01, 0.1, 19), 3) # - for 0.01 loss
     # learning_rates = learning_rates[learning_rates < 0.06]
-    learning_rates = np.round(np.logspace(np.log10(0.0005), np.logspace(0.1), 15), 3) # - for 0.001 loss
-    # learning_rates = np.round(np.linspace(0.005, 0.020, 16), 3) # - for 0.001 loss
+    # learning_rates = np.round(np.logspace(np.log10(0.005), np.log10(0.5), 15), 4) # - for 0.001 loss
+    #learning_rates = np.round(np.linspace(0.005, 0.020, 16), 3) # - for 0.001 loss
     
     print(f"Learning rate grid: {learning_rates}")
     
@@ -300,14 +300,14 @@ def main() -> None:
     rint = 10
     # Define algorithm specifications (without fixed lr and momentum if grid searching)
     algorithm_specs: List[Dict[str, Any]] = [
-        # dict(
-        #     name="NSGD",
-        #     optimizer_class=NormalizedMuon,
-        #     optimizer_kwargs=dict(nesterov=True, sgd_coeff=1),
-        #     num_iterations=600,
-        #     record_interval=rint,
-        #     use_momentum=momentums is not None,  # Use momentum grid search if momentums are provided
-        # ),
+        dict(
+            name="NSGD",
+            optimizer_class=NormalizedMuon,
+            optimizer_kwargs=dict(nesterov=True, sgd_coeff=1),
+            num_iterations=1500,
+            record_interval=rint,
+            use_momentum=momentums is not None,  # Use momentum grid search if momentums are provided
+        ),
         # dict(
         #     name="MLion",
         #     optimizer_class=MLion,
@@ -338,14 +338,14 @@ def main() -> None:
         #     record_interval=rint,
         #     use_momentum=momentums is not None,
         # ),
-        dict(
-            name="SignSGD",
-            optimizer_class=SignSGDMuon,
-            optimizer_kwargs=dict(nesterov=True, sign_lr_mult=0.01, sgd_coeff=1),
-            num_iterations=3500,
-            record_interval=rint,
[…]

### 2025-11-26 code/L_smooth_tests/run_benchmarks_new.py

@@ -98,8 +98,8 @@ def main() -> None:
     m, n = 500, 500
     # problem = LogisticRegressionProblem(m, n) 
     problem = SimpleQuadratic(m, n, device=device, seed=42, eig_range_M=(0,1), eig_range_N=(0,1), shift_scale=0)
-    iter_num_nuc = iter_num_op = iter_num_fro = 1200
-    record_period=25
+    iter_num_nuc = iter_num_op = iter_num_fro = 5000
+    record_period = 10
     # iter_num = 100000
     X0 = 0.1 * torch.randn(m, n, dtype=torch.float32, device=device)
     # Define optimizers and settings
@@ -120,16 +120,24 @@ def main() -> None:
     smuon_mom = 0.9
 
     # Define experiment specifications
-    colors = cm.tab20b(np.linspace(0, 1, 5))[::-1]
+    # Colors will be assigned based on base optimizer name
+    # Linestyles: plain = "-", F- = ":", S- = "--"
     experiment_specs: Dict[str, Dict[str, Any]] = {
         "NSGD": dict(
             optimizer_class=NormalizedMuon,
-            optimizer_kwargs=dict(lr=0.25, momentum=0.9, nesterov=True, sgd_coeff=1),
+            optimizer_kwargs=dict(lr=0.08, momentum=0.95, nesterov=True, sgd_coeff=1),
             num_iterations=iter_num_fro,
             record_interval=record_period,
             verbose=True,
             lr_scheduler=const_lr,
-            color="gray"
+        ),
+        "SignSGD": dict(
+            optimizer_class=SignSGDMuon,
+            optimizer_kwargs=dict(lr=0.016, momentum=0.95, nesterov=True, sign_lr_mult=0.01, sgd_coeff=1),
+            num_iterations=iter_num_fro,
+            record_interval=record_period,
+            verbose=True,
+            lr_scheduler=const_lr,
         ),
         #  "Random F-Muon": dict(
         #     optimizer_class=RandomNormalizedMuon,
@@ -167,7 +175,6 @@ def main() -> None:
             verbose=True,
             lr_scheduler=const_lr,
             requires_cuda=True,
-            color=colors[0]
         ),
         "F-Neon": dict(
             optimizer_class=Neon,
@@ -177,7 +184,15 @@ def main() -> None:
             verbose=True,
             lr_scheduler=const_lr,
             requires_cuda=True,
-            color=colors[0]
+        ),
+        "S-Neon": dict(
+            optimizer_class=Neon,
[…]

### 2025-11-26 code/optimizers.py

@@ -269,13 +269,14 @@ class MuonCringeMomentum(torch.optim.Optimizer):
 # this function also supports F-Neon
 class Neon(torch.optim.Optimizer):
     def __init__(self, params, lr=1e-3, momentum=0, k=1, tau = 0, nesterov=False,
-                 neon_mode='fast', iter_num=1, sgd_coeff=0, norm_weight=True):
+                 neon_mode='fast', iter_num=1, sgd_coeff=0, norm_weight=True, sign_lr_mult=None):
         self.tau = tau
         self.k = k # target number of SVD componenets which we preserve
         self.lanczos_iter_num = iter_num
         self.type = neon_mode # fast (vanilla Neon), accurate (old F*-Neon), Ky-Fan (~Dion)
         self.sgd_coeff = sgd_coeff
         self.norm_weight = norm_weight
+        self.sign_lr_mult = sign_lr_mult
         if lr < 0.0:
             raise ValueError(f"Invalid learning rate: {lr}")
         if momentum < 0.0:
@@ -326,7 +327,12 @@ class Neon(torch.optim.Optimizer):
                         u, s, vt = several_sv_svds_approximation(g_resh, self.k, self.lanczos_iter_num)
                         update = u @ vt
                         error = u @ torch.diag(s) @ vt
-                update2 = (1-self.sgd_coeff) * update.view(g.shape) + self.sgd_coeff * g / (g.norm() + 1e-12)
+                if self.sign_lr_mult is not None:
+                    # Use sign-based update when sign_lr_mult is provided
+                    update2 = (1-self.sgd_coeff) * update.view(g.shape) + self.sgd_coeff * g.sign() * self.sign_lr_mult
+                else:
+                    # Default: normalized gradient
+                    update2 = (1-self.sgd_coeff) * update.view(g.shape) + self.sgd_coeff * g / (g.norm() + 1e-12)
                 p.data.add_(update2, alpha=-lr)
 
 
@@ -594,8 +600,11 @@ class SignSGDMuon(torch.optim.Optimizer):
                 if self.norm_weight:
                     p.data.mul_(len(p.data)**0.5 / p.data.norm()) # normalize the weight
                 # update = zeropower_via_newtonschulz5(g.reshape(len(g), -1)).view(g.shape) # whiten the update
-                update_part = zeropower_via_newtonschulz5(g.reshape(len(g), -1)).view(g.shape)
[…]

### 2025-11-26 code/L_smooth_tests/benchmark_plotter.py
Improved style of the LLS plots

@@ -61,7 +61,7 @@ def plot_from_descriptions(
         if yscale:
             ax.set_yscale(yscale)
         ax.grid(True, alpha=0.3)
-        ax.legend(ncol=2, frameon=True, fancybox=True, shadow=True)
+        ax.legend(ncol=2, frameon=True, fancybox=True, loc='upper right', bbox_to_anchor=(1, 0.85), framealpha=0.95)
 
     # Hide any unused subplots
     for j in range(num_panels, len(axes_list)):
@@ -128,7 +128,7 @@ def plot_and_save_individual_panels(
         
         # Add grid and legend
         ax.grid(True, alpha=0.3)
-        ax.legend(ncol=2, frameon=True, fancybox=True, shadow=True)
+        ax.legend(ncol=2, frameon=True, fancybox=True, loc='upper right', bbox_to_anchor=(1, 0.85), framealpha=0.95)
         
         # Create filename from title (sanitized for filesystem)
         filename = base_title.lower().replace(" ", "_").replace("(", "").replace(")", "").replace("||", "").replace("∇", "grad").replace("f", "f").replace("X", "X").replace(":", "").replace("__", "_")
@@ -162,7 +162,13 @@ def build_default_panels(experiments: Dict[str, Dict[str, Any]]) -> List[Dict[st
 
     def collect_series(x_key: str, y_key: str) -> List[Dict[str, Any]]:
         series: List[Dict[str, Any]] = []
-        for name, res in experiments.items():
+        # Use preserved order if available, otherwise use dict iteration order
+        experiment_order = experiments.get("_experiment_order", list(experiments.keys()))
+        # Filter out the special _experiment_order key
+        experiment_order = [name for name in experiment_order if name != "_experiment_order" and name in experiments]
+        
+        for name in experiment_order:
+            res = experiments[name]
             # Pass through optional style fields if present in the experiment results
             series_item: Dict[str, Any] = {
                 "name": name,
@@ -318,6 +324,7 @@ def load_experiments_from_csv(csv_path: str) -> Dict[str, Dict[str, Any]]:
     """
     Load experiments from CSV produced by save_experiments_to_csv and
     reconstruct the original experiments dict structure expected by plotting helpers.
+    Preserves the order of experiments as they first appear in the CSV.
     """
     df = pd.read_csv(csv_path)
     # Ensure types
[…]

### 2025-11-26 conf_article/neon.tex
added text about signsgd in sections 1 and 4

@@
 \DeclarePairedDelimiter{\normn}{\|}{\|_{\mathrm{nuc}}}
+\DeclarePairedDelimiter{\normc}{\|}{\|_{\mathrm{C}}}
 \def\<#1,#2>{\langle #1,#2\rangle}
@@
 Using this formula it is easy to compute updates for algorithms induced by various norms $\norm{\cdot}$.
+
 For example, when the norm $\norm{\cdot}$ is the Frobenius norm $\normf{\cdot}$, \cref{eq:our_update} turns into
@@
 \end{equation} 
-which recovers Normalized SGD (NSGD). When one has the spectral norm $\norms{\cdot}$, one gets
+which recovers Normalized SGD (NSGD).
+
+And when the norm is the spectral norm $\norms{\cdot}$, we get
 \begin{equation}\label{eq:muon_update}
@@
 Muon can be recovered by taking the RMS-to-RMS operator norm $\sqrt{\frac{n}{m}}\norms{\cdot}$.
+
+When the norm is the Chebyshev norm $\normc{\cdot}$, we get
+\begin{equation}\label{eq:signsgd_update}
+	\mX^{k+1} = \mX^{k} - \eta_k \sign(\mM^k)\,,
+\end{equation}
+which recovers SignSGD~\citep{bernstein2018signsgd}.
+Here, $\sign(\mM^k)$ denotes the element-wise sign function.
+SignSGD is particularly notable for its communication efficiency in distributed training, as it compresses gradients to 1-bit per parameter.
 
@@
 the sum of the $k$ largest singular values of the matrix.
-Special cases of the Ky Fan $k$-norm are the Ky Fan $1$-norm, which is the sprectral norm, and the Ky Fan $\min\{m, n\}$-norm, which is the nuclear norm.
+Special cases of the Ky Fan $k$-norm are the Ky Fan $1$-norm, which is the spectral norm, and the Ky Fan $\min\{m, n\}$-norm, which is the nuclear norm.
 
@@
 	$$
-		\norm{x} := \sum_{i = 1}^n \alpha_i \norm{x}_{(i)}.
+		\norm{\cdot} := \sum_{i = 1}^n \alpha_i \norm{\cdot}_{(i)}.
 	$$
@@
 
-Applying it to optimization algorithms we obtain the following corollary.
+Applying it to optimization algorithms, we obtain the following corollary.
 
@@
 
-Combining Fanions-$k$ with another excellent LMO-algorithm, NSGD, we obtain the family of algorithms with updates defined by
+Combining Fanions for various $k$ with another excellent LMO-algorithm, NSGD, we obtain the family of algorithms with updates defined by
 \begin{equation}\label{eq:nsgd_muon_update}
-		\mX^{t+1} = \mX^{t} - \gamma \left(\alpha \sum_{i = 1}^k u_i v_i^\top + (1-\alpha)\frac{\mM^k}{\normf{\mM^k}}\right)\,.
+		\mX^{t+1} = \mX^{t} - \gamma_k \left(\alpha \sum_{i = 1}^k u_i v_i^\top + (1-\alpha)\frac{\mM^k}{\normf{\mM^k}}\right)\,.
 \end{equation}
-By \cref{corollary:lin_comb_alg}, this is an LMO-algorithm corresponding to the unit ball of the norm $\norm{\cdot}_{\mathrm{F-KF-k}}$, given by
-\begin{equation}\label{eq:fstar_norm}
-	\norm{\cdot}_{\mathrm{F-KF-k}}^\dagger = \alpha \normkfk{\cdot}^\dagger + (1-\alpha)\normf{\cdot}^\dagger\,.
+Recall that Fanion-$k$ update is governed by the norm dual to the Ky Fan-$k$ norm, and NSGD corresponds to the Frobenius norm, which is dual to itself.
+Thus, by \cref{corollary:lin_comb_alg}, our new algorithm is an LMO-algorithm corresponding to the unit ball of the norm $\norm{\cdot}_{\mathrm{F-KF-k}}$:
+\begin{equation}\label{eq:fkfk_norm}
+	\norm{\cdot}_{\mathrm{F-KF-k}}^\dagger = \alpha \normkfk{\cdot} + (1-\alpha)\normf{\cdot}\,.
 \end{equation}
-We name the derived algorithms \emph{F-Fanions}.
-A bit of information and visualizations related to the $\normfstar{\cdot}$ norm is presented in the appendix (see \cref{eq:normfstardual}, \cref{fig:fstardual}).
+We name the derived family of algorithm \emph{F-Fanions}.
 
-Two edge members of this family, with $k = 1$ and $k = \min\{m, n\}$ correspondingly, are \emph{F-Neon} and \emph{F-Muon}.
+Two edge members of this family, with $k = 1$ and $k = \min\{m, n\}$ correspondingly, are \emph{F-Neon} and \emph{F-Muon}, are of particular interest to us.
+A bit of information and visualizations related to the $\normfstar{\cdot} = \norm{\cdot}_{\mathrm{F-KF-1}}$ norm is presented in the appendix (see \cref{eq:normfstardual}, \cref{fig:fstardual}).
 
 \subsection{Add SignSGD: S-Muon and S-Neon}
-Trivial formulas, but we need more experiments.
+
+Another algorithm we consider is the SignSGD. Mixing its update with that of Fanion-$k$, we obtain the update of \emph{S-Fanion-$k$}:
+\begin{equation}\label{eq:sign_muon_update}
+		\mX^{t+1} = \mX^{t} - \gamma_k \left(\alpha \sum_{i = 1}^k u_i v_i^\top + (1-\alpha)\eta\sign(\mM^k)\right)\,,
+\end{equation}
+where $\eta$ is the special SignSGD learning rate coefficient.
+
+The norm $\norm{\cdot}_{\mathrm{C-KF-k}}$ which produces this update in the LMO framework is again given by \cref{corollary:lin_comb_alg}:
+\begin{equation}\label{eq:ckfk_norm}
+	\norm{\cdot}_{\mathrm{C-KF-k}}^\dagger = \alpha \normkfk{\cdot} + \frac{1-\alpha}{\eta}\normc{\cdot}^\dagger\,.
+\end{equation}
 

### 2025-11-26 conf_article/neon.tex
Added LLS

--- 2025-11-26 conf_article/neon.tex
@@
 That is what can be said from a practical point of view.
-From a theoretical perspective, a key innovation of Muon was its principled derivation of the update rule, which emerges as the solution to an optimization problem constrained by the RMS-to-RMS norm (scaled version of the spectral norm) \citep{bernstein2025deriving}
+From a theoretical perspective, a key innovation of Muon was its principled derivation of the update rule, which emerged as the solution to an optimization problem constrained by the RMS-to-RMS norm (scaled version of the spectral norm) \citep{bernstein2025deriving}
 
@@
 
-\textit{In deriving Muon's update step, why should one constrain by the spectral or any operator norm? How would alternative norms affect performance and computational cost?}
+\textit{In deriving Muon's update step, why should one constrain by the spectral or any other operator norm? How would alternative norms affect performance and computational cost?}
 
 In this article, we tackle this question by actually showing that there are many viable non-operator norms.
-We leverage the family of norms dual to Ky Fan $k$-norms to derive a new family of Fanions, algorithms with low-rank updates.
+We leverage the family of norms dual to Ky Fan $k$-norms to derive a new family of \textbf{Fanions}, algorithms with low-rank updates.
 This approach theoretically explains the backbone of Dion's update \citep{ahn2025dion} and generalizes the memory-motivated application of the nuclear norm to Sharpness-Aware Minimization \citep{pethick2025sam}.
-As it was done with Muon, we come up with an effective procedure for computing Fanions' updates, Lanczos algorithm as described in ~\cref{sec:matrix-side} is the most effective algorithm, which, however, lacks an effective GPU- and PyTorch-friendly implementation.
+As it was done with Muon, we come up with an effective procedure for computing Fanions' updates.
+Lanczos algorithm, which is described and compared with its competitors in ~\cref{sec:matrix-side}, is the most operation-effective algorithm, which, however, for now lacks an effective GPU- and PyTorch-friendly implementation.
 
-Working with duals to conic combinations of dual norms, we consturct the families of F-Fanions and S-Fanions, which are hybrids between Muon and NormalizedSGD and SignSGD, respectively.
+Working with duals to conic combinations of dual norms, we consturct the families of \textbf{F-Fanions} and \textbf{S-Fanions}, which are hybrids of Muon and NormalizedSGD and SignSGD, respectively.
 
-We then compare the performance of the algorithm families across various benchmarks (\cref{sec:experiments}):
+Then We compare the performances of the algorithm families on various model and real-world problems:
 \begin{itemize}
-	\item Synthetic least squares experiment,
-	\item CIFAR-10 airbench,
-	\item Pre-training NanoGPT on FineWeb dataset.
-	\item Fine-tuning NanoGPT on Tiny Stories
+	\item Synthetic least squares experiment \cref{subsec:lls_exps}
+	\item CIFAR-10 airbench \citep{cifar2023airbench}
+	\item Pre-training NanoGPT on FineWeb dataset \citep{modded_nanogpt_2024}
+	\item Fine-tuning NanoGPT on Tiny Stories \citep{eldan2023tinystoriessmalllanguagemodels}
 \end{itemize}
@@
 Our experiments reveal important insights into the role of matrix norms in optimization.
-First, we show on the example of Neon, a one-rank Fanion, that not every LMO-based algorithm is effective, despite the general bounds of \citet{kovalev2025understanding} and \citet{riabinin2025gluon}.
-
-\textcolor{red}{On a synthetic least squares problem, we observe a striking discrepancy: while some algorithms converge slowly in terms of loss, they may converge quickly in terms of gradient norm, and vice versa}.
-This suggests that existing theoretical guarantees may not fully explain practical algorithm performance.
+First, we show on the example of Neon, the one-rank Fanion, that not every LMO-based algorithm is effective, despite the same asymptotics in the general bounds of \citet{kovalev2025understanding} and \citet{riabinin2025gluon}.
+This suggests that existing theoretical guarantees should be reworked to explain empirical performance.
 
 Most notably, our experiments on real-world tasks demonstrate that the choice of underlying matrix norm is remarkably flexible.
-On CIFAR-10 airbench, properly-tuned F-Muon and S-Muon achieve $94.01 \pm 0.14\%$ and $94.02 \pm 0.13\%$ accuracy, essentially matching Muon's $94.01 \pm 0.13\%$ performance.
+On CIFAR-10 airbench, properly-tuned F-Muon and S-Muon achieve $94.04 \pm 0.13\%$ and $94.03 \pm 0.13\%$ accuracy, matching Muon's $94.01 \pm 0.13\%$ performance.
 On NanoGPT pre-training, F-Muon achieves 3.281 cross-entropy loss, while fully-tuned Muon achieves 3.279.
-Finally, S-Muon beats both Muon and F-Muon on fine-tuning of NanoGPT on Tiny Stories.
+Finally, S-Muon matches Muon on fine-tuning of NanoGPT on Tiny Stories, while F-Muon is far more resistant to the learning rate choice than Muon.
 These results show that Muon-like algorithms can maintain competitive performance even when the underlying norm constraint is significantly modified, answering affirmatively the central question posed above.
-This flexibility suggests potential for developing easier-to-compute variants of successful algorithms like Muon.
+Moreover, the tools from ~\ref{sec:conic_combo} give the researchers an unheard-of flexibility in designing algorithms that does not have to be modifications of Muon.
 
@@
 \subsection{Randomized Linear Least Squares}
+\label{subsec:lls_exps}
 First, we compare F-Fanions on the following convex $L$-smooth problem:
@@
 
-We run different Fanions and their respective F-Fanions with $\alpha=1/2$: Neon (Fanion-1), Fanion-2, Fanion-5, Fanion-100, and Muon (Fanion-500).
+We run different Fanions and their respective F-Fanions with $\alpha=1/2$: Neon (Fanion-1), Fanion-2, Fanion-10, Fanion-100, and Muon (Fanion-500).
 Also, we test S-Fanions with $\alpha=1/2$ and \texttt{sign\_lr\_coeff=0.01}.
@@
 
-Since theoretical bounds \citep{kovalev2025understanding, riabinin2025gluon} rely on a very loose norm bound $\norm{\cdot}\le \rho \normf{\cdot}$, we do not derive learning rate or momentum from the smoothness constants that also depend on the norm choice.
-Rather, we find such \texttt{(lr, momentum)} pair that the loss threshold $0.01$ is reached in a minimal number of iterations.
+Since theoretical bounds \citep{kovalev2025understanding, riabinin2025gluon} rely on a very loose norm bound $\norm{\cdot}\le \rho \normf{\cdot}$, we do not derive learning rate or Nesterov momentum from the smoothness constants that also depend on the norm choice.
+Rather, we find such \texttt{(lr, momentum)} pair that the loss threshold $0.001$ is reached in a minimal number of iterations.
 This setting is the most realistic, and at the same time aligns with the corollaries of the convergence theorems where constant learning rate and momentum coefficients are proposed.
 
-The results are presented in \cref{fig:lls} and ~\cref{sec:lls_plots_section}.
+We perform a grid search for \texttt{momentum} $\in \{0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 0.95\}$ and for \texttt{lr}.
+For Muon's, F-Muon's, S-Muon's \texttt{lr} $\in [0.005, 0.0020]$.
+For SignSGD, \texttt{lr} $\in 0.01 \times [0.005, 0.020]$, and for NSGD \texttt{lr} $\in [0.01, 0.10]$.
+The tuned parameters and the number of iterations to converge to $0.001$ are in \cref{tbl:lls_lrs}.
+\texttt{lr} and \texttt{momentum}, and \texttt{sign\_lr\_coeff} for Fanions, F-Fanions, and S-Fanions are the same as for Muon, F-Muon, and S-Muon respectively due to the impossibility to tune them properly: their losses decrease too slow to reach 0.001.
+
+The results are presented in \cref{fig:lls} and detailed in ~\cref{sec:lls_plots_section}.
 
@@
         \centering
-        \includegraphics[width=\linewidth]{figs/simple_lls/loss_vs_iteration_500x500.pdf}
+        \includegraphics[width=\linewidth]{figs/simple_lls/loss_vs_iteration_long_run.pdf}
         \caption{The loss}
@@
         \centering
-        \includegraphics[width=\linewidth]{figs/simple_lls/gradient_frobenius_norm_vs_iteration_500x500.pdf}
+        \includegraphics[width=\linewidth]{figs/simple_lls/gradient_frobenius_norm_vs_iteration_long_run.pdf}
         \caption{The Frobenius norm of the gradient}
@@
 Then we tune F-Muon with $\alpha = 0.5$.
-Tuned parameters are {\tt lr=0.4(1 - step/total\_steps), momentum=0.65, nesterov = True}.
-Tuned F-Muon reaches $94.01 \pm 0.14\%$ after 8 epochs (averaged by 200 iterations), matching Muon's accuracy and variance.
+Tuned parameters are {\tt lr=0.42(1 - step/total\_steps), momentum=0.63, nesterov = True}.
+Tuned F-Muon reaches $94.04 \pm 0.13\%$ after 8 epochs (averaged by 200 iterations), matching Muon's accuracy and variance.
 Once again we measure \texttt{val\_accuracy}($\alpha$) curve (\cref{fig:fmuon_alphas}).
@@
 
-We tune S-Muon with $\alpha=0.5$ for the best \texttt{(lr, momentum, sign\_lr\_coeff)} and by setting {\tt lr=0.42(1 - step/total\_steps), momentum=0.65, nesterov = True, sign\_lr\_coeff = 0.003} get $\bf{94.02} \pm 0.13\%$, which is even higher than for vanilla Muon.
+We tune S-Muon with $\alpha=0.5$ for the best \texttt{(lr, momentum, sign\_lr\_coeff)} and by setting {\tt lr=0.42(1 - step/total\_steps), momentum=0.63, nesterov = True, sign\_lr\_coeff = 0.003} get $\bf{94.03} \pm 0.13\%$, which is even higher than for vanilla Muon.
 
@@
 
+\subsection{NanoGPT Fine-Tuning}
+
+We fine-tuned the NanoGPT framework by Karpathy \citep{Karpathy2022} using the standard \textit{NanoGPT-medium} configuration, which corresponds to GPT-2 Medium.
+This model consists of 24 transformer layers, a hidden size of 1024, 16 attention heads, and approximately 345 million parameters.
+The training dataset was the \textit{Tiny Stories} corpus \citep{eldan2023tinystoriessmalllanguagemodels}, selected for its high entropy and structural diversity, which amplify and make more visible the differences in optimization dynamics.
+All training runs were initialized with the same random seed, weight initialization, and learning rate schedule to ensure that any differences in performance arose solely from the choice of optimizer.
+
+Experiments were conducted on a single NVIDIA RTX 4090 (24GB) GPU using PyTorch~2.8 with CUDA~12.8 and cuDNN~9.0, running on a workstation equipped with an Intel Core i9-14900KS CPU and 128 GB of RAM.
+No distributed or mixed-hardware training setups were used, ensuring a strictly controlled and unbiased comparison across optimizers.
+
+A uniform training protocol was applied to all runs, including identical batch size, warm-up and cosine decay learning rate schedule, gradient clipping strategy, and validation split.
+Our optimization methods were integrated into the NanoGPT training loop without modifying the model architecture or preprocessing pipeline.
+Model quality was primarily evaluated using validation loss tracked over 200 training steps.
+For all one-dimensional layers, AdamW was used with a learning rate of $1\!\times\!10^{-3}$.
+Since momentum exhibited negligible influence on training dynamics, it was held constant across all experiments.
+A comparative analysis of optimization performance is provided in Figure~\ref{fig:nanogptftexps}.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{figs/nanogpt/optimizers_comparison.pdf}
+    \caption{Comparison of Muon, F-Muon, and S-Muon across a range of learning rates. Stars denote the best learning rate identified for each optimizer.l validation loss.}
+    \label{fig:nanogptftexps}
+\end{figure}
+
+Figure~\ref{fig:nanogptftexps2} presents the train and validation loss curves at the optimal learning rate for each optimizer.
+While F-Muon and S-Muon maintain consistent behavior, vanilla Muon attains the lowest loss overall.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{figs/nanogpt/Nanogpt finetune results.pdf}
+    \caption{Trai and validation loss at the optimal learning rate for each optimizer.
+The standard Muon method achieves the lowest loss, while F-Muon and S-Muon remain stable but exhibit slightly inferior performance.}
+    \label{fig:nanogptftexps2}
+\end{figure}
+
 \section{Algorithms for Vectors $\rightarrow$ Algorithms for Matrices}
 
-One cannot but notice the similarity between the updates LMO optimizers in the vector $l_p$ and the matrix Schatten $S_p$ norms, which is illustrated by \cref{tbl:mat_vs_vec_lmo}.
-However, the parallels are not limited to this similarity and can be traced even in the empirical observations.
-SignSGD is very close to Adam, as noted in \citet{bernstein2024oldoptimizernewnorm}, and both Adam and Muon perform well in training large models.
+One cannot but notice the similarity between the updates of LMO optimizers in the vector $l_p$ and the matrix Schatten $S_p$ norms, which is illustrated by \cref{tbl:mat_vs_vec_lmo}.
+The parallels are not limited to mere similarity of the updates and can be traced even in the empirical observations.
+SignSGD is very close to Adam, as noted in \citet{bernstein2024oldoptimizernewnorm}, and both Adam and Muon perform well in training large models \citep{zhao2025deconstructing, liu2025muon}.
 NSGD is the same for both matrix and vector cases.
@@
 			\hline
-			Method                             & lmo constraint set $\mathcal D$ & lmo                                                         & Reference                            \\
+			Algorithm                             & lmo constraint set $\mathcal D$ & lmo                                                         & Reference                            \\
 			\hline
@@
 \section{Related Work and Discussion}
-Since Muon \citep{jordan2024muon} is a very successful and popular optimizer for functions of weight matrices, a lot of reseach has been put into, first, further improving its performance and, second, explaining its success.
+Since Muon \citep{jordan2024muon} is a very efficient optimizer for functions of weight matrices, a lot of reseach has been put into, first, further improving its performance and, second, explaining its success.
 
@@
 IO suggested using the nuclear norm in the \citet{bernstein2024oldoptimizernewnorm} framework.
-DM supervised the project and helped with editing the article.
+DM presented the problem at the MIPT optimization course and helped with editing the article.
 IK suggested using composite norms (though not the ones that induce F-Fanions and S-Fanions) and refactored \cref{sec:conic_combo}.
-NK suggested Lanczos algorithm as the most precise means to compute Fanions' updates, conducted experiments to prove it, and wrote ~\cref{sec:matrix-side}.
-AV tested Neon on the finetuning of NanoGPT on Tiny Stories.
-All other work was done by AK.
+NK suggested Lanczos algorithm as the most precise means to compute Fanions' updates, conducted experiments to prove this, and wrote ~\cref{sec:matrix-side}.
+AV conducted the finetuning of NanoGPT on Tiny Stories and wrote ~\cref{subsec:finetune}.
+All other work was done by AK: creating Fanions, F-Fanions, S-Fanions, experimenting on the Linear Least Squares, CIFAR-airbench and NanoGPT pretrain, and writing the manuscript.
 
@@
 		\includegraphics[width=\linewidth]{figs/norms/KyFan.pdf}
-		\caption{Ky Fan 2-norm}
+		\caption{The Ky Fan $2$-norm}
 		\label{fig:kyfan}
@@
 		\includegraphics[width=\linewidth]{figs/norms/KyFanDual.pdf}
-		\caption{Dual to Ky Fan 2-norm}
+		\caption{Dual to the Ky Fan $2$-norm}
 		\label{fig:kyfandual}
@@
 	  \hline
-	  Method & lr  & momentum           & Iterations to 0.01 loss \\
+	  Algorithm & \texttt{lr}  & \texttt{momentum}  & Iterations to 0.001 loss \\
 	  \hline
-	  NSGD & 0.25 & 0.9 & 350 \\
+	  Muon & 0.007 & 0.5 & 1060 \\
 	  \hdashline
-	  SignSGD & 0.055 $\cdot$ 0.01 & 0.95 & 700 \\
+	  NSGD & 0.08 & 0.95 & 1020 \\
+	  F-Muon & 0.015 & 0.7 & 910 \\
 	  \hdashline
-	  Muon & 0.025 & 0.5 & 290 \\
-	  \hdashline
-	  F-Muon & 0.035 & 0.5 & 320 \\
-	  \hdashline
-	  S-Muon & 0.035 & 0.8 & 270 \\
+	  SignSGD & 0.016 $\times$ 0.01 & 0.95 & 2650 \\
+	  S-Muon & 0.011 & 0.9 & 890 \\
 	  \hline
@@
         \centering
-        \includegraphics[width=\linewidth]{figs/simple_lls/gradient_spectral_norm_vs_iteration_500x500.pdf}
+        \includegraphics[width=\linewidth]{figs/simple_lls/gradient_spectral_norm_vs_iteration_long_run.pdf}
         \caption{The spectral norm of the gradient}
@@
         \centering
-        \includegraphics[width=\linewidth]{figs/simple_lls/gradient_nuclear_norm_vs_iteration_500x500.pdf}
+        \includegraphics[width=\linewidth]{figs/simple_lls/gradient_nuclear_norm_vs_iteration_long_run.pdf}
         \caption{The nuclear norm of the gradient}
@@
         \centering
-        \includegraphics[width=\linewidth]{figs/simple_lls/loss_vs_time_500x500.pdf}
+        \includegraphics[width=\linewidth]{figs/simple_lls/loss_vs_time_long_run.pdf}
         \caption{The loss over time}
@@
 
+\section{Under the hood of CNN on CIFAR-10}
+Most bounds for Muon and other LMO-algorithms are given for the norm of the gradient \citep{kovalev2025understanding, pethick2025training, riabinin2025gluon,kovalev2025noneuclideansgdstructuredoptimization}.
+Accordingly, one is immediately interested to measure those norms on a real deep learning problem.
+
+We compare Muon, F-Muon, S-Muon, NSGD, SignSGD, Neon, and F-Neon on CIFAR-airbench.
+To preserve the algorithms, we do not normalize the weights of the network each iteration.
+The parameters are in the table.
+
+\begin{table}[ht]
+	\centering
+	\begin{tabular}{|c|c|c|c|}
+	  \hline
+	  Method & lr  & momentum   & val\_accuracy, \% \\
+	  \hline
+	  NSGD & ? & ? & ? \\
+	  \hdashline
+	  SignSGD & ? $\cdot$ ? & ? & ? \\
+	  \hdashline
+	  Muon & 0.24 & 0.63 & $94.02 \pm 0.10$ \\
+	  \hdashline
+	  F-Muon & 0.42 & 0.63 & $94.04 \pm 0.13$\\
+	  \hdashline
+	  S-Muon & 0.42 & 0.63 & $94.03 \pm 0.13$ \\
+	  \hline
+	\end{tabular}
+  \caption{Parameters for CIFAR-airbench. \texttt{sign\_lr\_mult}=0.003}
+  \label{tbl:cifar_lrs}
+  \end{table}
+
 \section{Technical details of the experiments}

### 2025-11-26 conf_article/neon.tex
Added arrows
Fixed accuracies

@@
 Ivan Oseledets \\
-AIRI, Skoltech \\
+AIRI, Skoltech, \\
+Marchuk Institute of Numerical Mathematics \\
 \texttt{i.oseledets@skoltech.ru}
@@
 Most notably, our experiments on real-world tasks demonstrate that the choice of underlying matrix norm is remarkably flexible.
-On CIFAR-10 airbench, properly-tuned F-Muon and S-Muon achieve $94.04 \pm 0.13\%$ and $94.03 \pm 0.13\%$ accuracy, matching Muon's $94.01 \pm 0.13\%$ performance.
+On CIFAR-10 airbench, properly-tuned F-Muon and S-Muon achieve $94.02 \pm 0.13\%$ and $94.03 \pm 0.13\%$ accuracy, matching Muon's $94.01 \pm 0.13\%$ performance.
 On NanoGPT pre-training, F-Muon achieves 3.281 cross-entropy loss, while fully-tuned Muon achieves 3.279.
@@
         \centering
-        \includegraphics[width=\linewidth]{figs/simple_lls/loss_vs_iteration_long_run.pdf}
+        \includegraphics[width=\linewidth]{figs/simple_lls/loss_long_run_16_increase.pdf}
         \caption{The loss}
@@
 Then we tune F-Muon with $\alpha = 0.5$.
-Tuned parameters are {\tt lr=0.42(1 - step/total\_steps), momentum=0.63, nesterov = True}.
-Tuned F-Muon reaches $94.04 \pm 0.13\%$ after 8 epochs (averaged by 200 iterations), matching Muon's accuracy and variance.
+Tuned parameters are {\tt lr=0.4(1 - step/total\_steps), momentum=0.6, nesterov = True}.
+Tuned F-Muon reaches $94.02 \pm 0.13\%$ after 8 epochs (averaged by 200 iterations), matching Muon's accuracy and variance.
 Once again we measure \texttt{val\_accuracy}($\alpha$) curve (\cref{fig:fmuon_alphas}).
@@
 	  \hdashline
-	  Muon & 0.24 & 0.63 & $94.02 \pm 0.10$ \\
+	  Muon & 0.24 & 0.6 & $94.01 \pm 0.10$ \\
 	  \hdashline
-	  F-Muon & 0.42 & 0.63 & $94.04 \pm 0.13$\\
+	  F-Muon & 0.40 & 0.6 & $94.02 \pm 0.13$\\
 	  \hdashline

### 2025-11-26 article_final/main.tex
started transfering article to the new style

%\documentclass[wcp,gray]{jmlr} % test grayscale version
 %\documentclass[wcp]{jmlr}% former name JMLR W\&CP
\documentclass[pmlr]{jmlr}% new name PMLR (Proceedings of Machine Learning)

\RequirePackage{graphicx}
 % The following packages will be automatically loaded:
 % amsmath, amssymb, natbib, graphicx, url, algorithm2e
 \usepackage{booktabs}
 %\usepackage{rotating}% for sideways figures and tables
\usepackage{longtable}% for long tables

\makeatletter
\def\set@curr@file#1{\def\@curr@file{#1}} %temp workaround for 2019 latex release
\makeatother
\usepackage[load-configurations=version-1]{siunitx} % newer version

% change the arguments, as appropriate, in the following:
\jmlrvolume{tbd}
\jmlryear{2025}
\jmlrworkshop{International Conference on Computational Optimization}

\title[Ky Fan Norms for Matrix Optimization]{The Ky Fan Norms and Beyond: Dual Norms and Combinations for Matrix Optimization}

\author{\Name{Alexey Kravatskiy}
       \Email{kravtskii.aiu@phystech.edu}\\
       \addr Moscow Institute of Physics and Technology (MIPT)
       \AND
       \Name{Ivan Kozyrev}
       \Email{kozyrev.in@phystech.edu}\\
       \addr Moscow Institute of Physics and Technology (MIPT)\\
       Marchuk Institute of Numerical Mathematics
       \AND
       \Name{Alexander Vinogradov}
       \Email{vinogradov.am@phystech.edu}\\
       \addr Moscow Institute of Physics and Technology (MIPT)
       \AND
       \Name{Nikolai Kozlov}
       \Email{kozlov.na@phystech.edu}\\
       \addr Moscow Institute of Physics and Technology (MIPT)
       \AND
       \Name{Daniil Merkulov}
       \Email{daniil.merkulov@phystech.edu}\\
       \addr Moscow Institute of Physics and Technology (MIPT)\\
       Skoltech, HSE, AI4Science
       \AND
       \Name{Ivan Oseledets}
       \Email{i.oseledets@skoltech.ru}\\
       \addr AIRI, Skoltech\\
       Marchuk Institute of Numerical Mathematics}

\begin{document}

\maketitle

\begin{abstract}
	In this article, we explore the use of various matrix norms for optimizing functions of weight matrices, a crucial problem in training large language models. Moving beyond the spectral norm that underlies the Muon update, we leverage the duals to the Ky Fan $k$-norms to introduce a family of Muon-like algorithms we name \emph{Fanions}, which happen to be similar to Dion. Then working with the duals of convex combinations of the Ky Fan $k$-norms and the Frobenius norm or the $l_\infty$ norm, we construct the families of \emph{F-Fanions} and \emph{S-Fanions} respectively. Their most prominent members are \emph{F-Muon} and \emph{S-Muon}. We complement our theoretical analysis with an extensive empirical study of the algorithms across a wide range of tasks and settings, from which it follows that F-Muon and S-Muon are always on par with Muon, while on fine-tuning of NanoGPT and synthetic linear least squares they are even better than vanilla Muon optimizer.
\end{abstract}

\end{document}

### 2025-11-26 code/cifar_norms/cifar_norm_plotter.py
CIFAR under hood done

@@ -102,7 +102,11 @@ def plot_norm_comparison(
             ax.set_ylabel(_format_metric_name(metric_col), fontsize=12)
         
         # ax.set_title(_format_metric_name(metric_col), fontsize=14, fontweight='bold') - we need no title
-        ax.legend(fontsize=10)
+        # Sort legend alphabetically
+        handles, labels = ax.get_legend_handles_labels()
+        sorted_pairs = sorted(zip(labels, handles), key=lambda x: x[0])
+        sorted_labels, sorted_handles = zip(*sorted_pairs) if sorted_pairs else ([], [])
+        ax.legend(sorted_handles, sorted_labels, fontsize=10)
         ax.grid(True, alpha=0.3)
         
         # Save figure
@@ -187,7 +191,7 @@ import re
 from datetime import datetime
 
 
-def collect_grad_norm_files(folder_path, after_dt=datetime(2025, 11, 26, 10, 50, 0)):
+def collect_grad_norm_files(folder_path, after_dt=datetime(2025, 11, 27, 0, 50, 0)):
     """
     Collect grad_norm CSV logs whose timestamp is strictly after `after_dt`.
 

### 2025-11-26 article_final/main.tex

@@
        \AND
+       \Name{Nikolai Kozlov}
+       \Email{kozlov.na@phystech.edu}\\
+       \addr Moscow Institute of Physics and Technology (MIPT)
+       \AND
        \Name{Alexander Vinogradov}
        \Email{vinogradov.am@phystech.edu}\\
-       \addr Moscow Institute of Physics and Technology (MIPT)
-       \AND
-       \Name{Nikolai Kozlov}
-       \Email{kozlov.na@phystech.edu}\\
        \addr Moscow Institute of Physics and Technology (MIPT)
@@
 
+\section{Introduction}
+
+Minimizing loss functions in unprecedentedly high-dimensional spaces has recently become an integral and crucial part in training large language models.
+Hence, new scalable, time- and memory-efficient algorithms have been demanded.
+Besides well-known Adam \citep{kingma2014adam} and AdamW \citep{loshchilov2017decoupled}, recently proposed Muon \citep{jordan2024muon} has shown promising results on training very large models \citep{liu2025muon}.
+Its key difference from Adam and AdamW is that it has been constructed specifically for optimizing functions of weight matrices, which are common in deep learning.
+
+That is what can be said from a practical point of view.
+From a theoretical perspective, a key innovation of Muon was its principled derivation of the update rule, which emerged as the solution to an optimization problem constrained by the RMS-to-RMS norm (scaled version of the spectral norm) \citep{bernstein2025deriving}.
+
+Motivated by the success of Muon, many generalizations and variations of it were proposed.
+Among the notable ones are Scion \citep{pethick2025training}, Dion \citep{ahn2025dion} and Gluon \citep{riabinin2025gluon}.
+Those works try to explain Muon's efficiency and establish convergence bounds.
+One central question, however, remains unanswered:
+
+\textit{In deriving Muon's update step, why should one constrain by the spectral or any other operator norm? How would alternative norms affect performance and computational cost?}
+
+In this article, we tackle this question by actually showing that there are many viable non-operator norms.
+We leverage the family of norms dual to Ky Fan $k$-norms to derive a new family of \textbf{Fanions}, algorithms with low-rank updates.
+This approach theoretically explains the backbone of Dion's update \citep{ahn2025dion} and generalizes the memory-motivated application of the nuclear norm to Sharpness-Aware Minimization \citep{pethick2025sam}.
+As it was done with Muon, we come up with an effective procedure for computing Fanions' updates.
+Lanczos algorithm, which is described and compared with its competitors in \sectionref{sec:matrix-side}, is the most operation-effective algorithm, which, however, for now lacks an effective GPU- and PyTorch-friendly implementation.
+
+Working with duals to conic combinations of dual norms, we construct the families of \textbf{F-Fanions} and \textbf{S-Fanions}, which are hybrids of Muon and NormalizedSGD and SignSGD, respectively.
+
+Then we compare the performances of the algorithm families on various model and real-world problems:
+\begin{itemize}
+	\item Synthetic least squares experiment \sectionref{subsec:lls_exps}
+	\item CIFAR-10 airbench \citep{cifar2023airbench}
+	\item Pre-training NanoGPT on FineWeb dataset \citep{modded_nanogpt_2024}
+	\item Fine-tuning NanoGPT on Tiny Stories \citep{eldan2023tinystoriessmalllanguagemodels}
+\end{itemize}
+
+Our experiments reveal important insights into the role of matrix norms in optimization.
+First, we show on the example of Neon, the one-rank Fanion, that not every LMO-based algorithm is effective, despite the same asymptotics in the general bounds of \citet{kovalev2025understanding} and \citet{riabinin2025gluon}.
+This suggests that existing theoretical guarantees should be reworked to explain empirical performance.
+
+Most notably, our experiments on real-world tasks demonstrate that the choice of underlying matrix norm is remarkably flexible.
+On CIFAR-10 airbench, properly-tuned F-Muon and S-Muon achieve $94.02 \pm 0.13\%$ and $94.03 \pm 0.13\%$ accuracy, matching Muon's $94.01 \pm 0.13\%$ performance.
+On NanoGPT pre-training, F-Muon achieves 3.281 cross-entropy loss, while fully-tuned Muon achieves 3.279.
+Finally, S-Muon matches Muon on fine-tuning of NanoGPT on Tiny Stories, while F-Muon is far more resistant to the learning rate choice than Muon.
+These results show that Muon-like algorithms can maintain competitive performance even when the underlying norm constraint is significantly modified, answering affirmatively the central question posed above.
+Moreover, the tools from \sectionref{sec:conic_combo} give the researchers an unheard-of flexibility in designing algorithms that do not have to be modifications of Muon.
+
+\bibliography{icomp2024_conference}
+
 \end{document}

### 2025-11-26 article_final/main.tex
in the process of editing preliminaries

@@
 \usepackage[load-configurations=version-1]{siunitx} % newer version
+
+% Load required packages for custom math commands
+\usepackage{mathtools}
+\usepackage{bm}
+
+% Custom math commands
+% Blackboard bold
+\newcommand{\R}{\mathbb{R}}
+
+% Bold matrices
+\newcommand{\mA}{\bm{A}}
+\newcommand{\mB}{\bm{B}}
+\newcommand{\mD}{\bm{D}}
+\newcommand{\mF}{\bm{F}}
+\newcommand{\mG}{\bm{G}}
+\newcommand{\mM}{\bm{M}}
+\newcommand{\mN}{\bm{N}}
+\newcommand{\mS}{\bm{S}}
+\newcommand{\mU}{\bm{U}}
+\newcommand{\mV}{\bm{V}}
+\newcommand{\mX}{\bm{X}}
+\newcommand{\mSigma}{\bm{\Sigma}}
+
+% Calligraphic letters
+\newcommand{\cB}{\mathcal{B}}
+\newcommand{\cS}{\mathcal{S}}
+
+% Norms and operators
+\newcommand{\norm}[1]{\lVert #1\rVert}
+\DeclarePairedDelimiter{\normf}{\|}{\|_\mathrm{F}}
+\DeclarePairedDelimiter{\norms}{\|}{\|_{\mathrm{2}}}
+\DeclarePairedDelimiter{\normc}{\|}{\|_{\mathrm{C}}}
+
+% Inner product
+\newcommand{\<}{\langle}
+\renewcommand{\>}{\rangle}
+
+% Math operators
+\DeclareMathOperator*{\argmin}{arg\,min}
+\DeclareMathOperator{\sign}{sign}
+\DeclareMathOperator{\tr}{tr}
+\DeclareMathOperator{\diag}{diag}
+
+% Custom shortcuts
+\renewcommand{\epsilon}{\varepsilon}
+\newcommand{\Rmn}{\R^{m\times n}}
 
@@
 
+\section{Preliminaries: Linear Minimization Oracle Framework}
+
+Training a neural network is essentially an optimization of a function of several weight matrices and a few vectors.
+Let us start by considering the problem of minimizing a differentiable function of a {\it single} matrix:
+\begin{equation}\label{eq:matrix_problem}
+	F(\cdot)\colon \Rmn \to \R\,,\qquad F(\mX) \to \min_{\mX \in \Rmn}\,
+\end{equation}
+We equip the matrix space $\Rmn$ with a standard dot product $\<\cdot, \cdot> \to \R$ and a norm $\norm{\cdot}\colon \Rmn \to \R_+$, which does not have to coincide with the Frobenius norm $\normf{\cdot}$.
+The dual norm $\norm{\cdot}^\dagger\colon \Rmn \to \R_+$ that is associated with $\norm{\cdot}$ is defined as
+\begin{equation}\label{eq:dual_norm}
+	\norm{\mG}^\dagger = \sup_{\mD \in \Rmn\,:\norm{\mD}\leq 1} \<\mG,\mD>\,.
+\end{equation}
+
+Such problems can be solved with an iterative algorithm based on the Linear Minimization Oracle (LMO):
+\begin{equation}\label{eq:lmo}
+	\mathrm{LMO}(\mM^k) \in \argmin_{\mD \in \cS} \<\mM^k, \mD>\,,
+\end{equation}
+where $\mM^k$ is a gradient (or a momentum) of $F$ in $\mX^k$ and $\cS \subset \Rmn$ is some set. The update of the algorithm is defined as follows:
+\begin{equation}\label{eq:simple_update}
+	\mX^{k + 1} = \mX^k + \gamma_k\mathrm{LMO}\left(\mM^k\right)\,.
+\end{equation}
+
+We are particularly interested in the case when $\cS$ is a unit ball in the norm $\norm{\cdot}$:
+
+$$
+\cS = \cB_{\norm{\cdot}} = \{ \mD \in \Rmn \mid \norm{\mD} \leq 1 \}\,.
+$$
+In this case,
+$$
+\argmin_{\mD \in \cS} \<\mM^k, \mD> = - \{ \mD \in \cB_1 \mid \<\mM^k, \mD> = \norm{\mM^k}^\dagger\}\,,
+$$
+and update for $\mX^k$ in \cref{eq:simple_update} simplifies to
+\begin{equation}\label{eq:our_update}
+	\mX^{k+1} = \mX^{k} - \gamma_k \{\mD \in \cB_1 \mid \<\mM^k, \mD> = \norm{\mM^k}^\dagger\}\,.
+\end{equation}
+
+Using this formula it is easy to compute updates for algorithms induced by various norms $\norm{\cdot}$.
+
+For example, when the norm $\norm{\cdot}$ is the Frobenius norm $\normf{\cdot}$, \cref{eq:our_update} turns into
+\begin{equation}\label{eq:nsgd_update}
+	\mX^{k+1} = \mX^{k} - \eta_k \frac{\mM^k}{\normf{\mM^k}}\,,
+\end{equation}
+which recovers Normalized SGD (NSGD).
+
+And when the norm is the spectral norm $\norms{\cdot}$, we get
+\begin{equation}\label{eq:muon_update}
+	\mX^{k+1} = \mX^{k} - \eta_k U V^\top\,,
+\end{equation}
+which is Muon without the $\sqrt{m/n}$ factor.
+Here, $\mM^k = \mU \mSigma \mV^\top$ is the Singular Value Decomposition (SVD) of $\mM^k$ ($\mU = [u_1, u_2, \dots, u_r]$, $\mSigma = \diag(\sigma_1, \space \sigma_2, \dots, \sigma_r)$, and $\mV = [v_1, v_2, \dots, v_r]$).
+Muon can be recovered by taking the RMS-to-RMS operator norm $\sqrt{\frac{n}{m}}\norms{\cdot}$.
+
+When the norm is the Chebyshev norm $\normc{\cdot}$, we get
+\begin{equation}\label{eq:signsgd_update}
+	\mX^{k+1} = \mX^{k} - \eta_k \sign(\mM^k)\,,
+\end{equation}
+which recovers SignSGD~\citep{bernstein2018signsgd}.
+Here, $\sign(\mM^k)$ denotes the element-wise sign function.
+SignSGD is particularly notable for its communication efficiency in distributed training, as it compresses gradients to 1-bit per parameter.
+
 \bibliography{icomp2024_conference}

### 2025-11-26 conf_article/neon.tex
adopted pdf from remote origin
Minor fixes

@@
 Most notably, our experiments on real-world tasks demonstrate that the choice of underlying matrix norm is remarkably flexible.
-On CIFAR-10 airbench, properly-tuned F-Muon and S-Muon achieve $94.02 \pm 0.13\%$ and $94.03 \pm 0.13\%$ accuracy, matching Muon's $94.01 \pm 0.13\%$ performance.
+On CIFAR-10 airbench, properly-tuned F-Muon and S-Muon achieve $94.01 \pm 0.13\%$ and $94.03 \pm 0.13\%$ accuracy, matching Muon's $94.01 \pm 0.13\%$ performance.
 On NanoGPT pre-training, F-Muon achieves 3.281 cross-entropy loss, while fully-tuned Muon achieves 3.279.
@@
 We compare the algorithms on the CIFAR-10 airbench \citep{cifar2023airbench}.
-First, to test the impact of $\alpha$, we run F-Muon for different $\alpha$ with the same {\tt lr=0.24(1 - step/total\_steps), momentum=0.6, nesterov=True} and weight normalization, as have been finetuned by Keller Jordan, 10 repetitions for each $\alpha$.
+First, to test the impact of $\alpha$, we run F-Muon for different $\alpha$ with the same {\tt lr=0.24(1 - step/total\_steps), momentum=0.65, nesterov=True} and weight normalization, as have been finetuned by Keller Jordan, 10 repetitions for each $\alpha$.
 We record the accuracy after 8 epochs of training (\cref{fig:muon_alphas}).
@@
 We compare Muon, F-Muon, S-Muon, NSGD, SignSGD, Neon, and F-Neon on CIFAR-airbench.
+NSGD and SignSGD are not heavily tuned.
 To preserve the algorithms, we do not normalize the weights of the network each iteration.
@@
 	  \hline
-	  Method & lr  & momentum   & val\_accuracy, \% \\
+	  Method & \texttt{lr}  & \texttt{momentum}  & val\_accuracy, \% \\
 	  \hline
-	  NSGD & ? & ? & ? \\
+	  NSGD & 0.5 & 0.95 & $91.6 \pm 0.52$ \\
 	  \hdashline
-	  SignSGD & ? $\cdot$ ? & ? & ? \\
+	  SignSGD & 0.003 & 0.95 & $91.54 \pm 0.26$ \\
 	  \hdashline
@@
 	  \hdashline
-	  F-Muon & 0.40 & 0.6 & $94.02 \pm 0.13$\\
+	  F-Muon & 0.40 & 0.6 & $94.01 \pm 0.13$\\
 	  \hdashline

### 2025-11-26 article_final/main.tex
added preliminaries

@@
 
-% Inner product
-\newcommand{\<}{\langle}
-\renewcommand{\>}{\rangle}
+% Inner product notation
+\def\<#1,#2>{\langle #1,#2\rangle}
 
@@
 Training a neural network is essentially an optimization of a function of several weight matrices and a few vectors.
-Let us start by considering the problem of minimizing a differentiable function of a {\it single} matrix:
+Let us start by considering the problem of minimizing a differentiable function of a \emph{single} matrix:
 \begin{equation}\label{eq:matrix_problem}
@@
 
-$$
+\[
 \cS = \cB_{\norm{\cdot}} = \{ \mD \in \Rmn \mid \norm{\mD} \leq 1 \}\,.
-$$
+\]
 In this case,
-$$
+\[
 \argmin_{\mD \in \cS} \<\mM^k, \mD> = - \{ \mD \in \cB_1 \mid \<\mM^k, \mD> = \norm{\mM^k}^\dagger\}\,,
-$$
-and update for $\mX^k$ in \cref{eq:simple_update} simplifies to
+\]
+and update for $\mX^k$ in \equationref{eq:simple_update} simplifies to
 \begin{equation}\label{eq:our_update}
@@
 
-For example, when the norm $\norm{\cdot}$ is the Frobenius norm $\normf{\cdot}$, \cref{eq:our_update} turns into
+For example, when the norm $\norm{\cdot}$ is the Frobenius norm $\normf{\cdot}$, \equationref{eq:our_update} turns into
 \begin{equation}\label{eq:nsgd_update}
@@
 \begin{equation}\label{eq:muon_update}
-	\mX^{k+1} = \mX^{k} - \eta_k U V^\top\,,
+	\mX^{k+1} = \mX^{k} - \eta_k \mU \mV^\top\,,
 \end{equation}
 which is Muon without the $\sqrt{m/n}$ factor.
-Here, $\mM^k = \mU \mSigma \mV^\top$ is the Singular Value Decomposition (SVD) of $\mM^k$ ($\mU = [u_1, u_2, \dots, u_r]$, $\mSigma = \diag(\sigma_1, \space \sigma_2, \dots, \sigma_r)$, and $\mV = [v_1, v_2, \dots, v_r]$).
+Here, $\mM^k = \mU \mSigma \mV^\top$ is the Singular Value Decomposition (SVD) of $\mM^k$ ($\mU = [u_1, u_2, \dots, u_r]$, $\mSigma = \diag(\sigma_1, \sigma_2, \dots, \sigma_r)$, and $\mV = [v_1, v_2, \dots, v_r]$).
 Muon can be recovered by taking the RMS-to-RMS operator norm $\sqrt{\frac{n}{m}}\norms{\cdot}$.

### 2025-11-26 conf_article/neon.tex
Added train and val accs

@@
 \end{equation}
-We name the derived family of algorithm \emph{F-Fanions}.
+We name the derived family of algorithms \emph{F-Fanions}.
 
@@
 
-Experiments were conducted on a single NVIDIA RTX 4090 (24GB) GPU using PyTorch~2.8 with CUDA~12.8 and cuDNN~9.0, running on a workstation equipped with an Intel Core i9-14900KS CPU and 128 GB of RAM.
-No distributed or mixed-hardware training setups were used, ensuring a strictly controlled and unbiased comparison across optimizers.
-
-A uniform training protocol was applied to all runs, including identical batch size, warm-up and cosine decay learning rate schedule, gradient clipping strategy, and validation split.
-Our optimization methods were integrated into the NanoGPT training loop without modifying the model architecture or preprocessing pipeline.
-Model quality was primarily evaluated using validation loss tracked over 200 training steps.
 For all one-dimensional layers, AdamW was used with a learning rate of $1\!\times\!10^{-3}$.
@@
 
-%\section{Updates derivations}
-% we prove a general lemma instead
-%Proof of \cref{lemma:nsgd_muon_update}:
-%Since $\norm{\cdot}^\dagger = \normfstar{\cdot}^{\dagger\dagger} = \normfstar{\cdot}$, the goal is to reach $\normfstar{\mM^k} = \alpha \tr{\mSigma} + (1-\alpha) \normf{\mM^k}$.
-
-%Let us note that $\Delta = \alpha \mU \mV^\top + (1-\alpha) \frac{\mM^k}{\normf{\mM^k}}$ delivers this value.
-Indeed, by the trace property, $\<\mM^k, \Delta> =\<\mU \mSigma \mV^\top, \alpha \mU \mV^\top + (1-\alpha) \frac{\mU \mSigma \mV^\top}{\normf{\mM^k}}> = \alpha \tr \mSigma + (1-\alpha) \normf{\mM^k} = \normfstar{\mM^k}$, which completes the proof.
-
-% we prove a general lemma instead
-%Proof of \cref{lemma:nsgd_neon_update}:
-%Since $\norm{\cdot}^\dagger = \normftwo{\cdot}^{\dagger\dagger} = \normftwo{\cdot}$, the goal is to reach $\normftwo{\mM^k} = \alpha \sigma_1 + (1-\alpha) \normf{\mM^k}$.
-
-% Let us note that $\Delta = \alpha (u_1 v_1^\top) + (1-\alpha) \frac{\mM^k}{\normf{\mM^k}}$ delivers this value.
-Indeed, by the trace property and singular vectors orthogonality, $\<\mM^k, \Delta> =\<\mU \mSigma \mV^\top, \alpha u_1 v_1^\top + (1-\alpha) \frac{\mU \mSigma \mV^\top}{\normf{\mM^k}}> = \alpha \tr \diag({\sigma_1, 0, \dots, 0}) + (1-\alpha) \normf{\mM^k} = \normftwo{\mM^k}$, which completes the proof.
-
 \section{More data for Linear Least Squares}
@@
 Most bounds for Muon and other LMO-algorithms are given for the norm of the gradient \citep{kovalev2025understanding, pethick2025training, riabinin2025gluon,kovalev2025noneuclideansgdstructuredoptimization}.
-Accordingly, one is immediately interested to measure those norms on a real deep learning problem.
-
-We compare Muon, F-Muon, S-Muon, NSGD, SignSGD, Neon, and F-Neon on CIFAR-airbench.
-NSGD and SignSGD are not heavily tuned.
-To preserve the algorithms, we do not normalize the weights of the network each iteration.
-The parameters are in the table.
+Accordingly, it is natural to measure those norms on a real deep learning problem.
 
@@
 	  Muon & 0.24 & 0.6 & $94.01 \pm 0.10$ \\
-	  \hdashline
 	  F-Muon & 0.40 & 0.6 & $94.01 \pm 0.13$\\
@@
 	  S-Muon & 0.42 & 0.63 & $94.03 \pm 0.13$ \\
+	  \hdashline
+	  Neon & 0.24 & 0.6 & $69.8 \pm 0.5$ \\
+	  F-Neon & 0.40 & 0.6 & $87.15 \pm 0.24$\\
+	  \hdashline
+	  Fanion-5 & 0.24 & 0.6 & $80.69 \pm 1.25$ \\
+	  F-Fanion-5 & 0.40 & 0.6 & $86.66 \pm 0.65$\\
 	  \hline
 	\end{tabular}
-  \caption{Parameters for CIFAR-airbench. \texttt{sign\_lr\_mult}=0.003}
+  \caption{Parameters for CIFAR-airbench. \texttt{sign\_lr\_mult} for S-Muon is 0.003}
   \label{tbl:cifar_lrs}
@@
 
+We compare Muon, F-Muon, S-Muon, NSGD, SignSGD, Neon, and F-Neon, Fanion-5, and F-Fanion-5 on CIFAR-airbench.
+NSGD and SignSGD are not heavily tuned.
+The hyperparameters of Neon, F-Neon, Fanion-5, and F-Fanion-5 are taken from Muon and F-Muon.
+The parameters are in the table \cref{tbl:cifar_lrs}.
+val\_accuracy is for the the weighted variant.
+However, to preserve the algorithms, we do not normalize the weights of the network each iteration.
+
+We log train and validation accuracy, and nuclear, gradient, and spectral norms of the gradients of conv1.weight and conv2.weight for all the layers.
+
+\begin{figure}[h!]
+    \centering
+    \begin{subfigure}[t]{0.7\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{figs/cifar_logger/train_acc.pdf}
+        \caption{Train accuracy}
+		\label{fig:train_acc_cifar}
+    \end{subfigure}
+    \vspace{1em} % optional: add vertical spacing between subfigures
+    \begin{subfigure}[t]{0.7\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{figs/cifar_logger/val_acc.pdf}
+        \caption{Validation accuracy}
+		\label{fig:val_acc_cifar}
+    \end{subfigure}
+    \caption{Different optimizers on CIFAR airbench without weight normalization}
+	\label{fig:cifar_accs}
+\end{figure}
+
 \section{Technical details of the experiments}
-We compared TRLand with other methods in Google Colab. We used RTX A4000 for CIFAR airbench.
+\textbf{CIFAR airbench} We used NVIDIA RTX A4000.
+
+\textbf{NanoGPT pretrain} We used the standard setting of 8 $\times$ H100 as documented in \citet{modded_nanogpt_2024}.
+
+\textbf{NanoGPT Fine-tuning.} Experiments were conducted on a single NVIDIA RTX 4090 (24GB) GPU using PyTorch~2.8 with CUDA~12.8 and cuDNN~9.0, running on a workstation equipped with an Intel Core i9-14900KS CPU and 128 GB of RAM.
+No distributed or mixed-hardware training setups were used, ensuring a strictly controlled and unbiased comparison across optimizers.
+
+A uniform training protocol was applied to all runs, including identical batch size, warm-up and cosine decay learning rate schedule, gradient clipping strategy, and validation split.
+Our optimization methods were integrated into the NanoGPT training loop without modifying the model architecture or preprocessing pipeline.
+Model quality was primarily evaluated using validation loss tracked over 200 training steps.
 

### 2025-11-26 article_final/main.tex

@@
 
+% Vectors (note: \vec is already defined by jmlr class)
+% We define shortcuts for commonly used vectors
+\newcommand{\vu}{\vec{u}}
+\newcommand{\vv}{\vec{v}}
+
 % Calligraphic letters
@@
 \DeclarePairedDelimiter{\normc}{\|}{\|_{\mathrm{C}}}
+\DeclarePairedDelimiter{\normn}{\|}{\|_{*}}
+\newcommand{\normkfk}[1]{\|#1\|_{\text{KF-}k}}
 
@@
 
-Using this formula it is easy to compute updates for algorithms induced by various norms $\norm{\cdot}$.
+Using this formula it is easy to compute updates for algorithms induced by various norms $\norm{\cdot}$:
 
-For example, when the norm $\norm{\cdot}$ is the Frobenius norm $\normf{\cdot}$, \equationref{eq:our_update} turns into
+\paragraph{Frobenius norm and Normalized SGD}
+When the norm $\norm{\cdot}$ is the Frobenius norm $\normf{\cdot}$, \equationref{eq:our_update} turns into
 \begin{equation}\label{eq:nsgd_update}
@@
 
-And when the norm is the spectral norm $\norms{\cdot}$, we get
+\paragraph{Spectral norm and Muon}
+When the norm is the spectral norm $\norms{\cdot}$, we get
 \begin{equation}\label{eq:muon_update}
@@
 which is Muon without the $\sqrt{m/n}$ factor.
-Here, $\mM^k = \mU \mSigma \mV^\top$ is the Singular Value Decomposition (SVD) of $\mM^k$ ($\mU = [u_1, u_2, \dots, u_r]$, $\mSigma = \diag(\sigma_1, \sigma_2, \dots, \sigma_r)$, and $\mV = [v_1, v_2, \dots, v_r]$).
+Here, $\mM^k = \mU \mSigma \mV^\top$ is the Singular Value Decomposition (SVD) of $\mM^k$ ($\mU = [\vu_1, \vu_2, \dots, \vu_r]$, $\mSigma = \diag(\sigma_1, \sigma_2, \dots, \sigma_r)$, and $\mV = [\vv_1, \vv_2, \dots, \vv_r]$).
 Muon can be recovered by taking the RMS-to-RMS operator norm $\sqrt{\frac{n}{m}}\norms{\cdot}$.
 
+\paragraph{Chebyshev norm and SignSGD}
 When the norm is the Chebyshev norm $\normc{\cdot}$, we get
@@
 
+\section{Duals to Ky Fan Norms Instead of the Spectral Norm}
+
+\subsection{\texorpdfstring{$\normn{\mM^{k}}$ and Neon}{Neon}}
+After considering $\normf{\mM^k}$ and $\norms{\mM^k}$, it is natural to look at the nuclear norm $\normn{\mM^k}$.
+\begin{lemma}\label{lemma:neon_update}
+	When $\norm{\cdot} = \normn{\cdot}$, \equationref{eq:our_update} turns into:
+
+	\begin{equation}\label{eq:neon_update}
+		\mX^{k+1} = \mX^{k} - \eta_k \vu_1 \vv_1^\top
+	\end{equation}
+\end{lemma}
+\begin{proof}
+Since $\norm{\cdot}^\dagger = \norms{\cdot}$, the goal is to reach $\norms{\mM^k} = \sigma_1$.
+
+Let us note that $\Delta = \vu_1 \vv_1^\top$ delivers this value. Indeed, $\normn{\Delta}=1$ and by the trace property and orthogonality of the singular vectors,
+\[
+\<\mM^k, \Delta> =\<\mU \mSigma \mV^\top, \vu_1 \vv_1^\top> = \tr \diag({\sigma_1, 0, \dots, 0}) = \norms{\mM^k}\,,
+\]
+which completes the proof.
+\end{proof}
+
+We name the derived algorithm \emph{Neon}. In \sectionref{sec:matrix-side}, we will discuss how to compute the Neon's update.
+
+\subsection{\texorpdfstring{$\normkfk{\mM^k}^\dagger$ and Muon, Neon, and Centralized Dion without error feedback}{Muon, Neon, and Centralized Dion without error feedback}}
+
+Neon's and Muon's updates seem to be complete opposites: one has rank one, while the other is full-rank. It would be interesting to derive algorithms with updates of intermediate ranks.
+
+\subsubsection{Schatten norms}
+\citet{cesista2025schattenp} considered Schatten-$p$ norms: 
+\[\norm{\mM^k}_{S_p} = \left(\sum_{i=1}^{\min(m, n)}\sigma_i^p\right)^{1/p}\,,
+\]
+which produce the update
+\[
+\mX^{k+1} = \mX^k - \eta_k \mU \frac{\diag\left(\sigma_1^{q-1},\dots, \sigma_{\min(m,n)}^{q-1}\right)}{\left(\sum_{i=1}^{\min(m,n)}{\sigma_i^q}\right)^{\frac{q-1}{q}}}\mV^\top
+\]
+for $p$ and $q$ such that: $\frac{1}{p} + \frac{1}{q} = 1$.
+This formula recovers Neon when $p\rightarrow1$ provided that $\sigma_1 > \sigma_2$, which is true on real data; NSGD when $p=2$, and Muon when $p\rightarrow\infty$.
+
+However, Schatten norms do not fill the gap rank: indeed, when $p > 1$, the rank of the update is full, while when $p=1$ ($p \rightarrow 1$) it is one.
+Moreover, it is not clear how to calculate the update for $p \neq 1, 2, \infty$: it seems one has to know all $\sigma_i$ to compute the update, which makes the problem as hard as the full SVD.
+
+\subsubsection{Ky Fan norms}
+There is another family of matrix norms, which might help us: Ky Fan norms.
+For $k \in \{1, \dots, \min(m, n)\}$, the Ky Fan $k$-norm $\normkfk{\cdot}$ is $\sum_{i=1}^{k}\sigma_i$, i.e.
+the sum of the $k$ largest singular values of the matrix.
+Special cases of the Ky Fan $k$-norm are the Ky Fan $1$-norm, which is the spectral norm, and the Ky Fan $\min\{m, n\}$-norm, which is the nuclear norm.
+
+Let us derive the update for an arbitrary $k$.
+Since $\normkfk{\cdot}^\dagger = \max\{\frac{1}{k} \normn{\cdot}, \norms{\cdot}\}$ (see \citet{bhatia2013matrix}, p.
+96), the goal is to reach $\max\{\frac{1}{k}\sum_{i=1}^{\min(m,n)}\sigma_i, \sigma_1\}$.
+$\Delta = \vu_1 \vv_1^\top$ from Neon delivers $\sigma_1$, while $\Delta = \frac{1}{k}\sum_{i=1}^{\min(m,n)}\vu_i \vv_i^\top$ from Muon (with an extra factor of $1/k$) delivers $\frac{1}{k}\sum_{i=1}^{\min(m,n)}\sigma_i$.
+Thus, the update is either
+\[
+\mX^{t+1} = \mX^{t} - \eta_t \vu_1 \vv_1^\top\text{ or }\mX^{t+1} = \mX^{t} - \frac{\eta_t}{k}\mU \mV^\top\,,
+\]
+depending on which one better minimizes the linear function $L(\mX) \equiv F(\mX^t) + \mM^t(\mX - \mX^t)$.
+
+The rank gap is not closed by the Ky Fan norms because the obtained update is either one-rank or full-rank.
+
+\subsubsection{Duals to Ky Fan norms}
+While Schatten norms are closed under dualization and $\norm{\mM^k}_{S_p}^\dagger = \norm{\mM^k}_{S_q}$, for Ky Fan norms it is not generally the case.
+$k = 1$ and $k=\min(m, n)$ are exceptional: for $k=1$ the dual norm is $\normn{\cdot}$ with $k = \min(m, n)$ and for $k = \min(m, n)$ the dual norm is $\norms{\cdot}$ with $k=1$.
+For each other $k$, $\normkfk{\cdot} \neq \norm{\cdot}_{\text{KF-}k'}$ for any $k'$: $k'=\min(m, n)$ and $k'= 1$ correspond to the already discussed cases, while for other $k'$ one can change $\norm{\mM^k}_{\text{KF-}k'}$ by moving a small value between $\sigma_{k'}$ and $\sigma_{k'+1}$.
+During this change, $\norms{\mM^k}$ and $\normn{\mM^k}$ will remain constant, hence $\normkfk{\mM^k}$ will not change as well.
+
+\begin{lemma}\label{lemma:ky_fan_update}
+	When $\norm{\cdot} = \normkfk{\mM^t}^\dagger$, \equationref{eq:our_update} turns into:
+
+	\begin{equation}\label{eq:ky_fan_update}
+		\mX^{t+1} = \mX^{t} - \eta \sum_{i=1}^{k}\vu_i \vv_i^\top
+	\end{equation}
+\end{lemma}
+
+\begin{proof}
+Since $\norm{\cdot}^\dagger = \normkfk{\cdot}^{\dagger\dagger} = \normkfk{\cdot}$, the goal is to reach $\normkfk{\mM^t}$.
+
+Let us note that $\Delta = \sum_{i=1}^{k} \vu_i \vv_i^\top$ delivers the value. Indeed,
+\[
+\<\mM^t, \Delta> =\<\mU \mSigma \mV^\top, \sum_{i=1}^{k} \vu_i \vv_i^\top> = \sum_{i,j=1}^{r, k}\<\vu_i \sigma_i \vv_i^\top, \vu_j \vv_j^\top> = \sum_{i=1}^{k}\sigma_i = \normkfk{\mM^t}\,,
+\]
+which completes the proof.
+\end{proof}
+These updates constitute the family of \emph{Fanions}, LMO-based algorithms under the $\normkfk{\mM^t}^\dagger$ norms.
+The algorithm for a particular $k$ we will call \emph{Fanion-$k$}.
+Thus, Muon is a Fanion-$\min\{m,n\}$, while Neon is a Fanion-1.
+Moreover, the unsharded version of the rank-$r$ Dion (Algorithm 1 from \citet{ahn2025dion}) without the error feedback and without scaling of the update is actually a Fanion-$r$ (see \citet{pethick2025understandingdion}, where Dion is written down in a notation more similar to ours).
+
 \bibliography{icomp2024_conference}

### 2025-11-26 article_final/main.tex
added section 5

@@
 \usepackage{longtable}% for long tables
+\usepackage{arydshln}% for dashed lines in tables
 
@@
 \newcommand{\normkfk}[1]{\|#1\|_{\text{KF-}k}}
+\DeclarePairedDelimiter{\normfstar}{\|}{\|_\mathrm{F*}}
+\DeclarePairedDelimiter{\normftwo}{\|}{\|_\mathrm{F2}}
 
@@
 \begin{lemma}\label{lemma:neon_update}
-	When $\norm{\cdot} = \normn{\cdot}$, \equationref{eq:our_update} turns into:
-
+	When $\norm{\cdot} = \normn{\cdot}$, \equationref{eq:our_update} becomes
 	\begin{equation}\label{eq:neon_update}
-		\mX^{k+1} = \mX^{k} - \eta_k \vu_1 \vv_1^\top
+		\mX^{k+1} = \mX^{k} - \eta_k \vu_1 \vv_1^\top\,.
 	\end{equation}
@@
 Since $\norm{\cdot}^\dagger = \norms{\cdot}$, the goal is to reach $\norms{\mM^k} = \sigma_1$.
-
-Let us note that $\Delta = \vu_1 \vv_1^\top$ delivers this value. Indeed, $\normn{\Delta}=1$ and by the trace property and orthogonality of the singular vectors,
+Let us note that $\Delta = \vu_1 \vv_1^\top$ delivers this value.
+Indeed, $\normn{\Delta}=1$ and by the trace property and orthogonality of the singular vectors,
 \[
@@
 
+\section{Conic Combination of LMO-algorithms is an LMO-algorithm}
+\label{sec:conic_combo}
+
+Simply applying norm dual to the Ky Fan $k$-norm however, is not enough to provide family of algorithms diverse enough for our cause.
+Next we consider linear combinations of algorithms from LMO framework, which happen to be LMO-algorithms too, as we show in succeeding paragraphs.
+
+\subsection{General Case}
+First, we need a well-known fact mentioned, for example, in \citet[Table 1]{yu2012arithmetic}. For the sake of completeness, we provide a proof here.
+
+\begin{lemma}\label{lemma:dual_to_conv_comb}
+	Let $\norm{\cdot}_{(1)},\dots,\norm{\cdot}_{(n)}$ be norms on a finite-dimensional Euclidean space, and let $\alpha_1,\dots,\alpha_n$ be non-negative reals. Define
+	\[
+		\norm{\cdot} := \sum_{i = 1}^n \alpha_i \norm{\cdot}_{(i)}.
+	\]
+	Then the dual unit ball of $\norm{\cdot}$ satisfies
+	\[
+		\cB_{\norm{\cdot}^\dagger}
+		= \sum_{i = 1}^n \alpha_i \cB_{\norm{\cdot}_{(i)}^\dagger}
+	\]
+	where $\sum$ denotes the Minkowski sum and $\cB_{\norm{\cdot}_{(i)}^\dagger}$ is the unit ball of the dual norm $\norm{\cdot}_{(i)}^\dagger$.
+\end{lemma}
+
+\begin{proof}
+	Let us first prove the lemma for the case $n = 2$. Denote $f(x) = \alpha_1 \norm{x}_{(1)}$ and $g(x) = \alpha_2 \norm{x}_{(2)}$, such that $\norm{x} = f(x) + g(x)$. Recall two standard facts:
+	\begin{enumerate}
+		\item For any norm $\norm{\cdot}$ and $\lambda>0$,
+		      \[
+			      (\lambda \norm{\cdot})^*(y) =
+			      \sup_{x}\bigl( \langle y,x \rangle - \lambda \norm{x}\bigr)
+			      = \delta_{\lambda \cB_{\norm{\cdot}^\dagger}}(y)\,,
+		      \]
+		      i.e., the indicator function of the scaled dual ball.
+		\item The Fenchel conjugate of a sum satisfies
+		      \[
+			      (f+g)^*(y) = \inf_{u+v=y} \bigl(f^*(u) + g^*(v)\bigr)\,.
+		      \]
+	\end{enumerate}
+	Applying these to $f$ and $g$, we have
+	\[
+		f^*(u) = \delta_{\alpha_1 \cB_{\norm{\cdot}_{(1)}^\dagger}}(u)\,,
+		\quad
+		g^*(v) = \delta_{\alpha_2 \cB_{\norm{\cdot}_{(2)}^\dagger}}(v)\,.
+	\]
+	Thus,
+	\[
+		\norm{\cdot}^*(y)
+		= (f+g)^*(y)
+		= \inf_{u+v=y}
+		\bigl(
+		\delta_{\alpha_1 \cB_{\norm{\cdot}_{(1)}^\dagger}}(u)
+		+
+		\delta_{\alpha_2 \cB_{\norm{\cdot}_{(2)}^\dagger}}(v)
+		\bigr)
+		=
+		\delta_{\alpha_1 \cB_{\norm{\cdot}_{(1)}^\dagger}+\alpha_2 \cB_{\norm{\cdot}_{(2)}^\dagger}}(y).
+	\]
+	By definition, the conjugate of a norm is exactly the indicator of its dual unit ball:
+	\[
+	\norm{\cdot}^*(y) = \delta_{\cB_{\norm{\cdot}^\dagger}}(y)\,.
+	\]
+	Therefore, $\cB_{\norm{\cdot}^\dagger} = \alpha_1 \cB_{\norm{\cdot}_{(1)}^\dagger} + \alpha_2 \cB_{\norm{\cdot}_{(2)}^\dagger}$.
+
+	Now we prove the general case by induction. The base case ($n=2$) is already proven. Suppose that the assumption of the lemma holds for $n = k$. Then, for $n = k + 1$,
+	\[
+	\norm{x} = \sum_{i = 1}^k \alpha_i \norm{x}_{(i)} + \alpha_{k + 1}\norm{x}_{(k + 1)} = \norm{x}_{(1:k)} + \alpha_{k + 1}\norm{x}_{(k + 1)}\,.
+	\]
+	Applying the result for $n=2$ combined with the induction assumption, we obtain
+	\[
+	\cB_{\norm{\cdot}^\dagger} = \cB_{\norm{\cdot}_{(1:k)}^\dagger} + \alpha_{k + 1} \cB_{\norm{\cdot}_{(k + 1)}^\dagger} = \sum_{i = 1}^{k + 1} \alpha_i \cB_{\norm{\cdot}_{(i)}^\dagger}\,,
+	\]
+	which proves the lemma.
+\end{proof}
+
+\begin{lemma}\label{lemma:lin_comb_lmo}
+	Let $\norm{\cdot}_{(1)}, \dots, \norm{\cdot}_{(n)}$ be norms on a finite-dimensional Euclidean space, and let $\alpha_1, \dots, \alpha_n$ be non-negative reals.
+Consider Linear Minimization Oracles $\mathrm{LMO}_{1}, \dots, \mathrm{LMO}_{n}$, corresponding to the unit balls of these norms.
+Then, $\sum_{i = 1}^n \alpha_i \mathrm{LMO}_{i}$ is the LMO corresponding to the norm $\norm{\cdot}$ dual to the $\sum_{i = 1}^n \alpha_i \norm{\cdot}_{(i)}^\dagger$.
+\end{lemma}
+
+\begin{proof}
+	Using \lemmaref{lemma:dual_to_conv_comb} and the fact $\norm{\cdot}^{\dagger \dagger} = \norm{\cdot}$, we can obtain general form of the unit ball in the $\norm{\cdot}$ norm: $\cB_{\norm{\cdot}} = \sum_{i = 1}^n \alpha_i \cB_{\norm{\cdot}_{(i)}}$.
+Thus, the linear function minimization objective on a $\norm{\cdot}$ norm ball can be transformed as follows:
+	\[
+	\argmin_{\mD \in \cB_{\norm{\cdot}}}\<\mM, \mD> = \argmin_{\mD_1 \in \alpha_i \cB_{\norm{\cdot}_{(1)}},\dots,\mD_n \in \alpha_n\cB_{\norm{\cdot}_{(n)}}}\<\mM, \sum_{i = 1}^n\mD_i> = \sum_{i = 1}^n \argmin_{D_i \in \cB_{\norm{\cdot}_{(i)}}}\<\mM, \mD_i>\,,
+	\]
+	where the last summation denotes the Minkowski sum.
+This immediately implies $\sum_{i = 1}^n \alpha_i \mathrm{LMO}_{i} \in \argmin_{\mD \in \cB_{\norm{\cdot}}}\<\mM, \mD>$, which proves the claim of the lemma.
+\end{proof}
+
+Applying it to optimization algorithms, we obtain the following corollary.
+
+\begin{corollary}\label{corollary:lin_comb_alg}
+	Let there be a finite family of LMO based algorithms indexed by $i = 1,\dots,n$, where the update of the $i$-th algorithm is defined by
+	\[
+	\mX^{k + 1} - \mX^{k} = \gamma_k \mathrm{LMO}_{i}(\mM^{k})\,,
+	\]
+	and $\mathrm{LMO}_i$ corresponds to the unit ball of norm $\norm{\cdot}_i$. For arbitrary non-negative $\alpha_1,\dots,\alpha_n$, the algorithm with the update given by
+	\[
+	\mX^{k + 1} - \mX^{k} = \gamma_k \sum_{i = 1}^n \alpha_i \mathrm{LMO}_{\norm{\cdot}_{(i)}}(\mM^k)
+	\]
+	is an LMO-algorithm itself, with LMO corresponding to the unit ball of the norm $\norm{\cdot}$ dual to the norm given by $\sum_{i = 1}^n \alpha_i \norm{\cdot}_{(i)}^\dagger$.
+\end{corollary}
+
+\subsection{Frobeniusize Them: F-Muon and F-Neon}
+Let us construct the concrete examples of algorithms given by linear combinations of LMO-algorithms.
+It follows from \corollaryref{corollary:lin_comb_alg} that those algorithms can also be viewed as LMO-algorithms.
+
+Combining Fanions for various $k$ with another excellent LMO-algorithm, NSGD, we obtain the family of algorithms with updates defined by
+\begin{equation}\label{eq:nsgd_muon_update}
+		\mX^{t+1} = \mX^{t} - \gamma_k \left(\alpha \sum_{i = 1}^k \vu_i \vv_i^\top + (1-\alpha)\frac{\mM^k}{\normf{\mM^k}}\right)\,.
+\end{equation}
+Recall that Fanion-$k$ update is governed by the norm dual to the Ky Fan-$k$ norm, and NSGD corresponds to the Frobenius norm, which is dual to itself.
+Thus, by \corollaryref{corollary:lin_comb_alg}, our new algorithm is an LMO-algorithm corresponding to the unit ball of the norm $\norm{\cdot}_{\mathrm{F-KF-k}}$:
+\begin{equation}\label{eq:fkfk_norm}
+	\norm{\cdot}_{\mathrm{F-KF-k}}^\dagger = \alpha \normkfk{\cdot} + (1-\alpha)\normf{\cdot}\,.
+\end{equation}
+We name the derived family of algorithms \emph{F-Fanions}.
+
+Two edge members of this family, with $k = 1$ and $k = \min\{m, n\}$ correspondingly, \emph{F-Neon} and \emph{F-Muon}, are of particular interest to us.
+A bit of information and visualizations related to the $\normfstar{\cdot} = \norm{\cdot}_{\mathrm{F-KF-1}}$ norm is presented in the appendix (see \equationref{eq:normfstardual}, \figureref{fig:fstardual}).
+
+\subsection{Add SignSGD: S-Muon and S-Neon}
+
+Another algorithm we consider is the SignSGD. Mixing its update with that of Fanion-$k$, we obtain the update of \emph{S-Fanion-$k$}:
+\begin{equation}\label{eq:sign_muon_update}
+		\mX^{t+1} = \mX^{t} - \gamma_k \left(\alpha \sum_{i = 1}^k \vu_i \vv_i^\top + (1-\alpha)\eta\sign(\mM^k)\right)\,,
+\end{equation}
+where $\eta$ is the special SignSGD learning rate coefficient.
+
+The norm $\norm{\cdot}_{\mathrm{C-KF-k}}$ which produces this update in the LMO framework is again given by \corollaryref{corollary:lin_comb_alg}:
+\begin{equation}\label{eq:ckfk_norm}
+	\norm{\cdot}_{\mathrm{C-KF-k}}^\dagger = \alpha \normkfk{\cdot} + \frac{1-\alpha}{\eta}\normc{\cdot}^\dagger\,.
+\end{equation}
+
+\section{Computing the Updates}\label{sec:matrix-side}
+We employ the thick-restarted Lanczos method for the symmetric eigenvalue problem (thick-restart Lanczos, TRLan) to compute the low-rank updates of Fanions.
+We apply TRLan to either $\mM^{k\top}\mM^k$ or $\mM^k \mM^{k\top}$ (whichever matrix is smaller).
+We use the CuPy implementation of \texttt{cupy.sparse.linalg.svds} \citep{cupy_svds_ref} which internally relies on TRLan \citep{simonz1998thick}.
+
+TRLan is specifically designed for efficiently approximating the largest singular values and corresponding singular vectors of very large matrices.
+The thick-restart strategy retains the most informative Ritz vectors across restarts, which dramatically accelerates convergence while keeping memory consumption moderate.
+TRLan is particularly attractive in our GPU setting because its dominant cost is a modest number of highly parallelizable matrix-vector multiplications (matvecs) and it avoids full reorthogonalization against the entire Krylov basis by using short recurrence relations combined with thick restarting.
+
+The per-cycle complexity is $\mathcal{O}(mn^2 + n^2k + nk^2)$, where $m \geq n$ are the dimensions of the target matrix and $k$ is the size of the retained subspace (typically $k \ll n$).
+
+In \tableref{tbl:matrix_methods}, we compare TRLan against randomized SVD (RSVD) and simple power iterations when computing the rank-$k$ update used in Fanion-$k$ and related algorithms.
+Experiments are performed on dense random matrices with i.i.d.
+$\mathcal{N}(0,1)$ entries and use CPU implementations for fair comparison.
+We report:
+\begin{itemize}
+  \item $err_1$: relative error in the Frobenius norm of $\sum_i^k \vu_i \vv_i^T$,
+  \item $err_2$: relative error in the Frobenius norm of $\sum_i^k \sigma_i \vu_i \vv_i^T$.
+\end{itemize}
+
+On $500\times500$ matrices, TRLan and RSVD require comparable wall-clock time, but TRLan delivers orders-of-magnitude lower error and far fewer matvecs.
+On larger $5000\times5000$ matrices the advantage becomes even more pronounced: TRLan is 3-4 times faster than RSVD while using $\sim$30 times fewer matvecs at comparable or better accuracy.
+
+An interesting empirical observation is that RSVD tends to approximate the \emph{singular values} themselves reasonably well but the reconstructed low-rank matrix noticeably deviates from the truncated SVD, whereas TRLan provides an excellent approximation to the truncated SVD matrix itself (low $err_2$) at the cost of occasionally less accurate individual singular values.
+This makes TRLan the preferred choice for algorithms like Neon/Fanion-$k$ that only need the low-rank term $\sum \sigma_i \vu_i \vv_i^\top$, but less ideal for methods (e.g., Dion) that explicitly require accurate $\sigma_i$ for error feedback or step-size control.
+
+A current practical limitation is the lack of a native PyTorch implementation of thick-restart Lanczos; existing PyTorch-based randomized SVD routines cannot match TRLan's accuracy/efficiency combination for the matrix reconstruction task.
+
+For reference, \tableref{tbl:newton_schulz_reference} shows results for the Newton-Schulz polar decomposition iteration on the same matrices, $err_1$ is the relative error of $\mU\mV^T$ (29-30 iterations to converge, significantly higher matvec count than TRLan).
+
+\begin{table}[ht]
+	\centering
+	\begin{tabular}{|c|c|c|c|c|c|c|c|}
+	  \hline
+	  Matrix sizes & $k$  & Method           & Time (s) & Matvecs & Iterations & $err_1$         & $err_2$       \\
+	  \hline
+	  $500 \times 500$      & 5  & Power Iterations           & 0.062   & 2005    & 200        & 9.2$\cdot 10^{-3}$ & 9.1$\cdot10^{-3}$\\
+	  $500 \times 500$      & 5  & RSVD             & 0.017   & 1170    & 38         & 9.8$\cdot 10^{-3}$ & 9.6$\cdot10^{-3}$\\
+	  $500 \times 500$      & 5  & TRLan            & 0.018   & 131     & 65         & 9.6$\cdot 10^{-5}$ & 9.4$\cdot10^{-5}$\\
+	  \hdashline
+	  $500 \times 500$      & 50 & Power Iterations           & 0.44    & 43750   & 437      & 9.9$\cdot 10^{-3}$ & 9.0$\cdot10^{-3}$\\
+	  $500 \times 500$      & 50 & RSVD             & 0.61    & 6120    & 50      & 9.9$\cdot 10^{-3}$ & 9.1$\cdot10^{-3}$\\
+	  $500 \times 500$      & 50 & TRLan            & 0.16    & 462     & 231      & 3.3$\cdot 10^{-7}$ & 3.0$\cdot10^{-7}$\\
+	  \hdashline
+	  $5000 \times 5000$    & 5  & Power Iterations           & 9.6     & 9065    & 906      & 8.6$\cdot 10^{-3}$ & 8.6$\cdot10^{-3}$\\
+	  $5000 \times 5000$    & 5  & RSVD             & 2.1     & 5640    & 187      & 9.7$\cdot 10^{-3}$ & 9.7$\cdot10^{-3}$\\
+	  $5000 \times 5000$    & 5  & TRLan            & 0.70    & 205     & 102      & 7.7$\cdot 10^{-3}$ & 7.7$\cdot10^{-3}$\\
+	  \hline
+	\end{tabular}
+  \caption{Comparison of methods for computing rank-$k$ updates on dense random matrices (CPU, double precision). Lower is better in all columns.}
+  \label{tbl:matrix_methods}
+  \end{table}
+
+  \begin{table}[ht]
+  \centering
+  \begin{tabular}{|c|c|c|c|c|}
+  \hline
+  Matrix size    & Time (s) & Matvecs & Iterations & $err_1$ \\
+  \hline
+  $500 \times 500$     & 0.041    & 27 000  & 27         & 4.8e-3  \\
+  $5000 \times 5000$   & 26.4     & 290 000 & 29         & 6.5e-3  \\
+  \hline
+  \end{tabular}
+  \caption{Newton–Schulz iteration on the same matrices (for reference).}
+  \label{tbl:newton_schulz_reference}
+  \end{table}
+
 \bibliography{icomp2024_conference}

### 2025-11-26 conf_article/neon.tex
Added loads of pictures for CIFAR

@@
 We log train and validation accuracy, and nuclear, gradient, and spectral norms of the gradients of conv1.weight and conv2.weight for all the layers.
+The gradient norms decrease only by the order of ten maximum during training, while train accuracy reaches full 100\% for Muon, S-Muon, and F-Muon.
 
@@
 
+\begin{figure}[h!]
+    \centering
+    \begin{subfigure}[t]{0.7\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{figs/cifar_logger/total_frobenius.pdf}
+        \caption{Total Frobenius norm}
+		\label{fig:total_fro_norm_cifar}
+    \end{subfigure}
+    \vspace{1em} % optional: add vertical spacing between subfigures
+    \begin{subfigure}[t]{0.7\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{figs/cifar_logger/total_spectral.pdf}
+        \caption{Total spectral norm}
+		\label{fig:total_spectral_norm_cifar}
+    \end{subfigure}
+	\vspace{1em} % optional: add vertical spacing between subfigures
+    \begin{subfigure}[t]{0.7\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{figs/cifar_logger/total_nuclear.pdf}
+        \caption{Total nuclear norm}
+		\label{fig:total_nuclear_norm_cifar}
+    \end{subfigure}
+    \caption{Matrix norms of the whole gradient of CIFAR CNN}
+	\label{fig:total_norms_cifar}
+\end{figure}
+
+% Juxtaposed conv1 (left) vs conv2 (right) norms
+
+\begin{figure}[h!]
+    \centering
+    %--- First row: Frobenius ---
+    \begin{subfigure}[t]{0.45\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{figs/cifar_logger/layers_2_conv1_weight_frobenius.pdf}
+        \caption{Frobenius norm of layer2.conv1}
+        \label{fig:fro_conv1}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.45\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{figs/cifar_logger/layers_2_conv2_weight_frobenius.pdf}
+        \caption{Frobenius norm of layer2.conv2}
+        \label{fig:fro_conv2}
+    \end{subfigure}
+
+    \vspace{1em}
+
+    %--- Second row: Spectral ---
+    \begin{subfigure}[t]{0.45\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{figs/cifar_logger/layers_2_conv1_weight_spectral.pdf}
+        \caption{Spectral norm of layer2.conv1}
+        \label{fig:spec_conv1}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.45\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{figs/cifar_logger/layers_2_conv2_weight_spectral.pdf}
+        \caption{Spectral norm of layer2.conv2}
+        \label{fig:spec_conv2}
+    \end{subfigure}
+
+    \vspace{1em}
+
+    %--- Third row: Nuclear ---
+    \begin{subfigure}[t]{0.45\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{figs/cifar_logger/layers_2_conv1_weight_nuclear.pdf}
+        \caption{Nuclear norm of layer2.conv1}
+        \label{fig:nuc_conv1}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[t]{0.45\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{figs/cifar_logger/layers_2_conv2_weight_nuclear.pdf}
+        \caption{Nuclear norm of layer2.conv2}
+        \label{fig:nuc_conv2}
+    \end{subfigure}
+
+    \caption{Matrix norms of layer2.conv1 (left) and layer2.conv2 (right) gradients of CIFAR CNN}
+    \label{fig:norms_conv1_vs_conv2}
+\end{figure}
+
 \section{Technical details of the experiments}

### 2025-11-27 article_final/main.tex
New style file reached

@@
 \documentclass[pmlr]{jmlr}% new name PMLR (Proceedings of Machine Learning)
-
+\input{math_commands.tex}
 \RequirePackage{graphicx}
@@
 \usepackage{longtable}% for long tables
+\usepackage{xcolor}% required by colortbl
+\usepackage{colortbl}% for colored table rules (required by arydshln)
 \usepackage{arydshln}% for dashed lines in tables
@@
 
-% Custom math commands
-% Blackboard bold
-\newcommand{\R}{\mathbb{R}}
-
-% Bold matrices
-\newcommand{\mA}{\bm{A}}
-\newcommand{\mB}{\bm{B}}
-\newcommand{\mD}{\bm{D}}
-\newcommand{\mF}{\bm{F}}
-\newcommand{\mG}{\bm{G}}
-\newcommand{\mM}{\bm{M}}
-\newcommand{\mN}{\bm{N}}
-\newcommand{\mS}{\bm{S}}
-\newcommand{\mU}{\bm{U}}
-\newcommand{\mV}{\bm{V}}
-\newcommand{\mX}{\bm{X}}
-\newcommand{\mSigma}{\bm{\Sigma}}
-
-% Vectors (note: \vec is already defined by jmlr class)
-% We define shortcuts for commonly used vectors
-\newcommand{\vu}{\vec{u}}
-\newcommand{\vv}{\vec{v}}
-
-% Calligraphic letters
-\newcommand{\cB}{\mathcal{B}}
-\newcommand{\cS}{\mathcal{S}}
-
 % Norms and operators
@@
 \DeclarePairedDelimiter{\normftwo}{\|}{\|_\mathrm{F2}}
+\DeclarePairedDelimiter{\abs}{\lvert}{\rvert}
 
@@
 % Math operators
-\DeclareMathOperator*{\argmin}{arg\,min}
-\DeclareMathOperator{\sign}{sign}
 \DeclareMathOperator{\tr}{tr}
@@
 \begin{table}[ht]
-	\centering
+\floatconts
+  {tbl:matrix_methods}
+  {\caption{Comparison of methods for computing rank-$k$ updates on dense random matrices (CPU, double precision). Lower is better in all columns.}}
+  {%
     \begin{tabular}{|c|c|c|c|c|c|c|c|}
-	  \hline
+      \arrayrulecolor{black}\hline
       Matrix sizes & $k$  & Method           & Time (s) & Matvecs & Iterations & $err_1$         & $err_2$       \\
-	  \hline
+      \arrayrulecolor{black}\hline
       $500 \times 500$      & 5  & Power Iterations           & 0.062   & 2005    & 200        & 9.2$\cdot 10^{-3}$ & 9.1$\cdot10^{-3}$\\
@@
       $5000 \times 5000$    & 5  & TRLan            & 0.70    & 205     & 102      & 7.7$\cdot 10^{-3}$ & 7.7$\cdot10^{-3}$\\
-	  \hline
-	\end{tabular}
-  \caption{Comparison of methods for computing rank-$k$ updates on dense random matrices (CPU, double precision). Lower is better in all columns.}
-  \label{tbl:matrix_methods}
+      \arrayrulecolor{black}\hline
+    \end{tabular}%
+  }
 \end{table}
@@
 \begin{table}[ht]
-  \centering
+\floatconts
+  {tbl:newton_schulz_reference}
+  {\caption{Newton–Schulz iteration on the same matrices (for reference).}}
+  {%
     \begin{tabular}{|c|c|c|c|c|}
-  \hline
+      \arrayrulecolor{black}\hline
       Matrix size    & Time (s) & Matvecs & Iterations & $err_1$ \\
-  \hline
+      \arrayrulecolor{black}\hline
       $500 \times 500$     & 0.041    & 27 000  & 27         & 4.8e-3  \\
       $5000 \times 5000$   & 26.4     & 290 000 & 29         & 6.5e-3  \\
-  \hline
-  \end{tabular}
-  \caption{Newton–Schulz iteration on the same matrices (for reference).}
-  \label{tbl:newton_schulz_reference}
+      \arrayrulecolor{black}\hline
+    \end{tabular}%
+  }
 \end{table}
+
+\section{Experiments}\label{sec:experiments}
+
+\subsection{Randomized Linear Least Squares}
+\label{subsec:lls_exps}
+
+First, we compare F-Fanions on the following convex $L$-smooth problem:
+\begin{equation}\label{eq:lls}
+  F(\mX) = \frac{1}{2} \<(\mX-\mS), \mM (\mX - \mS) \mN> \to \min_{\mX \in \Rmn}\,
+\end{equation}
+where $\mX \in \Rmn$, $m=500$ and $n=500$ are typical dimensions of a neural network weight matrix, $\mS \in \Rmn$, $\mM \in \mathbb{S}^m_{+}$ and $\mN \in \mathbb{S}^n_{+}$ are positive-semidefinite matrices.
+The spectra of $\mM$ and $\mN$ are uniformly distributed in the $(0, 1)$ interval.
+We set $\mS=0$, and $\mX^0 \sim 0.1 \cN(0, \mI)$.
+
+We run different Fanions and their respective F-Fanions with $\alpha=1/2$: Neon (Fanion-1), Fanion-2, Fanion-10, Fanion-100, and Muon (Fanion-500).
+Also, we test S-Fanions with $\alpha=1/2$ and \texttt{sign\_lr\_coeff=0.01}.
+We test them against Normalized SGD and SignSGD, which are also F-Fanion and S-Fanion respectively with $\alpha = 0$ and an arbitrary $k$.
+
+Since theoretical bounds \citep{kovalev2025understanding, riabinin2025gluon} rely on a very loose norm bound $\norm{\cdot}\le \rho \normf{\cdot}$, we do not derive learning rate or Nesterov momentum from the smoothness constants that also depend on the norm choice.
+Rather, we find such \texttt{(lr, momentum)} pair that the loss threshold $0.001$ is reached in a minimal number of iterations.
+This setting is the most realistic, and at the same time aligns with the corollaries of the convergence theorems where constant learning rate and momentum coefficients are proposed.
+
+We perform a grid search for \texttt{momentum} $\in \{0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 0.95\}$ and for \texttt{lr}.
+For Muon's, F-Muon's, S-Muon's \texttt{lr} $\in [0.005, 0.0020]$.
+For SignSGD, \texttt{lr} $\in 0.01 \times [0.005, 0.020]$, and for NSGD \texttt{lr} $\in [0.01, 0.10]$.
+The tuned parameters and the number of iterations to converge to $0.001$ are in \tableref{tbl:lls_lrs}.
+\texttt{lr} and \texttt{momentum}, and \texttt{sign\_lr\_coeff} for Fanions, F-Fanions, and S-Fanions are the same as for Muon, F-Muon, and S-Muon respectively due to the impossibility to tune them properly: their losses decrease too slow to reach 0.001.
+
+The results are presented in \figureref{fig:lls} and detailed in \sectionref{sec:lls_plots_section}.
+
+Both F-Muon and S-Muon converge faster and to a lower loss and a lower Frobenius norm of the gradient than NSGD, Muon, or SignSGD.
+
+\begin{table}[htbp]
+\floatconts
+  {tbl:lls_lrs}
+  {\caption{Tuning learning rates and momentum coefficients}}
+  {%
+    \begin{tabular}{|c|c|c|c|}
+      \arrayrulecolor{black}\hline
+      Algorithm & \texttt{lr}  & \texttt{momentum}  & Iterations to 0.001 loss \\
+      \arrayrulecolor{black}\hline
+      Muon & 0.007 & 0.5 & 1060 \\
+      \hdashline
+      NSGD & 0.08 & 0.95 & 1020 \\
+      F-Muon & 0.015 & 0.7 & 910 \\
+      \hdashline
+      SignSGD & 0.016 $\times$ 0.01 & 0.95 & 2650 \\
+      S-Muon & 0.011 & 0.9 & 890 \\
+      \arrayrulecolor{black}\hline
+    \end{tabular}%
+  }
+\end{table}
+
+\begin{figure}[htbp]
+\floatconts
+  {fig:lls}
+  {\caption{Linear least squares problem for a 500x500 matrix}}
+  {%
+    \subfigure[The loss]{\label{fig:lls_loss}%
+      \includegraphics[width=0.8\linewidth]{figs/simple_lls/loss_long_run_16_increase.pdf}}%
+    \vspace{1em}
+    \subfigure[The Frobenius norm of the gradient]{\label{fig:lls_fro_grad_norm}%
+      \includegraphics[width=0.8\linewidth]{figs/simple_lls/gradient_frobenius_norm_vs_iteration_long_run.pdf}}
+  }
+\end{figure}
+
+\subsection{CIFAR-10 airbench}
+
+We compare the algorithms on the CIFAR-10 airbench \citep{cifar2023airbench}.
+First, to test the impact of $\alpha$, we run F-Muon for different $\alpha$ with the same \texttt{lr=0.24(1 - step/total\_steps), momentum=0.65, nesterov=True} and weight normalization, as have been finetuned by Keller Jordan, 10 repetitions for each $\alpha$.
+We record the accuracy after 8 epochs of training (\figureref{fig:muon_alphas}).
+
+Then we tune F-Muon with $\alpha = 0.5$.
+Tuned parameters are \texttt{lr=0.4(1 - step/total\_steps), momentum=0.6, nesterov = True}.
+Tuned F-Muon reaches $94.02 \pm 0.13\%$ after 8 epochs (averaged by 200 iterations), matching Muon's accuracy and variance.
+Once again we measure \texttt{val\_accuracy}($\alpha$) curve (\figureref{fig:fmuon_alphas}).
+We notice that even when $\alpha=0.1$, the accuracy is much higher than in the case of vanilla NSGD.
+
+We tune S-Muon with $\alpha=0.5$ for the best \texttt{(lr, momentum, sign\_lr\_coeff)} and by setting \texttt{lr=0.42(1 - step/total\_steps), momentum=0.63, nesterov = True, sign\_lr\_coeff = 0.003} get $\mathbf{94.03} \pm 0.13\%$, which is even higher than for vanilla Muon.
+
+\begin{figure}[htbp]
+\floatconts
+  {fig:diff_alphas}
+  {\caption{Mean validation accuracies for F-Muon with different $\alpha$}}
+  {%
+    \subfigure[With parameters tuned for Muon]{\label{fig:muon_alphas}%
+      \includegraphics[width=0.48\linewidth]{figs/muon_tuned_diff_alpha.pdf}}%
+    \hfill
+    \subfigure[With parameters tuned for F-Muon]{\label{fig:fmuon_alphas}%
+      \includegraphics[width=0.48\linewidth]{figs/fmuon_tuned_diff_alpha.pdf}}
+  }
+\end{figure}
+
+The Muon-level performance is surprising for the contorted geometry of F-Muon and S-Muon.
+F-Muon can be visualized by \figureref{fig:cifar_ball}, where the LMO balls of Muon and F-Muon for real \texttt{lr} are plotted in a 2D space of singular values.
+S-Muon cannot be visualized because the $l_\infty$ norm is not a function of singular values, while $\norms{\cdot}$, when plotted for the $\Rmn$ space with $m, n > 2$, requires at least a 4D space.
+However, it is plain that the LMO ball of S-Muon significantly differs from the ball of Muon because the $\texttt{lr}(1-\alpha)\texttt{sign\_lr\_coeff} = 6.3 \times 10^{-4}$, which is comparable to the usual \texttt{lr} of SignSGD for deep learning problems.
+
+The most pathological case is $\alpha > 1$, which corresponds to the LMO with an area that is not a norm ball!
+Despite this violation, the mixture of algorithms reaches almost the same accuracy as vanilla Muon.
+
+These observations raise the question of how much LMO-based algorithms are sensitive to the constraint area.
+
+\begin{figure}[htbp]
+\floatconts
+  {fig:cifar_ball}
+  {\caption{Visualization of the LMO balls for Muon and F-Muon.}}
+  {\includegraphics[width=0.6\linewidth]{figs/norms/fstardual_cifar.pdf}}
+\end{figure}
+
+\subsection{NanoGPT speedrun}
+
+We test F-Muon on the NanoGPT speedrun \citep{modded_nanogpt_2024}.
+For $\alpha = 0.5$, the optimal parameters are \texttt{lr=0.07, momentum=0.95}, while for Muon they were \texttt{lr=0.05, momentum=0.95}.
+After testing for 1750 steps we get 3.281 cross-entropy loss.
+For Muon, it falls below the target threshold 3.28 and reaches 3.279.
+However, this difference is negligible if one looks at \figureref{fig:gpt_loss}.
+It is even more striking, considering the fact that this F-Muon is an average of Muon and NSGD, and the latter performed quite poorly.
+
+\begin{figure}[htbp]
+\floatconts
+  {fig:gpt_loss}
+  {\caption{The validation loss for NanoGPT}}
+  {\includegraphics[width=0.7\linewidth]{figs/nanogpt/val_loss_vs_step.pdf}}
+\end{figure}
+
+Again, as on the CIFAR airbench, if we set $\alpha=-0.1$ that corresponds to no ball, we get a 3.2818 loss, which is a small difference from the baseline.
+
+\subsection{NanoGPT Fine-Tuning}
+\label{subsec:finetune}
+
+We fine-tuned the NanoGPT framework by Karpathy \citep{Karpathy2022} using the standard \textit{NanoGPT-medium} configuration, which corresponds to GPT-2 Medium.
+This model consists of 24 transformer layers, a hidden size of 1024, 16 attention heads, and approximately 345 million parameters.
+The training dataset was the \textit{Tiny Stories} corpus \citep{eldan2023tinystoriessmalllanguagemodels}, selected for its high entropy and structural diversity, which amplify and make more visible the differences in optimization dynamics.
+All training runs were initialized with the same random seed, weight initialization, and learning rate schedule to ensure that any differences in performance arose solely from the choice of optimizer.
+
+For all one-dimensional layers, AdamW was used with a learning rate of $1\!\times\!10^{-3}$.
+Since momentum exhibited negligible influence on training dynamics, it was held constant across all experiments.
+A comparative analysis of optimization performance is provided in \figureref{fig:nanogptftexps}.
+
+\begin{figure}[htbp]
+\floatconts
+  {fig:nanogptftexps}
+  {\caption{Comparison of Muon, F-Muon, and S-Muon across a range of learning rates. Stars denote the best learning rate identified for each optimizer. Validation loss.}}
+  {\includegraphics[width=\linewidth]{figs/nanogpt/optimizers_comparison.pdf}}
+\end{figure}
+
+\figureref{fig:nanogptftexps2} presents the train and validation loss curves at the optimal learning rate for each optimizer.
+While F-Muon and S-Muon maintain consistent behavior, vanilla Muon attains the lowest loss overall.
+
+\begin{figure}[htbp]
+\floatconts
+  {fig:nanogptftexps2}
+  {\caption{Train and validation loss at the optimal learning rate for each optimizer.
+The standard Muon method achieves the lowest loss, while F-Muon and S-Muon remain stable but exhibit slightly inferior performance.}}
+  {\includegraphics[width=\linewidth]{figs/nanogpt/Nanogpt finetune results.pdf}}
+\end{figure}
+
+\section{Algorithms for Vectors $\rightarrow$ Algorithms for Matrices}
+
+One cannot but notice the similarity between the updates of LMO optimizers in the vector $l_p$ and the matrix Schatten $S_p$ norms, which is illustrated by \tableref{tbl:mat_vs_vec_lmo}.
+The parallels are not limited to mere similarity of the updates and can be traced even in the empirical observations.
+SignSGD is very close to Adam, as noted in \citet{bernstein2024oldoptimizernewnorm}, and both Adam and Muon perform well in training large models \citep{zhao2025deconstructing, liu2025muon}.
+NSGD is the same for both matrix and vector cases.
+Greedy Coordinate Descent is not applied to high-dimensional problems, so from this perspective it is not surprising that one-rank Neon underperforms on such problems.
+
+\begin{table}[t]
+\floatconts
+  {tbl:mat_vs_vec_lmo}
+  {\caption{LMO optimizers in Schatten $S_p$ norms and in $l_p$ norms. $g$ is the gradient. When it is a matrix, $g = \mU \mSigma \mV^\top$}}
+  {%
+    \begingroup
+    \def\arraystretch{1.2}
+    \resizebox{\linewidth}{!}{%
+      \begin{tabular}{|l|c|c|c|}
+        \arrayrulecolor{black}\hline
+        Algorithm                             & LMO constraint set $\mathcal{D}$ & LMO                                                         & Reference                            \\
+        \arrayrulecolor{black}\hline
+        \arrayrulecolor{black}\hline
+        Normalized SGD                     & $l_2$-ball, $S_2$-ball          & $-\eta \tfrac{g}{\norm{g}_2} = -\eta \tfrac{g}{\normf{g}}$ & \citep{hazan2015beyond}              \\
+        Momentum Normalized SGD            & Ball in $l_2$, or Ball in $S_2$ & $-\eta \tfrac{g}{\norm{g}_2} = -\eta\tfrac{g}{\normf{g}}$  & \citep{cutkosky2020momentum}         \\
+        \arrayrulecolor{black}\hline
+        SignSGD                            & Ball in Max-norm $l_\infty$     & $-\eta \sign(g)$                                            & \citep[Thm. 1]{bernstein2018signsgd} \\
+        Signum                             & Ball in Max-norm $l_\infty$     & $-\eta \sign(g)$                                            & \citep[Thm. 3]{bernstein2018signsgd} \\
+        \hdashline
+        Muon                               & Ball in Spectral $S_\infty$     & $-\eta \mU\mV^\top$                                             & \citep{jordan2024muon}               \\
+        \arrayrulecolor{black}\hline
+        Gauss-Southwell Coordinate Descent & Ball in $l_1$                   & $-\eta \sum_{i \in \argmax|g_i^t|} \sign(g_i^t)e_i$                       & \citep[p. 19]{shi2016primer}         \\
+        \hdashline
+        Neon                               & Ball in Nuclear $S_1$           & $-\eta \vu_1 \vv_1^\top$                                        & This work                            \\
+        \arrayrulecolor{black}\hline
+      \end{tabular}%
+    }%
+    \endgroup
+  }
+\end{table}
+
+\section{Related Work and Discussion}
+
+Since Muon \citep{jordan2024muon} is a very efficient optimizer for functions of weight matrices, a lot of research has been put into, first, further improving its performance, and second, explaining its success.
+
+\textbf{Improvements of Muon.} Regarding the first point, a large number of applications and improvements of Muon has been proposed in less than a year.
+\citet{liu2025muon} adapted the algorithm for training language models larger than NanoGPT.
+\citet{shah2025practical} organized efficient hyperparameter transfer by combining Muon with maximal update parametrization.
+To construct their COSMOS optimizer, \citet{chen2025cosmoshybridadaptive} applied computationally intensive updates of SOAP optimizer \citep{vyas2025soap} to a low-dimensional ``leading eigensubspace'' while using memory-efficient methods like Muon for the remaining parameters.
+\citet{amsel2025polar, grishina2025chebyshev} proposed more efficient alternatives to Newton-Schulz algorithm.
+\citet{si2025adamuon} introduced AdaMuon which combines element-wise adaptivity with orthogonal updates.
+We suppose that the described above or similar techniques can be applied to our optimizers as well.
+For example, F-Muon and S-Muon also benefit from faster alternatives to Newton-Schulz iterations, and Fanions may be a great substitute to Muon in COSMOS, because, as we have shown in \sectionref{sec:matrix-side}, Lanczos algorithm is much faster than Newton-Schulz iterations on very large matrices.
+
+\textbf{Theory behind Muon.} Regarding the second point, there has been a prolonged gap in theory behind Muon, simplistic derivation of \citet{bernstein2025deriving} based on \citet{bernstein2024oldoptimizernewnorm} excluded.
+This gap, as it seems to us, is not even now completely closed.
+For example, \citet{kovalev2025understanding} has provided convergence guarantees of Muon and in various settings, from which, however, Muon's supremacy cannot be recovered.
+Indeed, although the obtained bounds depend on the norm choice, the asymptotics of the convergence remain the same as for NSGD and other optimizers, $K = \cO(\epsilon^{-4})$ in an $L$-smooth stochastic case.
+
+A similar drawback has the article \citet{riabinin2025gluon}, where $L$-smoothness assumption is replaced with a more practical $(L_0, L_1)$-smoothness.
+By estimating smoothness and substituting it into their Theorem 1, the authors recovered the optimal fine-tuned stepsizes reported by \citet{pethick2025training}.
+However, they have not showed the optimality of the spectral or the RMS-to-RMS norm, which is observed in practice, as our comparison with NSGD highlights.
+
+Common important drawback of the analyses is the consideration of the convergence by the norm of the gradient.
+As we showed in our experiments on CIFAR, the gradient norm may decrease only by a factor of ten, when the accuracy reaches 100\%.
+
+We suppose that the reason for the recorded stark discrepancy between Neon and Muon performance, both of which are described by Stochastic Conditional Gradient \citep{pethick2025training} or Gluon frameworks, lies in the structure of the norm ball or in the preconditioner interpretation \citep{pethick2025trainingneuralnetworksscale}, which must be an object of further research.
+
+\textbf{The LMO and Error Feedback.} We already mentioned that rank-$k$ unsharded Dion without error feedback and scaling of the update is Fanion-$k$.
+Since error feedback for Dion is very important as discovered by the ablation study in \citet{ahn2025dion}, F-Fanions and S-Fanions will benefit from it as well.
+In the context of federated learning, error feedback is effective even for compressed Muon \citep{gruntkowska2025efmuon}.
+The advantage of Fanions and S-Fanions is that less bits are required for transmission: $\sum_{i=1}^k \vu_i \vv_i^\top$ is easily transmitted as $\{\vu_1, \vv_1, \dots, \vu_k, \vv_k\}$ ($(m+n)\times k$ floats), while the sign part is as usual coded in $m \times n$ bits.
+Thus, the compression is ``built-in''.
+Moreover, there is an intriguing possibility to construct differentially private Fanions and S-Fanions with the specific more optimal non-Gaussian noise, as was done with DP-signSGD \citep{jang2024dplogsign}.
+This we leave for future research.
+
+\textbf{The nuclear norm in the LMO.} As we found out only when transforming our results into an article, the nuclear norm has already been explored in the context of the linear minimization oracle.
+\citet{pethick2025sam} applied it to create $\nu$SAM, a new sharpness-aware minimization technique.
+It would be interesting to substitute $\normn{\cdot}$ with $\normkfk{\cdot}^\dagger$ in their approach.
+Since the SAM-neighborhood becomes more diverse, $k > 1$ might add to the accuracy boost, while preserving a small memory footprint and a small time overhead, if Dion-style power iterations are used.
+
+\textbf{Orthogonal research.} Fanions, F-Fanions, and S-Fanions benefit from the general theoretical description of \citet{riabinin2025gluon}, and better learning rates can be predicted by calculating the trajectory smoothness.
+They could be transformed to Drop-Fanions by updating only some layers as in \citet{gruntkowska2025dropmuon}.
+They could be viewed as approximations of the Non-Euclidean Broximal Point Method for the corresponding norms \citep{gruntkowska2025noneuclideanbroximal}.
+One can clip them to produce ClippedScion-like algorithms \citep{pethick2025generalizedgradientnormclipping}.
+They could be made more memory-effective by the zero-order techniques that worked for Muon \citep{petrov2025zeromuon}.
+Finally, the results from \citet{shulgin2025inexactmuon} can be used to explain the robustness of Muon to the norm change that occurred in our experiments and to theoretically derive faster yet working approximate schemes to calculate the LMO; power iterations with a very limited number of iterations is a good algorithm to consider for the analysis.
+
+\section{Conclusion}
+
+In this article, we have generalized several successful algorithms, like Muon and Dion, to the LMO-based algorithms in the $\normkfk{\cdot}^\dagger$ norm.
+Also we have proposed the technique of ``regularizing'' the updates with NSGD, a trick to increase the robustness of the algorithms, which is motivated by the consideration of the $\normfstar{\cdot}$, $\normftwo{\cdot}$, and general $\norm{\cdot}_{\mathrm{F-KF-k}}$ norms.
+Generalizations of well-known norms and subsequent combinations of them may further improve performance of LMO-based algorithms.
+If a theory is developed that explains the discrepancy between performance of different algorithms based on the matrix norms, one will probably be able to intentionally construct norms optimal to given architectures and probably even their parameters.
+
+We suggest that future works that deal with the non-Euclidean LMO explain in corollaries to their convergence theorems the empirical superiority of Muon to other Fanions.
+Without that, it is hard to believe that the bounds are relevant for practitioners.
+
+\section{Author Contributions}
+
+IO suggested using the nuclear norm in the \citet{bernstein2024oldoptimizernewnorm} framework.
+DM presented the problem at the MIPT optimization course and helped with editing the article.
+IK suggested using composite norms (though not the ones that induce F-Fanions and S-Fanions) and refactored \sectionref{sec:conic_combo}.
+NK suggested Lanczos algorithm as the most precise means to compute Fanions' updates, conducted experiments to prove this, and wrote \sectionref{sec:matrix-side}.
+AV conducted the finetuning of NanoGPT on Tiny Stories and wrote \sectionref{subsec:finetune}.
+All other work was done by AK: creating Fanions, F-Fanions, S-Fanions, experimenting on the Linear Least Squares, CIFAR-airbench and NanoGPT pretrain, and writing the manuscript.
 
@@
 
+\appendix
+\section{LMO for Neural Networks}
+
+In a typical neural network, the objective function $F$ depends on a set of weight matrices $\{ \mW_1, \mW_2, \ldots \}$.
+The optimization framework we have described is applied in a layer-wise fashion.
+At each iteration $k$, a stochastic gradient $g(\mX^k, \xi^k)$ is computed using a mini-batch of data with the $\xi^k$ noise via backpropagation.
+This yields a separate gradient component, $\mG_i^k$, for each matrix $\mX_i$.
+The LMO-based update rule is then applied to each matrix $\mX_i$ using its corresponding gradient component $\mG_i^k$.
+
+The update rule used in Muon optimizer is uSCG \citep{pethick2025training}. In the most general case, which involves momentum, it can be written as
+\begin{equation}\label{eq:usgd_update}
+  \begin{aligned}
+    \mM^{k}     & = \alpha_{k} g(\mX^k, \xi^k) + (1 - \alpha_k)\mM^{k - 1}\,, \\
+    \mX^{k + 1} & = \mX^k + \gamma_k\mathrm{LMO}(\mM^{k})\,.
+  \end{aligned}
+\end{equation}
+
+\section{Norms $\normfstar{\cdot}^\dagger$ and $\normftwo{\cdot}^\dagger$}
+
+Based on the lemma, we immediately find $\normfstar{\cdot}^\dagger$, which is related to F-Muon update.
+Indeed, after setting $\beta=1-\alpha$ and remembering that for smooth and bounded cases we can use $\min$ instead of $\inf$, we get
+\begin{equation}\label{eq:normfstardual}
+  \normfstar{\mY}^\dagger = \min_{\mZ} \min_t\{t, s.t. \norms{\mZ}\leq \alpha t, \normf{\mY-\mZ}\leq (1-\alpha) t\}
+\end{equation}
+
+If $\alpha = 1$, then $\mZ = \mY$, and we get $\normfstar{\mY}^\dagger = \norms{\mY}$. If $\alpha = 0$, then $\mZ = 0$, and we get $\normfstar{\mY}^\dagger = \normf{\mY}$.
+
+Similarly, we find $\normftwo{\cdot}^\dagger$, which is related to F-Neon update:
+\begin{equation}\label{eq:normftwodual}
+  \normftwo{\mY}^\dagger = \min_{\mZ} \min_t\{t, s.t. \normn{\mZ}\leq \alpha t, \normf{\mY-\mZ}\leq (1-\alpha) t\}
+\end{equation}
+
+If $\alpha = 1$, then $\mZ = \mY$, and we get $\normftwo{\mY}^\dagger = \normn{\mY}$. If $\alpha = 0$, then $\mZ = 0$, and we get $\normftwo{\mY}^\dagger = \normf{\mY}$.
+
+Unfortunately, $\normfstar{\cdot}^\dagger$ and $\normftwo{\cdot}^\dagger$ do not have a closed-form expression.
+
+\section{Visualization of different matrix norms}
+\subsection{Duals to F* and F2 norms}
+
+It follows from \lemmaref{lemma:dual_to_conv_comb} that the norm ball in $\normfstar{\cdot}^\dagger$ is the Minkowski sum of the norm ball in $\alpha\normn{\cdot}$ and $(1-\alpha)\normf{\cdot}$ and the norm ball in $\normftwo{\cdot}^\dagger$ is the Minkowski sum of the norm ball in $\alpha\norms{\cdot}$ and $(1-\alpha)\normf{\cdot}$.
+
+In \figureref{fig:fduals} we plot these norms. On x-axis and y-axis, there are singular values $\sigma_1$, $\sigma_2$ respectively of a matrix from $\R^{m\times n}$ with $\min\{m,n\}=2$.
+
+\begin{figure}[htbp]
+\floatconts
+  {fig:fduals}
+  {\caption{Balls in the duals to F* and F2 norms for different $\alpha$}}
+  {%
+    \subfigure[lmo balls for F-Muon for different $\alpha$]{\label{fig:fstardual}%
+      \includegraphics[width=0.48\linewidth]{figs/norms/fstardualball.pdf}}%
+    \hfill
+    \subfigure[lmo ball for F-Neon for different $\alpha$]{\label{fig:ftwodual}%
+      \includegraphics[width=0.48\linewidth]{figs/norms/ftwodualball.pdf}}
+  }
+\end{figure}
+
+\subsection{The Ky Fan norm and its dual}
+
+1-balls in $l_\infty$, $l_1$ and $l_2$ norms are well-known from textbooks. But what about the Ky Fan $k$-norm? How can it be represented?
+
+To showcase the complex structure of the Ky Fan $k$-norm and its dual, we suggest the illustrations \figureref{fig:kyfan_combined} with the ball in the Ky Fan $2$-norm in \figureref{fig:kyfan} and its dual in \figureref{fig:kyfandual}.
+On x-, y-, and z-axes, there are singular values $\sigma_1$, $\sigma_2$, and $\sigma_3$ respectively of a matrix from $\R^{m\times n}$ with $\min\{m,n\}=3$.
+In this particular case, we do not sort the singular values.
+In the proposed representation, we actually plot balls in the Top-$2$ norm $\max\{\abs{x}+\abs{y}, \abs{x}+\abs{z}, \abs{y}+\abs{z}\}$ and its dual norm $\max\{\max(\abs(x),\abs{y},\abs{z}), \frac{1}{2} (\abs{x}+\abs{y}+\abs{z})\}$.
+The resulting balls are much more complex than balls in $l_\infty$, $l_1$ and $l_2$ norms.
+
+In fact, those balls can be described easier if we use the results from \citet{yu2012arithmetic}.
+The Ky Fan $2$-norm ball is an intersection of three $l_1$ balls in $(x,y)$, $(x, z)$, and $(y,z)$ spaces.
+The 1-ball in the dual Ky Fan $2$-norm is an intersection of 1-ball the in $l_\infty$ norm and $\frac{1}{2}$-ball in the $l_1$ norm.
+
+\begin{figure}[htbp]
+\floatconts
+  {fig:kyfan_combined}
+  {\caption{Ky Fan 2-norm and its dual}}
+  {%
+    \subfigure[The Ky Fan $2$-norm]{\label{fig:kyfan}%
+      \includegraphics[width=0.45\linewidth]{figs/norms/KyFan.pdf}}%
+    \hfill
+    \subfigure[Dual to the Ky Fan $2$-norm]{\label{fig:kyfandual}%
+      \includegraphics[width=0.45\linewidth]{figs/norms/KyFanDual.pdf}}
+  }
+\end{figure}
+
+\section{More data for Linear Least Squares}
+\label{sec:lls_plots_section}
+
+\begin{table}[htbp]
+\floatconts
+  {tbl:lls_lrs_app}
+  {\caption{Tuning lrs and momentum coefficients}}
+  {%
+    \begin{tabular}{|c|c|c|c|}
+      \arrayrulecolor{black}\hline
+      Algorithm & \texttt{lr}  & \texttt{momentum}  & Iterations to 0.001 loss \\
+      \arrayrulecolor{black}\hline
+      Muon & 0.007 & 0.5 & 1060 \\
+      \hdashline
+      NSGD & 0.08 & 0.95 & 1020 \\
+      F-Muon & 0.015 & 0.7 & 910 \\
+      \hdashline
+      SignSGD & 0.016 $\times$ 0.01 & 0.95 & 2650 \\
+      S-Muon & 0.011 & 0.9 & 890 \\
+      \arrayrulecolor{black}\hline
+    \end{tabular}%
+  }
+\end{table}
+
+\begin{figure}[htbp]
+\floatconts
+  {fig:app_lls}
+  {\caption{More images for Linear least squares problem for a 500x500 matrix}}
+  {%
+    \subfigure[The spectral norm of the gradient]{\label{fig:lls_op_grad_norm}%
+      \includegraphics[width=0.6\linewidth]{figs/simple_lls/gradient_spectral_norm_vs_iteration_long_run.pdf}}%
+    \vspace{1em}
+    \subfigure[The nuclear norm of the gradient]{\label{fig:lls_nuclear_grad_norm}%
+      \includegraphics[width=0.6\linewidth]{figs/simple_lls/gradient_nuclear_norm_vs_iteration_long_run.pdf}}%
+    \vspace{1em}
+    \subfigure[The loss over time]{\label{fig:lls_loss_time}%
+      \includegraphics[width=0.6\linewidth]{figs/simple_lls/loss_vs_time_long_run.pdf}}
+  }
+\end{figure}
+
+\section{Under the hood of CNN on CIFAR-10}
+
+Most bounds for Muon and other LMO-algorithms are given for the norm of the gradient \citep{kovalev2025understanding, pethick2025training, riabinin2025gluon,kovalev2025noneuclideansgdstructuredoptimization}.
+Accordingly, it is natural to measure those norms on a real deep learning problem.
+
+\begin{table}[htbp]
+\floatconts
+  {tbl:cifar_lrs}
+  {\caption{Parameters for CIFAR-airbench. \texttt{sign\_lr\_mult} for S-Muon is 0.003}}
+  {%
+    \begin{tabular}{|c|c|c|c|}
+      \arrayrulecolor{black}\hline
+      Method & \texttt{lr}  & \texttt{momentum}  & val\_accuracy, \% \\
+      \arrayrulecolor{black}\hline
+      NSGD & 0.5 & 0.95 & $91.6 \pm 0.52$ \\
+      \hdashline
+      SignSGD & 0.003 & 0.95 & $91.54 \pm 0.26$ \\
+      \hdashline
+      Muon & 0.24 & 0.6 & $94.01 \pm 0.10$ \\
+      F-Muon & 0.40 & 0.6 & $94.01 \pm 0.13$\\
+      \hdashline
+      S-Muon & 0.42 & 0.63 & $94.03 \pm 0.13$ \\
+      \hdashline
+      Neon & 0.24 & 0.6 & $69.8 \pm 0.5$ \\
+      F-Neon & 0.40 & 0.6 & $87.15 \pm 0.24$\\
+      \hdashline
+      Fanion-5 & 0.24 & 0.6 & $80.69 \pm 1.25$ \\
+      F-Fanion-5 & 0.40 & 0.6 & $86.66 \pm 0.65$\\
+      \arrayrulecolor{black}\hline
+    \end{tabular}%
+  }
+\end{table}
+
+We compare Muon, F-Muon, S-Muon, NSGD, SignSGD, Neon, and F-Neon, Fanion-5, and F-Fanion-5 on CIFAR-airbench.
+NSGD and SignSGD are not heavily tuned.
+The hyperparameters of Neon, F-Neon, Fanion-5, and F-Fanion-5 are taken from Muon and F-Muon.
+The parameters are in the table \tableref{tbl:cifar_lrs}.
+val\_accuracy is for the the weighted variant.
+However, to preserve the algorithms, we do not normalize the weights of the network each iteration.
+
+We log train and validation accuracy, and nuclear, gradient, and spectral norms of the gradients of conv1.weight and conv2.weight for all the layers.
+The gradient norms decrease only by the order of ten maximum during training, while train accuracy reaches full 100\% for Muon, S-Muon, and F-Muon.
+
+\begin{figure}[htbp]
+\floatconts
+  {fig:cifar_accs}
+  {\caption{Different optimizers on CIFAR airbench without weight normalization}}
+  {%
+    \subfigure[Train accuracy]{\label{fig:train_acc_cifar}%
+      \includegraphics[width=0.7\linewidth]{figs/cifar_logger/train_acc.pdf}}%
+    \vspace{1em}
+    \subfigure[Validation accuracy]{\label{fig:val_acc_cifar}%
+      \includegraphics[width=0.7\linewidth]{figs/cifar_logger/val_acc.pdf}}
+  }
+\end{figure}
+
+\begin{figure}[htbp]
+\floatconts
+  {fig:total_norms_cifar}
+  {\caption{Matrix norms of the whole gradient of CIFAR CNN}}
+  {%
+    \subfigure[Total Frobenius norm]{\label{fig:total_fro_norm_cifar}%
+      \includegraphics[width=0.7\linewidth]{figs/cifar_logger/total_frobenius.pdf}}%
+    \vspace{1em}
+    \subfigure[Total spectral norm]{\label{fig:total_spectral_norm_cifar}%
+      \includegraphics[width=0.7\linewidth]{figs/cifar_logger/total_spectral.pdf}}%
+    \vspace{1em}
+    \subfigure[Total nuclear norm]{\label{fig:total_nuclear_norm_cifar}%
+      \includegraphics[width=0.7\linewidth]{figs/cifar_logger/total_nuclear.pdf}}
+  }
+\end{figure}
+
+\begin{figure}[htbp]
+\floatconts
+  {fig:norms_conv1_vs_conv2}
+  {\caption{Matrix norms of layer2.conv1 (left) and layer2.conv2 (right) gradients of CIFAR CNN}}
+  {%
+    \subfigure[Frobenius norm of layer2.conv1]{\label{fig:fro_conv1}%
+      \includegraphics[width=0.45\linewidth]{figs/cifar_logger/layers_2_conv1_weight_frobenius.pdf}}%
+    \hfill
+    \subfigure[Frobenius norm of layer2.conv2]{\label{fig:fro_conv2}%
+      \includegraphics[width=0.45\linewidth]{figs/cifar_logger/layers_2_conv2_weight_frobenius.pdf}}%
+    \vspace{1em}
+    \subfigure[Spectral norm of layer2.conv1]{\label{fig:spec_conv1}%
+      \includegraphics[width=0.45\linewidth]{figs/cifar_logger/layers_2_conv1_weight_spectral.pdf}}%
+    \hfill
+    \subfigure[Spectral norm of layer2.conv2]{\label{fig:spec_conv2}%
+      \includegraphics[width=0.45\linewidth]{figs/cifar_logger/layers_2_conv2_weight_spectral.pdf}}%
+    \vspace{1em}
+    \subfigure[Nuclear norm of layer2.conv1]{\label{fig:nuc_conv1}%
+      \includegraphics[width=0.45\linewidth]{figs/cifar_logger/layers_2_conv1_weight_nuclear.pdf}}%
+    \hfill
+    \subfigure[Nuclear norm of layer2.conv2]{\label{fig:nuc_conv2}%
+      \includegraphics[width=0.45\linewidth]{figs/cifar_logger/layers_2_conv2_weight_nuclear.pdf}}
+  }
+\end{figure}
+
+\section{Technical details of the experiments}
+
+\textbf{CIFAR airbench} We used NVIDIA RTX A4000.
+
+\textbf{NanoGPT pretrain} We used the standard setting of 8 $\times$ H100 as documented in \citet{modded_nanogpt_2024}.
+
+\textbf{NanoGPT Fine-tuning.} Experiments were conducted on a single NVIDIA RTX 4090 (24GB) GPU using PyTorch~2.8 with CUDA~12.8 and cuDNN~9.0, running on a workstation equipped with an Intel Core i9-14900KS CPU and 128 GB of RAM.
+No distributed or mixed-hardware training setups were used, ensuring a strictly controlled and unbiased comparison across optimizers.
+
+A uniform training protocol was applied to all runs, including identical batch size, warm-up and cosine decay learning rate schedule, gradient clipping strategy, and validation split.
+Our optimization methods were integrated into the NanoGPT training loop without modifying the model architecture or preprocessing pipeline.
+Model quality was primarily evaluated using validation loss tracked over 200 training steps.
+
 \end{document}

### 2025-11-27 conf_article/neon.tex
New style file reached

@@
 Their most prominent members are {\it F-Muon} and {\it S-Muon}.
-We complement our theoretical analysis with an extensive empirical study of the algorithms across a wide range of tasks and settings, from which it follows that F-Muon and S-Muon are always on par with Muon and S-Muon, while on fine-tuning of NanoGPT and synthetic linear least squares even better than vanilla Muon.
+We complement our theoretical analysis with an extensive empirical study of the algorithms across a wide range of tasks and settings, from which it follows that F-Muon and S-Muon are always on par with Muon and S-Muon, while on synthetic linear least squares they are even better than vanilla Muon.
 

### 2025-11-27 code/cifar_bench/airbench_muon.py
Airbench done

@@ -25,7 +25,7 @@ import torchvision.transforms as T
 import math
 torch.backends.cudnn.benchmark = True
 
-from optimizers import MuonOrSign
+from optimizers import MuonOrSign, Neon
 
 #############################################
 #               Muon optimizer              #
@@ -428,12 +428,24 @@ def main(run, model):
                      dict(params=[model.head.weight], lr=head_lr, weight_decay=wd/head_lr)]
     optimizer1 = torch.optim.SGD(param_configs, momentum=0.85, nesterov=True, fused=True)
     # optimizer2 =  Muon(filter_params, lr=0.24, momentum=0.63, nesterov=True)
-    optimizer2 = NormalizedMuon(filter_params, lr=0.5, momentum=0.95, sgd_coeff=1, nesterov=True)
+    # optimizer2 = NormalizedMuon(filter_params, lr=0.5, momentum=0.95, sgd_coeff=1, nesterov=True)
     # optimizer2 = SignSGDMuon(filter_params, lr=1, momentum=0.95, nesterov=True, sgd_coeff=1, sign_lr_mult=0.003) # 91.54 +- 0.26%
 
     # optimizer2 =  SignSGDMuon(filter_params, lr=0.42, momentum=0.63, nesterov=True, sgd_coeff=0.5, sign_lr_mult=0.003)
     # optimizer2 = MuonOrSign(filter_params, lr=0.42, momentum=0.65, nesterov=True, sign_coeff=0.003) # 94.00%, for now a bad idea
 
+    # optimizer2 = Neon(
+    #         filter_params, neon_mode='kyfan', lr=0.24, momentum=0.6, nesterov=True, norm_weight=False # 69.8 +-0.50%
+    #     )
+    # optimizer2 = Neon(
+    #         filter_params, neon_mode='kyfan', lr=0.4, momentum=0.65, nesterov=True, sgd_coeff=0.5 # 87.15 +-0.24%
+    #     )
+    # optimizer2 = Neon(
+    #         filter_params, neon_mode='kyfan', lr=0.24, momentum=0.6, nesterov=True, sgd_coeff=0, k=5,
+    #     )
+    optimizer2= Neon(
+            filter_params, neon_mode='kyfan', lr=0.4, momentum=0.65, nesterov=True, sgd_coeff=0.5, k=5, #86.66 +-0.65%
+        )
     optimizers = [optimizer1, optimizer2]
     for opt in optimizers:
         for group in opt.param_groups:
[…]

### 2025-11-27 article_final/main.tex

@@
 Their most prominent members are \emph{F-Muon} and \emph{S-Muon}.
-We complement our theoretical analysis with an extensive empirical study of the algorithms across a wide range of tasks and settings, from which it follows that F-Muon and S-Muon are always on par with Muon, while on fine-tuning of NanoGPT and synthetic linear least squares they are even better than vanilla Muon optimizer.
+We complement our theoretical analysis with an extensive empirical study of the algorithms across a wide range of tasks and settings, from which it follows that F-Muon and S-Muon are always on par with Muon, while on synthetic linear least squares they are even better than vanilla Muon optimizer.
 \end{abstract}
@@
 Recall that Fanion-$k$ update is governed by the norm dual to the Ky Fan-$k$ norm, and NSGD corresponds to the Frobenius norm, which is dual to itself.
-Thus, by \corollaryref{corollary:lin_comb_alg}, our new algorithm is an LMO-algorithm corresponding to the unit ball of the norm $\norm{\cdot}_{\mathrm{F-KF-k}}$:
+Thus, by \corollaryref{corollary:lin_comb_alg}, our new algorithm is an LMO-algorithm corresponding to the unit ball of the norm $\norm{\cdot}_{\mathrm{F-KF-k}}^\dagger$:
 \begin{equation}\label{eq:fkfk_norm}
-	\norm{\cdot}_{\mathrm{F-KF-k}}^\dagger = \alpha \normkfk{\cdot} + (1-\alpha)\normf{\cdot}\,.
+	\norm{\cdot}_{\mathrm{F-KF-k}} = \alpha \normkfk{\cdot} + (1-\alpha)\normf{\cdot}\,.
 \end{equation}
@@
 
-The norm $\norm{\cdot}_{\mathrm{C-KF-k}}$ which produces this update in the LMO framework is again given by \corollaryref{corollary:lin_comb_alg}:
+The norm $\norm{\cdot}_{\mathrm{C-KF-k}}^\dagger$ that produces this update in the LMO framework is again given by \corollaryref{corollary:lin_comb_alg}:
 \begin{equation}\label{eq:ckfk_norm}
-	\norm{\cdot}_{\mathrm{C-KF-k}}^\dagger = \alpha \normkfk{\cdot} + \frac{1-\alpha}{\eta}\normc{\cdot}^\dagger\,.
+	\norm{\cdot}_{\mathrm{C-KF-k}} = \alpha \normkfk{\cdot} + \frac{1-\alpha}{\eta}\normc{\cdot}^\dagger\,.
 \end{equation}
@@
 
-In this article, we have generalized several successful algorithms, like Muon and Dion, to the LMO-based algorithms in the $\normkfk{\cdot}^\dagger$ norm.
-Also we have proposed the technique of ``regularizing'' the updates with NSGD, a trick to increase the robustness of the algorithms, which is motivated by the consideration of the $\normfstar{\cdot}$, $\normftwo{\cdot}$, and general $\norm{\cdot}_{\mathrm{F-KF-k}}$ norms.
-Generalizations of well-known norms and subsequent combinations of them may further improve performance of LMO-based algorithms.
-If a theory is developed that explains the discrepancy between performance of different algorithms based on the matrix norms, one will probably be able to intentionally construct norms optimal to given architectures and probably even their parameters.
-
-We suggest that future works that deal with the non-Euclidean LMO explain in corollaries to their convergence theorems the empirical superiority of Muon to other Fanions.
-Without that, it is hard to believe that the bounds are relevant for practitioners.
+In this article, we have generalized several successful algorithms, like Muon, Dion and $\nu$SAM, to the LMO-based algorithms in the $\normkfk{\cdot}^\dagger$ norm.
+Also we have proposed the technique of ``regularizing'' the updates with NSGD or SignSGD, a trick to increase the robustness of the algorithms, which is motivated by the consideration of the $\norm{\cdot}_{\mathrm{F-KF-k}}$ and $\norm{\cdot}_{\mathrm{C^\dagger-KF-k}}$ norms.
+We suggest that future works that deal with the non-Euclidean LMO explain in the corollaries to their convergence theorems the empirical superiority of Muon to other Fanions, and ideally accont for S-Fanion and F-Muon robustness as well.
+Without that, it is hard to believe that the bounds are relevant for practitioners, as we see by the notorious example of one-rank Neon.
 
@@
 IO suggested using the nuclear norm in the \citet{bernstein2024oldoptimizernewnorm} framework.
-DM presented the problem at the MIPT optimization course and helped with editing the article.
+DM presented the problem at the MIPT optimization course.
 IK suggested using composite norms (though not the ones that induce F-Fanions and S-Fanions) and refactored \sectionref{sec:conic_combo}.
@@
 AV conducted the finetuning of NanoGPT on Tiny Stories and wrote \sectionref{subsec:finetune}.
-All other work was done by AK: creating Fanions, F-Fanions, S-Fanions, experimenting on the Linear Least Squares, CIFAR-airbench and NanoGPT pretrain, and writing the manuscript.
+All other work was done by AK: constructing Fanions, F-Fanions, S-Fanions, experimenting on the Linear Least Squares, CIFAR-airbench and NanoGPT pretrain, and writing the manuscript.
 
@@
 
-\textbf{CIFAR airbench} We used NVIDIA RTX A4000.
+\paragraph{CIFAR airbench} We used NVIDIA RTX A4000.
 
-\textbf{NanoGPT pretrain} We used the standard setting of 8 $\times$ H100 as documented in \citet{modded_nanogpt_2024}.
+\paragraph{NanoGPT pretrain} We used the standard setting of 8 $\times$ H100 as documented in \citet{modded_nanogpt_2024}.
 
-\textbf{NanoGPT Fine-tuning.} Experiments were conducted on a single NVIDIA RTX 4090 (24GB) GPU using PyTorch~2.8 with CUDA~12.8 and cuDNN~9.0, running on a workstation equipped with an Intel Core i9-14900KS CPU and 128 GB of RAM.
+\paragraph{NanoGPT Fine-tuning.} Experiments were conducted on a single NVIDIA RTX 4090 (24GB) GPU using PyTorch~2.8 with CUDA~12.8 and cuDNN~9.0, running on a workstation equipped with an Intel Core i9-14900KS CPU and 128 GB of RAM.
 No distributed or mixed-hardware training setups were used, ensuring a strictly controlled and unbiased comparison across optimizers.

### 2025-11-27 article_final/main.tex
refactored abstract and introduction a little bit

@@
 	In this article, we explore the use of various matrix norms for optimizing functions of weight matrices, a crucial problem in training large language models.
-Moving beyond the spectral norm that underlies the Muon update, we leverage the duals to the Ky Fan $k$-norms to introduce a family of Muon-like algorithms we name \emph{Fanions}, which happen to be similar to Dion.
-Then working with the duals of convex combinations of the Ky Fan $k$-norms and the Frobenius norm or the $l_\infty$ norm, we construct the families of \emph{F-Fanions} and \emph{S-Fanions} respectively.
+Moving beyond the spectral norm underlying the Muon update, we leverage duals of the Ky Fan $k$-norms to introduce a family of Muon-like algorithms we name \emph{Fanions}, which are closely related to Dion.
+By working with duals of convex combinations of the Ky Fan $k$-norms with either the Frobenius norm or the $l_\infty$ norm, we construct the families of \emph{F-Fanions} and \emph{S-Fanions}, respectively.
 Their most prominent members are \emph{F-Muon} and \emph{S-Muon}.
-We complement our theoretical analysis with an extensive empirical study of the algorithms across a wide range of tasks and settings, from which it follows that F-Muon and S-Muon are always on par with Muon, while on synthetic linear least squares they are even better than vanilla Muon optimizer.
+We complement our theoretical analysis with an extensive empirical study of these algorithms across a wide range of tasks and settings, demonstrating that F-Muon and S-Muon consistently match Muon's performance, while outperforming vanilla Muon on synthetic linear least squares problems.
 \end{abstract}
@@
 
-That is what can be said from a practical point of view.
-From a theoretical perspective, a key innovation of Muon was its principled derivation of the update rule, which emerged as the solution to an optimization problem constrained by the RMS-to-RMS norm (scaled version of the spectral norm) \citep{bernstein2025deriving}.
+From a theoretical perspective, a key innovation of Muon was its principled derivation of the update rule, which emerged as the solution to an optimization problem constrained by the RMS-to-RMS norm (a scaled version of the spectral norm) \citep{bernstein2025deriving}.
 
@@
 
-\textit{In deriving Muon's update step, why should one constrain by the spectral or any other operator norm? How would alternative norms affect performance and computational cost?}
+\emph{In deriving Muon's update step, why should one constrain by the spectral or any other operator norm? How would alternative norms affect performance and computational cost?}
 
-In this article, we tackle this question by actually showing that there are many viable non-operator norms.
-We leverage the family of norms dual to Ky Fan $k$-norms to derive a new family of \textbf{Fanions}, algorithms with low-rank updates.
+In this article, we tackle this question by showing that there are many viable non-operator norms.
+We leverage the family of norms dual to Ky Fan $k$-norms to derive a new family of \emph{Fanions}, algorithms with low-rank updates.
 This approach theoretically explains the backbone of Dion's update \citep{ahn2025dion} and generalizes the memory-motivated application of the nuclear norm to Sharpness-Aware Minimization \citep{pethick2025sam}.
-As it was done with Muon, we come up with an effective procedure for computing Fanions' updates.
-Lanczos algorithm, which is described and compared with its competitors in \sectionref{sec:matrix-side}, is the most operation-effective algorithm, which, however, for now lacks an effective GPU- and PyTorch-friendly implementation.
+As was done with Muon, we develop an effective procedure for computing Fanions' updates.
+The Lanczos algorithm, which is described and compared with its competitors in \sectionref{sec:matrix-side}, is the most operation-efficient algorithm, though it currently lacks an effective GPU- and PyTorch-friendly implementation.
 
-Working with duals to conic combinations of dual norms, we construct the families of \textbf{F-Fanions} and \textbf{S-Fanions}, which are hybrids of Muon and NormalizedSGD and SignSGD, respectively.
+Working with dual norms and various convex combinations of norms, we construct the families of \emph{F-Fanions} and \emph{S-Fanions}, which are hybrids of Muon with NormalizedSGD and SignSGD, respectively.
 
-Then we compare the performances of the algorithm families on various model and real-world problems:
+We then compare the performance of these algorithm families on various model and real-world problems:
 \begin{itemize}
@@
 Our experiments reveal important insights into the role of matrix norms in optimization.
-First, we show on the example of Neon, the one-rank Fanion, that not every LMO-based algorithm is effective, despite the same asymptotics in the general bounds of \citet{kovalev2025understanding} and \citet{riabinin2025gluon}.
-This suggests that existing theoretical guarantees should be reworked to explain empirical performance.
+First, using the example of Neon (the rank-one Fanion), we show that not every LMO-based algorithm is effective, despite sharing the same asymptotics in the general bounds of \citet{kovalev2025understanding} and \citet{riabinin2025gluon}.
+This suggests that existing theoretical guarantees should be refined to better explain empirical performance.
 
@@
 On NanoGPT pre-training, F-Muon achieves 3.281 cross-entropy loss, while fully-tuned Muon achieves 3.279.
-Finally, S-Muon matches Muon on fine-tuning of NanoGPT on Tiny Stories, while F-Muon is far more resistant to the learning rate choice than Muon.
-These results show that Muon-like algorithms can maintain competitive performance even when the underlying norm constraint is significantly modified, answering affirmatively the central question posed above.
-Moreover, the tools from \sectionref{sec:conic_combo} give the researchers an unheard-of flexibility in designing algorithms that do not have to be modifications of Muon.
+Finally, S-Muon matches Muon on fine-tuning NanoGPT on Tiny Stories, while F-Muon is far more resistant to learning rate choice than Muon.
+These results show that Muon-like algorithms can maintain competitive performance even when the underlying norm constraint is significantly modified, providing an affirmative answer to the central question posed above.
+Moreover, the tools from \sectionref{sec:conic_combo} give researchers considerable flexibility in designing algorithms that need not be direct modifications of Muon.
 
@@
   {tbl:newton_schulz_reference}
-  {\caption{Newton–Schulz iteration on the same matrices (for reference).}}
+  {\caption{Newton-Schulz iterations on the same matrices (for reference).}}
   {%
@@
 
-\textbf{Improvements of Muon.} Regarding the first point, a large number of applications and improvements of Muon has been proposed in less than a year.
+\paragraph{Improvements of Muon} Regarding the first point, a large number of applications and improvements of Muon has been proposed in less than a year.
 \citet{liu2025muon} adapted the algorithm for training language models larger than NanoGPT.
@@
 
-\textbf{Theory behind Muon.} Regarding the second point, there has been a prolonged gap in theory behind Muon, simplistic derivation of \citet{bernstein2025deriving} based on \citet{bernstein2024oldoptimizernewnorm} excluded.
+\paragraph{Theory behind Muon} Regarding the second point, there has been a prolonged gap in theory behind Muon, simplistic derivation of \citet{bernstein2025deriving} based on \citet{bernstein2024oldoptimizernewnorm} excluded.
 This gap, as it seems to us, is not even now completely closed.
@@
 
-\textbf{The LMO and Error Feedback.} We already mentioned that rank-$k$ unsharded Dion without error feedback and scaling of the update is Fanion-$k$.
+\paragraph{The LMO and Error Feedback} We already mentioned that rank-$k$ unsharded Dion without error feedback and scaling of the update is Fanion-$k$.
 Since error feedback for Dion is very important as discovered by the ablation study in \citet{ahn2025dion}, F-Fanions and S-Fanions will benefit from it as well.
@@
 
-\textbf{The nuclear norm in the LMO.} As we found out only when transforming our results into an article, the nuclear norm has already been explored in the context of the linear minimization oracle.
+\paragraph{The nuclear norm in the LMO} As we found out only when transforming our results into an article, the nuclear norm has already been explored in the context of the linear minimization oracle.
 \citet{pethick2025sam} applied it to create $\nu$SAM, a new sharpness-aware minimization technique.
@@
 
-\textbf{Orthogonal research.} Fanions, F-Fanions, and S-Fanions benefit from the general theoretical description of \citet{riabinin2025gluon}, and better learning rates can be predicted by calculating the trajectory smoothness.
+\paragraph{Orthogonal research} Fanions, F-Fanions, and S-Fanions benefit from the general theoretical description of \citet{riabinin2025gluon}, and better learning rates can be predicted by calculating the trajectory smoothness.
 They could be transformed to Drop-Fanions by updating only some layers as in \citet{gruntkowska2025dropmuon}.

### 2025-11-28 article_final/main.tex
made some corrections considering language and notation

@@
 \end{equation}
-We equip the matrix space $\Rmn$ with a standard dot product $\<\cdot, \cdot> \to \R$ and a norm $\norm{\cdot}\colon \Rmn \to \R_+$, which does not have to coincide with the Frobenius norm $\normf{\cdot}$.
+We equip the matrix space $\Rmn$ with a standard dot product $\<\cdot, \cdot> \to \R$ and a norm $\norm{\cdot}\colon \Rmn \to \R_+$, which does not have to coincide with the Frobenius norm $\normf{\cdot} = \<\cdot,\cdot>$.
 The dual norm $\norm{\cdot}^\dagger\colon \Rmn \to \R_+$ that is associated with $\norm{\cdot}$ is defined as
@@
 \end{equation}
-where $\mM^k$ is a gradient (or a momentum) of $F$ in $\mX^k$ and $\cS \subset \Rmn$ is some set. The update of the algorithm is defined as follows:
+where $\mM^k$ is the gradient (or momentum) of $F$ at $\mX^k$ and $\cS \subset \Rmn$ is some set. The update of the algorithm is defined as follows:
 \begin{equation}\label{eq:simple_update}
@@
 
-We are particularly interested in the case when $\cS$ is a unit ball in the norm $\norm{\cdot}$:
-
+We are particularly interested in the case when $\cS$ is a unit ball in some norm $\norm{\cdot}$:
 \[
@@
 \]
-and update for $\mX^k$ in \equationref{eq:simple_update} simplifies to
+and the update for $\mX^k$ in \equationref{eq:simple_update} simplifies to
 \begin{equation}\label{eq:our_update}
@@
 
-Using this formula it is easy to compute updates for algorithms induced by various norms $\norm{\cdot}$:
+Using this formula, it is easy to compute updates for algorithms induced by various norms $\norm{\cdot}$:
 
@@
 \begin{equation}\label{eq:nsgd_update}
-	\mX^{k+1} = \mX^{k} - \eta_k \frac{\mM^k}{\normf{\mM^k}}\,,
+	\mX^{k+1} = \mX^{k} - \gamma_k \frac{\mM^k}{\normf{\mM^k}}\,,
 \end{equation}
@@
 \begin{equation}\label{eq:muon_update}
-	\mX^{k+1} = \mX^{k} - \eta_k \mU \mV^\top\,,
+	\mX^{k+1} = \mX^{k} - \gamma_k \mU \mV^\top\,,
 \end{equation}
@@
 \begin{equation}\label{eq:signsgd_update}
-	\mX^{k+1} = \mX^{k} - \eta_k \sign(\mM^k)\,,
+	\mX^{k+1} = \mX^{k} - \gamma_k \sign(\mM^k)\,,
 \end{equation}
@@
 	\begin{equation}\label{eq:neon_update}
-		\mX^{k+1} = \mX^{k} - \eta_k \vu_1 \vv_1^\top\,.
+		\mX^{k+1} = \mX^{k} - \gamma_k \vu_1 \vv_1^\top\,.
 	\end{equation}
@@
 
-We name the derived algorithm \emph{Neon}. In \sectionref{sec:matrix-side}, we will discuss how to compute the Neon's update.
+We name the derived algorithm \emph{Neon}. In \sectionref{sec:matrix-side}, we will discuss how to compute Neon's update.
 
@@
 
-However, Schatten norms do not fill the gap rank: indeed, when $p > 1$, the rank of the update is full, while when $p=1$ ($p \rightarrow 1$) it is one.
-Moreover, it is not clear how to calculate the update for $p \neq 1, 2, \infty$: it seems one has to know all $\sigma_i$ to compute the update, which makes the problem as hard as the full SVD.
+However, Schatten norms do not fill the rank gap: indeed, when $p > 1$, the rank of the update is full, while when $p=1$ ($p \rightarrow 1$) it is one.
+Moreover, it is not clear how to calculate the update for $p \neq 1, 2, \infty$: it seems one has to know all $\sigma_i$ to compute the update, which makes the problem as hard as computing the full SVD.
 
 \subsubsection{Ky Fan norms}
-There is another family of matrix norms, which might help us: Ky Fan norms.
-For $k \in \{1, \dots, \min(m, n)\}$, the Ky Fan $k$-norm $\normkfk{\cdot}$ is $\sum_{i=1}^{k}\sigma_i$, i.e.
-the sum of the $k$ largest singular values of the matrix.
+There is another family of matrix norms that might help us: Ky Fan norms.
+For $k \in \{1, \dots, \min(m, n)\}$, the Ky Fan $k$-norm $\normkfk{\cdot}$ is $\sum_{i=1}^{k}\sigma_i$, i.e., the sum of the $k$ largest singular values of the matrix.
 Special cases of the Ky Fan $k$-norm are the Ky Fan $1$-norm, which is the spectral norm, and the Ky Fan $\min\{m, n\}$-norm, which is the nuclear norm.
@@
 These updates constitute the family of \emph{Fanions}, LMO-based algorithms under the $\normkfk{\mM^t}^\dagger$ norms.
-The algorithm for a particular $k$ we will call \emph{Fanion-$k$}.
-Thus, Muon is a Fanion-$\min\{m,n\}$, while Neon is a Fanion-1.
-Moreover, the unsharded version of the rank-$r$ Dion (Algorithm 1 from \citet{ahn2025dion}) without the error feedback and without scaling of the update is actually a Fanion-$r$ (see \citet{pethick2025understandingdion}, where Dion is written down in a notation more similar to ours).
+We call the algorithm for a particular $k$ \emph{Fanion-$k$}.
+Thus, Muon is Fanion-$\min\{m,n\}$, while Neon is Fanion-1.
+Moreover, the unsharded version of rank-$r$ Dion (Algorithm 1 from \citet{ahn2025dion}) without error feedback and without scaling of the update is actually Fanion-$r$ (see \citet{pethick2025understandingdion}, where Dion is written in a notation more similar to ours).
 
@@
 
-Simply applying norm dual to the Ky Fan $k$-norm however, is not enough to provide family of algorithms diverse enough for our cause.
-Next we consider linear combinations of algorithms from LMO framework, which happen to be LMO-algorithms too, as we show in succeeding paragraphs.
+Simply applying the norm dual to the Ky Fan $k$-norm, however, is not enough to provide a family of algorithms diverse enough for our purposes.
+Next, we consider linear combinations of algorithms from the LMO framework, which happen to be LMO-algorithms as well, as we show in the following paragraphs.
 

### 2025-11-28 article_final/main.tex
maked notation more consistent by always using t for iteration number

@@
 \begin{equation}\label{eq:lmo}
-	\mathrm{LMO}(\mM^k) \in \argmin_{\mD \in \cS} \<\mM^k, \mD>\,,
+	\mathrm{LMO}(\mM^t) \in \argmin_{\mD \in \cS} \<\mM^t, \mD>\,,
 \end{equation}
-where $\mM^k$ is the gradient (or momentum) of $F$ at $\mX^k$ and $\cS \subset \Rmn$ is some set. The update of the algorithm is defined as follows:
+where $\mM^t$ is the gradient (or momentum) of $F$ at $\mX^t$ and $\cS \subset \Rmn$ is some set. The update of the algorithm is defined as follows:
 \begin{equation}\label{eq:simple_update}
-	\mX^{k + 1} = \mX^k + \gamma_k\mathrm{LMO}\left(\mM^k\right)\,.
+	\mX^{t + 1} = \mX^t + \gamma_t\mathrm{LMO}\left(\mM^t\right)\,.
 \end{equation}
@@
 \[
-\argmin_{\mD \in \cS} \<\mM^k, \mD> = - \{ \mD \in \cB_1 \mid \<\mM^k, \mD> = \norm{\mM^k}^\dagger\}\,,
+\argmin_{\mD \in \cS} \<\mM^t, \mD> = - \{ \mD \in \cB_1 \mid \<\mM^t, \mD> = \norm{\mM^t}^\dagger\}\,,
 \]
-and the update for $\mX^k$ in \equationref{eq:simple_update} simplifies to
+and the update for $\mX^t$ in \equationref{eq:simple_update} simplifies to
 \begin{equation}\label{eq:our_update}
-	\mX^{k+1} = \mX^{k} - \gamma_k \{\mD \in \cB_1 \mid \<\mM^k, \mD> = \norm{\mM^k}^\dagger\}\,.
+	\mX^{t+1} = \mX^{t} - \gamma_t \{\mD \in \cB_1 \mid \<\mM^t, \mD> = \norm{\mM^t}^\dagger\}\,.
 \end{equation}
@@
 \begin{equation}\label{eq:nsgd_update}
-	\mX^{k+1} = \mX^{k} - \gamma_k \frac{\mM^k}{\normf{\mM^k}}\,,
+	\mX^{t+1} = \mX^{t} - \gamma_t \frac{\mM^t}{\normf{\mM^t}}\,,
 \end{equation}
@@
 \begin{equation}\label{eq:muon_update}
-	\mX^{k+1} = \mX^{k} - \gamma_k \mU \mV^\top\,,
+	\mX^{t+1} = \mX^{t} - \gamma_t \mU \mV^\top\,,
 \end{equation}
 which is Muon without the $\sqrt{m/n}$ factor.
-Here, $\mM^k = \mU \mSigma \mV^\top$ is the Singular Value Decomposition (SVD) of $\mM^k$ ($\mU = [\vu_1, \vu_2, \dots, \vu_r]$, $\mSigma = \diag(\sigma_1, \sigma_2, \dots, \sigma_r)$, and $\mV = [\vv_1, \vv_2, \dots, \vv_r]$).
+Here, $\mM^t = \mU \mSigma \mV^\top$ is the Singular Value Decomposition (SVD) of $\mM^t$ ($\mU = [\vu_1, \vu_2, \dots, \vu_r]$, $\mSigma = \diag(\sigma_1, \sigma_2, \dots, \sigma_r)$, and $\mV = [\vv_1, \vv_2, \dots, \vv_r]$).
 Muon can be recovered by taking the RMS-to-RMS operator norm $\sqrt{\frac{n}{m}}\norms{\cdot}$.
@@
 \begin{equation}\label{eq:signsgd_update}
-	\mX^{k+1} = \mX^{k} - \gamma_k \sign(\mM^k)\,,
+	\mX^{t+1} = \mX^{t} - \gamma_t \sign(\mM^t)\,,
 \end{equation}
 which recovers SignSGD~\citep{bernstein2018signsgd}.
-Here, $\sign(\mM^k)$ denotes the element-wise sign function.
+Here, $\sign(\mM^t)$ denotes the element-wise sign function.
 SignSGD is particularly notable for its communication efficiency in distributed training, as it compresses gradients to 1-bit per parameter.
@@
 
-\subsection{\texorpdfstring{$\normn{\mM^{k}}$ and Neon}{Neon}}
-After considering $\normf{\mM^k}$ and $\norms{\mM^k}$, it is natural to look at the nuclear norm $\normn{\mM^k}$.
+\subsection{\texorpdfstring{$\normn{\mM^{t}}$ and Neon}{Neon}}
+After considering $\normf{\mM^t}$ and $\norms{\mM^t}$, it is natural to look at the nuclear norm $\normn{\mM^t}$.
 \begin{lemma}\label{lemma:neon_update}
@@
 	\begin{equation}\label{eq:neon_update}
-		\mX^{k+1} = \mX^{k} - \gamma_k \vu_1 \vv_1^\top\,.
+		\mX^{t+1} = \mX^{t} - \gamma_t \vu_1 \vv_1^\top\,.
 	\end{equation}
@@
 \begin{proof}
-Since $\norm{\cdot}^\dagger = \norms{\cdot}$, the goal is to reach $\norms{\mM^k} = \sigma_1$.
+Since $\norm{\cdot}^\dagger = \norms{\cdot}$, the goal is to reach $\norms{\mM^t} = \sigma_1$.
 Let us note that $\Delta = \vu_1 \vv_1^\top$ delivers this value.
@@
 \[
-\<\mM^k, \Delta> =\<\mU \mSigma \mV^\top, \vu_1 \vv_1^\top> = \tr \diag({\sigma_1, 0, \dots, 0}) = \norms{\mM^k}\,,
+\<\mM^t, \Delta> =\<\mU \mSigma \mV^\top, \vu_1 \vv_1^\top> = \tr \diag({\sigma_1, 0, \dots, 0}) = \norms{\mM^t}\,,
 \]
@@
 
-\subsection{\texorpdfstring{$\normkfk{\mM^k}^\dagger$ and Muon, Neon, and Centralized Dion without error feedback}{Muon, Neon, and Centralized Dion without error feedback}}
+\subsection{\texorpdfstring{$\normkfk{\mM^t}^\dagger$ and Muon, Neon, and Centralized Dion without error feedback}{Muon, Neon, and Centralized Dion without error feedback}}
 
@@
 \citet{cesista2025schattenp} considered Schatten-$p$ norms:
-\[\norm{\mM^k}_{S_p} = \left(\sum_{i=1}^{\min(m, n)}\sigma_i^p\right)^{1/p}\,,
+\[\norm{\mM^t}_{S_p} = \left(\sum_{i=1}^{\min(m, n)}\sigma_i^p\right)^{1/p}\,,
 \]
@@
 \[
-\mX^{k+1} = \mX^k - \eta_k \mU \frac{\diag\left(\sigma_1^{q-1},\dots, \sigma_{\min(m,n)}^{q-1}\right)}{\left(\sum_{i=1}^{\min(m,n)}{\sigma_i^q}\right)^{\frac{q-1}{q}}}\mV^\top
+\mX^{t+1} = \mX^t - \eta_t \mU \frac{\diag\left(\sigma_1^{q-1},\dots, \sigma_{\min(m,n)}^{q-1}\right)}{\left(\sum_{i=1}^{\min(m,n)}{\sigma_i^q}\right)^{\frac{q-1}{q}}}\mV^\top
 \]
@@
 	\[
-	\mX^{k + 1} - \mX^{k} = \gamma_k \mathrm{LMO}_{i}(\mM^{k})\,,
+	\mX^{t + 1} - \mX^{t} = \gamma_t \mathrm{LMO}_{i}(\mM^{t})\,,
 	\]
@@
 	\[
-	\mX^{k + 1} - \mX^{k} = \gamma_k \sum_{i = 1}^n \alpha_i \mathrm{LMO}_{\norm{\cdot}_{(i)}}(\mM^k)
+	\mX^{t + 1} - \mX^{t} = \gamma_t \sum_{i = 1}^n \alpha_i \mathrm{LMO}_{\norm{\cdot}_{(i)}}(\mM^t)
 	\]
@@
 \begin{equation}\label{eq:nsgd_muon_update}
-		\mX^{t+1} = \mX^{t} - \gamma_k \left(\alpha \sum_{i = 1}^k \vu_i \vv_i^\top + (1-\alpha)\frac{\mM^k}{\normf{\mM^k}}\right)\,.
+		\mX^{t+1} = \mX^{t} - \gamma_t \left(\alpha \sum_{i = 1}^k \vu_i \vv_i^\top + (1-\alpha)\frac{\mM^t}{\normf{\mM^t}}\right)\,.
 \end{equation}
@@
 \begin{equation}\label{eq:sign_muon_update}
-		\mX^{t+1} = \mX^{t} - \gamma_k \left(\alpha \sum_{i = 1}^k \vu_i \vv_i^\top + (1-\alpha)\eta\sign(\mM^k)\right)\,,
+		\mX^{t+1} = \mX^{t} - \gamma_t \left(\alpha \sum_{i = 1}^k \vu_i \vv_i^\top + (1-\alpha)\eta\sign(\mM^t)\right)\,,
 \end{equation}
@@
 We employ the thick-restarted Lanczos method for the symmetric eigenvalue problem (thick-restart Lanczos, TRLan) to compute the low-rank updates of Fanions.
-We apply TRLan to either $\mM^{k\top}\mM^k$ or $\mM^k \mM^{k\top}$ (whichever matrix is smaller).
+We apply TRLan to either $\mM^{t\top}\mM^t$ or $\mM^t \mM^{t\top}$ (whichever matrix is smaller).
 We use the CuPy implementation of \texttt{cupy.sparse.linalg.svds} \citep{cupy_svds_ref} which internally relies on TRLan \citep{simonz1998thick}.
@@
 The optimization framework we have described is applied in a layer-wise fashion.
-At each iteration $k$, a stochastic gradient $g(\mX^k, \xi^k)$ is computed using a mini-batch of data with the $\xi^k$ noise via backpropagation.
-This yields a separate gradient component, $\mG_i^k$, for each matrix $\mX_i$.
-The LMO-based update rule is then applied to each matrix $\mX_i$ using its corresponding gradient component $\mG_i^k$.
+At each iteration $t$, a stochastic gradient $g(\mX^t, \xi^t)$ is computed using a mini-batch of data with the $\xi^t$ noise via backpropagation.
+This yields a separate gradient component, $\mG_i^t$, for each matrix $\mX_i$.
+The LMO-based update rule is then applied to each matrix $\mX_i$ using its corresponding gradient component $\mG_i^t$.
 
@@
   \begin{aligned}
-    \mM^{k}     & = \alpha_{k} g(\mX^k, \xi^k) + (1 - \alpha_k)\mM^{k - 1}\,, \\
-    \mX^{k + 1} & = \mX^k + \gamma_k\mathrm{LMO}(\mM^{k})\,.
+    \mM^{t}     & = \alpha_{t} g(\mX^t, \xi^t) + (1 - \alpha_t)\mM^{t - 1}\,, \\
+    \mX^{t + 1} & = \mX^t + \gamma_t\mathrm{LMO}(\mM^{t})\,.
   \end{aligned}

### 2025-11-28 article_final/main.tex
Merging

@@
 \[
-\argmin_{\mD \in \cS} \<\mM^t, \mD> = - \{ \mD \in \cB_1 \mid \<\mM^t, \mD> = \norm{\mM^t}^\dagger\}\,,
+\argmin_{\mD \in \cS} \<\mM^t, \mD> = - \{ \mD \in \cB_{\norm{\cdot}} \mid \<\mM^t, \mD> = \norm{\mM^t}^\dagger\}\,,
 \]
@@
 \[
-\mX^{t+1} = \mX^t - \eta_t \mU \frac{\diag\left(\sigma_1^{q-1},\dots, \sigma_{\min(m,n)}^{q-1}\right)}{\left(\sum_{i=1}^{\min(m,n)}{\sigma_i^q}\right)^{\frac{q-1}{q}}}\mV^\top
+\mX^{t+1} = \mX^t - \gamma_t \mU \frac{\diag\left(\sigma_1^{q-1},\dots, \sigma_{\min(m,n)}^{q-1}\right)}{\left(\sum_{i=1}^{\min(m,n)}{\sigma_i^q}\right)^{\frac{q-1}{q}}}\mV^\top
 \]
@@
 \[
-\mX^{t+1} = \mX^{t} - \eta_t \vu_1 \vv_1^\top\text{ or }\mX^{t+1} = \mX^{t} - \frac{\eta_t}{k}\mU \mV^\top\,,
+\mX^{t+1} = \mX^{t} - \gamma_t \vu_1 \vv_1^\top\text{ or }\mX^{t+1} = \mX^{t} - \frac{\gamma_t}{k}\mU \mV^\top\,,
 \]
@@
 \subsubsection{Duals to Ky Fan norms}
-While Schatten norms are closed under dualization and $\norm{\mM^k}_{S_p}^\dagger = \norm{\mM^k}_{S_q}$, for Ky Fan norms it is not generally the case.
+While Schatten norms are closed under dualization and $\norm{\mM^t}_{S_p}^\dagger = \norm{\mM^t}_{S_q}$, for Ky Fan norms it is not generally the case.
 $k = 1$ and $k=\min(m, n)$ are exceptional: for $k=1$ the dual norm is $\normn{\cdot}$ with $k = \min(m, n)$ and for $k = \min(m, n)$ the dual norm is $\norms{\cdot}$ with $k=1$.
-For each other $k$, $\normkfk{\cdot} \neq \norm{\cdot}_{\text{KF-}k'}$ for any $k'$: $k'=\min(m, n)$ and $k'= 1$ correspond to the already discussed cases, while for other $k'$ one can change $\norm{\mM^k}_{\text{KF-}k'}$ by moving a small value between $\sigma_{k'}$ and $\sigma_{k'+1}$.
-During this change, $\norms{\mM^k}$ and $\normn{\mM^k}$ will remain constant, hence $\normkfk{\mM^k}$ will not change as well.
+For each other $k$, $\normkfk{\cdot} \neq \norm{\cdot}_{\text{KF-}k'}$ for any $k'$: $k'=\min(m, n)$ and $k'= 1$ correspond to the already discussed cases, while for other $k'$ one can change $\norm{\mM^t}_{\text{KF-}k'}$ by moving a small value between $\sigma_{k'}$ and $\sigma_{k'+1}$.
+During this change, $\norms{\mM^t}$ and $\normn{\mM^t}$ will remain constant, hence $\normkfk{\mM^t}$ will not change as well.
 
@@
 	\begin{equation}\label{eq:ky_fan_update}
-		\mX^{t+1} = \mX^{t} - \eta \sum_{i=1}^{k}\vu_i \vv_i^\top
+		\mX^{t+1} = \mX^{t} - \gamma_t \sum_{i=1}^{k}\vu_i \vv_i^\top
 	\end{equation}

### 2025-11-28 code/cifar_bench/base_airbench.py

@@ -417,7 +417,7 @@ def main(run, model):
     # optimizer2 = Dion(filter_params, lr=0.45, momentum=0.65, rank=10, momentum_decay=0.9, sgd_coeff=0) # 84.5%, with 0.4% variance - wrong EF
     # optimizer2 = Dion(filter_params, lr=0.45, momentum=0.65, rank=20, momentum_decay=0.9, sgd_coeff=0) # 89.3%, with 0.2% variance - wrong EF
     
-    optimizer2 = SignSGDMuon(filter_params, lr=1, momentum=0.95, nesterov=True, sgd_coeff=1, sign_lr_mult=0.003) # 91.41 +- 0.33% BASELINE
+    # optimizer2 = SignSGDMuon(filter_params, lr=1, momentum=0.95, nesterov=True, sgd_coeff=1, sign_lr_mult=0.003) # 91.41 +- 0.33% BASELINE
 
     # optimizer2 = SignSGDMuon(filter_params, lr=0.4, momentum=0.6, nesterov=True, sgd_coeff=1, sign_lr_mult=0.005) # 90.94 +- 0.4%
     # optimizer2 = SignSGDMuon(filter_params, lr=0.003, momentum=0.6, nesterov=True, sgd_coeff=1, sign_lr_mult=1) # 90.94 +- 0.4%
@@ -426,8 +426,7 @@ def main(run, model):
     # optimizer2 = SignSGDMuon(filter_params, lr=0.4, momentum=0.6, nesterov=True, sgd_coeff=1, sign_lr_mult=0.009) # 89.66 +- 0.33%
     # optimizer2 = SGDMuon(filter_params, lr=0.24, momentum=0.6, nesterov=True, sgd_coeff=0.1) # 87% - does not work well, because it's not an LMO algorithm
     # optimizer2 = MuonOrSign(filter_params, lr=0.42, momentum=0.65, nesterov=True, sign_coeff=0.003) # 94.00% - see airbench_muon.py - no benefit
-    # optimizer2 = MuonSignedUpdate(filter_params, lr=0.003, momentum=0.8, nesterov=True) # 93.14 +- 0.17%
-    
+    optimizer2 = MuonSignedUpdate(filter_params, lr=0.005, momentum=0.8, nesterov=True) # 93.57 +- 0.11 - AWESOME
 
[…]

### 2025-11-28 code/gpt_pretrain/medium_f_s_muon.py

@@ -1,0 +1,698 @@
+import os
+import sys
+with open(sys.argv[0]) as f:
+    code = f.read() # read the code of this file ASAP, for logging
+import uuid
+import time
+import copy
+from dataclasses import dataclass
+from functools import lru_cache
+from pathlib import Path
+
+os.environ["PYTORCH_CUDA_ALLOC_CONF"] = "expandable_segments:True"
+import torch
+torch.empty(1, device="cuda", requires_grad=True).backward() # prevents a bug on some systems
+from torch import Tensor, nn
+import torch.nn.functional as F
+import torch.distributed as dist
+# use of FlexAttention contributed by @KoszarskyB
+from torch.nn.attention.flex_attention import BlockMask, flex_attention
+torch._inductor.config.coordinate_descent_tuning = True # we have banned this flag for new records because it causes compilation to take 30min
+torch._dynamo.config.compiled_autograd = True
+
+# -----------------------------------------------------------------------------
+# Muon optimizer
+
+def zeropower_via_newtonschulz5(G: Tensor) -> Tensor:
+    """
+    Newton-Schulz iteration to compute the zeroth power / orthogonalization of G. We opt to use a
+    quintic iteration whose coefficients are selected to maximize the slope at zero. For the purpose
+    of minimizing steps, it turns out to be empirically effective to keep increasing the slope at
+    zero even beyond the point where the iteration no longer converges all the way to one everywhere
+    on the interval. This iteration therefore does not produce UV^T but rather something like US'V^T
+    where S' is diagonal with S_{ii}' ∈ [1 - l, 1 + r], which turns out not to hurt model
+    performance at all relative to UV^T, where USV^T = G is the SVD.
+    """
+    assert G.ndim >= 2 # batched Muon implementation by @scottjmaddox, and put into practice in the record by @YouJiacheng
+    X = G.bfloat16()
+    if G.size(-2) > G.size(-1):
+        X = X.mT
+
+    # Ensure spectral norm is at most 1
+    X = X / (X.norm(dim=(-2, -1), keepdim=True) + 1e-7)
+    # Perform the NS iterations
+    for a, b, c in [
+        (4.0848, -6.8946, 2.9270),
+        (3.9505, -6.3029, 2.6377),
+        (3.7418, -5.5913, 2.3037),
+        (2.8769, -3.1427, 1.2046),
[…]

### 2025-11-28 article_final/main.tex
Minor improvements

@@
 
-\title[Ky Fan Norms for Matrix Optimization]{The Ky Fan Norms and Beyond: Dual Norms and Combinations for Matrix Optimization}
+\title[Ky Fan Norms, Duals, and Norm Combinations for Matrix Optimization]{The Ky Fan Norms and Beyond: Dual Norms and Combinations for Matrix Optimization}
 
@@
 \begin{equation}\label{eq:our_update}
-	\mX^{t+1} = \mX^{t} - \gamma_t \{\mD \in \cB_1 \mid \<\mM^t, \mD> = \norm{\mM^t}^\dagger\}\,.
+	\mX^{t+1} = \mX^{t} - \gamma_t \{\mD \in \cB_{\norm{\cdot}} \mid \<\mM^t, \mD> = \norm{\mM^t}^\dagger\}\,.
 \end{equation}
@@
 \]
-depending on which one better minimizes the linear function $L(\mX) \equiv F(\mX^t) + \mM^t(\mX - \mX^t)$.
+depending on which one better minimizes the linear function $L(\mX) \equiv F(\mX^t) + \<\mM^t, \mX - \mX^t>$.
 

### 2025-11-28 article_final/main.tex
merged main.tex

@@
 
-\section{Duals to Ky Fan Norms Instead of the Spectral Norm}
+\section{Beyond the Spectral Norm: Fanions}
 
-\subsection{\texorpdfstring{$\normn{\mM^{t}}$ and Neon}{Neon}}
-After considering $\normf{\mM^t}$ and $\norms{\mM^t}$, it is natural to look at the nuclear norm $\normn{\mM^t}$.
+\subsection{The Rank Gap Problem}
+
+Having examined the Frobenius norm (NSGD), spectral norm (Muon), and Chebyshev norm (SignSGD), all of which produce full-rank updates, we naturally turn to the nuclear norm $\normn{\cdot}$.
+
 \begin{lemma}\label{lemma:neon_update}
@@
 \begin{proof}
-Since $\norm{\cdot}^\dagger = \norms{\cdot}$, the goal is to reach $\norms{\mM^t} = \sigma_1$.
-Let us note that $\Delta = \vu_1 \vv_1^\top$ delivers this value.
+Since $\normn{\cdot}^\dagger = \norms{\cdot}$, the goal is to reach $\norms{\mM^t} = \sigma_1$.
+Note that $\Delta = \vu_1 \vv_1^\top$ delivers this value.
 Indeed, $\normn{\Delta}=1$ and by the trace property and orthogonality of the singular vectors,
@@
 
-We name the derived algorithm \emph{Neon}. In \sectionref{sec:matrix-side}, we will discuss how to compute Neon's update.
+We call this algorithm \emph{Neon}.
+The nuclear norm thus yields rank-one updates, in stark contrast to the full-rank updates of Muon, NSGD, and SignSGD.
+This raises a natural question: can we derive algorithms with updates of intermediate ranks?
 
-\subsection{\texorpdfstring{$\normkfk{\mM^t}^\dagger$ and Muon, Neon, and Centralized Dion without error feedback}{Muon, Neon, and Centralized Dion without error feedback}}
-
-Neon's and Muon's updates seem to be complete opposites: one has rank one, while the other is full-rank. It would be interesting to derive algorithms with updates of intermediate ranks.
-
-\subsubsection{Schatten norms}
+\paragraph{Schatten norms}
 \citet{cesista2025schattenp} considered Schatten-$p$ norms:
@@
 \]
-for $p$ and $q$ such that: $\frac{1}{p} + \frac{1}{q} = 1$.
-This formula recovers Neon when $p\rightarrow1$ provided that $\sigma_1 > \sigma_2$, which is true on real data; NSGD when $p=2$, and Muon when $p\rightarrow\infty$.
+where $p$ and $q$ satisfy $\frac{1}{p} + \frac{1}{q} = 1$.
+This formula recovers Neon when $p\rightarrow1$ (provided $\sigma_1 > \sigma_2$, which holds on real data), NSGD when $p=2$, and Muon when $p\rightarrow\infty$.
 
-However, Schatten norms do not fill the rank gap: indeed, when $p > 1$, the rank of the update is full, while when $p=1$ ($p \rightarrow 1$) it is one.
-Moreover, it is not clear how to calculate the update for $p \neq 1, 2, \infty$: it seems one has to know all $\sigma_i$ to compute the update, which makes the problem as hard as computing the full SVD.
+However, Schatten norms do not fill the rank gap: when $p > 1$, the update has full rank, while when $p=1$ (or $p \rightarrow 1$), it has rank one.
+Moreover, computing the update for $p \neq 1, 2, \infty$ appears to require knowing all $\sigma_i$, making the problem as hard as computing the full SVD.
 
-\subsubsection{Ky Fan norms}
-There is another family of matrix norms that might help us: Ky Fan norms.
-For $k \in \{1, \dots, \min(m, n)\}$, the Ky Fan $k$-norm $\normkfk{\cdot}$ is $\sum_{i=1}^{k}\sigma_i$, i.e., the sum of the $k$ largest singular values of the matrix.
-Special cases of the Ky Fan $k$-norm are the Ky Fan $1$-norm, which is the spectral norm, and the Ky Fan $\min\{m, n\}$-norm, which is the nuclear norm.
+\paragraph{Ky Fan norms}
+Another family of matrix norms offers a potential solution: the Ky Fan norms.
+For $k \in \{1, \dots, \min(m, n)\}$, the Ky Fan $k$-norm $\normkfk{\cdot}$ equals $\sum_{i=1}^{k}\sigma_i$, the sum of the $k$ largest singular values.
+Notable special cases include the Ky Fan $1$-norm (the spectral norm) and the Ky Fan $\min\{m, n\}$-norm (the nuclear norm).
 
-Let us derive the update for an arbitrary $k$.
-Since $\normkfk{\cdot}^\dagger = \max\{\frac{1}{k} \normn{\cdot}, \norms{\cdot}\}$ (see \citet{bhatia2013matrix}, p.
+Deriving the update for arbitrary $k$: Since $\normkfk{\cdot}^\dagger = \max\{\frac{1}{k} \normn{\cdot}, \norms{\cdot}\}$ (see \citet{bhatia2013matrix}, p.
 96), the goal is to reach $\max\{\frac{1}{k}\sum_{i=1}^{\min(m,n)}\sigma_i, \sigma_1\}$.
-$\Delta = \vu_1 \vv_1^\top$ from Neon delivers $\sigma_1$, while $\Delta = \frac{1}{k}\sum_{i=1}^{\min(m,n)}\vu_i \vv_i^\top$ from Muon (with an extra factor of $1/k$) delivers $\frac{1}{k}\sum_{i=1}^{\min(m,n)}\sigma_i$.
+The Neon update $\Delta = \vu_1 \vv_1^\top$ achieves $\sigma_1$, while the scaled Muon update $\Delta = \frac{1}{k}\sum_{i=1}^{\min(m,n)}\vu_i \vv_i^\top$ achieves $\frac{1}{k}\sum_{i=1}^{\min(m,n)}\sigma_i$.
 Thus, the update is either
@@
 \]
-depending on which one better minimizes the linear function $L(\mX) \equiv F(\mX^t) + \<\mM^t, \mX - \mX^t>$.
+depending on which better minimizes the linear function $L(\mX) \equiv F(\mX^t) + \mM^t(\mX - \mX^t)$.
 
-The rank gap is not closed by the Ky Fan norms because the obtained update is either one-rank or full-rank.
+The Ky Fan norms thus fail to close the rank gap: the resulting update is either rank-one or full-rank.
 
-\subsubsection{Duals to Ky Fan norms}
+\subsection{Solution: Duals to Ky Fan Norms}
+
 While Schatten norms are closed under dualization and $\norm{\mM^t}_{S_p}^\dagger = \norm{\mM^t}_{S_q}$, for Ky Fan norms it is not generally the case.
@@
 During this change, $\norms{\mM^t}$ and $\normn{\mM^t}$ will remain constant, hence $\normkfk{\mM^t}$ will not change as well.
+
+The key insight is to work with the duals of Ky Fan norms:
 
@@
 \end{proof}
+
 These updates constitute the family of \emph{Fanions}, LMO-based algorithms under the $\normkfk{\mM^t}^\dagger$ norms.
 We call the algorithm for a particular $k$ \emph{Fanion-$k$}.
-Thus, Muon is Fanion-$\min\{m,n\}$, while Neon is Fanion-1.
-Moreover, the unsharded version of rank-$r$ Dion (Algorithm 1 from \citet{ahn2025dion}) without error feedback and without scaling of the update is actually Fanion-$r$ (see \citet{pethick2025understandingdion}, where Dion is written in a notation more similar to ours).
+This family elegantly bridges the rank gap with updates of any intermediate rank $k$.
+
+\paragraph{Connection to existing algorithms}
+The Fanion family unifies several known algorithms as special cases:
+\begin{itemize}
+	\item \textbf{Neon} is Fanion-1 (rank-one updates)
+	\item \textbf{Muon} is Fanion-$\min\{m,n\}$ (full-rank updates)
+	\item \textbf{Dion} (unsharded): The rank-$r$ Dion (Algorithm 1 from \citet{ahn2025dion}) without error feedback and without scaling of the update is actually Fanion-$r$ (see \citet{pethick2025understandingdion}, where Dion is written in a notation more similar to ours)
+\end{itemize}
+
+In \sectionref{sec:matrix-side}, we will discuss how to efficiently compute Fanion updates using the Lanczos algorithm.
 

### 2025-11-28 article_final/main.tex

@@
 Our experiments reveal important insights into the role of matrix norms in optimization.
-First, using the example of Neon (the rank-one Fanion), we show that not every LMO-based algorithm is effective, despite sharing the same asymptotics in the general bounds of \citet{kovalev2025understanding} and \citet{riabinin2025gluon}.
+First, using the example of Neon (the rank-one Fanion), we show that not every LMO-based algorithm is effective, despite sharing the same asymptotics in the general bounds of \citep{kovalev2025understanding} and \citep{riabinin2025gluon}.
 This suggests that existing theoretical guarantees should be refined to better explain empirical performance.
@@
 \paragraph{Schatten norms}
-\citet{cesista2025schattenp} considered Schatten-$p$ norms:
+\citep{cesista2025schattenp} considered Schatten-$p$ norms:
 \[\norm{\mM^t}_{S_p} = \left(\sum_{i=1}^{\min(m, n)}\sigma_i^p\right)^{1/p}\,,
@@
 
-Deriving the update for arbitrary $k$: Since $\normkfk{\cdot}^\dagger = \max\{\frac{1}{k} \normn{\cdot}, \norms{\cdot}\}$ (see \citet{bhatia2013matrix}, p.
+Deriving the update for arbitrary $k$: Since $\normkfk{\cdot}^\dagger = \max\{\frac{1}{k} \normn{\cdot}, \norms{\cdot}\}$ (see \citep{bhatia2013matrix}, p.
 96), the goal is to reach $\max\{\frac{1}{k}\sum_{i=1}^{\min(m,n)}\sigma_i, \sigma_1\}$.
@@
 	\item \textbf{Muon} is Fanion-$\min\{m,n\}$ (full-rank updates)
-	\item \textbf{Dion} (unsharded): The rank-$r$ Dion (Algorithm 1 from \citet{ahn2025dion}) without error feedback and without scaling of the update is actually Fanion-$r$ (see \citet{pethick2025understandingdion}, where Dion is written in a notation more similar to ours)
+	\item \textbf{Dion} (unsharded): The rank-$r$ Dion (Algorithm 1 from \citep{ahn2025dion}) without error feedback and without scaling of the update is actually Fanion-$r$ (see \citep{pethick2025understandingdion}, where Dion is written in a notation more similar to ours)
 \end{itemize}
@@
 \subsection{General Case}
-First, we need a well-known fact mentioned, for example, in \citet[Table 1]{yu2012arithmetic}. For the sake of completeness, we provide a proof here.
+First, we need a well-known fact mentioned, for example, in \citep[Table 1]{yu2012arithmetic}. For the sake of completeness, we provide a proof here.
 
@@
 The spectra of $\mM$ and $\mN$ are uniformly distributed in the $(0, 1)$ interval.
-We set $\mS=0$, and $\mX^0 \sim 0.1 \cN(0, \mI)$.
+We set $\mS=0$, and elements of $\mX^0$ are instances of random value with distribution $\sim 0.1 \cN(0, 1)$.
 
@@
 The parallels are not limited to mere similarity of the updates and can be traced even in the empirical observations.
-SignSGD is very close to Adam, as noted in \citet{bernstein2024oldoptimizernewnorm}, and both Adam and Muon perform well in training large models \citep{zhao2025deconstructing, liu2025muon}.
+SignSGD is very close to Adam, as noted in \citep{bernstein2024oldoptimizernewnorm}, and both Adam and Muon perform well in training large models \citep{zhao2025deconstructing, liu2025muon}.
 NSGD is the same for both matrix and vector cases.
@@
 \paragraph{Improvements of Muon} Regarding the first point, a large number of applications and improvements of Muon has been proposed in less than a year.
-\citet{liu2025muon} adapted the algorithm for training language models larger than NanoGPT.
-\citet{shah2025practical} organized efficient hyperparameter transfer by combining Muon with maximal update parametrization.
-To construct their COSMOS optimizer, \citet{chen2025cosmoshybridadaptive} applied computationally intensive updates of SOAP optimizer \citep{vyas2025soap} to a low-dimensional ``leading eigensubspace'' while using memory-efficient methods like Muon for the remaining parameters.
-\citet{amsel2025polar, grishina2025chebyshev} proposed more efficient alternatives to Newton-Schulz algorithm.
-\citet{si2025adamuon} introduced AdaMuon which combines element-wise adaptivity with orthogonal updates.
+\citep{liu2025muon} adapted the algorithm for training language models larger than NanoGPT.
+\citep{shah2025practical} organized efficient hyperparameter transfer by combining Muon with maximal update parametrization.
+To construct their COSMOS optimizer, \citep{chen2025cosmoshybridadaptive} applied computationally intensive updates of SOAP optimizer \citep{vyas2025soap} to a low-dimensional ``leading eigensubspace'' while using memory-efficient methods like Muon for the remaining parameters.
+\citep{amsel2025polar, grishina2025chebyshev} proposed more efficient alternatives to Newton-Schulz algorithm.
+\citep{si2025adamuon} introduced AdaMuon which combines element-wise adaptivity with orthogonal updates.
 We suppose that the described above or similar techniques can be applied to our optimizers as well.
@@
 
-\paragraph{Theory behind Muon} Regarding the second point, there has been a prolonged gap in theory behind Muon, simplistic derivation of \citet{bernstein2025deriving} based on \citet{bernstein2024oldoptimizernewnorm} excluded.
+\paragraph{Theory behind Muon} Regarding the second point, there has been a prolonged gap in theory behind Muon, simplistic derivation of \citep{bernstein2025deriving} based on \citep{bernstein2024oldoptimizernewnorm} excluded.
 This gap, as it seems to us, is not even now completely closed.
-For example, \citet{kovalev2025understanding} has provided convergence guarantees of Muon and in various settings, from which, however, Muon's supremacy cannot be recovered.
+For example, \citep{kovalev2025understanding} has provided convergence guarantees of Muon and in various settings, from which, however, Muon's supremacy cannot be recovered.
 Indeed, although the obtained bounds depend on the norm choice, the asymptotics of the convergence remain the same as for NSGD and other optimizers, $K = \cO(\epsilon^{-4})$ in an $L$-smooth stochastic case.
 
-A similar drawback has the article \citet{riabinin2025gluon}, where $L$-smoothness assumption is replaced with a more practical $(L_0, L_1)$-smoothness.
-By estimating smoothness and substituting it into their Theorem 1, the authors recovered the optimal fine-tuned stepsizes reported by \citet{pethick2025training}.
+A similar drawback has the article \citep{riabinin2025gluon}, where $L$-smoothness assumption is replaced with a more practical $(L_0, L_1)$-smoothness.
+By estimating smoothness and substituting it into their Theorem 1, the authors recovered the optimal fine-tuned stepsizes reported by \citep{pethick2025training}.
 However, they have not showed the optimality of the spectral or the RMS-to-RMS norm, which is observed in practice, as our comparison with NSGD highlights.
@@
 \paragraph{The LMO and Error Feedback} We already mentioned that rank-$k$ unsharded Dion without error feedback and scaling of the update is Fanion-$k$.
-Since error feedback for Dion is very important as discovered by the ablation study in \citet{ahn2025dion}, F-Fanions and S-Fanions will benefit from it as well.
+Since error feedback for Dion is very important as discovered by the ablation study in \citep{ahn2025dion}, F-Fanions and S-Fanions will benefit from it as well.
 In the context of federated learning, error feedback is effective even for compressed Muon \citep{gruntkowska2025efmuon}.
@@
 \paragraph{The nuclear norm in the LMO} As we found out only when transforming our results into an article, the nuclear norm has already been explored in the context of the linear minimization oracle.
-\citet{pethick2025sam} applied it to create $\nu$SAM, a new sharpness-aware minimization technique.
+\citep{pethick2025sam} applied it to create $\nu$SAM, a new sharpness-aware minimization technique.
 It would be interesting to substitute $\normn{\cdot}$ with $\normkfk{\cdot}^\dagger$ in their approach.
@@
 
-\paragraph{Orthogonal research} Fanions, F-Fanions, and S-Fanions benefit from the general theoretical description of \citet{riabinin2025gluon}, and better learning rates can be predicted by calculating the trajectory smoothness.
-They could be transformed to Drop-Fanions by updating only some layers as in \citet{gruntkowska2025dropmuon}.
+\paragraph{Orthogonal research} Fanions, F-Fanions, and S-Fanions benefit from the general theoretical description of \citep{riabinin2025gluon}, and better learning rates can be predicted by calculating the trajectory smoothness.
+They could be transformed to Drop-Fanions by updating only some layers as in \citep{gruntkowska2025dropmuon}.
 They could be viewed as approximations of the Non-Euclidean Broximal Point Method for the corresponding norms \citep{gruntkowska2025noneuclideanbroximal}.
@@
 They could be made more memory-effective by the zero-order techniques that worked for Muon \citep{petrov2025zeromuon}.
-Finally, the results from \citet{shulgin2025inexactmuon} can be used to explain the robustness of Muon to the norm change that occurred in our experiments and to theoretically derive faster yet working approximate schemes to calculate the LMO; power iterations with a very limited number of iterations is a good algorithm to consider for the analysis.
+Finally, the results from \citep{shulgin2025inexactmuon} can be used to explain the robustness of Muon to the norm change that occurred in our experiments and to theoretically derive faster yet working approximate schemes to calculate the LMO; power iterations with a very limited number of iterations is a good algorithm to consider for the analysis.
 
@@
 
-IO suggested using the nuclear norm in the \citet{bernstein2024oldoptimizernewnorm} framework.
+IO suggested using the nuclear norm in the \citep{bernstein2024oldoptimizernewnorm} framework.
 DM presented the problem at the MIPT optimization course.
@@
 
-In fact, those balls can be described easier if we use the results from \citet{yu2012arithmetic}.
+In fact, those balls can be described easier if we use the results from \citep{yu2012arithmetic}.
 The Ky Fan $2$-norm ball is an intersection of three $l_1$ balls in $(x,y)$, $(x, z)$, and $(y,z)$ spaces.
@@
 
-\paragraph{NanoGPT pretrain} We used the standard setting of 8 $\times$ H100 as documented in \citet{modded_nanogpt_2024}.
+\paragraph{NanoGPT pretrain} We used the standard setting of 8 $\times$ H100 as documented in \citep{modded_nanogpt_2024}.
 

### 2025-11-28 article_final/main.tex
reworked the proofs in section 3 a little

--- 2025-11-28 article_final/main.tex
@@
 Here, $\mM^t = \mU \mSigma \mV^\top$ is the Singular Value Decomposition (SVD) of $\mM^t$ ($\mU = [\vu_1, \vu_2, \dots, \vu_r]$, $\mSigma = \diag(\sigma_1, \sigma_2, \dots, \sigma_r)$, and $\mV = [\vv_1, \vv_2, \dots, \vv_r]$).
-Muon can be recovered by taking the RMS-to-RMS operator norm $\sqrt{\frac{n}{m}}\norms{\cdot}$.
+Muon can be recovered by taking the RMS-to-RMS norm: $\sqrt{n/m}\norms{\cdot}$.
 
@@
 
-Having examined the Frobenius norm (NSGD), spectral norm (Muon), and Chebyshev norm (SignSGD), all of which produce full-rank updates, we naturally turn to the nuclear norm $\normn{\cdot}$.
+Having examined the Frobenius norm (NSGD), spectral norm (Muon), and Chebyshev norm (SignSGD), all of which produce full-rank updates, we turn to the nuclear norm $\normn{\cdot}$.
 
@@
 \begin{proof}
-Since $\normn{\cdot}^\dagger = \norms{\cdot}$, the goal is to reach $\norms{\mM^t} = \sigma_1$.
-Note that $\Delta = \vu_1 \vv_1^\top$ delivers this value.
-Indeed, $\normn{\Delta}=1$ and by the trace property and orthogonality of the singular vectors,
+Since $\normn{\cdot}^\dagger = \norms{\cdot}$, the goal is to reach $\<\mM^t, \mD> = \sigma_1$ in \equationref{eq:our_update}.
+Note that $\mD = \vu_1 \vv_1^\top$ delivers this value.
+Indeed, $\normn{\mD}=1$ and by the trace property and orthogonality of the singular vectors,
 \[
-\<\mM^t, \Delta> =\<\mU \mSigma \mV^\top, \vu_1 \vv_1^\top> = \tr \diag({\sigma_1, 0, \dots, 0}) = \norms{\mM^t}\,,
+\<\mM^t, \mD> =\<\mU \mSigma \mV^\top, \vu_1 \vv_1^\top> = \tr \diag({\sigma_1, 0, \dots, 0}) = \norms{\mM^t}\,,
 \]
@@
 \]
-where $p$ and $q$ satisfy $\frac{1}{p} + \frac{1}{q} = 1$.
+where $p$ and $q$ satisfy $p^{-1} + q^{-1} = 1$.
 This formula recovers Neon when $p\rightarrow1$ (provided $\sigma_1 > \sigma_2$, which holds on real data), NSGD when $p=2$, and Muon when $p\rightarrow\infty$.
 
-However, Schatten norms do not fill the rank gap: when $p > 1$, the update has full rank, while when $p=1$ (or $p \rightarrow 1$), it has rank one.
+However, Schatten norms do not fill the rank gap: when $p > 1$, the update has full rank, while when $p=1$, it has rank one.
 Moreover, computing the update for $p \neq 1, 2, \infty$ appears to require knowing all $\sigma_i$, making the problem as hard as computing the full SVD.
@@
 
-Deriving the update for arbitrary $k$: Since $\normkfk{\cdot}^\dagger = \max\{\frac{1}{k} \normn{\cdot}, \norms{\cdot}\}$ (see \citet{bhatia2013matrix}, p.
-96), the goal is to reach $\max\{\frac{1}{k}\sum_{i=1}^{\min(m,n)}\sigma_i, \sigma_1\}$.
-The Neon update $\Delta = \vu_1 \vv_1^\top$ achieves $\sigma_1$, while the scaled Muon update $\Delta = \frac{1}{k}\sum_{i=1}^{\min(m,n)}\vu_i \vv_i^\top$ achieves $\frac{1}{k}\sum_{i=1}^{\min(m,n)}\sigma_i$.
+To derive the update formula for arbitrary $k$, we use the expression for the norm dual to the Ky Fan $k$-norm (see, for instance \citet{bhatia2013matrix}, p. 96):
+\[
+    \normkfk{\cdot}^\dagger = \max\left\{\frac{1}{k} \normn{\cdot}, \norms{\cdot}\right\}\,.
+\]
+According to \equationref{eq:our_update}, we need to find $\mD$ such that $\normkfk{\mD} = 1$ and
+\[
+\<\mM^t, \mD> = \max\left\{\frac{1}{k}\sum_{i=1}^{\min(m,n)}\sigma_i,\ \sigma_1\right\}\,.
+\]
+The Neon update $\mD = \vu_1 \vv_1^\top$ achieves $\sigma_1$, while the scaled Muon update $\mD = \frac{1}{k}\sum_{i=1}^{\min(m,n)}\vu_i \vv_i^\top$ achieves $\frac{1}{k}\sum_{i=1}^{\min(m,n)}\sigma_i$.
 Thus, the update is either
@@
 \]
-depending on which better minimizes the linear function $L(\mX) \equiv F(\mX^t) + \mM^t(\mX - \mX^t)$.
-
+depending on which of those two expressions achieves smaller residual on the linear function $L(\mX) := F(\mX^t) + \<\mM^t, \mX - \mX^t>$.
 The Ky Fan norms thus fail to close the rank gap: the resulting update is either rank-one or full-rank.
@@
 
-While Schatten norms are closed under dualization and $\norm{\mM^t}_{S_p}^\dagger = \norm{\mM^t}_{S_q}$, for Ky Fan norms it is not generally the case.
-$k = 1$ and $k=\min(m, n)$ are exceptional: for $k=1$ the dual norm is $\normn{\cdot}$ with $k = \min(m, n)$ and for $k = \min(m, n)$ the dual norm is $\norms{\cdot}$ with $k=1$.
-For each other $k$, $\normkfk{\cdot} \neq \norm{\cdot}_{\text{KF-}k'}$ for any $k'$: $k'=\min(m, n)$ and $k'= 1$ correspond to the already discussed cases, while for other $k'$ one can change $\norm{\mM^t}_{\text{KF-}k'}$ by moving a small value between $\sigma_{k'}$ and $\sigma_{k'+1}$.
-During this change, $\norms{\mM^t}$ and $\normn{\mM^t}$ will remain constant, hence $\normkfk{\mM^t}$ will not change as well.
+Unlike Schatten norms, which satisfy $\norm{\mM^t}_{S_p}^\dagger = \norm{\mM^t}_{S_q}$ (for $p^{-1} + q^{-1} = 1$) and are thus closed under dualization, Ky Fan norms are generally not.
+Only two exceptional cases exist: the dual of the Ky Fan $1$-norm $\norm{\cdot}_{\text{KF-}1}$ (the spectral norm) is the Ky Fan $\min(m,n)$-norm $\norm{\cdot}_{\text{KF-}\min(m,n)}$ (the nuclear norm), and vice versa.
 
-The key insight is to work with the duals of Ky Fan norms:
+Thus working with duals of Ky Fan norms can open new possibilities:
 
@@
 	When $\norm{\cdot} = \normkfk{\mM^t}^\dagger$, \equationref{eq:our_update} turns into:
-
 	\begin{equation}\label{eq:ky_fan_update}
-		\mX^{t+1} = \mX^{t} - \gamma_t \sum_{i=1}^{k}\vu_i \vv_i^\top
+		\mX^{t+1} = \mX^{t} - \gamma_t \sum_{i=1}^{k}\vu_i \vv_i^\top\,.
 	\end{equation}
@@
 \begin{proof}
-Since $\norm{\cdot}^\dagger = \normkfk{\cdot}^{\dagger\dagger} = \normkfk{\cdot}$, the goal is to reach $\normkfk{\mM^t}$.
-
-Let us note that $\Delta = \sum_{i=1}^{k} \vu_i \vv_i^\top$ delivers the value. Indeed,
+Since $\norm{\cdot}^\dagger = \normkfk{\cdot}^{\dagger\dagger} = \normkfk{\cdot}$, the goal is to reach $\<\mM^t, \mD> = \normkfk{\mM^t}$ in \equationref{eq:our_update}.
+Note that $\mD = \sum_{i=1}^{k} \vu_i \vv_i^\top$ attains this value.
+Indeed,
 \[
-\<\mM^t, \Delta> =\<\mU \mSigma \mV^\top, \sum_{i=1}^{k} \vu_i \vv_i^\top> = \sum_{i,j=1}^{r, k}\<\vu_i \sigma_i \vv_i^\top, \vu_j \vv_j^\top> = \sum_{i=1}^{k}\sigma_i = \normkfk{\mM^t}\,,
+\<\mM^t, \mD> =\<\mU \mSigma \mV^\top, \sum_{i=1}^{k} \vu_i \vv_i^\top> = \sum_{i,j=1}^{r, k}\<\vu_i \sigma_i \vv_i^\top, \vu_j \vv_j^\top> = \sum_{i=1}^{k}\sigma_i = \normkfk{\mM^t}\,,
 \]
@@
 
-These updates constitute the family of \emph{Fanions}, LMO-based algorithms under the $\normkfk{\mM^t}^\dagger$ norms.
-We call the algorithm for a particular $k$ \emph{Fanion-$k$}.
-This family elegantly bridges the rank gap with updates of any intermediate rank $k$.
+These updates define the \emph{Fanion} family of LMO-based algorithms, each operating under a norm $\normkfk{\mM^t}^\dagger$.
+We denote the algorithm for a particular $k$ as \emph{Fanion-$k$}.
+This family elegantly bridges the rank gap, providing updates of any intermediate rank $k$.
 

### 2025-11-28 article_final/main.tex
fixed a ffiliations

@@
        \Email{kravtskii.aiu@phystech.edu}\\
-       \addr Moscow Institute of Physics and Technology (MIPT)
+       \addr MIPT
        \AND
@@
        \Email{kozyrev.in@phystech.edu}\\
-       \addr Moscow Institute of Physics and Technology (MIPT)\\
-       Marchuk Institute of Numerical Mathematics
+       \addr MIPT, INM RAS
        \AND
@@
        \Email{kozlov.na@phystech.edu}\\
-       \addr Moscow Institute of Physics and Technology (MIPT)
+       \addr MIPT
        \AND
@@
        \Email{vinogradov.am@phystech.edu}\\
-       \addr Moscow Institute of Physics and Technology (MIPT)
+       \addr MIPT
        \AND
@@
        \Email{daniil.merkulov@phystech.edu}\\
-       \addr Moscow Institute of Physics and Technology (MIPT)\\
-       Skoltech, HSE, AI4Science
+       \addr MIPT, Skoltech, HSE, AI4Science
        \AND
@@
        \Email{i.oseledets@skoltech.ru}\\
-       \addr AIRI, Skoltech\\
-       Marchuk Institute of Numerical Mathematics}
+       \addr AIRI, Skoltech, INM RAS
+}
 
@@
 Only two exceptional cases exist: the dual of the Ky Fan $1$-norm $\norm{\cdot}_{\text{KF-}1}$ (the spectral norm) is the Ky Fan $\min(m,n)$-norm $\norm{\cdot}_{\text{KF-}\min(m,n)}$ (the nuclear norm), and vice versa.
-
-Thus working with duals of Ky Fan norms can open new possibilities:
+Thus working with duals of Ky Fan norms can open up new possibilities:
 
@@
 \paragraph{Connection to existing algorithms}
-The Fanion family unifies several known algorithms as special cases:
+The Fanion family unifies several known algorithms:
 \begin{itemize}

### 2025-11-28 article_final/main.tex
refined text in section 4

--- 2025-11-28 article_final/main.tex
@@
 
-Deriving the update for arbitrary $k$: Since $\normkfk{\cdot}^\dagger = \max\{\frac{1}{k} \normn{\cdot}, \norms{\cdot}\}$ (see \citep{bhatia2013matrix}, p.
-96), the goal is to reach $\max\{\frac{1}{k}\sum_{i=1}^{\min(m,n)}\sigma_i, \sigma_1\}$.
-The Neon update $\Delta = \vu_1 \vv_1^\top$ achieves $\sigma_1$, while the scaled Muon update $\Delta = \frac{1}{k}\sum_{i=1}^{\min(m,n)}\vu_i \vv_i^\top$ achieves $\frac{1}{k}\sum_{i=1}^{\min(m,n)}\sigma_i$.
+To derive the update formula for arbitrary $k$, we use the expression for the norm dual to the Ky Fan $k$-norm (see, for instance \citep{bhatia2013matrix}, p. 96):
+\[
+    \normkfk{\cdot}^\dagger = \max\left\{\frac{1}{k} \normn{\cdot}, \norms{\cdot}\right\}\,.
+\]
+According to \equationref{eq:our_update}, we need to find $\mD$ such that $\normkfk{\mD} = 1$ and
+\[
+\<\mM^t, \mD> = \max\left\{\frac{1}{k}\sum_{i=1}^{\min(m,n)}\sigma_i,\ \sigma_1\right\}\,.
+\]
+The Neon update $\mD = \vu_1 \vv_1^\top$ achieves $\sigma_1$, while the scaled Muon update $\mD = \frac{1}{k}\sum_{i=1}^{\min(m,n)}\vu_i \vv_i^\top$ achieves $\frac{1}{k}\sum_{i=1}^{\min(m,n)}\sigma_i$.
 Thus, the update is either
@@
 
-While Schatten norms are closed under dualization and $\norm{\mM^t}_{S_p}^\dagger = \norm{\mM^t}_{S_q}$, for Ky Fan norms it is not generally the case.
-$k = 1$ and $k=\min(m, n)$ are exceptional: for $k=1$ the dual norm is $\normn{\cdot}$ with $k = \min(m, n)$ and for $k = \min(m, n)$ the dual norm is $\norms{\cdot}$ with $k=1$.
-For each other $k$, $\normkfk{\cdot} \neq \norm{\cdot}_{\text{KF-}k'}$ for any $k'$: $k'=\min(m, n)$ and $k'= 1$ correspond to the already discussed cases, while for other $k'$ one can change $\norm{\mM^t}_{\text{KF-}k'}$ by moving a small value between $\sigma_{k'}$ and $\sigma_{k'+1}$.
-During this change, $\norms{\mM^t}$ and $\normn{\mM^t}$ will remain constant, hence $\normkfk{\mM^t}$ will not change as well.
-
-The key insight is to work with the duals of Ky Fan norms:
+Unlike Schatten norms, which satisfy $\norm{\mM^t}_{S_p}^\dagger = \norm{\mM^t}_{S_q}$ (for $p^{-1} + q^{-1} = 1$) and are thus closed under dualization, Ky Fan norms are generally not.
+Only two exceptional cases exist: the dual of the Ky Fan $1$-norm $\norm{\cdot}_{\text{KF-}1}$ (the spectral norm) is the Ky Fan $\min(m,n)$-norm $\norm{\cdot}_{\text{KF-}\min(m,n)}$ (the nuclear norm), and vice versa.
+Thus working with duals of Ky Fan norms can open up new possibilities:
 
@@
 
-These updates constitute the family of \emph{Fanions}, LMO-based algorithms under the $\normkfk{\mM^t}^\dagger$ norms.
-We call the algorithm for a particular $k$ \emph{Fanion-$k$}.
-This family elegantly bridges the rank gap with updates of any intermediate rank $k$.
+These updates define the \emph{Fanion} family of LMO-based algorithms, each operating under a norm $\normkfk{\mM^t}^\dagger$.
+We denote the algorithm for a particular $k$ as \emph{Fanion-$k$}.
+This family elegantly bridges the rank gap, providing updates of any intermediate rank $k$.
 
 \paragraph{Connection to existing algorithms}
-The Fanion family unifies several known algorithms as special cases:
+The Fanion family unifies several known algorithms:
 \begin{itemize}
@@
 
-Simply applying the norm dual to the Ky Fan $k$-norm, however, is not enough to provide a family of algorithms diverse enough for our purposes.
-Next, we consider linear combinations of algorithms from the LMO framework, which happen to be LMO-algorithms as well, as we show in the following paragraphs.
+Simply applying norms dual to Ky Fan $k$-norms, however, does not provide sufficient algorithmic diversity for our purposes.
+We now consider linear combinations of LMO-based algorithms and show that these combinations are themselves LMO-algorithms.
 
 \subsection{General Case}
-First, we need a well-known fact mentioned, for example, in \citep[Table 1]{yu2012arithmetic}. For the sake of completeness, we provide a proof here.
+We begin with a well-known result on dual norms (see, e.g., \citep[Table 1]{yu2012arithmetic}).
 
@@
 
-\begin{proof}
-	Let us first prove the lemma for the case $n = 2$. Denote $f(x) = \alpha_1 \norm{x}_{(1)}$ and $g(x) = \alpha_2 \norm{x}_{(2)}$, such that $\norm{x} = f(x) + g(x)$. Recall two standard facts:
-	\begin{enumerate}
-		\item For any norm $\norm{\cdot}$ and $\lambda>0$,
-		      \[
-			      (\lambda \norm{\cdot})^*(y) =
-			      \sup_{x}\bigl( \langle y,x \rangle - \lambda \norm{x}\bigr)
-			      = \delta_{\lambda \cB_{\norm{\cdot}^\dagger}}(y)\,,
-		      \]
-		      i.e., the indicator function of the scaled dual ball.
-		\item The Fenchel conjugate of a sum satisfies
-		      \[
-			      (f+g)^*(y) = \inf_{u+v=y} \bigl(f^*(u) + g^*(v)\bigr)\,.
-		      \]
-	\end{enumerate}
-	Applying these to $f$ and $g$, we have
-	\[
-		f^*(u) = \delta_{\alpha_1 \cB_{\norm{\cdot}_{(1)}^\dagger}}(u)\,,
-		\quad
-		g^*(v) = \delta_{\alpha_2 \cB_{\norm{\cdot}_{(2)}^\dagger}}(v)\,.
-	\]
-	Thus,
-	\[
-		\norm{\cdot}^*(y)
-		= (f+g)^*(y)
-		= \inf_{u+v=y}
-		\bigl(
-		\delta_{\alpha_1 \cB_{\norm{\cdot}_{(1)}^\dagger}}(u)
-		+
-		\delta_{\alpha_2 \cB_{\norm{\cdot}_{(2)}^\dagger}}(v)
-		\bigr)
-		=
-		\delta_{\alpha_1 \cB_{\norm{\cdot}_{(1)}^\dagger}+\alpha_2 \cB_{\norm{\cdot}_{(2)}^\dagger}}(y).
-	\]
-	By definition, the conjugate of a norm is exactly the indicator of its dual unit ball:
-	\[
-	\norm{\cdot}^*(y) = \delta_{\cB_{\norm{\cdot}^\dagger}}(y)\,.
-	\]
-	Therefore, $\cB_{\norm{\cdot}^\dagger} = \alpha_1 \cB_{\norm{\cdot}_{(1)}^\dagger} + \alpha_2 \cB_{\norm{\cdot}_{(2)}^\dagger}$.
-
-	Now we prove the general case by induction. The base case ($n=2$) is already proven. Suppose that the assumption of the lemma holds for $n = k$. Then, for $n = k + 1$,
-	\[
-	\norm{x} = \sum_{i = 1}^k \alpha_i \norm{x}_{(i)} + \alpha_{k + 1}\norm{x}_{(k + 1)} = \norm{x}_{(1:k)} + \alpha_{k + 1}\norm{x}_{(k + 1)}\,.
-	\]
-	Applying the result for $n=2$ combined with the induction assumption, we obtain
-	\[
-	\cB_{\norm{\cdot}^\dagger} = \cB_{\norm{\cdot}_{(1:k)}^\dagger} + \alpha_{k + 1} \cB_{\norm{\cdot}_{(k + 1)}^\dagger} = \sum_{i = 1}^{k + 1} \alpha_i \cB_{\norm{\cdot}_{(i)}^\dagger}\,,
-	\]
-	which proves the lemma.
-\end{proof}
+A proof is provided in the appendix (see \sectionref{sec:proof_dual_conv_comb}).
 
@@
 \begin{proof}
-	Using \lemmaref{lemma:dual_to_conv_comb} and the fact $\norm{\cdot}^{\dagger \dagger} = \norm{\cdot}$, we can obtain general form of the unit ball in the $\norm{\cdot}$ norm: $\cB_{\norm{\cdot}} = \sum_{i = 1}^n \alpha_i \cB_{\norm{\cdot}_{(i)}}$.
-Thus, the linear function minimization objective on a $\norm{\cdot}$ norm ball can be transformed as follows:
+	Using \lemmaref{lemma:dual_to_conv_comb} and the biduality property $\norm{\cdot}^{\dagger \dagger} = \norm{\cdot}$, we obtain the unit ball representation: $\cB_{\norm{\cdot}} = \sum_{i = 1}^n \alpha_i \cB_{\norm{\cdot}_{(i)}}$.
+The linear minimization problem over this ball can thus be transformed as follows:
 	\[
@@
 	\]
-	where the last summation denotes the Minkowski sum.
-This immediately implies $\sum_{i = 1}^n \alpha_i \mathrm{LMO}_{i} \in \argmin_{\mD \in \cB_{\norm{\cdot}}}\<\mM, \mD>$, which proves the claim of the lemma.
+	where the last summation denotes the Minkowski sum. This immediately implies $\sum_{i = 1}^n \alpha_i \mathrm{LMO}_{i} \in \argmin_{\mD \in \cB_{\norm{\cdot}}}\<\mM, \mD>$, completing the proof.
 \end{proof}
 
-Applying it to optimization algorithms, we obtain the following corollary.
+Applying this result to optimization algorithms yields the following corollary.
 
@@
 \subsection{Frobeniusize Them: F-Muon and F-Neon}
-Let us construct the concrete examples of algorithms given by linear combinations of LMO-algorithms.
-It follows from \corollaryref{corollary:lin_comb_alg} that those algorithms can also be viewed as LMO-algorithms.
+We now construct concrete examples of algorithms obtained via linear combinations of LMO-algorithms. By \corollaryref{corollary:lin_comb_alg}, these combinations are themselves LMO-algorithms.
 
-Combining Fanions for various $k$ with another excellent LMO-algorithm, NSGD, we obtain the family of algorithms with updates defined by
+Combining Fanion-$k$ with NSGD yields a family of algorithms with updates
 \begin{equation}\label{eq:nsgd_muon_update}
@@
 \end{equation}
-Recall that Fanion-$k$ update is governed by the norm dual to the Ky Fan-$k$ norm, and NSGD corresponds to the Frobenius norm, which is dual to itself.
-Thus, by \corollaryref{corollary:lin_comb_alg}, our new algorithm is an LMO-algorithm corresponding to the unit ball of the norm $\norm{\cdot}_{\mathrm{F-KF-k}}^\dagger$:
+Recall that Fanion-$k$ operates under the dual to the Ky Fan $k$-norm, while NSGD operates under the self-dual Frobenius norm.
+By \corollaryref{corollary:lin_comb_alg}, this combination defines an LMO-algorithm with norm $\norm{\cdot}_{\mathrm{F-KF-k}}^\dagger$, where
 \begin{equation}\label{eq:fkfk_norm}
@@
 \end{equation}
-We name the derived family of algorithms \emph{F-Fanions}.
+We call this family \emph{F-Fanions}.
 
-Two edge members of this family, with $k = 1$ and $k = \min\{m, n\}$ correspondingly, \emph{F-Neon} and \emph{F-Muon}, are of particular interest to us.
-A bit of information and visualizations related to the $\normfstar{\cdot} = \norm{\cdot}_{\mathrm{F-KF-1}}$ norm is presented in the appendix (see \equationref{eq:normfstardual}, \figureref{fig:fstardual}).
+The extremal members of this family are \emph{F-Neon} (with $k = 1$) and \emph{F-Muon} (with $k = \min\{m, n\}$).
+Additional information and visualizations for the $\normfstar{\cdot} = \norm{\cdot}_{\mathrm{F-KF-1}}$ norm appear in the appendix (see \equationref{eq:normfstardual}, \figureref{fig:fstardual}).
 
@@
 
-Another algorithm we consider is the SignSGD. Mixing its update with that of Fanion-$k$, we obtain the update of \emph{S-Fanion-$k$}:
+Similarly, combining Fanion-$k$ with SignSGD yields \emph{S-Fanion-$k$} with update
 \begin{equation}\label{eq:sign_muon_update}
@@
 \end{equation}
-where $\eta$ is the special SignSGD learning rate coefficient.
+where $\eta$ is a scaling coefficient specific to SignSGD.
 
-The norm $\norm{\cdot}_{\mathrm{C-KF-k}}^\dagger$ that produces this update in the LMO framework is again given by \corollaryref{corollary:lin_comb_alg}:
+By \corollaryref{corollary:lin_comb_alg}, this defines an LMO-algorithm with norm $\norm{\cdot}_{\mathrm{C-KF-k}}^\dagger$, where
 \begin{equation}\label{eq:ckfk_norm}
@@
 \end{equation}
+
+The extremal members of this family are \emph{S-Neon} (with $k = 1$) and \emph{S-Muon} (with $k = \min\{m, n\}$).
 
@@
 The spectra of $\mM$ and $\mN$ are uniformly distributed in the $(0, 1)$ interval.
-We set $\mS=0$, and elements of $\mX^0$ are instances of random value with distribution $\sim 0.1 \cN(0, 1)$.
+We set $\mS=0$, and $\mX^0 \sim 0.1 \cN(0, \mI)$.
 
@@
 
+\section{Proof of Lemma on Dual of Convex Combinations}\label{sec:proof_dual_conv_comb}
+
+We provide here the proof of \lemmaref{lemma:dual_to_conv_comb}.
+
+\begin{proof}
+	Let us first prove the lemma for the case $n = 2$. Denote $f(x) = \alpha_1 \norm{x}_{(1)}$ and $g(x) = \alpha_2 \norm{x}_{(2)}$, such that $\norm{x} = f(x) + g(x)$. Recall two standard facts:
+	\begin{enumerate}
+		\item For any norm $\norm{\cdot}$ and $\lambda>0$,
+		      \[
+			      (\lambda \norm{\cdot})^*(y) =
+			      \sup_{x}\bigl( \langle y,x \rangle - \lambda \norm{x}\bigr)
+			      = \delta_{\lambda \cB_{\norm{\cdot}^\dagger}}(y)\,,
+		      \]
+		      i.e., the indicator function of the scaled dual ball.
+		\item The Fenchel conjugate of a sum satisfies
+		      \[
+			      (f+g)^*(y) = \inf_{u+v=y} \bigl(f^*(u) + g^*(v)\bigr)\,.
+		      \]
+	\end{enumerate}
+	Applying these to $f$ and $g$, we have
+	\[
+		f^*(u) = \delta_{\alpha_1 \cB_{\norm{\cdot}_{(1)}^\dagger}}(u)\,,
+		\quad
+		g^*(v) = \delta_{\alpha_2 \cB_{\norm{\cdot}_{(2)}^\dagger}}(v)\,.
+	\]
+	Thus,
+	\[
+		\norm{\cdot}^*(y)
+		= (f+g)^*(y)
+		= \inf_{u+v=y}
+		\bigl(
+		\delta_{\alpha_1 \cB_{\norm{\cdot}_{(1)}^\dagger}}(u)
+		+
+		\delta_{\alpha_2 \cB_{\norm{\cdot}_{(2)}^\dagger}}(v)
+		\bigr)
+		=
+		\delta_{\alpha_1 \cB_{\norm{\cdot}_{(1)}^\dagger}+\alpha_2 \cB_{\norm{\cdot}_{(2)}^\dagger}}(y).
+	\]
+	By definition, the conjugate of a norm is exactly the indicator of its dual unit ball:
+	\[
+	\norm{\cdot}^*(y) = \delta_{\cB_{\norm{\cdot}^\dagger}}(y)\,.
+	\]
+	Therefore, $\cB_{\norm{\cdot}^\dagger} = \alpha_1 \cB_{\norm{\cdot}_{(1)}^\dagger} + \alpha_2 \cB_{\norm{\cdot}_{(2)}^\dagger}$.
+
+	Now we prove the general case by induction. The base case ($n=2$) is already proven. Suppose that the assumption of the lemma holds for $n = k$. Then, for $n = k + 1$,
+	\[
+	\norm{x} = \sum_{i = 1}^k \alpha_i \norm{x}_{(i)} + \alpha_{k + 1}\norm{x}_{(k + 1)} = \norm{x}_{(1:k)} + \alpha_{k + 1}\norm{x}_{(k + 1)}\,.
+	\]
+	Applying the result for $n=2$ combined with the induction assumption, we obtain
+	\[
+	\cB_{\norm{\cdot}^\dagger} = \cB_{\norm{\cdot}_{(1:k)}^\dagger} + \alpha_{k + 1} \cB_{\norm{\cdot}_{(k + 1)}^\dagger} = \sum_{i = 1}^{k + 1} \alpha_i \cB_{\norm{\cdot}_{(i)}^\dagger}\,,
+	\]
+	which proves the lemma.
+\end{proof}
+
 \section{Visualization of different matrix norms}

### 2025-11-30 article_final/main.tex
Started expanding the article

@@
 
-\subsection{NanoGPT speedrun}
-
-We test F-Muon on the NanoGPT speedrun \citep{modded_nanogpt_2024}.
-For $\alpha = 0.5$, the optimal parameters are \texttt{lr=0.07, momentum=0.95}, while for Muon they were \texttt{lr=0.05, momentum=0.95}.
-After testing for 1750 steps we get 3.281 cross-entropy loss.
+\subsection{NanoGPT Speedrun}
+
+We test F-Muon and S-Muon with $\alpha=0.5$ on the NanoGPT speedrun \citep{modded_nanogpt_2024}.
+The optimal parameters are in the legend \figureref{fig:gpt_loss}, \texttt{sign\_sgd\_coeff=0.0003/0.07}.
+After testing for 1750 steps we get 3.281 cross-entropy loss for F-Muon and 3.287 for S-Muon.
 For Muon, it falls below the target threshold 3.28 and reaches 3.279.
@@
 It is even more striking, considering the fact that this F-Muon is an average of Muon and NSGD, and the latter performed quite poorly.
+The same is true for S-Muon.
 
@@
 
-Again, as on the CIFAR airbench, if we set $\alpha=-0.1$ that corresponds to no ball, we get a 3.2818 loss, which is a small difference from the baseline.
+Again, as on the CIFAR airbench, if for F-Muon we set $\alpha=-0.1$ that corresponds to no ball, we get a 3.2818 loss, which is a small difference from the baseline.
+
+It is noteworthy that F-Muon with NSGD-part that is not scaled by Muon's rule gives 3.288 loss (0.007 increase), while S-Muon scaled with Muon's rule gives 3.292 loss (0.005 increase)
+
+\subsection{GPT-2 Medium Speedrun}
+We scale from NanoGPT to GPT-2 Medium.
+After testing for 5960 steps with F-Muon we get 2.9215 loss, while for Muon it is 2.9198.
+In ?
+steps, F-Muon reaches <2.92 loss, thus needing only ?\% more iterations than Muon.
 

### 2025-11-30 article_final/main.tex
made language in section 5 a bit better

@@
 \section{Computing the Updates}\label{sec:matrix-side}
-We employ the thick-restarted Lanczos method for the symmetric eigenvalue problem (thick-restart Lanczos, TRLan) to compute the low-rank updates of Fanions.
-We apply TRLan to either $\mM^{t\top}\mM^t$ or $\mM^t \mM^{t\top}$ (whichever matrix is smaller).
-We use the CuPy implementation of \texttt{cupy.sparse.linalg.svds} \citep{cupy_svds_ref} which internally relies on TRLan \citep{simonz1998thick}.
+We employ the thick-restart Lanczos method for the symmetric eigenvalue problem (TRLan) to compute the low-rank updates of Fanions.
+We apply TRLan to either $\mM^{t\top}\mM^t$ or $\mM^t \mM^{t\top}$, selecting whichever matrix is smaller.
+We use the CuPy implementation of \texttt{cupy.sparse.linalg.svds} \citep{cupy_svds_ref}, which internally relies on TRLan \citep{simonz1998thick}.
 
 TRLan is specifically designed for efficiently approximating the largest singular values and corresponding singular vectors of very large matrices.
-The thick-restart strategy retains the most informative Ritz vectors across restarts, which dramatically accelerates convergence while keeping memory consumption moderate.
-TRLan is particularly attractive in our GPU setting because its dominant cost is a modest number of highly parallelizable matrix-vector multiplications (matvecs) and it avoids full reorthogonalization against the entire Krylov basis by using short recurrence relations combined with thick restarting.
+The thick-restart strategy retains the most informative Ritz vectors across restarts, which dramatically accelerates convergence while maintaining moderate memory consumption.
+TRLan is particularly well-suited to our GPU setting because its dominant computational cost consists of a modest number of highly parallelizable matrix-vector multiplications (matvecs), and it avoids full reorthogonalization against the entire Krylov basis by employing short recurrence relations combined with thick restarting.
 
@@
 
-In \tableref{tbl:matrix_methods}, we compare TRLan against randomized SVD (RSVD) and simple power iterations when computing the rank-$k$ update used in Fanion-$k$ and related algorithms.
+In \tableref{tbl:matrix_methods}, we compare TRLan against randomized SVD (RSVD) and simple power iterations for computing the rank-$k$ update used in Fanion-$k$ and related algorithms.
 Experiments are performed on dense random matrices with i.i.d.
-$\mathcal{N}(0,1)$ entries and use CPU implementations for fair comparison.
+$\mathcal{N}(0,1)$ entries using CPU implementations for fair comparison.
 We report:
@@
 
-On $500\times500$ matrices, TRLan and RSVD require comparable wall-clock time, but TRLan delivers orders-of-magnitude lower error and far fewer matvecs.
-On larger $5000\times5000$ matrices the advantage becomes even more pronounced: TRLan is 3-4 times faster than RSVD while using $\sim$30 times fewer matvecs at comparable or better accuracy.
-
-An interesting empirical observation is that RSVD tends to approximate the \emph{singular values} themselves reasonably well but the reconstructed low-rank matrix noticeably deviates from the truncated SVD, whereas TRLan provides an excellent approximation to the truncated SVD matrix itself (low $err_2$) at the cost of occasionally less accurate individual singular values.
-This makes TRLan the preferred choice for algorithms like Neon/Fanion-$k$ that only need the low-rank term $\sum \sigma_i \vu_i \vv_i^\top$, but less ideal for methods (e.g., Dion) that explicitly require accurate $\sigma_i$ for error feedback or step-size control.
-
-A current practical limitation is the lack of a native PyTorch implementation of thick-restart Lanczos; existing PyTorch-based randomized SVD routines cannot match TRLan's accuracy/efficiency combination for the matrix reconstruction task.
-
-For reference, \tableref{tbl:newton_schulz_reference} shows results for the Newton-Schulz polar decomposition iteration on the same matrices, $err_1$ is the relative error of $\mU\mV^T$ (29-30 iterations to converge, significantly higher matvec count than TRLan).
+On $500\times500$ matrices, TRLan and RSVD require comparable wall-clock time, but TRLan delivers orders-of-magnitude lower error with far fewer matvecs.
+On larger $5000\times5000$ matrices, this advantage becomes even more pronounced: TRLan is 3-4 times faster than RSVD while using $\sim$30 times fewer matvecs at comparable or superior accuracy.
+
+An interesting empirical observation is that RSVD tends to approximate the \emph{singular values} themselves reasonably well, but the reconstructed low-rank matrix exhibits noticeable deviation from the truncated SVD.
+In contrast, TRLan provides an excellent approximation to the truncated SVD matrix itself (low $err_2$), albeit at the cost of occasionally less accurate individual singular values.
+This makes TRLan the preferred choice for algorithms like Neon/Fanion-$k$ that only require the low-rank term $\sum \sigma_i \vu_i \vv_i^\top$, but less suitable for methods (e.g., Dion) that explicitly require accurate $\sigma_i$ for error feedback or step-size control.
+
+A current practical limitation is the absence of a native PyTorch implementation of thick-restart Lanczos; existing PyTorch-based randomized SVD routines cannot match TRLan's accuracy-efficiency combination for the matrix reconstruction task.
+
+For reference, \tableref{tbl:newton_schulz_reference} presents results for the Newton-Schulz polar decomposition iteration on the same matrices, where $err_1$ is the relative error of $\mU\mV^T$ (29-30 iterations to converge, resulting in a significantly higher matvec count than TRLan).
 

### 2025-11-30 article_final/main.tex
Added Medium logs + minor changes in the article

@@
 	\item CIFAR-10 airbench \citep{cifar2023airbench}
-	\item Pre-training NanoGPT on FineWeb dataset \citep{modded_nanogpt_2024}
+	\item Pre-training NanoGPT and GPT-2 Medium on FineWeb dataset \citep{modded_nanogpt_2024}
 	\item Fine-tuning NanoGPT on Tiny Stories \citep{eldan2023tinystoriessmalllanguagemodels}
@@
 \begin{equation}\label{eq:dual_norm}
-	\norm{\mG}^\dagger = \sup_{\mD \in \Rmn\,:\norm{\mD}\leq 1} \<\mG,\mD>\,.
+	\norm{\mM}^\dagger = \sup_{\mD \in \Rmn\,:\norm{\mD}\leq 1} \<\mM,\mD>\,.
 \end{equation}
@@
 \begin{equation}\label{eq:lmo}
-	\mathrm{LMO}(\mM^t) \in \argmin_{\mD \in \cS} \<\mM^t, \mD>\,,
-\end{equation}
-where $\mM^t$ is the gradient (or momentum) of $F$ at $\mX^t$ and $\cS \subset \Rmn$ is some set. The update of the algorithm is defined as follows:
+	\mathrm{LMO}(\mM^t) \in \argmin_{\mD \in \cS} \langle\mM^t, \mD\rangle\,,
+\end{equation}
+where $\mM^t$ is the effective update direction and $\cS \subset \Rmn$ is a constraint set. The algorithm proceeds as:
 \begin{equation}\label{eq:simple_update}
-	\mX^{t + 1} = \mX^t + \gamma_t\mathrm{LMO}\left(\mM^t\right)\,.
-\end{equation}
+  \begin{aligned}
+  \mB^t &= \beta \mB^{t-1} + (1 - \beta) \nabla F(\mX^t) \quad \text{(momentum buffer)},\\
+  \mM^t &= \begin{cases}
+  \nabla F(\mX^t) & \text{(no momentum)},\\
+  \mB^t & \text{(heavy ball)},\\
+  \nabla F(\mX^t) + \beta \mB^t & \text{(approximate Nesterov)},
+  \end{cases}\\
+  \mX^{t + 1} &= \mX^t + \gamma_t\mathrm{LMO}\left(\mM^t\right)\,,
+  \end{aligned}
+\end{equation}
+where $\beta \in [0, 1)$ is the momentum coefficient.
+Throughout our experiments, we employ approximate Nesterov momentum, which matches the implementation in Muon and PyTorch SGD (\texttt{nesterov=True}).
+It uses the current gradient rather than the true lookahead gradient, trading theoretical guarantees for computational efficiency.
 
@@
 \]
-and the update for $\mX^t$ in \equationref{eq:simple_update} simplifies to
+and the update for $\mX^{t + 1}$ in \equationref{eq:simple_update} simplifies to
 \begin{equation}\label{eq:our_update}
@@
 \begin{corollary}\label{corollary:lin_comb_alg}
-	Let there be a finite family of LMO based algorithms indexed by $i = 1,\dots,n$, where the update of the $i$-th algorithm is defined by
+	Let there be a finite family of LMO based algorithms indexed by $i = 1,\dots,n$, where the $\mX^{t + 1}$ update in the $i$-th algorithm is defined by
 	\[
@@
 \end{equation}
-where $\eta$ is a scaling coefficient specific to SignSGD.
+where \texttt{sign\_lr\_coeff} $\eta$ is a scaling coefficient specific to SignSGD.
 
@@
 
-First, we compare F-Fanions on the following convex $L$-smooth problem:
+We begin by evaluating F-Fanions on the following convex $L$-smooth problem:
 \begin{equation}\label{eq:lls}
@@
 \end{equation}
-where $\mX \in \Rmn$, $m=500$ and $n=500$ are typical dimensions of a neural network weight matrix, $\mS \in \Rmn$, $\mM \in \mathbb{S}^m_{+}$ and $\mN \in \mathbb{S}^n_{+}$ are positive-semidefinite matrices.
-The spectra of $\mM$ and $\mN$ are uniformly distributed in the $(0, 1)$ interval.
-We set $\mS=0$, and $\mX^0 \sim 0.1 \cN(0, \mI)$.
-
-We run different Fanions and their respective F-Fanions with $\alpha=1/2$: Neon (Fanion-1), Fanion-2, Fanion-10, Fanion-100, and Muon (Fanion-500).
-Also, we test S-Fanions with $\alpha=1/2$ and \texttt{sign\_lr\_coeff=0.01}.
-We test them against Normalized SGD and SignSGD, which are also F-Fanion and S-Fanion respectively with $\alpha = 0$ and an arbitrary $k$.
-
-Since theoretical bounds \citep{kovalev2025understanding, riabinin2025gluon} rely on a very loose norm bound $\norm{\cdot}\le \rho \normf{\cdot}$, we do not derive learning rate or Nesterov momentum from the smoothness constants that also depend on the norm choice.
-Rather, we find such \texttt{(lr, momentum)} pair that the loss threshold $0.001$ is reached in a minimal number of iterations.
-This setting is the most realistic, and at the same time aligns with the corollaries of the convergence theorems where constant learning rate and momentum coefficients are proposed.
-
-We perform a grid search for \texttt{momentum} $\in \{0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 0.95\}$ and for \texttt{lr}.
-For Muon's, F-Muon's, S-Muon's \texttt{lr} $\in [0.005, 0.0020]$.
-For SignSGD, \texttt{lr} $\in 0.01 \times [0.005, 0.020]$, and for NSGD \texttt{lr} $\in [0.01, 0.10]$.
-The tuned parameters and the number of iterations to converge to $0.001$ are in \tableref{tbl:lls_lrs}.
-\texttt{lr} and \texttt{momentum}, and \texttt{sign\_lr\_coeff} for Fanions, F-Fanions, and S-Fanions are the same as for Muon, F-Muon, and S-Muon respectively due to the impossibility to tune them properly: their losses decrease too slow to reach 0.001.
-
-The results are presented in \figureref{fig:lls} and detailed in \sectionref{sec:lls_plots_section}.
-
-Both F-Muon and S-Muon converge faster and to a lower loss and a lower Frobenius norm of the gradient than NSGD, Muon, or SignSGD.
+where $\mX \in \Rmn$, $m=500$ and $n=500$ represent typical dimensions of a neural network weight matrix, $\mS \in \Rmn$, and $\mM \in \mathbb{S}^m_{+}$ and $\mN \in \mathbb{S}^n_{+}$ are positive-semidefinite matrices.
+The spectra of $\mM$ and $\mN$ are uniformly distributed on the interval $(0, 1)$.
+We set $\mS=0$ and initialize $\mX^0$ with entries drawn independently from $\cN(0, 0.01)$.
+
+We evaluate several algorithms from the Fanion family and their convex combinations, using the update rule from \equationref{eq:simple_update} with approximate Nesterov momentum:
+\begin{itemize}
+  \item \textbf{Fanions:} Neon (Fanion-1), Fanion-2, Fanion-10, Fanion-100, and Muon (Fanion-500).
+  \item \textbf{F-Fanions:} F-Neon (F-Fanion-1), F-Fanion-2, F-Fanion-10, F-Fanion-100, and F-Muon (F-Fanion-500) with $\alpha=1/2$.
+  \item \textbf{S-Fanions:} S-Neon (S-Fanion-1), S-Fanion-2, S-Fanion-10, S-Fanion-100, and S-Muon (S-Fanion-500) with $\alpha=1/2$ and \texttt{sign\_lr\_coeff=0.01}.
+\end{itemize}
+We also include baselines Normalized SGD and SignSGD, which correspond to F-Fanion and S-Fanion respectively with $\alpha = 0$ and arbitrary $k$.
+
+Since theoretical bounds \citep{kovalev2025understanding, riabinin2025gluon} rely on a loose norm bound $\norm{\cdot}\le \rho \normf{\cdot}$, we do not derive the learning rate or Nesterov momentum from the smoothness constants, which also depend on the choice of norm.
+Instead, we identify the \texttt{(lr, momentum)} pair that reaches the loss threshold of $0.001$ in the minimal number of iterations.
+This setting is both realistic and consistent with the corollaries of convergence theorems that propose constant learning rate and momentum coefficients.
+
+We perform a grid search over the following hyperparameter ranges:
+\begin{itemize}
+  \item \texttt{momentum} $\in \{0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 0.95\}$ for all algorithms.
+  \item \texttt{lr} $\in [0.005, 0.020]$ with step $0.001$ for Muon, F-Muon, and S-Muon.
+  \item \texttt{lr} $\in [5\mathrm{e}{-5},\, 2\mathrm{e}{-4}]$ with step $1\mathrm{e}{-5}$.
+  \item \texttt{lr} $\in [0.01, 0.10]$ with step $0.001$ for NSGD.
+\end{itemize}
+The tuned parameters and the number of iterations required to converge to $0.001$ are presented in \tableref{tbl:lls_lrs}.
+For intermediate-rank Fanions, F-Fanions, and S-Fanions, the hyperparameters \texttt{lr}, \texttt{momentum}, and \texttt{sign\_lr\_coeff} are set equal to those of Muon, F-Muon, and S-Muon respectively, as their loss decreases too slowly to reach $0.001$ within a reasonable tuning budget.
 
@@
 \end{table}
+
+The results are presented in \figureref{fig:lls} with additional details in \sectionref{sec:lls_plots_section}.
 
@@
 
+Both F-Muon and S-Muon converge faster to lower loss values and achieve lower Frobenius norms of the gradient than NSGD, Muon, or SignSGD.
+
 \subsection{CIFAR-10 airbench}
 
-We compare the algorithms on the CIFAR-10 airbench \citep{cifar2023airbench}.
-First, to test the impact of $\alpha$, we run F-Muon for different $\alpha$ with the same \texttt{lr=0.24(1 - step/total\_steps), momentum=0.65, nesterov=True} and weight normalization, as have been finetuned by Keller Jordan, 10 repetitions for each $\alpha$.
-We record the accuracy after 8 epochs of training (\figureref{fig:muon_alphas}).
-
-Then we tune F-Muon with $\alpha = 0.5$.
-Tuned parameters are \texttt{lr=0.4(1 - step/total\_steps), momentum=0.6, nesterov = True}.
-Tuned F-Muon reaches $94.02 \pm 0.13\%$ after 8 epochs (averaged by 200 iterations), matching Muon's accuracy and variance.
-Once again we measure \texttt{val\_accuracy}($\alpha$) curve (\figureref{fig:fmuon_alphas}).
-We notice that even when $\alpha=0.1$, the accuracy is much higher than in the case of vanilla NSGD.
-
-We tune S-Muon with $\alpha=0.5$ for the best \texttt{(lr, momentum, sign\_lr\_coeff)} and by setting \texttt{lr=0.42(1 - step/total\_steps), momentum=0.63, nesterov = True, sign\_lr\_coeff = 0.003} get $\mathbf{94.03} \pm 0.13\%$, which is even higher than for vanilla Muon.
+We evaluate the algorithms on the CIFAR-10 airbench \citep{cifar2023airbench}.
+To assess the impact of the mixing parameter $\alpha$, we first run F-Muon for different values of $\alpha$ using hyperparameters tuned for vanilla Muon by Keller Jordan: \texttt{lr=0.24(1 - step/total\_steps), momentum=0.65, nesterov=True} with weight normalization.
+We perform 10 repetitions for each $\alpha$ value and record the accuracy after 8 epochs of training (\figureref{fig:muon_alphas}).
+
+Next, we tune F-Muon specifically with $\alpha = 0.5$, obtaining optimal hyperparameters \texttt{lr=0.4(1 - step/total\_steps), momentum=0.6, nesterov = True}.
+Tuned F-Muon achieves $94.02 \pm 0.13\%$ validation accuracy after 8 epochs (averaged over 200 runs), matching Muon's accuracy and variance.
+We again measure the validation accuracy as a function of $\alpha$ (\figureref{fig:fmuon_alphas}) and observe that even at $\alpha=0.1$, the accuracy substantially exceeds that of vanilla NSGD.
+
+For S-Muon with $\alpha=0.5$, we tune all hyperparameters to find the optimal configuration \texttt{lr=0.42(1 - step/total\_steps), momentum=0.63, nesterov = True, sign\_lr\_coeff = 0.003}, achieving $\mathbf{94.03} \pm 0.13\%$ validation accuracy, which slightly exceeds vanilla Muon's performance.
 
@@
 
-The Muon-level performance is surprising for the contorted geometry of F-Muon and S-Muon.
-F-Muon can be visualized by \figureref{fig:cifar_ball}, where the LMO balls of Muon and F-Muon for real \texttt{lr} are plotted in a 2D space of singular values.
-S-Muon cannot be visualized because the $l_\infty$ norm is not a function of singular values, while $\norms{\cdot}$, when plotted for the $\Rmn$ space with $m, n > 2$, requires at least a 4D space.
-However, it is plain that the LMO ball of S-Muon significantly differs from the ball of Muon because the $\texttt{lr}(1-\alpha)\texttt{sign\_lr\_coeff} = 6.3 \times 10^{-4}$, which is comparable to the usual \texttt{lr} of SignSGD for deep learning problems.
-
-The most pathological case is $\alpha > 1$, which corresponds to the LMO with an area that is not a norm ball!
-Despite this violation, the mixture of algorithms reaches almost the same accuracy as vanilla Muon.
-
-These observations raise the question of how much LMO-based algorithms are sensitive to the constraint area.
+The Muon-level performance is remarkable given that F-Muon and S-Muon operate with substantially different constraint geometries.
+\figureref{fig:cifar_ball} visualizes F-Muon by plotting the LMO balls of Muon and F-Muon (scaled by actual learning rates) in the 2D space of singular values.
+S-Muon cannot be visualized in this manner because the $l_\infty$ norm is not a function of singular values alone, and visualizing $\norms{\cdot}$ for $\Rmn$ with $m, n > 2$ requires at least 4D space.
+Nevertheless, the LMO ball of S-Muon clearly differs substantially from Muon's, as the effective SignSGD learning rate $\texttt{lr}(1-\alpha)\texttt{sign\_lr\_coeff} = 6.3 \times 10^{-4}$ is comparable to typical SignSGD learning rates in deep learning.
+
+Notably, even the pathological case $\alpha > 1$—which corresponds to an LMO constraint region that is not a norm ball—achieves accuracy nearly identical to vanilla Muon.
+
+These observations raise fundamental questions about the sensitivity of LMO-based algorithms to the shape of the constraint region.
 
@@
 
-We test F-Muon and S-Muon with $\alpha=0.5$ on the NanoGPT speedrun \citep{modded_nanogpt_2024}.
-The optimal parameters are in the legend \figureref{fig:gpt_loss}, \texttt{sign\_sgd\_coeff=0.0003/0.07}.
-After testing for 1750 steps we get 3.281 cross-entropy loss for F-Muon and 3.287 for S-Muon.
-For Muon, it falls below the target threshold 3.28 and reaches 3.279.
-However, this difference is negligible if one looks at \figureref{fig:gpt_loss}.
-It is even more striking, considering the fact that this F-Muon is an average of Muon and NSGD, and the latter performed quite poorly.
-The same is true for S-Muon.
+We evaluate F-Muon and S-Muon with $\alpha=0.5$ against Muon, SignSGD, and NSGD on the NanoGPT speedrun benchmark \citep{modded_nanogpt_2024}.
+The optimal hyperparameters are shown in the legend of \figureref{fig:gpt_loss}, with \texttt{sign\_sgd\_coeff} values of 0.0003/0.07 for S-Muon.
+After 1750 training steps, F-Muon achieves a cross-entropy loss of 3.281, S-Muon achieves 3.287, and Muon achieves 3.279, falling below the target threshold of 3.280.
+As illustrated in \figureref{fig:gpt_loss}, these differences are negligible.
+Notably, F-Muon represents a convex combination of Muon and NSGD, despite NSGD performing poorly in isolation.
+The same observation holds for S-Muon.
 
@@
 \subsection{GPT-2 Medium Speedrun}
-We scale from NanoGPT to GPT-2 Medium.
-After testing for 5960 steps with F-Muon we get 2.9215 loss, while for Muon it is 2.9198.
-In ?
-steps, F-Muon reaches <2.92 loss, thus needing only ?\% more iterations than Muon.
+We scale from NanoGPT to GPT-2 Medium. After testing for 5960 steps we get 2.9215 loss for F-Muon and 2.9235 for S-Muon, while for Muon it is 2.9198 (the speedrun threshold is 2.92).
 
@@
 Common important drawback of the analyses is the consideration of the convergence by the norm of the gradient.
-As we showed in our experiments on CIFAR, the gradient norm may decrease only by a factor of ten, when the accuracy reaches 100\%.
+As we showed in our experiments on CIFAR \cref{sec:cifar_underhood}, the gradient norm may decrease only by a factor of ten even when the accuracy reaches 100\%.
 
@@
 \section{Under the hood of CNN on CIFAR-10}
+\label{sec:cifar_underhood}
 

### 2025-11-30 article_final/main.tex
updated subsection 6.1. IMPORTANT: Check the formula for nesterov update

--- 2025-11-30 article_final/main.tex
@@
 \end{equation}
-where $\mM^t$ is the gradient (or momentum) of $F$ at $\mX^t$ and $\cS \subset \Rmn$ is some set. The update of the algorithm is defined as follows:
+where $\mM^t$ is the gradient or momentum buffer and $\cS \subset \Rmn$ is some constraint set. The algorithm proceeds as follows: first compute the momentum buffer $\mM^t$, then update $\mX^{t+1}$:
 \begin{equation}\label{eq:simple_update}
-	\mX^{t + 1} = \mX^t + \gamma_t\mathrm{LMO}\left(\mM^t\right)\,.
-\end{equation}
+	\begin{aligned}
+	\mM^{t} &= \begin{cases}
+		\nabla F(\mX^t) & \text{(no momentum)},\\
+		\beta \mM^{t-1} + (1 - \beta) \nabla F(\mX^t) & \text{(heavy ball)},\\
+    \beta \mM^{t-1} + (1 - \beta) \nabla F\left(\mX^{t - 1} + \gamma_t\mathrm{LMO}\left(\mM^{t-1}\right)\right) & \text{(Nesterov)},
+	\end{cases}\\
+	\mX^{t + 1} &= \mX^t + \gamma_t\mathrm{LMO}\left(\mM^t\right)\,.\\
+	\end{aligned}
+\end{equation}
+Here $\beta \in [0, 1)$ is the momentum coefficient. Throughout our experiments, we employ Nesterov momentum unless otherwise specified.
 
@@
 \end{equation}
-where $\mX \in \Rmn$, $m=500$ and $n=500$ are typical dimensions of a neural network weight matrix, $\mS \in \Rmn$, $\mM \in \mathbb{S}^m_{+}$ and $\mN \in \mathbb{S}^n_{+}$ are positive-semidefinite matrices.
-The spectra of $\mM$ and $\mN$ are uniformly distributed in the $(0, 1)$ interval.
-We set $\mS=0$, and $\mX^0 \sim 0.1 \cN(0, \mI)$.
-
-We run different Fanions and their respective F-Fanions with $\alpha=1/2$: Neon (Fanion-1), Fanion-2, Fanion-10, Fanion-100, and Muon (Fanion-500).
-Also, we test S-Fanions with $\alpha=1/2$ and \texttt{sign\_lr\_coeff=0.01}.
-We test them against Normalized SGD and SignSGD, which are also F-Fanion and S-Fanion respectively with $\alpha = 0$ and an arbitrary $k$.
-
-Since theoretical bounds \citep{kovalev2025understanding, riabinin2025gluon} rely on a very loose norm bound $\norm{\cdot}\le \rho \normf{\cdot}$, we do not derive learning rate or Nesterov momentum from the smoothness constants that also depend on the norm choice.
-Rather, we find such \texttt{(lr, momentum)} pair that the loss threshold $0.001$ is reached in a minimal number of iterations.
-This setting is the most realistic, and at the same time aligns with the corollaries of the convergence theorems where constant learning rate and momentum coefficients are proposed.
-
-We perform a grid search for \texttt{momentum} $\in \{0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 0.95\}$ and for \texttt{lr}.
-For Muon's, F-Muon's, S-Muon's \texttt{lr} $\in [0.005, 0.0020]$.
-For SignSGD, \texttt{lr} $\in 0.01 \times [0.005, 0.020]$, and for NSGD \texttt{lr} $\in [0.01, 0.10]$.
-The tuned parameters and the number of iterations to converge to $0.001$ are in \tableref{tbl:lls_lrs}.
-\texttt{lr} and \texttt{momentum}, and \texttt{sign\_lr\_coeff} for Fanions, F-Fanions, and S-Fanions are the same as for Muon, F-Muon, and S-Muon respectively due to the impossibility to tune them properly: their losses decrease too slow to reach 0.001.
-
-The results are presented in \figureref{fig:lls} and detailed in \sectionref{sec:lls_plots_section}.
-
-Both F-Muon and S-Muon converge faster and to a lower loss and a lower Frobenius norm of the gradient than NSGD, Muon, or SignSGD.
+where $\mX \in \Rmn$, $m=500$ and $n=500$ represent typical dimensions of a neural network weight matrix, $\mS \in \Rmn$, and $\mM \in \mathbb{S}^m_{+}$ and $\mN \in \mathbb{S}^n_{+}$ are positive-semidefinite matrices.
+The spectra of $\mM$ and $\mN$ are uniformly distributed on the interval $(0, 1)$.
+We set $\mS=0$ and initialize $\mX^0$ with entries drawn independently from $\cN(0, 0.01)$.
+
+We evaluate several algorithms from the Fanion family and their convex combinations, using the update rule from \equationref{eq:simple_update} with Nesterov momentum:
+\begin{itemize}
+  \item \textbf{Fanions:} Neon (Fanion-1), Fanion-2, Fanion-10, Fanion-100, and Muon (Fanion-500).
+  \item \textbf{F-Fanions:} F-Neon (F-Fanion-1), F-Fanion-2, F-Fanion-10, F-Fanion-100, and F-Muon (F-Fanion-500) with $\alpha=1/2$.
+  \item \textbf{S-Fanions:} S-Neon (S-Fanion-1), S-Fanion-2, S-Fanion-10, S-Fanion-100, and S-Muon (S-Fanion-500) with $\alpha=1/2$ and \texttt{sign\_lr\_coeff=0.01}.
+\end{itemize}
+We also include baselines Normalized SGD and SignSGD, which correspond to F-Fanion and S-Fanion respectively with $\alpha = 0$ and arbitrary $k$.
+
+Since theoretical bounds \citep{kovalev2025understanding, riabinin2025gluon} rely on a loose norm bound $\norm{\cdot}\le \rho \normf{\cdot}$, we do not derive the learning rate or Nesterov momentum from the smoothness constants, which also depend on the choice of norm.
+Instead, we identify the \texttt{(lr, momentum)} pair that reaches the loss threshold of $0.001$ in the minimal number of iterations.
+This setting is both realistic and consistent with the corollaries of convergence theorems that propose constant learning rate and momentum coefficients.
+
+We perform a grid search over the following hyperparameter ranges:
+\begin{itemize}
+  \item \texttt{momentum} $\in \{0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 0.95\}$ for all algorithms.
+  \item \texttt{lr} $\in [0.005, 0.020]$ with step $0.001$ for Muon, F-Muon, and S-Muon.
+  \item \texttt{lr} $\in [0.00005, 0.00020]$ with step $0.00001$ for SignSGD.
+  \item \texttt{lr} $\in [0.01, 0.10]$ with step $0.001$ for NSGD.
+\end{itemize}
+The tuned parameters and the number of iterations required to converge to $0.001$ are presented in \tableref{tbl:lls_lrs}.
+For intermediate-rank Fanions, F-Fanions, and S-Fanions, the hyperparameters \texttt{lr}, \texttt{momentum}, and \texttt{sign\_lr\_coeff} are set equal to those of Muon, F-Muon, and S-Muon respectively, as their loss decreases too slowly to reach $0.001$ within a reasonable tuning budget.
 
@@
 \end{figure}
+
+Both F-Muon and S-Muon converge faster to lower loss values and achieve lower Frobenius norms of the gradient than NSGD, Muon, or SignSGD.
 

### 2025-11-30 article_final/main.tex

--- 2025-11-30 article_final/main.tex
@@
 
-We then compare the performance of these algorithm families on various model and real-world problems:
+After that in \sectionref{sec:experiments} compare the performance of these algorithm families on various model and real-world problems:
 \begin{itemize}
-	\item Synthetic least squares experiment \sectionref{subsec:lls_exps}
+	\item Synthetic least squares experiment
 	\item CIFAR-10 airbench \citep{cifar2023airbench}
 	\item Pre-training NanoGPT and GPT-2 Medium on FineWeb dataset \citep{modded_nanogpt_2024}
-	\item Fine-tuning NanoGPT on Tiny Stories \citep{eldan2023tinystoriessmalllanguagemodels}
+	\item Fine-tuning NanoGPT on Tiny Stories dataset \citep{eldan2023tinystoriessmalllanguagemodels}
 \end{itemize}
@@
 Training a neural network is essentially an optimization of a function of several weight matrices and a few vectors.
-Let us start by considering the problem of minimizing a differentiable function of a \emph{single} matrix:
+Let us start by considering the problem of minimizing a differentiable function of a \emph{single} matrix (the connection to the general case is presented in \sectionref{sec:lmo_for_nn}):
 \begin{equation}\label{eq:matrix_problem}
@@
 \]
-The Neon update $\mD = \vu_1 \vv_1^\top$ achieves $\sigma_1$, while the scaled Muon update $\mD = \frac{1}{k}\sum_{i=1}^{\min(m,n)}\vu_i \vv_i^\top$ achieves $\frac{1}{k}\sum_{i=1}^{\min(m,n)}\sigma_i$.
+The Neon update $\mD = \vu_1 \vv_1^\top$ achieves $\sigma_1$, while the scaled Muon update $\mD = \frac{1}{k}\mU\mV^\top$ achieves $\frac{1}{k}\sum_{i=1}^{\min(m,n)}\sigma_i$.
 Thus, the update is either
@@
 	\]
-	where the last summation denotes the Minkowski sum. This immediately implies $\sum_{i = 1}^n \alpha_i \mathrm{LMO}_{i} \in \argmin_{\mD \in \cB_{\norm{\cdot}}}\<\mM, \mD>$, completing the proof.
+	where the last summation denotes the Minkowski sum. This immediately implies 
+  \[
+  \sum_{i = 1}^n \alpha_i \mathrm{LMO}_{i} \in \argmin_{\mD \in \cB_{\norm{\cdot}}}\<\mM, \mD>\,,
+  \] 
+  completing the proof.
 \end{proof}
@@
 \begin{equation}\label{eq:ckfk_norm}
-	\norm{\cdot}_{\mathrm{C-KF-k}} = \alpha \normkfk{\cdot} + \frac{1-\alpha}{\eta}\normc{\cdot}^\dagger\,.
+	\norm{\cdot}_{\mathrm{C-KF-k}} = \alpha \normkfk{\cdot} + \frac{1-\alpha}{\eta}\normc{\cdot}^\dagger\,.sec:lmo_for_nn
 \end{equation}
@@
 
-Next, we tune F-Muon specifically with $\alpha = 0.5$, obtaining optimal hyperparameters \texttt{lr=0.4(1 - step/total\_steps), momentum=0.6, nesterov = True}.
+Next, we tune F-Muon specifically with $\alpha = 0.5$, obtaining optimal hyperparameters \texttt{lr=0.4(1 - step/\allowbreak{}total\_steps)}, \texttt{momentum=0.6}, \texttt{nesterov = True}.
 Tuned F-Muon achieves $94.02 \pm 0.13\%$ validation accuracy after 8 epochs (averaged over 200 runs), matching Muon's accuracy and variance.
@@
 
-For S-Muon with $\alpha=0.5$, we tune all hyperparameters to find the optimal configuration \texttt{lr=0.42(1 - step/total\_steps), momentum=0.63, nesterov = True, sign\_lr\_coeff = 0.003}, achieving $\mathbf{94.03} \pm 0.13\%$ validation accuracy, which slightly exceeds vanilla Muon's performance.
+For S-Muon with $\alpha=0.5$, we tune all hyperparameters to find the optimal configuration \texttt{lr=0.42(1 - step/\allowbreak{}total\_steps)}, \texttt{momentum=0.63}, \texttt{nesterov = True}, \texttt{sign\_lr\_coeff = 0.003}, achieving $94.03 \pm 0.13\%$ validation accuracy, which slightly exceeds vanilla Muon's performance.
 
@@
 
-Again, as on the CIFAR airbench, if for F-Muon we set $\alpha=-0.1$ that corresponds to no ball, we get a 3.2818 loss, which is a small difference from the baseline.
-
-It is noteworthy that F-Muon with NSGD-part that is not scaled by Muon's rule gives 3.288 loss (0.007 increase), while S-Muon scaled with Muon's rule gives 3.292 loss (0.005 increase)
+As observed in the CIFAR airbench experiments, setting $\alpha=-0.1$ for F-Muon (corresponding to the no-ball configuration) yields a loss of 3.2818, representing only a marginal difference from the baseline.
+
+Notably, F-Muon with the NSGD component not scaled by Muon's rule yields a loss of 3.288 (an increase of 0.007), while S-Muon scaled with Muon's rule achieves a loss of 3.292 (an increase of 0.005).
 
@@
 
-We fine-tuned the NanoGPT framework by Karpathy \citep{Karpathy2022} using the standard \textit{NanoGPT-medium} configuration, which corresponds to GPT-2 Medium.
-This model consists of 24 transformer layers, a hidden size of 1024, 16 attention heads, and approximately 345 million parameters.
-The training dataset was the \textit{Tiny Stories} corpus \citep{eldan2023tinystoriessmalllanguagemodels}, selected for its high entropy and structural diversity, which amplify and make more visible the differences in optimization dynamics.
-All training runs were initialized with the same random seed, weight initialization, and learning rate schedule to ensure that any differences in performance arose solely from the choice of optimizer.
-
-For all one-dimensional layers, AdamW was used with a learning rate of $1\!\times\!10^{-3}$.
+We fine-tuned the NanoGPT framework by Karpathy \citep{Karpathy2022} using the standard \emph{NanoGPT-medium} configuration, which corresponds to GPT-2 Medium.
+This model consists of 24 transformer layers, a hidden dimension of 1024, 16 attention heads, and approximately 345 million parameters.
+We selected the \emph{Tiny Stories} corpus \citep{eldan2023tinystoriessmalllanguagemodels} as the training dataset for its high entropy and structural diversity, which amplify the differences in optimization dynamics.
+All training runs were initialized with the same random seed, weight initialization, and learning rate schedule to ensure that performance differences arose solely from the choice of optimizer.
+
+For all one-dimensional layers, we used AdamW with a learning rate of $1\!\times\!10^{-3}$.
 Since momentum exhibited negligible influence on training dynamics, it was held constant across all experiments.
-A comparative analysis of optimization performance is provided in \figureref{fig:nanogptftexps}.
+\figureref{fig:nanogptftexps} presents a comparative analysis of optimization performance.
 
@@
   {fig:nanogptftexps}
-  {\caption{Comparison of Muon, F-Muon, and S-Muon across a range of learning rates. Stars denote the best learning rate identified for each optimizer. Validation loss.}}
-  {\includegraphics[width=\linewidth]{figs/nanogpt/optimizers_comparison.pdf}}
-\end{figure}
-
-\figureref{fig:nanogptftexps2} presents the train and validation loss curves at the optimal learning rate for each optimizer.
-While F-Muon and S-Muon maintain consistent behavior, vanilla Muon attains the lowest loss overall.
+  {\caption{Comparison of validation loss for Muon, F-Muon, and S-Muon across a range of learning rates. Stars denote the best learning rate identified for each optimizer.}}
+  {\includegraphics[width=0.8\linewidth]{figs/nanogpt/optimizers_comparison.pdf}}
+\end{figure}
+
+\figureref{fig:nanogptftexps2} presents the training and validation loss curves at the optimal learning rate for each optimizer.
+While F-Muon and S-Muon maintain consistent behavior, vanilla Muon achieves the lowest loss overall, outperforming other algorithms by a small margin.
 
@@
   {fig:nanogptftexps2}
-  {\caption{Train and validation loss at the optimal learning rate for each optimizer.
-The standard Muon method achieves the lowest loss, while F-Muon and S-Muon remain stable but exhibit slightly inferior performance.}}
+  {\caption{Train and validation loss at the optimal learning rate for each optimizer.}}
   {\includegraphics[width=\linewidth]{figs/nanogpt/Nanogpt finetune results.pdf}}
@@
 
-\section{Algorithms for Vectors $\rightarrow$ Algorithms for Matrices}
-
-One cannot but notice the similarity between the updates of LMO optimizers in the vector $l_p$ and the matrix Schatten $S_p$ norms, which is illustrated by \tableref{tbl:mat_vs_vec_lmo}.
-The parallels are not limited to mere similarity of the updates and can be traced even in the empirical observations.
-SignSGD is very close to Adam, as noted in \citep{bernstein2024oldoptimizernewnorm}, and both Adam and Muon perform well in training large models \citep{zhao2025deconstructing, liu2025muon}.
-NSGD is the same for both matrix and vector cases.
-Greedy Coordinate Descent is not applied to high-dimensional problems, so from this perspective it is not surprising that one-rank Neon underperforms on such problems.
-
-\begin{table}[t]
+\section{Related Work and Discussion}
+
+\subsection{Algorithms for Vectors $\rightarrow$ Algorithms for Matrices}
+
+The updates of LMO optimizers exhibit striking similarities between vector $l_p$ norms and matrix Schatten $S_p$ norms, as illustrated in \tableref{tbl:mat_vs_vec_lmo}.
+These parallels extend beyond the update formulas themselves to empirical performance characteristics.
+SignSGD closely resembles Adam, as noted in \citep{bernstein2024oldoptimizernewnorm}, and both Adam and Muon perform well when training large models \citep{zhao2025deconstructing, liu2025muon}.
+NSGD maintains identical updates in both vector and matrix cases.
+Greedy Coordinate Descent is rarely applied to high-dimensional problems, which provides perspective on why one-rank Neon underperforms in such settings.
+
+\begin{table}[htbp]
 \floatconts
@@
 
-\section{Related Work and Discussion}
-
-Since Muon \citep{jordan2024muon} is a very efficient optimizer for functions of weight matrices, a lot of research has been put into, first, further improving its performance, and second, explaining its success.
-
-\paragraph{Improvements of Muon} Regarding the first point, a large number of applications and improvements of Muon has been proposed in less than a year.
+\subsection{Improvements of Muon}
+
+Since Muon \citep{jordan2024muon} is a highly efficient optimizer for functions of weight matrices, substantial research has focused on two objectives: further improving its performance and explaining its success.
+
+A large number of applications and improvements of Muon have been proposed in less than a year.
 \citep{liu2025muon} adapted the algorithm for training language models larger than NanoGPT.
-\citep{shah2025practical} organized efficient hyperparameter transfer by combining Muon with maximal update parametrization.
-To construct their COSMOS optimizer, \citep{chen2025cosmoshybridadaptive} applied computationally intensive updates of SOAP optimizer \citep{vyas2025soap} to a low-dimensional ``leading eigensubspace'' while using memory-efficient methods like Muon for the remaining parameters.
-\citep{amsel2025polar, grishina2025chebyshev} proposed more efficient alternatives to Newton-Schulz algorithm.
-\citep{si2025adamuon} introduced AdaMuon which combines element-wise adaptivity with orthogonal updates.
-We suppose that the described above or similar techniques can be applied to our optimizers as well.
-For example, F-Muon and S-Muon also benefit from faster alternatives to Newton-Schulz iterations, and Fanions may be a great substitute to Muon in COSMOS, because, as we have shown in \sectionref{sec:matrix-side}, Lanczos algorithm is much faster than Newton-Schulz iterations on very large matrices.
-
-\paragraph{Theory behind Muon} Regarding the second point, there has been a prolonged gap in theory behind Muon, simplistic derivation of \citep{bernstein2025deriving} based on \citep{bernstein2024oldoptimizernewnorm} excluded.
-This gap, as it seems to us, is not even now completely closed.
-For example, \citep{kovalev2025understanding} has provided convergence guarantees of Muon and in various settings, from which, however, Muon's supremacy cannot be recovered.
-Indeed, although the obtained bounds depend on the norm choice, the asymptotics of the convergence remain the same as for NSGD and other optimizers, $K = \cO(\epsilon^{-4})$ in an $L$-smooth stochastic case.
-
-A similar drawback has the article \citep{riabinin2025gluon}, where $L$-smoothness assumption is replaced with a more practical $(L_0, L_1)$-smoothness.
-By estimating smoothness and substituting it into their Theorem 1, the authors recovered the optimal fine-tuned stepsizes reported by \citep{pethick2025training}.
-However, they have not showed the optimality of the spectral or the RMS-to-RMS norm, which is observed in practice, as our comparison with NSGD highlights.
-
-Common important drawback of the analyses is the consideration of the convergence by the norm of the gradient.
-As we showed in our experiments on CIFAR \cref{sec:cifar_underhood}, the gradient norm may decrease only by a factor of ten even when the accuracy reaches 100\%.
-
-We suppose that the reason for the recorded stark discrepancy between Neon and Muon performance, both of which are described by Stochastic Conditional Gradient \citep{pethick2025training} or Gluon frameworks, lies in the structure of the norm ball or in the preconditioner interpretation \citep{pethick2025trainingneuralnetworksscale}, which must be an object of further research.
-
-\paragraph{The LMO and Error Feedback} We already mentioned that rank-$k$ unsharded Dion without error feedback and scaling of the update is Fanion-$k$.
-Since error feedback for Dion is very important as discovered by the ablation study in \citep{ahn2025dion}, F-Fanions and S-Fanions will benefit from it as well.
-In the context of federated learning, error feedback is effective even for compressed Muon \citep{gruntkowska2025efmuon}.
-The advantage of Fanions and S-Fanions is that less bits are required for transmission: $\sum_{i=1}^k \vu_i \vv_i^\top$ is easily transmitted as $\{\vu_1, \vv_1, \dots, \vu_k, \vv_k\}$ ($(m+n)\times k$ floats), while the sign part is as usual coded in $m \times n$ bits.
-Thus, the compression is ``built-in''.
-Moreover, there is an intriguing possibility to construct differentially private Fanions and S-Fanions with the specific more optimal non-Gaussian noise, as was done with DP-signSGD \citep{jang2024dplogsign}.
-This we leave for future research.
-
-\paragraph{The nuclear norm in the LMO} As we found out only when transforming our results into an article, the nuclear norm has already been explored in the context of the linear minimization oracle.
-\citep{pethick2025sam} applied it to create $\nu$SAM, a new sharpness-aware minimization technique.
+\citep{shah2025practical} enabled efficient hyperparameter transfer by combining Muon with maximal update parametrization.
+To construct their COSMOS optimizer, \citep{chen2025cosmoshybridadaptive} applied computationally intensive updates of the SOAP optimizer \citep{vyas2025soap} to a low-dimensional ``leading eigensubspace'' while using memory-efficient methods like Muon for the remaining parameters.
+\citep{amsel2025polar, grishina2025chebyshev} proposed more efficient alternatives to the Newton-Schulz algorithm.
+\citep{si2025adamuon} introduced AdaMuon, which combines element-wise adaptivity with orthogonal updates.
+We expect that similar techniques can be applied to our optimizers as well.
+For example, F-Muon and S-Muon also benefit from faster alternatives to Newton-Schulz iterations, and Fanions may serve as an effective substitute for Muon in COSMOS, as we have shown in \sectionref{sec:matrix-side} that the Lanczos algorithm is substantially faster than Newton-Schulz iterations on very large matrices.
+
+\subsection{Theory behind Muon}
+
+There has been a prolonged gap in the theoretical understanding of Muon, excluding the simplistic derivation of \citep{bernstein2025deriving} based on \citep{bernstein2024oldoptimizernewnorm}.
+This gap, in our view, remains incompletely closed.
+For example, \citep{kovalev2025understanding} provided convergence guarantees for Muon in various settings, from which, however, Muon's empirical supremacy cannot be recovered.
+Indeed, although the obtained bounds depend on the norm choice, the convergence asymptotics remain the same as for NSGD and other optimizers: $K = \cO(\epsilon^{-4})$ in the $L$-smooth stochastic case.
+
+A similar limitation affects \citep{riabinin2025gluon}, where the $L$-smoothness assumption is replaced with a more practical $(L_0, L_1)$-smoothness.
+By estimating smoothness and substituting it into their Theorem 1, the authors recovered the optimal fine-tuned step sizes reported by \citep{pethick2025training}.
+However, they did not demonstrate the optimality of the spectral or RMS-to-RMS norm, which is observed in practice, as our comparison with NSGD highlights.
+
+A common limitation of these analyses is their focus on convergence measured by the gradient norm.
+As we showed in our CIFAR experiments \cref{sec:cifar_underhood}, the gradient norm may decrease by only a factor of ten when the accuracy reaches 100\%.
+
+We hypothesize that the stark performance discrepancy between Neon and Muon, both of which are described by the Stochastic Conditional Gradient \citep{pethick2025training} or Gluon frameworks, lies in the structure of the norm ball or in the preconditioner interpretation \citep{pethick2025trainingneuralnetworksscale}, which warrants further investigation.
+
+\subsection{The LMO and Error Feedback}
+
+As previously mentioned, rank-$k$ unsharded Dion without error feedback and update scaling is equivalent to Fanion-$k$.
+Since error feedback is crucial for Dion, as demonstrated by the ablation study in \citep{ahn2025dion}, F-Fanions and S-Fanions would benefit from it as well.
+In federated learning, error feedback proves effective even for compressed Muon \citep{gruntkowska2025efmuon}.
+Fanions and S-Fanions offer a transmission advantage, requiring fewer bits: $\sum_{i=1}^k \vu_i \vv_i^\top$ can be efficiently transmitted as $\{\vu_1, \vv_1, \dots, \vu_k, \vv_k\}$ ($(m+n)\times k$ floats), while the sign component can be encoded in $m \times n$ bits.
+Thus, compression is inherently built into the representation.
+Moreover, there is an intriguing possibility to construct differentially private Fanions and S-Fanions using more optimal non-Gaussian noise, as was done with DP-SignSGD \citep{jang2024dplogsign}.
+We leave this for future research.
+
+\subsection{The nuclear norm in the LMO}
+
+We discovered during the preparation of this article that the nuclear norm has already been explored in the context of the linear minimization oracle.
+\citep{pethick2025sam} applied it to create $\nu$SAM, a novel sharpness-aware minimization technique.
 It would be interesting to substitute $\normn{\cdot}$ with $\normkfk{\cdot}^\dagger$ in their approach.
-Since the SAM-neighborhood becomes more diverse, $k > 1$ might add to the accuracy boost, while preserving a small memory footprint and a small time overhead, if Dion-style power iterations are used.
-
-\paragraph{Orthogonal research} Fanions, F-Fanions, and S-Fanions benefit from the general theoretical description of \citep{riabinin2025gluon}, and better learning rates can be predicted by calculating the trajectory smoothness.
-They could be transformed to Drop-Fanions by updating only some layers as in \citep{gruntkowska2025dropmuon}.
-They could be viewed as approximations of the Non-Euclidean Broximal Point Method for the corresponding norms \citep{gruntkowska2025noneuclideanbroximal}.
-One can clip them to produce ClippedScion-like algorithms \citep{pethick2025generalizedgradientnormclipping}.
-They could be made more memory-effective by the zero-order techniques that worked for Muon \citep{petrov2025zeromuon}.
-Finally, the results from \citep{shulgin2025inexactmuon} can be used to explain the robustness of Muon to the norm change that occurred in our experiments and to theoretically derive faster yet working approximate schemes to calculate the LMO; power iterations with a very limited number of iterations is a good algorithm to consider for the analysis.
+Since the SAM neighborhood becomes more diverse, using $k > 1$ might enhance the accuracy boost while preserving a small memory footprint and minimal time overhead when Dion-style power iterations are employed.
+
+\subsection{Orthogonal research}
+
+Fanions, F-Fanions, and S-Fanions benefit from the general theoretical description in \citep{riabinin2025gluon}, where better learning rates can be predicted by calculating the trajectory smoothness.
+They could be transformed into Drop-Fanions by updating only selected layers, as in \citep{gruntkowska2025dropmuon}.
+They can be viewed as approximations of the Non-Euclidean Proximal Point Method for the corresponding norms \citep{gruntkowska2025noneuclideanbroximal}.
+They can be clipped to produce ClippedScion-like algorithms \citep{pethick2025generalizedgradientnormclipping}.
+They could be made more memory-efficient through the zero-order techniques that proved effective for Muon \citep{petrov2025zeromuon}.
+Finally, the results from \citep{shulgin2025inexactmuon} can be used to explain the robustness of Muon to the norm changes observed in our experiments and to theoretically derive faster yet effective approximate schemes for calculating the LMO; power iterations with a limited number of iterations represent a promising direction for this analysis.
 
@@
 
-In this article, we have generalized several successful algorithms, like Muon, Dion and $\nu$SAM, to the LMO-based algorithms in the $\normkfk{\cdot}^\dagger$ norm.
-Also we have proposed the technique of ``regularizing'' the updates with NSGD or SignSGD, a trick to increase the robustness of the algorithms, which is motivated by the consideration of the $\norm{\cdot}_{\mathrm{F-KF-k}}$ and $\norm{\cdot}_{\mathrm{C^\dagger-KF-k}}$ norms.
-We suggest that future works that deal with the non-Euclidean LMO explain in the corollaries to their convergence theorems the empirical superiority of Muon to other Fanions, and ideally accont for S-Fanion and F-Muon robustness as well.
-Without that, it is hard to believe that the bounds are relevant for practitioners, as we see by the notorious example of one-rank Neon.
+In this article, we addressed the central question of why one should constrain by the spectral or any other operator norm in deriving Muon-like updates, and how alternative norms affect performance and computational cost.
+Our answer is that the choice of matrix norm is remarkably flexible: properly-tuned variants based on alternative norms can match or even slightly exceed Muon's performance on real-world tasks, while offering additional benefits such as improved learning rate robustness.
+
+We generalized several successful algorithms, including Muon, Dion, and $\nu$SAM, to LMO-based algorithms using the family of norms dual to Ky Fan $k$-norms, yielding the Fanion family with low-rank updates.
+We further proposed the technique of regularizing these updates by combining them with NSGD or SignSGD, creating the F-Fanion and S-Fanion families through the $\norm{\cdot}_{\mathrm{F-KF-k}}$ and $\norm{\cdot}_{\mathrm{C^\dagger-KF-k}}$ norms.
+Our experiments demonstrate that F-Muon and S-Muon achieve competitive performance with Muon on CIFAR-10 airbench and NanoGPT tasks, confirming that the underlying norm constraint can be significantly modified without sacrificing effectiveness.
+
+However, our results also reveal that not every LMO-based algorithm is effective: Neon (rank-one Fanion) underperforms despite sharing the same theoretical convergence asymptotics as Muon in existing bounds.
+We suggest that future work on non-Euclidean LMO algorithms should explain in the corollaries to their convergence theorems the empirical superiority of Muon over other Fanions, and ideally account for the robustness of F-Muon and S-Muon as well.
+Without such refinements, it is difficult to believe that current theoretical bounds are relevant for practitioners.
 
@@
 DM presented the problem at the MIPT optimization course.
-IK suggested using composite norms (though not the ones that induce F-Fanions and S-Fanions) and refactored \sectionref{sec:conic_combo}.
+IK suggested using composite norms (though not the ones that induce F-Fanions and S-Fanions) and helped to draft and rewise the manuscript.
 NK suggested Lanczos algorithm as the most precise means to compute Fanions' updates, conducted experiments to prove this, and wrote \sectionref{sec:matrix-side}.
@@
 \appendix
-\section{LMO for Neural Networks}
+\section{LMO for Neural Networks}\label{sec:lmo_for_nn}
 
@@
 The LMO-based update rule is then applied to each matrix $\mX_i$ using its corresponding gradient component $\mG_i^t$.
-
-The update rule used in Muon optimizer is uSCG \citep{pethick2025training}. In the most general case, which involves momentum, it can be written as
-\begin{equation}\label{eq:usgd_update}
-  \begin{aligned}
-    \mM^{t}     & = \alpha_{t} g(\mX^t, \xi^t) + (1 - \alpha_t)\mM^{t - 1}\,, \\
-    \mX^{t + 1} & = \mX^t + \gamma_t\mathrm{LMO}(\mM^{t})\,.
-  \end{aligned}
-\end{equation}
-
-\section{Norms $\normfstar{\cdot}^\dagger$ and $\normftwo{\cdot}^\dagger$}
-
-Based on the lemma, we immediately find $\normfstar{\cdot}^\dagger$, which is related to F-Muon update.
-Indeed, after setting $\beta=1-\alpha$ and remembering that for smooth and bounded cases we can use $\min$ instead of $\inf$, we get
-\begin{equation}\label{eq:normfstardual}
-  \normfstar{\mY}^\dagger = \min_{\mZ} \min_t\{t, s.t. \norms{\mZ}\leq \alpha t, \normf{\mY-\mZ}\leq (1-\alpha) t\}
-\end{equation}
-
-If $\alpha = 1$, then $\mZ = \mY$, and we get $\normfstar{\mY}^\dagger = \norms{\mY}$. If $\alpha = 0$, then $\mZ = 0$, and we get $\normfstar{\mY}^\dagger = \normf{\mY}$.
-
-Similarly, we find $\normftwo{\cdot}^\dagger$, which is related to F-Neon update:
-\begin{equation}\label{eq:normftwodual}
-  \normftwo{\mY}^\dagger = \min_{\mZ} \min_t\{t, s.t. \normn{\mZ}\leq \alpha t, \normf{\mY-\mZ}\leq (1-\alpha) t\}
-\end{equation}
-
-If $\alpha = 1$, then $\mZ = \mY$, and we get $\normftwo{\mY}^\dagger = \normn{\mY}$. If $\alpha = 0$, then $\mZ = 0$, and we get $\normftwo{\mY}^\dagger = \normf{\mY}$.
-
-Unfortunately, $\normfstar{\cdot}^\dagger$ and $\normftwo{\cdot}^\dagger$ do not have a closed-form expression.
 
@@
 \end{proof}
+
+\section{Norms $\normfstar{\cdot}^\dagger$ and $\normftwo{\cdot}^\dagger$}
+
+Based on the \lemmaref{lemma:dual_to_conv_comb}, we immediately find $\normfstar{\cdot}^\dagger$, which is related to F-Muon update.
+Indeed, after setting $\beta=1-\alpha$ and remembering that for smooth and bounded cases we can use $\min$ instead of $\inf$, we get
+\begin{equation}\label{eq:normfstardual}
+  \normfstar{\mY}^\dagger = \min_{\mZ} \min_t\{t, s.t. \norms{\mZ}\leq \alpha t, \normf{\mY-\mZ}\leq (1-\alpha) t\}
+\end{equation}
+
+If $\alpha = 1$, then $\mZ = \mY$, and we get $\normfstar{\mY}^\dagger = \norms{\mY}$. If $\alpha = 0$, then $\mZ = 0$, and we get $\normfstar{\mY}^\dagger = \normf{\mY}$.
+
+Similarly, we find $\normftwo{\cdot}^\dagger$, which is related to F-Neon update:
+\begin{equation}\label{eq:normftwodual}
+  \normftwo{\mY}^\dagger = \min_{\mZ} \min_t\{t, s.t. \normn{\mZ}\leq \alpha t, \normf{\mY-\mZ}\leq (1-\alpha) t\}
+\end{equation}
+
+If $\alpha = 1$, then $\mZ = \mY$, and we get $\normftwo{\mY}^\dagger = \normn{\mY}$. If $\alpha = 0$, then $\mZ = 0$, and we get $\normftwo{\mY}^\dagger = \normf{\mY}$.
+
+Unfortunately, $\normfstar{\cdot}^\dagger$ and $\normftwo{\cdot}^\dagger$ do not have a closed-form expression.
 

### 2025-11-30 article_final/main.tex
fixed Nesterov formulation

--- 2025-11-30 article_final/main.tex
@@
 	\item CIFAR-10 airbench \citep{cifar2023airbench}
-	\item Pre-training NanoGPT and GPT-2 Medium on FineWeb dataset \citep{modded_nanogpt_2024}
+	\item Pre-training NanoGPT on FineWeb dataset \citep{modded_nanogpt_2024}
 	\item Fine-tuning NanoGPT on Tiny Stories \citep{eldan2023tinystoriessmalllanguagemodels}
@@
 \end{equation}
-where \texttt{sign\_lr\_coeff} $\eta$ is a scaling coefficient specific to SignSGD.
+where $\eta$ is a scaling coefficient specific to SignSGD.
 
@@
   \item \texttt{lr} $\in [0.005, 0.020]$ with step $0.001$ for Muon, F-Muon, and S-Muon.
-  \item \texttt{lr} $\in [5\mathrm{e}{-5},\, 2\mathrm{e}{-4}]$ with step $1\mathrm{e}{-5}$.
+  \item \texttt{lr} $\in [0.00005, 0.00020]$ with step $0.00001$ for SignSGD.
   \item \texttt{lr} $\in [0.01, 0.10]$ with step $0.001$ for NSGD.
@@
 
-We evaluate F-Muon and S-Muon with $\alpha=0.5$ against Muon, SignSGD, and NSGD on the NanoGPT speedrun benchmark \citep{modded_nanogpt_2024}.
-The optimal hyperparameters are shown in the legend of \figureref{fig:gpt_loss}, with \texttt{sign\_sgd\_coeff} values of 0.0003/0.07 for S-Muon.
-After 1750 training steps, F-Muon achieves a cross-entropy loss of 3.281, S-Muon achieves 3.287, and Muon achieves 3.279, falling below the target threshold of 3.280.
+We evaluate F-Muon and S-Muon with $\alpha=0.5$ on the NanoGPT speedrun benchmark \citep{modded_nanogpt_2024}.
+The optimal hyperparameters are shown in the legend of \figureref{fig:gpt_loss}, with \texttt{sign\_sgd\_coeff} values of 0.0003 for F-Muon and 0.07 for S-Muon.
+After 1750 training steps, F-Muon achieves a cross-entropy loss of 3.281, S-Muon achieves 3.287, and Muon achieves 3.279, falling below the target threshold of 3.28.
 As illustrated in \figureref{fig:gpt_loss}, these differences are negligible.
@@
 \subsection{GPT-2 Medium Speedrun}
-We scale from NanoGPT to GPT-2 Medium. After testing for 5960 steps we get 2.9215 loss for F-Muon and 2.9235 for S-Muon, while for Muon it is 2.9198 (the speedrun threshold is 2.92).
+We scale from NanoGPT to GPT-2 Medium.
+After testing for 5960 steps with F-Muon we get 2.9215 loss, while for Muon it is 2.9198.
+In ?
+steps, F-Muon reaches <2.92 loss, thus needing only ?\% more iterations than Muon.
 
@@
 Common important drawback of the analyses is the consideration of the convergence by the norm of the gradient.
-As we showed in our experiments on CIFAR \cref{sec:cifar_underhood}, the gradient norm may decrease only by a factor of ten even when the accuracy reaches 100\%.
+As we showed in our experiments on CIFAR, the gradient norm may decrease only by a factor of ten, when the accuracy reaches 100\%.
 
@@
 \section{Under the hood of CNN on CIFAR-10}
-\label{sec:cifar_underhood}
 

### 2025-11-30 article_final/main.tex
Added charts for Medium GPT

--- 2025-11-30 article_final/main.tex
@@
 Their most prominent members are \emph{F-Muon} and \emph{S-Muon}.
-We complement our theoretical analysis with an extensive empirical study of these algorithms across a wide range of tasks and settings, demonstrating that F-Muon and S-Muon consistently match Muon's performance, while outperforming vanilla Muon on synthetic linear least squares problems.
+We complement our theoretical analysis with an extensive empirical study of these algorithms across a wide range of tasks and settings, demonstrating that F-Muon and S-Muon consistently match Muon's performance, while outperforming vanilla Muon on a synthetic linear least squares problem.
 \end{abstract}
@@
 On CIFAR-10 airbench, properly-tuned F-Muon and S-Muon achieve $94.02 \pm 0.13\%$ and $94.03 \pm 0.13\%$ accuracy, matching Muon's $94.01 \pm 0.13\%$ performance.
-On NanoGPT pre-training, F-Muon achieves 3.281 cross-entropy loss, while fully-tuned Muon achieves 3.279.
+On pre-training of NanoGPT F-Muon achieves 3.281 cross-entropy loss, while fully-tuned Muon achieves 3.279; of GPT-2 Medium - F-Muon 2.9215 and Muon 2.9198.
 Finally, S-Muon matches Muon on fine-tuning NanoGPT on Tiny Stories, while F-Muon is far more resistant to learning rate choice than Muon.
@@
 Unlike Schatten norms, which satisfy $\norm{\mM^t}_{S_p}^\dagger = \norm{\mM^t}_{S_q}$ (for $p^{-1} + q^{-1} = 1$) and are thus closed under dualization, Ky Fan norms are generally not.
-Only two exceptional cases exist: the dual of the Ky Fan $1$-norm $\norm{\cdot}_{\text{KF-}1}$ (the spectral norm) is the Ky Fan $\min(m,n)$-norm $\norm{\cdot}_{\text{KF-}\min(m,n)}$ (the nuclear norm), and vice versa.
+Only two exceptional cases exist: the dual to the Ky Fan $1$-norm $\norm{\cdot}_{\text{KF-}1}$ (the spectral norm) is the Ky Fan $\min(m,n)$-norm $\norm{\cdot}_{\text{KF-}\min(m,n)}$ (the nuclear norm), and vice versa.
 Thus working with duals of Ky Fan norms can open up new possibilities:
@@
 \begin{equation}\label{eq:ckfk_norm}
-	\norm{\cdot}_{\mathrm{C-KF-k}} = \alpha \normkfk{\cdot} + \frac{1-\alpha}{\eta}\normc{\cdot}^\dagger\,.sec:lmo_for_nn
+	\norm{\cdot}_{\mathrm{C-KF-k}} = \alpha \normkfk{\cdot} + \frac{1-\alpha}{\eta}\normc{\cdot}^\dagger\,.
 \end{equation}
@@
 
-\subsection{Randomized Linear Least Squares}
+\subsection{Linear Least Squares}
 \label{subsec:lls_exps}
@@
   {fig:cifar_ball}
-  {\caption{Visualization of the LMO balls for Muon and F-Muon.}}
+  {\caption{Visualization of the LMO balls for Muon and F-Muon for CNN training.}}
   {\includegraphics[width=0.6\linewidth]{figs/norms/fstardual_cifar.pdf}}
@@
 We evaluate F-Muon and S-Muon with $\alpha=0.5$ against Muon, SignSGD, and NSGD on the NanoGPT speedrun benchmark \citep{modded_nanogpt_2024}.
-The optimal hyperparameters are shown in the legend of \figureref{fig:gpt_loss}, with \texttt{sign\_sgd\_coeff} values of 0.0003/0.07 for S-Muon.
+The optimal hyperparameters are shown in the legend of \figureref{fig:nano_gpt_loss}, with \texttt{sign\_sgd\_coeff} values of 3e-4/0.07 for S-Muon.
 After 1750 training steps, F-Muon achieves a cross-entropy loss of 3.281, S-Muon achieves 3.287, and Muon achieves 3.279, falling below the target threshold of 3.280.
-As illustrated in \figureref{fig:gpt_loss}, these differences are negligible.
+As illustrated in \figureref{fig:nano_gpt_loss}, these differences are negligible.
 Notably, F-Muon represents a convex combination of Muon and NSGD, despite NSGD performing poorly in isolation.
@@
 \floatconts
-  {fig:gpt_loss}
+  {fig:nano_gpt_loss}
   {\caption{The validation loss for NanoGPT}}
-  {\includegraphics[width=0.7\linewidth]{figs/nanogpt/val_loss_vs_step.pdf}}
-\end{figure}
-
-As observed in the CIFAR airbench experiments, setting $\alpha=-0.1$ for F-Muon (corresponding to the no-ball configuration) yields a loss of 3.2818, representing only a marginal difference from the baseline.
-
-Notably, F-Muon with the NSGD component not scaled by Muon's rule yields a loss of 3.288 (an increase of 0.007), while S-Muon scaled with Muon's rule achieves a loss of 3.292 (an increase of 0.005).
+  {\includegraphics[width=0.7\linewidth]{figs/nanogpt/nano/val_loss_vs_step.pdf}}
+\end{figure}
+
+As observed in the CIFAR airbench experiments, setting $\alpha=1.1$ for F-Muon (corresponding to the no-ball configuration) yields a loss of 3.2818, representing only a marginal difference from the baseline.
+
+Notably, F-Muon with the NSGD component not scaled by Muon's rule yields a loss of 3.288 (an increase of 0.007), while S-Muon scaled with this rule achieves a loss of 3.292 (an increase of 0.005).
 
@@
 We scale from NanoGPT to GPT-2 Medium. After testing for 5960 steps we get 2.9215 loss for F-Muon and 2.9235 for S-Muon, while for Muon it is 2.9198 (the speedrun threshold is 2.92).
+
+\begin{figure}[htbp]
+  \floatconts
+    {fig:med_gpt_loss}
+    {\caption{The validation loss for GPT-2 Medium}}
+    {\includegraphics[width=0.7\linewidth]{figs/nanogpt/medium/val_loss_vs_step.pdf}}
+  \end{figure}
 
@@
 These parallels extend beyond the update formulas themselves to empirical performance characteristics.
-SignSGD closely resembles Adam, as noted in \citep{bernstein2024oldoptimizernewnorm}, and both Adam and Muon perform well when training large models \citep{zhao2025deconstructing, liu2025muon}.
+SignSGD closely resembles Adam, as noted in \citep{bernstein2024oldoptimizernewnorm}, and both Adam and Muon perform well during training large models \citep{zhao2025deconstructing, liu2025muon}.
 NSGD maintains identical updates in both vector and matrix cases.
@@
 A common limitation of these analyses is their focus on convergence measured by the gradient norm.
-As we showed in our CIFAR experiments \cref{sec:cifar_underhood}, the gradient norm may decrease by only a factor of ten when the accuracy reaches 100\%.
+As we showed in our CIFAR experiments \sectionref{sec:cifar_underhood}, the gradient norm may decrease by only a factor of ten when the accuracy reaches 100\%.
 
@@
 We generalized several successful algorithms, including Muon, Dion, and $\nu$SAM, to LMO-based algorithms using the family of norms dual to Ky Fan $k$-norms, yielding the Fanion family with low-rank updates.
-We further proposed the technique of regularizing these updates by combining them with NSGD or SignSGD, creating the F-Fanion and S-Fanion families through the $\norm{\cdot}_{\mathrm{F-KF-k}}$ and $\norm{\cdot}_{\mathrm{C^\dagger-KF-k}}$ norms.
+We further proposed the technique of regularizing these updates by combining them with NSGD or SignSGD, creating the F-Fanion and S-Fanion families through the $\norm{\cdot}_{\mathrm{F-KF-k}}^\dagger$ and $\norm{\cdot}_{\mathrm{C^\dagger-KF-k}}^\dagger$ norms.
 Our experiments demonstrate that F-Muon and S-Muon achieve competitive performance with Muon on CIFAR-10 airbench and NanoGPT tasks, confirming that the underlying norm constraint can be significantly modified without sacrificing effectiveness.
@@
 
-\section{Author Contributions}
+\section*{Author Contributions}
 
@@
 DM presented the problem at the MIPT optimization course.
-IK suggested using composite norms (though not the ones that induce F-Fanions and S-Fanions) and helped to draft and rewise the manuscript.
+IK suggested using composite norms (though not the ones that induce Fanions, F-Fanions or S-Fanions) and helped to draft and rewise the manuscript.
 NK suggested Lanczos algorithm as the most precise means to compute Fanions' updates, conducted experiments to prove this, and wrote \sectionref{sec:matrix-side}.
 AV conducted the finetuning of NanoGPT on Tiny Stories and wrote \sectionref{subsec:finetune}.
-All other work was done by AK: constructing Fanions, F-Fanions, S-Fanions, experimenting on the Linear Least Squares, CIFAR-airbench and NanoGPT pretrain, and writing the manuscript.
+All other work was done by AK: constructing Fanions, F-Fanions, S-Fanions, experimenting on the Linear Least Squares and CIFAR-airbench, pretraining NanoGPT and GPT-2 Medium, and writing the manuscript.
+
+\section*{Acknowledgements}
+We thank Dmitry Kovalev and Egor Shulgin for helpful discussions.
 
@@
 
-\section{Proof of Lemma on Dual of Convex Combinations}\label{sec:proof_dual_conv_comb}
+\section{Proof of Lemma on the Dual to Convex Combinations}\label{sec:proof_dual_conv_comb}
 
@@
     \subfigure[The spectral norm of the gradient]{\label{fig:lls_op_grad_norm}%
-      \includegraphics[width=0.6\linewidth]{figs/simple_lls/gradient_spectral_norm_vs_iteration_long_run.pdf}}%
+      \includegraphics[width=0.5\linewidth]{figs/simple_lls/gradient_spectral_norm_vs_iteration_long_run.pdf}}%
     \vspace{1em}
     \subfigure[The nuclear norm of the gradient]{\label{fig:lls_nuclear_grad_norm}%
-      \includegraphics[width=0.6\linewidth]{figs/simple_lls/gradient_nuclear_norm_vs_iteration_long_run.pdf}}%
+      \includegraphics[width=0.5\linewidth]{figs/simple_lls/gradient_nuclear_norm_vs_iteration_long_run.pdf}}%
     \vspace{1em}
     \subfigure[The loss over time]{\label{fig:lls_loss_time}%
-      \includegraphics[width=0.6\linewidth]{figs/simple_lls/loss_vs_time_long_run.pdf}}
+      \includegraphics[width=0.5\linewidth]{figs/simple_lls/loss_vs_time_long_run.pdf}}
   }
@@
 
-Most bounds for Muon and other LMO-algorithms are given for the norm of the gradient \citep{kovalev2025understanding, pethick2025training, riabinin2025gluon,kovalev2025noneuclideansgdstructuredoptimization}.
+Most bounds for Muon and other LMO-algorithms are given for the norm of the gradient \citep{li2025noteconvergencemuon, kovalev2025understanding, pethick2025training, riabinin2025gluon,kovalev2025noneuclideansgdstructuredoptimization}.
 Accordingly, it is natural to measure those norms on a real deep learning problem.
@@
 
-We compare Muon, F-Muon, S-Muon, NSGD, SignSGD, Neon, and F-Neon, Fanion-5, and F-Fanion-5 on CIFAR-airbench.
+We compare Muon, F-Muon, S-Muon, NSGD, SignSGD, Neon, F-Neon, Fanion-5, and F-Fanion-5 on CIFAR-airbench.
 NSGD and SignSGD are not heavily tuned.
-The hyperparameters of Neon, F-Neon, Fanion-5, and F-Fanion-5 are taken from Muon and F-Muon.
-The parameters are in the table \tableref{tbl:cifar_lrs}.
-val\_accuracy is for the the weighted variant.
+The hyperparameters of Neon, F-Neon, Fanion-5, and F-Fanion-5 are taken from Muon and F-Muon (see \tableref{tbl:cifar_lrs}).
+In the table, val\_accuracy is for the the weighted variant.
 However, to preserve the algorithms, we do not normalize the weights of the network each iteration.

### 2025-11-30 article_final/main.tex
Some small changes in figure captions

--- 2025-11-30 article_final/main.tex
@@
 	\item CIFAR-10 airbench \citep{cifar2023airbench}
-	\item Pre-training NanoGPT and GPT-2 Medium on FineWeb dataset \citep{modded_nanogpt_2024}
+	\item Pre-training NanoGPT on FineWeb dataset \citep{modded_nanogpt_2024}
+  \item Pre-training GPT-2 medium 
 	\item Fine-tuning NanoGPT on Tiny Stories dataset \citep{eldan2023tinystoriessmalllanguagemodels}
@@
 Training a neural network is essentially an optimization of a function of several weight matrices and a few vectors.
-Let us start by considering the problem of minimizing a differentiable function of a \emph{single} matrix (the connection to the general case is presented in \sectionref{sec:lmo_for_nn}):
+Let us start by considering the problem of minimizing a differentiable function of a \emph{single} matrix:
 \begin{equation}\label{eq:matrix_problem}
@@
 \]
-The Neon update $\mD = \vu_1 \vv_1^\top$ achieves $\sigma_1$, while the scaled Muon update $\mD = \frac{1}{k}\mU\mV^\top$ achieves $\frac{1}{k}\sum_{i=1}^{\min(m,n)}\sigma_i$.
+The Neon update $\mD = \vu_1 \vv_1^\top$ achieves $\sigma_1$, while the scaled Muon update $\mD = \frac{1}{k}\sum_{i=1}^{\min(m,n)}\vu_i \vv_i^\top$ achieves $\frac{1}{k}\sum_{i=1}^{\min(m,n)}\sigma_i$.
 Thus, the update is either
@@
 
-Next, we tune F-Muon specifically with $\alpha = 0.5$, obtaining optimal hyperparameters \texttt{lr=0.4(1 - step/\allowbreak{}total\_steps)}, \texttt{momentum=0.6}, \texttt{nesterov = True}.
+Next, we tune F-Muon specifically with $\alpha = 0.5$, obtaining optimal hyperparameters \texttt{lr=0.4(1 - step/total\_steps), momentum=0.6, nesterov = True}.
 Tuned F-Muon achieves $94.02 \pm 0.13\%$ validation accuracy after 8 epochs (averaged over 200 runs), matching Muon's accuracy and variance.
@@
 
-For S-Muon with $\alpha=0.5$, we tune all hyperparameters to find the optimal configuration \texttt{lr=0.42(1 - step/\allowbreak{}total\_steps)}, \texttt{momentum=0.63}, \texttt{nesterov = True}, \texttt{sign\_lr\_coeff = 0.003}, achieving $94.03 \pm 0.13\%$ validation accuracy, which slightly exceeds vanilla Muon's performance.
+For S-Muon with $\alpha=0.5$, we tune all hyperparameters to find the optimal configuration \texttt{lr=0.42(1 - step/total\_steps), momentum=0.63, nesterov = True, sign\_lr\_coeff = 0.003}, achieving $\mathbf{94.03} \pm 0.13\%$ validation accuracy, which slightly exceeds vanilla Muon's performance.
 
@@
 
-We fine-tuned the NanoGPT framework by Karpathy \citep{Karpathy2022} using the standard \emph{NanoGPT-medium} configuration, which corresponds to GPT-2 Medium.
+We fine-tuned the NanoGPT framework by Karpathy \citep{Karpathy2022} using the standard \textit{NanoGPT-medium} configuration, which corresponds to GPT-2 Medium.
 This model consists of 24 transformer layers, a hidden dimension of 1024, 16 attention heads, and approximately 345 million parameters.
-We selected the \emph{Tiny Stories} corpus \citep{eldan2023tinystoriessmalllanguagemodels} as the training dataset for its high entropy and structural diversity, which amplify the differences in optimization dynamics.
+We selected the \textit{Tiny Stories} corpus \citep{eldan2023tinystoriessmalllanguagemodels} as the training dataset for its high entropy and structural diversity, which amplify the differences in optimization dynamics.
 All training runs were initialized with the same random seed, weight initialization, and learning rate schedule to ensure that performance differences arose solely from the choice of optimizer.
@@
 A common limitation of these analyses is their focus on convergence measured by the gradient norm.
-As we showed in our CIFAR experiments \cref{sec:cifar_underhood}, the gradient norm may decrease by only a factor of ten when the accuracy reaches 100\%.
+As we showed in our CIFAR experiments, the gradient norm may decrease by only a factor of ten when the accuracy reaches 100\%.
 
@@
 \appendix
-\section{LMO for Neural Networks}\label{sec:lmo_for_nn}
+\section{LMO for Neural Networks}
 
@@
 The LMO-based update rule is then applied to each matrix $\mX_i$ using its corresponding gradient component $\mG_i^t$.
+
+The update rule used in Muon optimizer is uSCG \citep{pethick2025training}. In the most general case, which involves momentum, it can be written as
+\begin{equation}\label{eq:usgd_update}
+  \begin{aligned}
+    \mM^{t}     & = \alpha_{t} g(\mX^t, \xi^t) + (1 - \alpha_t)\mM^{t - 1}\,, \\
+    \mX^{t + 1} & = \mX^t + \gamma_t\mathrm{LMO}(\mM^{t})\,.
+  \end{aligned}
+\end{equation}
+
+\section{Norms $\normfstar{\cdot}^\dagger$ and $\normftwo{\cdot}^\dagger$}
+
+Based on the lemma, we immediately find $\normfstar{\cdot}^\dagger$, which is related to F-Muon update.
+Indeed, after setting $\beta=1-\alpha$ and remembering that for smooth and bounded cases we can use $\min$ instead of $\inf$, we get
+\begin{equation}\label{eq:normfstardual}
+  \normfstar{\mY}^\dagger = \min_{\mZ} \min_t\{t, s.t. \norms{\mZ}\leq \alpha t, \normf{\mY-\mZ}\leq (1-\alpha) t\}
+\end{equation}
+
+If $\alpha = 1$, then $\mZ = \mY$, and we get $\normfstar{\mY}^\dagger = \norms{\mY}$. If $\alpha = 0$, then $\mZ = 0$, and we get $\normfstar{\mY}^\dagger = \normf{\mY}$.
+
+Similarly, we find $\normftwo{\cdot}^\dagger$, which is related to F-Neon update:
+\begin{equation}\label{eq:normftwodual}
+  \normftwo{\mY}^\dagger = \min_{\mZ} \min_t\{t, s.t. \normn{\mZ}\leq \alpha t, \normf{\mY-\mZ}\leq (1-\alpha) t\}
+\end{equation}
+
+If $\alpha = 1$, then $\mZ = \mY$, and we get $\normftwo{\mY}^\dagger = \normn{\mY}$. If $\alpha = 0$, then $\mZ = 0$, and we get $\normftwo{\mY}^\dagger = \normf{\mY}$.
+
+Unfortunately, $\normfstar{\cdot}^\dagger$ and $\normftwo{\cdot}^\dagger$ do not have a closed-form expression.
 
@@
 
-\section{Norms $\normfstar{\cdot}^\dagger$ and $\normftwo{\cdot}^\dagger$}
-
-Based on the \lemmaref{lemma:dual_to_conv_comb}, we immediately find $\normfstar{\cdot}^\dagger$, which is related to F-Muon update.
-Indeed, after setting $\beta=1-\alpha$ and remembering that for smooth and bounded cases we can use $\min$ instead of $\inf$, we get
-\begin{equation}\label{eq:normfstardual}
-  \normfstar{\mY}^\dagger = \min_{\mZ} \min_t\{t, s.t. \norms{\mZ}\leq \alpha t, \normf{\mY-\mZ}\leq (1-\alpha) t\}
-\end{equation}
-
-If $\alpha = 1$, then $\mZ = \mY$, and we get $\normfstar{\mY}^\dagger = \norms{\mY}$. If $\alpha = 0$, then $\mZ = 0$, and we get $\normfstar{\mY}^\dagger = \normf{\mY}$.
-
-Similarly, we find $\normftwo{\cdot}^\dagger$, which is related to F-Neon update:
-\begin{equation}\label{eq:normftwodual}
-  \normftwo{\mY}^\dagger = \min_{\mZ} \min_t\{t, s.t. \normn{\mZ}\leq \alpha t, \normf{\mY-\mZ}\leq (1-\alpha) t\}
-\end{equation}
-
-If $\alpha = 1$, then $\mZ = \mY$, and we get $\normftwo{\mY}^\dagger = \normn{\mY}$. If $\alpha = 0$, then $\mZ = 0$, and we get $\normftwo{\mY}^\dagger = \normf{\mY}$.
-
-Unfortunately, $\normfstar{\cdot}^\dagger$ and $\normftwo{\cdot}^\dagger$ do not have a closed-form expression.
-
 \section{Visualization of different matrix norms}

### 2025-11-30 article_final/main.tex
small fixes in appendix

@@
   \item \texttt{lr} $\in [0.005, 0.020]$ with step $0.001$ for Muon, F-Muon, and S-Muon.
-  \item \texttt{lr} $\in [0.00005, 0.00020]$ with step $0.00001$ for SignSGD.
+  \item \texttt{lr} $\in [5 \times 10^{-5}, 20 \times 10^{-5}]$ with step $10^{-5}$ for SignSGD.
   \item \texttt{lr} $\in [0.01, 0.10]$ with step $0.001$ for NSGD.
@@
 
-We fine-tuned the NanoGPT framework by Karpathy \citep{Karpathy2022} using the standard \textit{NanoGPT-medium} configuration, which corresponds to GPT-2 Medium.
+We fine-tuned the NanoGPT framework by Karpathy \citep{Karpathy2022} using the standard \emph{NanoGPT-medium} configuration, which corresponds to GPT-2 Medium.
 This model consists of 24 transformer layers, a hidden dimension of 1024, 16 attention heads, and approximately 345 million parameters.
-We selected the \textit{Tiny Stories} corpus \citep{eldan2023tinystoriessmalllanguagemodels} as the training dataset for its high entropy and structural diversity, which amplify the differences in optimization dynamics.
+We selected the \emph{Tiny Stories} corpus \citep{eldan2023tinystoriessmalllanguagemodels} as the training dataset for its high entropy and structural diversity, which amplify the differences in optimization dynamics.
 All training runs were initialized with the same random seed, weight initialization, and learning rate schedule to ensure that performance differences arose solely from the choice of optimizer.

### 2025-11-30 article_final/main.tex
Expanded appendix

--- 2025-11-30 article_final/main.tex
@@
 The extremal members of this family are \emph{F-Neon} (with $k = 1$) and \emph{F-Muon} (with $k = \min\{m, n\}$).
-Additional information and visualizations for the $\normfstar{\cdot} = \norm{\cdot}_{\mathrm{F-KF-1}}$ norm appear in the appendix (see \equationref{eq:normfstardual}, \figureref{fig:fstardual}).
+Additional information and visualizations for the F-Muon generating $\normfstar{\cdot}^\dagger = \norm{\cdot}_{\mathrm{F-KF-1}}^\dagger$ and the F-Neon generating $\normftwo{\cdot}^\dagger = \norm{\cdot}_{\mathrm{F-KF-\min\{m, n\}}}^\dagger$ norms appear in the appendix (see \equationref{eq:normfstardual}, \equationref{eq:normftwodual}, \figureref{fig:fduals}).
 
@@
 
-Notably, even the pathological case $\alpha > 1$—which corresponds to an LMO constraint region that is not a norm ball—achieves accuracy nearly identical to vanilla Muon.
+Notably, even the pathological case $\alpha > 1$, which corresponds to an LMO constraint region that is not a norm ball, achieves accuracy nearly identical to vanilla Muon.
 
 These observations raise fundamental questions about the sensitivity of LMO-based algorithms to the shape of the constraint region.
+
+For gradient norm profiling and results for training with Fanions and F-Fanions, please see \sectionref{sec:cifar_underhood}.
 
@@
 
-\section{Norms $\normfstar{\cdot}^\dagger$ and $\normftwo{\cdot}^\dagger$}
+\section[Norms for F-Muon and F-Neon]%
+        {Norms $\normfstar{\cdot}^\dagger$ and $\normftwo{\cdot}^\dagger$}
+\label{sec:f2_fstar_norms}
 
@@
 
-Unfortunately, $\normfstar{\cdot}^\dagger$ and $\normftwo{\cdot}^\dagger$ do not have a closed-form expression.
+Unfortunately, $\normfstar{\cdot}^\dagger$ and $\normftwo{\cdot}^\dagger$ do not have closed-form expressions and cannot be found as easy as their duals.
 
@@
 
-\section{More data for Linear Least Squares}
+\section{More details for Linear Least Squares}
 \label{sec:lls_plots_section}
+Tuned \texttt{lr} and \texttt{momentum} coefficients are in the \tableref{tbl:lls_lrs_app}.
+Spectral, nuclear norm of the gradients over iterations and loss over time are in the \figureref{fig:app_lls}.
+Very low speed of Fanions and F-Fanions for large $k$ is probably caused by ineffective implementation of TRLan by \cite{cupy_svds_ref}, since $\mathcal{O}(mn^2 + n^2k + nk^2)$ complexity is actually $\mathcal{O}(mn^2)$.
 
@@
   {tbl:lls_lrs_app}
-  {\caption{Tuning lrs and momentum coefficients}}
+  {\caption{Tuning \texttt{lr} and \texttt{momentum} coefficients}}
   {%
@@
   {fig:app_lls}
-  {\caption{More images for Linear least squares problem for a 500x500 matrix}}
+  {\caption{More plots for linear least squares problem for a 500x500 matrix}}
   {%
@@
 The hyperparameters of Neon, F-Neon, Fanion-5, and F-Fanion-5 are taken from Muon and F-Muon (see \tableref{tbl:cifar_lrs}).
-In the table, val\_accuracy is for the the weighted variant.
-However, to preserve the algorithms, we do not normalize the weights of the network each iteration.
-
-We log train and validation accuracy, and nuclear, gradient, and spectral norms of the gradients of conv1.weight and conv2.weight for all the layers.
-The gradient norms decrease only by the order of ten maximum during training, while train accuracy reaches full 100\% for Muon, S-Muon, and F-Muon.
+In the table, val\_accuracy after 8 epochs is for the the airbench variant with weight normalization at each step.
+However, in order to be faithful to the conditions of the convergence theorems \citep{kovalev2025understanding,riabinin2025gluon}, we do not normalize the weights of the network when logging gradient norms.
+
+We log train and validation accuracies (\figureref{fig:cifar_accs}), and nuclear, gradient, and spectral norms of the gradients of conv1.weight and conv2.weight for all the layers (see \figureref{fig:norms_conv1_vs_conv2} for norms in the layer 2, and \figureref{fig:total_norms_cifar} for total norms of the weights).
+During training, the gradient norms decrease only by the order of ten maximum, while train accuracy reaches full 100\% for Muon, S-Muon, and F-Muon.
+This means that the bounds of the $\max\{\frac{A}{\epsilon}, \frac{B}{\epsilon^2}, \frac{C}{\epsilon^3}, \frac{D}{\epsilon^4}\}$ type (like Corollary 2 from \citet{kovalev2025understanding}) probably should not be reduced to $\frac{D}{\epsilon^4}$ when selecting the optimal algorithm for a practical problem.
 
@@
       \includegraphics[width=0.7\linewidth]{figs/cifar_logger/val_acc.pdf}}
-  }
-\end{figure}
-
-\begin{figure}[htbp]
-\floatconts
-  {fig:total_norms_cifar}
-  {\caption{Matrix norms of the whole gradient of CIFAR CNN}}
-  {%
-    \subfigure[Total Frobenius norm]{\label{fig:total_fro_norm_cifar}%
-      \includegraphics[width=0.7\linewidth]{figs/cifar_logger/total_frobenius.pdf}}%
-    \vspace{1em}
-    \subfigure[Total spectral norm]{\label{fig:total_spectral_norm_cifar}%
-      \includegraphics[width=0.7\linewidth]{figs/cifar_logger/total_spectral.pdf}}%
-    \vspace{1em}
-    \subfigure[Total nuclear norm]{\label{fig:total_nuclear_norm_cifar}%
-      \includegraphics[width=0.7\linewidth]{figs/cifar_logger/total_nuclear.pdf}}
   }
@@
 
+\begin{figure}[htbp]
+  \floatconts
+    {fig:total_norms_cifar}
+    {\caption{Matrix norms of the whole gradient of CIFAR CNN}}
+    {%
+      \subfigure[Total Frobenius norm]{\label{fig:total_fro_norm_cifar}%
+        \includegraphics[width=0.7\linewidth]{figs/cifar_logger/total_frobenius.pdf}}%
+      \vspace{1em}
+      \subfigure[Total spectral norm]{\label{fig:total_spectral_norm_cifar}%
+        \includegraphics[width=0.7\linewidth]{figs/cifar_logger/total_spectral.pdf}}%
+      \vspace{1em}
+      \subfigure[Total nuclear norm]{\label{fig:total_nuclear_norm_cifar}%
+        \includegraphics[width=0.7\linewidth]{figs/cifar_logger/total_nuclear.pdf}}
+    }
+  \end{figure}
+
 \section{Technical details of the experiments}
 
-\paragraph{CIFAR airbench} We used NVIDIA RTX A4000.
-
-\paragraph{NanoGPT pretrain} We used the standard setting of 8 $\times$ H100 as documented in \citep{modded_nanogpt_2024}.
+\paragraph{Linear Least Squares and CIFAR airbench} Experiments were conducted on a single NVIDIA RTX A4000. The code was run with Python 3.10.13 and Pytorch~2.6.0+cu126.
+
+\paragraph{NanoGPT and GPT-2 Medium Pretrain} We used the standard setting of 8 $\times$ NVIDIA H100 as documented in \citep{modded_nanogpt_2024}.
+The code was run with Python 3.10.18, Pytorch~2.10.0.dev20251124+cu126 on a computer with Xeon® Platinum 8462Y+ processor and 130 GB of RAM.
 

### 2025-11-30 article_final/main.tex
Edited complexity bound

@@
 
-The per-cycle complexity is $\mathcal{O}(mn^2 + n^2k + nk^2)$, where $m \geq n$ are the dimensions of the target matrix and $k$ is the size of the retained subspace (typically $k \ll n$).
+The per-cycle complexity is $\mathcal{O}(mn^2 + n^2d + nd^2)$, where $m \geq n$ are the dimensions of the target matrix and $d$ is the size of the retained subspace (typically $d \ll n$).
 
@@
 Spectral, nuclear norm of the gradients over iterations and loss over time are in the \figureref{fig:app_lls}.
-Very low speed of Fanions and F-Fanions for large $k$ is probably caused by ineffective implementation of TRLan by \cite{cupy_svds_ref}, since $\mathcal{O}(mn^2 + n^2k + nk^2)$ complexity is actually $\mathcal{O}(mn^2)$.
+Very low speed of Fanions and F-Fanions for large $k$ is probably caused by ineffective implementation of TRLan by \cite{cupy_svds_ref}.
 
@@
 
-\paragraph{Linear Least Squares and CIFAR airbench} Experiments were conducted on a single NVIDIA RTX A4000. The code was run with Python 3.10.13 and Pytorch~2.6.0+cu126.
+\paragraph{Linear Least Squares and CIFAR airbench} Experiments were conducted on a single NVIDIA RTX A4000.
+The code was run with Python 3.10.13 and Pytorch~2.6.0+cu126 on a computer with an EPYC 7543 processor.
 

### 2025-12-01 article_final/main.tex
many changes for notational consistency and clarity

@@
 \DeclarePairedDelimiter{\normn}{\|}{\|_{*}}
-\newcommand{\normkfk}[1]{\|#1\|_{\text{KF-}k}}
+\DeclarePairedDelimiter{\normkfk}{\|}{\|_{\text{KF-}k}}
 \DeclarePairedDelimiter{\normfstar}{\|}{\|_\mathrm{F*}}
@@
 	\[
-		\norm{\cdot} := \sum_{i = 1}^n \alpha_i \norm{\cdot}_{(i)}.
+		\norm{\cdot} := \sum_{i = 1}^n \alpha_i \norm{\cdot}_{(i)}\,.
 	\]
@@
 		\cB_{\norm{\cdot}^\dagger}
-		= \sum_{i = 1}^n \alpha_i \cB_{\norm{\cdot}_{(i)}^\dagger}
+		= \sum_{i = 1}^n \alpha_i \cB_{\norm{\cdot}_{(i)}^\dagger}\,,
 	\]
@@
       \arrayrulecolor{black}\hline
-      $500 \times 500$      & 5  & Power Iterations           & 0.062   & 2005    & 200        & 9.2$\cdot 10^{-3}$ & 9.1$\cdot10^{-3}$\\
-      $500 \times 500$      & 5  & RSVD             & 0.017   & 1170    & 38         & 9.8$\cdot 10^{-3}$ & 9.6$\cdot10^{-3}$\\
-      $500 \times 500$      & 5  & TRLan            & 0.018   & 131     & 65         & 9.6$\cdot 10^{-5}$ & 9.4$\cdot10^{-5}$\\
-      \hdashline
-      $500 \times 500$      & 50 & Power Iterations           & 0.44    & 43750   & 437      & 9.9$\cdot 10^{-3}$ & 9.0$\cdot10^{-3}$\\
-      $500 \times 500$      & 50 & RSVD             & 0.61    & 6120    & 50      & 9.9$\cdot 10^{-3}$ & 9.1$\cdot10^{-3}$\\
-      $500 \times 500$      & 50 & TRLan            & 0.16    & 462     & 231      & 3.3$\cdot 10^{-7}$ & 3.0$\cdot10^{-7}$\\
-      \hdashline 
-      $5000 \times 5000$    & 5  & Power Iterations           & 9.6     & 9065    & 906      & 8.6$\cdot 10^{-3}$ & 8.6$\cdot10^{-3}$\\
-      $5000 \times 5000$    & 5  & RSVD             & 2.1     & 5640    & 187      & 9.7$\cdot 10^{-3}$ & 9.7$\cdot10^{-3}$\\
-      $5000 \times 5000$    & 5  & TRLan            & 0.70    & 205     & 102      & 7.7$\cdot 10^{-3}$ & 7.7$\cdot10^{-3}$\\
+      $500 \times 500$      & 5  & Power Iterations           & 0.062   & 2005    & 200        & $9.2 \cdot 10^{-3}$ & $9.1 \cdot 10^{-3}$\\
+      $500 \times 500$      & 5  & RSVD             & 0.017   & 1170    & 38         & $9.8 \cdot 10^{-3}$ & $9.6 \cdot 10^{-3}$\\
+      $500 \times 500$      & 5  & TRLan            & 0.018   & 131     & 65         & $9.6 \cdot 10^{-5}$ & $9.4 \cdot 10^{-5}$\\
+      \hdashline
+      $500 \times 500$      & 50 & Power Iterations           & 0.44    & 43750   & 437      & $9.9 \cdot 10^{-3}$ & $9.0 \cdot 10^{-3}$\\
+      $500 \times 500$      & 50 & RSVD             & 0.61    & 6120    & 50      & $9.9 \cdot 10^{-3}$ & $9.1 \cdot 10^{-3}$\\
+      $500 \times 500$      & 50 & TRLan            & 0.16    & 462     & 231      & $3.3 \cdot 10^{-7}$ & $3.0 \cdot 10^{-7}$\\
+      \hdashline
+      $5000 \times 5000$    & 5  & Power Iterations           & 9.6     & 9065    & 906      & $8.6 \cdot 10^{-3}$ & $8.6 \cdot 10^{-3}$\\
+      $5000 \times 5000$    & 5  & RSVD             & 2.1     & 5640    & 187      & $9.7 \cdot 10^{-3}$ & $9.7 \cdot 10^{-3}$\\
+      $5000 \times 5000$    & 5  & TRLan            & 0.70    & 205     & 102      & $7.7 \cdot 10^{-3}$ & $7.7 \cdot 10^{-3}$\\
       \arrayrulecolor{black}\hline
@@
   {tbl:newton_schulz_reference}
-  {\caption{Newton-Schulz iterations on the same matrices (for reference).}}
+  {\caption{Newton-Schulz iterations performance on random dense matrices (for reference).}}
   {%
@@
       \arrayrulecolor{black}\hline
-      $500 \times 500$     & 0.041    & 27 000  & 27         & 4.8e-3  \\
-      $5000 \times 5000$   & 26.4     & 290 000 & 29         & 6.5e-3  \\
+      $500 \times 500$     & 0.041    & 27 000  & 27         & $4.8 \cdot 10^{-3}$  \\
+      $5000 \times 5000$   & 26.4     & 290 000 & 29         & $6.5 \cdot 10^{-3}$  \\
       \arrayrulecolor{black}\hline
@@
 \begin{equation}\label{eq:lls}
-  F(\mX) = \frac{1}{2} \<(\mX-\mS), \mM (\mX - \mS) \mN> \to \min_{\mX \in \Rmn}\,
+  F(\mX) = \frac{1}{2} \<(\mX-\mS), \mM (\mX - \mS) \mN> \to \min_{\mX \in \Rmn}\,,
 \end{equation}
@@
   \item \texttt{lr} $\in [0.005, 0.020]$ with step $0.001$ for Muon, F-Muon, and S-Muon.
-  \item \texttt{lr} $\in [5\mathrm{e}{-5},\, 2\mathrm{e}{-4}]$ with step $1\mathrm{e}{-5}$.
+  \item \texttt{lr} $\in [5 \cdot 10^{-5},\, 2 \cdot 10^{-4}]$ with step $1 \cdot 10^{-5}$.
   \item \texttt{lr} $\in [0.01, 0.10]$ with step $0.001$ for NSGD.
@@
   {tbl:lls_lrs}
-  {\caption{Tuning learning rates and momentum coefficients}}
+  {\caption{Tuned learning rates and momentum coefficients in the linear least squares experiment.}}
   {%
@@
   {fig:lls}
-  {\caption{Linear least squares problem for a 500x500 matrix}}
+  {\caption{Linear least squares problem for a 500x500 matrix.}}
   {%
@@
 We evaluate the algorithms on the CIFAR-10 airbench \citep{cifar2023airbench}.
-To assess the impact of the mixing parameter $\alpha$, we first run F-Muon for different values of $\alpha$ using hyperparameters tuned for vanilla Muon by Keller Jordan: \texttt{lr=0.24(1 - step/total\_steps), momentum=0.65, nesterov=True} with weight normalization.
+To assess the impact of the mixing parameter $\alpha$, we first run F-Muon for different values of $\alpha$ using hyperparameters tuned for vanilla Muon by Keller Jordan: \texttt{lr=0.24(1 - step/total\_steps)}, \texttt{momentum=0.65}, \texttt{nesterov=True} with weight normalization.
 We perform 10 repetitions for each $\alpha$ value and record the accuracy after 8 epochs of training (\figureref{fig:muon_alphas}).
 
-Next, we tune F-Muon specifically with $\alpha = 0.5$, obtaining optimal hyperparameters \texttt{lr=0.4(1 - step/\allowbreak{}total\_steps)}, \texttt{momentum=0.6}, \texttt{nesterov = True}.
+Next, we tune F-Muon specifically with $\alpha = 0.5$, obtaining optimal hyperparameters \texttt{lr=0.4(1 - step/total\_steps)}, \texttt{momentum=0.6}, \texttt{nesterov = True}.
 Tuned F-Muon achieves $94.02 \pm 0.13\%$ validation accuracy after 8 epochs (averaged over 200 runs), matching Muon's accuracy and variance.
@@
 
-For S-Muon with $\alpha=0.5$, we tune all hyperparameters to find the optimal configuration \texttt{lr=0.42(1 - step/\allowbreak{}total\_steps)}, \texttt{momentum=0.63}, \texttt{nesterov = True}, \texttt{sign\_lr\_coeff = 0.003}, achieving $94.03 \pm 0.13\%$ validation accuracy, which slightly exceeds vanilla Muon's performance.
+For S-Muon with $\alpha=0.5$, we tune all hyperparameters to find the optimal configuration \texttt{lr=0.42(1 - step/\allowbreak{}total\_steps)}, \texttt{momentum=0.63}, \texttt{nesterov = True}, \texttt{sign\_lr\_coeff = 0.003}, achieving $94.03 \pm 0.13\%$ validation accuracy (\figureref{fig:diff_alphas}), which slightly exceeds vanilla Muon's performance.
 
@@
   {fig:diff_alphas}
-  {\caption{Mean validation accuracies for F-Muon with different $\alpha$}}
+  {\caption{Mean validation accuracies for F-Muon with different $\alpha$.}}
   {%
@@
 The Muon-level performance is remarkable given that F-Muon and S-Muon operate with substantially different constraint geometries.
-\figureref{fig:cifar_ball} visualizes F-Muon by plotting the LMO balls of Muon and F-Muon (scaled by actual learning rates) in the 2D space of singular values.
-S-Muon cannot be visualized in this manner because the $l_\infty$ norm is not a function of singular values alone, and visualizing $\norms{\cdot}$ for $\Rmn$ with $m, n > 2$ requires at least 4D space.
-Nevertheless, the LMO ball of S-Muon clearly differs substantially from Muon's, as the effective SignSGD learning rate $\texttt{lr}(1-\alpha)\texttt{sign\_lr\_coeff} = 6.3 \times 10^{-4}$ is comparable to typical SignSGD learning rates in deep learning.
-
-Notably, even the pathological case $\alpha > 1$, which corresponds to an LMO constraint region that is not a norm ball, achieves accuracy nearly identical to vanilla Muon.
-
-These observations raise fundamental questions about the sensitivity of LMO-based algorithms to the shape of the constraint region.
-
-For gradient norm profiling and results for training with Fanions and F-Fanions, please see \sectionref{sec:cifar_underhood}.
+\figureref{fig:cifar_ball} visualizes F-Muon by plotting the LMO balls of Muon and F-Muon (scaled by actual learning rates) in the 2D space of singular values (with all other singular values set to zero).
+S-Muon cannot be visualized in this manner because the Chebyshev norm is not a function of singular values alone, and visualizing $\norms{\cdot}$ for $\Rmn$ with $m, n > 2$ requires at least 4D space.
+Nevertheless, the LMO ball of S-Muon clearly differs substantially from Muon's, as the effective SignSGD learning rate $\texttt{lr} = (1-\alpha)\texttt{sign\_lr\_coeff} = 6.3 \cdot 10^{-4}$ is comparable to typical SignSGD learning rates in deep learning.
 
@@
 
+Notably, even the pathological case $\alpha > 1$, which corresponds to an LMO constraint region that is not a norm ball, achieves accuracy nearly identical to vanilla Muon.
+These observations raise fundamental questions about the sensitivity of LMO-based algorithms to the shape of the constraint region.
+
+Additional results, including gradient norm profiling and performance comparisons with Fanions and F-Fanions, are provided in \sectionref{sec:cifar_underhood}.
+
 \subsection{NanoGPT Speedrun}
@@
 We evaluate F-Muon and S-Muon with $\alpha=0.5$ against Muon, SignSGD, and NSGD on the NanoGPT speedrun benchmark \citep{modded_nanogpt_2024}.
-The optimal hyperparameters are shown in the legend of \figureref{fig:nano_gpt_loss}, with \texttt{sign\_sgd\_coeff} values of 3e-4/0.07 for S-Muon.
+The optimal hyperparameters are shown in the legend of \figureref{fig:nano_gpt_loss}, with \texttt{sign\_sgd\_coeff} values of $3 \cdot 10^{-4}$ and $0.07$ for S-Muon.
 After 1750 training steps, F-Muon achieves a cross-entropy loss of 3.281, S-Muon achieves 3.287, and Muon achieves 3.279, falling below the target threshold of 3.280.
@@
   {fig:nano_gpt_loss}
-  {\caption{The validation loss for NanoGPT}}
+  {\caption{The validation loss for NanoGPT.}}
   {\includegraphics[width=0.7\linewidth]{figs/nanogpt/nano/val_loss_vs_step.pdf}}
@@
 \subsection{GPT-2 Medium Speedrun}
-We scale from NanoGPT to GPT-2 Medium. After testing for 5960 steps we get 2.9215 loss for F-Muon and 2.9235 for S-Muon, while for Muon it is 2.9198 (the speedrun threshold is 2.92).
+
+We scale from NanoGPT to GPT-2 Medium (24 transformer layers, 1024 hidden dimension, 16 attention heads, approximately 345 million parameters), evaluating the same algorithms with $\alpha=0.5$ on the FineWeb dataset.
+After 5960 training steps, Muon achieves a validation loss of 2.9198, successfully reaching the speedrun threshold of 2.92.
+F-Muon achieves 2.9215 and S-Muon achieves 2.9235, narrowly missing the threshold.
+As illustrated in \figureref{fig:med_gpt_loss}, the performance gap remains remarkably small: F-Muon trails Muon by only 0.0017, while S-Muon trails by 0.0037.
+These minimal differences demonstrate that alternative norm constraints maintain competitive performance even at this significantly larger model scale.
 
@@
     {fig:med_gpt_loss}
-    {\caption{The validation loss for GPT-2 Medium}}
+    {\caption{The validation loss for GPT-2 Medium.}}
     {\includegraphics[width=0.7\linewidth]{figs/nanogpt/medium/val_loss_vs_step.pdf}}
@@
 
-We fine-tuned the NanoGPT framework by Karpathy \citep{Karpathy2022} using the standard \emph{NanoGPT-medium} configuration, which corresponds to GPT-2 Medium.
-This model consists of 24 transformer layers, a hidden dimension of 1024, 16 attention heads, and approximately 345 million parameters.
+We fine-tuned the NanoGPT framework by Karpathy \citep{Karpathy2022} using the standard \emph{NanoGPT-medium} configuration, which again corresponds to GPT-2 Medium.
 We selected the \emph{Tiny Stories} corpus \citep{eldan2023tinystoriessmalllanguagemodels} as the training dataset for its high entropy and structural diversity, which amplify the differences in optimization dynamics.
@@
 
-For all one-dimensional layers, we used AdamW with a learning rate of $1\!\times\!10^{-3}$.
+For all one-dimensional layers, we used AdamW with a learning rate of $1 \cdot 10^{-3}$.
 Since momentum exhibited negligible influence on training dynamics, it was held constant across all experiments.
@@
   {\caption{Comparison of validation loss for Muon, F-Muon, and S-Muon across a range of learning rates. Stars denote the best learning rate identified for each optimizer.}}
-  {\includegraphics[width=0.8\linewidth]{figs/nanogpt/optimizers_comparison.pdf}}
+  {\includegraphics[width=0.7\linewidth]{figs/nanogpt/optimizers_comparison.pdf}}
 \end{figure}
@@
   {tbl:mat_vs_vec_lmo}
-  {\caption{LMO optimizers in Schatten $S_p$ norms and in $l_p$ norms. $g$ is the gradient. When it is a matrix, $g = \mU \mSigma \mV^\top$}}
+  {\caption{LMO optimizers in Schatten $S_p$ norms and in $l_p$ norms. $g$ is the gradient. When it is a matrix, $g = \mU \mSigma \mV^\top$.}}
   {%
@@
 
-Based on the \lemmaref{lemma:dual_to_conv_comb}, we immediately find $\normfstar{\cdot}^\dagger$, which is related to F-Muon update.
-Indeed, after setting $\beta=1-\alpha$ and remembering that for smooth and bounded cases we can use $\min$ instead of $\inf$, we get
+Based on \lemmaref{lemma:dual_to_conv_comb}, we immediately find $\normfstar{\cdot}^\dagger$, which is related to the F-Muon update.
+Indeed, after setting $\beta=1-\alpha$ and remembering that for smooth and bounded cases we can use $\min$ instead of $\inf$, we obtain
 \begin{equation}\label{eq:normfstardual}
-  \normfstar{\mY}^\dagger = \min_{\mZ} \min_t\{t, s.t. \norms{\mZ}\leq \alpha t, \normf{\mY-\mZ}\leq (1-\alpha) t\}
+  \normfstar{\mY}^\dagger = \min_{\mZ} \min_t\{t, s.t. \norms{\mZ}\leq \alpha t, \normf{\mY-\mZ}\leq (1-\alpha) t\}\,.
 \end{equation}
@@
 
-Similarly, we find $\normftwo{\cdot}^\dagger$, which is related to F-Neon update:
+Similarly, we find $\normftwo{\cdot}^\dagger$, which is related to the F-Neon update:
 \begin{equation}\label{eq:normftwodual}
-  \normftwo{\mY}^\dagger = \min_{\mZ} \min_t\{t, s.t. \normn{\mZ}\leq \alpha t, \normf{\mY-\mZ}\leq (1-\alpha) t\}
+  \normftwo{\mY}^\dagger = \min_{\mZ} \min_t\{t, s.t. \normn{\mZ}\leq \alpha t, \normf{\mY-\mZ}\leq (1-\alpha) t\}\,.
 \end{equation}
@@
 
-Unfortunately, $\normfstar{\cdot}^\dagger$ and $\normftwo{\cdot}^\dagger$ do not have closed-form expressions and cannot be found as easy as their duals.
+Unfortunately, $\normfstar{\cdot}^\dagger$ and $\normftwo{\cdot}^\dagger$ do not have simple closed-form expressions and cannot be computed as easily as their duals.
 
@@
 
-In \figureref{fig:fduals} we plot these norms. On x-axis and y-axis, there are singular values $\sigma_1$, $\sigma_2$ respectively of a matrix from $\R^{m\times n}$ with $\min\{m,n\}=2$.
+In \figureref{fig:fduals}, we plot these norms. The x-axis and y-axis represent the singular values $\sigma_1$ and $\sigma_2$, respectively, of a matrix from $\R^{m\times n}$ with $\min\{m,n\}=2$.
 
@@
   {fig:fduals}
-  {\caption{Balls in the duals to F* and F2 norms for different $\alpha$}}
+  {\caption{Balls in the duals to F* and F2 norms for different $\alpha$.}}
   {%
@@
 
-1-balls in $l_\infty$, $l_1$ and $l_2$ norms are well-known from textbooks. But what about the Ky Fan $k$-norm? How can it be represented?
-
-To showcase the complex structure of the Ky Fan $k$-norm and its dual, we suggest the illustrations \figureref{fig:kyfan_combined} with the ball in the Ky Fan $2$-norm in \figureref{fig:kyfan} and its dual in \figureref{fig:kyfandual}.
-On x-, y-, and z-axes, there are singular values $\sigma_1$, $\sigma_2$, and $\sigma_3$ respectively of a matrix from $\R^{m\times n}$ with $\min\{m,n\}=3$.
-In this particular case, we do not sort the singular values.
-In the proposed representation, we actually plot balls in the Top-$2$ norm $\max\{\abs{x}+\abs{y}, \abs{x}+\abs{z}, \abs{y}+\abs{z}\}$ and its dual norm $\max\{\max(\abs(x),\abs{y},\abs{z}), \frac{1}{2} (\abs{x}+\abs{y}+\abs{z})\}$.
-The resulting balls are much more complex than balls in $l_\infty$, $l_1$ and $l_2$ norms.
-
-In fact, those balls can be described easier if we use the results from \citep{yu2012arithmetic}.
-The Ky Fan $2$-norm ball is an intersection of three $l_1$ balls in $(x,y)$, $(x, z)$, and $(y,z)$ spaces.
-The 1-ball in the dual Ky Fan $2$-norm is an intersection of 1-ball the in $l_\infty$ norm and $\frac{1}{2}$-ball in the $l_1$ norm.
+While 1-balls in $l_\infty$, $l_1$, and $l_2$ norms are well-known from textbooks, the Ky Fan $k$-norm presents a more intricate structure.
+
+To showcase the complex geometry of the Ky Fan $k$-norm and its dual, we present \figureref{fig:kyfan_combined}, which displays the unit ball in the Ky Fan $2$-norm (\figureref{fig:kyfan}) and its dual (\figureref{fig:kyfandual}).
+The x-, y-, and z-axes represent the singular values $\sigma_1$, $\sigma_2$, and $\sigma_3$, respectively, of a matrix from $\R^{m\times n}$ with $\min\{m,n\}=3$.
+In this representation, we do not sort the singular values.
+We plot unit balls in the Top-$2$ norm $\max\{\abs{x}+\abs{y}, \abs{x}+\abs{z}, \abs{y}+\abs{z}\}$ and its dual norm $\max\{\max(\abs{x},\abs{y},\abs{z}), (\abs{x}+\abs{y}+\abs{z})/2\}$.
+The resulting balls exhibit significantly greater complexity than those in $l_\infty$, $l_1$, and $l_2$ norms.
 
@@
   {fig:kyfan_combined}
-  {\caption{Ky Fan 2-norm and its dual}}
+  {\caption{Ky Fan 2-norm and its dual in the space of 3 singular values.}}
   {%
@@
 
+These balls can be described more simply using the results from \citet{yu2012arithmetic}.
+The Ky Fan $2$-norm ball is an intersection of three $l_1$ balls in $(x,y)$, $(x, z)$, and $(y,z)$ spaces.
+The unit ball in the dual Ky Fan $2$-norm is an intersection of the unit ball in the $l_\infty$ norm and the $1/2$-ball in the $l_1$ norm.
+
 \section{More details for Linear Least Squares}
 \label{sec:lls_plots_section}
-Tuned \texttt{lr} and \texttt{momentum} coefficients are in the \tableref{tbl:lls_lrs_app}.
-Spectral, nuclear norm of the gradients over iterations and loss over time are in the \figureref{fig:app_lls}.
-Very low speed of Fanions and F-Fanions for large $k$ is probably caused by ineffective implementation of TRLan by \cite{cupy_svds_ref}.
+Tuned \texttt{lr} and \texttt{momentum} coefficients are presented in \tableref{tbl:lls_lrs_app}.
+The spectral and nuclear norms of the gradients over iterations, as well as the loss over time, are shown in \figureref{fig:app_lls}.
+The poor performance of Fanions and F-Fanions in terms of speed for large $k$ is likely caused by the inefficient implementation of TRLan.
 
@@
   {tbl:lls_lrs_app}
-  {\caption{Tuning \texttt{lr} and \texttt{momentum} coefficients}}
+  {\caption{Tuned \texttt{lr} and \texttt{momentum} coefficients in the lease squares experiment.}}
   {%
@@
   {fig:app_lls}
-  {\caption{More plots for linear least squares problem for a 500x500 matrix}}
+  {\caption{More plots for linear least squares problem for a 500x500 matrix.}}
   {%
@@
 
-\section{Under the hood of CNN on CIFAR-10}
+\section{Gradient Norm Analysis for CIFAR-10 Experiments}
 \label{sec:cifar_underhood}
 
-Most bounds for Muon and other LMO-algorithms are given for the norm of the gradient \citep{li2025noteconvergencemuon, kovalev2025understanding, pethick2025training, riabinin2025gluon,kovalev2025noneuclideansgdstructuredoptimization}.
-Accordingly, it is natural to measure those norms on a real deep learning problem.
+Theoretical convergence bounds for Muon and other LMO-based algorithms are typically expressed in terms of gradient norms \citep{li2025noteconvergencemuon, kovalev2025understanding, pethick2025training, riabinin2025gluon,kovalev2025noneuclideansgdstructuredoptimization}.
+Accordingly, we measure these norms during training on a real deep learning problem to validate theoretical predictions.
 
@@
   {tbl:cifar_lrs}
-  {\caption{Parameters for CIFAR-airbench. \texttt{sign\_lr\_mult} for S-Muon is 0.003}}
+  {\caption{Parameters for CIFAR-airbench. \texttt{sign\_lr\_mult} for S-Muon is 0.003.}}
   {%
@@
 The hyperparameters of Neon, F-Neon, Fanion-5, and F-Fanion-5 are taken from Muon and F-Muon (see \tableref{tbl:cifar_lrs}).
-In the table, val\_accuracy after 8 epochs is for the the airbench variant with weight normalization at each step.
-However, in order to be faithful to the conditions of the convergence theorems \citep{kovalev2025understanding,riabinin2025gluon}, we do not normalize the weights of the network when logging gradient norms.
-
-We log train and validation accuracies (\figureref{fig:cifar_accs}), and nuclear, gradient, and spectral norms of the gradients of conv1.weight and conv2.weight for all the layers (see \figureref{fig:norms_conv1_vs_conv2} for norms in the layer 2, and \figureref{fig:total_norms_cifar} for total norms of the weights).
-During training, the gradient norms decrease only by the order of ten maximum, while train accuracy reaches full 100\% for Muon, S-Muon, and F-Muon.
-This means that the bounds of the $\max\{\frac{A}{\epsilon}, \frac{B}{\epsilon^2}, \frac{C}{\epsilon^3}, \frac{D}{\epsilon^4}\}$ type (like Corollary 2 from \citet{kovalev2025understanding}) probably should not be reduced to $\frac{D}{\epsilon^4}$ when selecting the optimal algorithm for a practical problem.
+The validation accuracies reported after 8 epochs correspond to the airbench variant with weight normalization applied at each step.
+However, to remain faithful to the conditions assumed in convergence theorems \citep{kovalev2025understanding,riabinin2025gluon}, we do not normalize network weights when logging gradient norms.
+
+We log train and validation accuracies (\figureref{fig:cifar_accs}), as well as the Frobenius, spectral, and nuclear norms of the gradients for \texttt{conv1.weight} and \texttt{conv2.weight} across all layers (see \figureref{fig:norms_conv1_vs_conv2} for norms in layer 2, and \figureref{fig:total_norms_cifar} for total gradient norms).
+During training, gradient norms decrease by at most one order of magnitude, while training accuracy reaches 100\% for Muon, S-Muon, and F-Muon.
+This observation suggests that convergence bounds of the form $\max\{\frac{A}{\epsilon}, \frac{B}{\epsilon^2}, \frac{C}{\epsilon^3}, \frac{D}{\epsilon^4}\}$ (such as Corollary 2 from \citet{kovalev2025understanding}) should not be simplified to $\frac{D}{\epsilon^4}$ when selecting an optimal algorithm for practical applications.
 
@@
   {fig:cifar_accs}
-  {\caption{Different optimizers on CIFAR airbench without weight normalization}}
+  {\caption{Different optimizers on CIFAR airbench without weight normalization.}}
   {%
@@
   {fig:norms_conv1_vs_conv2}
-  {\caption{Matrix norms of layer2.conv1 (left) and layer2.conv2 (right) gradients of CIFAR CNN}}
+  {\caption{Matrix norms of layer2.conv1 (left) and layer2.conv2 (right) gradients of CIFAR CNN.}}
   {%
@@
     {fig:total_norms_cifar}
-    {\caption{Matrix norms of the whole gradient of CIFAR CNN}}
+    {\caption{Matrix norms of the whole gradient of CIFAR CNN.}}
     {%

### 2025-12-01 article_final/main.tex
rearranged the figures a little bit

@@
     \subfigure[The spectral norm of the gradient]{\label{fig:lls_op_grad_norm}%
-      \includegraphics[width=0.5\linewidth]{figs/simple_lls/gradient_spectral_norm_vs_iteration_long_run.pdf}}%
-    \vspace{1em}
+      \includegraphics[width=0.49\linewidth]{figs/simple_lls/gradient_spectral_norm_vs_iteration_long_run.pdf}}%
+    \hfill
     \subfigure[The nuclear norm of the gradient]{\label{fig:lls_nuclear_grad_norm}%
-      \includegraphics[width=0.5\linewidth]{figs/simple_lls/gradient_nuclear_norm_vs_iteration_long_run.pdf}}%
+      \includegraphics[width=0.49\linewidth]{figs/simple_lls/gradient_nuclear_norm_vs_iteration_long_run.pdf}}%
     \vspace{1em}
     \subfigure[The loss over time]{\label{fig:lls_loss_time}%
-      \includegraphics[width=0.5\linewidth]{figs/simple_lls/loss_vs_time_long_run.pdf}}
+      \includegraphics[width=0.49\linewidth]{figs/simple_lls/loss_vs_time_long_run.pdf}}
   }

### 2025-12-01 article_final/main.tex
Removed redundant table

@@
 \label{sec:lls_plots_section}
-Tuned \texttt{lr} and \texttt{momentum} coefficients are presented in \tableref{tbl:lls_lrs_app}.
 The spectral and nuclear norms of the gradients over iterations, as well as the loss over time, are shown in \figureref{fig:app_lls}.
 The poor performance of Fanions and F-Fanions in terms of speed for large $k$ is likely caused by the inefficient implementation of TRLan.
-
-\begin{table}[htbp]
-\floatconts
-  {tbl:lls_lrs_app}
-  {\caption{Tuned \texttt{lr} and \texttt{momentum} coefficients in the lease squares experiment.}}
-  {%
-    \begin{tabular}{|c|c|c|c|}
-      \arrayrulecolor{black}\hline
-      Algorithm & \texttt{lr}  & \texttt{momentum}  & Iterations to 0.001 loss \\
-      \arrayrulecolor{black}\hline
-      Muon & 0.007 & 0.5 & 1060 \\
-      \hdashline
-      NSGD & 0.08 & 0.95 & 1020 \\
-      F-Muon & 0.015 & 0.7 & 910 \\
-      \hdashline
-      SignSGD & 0.016 $\times$ 0.01 & 0.95 & 2650 \\
-      S-Muon & 0.011 & 0.9 & 890 \\
-      \arrayrulecolor{black}\hline
-    \end{tabular}%
-  }
-\end{table}
 

### 2025-12-01 article_final/main.tex

@@
 
-The per-cycle complexity is $\mathcal{O}(mn^2 + n^2d + nd^2)$, where $m \geq n$ are the dimensions of the target matrix and $d$ is the size of the retained subspace (typically $d \ll n$).
+The per-iteration complexity of TRLan is $\mathcal{O}(mn)$, where $m \geq n$ are the dimensions of the target matrix.
 

### 2025-12-01 article_final/main.tex
Final version done

--- 2025-12-01 article_final/main.tex
@@
  \usepackage{booktabs}
+\AtBeginEnvironment{table}{\setlength{\extrarowheight}{2pt}}
+\AtEndEnvironment{table}{\setlength{\extrarowheight}{1pt}}
  %\usepackage{rotating}% for sideways figures and tables
@@
 
-\emph{In deriving Muon's update step, why should one constrain by the spectral or any other operator norm? How would alternative norms affect performance and computational cost?}
+\emph{Can the employment of a norm other than the operator norm in Muon-like algorithms yield comparable or superior empirical results?}
 
@@
 This approach theoretically explains the backbone of Dion's update \citep{ahn2025dion} and generalizes the memory-motivated application of the nuclear norm to Sharpness-Aware Minimization \citep{pethick2025sam}.
-As was done with Muon, we develop an effective procedure for computing Fanions' updates.
-The Lanczos algorithm, which is described and compared with its competitors in \sectionref{sec:matrix-side}, is the most operation-efficient algorithm, though it currently lacks an effective GPU- and PyTorch-friendly implementation.
+As was done with Muon, we develop an effective procedure for computing Fanions' updates using the Lanczos algorithm (see \sectionref{sec:matrix-side}).
 
@@
 
-After that in \sectionref{sec:experiments} compare the performance of these algorithm families on various model and real-world problems:
+In \sectionref{sec:experiments}, we compare the performance of these algorithm families on various model and real-world problems:
 \begin{itemize}
@@
 On CIFAR-10 airbench, properly-tuned F-Muon and S-Muon achieve $94.02 \pm 0.13\%$ and $94.03 \pm 0.13\%$ accuracy, matching Muon's $94.01 \pm 0.13\%$ performance.
-On pre-training of NanoGPT F-Muon achieves 3.281 cross-entropy loss, while fully-tuned Muon achieves 3.279; of GPT-2 Medium - F-Muon 2.9215 and Muon 2.9198.
+After 1750 iterations on NanoGPT training, F-Muon achieves 3.281 cross-entropy loss, while Muon achieves 3.279.
+Similarly, after 5960 iterations of GPT-2 Medium training we get 2.9215 with F-Muon and 2.9198 with Muon.
 Finally, S-Muon matches Muon on fine-tuning NanoGPT on Tiny Stories, while F-Muon is far more resistant to learning rate choice than Muon.
@@
 \end{equation}
-where $\beta \in [0, 1)$ is the momentum coefficient.
+where $\beta \in [0, 1)$ is the momentum coefficient and $\nabla F(\mX^t)$ is either full or stochastic gradient.
 Throughout our experiments, we employ approximate Nesterov momentum, which matches the implementation in Muon and PyTorch SGD (\texttt{nesterov=True}).
 It uses the current gradient rather than the true lookahead gradient, trading theoretical guarantees for computational efficiency.
+From here on, we refer to stochastic gradient as just gradient if not otherwise specified.
 
@@
 
-Using this formula, it is easy to compute updates for algorithms induced by various norms $\norm{\cdot}$:
+Using this formula, it is easy \citep{bernstein2024oldoptimizernewnorm, pethick2025training} to compute updates for algorithms induced by various norms $\norm{\cdot}$:
 
@@
 where $p$ and $q$ satisfy $p^{-1} + q^{-1} = 1$.
-This formula recovers Neon when $p\rightarrow1$ (provided $\sigma_1 > \sigma_2$, which holds on real data), NSGD when $p=2$, and Muon when $p\rightarrow\infty$.
-
-However, Schatten norms do not fill the rank gap: when $p > 1$, the update has full rank, while when $p=1$, it has rank one.
+This formula recovers Neon when $p\rightarrow1$ (provided that $\sigma_1 > \sigma_2$, which holds on real data), NSGD when $p=2$, and Muon when $p\rightarrow\infty$.
+
+However, Schatten norms do not fill the rank gap: when $p > 1$, the update has full rank, while when $p=1$, its rank equals 1.
 Moreover, computing the update for $p \neq 1, 2, \infty$ appears to require knowing all $\sigma_i$, making the problem as hard as computing the full SVD.
@@
 \]
-depending on which of those two expressions achieves smaller residual on the linear function $L(\mX) := F(\mX^t) + \<\mM^t, \mX - \mX^t>$.
+depending on which of those two expressions achieves smaller residual on the linear function $L(\mX^{t+1}) := F(\mX^t) + \<\mM^t, \mX^{t+1} - \mX^t>$.
 The Ky Fan norms thus fail to close the rank gap: the resulting update is either rank-one or full-rank.
@@
 
-In \sectionref{sec:matrix-side}, we will discuss how to efficiently compute Fanion updates using the Lanczos algorithm.
+In \sectionref{sec:matrix-side}, we discuss how to efficiently compute Fanion updates using the Lanczos algorithm.
 
@@
 
-Simply applying norms dual to Ky Fan $k$-norms, however, does not provide sufficient algorithmic diversity for our purposes.
+%Simply applying norms dual to Ky Fan $k$-norms, however, does not provide sufficient algorithmic diversity for our purposes.
+The approaches to design new LMO algorithms are not limited to applying norms dual to Ky Fan $k$-norms.
 We now consider linear combinations of LMO-based algorithms and show that these combinations are themselves LMO-algorithms.
@@
 \begin{corollary}\label{corollary:lin_comb_alg}
-	Let there be a finite family of LMO based algorithms indexed by $i = 1,\dots,n$, where the $\mX^{t + 1}$ update in the $i$-th algorithm is defined by
+	Let there be a finite family of LMO-based algorithms indexed by $i = 1,\dots,n$, where the $\mX^{t + 1}$ update in the $i$-th algorithm is defined by
 	\[
@@
 
-The extremal members of this family are \emph{F-Neon} (with $k = 1$) and \emph{F-Muon} (with $k = \min\{m, n\}$).
+The extreme members of this family are \emph{F-Neon} (with $k = 1$) and \emph{F-Muon} (with $k = \min\{m, n\}$).
 Additional information and visualizations for the F-Muon generating $\normfstar{\cdot}^\dagger = \norm{\cdot}_{\mathrm{F-KF-1}}^\dagger$ and the F-Neon generating $\normftwo{\cdot}^\dagger = \norm{\cdot}_{\mathrm{F-KF-\min\{m, n\}}}^\dagger$ norms appear in the appendix (see \equationref{eq:normfstardual}, \equationref{eq:normftwodual}, \figureref{fig:fduals}).
@@
 
-The extremal members of this family are \emph{S-Neon} (with $k = 1$) and \emph{S-Muon} (with $k = \min\{m, n\}$).
+The extreme members of this family are \emph{S-Neon} (with $k = 1$) and \emph{S-Muon} (with $k = \min\{m, n\}$).
 
 \section{Computing the Updates}\label{sec:matrix-side}
-We employ the thick-restart Lanczos method for the symmetric eigenvalue problem (TRLan) to compute the low-rank updates of Fanions.
+We employ the thick-restart the Lanczos method for the symmetric eigenvalue problem (TRLan) to compute the low-rank updates of Fanions.
 We apply TRLan to either $\mM^{t\top}\mM^t$ or $\mM^t \mM^{t\top}$, selecting whichever matrix is smaller.
@@
   {%
-    \begin{tabular}{|c|c|c|c|c|c|c|c|}
-      \arrayrulecolor{black}\hline
+    \begin{tabular}{cccccccc}
+      \arrayrulecolor{black}\toprule
       Matrix sizes & $k$  & Method           & Time (s) & Matvecs & Iterations & $err_1$         & $err_2$       \\
@@
       $5000 \times 5000$    & 5  & TRLan            & 0.70    & 205     & 102      & $7.7 \cdot 10^{-3}$ & $7.7 \cdot 10^{-3}$\\
-      \arrayrulecolor{black}\hline
+      \arrayrulecolor{black}\bottomrule
     \end{tabular}%
@@
   {%
-    \begin{tabular}{|c|c|c|c|c|}
-      \arrayrulecolor{black}\hline
+    \begin{tabular}{ccccc}
+      \arrayrulecolor{black}\toprule
       Matrix size    & Time (s) & Matvecs & Iterations & $err_1$ \\
@@
       $5000 \times 5000$   & 26.4     & 290 000 & 29         & $6.5 \cdot 10^{-3}$  \\
-      \arrayrulecolor{black}\hline
+      \arrayrulecolor{black}\bottomrule
     \end{tabular}%
@@
 
-We begin by evaluating F-Fanions on the following convex $L$-smooth problem:
+We begin by evaluating F-Fanions and S-Fanions on the following convex $L$-smooth problem:
 \begin{equation}\label{eq:lls}
@@
   {%
-    \begin{tabular}{|c|c|c|c|}
-      \arrayrulecolor{black}\hline
+    \begin{tabular}{cccc}
+      \arrayrulecolor{black}\toprule
       Algorithm & \texttt{lr}  & \texttt{momentum}  & Iterations to 0.001 loss \\
@@
       S-Muon & 0.011 & 0.9 & 890 \\
-      \arrayrulecolor{black}\hline
+      \arrayrulecolor{black}\bottomrule
     \end{tabular}%
@@
     \vspace{1em}
-    \subfigure[The Frobenius norm of the gradient]{\label{fig:lls_fro_grad_norm}%
+    \subfigure[The Frobenius norm of the full gradient]{\label{fig:lls_fro_grad_norm}%
       \includegraphics[width=0.8\linewidth]{figs/simple_lls/gradient_frobenius_norm_vs_iteration_long_run.pdf}}
@@
 
-Both F-Muon and S-Muon converge faster to lower loss values and achieve lower Frobenius norms of the gradient than NSGD, Muon, or SignSGD.
+Both F-Muon and S-Muon converge faster to lower loss values and achieve lower Frobenius norms of the full gradient than NSGD, Muon, or SignSGD.
 
@@
     \resizebox{\linewidth}{!}{%
-      \begin{tabular}{|l|c|c|c|}
-        \arrayrulecolor{black}\hline
+      \begin{tabular}{lccc}
+        \arrayrulecolor{black}\toprule
         Algorithm                             & LMO constraint set $\mathcal{D}$ & LMO                                                         & Reference                            \\
@@
         Neon                               & Ball in Nuclear $S_1$           & $-\eta \vu_1 \vv_1^\top$                                        & This work                            \\
-        \arrayrulecolor{black}\hline
+        \arrayrulecolor{black}\bottomrule
       \end{tabular}%
@@
 A common limitation of these analyses is their focus on convergence measured by the gradient norm.
-As we showed in our CIFAR experiments \sectionref{sec:cifar_underhood}, the gradient norm may decrease by only a factor of ten when the accuracy reaches 100\%.
+As we showed in our CIFAR experiments \sectionref{sec:cifar_underhood}, the stochastic gradient norm may decrease by only a factor of ten when the accuracy reaches 100\%.
 
@@
 
-\subsection{The nuclear norm in the LMO}
+\subsection{The Nuclear Norm in the LMO}
 
@@
 
-\subsection{Orthogonal research}
+\subsection{Possible Improvements}
 
@@
 IO suggested using the nuclear norm in the \citep{bernstein2024oldoptimizernewnorm} framework.
-DM presented the problem at the MIPT optimization course.
-IK suggested using composite norms (though not the ones that induce Fanions, F-Fanions or S-Fanions) and helped to draft and rewise the manuscript.
-NK suggested Lanczos algorithm as the most precise means to compute Fanions' updates, conducted experiments to prove this, and wrote \sectionref{sec:matrix-side}.
+DM presented the problem at the MIPT optimization course, supervised the project, and helped to revise the manuscript.
+IK suggested using composite norms (though not the ones that induce Fanions, F-Fanions or S-Fanions) and helped to draft and revise the manuscript.
+NK suggested the Lanczos algorithm as the most precise means to compute Fanions' updates, conducted experiments to prove this, and wrote \sectionref{sec:matrix-side}.
 AV conducted the finetuning of NanoGPT on Tiny Stories and wrote \sectionref{subsec:finetune}.
@@
 
-\section{Visualization of different matrix norms}
-\subsection{Duals to F* and F2 norms}
+\section{Visualization of Different Matrix Norms}
+\subsection{Duals to F* and F2 Norms}
 
@@
 
-\subsection{The Ky Fan norm and its dual}
+\subsection{The Ky Fan Norm and Its Dual}
 
@@
 
-\section{More details for Linear Least Squares}
+\section{More Details for Linear Least Squares}
 \label{sec:lls_plots_section}
-The spectral and nuclear norms of the gradients over iterations, as well as the loss over time, are shown in \figureref{fig:app_lls}.
+The spectral and nuclear norms of the full gradients over iterations, as well as the loss over time, are shown in \figureref{fig:app_lls}.
 The poor performance of Fanions and F-Fanions in terms of speed for large $k$ is likely caused by the inefficient implementation of TRLan.
@@
   {%
-    \subfigure[The spectral norm of the gradient]{\label{fig:lls_op_grad_norm}%
+    \subfigure[The spectral norm of the full gradient]{\label{fig:lls_op_grad_norm}%
       \includegraphics[width=0.49\linewidth]{figs/simple_lls/gradient_spectral_norm_vs_iteration_long_run.pdf}}%
     \hfill
-    \subfigure[The nuclear norm of the gradient]{\label{fig:lls_nuclear_grad_norm}%
+    \subfigure[The nuclear norm of the full gradient]{\label{fig:lls_nuclear_grad_norm}%
       \includegraphics[width=0.49\linewidth]{figs/simple_lls/gradient_nuclear_norm_vs_iteration_long_run.pdf}}%
@@
   {%
-    \begin{tabular}{|c|c|c|c|}
-      \arrayrulecolor{black}\hline
+    \begin{tabular}{cccc}
+      \arrayrulecolor{black}\toprule
       Method & \texttt{lr}  & \texttt{momentum}  & val\_accuracy, \% \\
@@
       F-Fanion-5 & 0.40 & 0.6 & $86.66 \pm 0.65$\\
-      \arrayrulecolor{black}\hline
+      \arrayrulecolor{black}\bottomrule
     \end{tabular}%
@@
 
-\section{Technical details of the experiments}
+\section{Technical Details of the Experiments}
 

### 2025-12-01 article_final/main.tex
Edited with Grammarly

@@
 
-Minimizing loss functions in unprecedentedly high-dimensional spaces has recently become an integral and crucial part in training large language models.
+ Minimizing loss functions in unprecedentedly high-dimensional spaces has recently become an integral and crucial part of training large language models.
 Hence, new scalable, time- and memory-efficient algorithms have been demanded.
@@
 After 1750 iterations on NanoGPT training, F-Muon achieves 3.281 cross-entropy loss, while Muon achieves 3.279.
-Similarly, after 5960 iterations of GPT-2 Medium training we get 2.9215 with F-Muon and 2.9198 with Muon.
-Finally, S-Muon matches Muon on fine-tuning NanoGPT on Tiny Stories, while F-Muon is far more resistant to learning rate choice than Muon.
+Similarly, after 5960 iterations of GPT-2 Medium training, we get 2.9215 with F-Muon and 2.9198 with Muon.
+Finally, S-Muon matches Muon in fine-tuning NanoGPT on Tiny Stories, while F-Muon is far more resistant to learning rate choice than Muon.
 These results show that Muon-like algorithms can maintain competitive performance even when the underlying norm constraint is significantly modified, providing an affirmative answer to the central question posed above.
@@
  \]
-depending on which of those two expressions achieves smaller residual on the linear function $L(\mX^{t+1}) := F(\mX^t) + \<\mM^t, \mX^{t+1} - \mX^t>$.
+ depending on which of those two expressions achieves a smaller residual on the linear function $L(\mX^{t+1}) := F(\mX^t) + \<\mM^t, \mX^{t+1} - \mX^t>$.
 The Ky Fan norms thus fail to close the rank gap: the resulting update is either rank-one or full-rank.
@@
 Only two exceptional cases exist: the dual to the Ky Fan $1$-norm $\norm{\cdot}_{\text{KF-}1}$ (the spectral norm) is the Ky Fan $\min(m,n)$-norm $\norm{\cdot}_{\text{KF-}\min(m,n)}$ (the nuclear norm), and vice versa.
-Thus working with duals of Ky Fan norms can open up new possibilities:
+Thus, working with duals of Ky Fan norms can open up new possibilities:
 
@@
  %Simply applying norms dual to Ky Fan $k$-norms, however, does not provide sufficient algorithmic diversity for our purposes.
-The approaches to design new LMO algorithms are not limited to applying norms dual to Ky Fan $k$-norms.
+ The approaches to designing new LMO algorithms are not limited to applying norms dual to Ky Fan $k$-norms.
 We now consider linear combinations of LMO-based algorithms and show that these combinations are themselves LMO-algorithms.
@@
  \end{itemize}
-We also include baselines Normalized SGD and SignSGD, which correspond to F-Fanion and S-Fanion respectively with $\alpha = 0$ and arbitrary $k$.
+ We also include baselines Normalized SGD and SignSGD, which correspond to F-Fanion and S-Fanion, respectively, with $\alpha = 0$ and arbitrary $k$.
 
@@
  The tuned parameters and the number of iterations required to converge to $0.001$ are presented in \tableref{tbl:lls_lrs}.
-For intermediate-rank Fanions, F-Fanions, and S-Fanions, the hyperparameters \texttt{lr}, \texttt{momentum}, and \texttt{sign\_lr\_coeff} are set equal to those of Muon, F-Muon, and S-Muon respectively, as their loss decreases too slowly to reach $0.001$ within a reasonable tuning budget.
+For intermediate-rank Fanions, F-Fanions, and S-Fanions, the hyperparameters \texttt{lr}, \texttt{momentum}, and \texttt{sign\_lr\_coeff} are set equal to those of Muon, F-Muon, and S-Muon, respectively, as their loss decreases too slowly to reach $0.001$ within a reasonable tuning budget.
 
@@
    {fig:app_lls}
-  {\caption{More plots for linear least squares problem for a 500x500 matrix.}}
+   {\caption{More plots for the linear least squares problem for a 500x500 matrix.}}
    {%
@@
 
-A uniform training protocol was applied to all runs, including identical batch size, warm-up and cosine decay learning rate schedule, gradient clipping strategy, and validation split.
+ A uniform training protocol was applied to all runs, including an identical batch size, warm-up and cosine decay learning rate schedule, gradient clipping strategy, and validation split.
 Our optimization methods were integrated into the NanoGPT training loop without modifying the model architecture or preprocessing pipeline.

### 2025-12-01 article_final/main.tex
Resized the table

@@
    {\caption{Comparison of methods for computing rank-$k$ updates on dense random matrices (CPU, double precision). Lower is better in all columns.}}
-   {%
+   {\resizebox{\textwidth}{!}{%
      \begin{tabular}{cccccccc}
@@
      \end{tabular}%
+   }
    }
@@
  We generalized several successful algorithms, including Muon, Dion, and $\nu$SAM, to LMO-based algorithms using the family of norms dual to Ky Fan $k$-norms, yielding the Fanion family with low-rank updates.
-We further proposed the technique of regularizing these updates by combining them with NSGD or SignSGD, creating the F-Fanion and S-Fanion families through the $\norm{\cdot}_{\mathrm{F-KF-k}}^\dagger$ and $\norm{\cdot}_{\mathrm{C^\dagger-KF-k}}^\dagger$ norms.
+We further proposed the technique of combining them with NSGD or SignSGD, creating the F-Fanion and S-Fanion families through the $\norm{\cdot}_{\mathrm{F-KF-k}}^\dagger$ and $\norm{\cdot}_{\mathrm{C^\dagger-KF-k}}^\dagger$ norms.
 Our experiments demonstrate that F-Muon and S-Muon achieve competitive performance with Muon on CIFAR-10 airbench and NanoGPT tasks, confirming that the underlying norm constraint can be significantly modified without sacrificing effectiveness.
@@
  However, our results also reveal that not every LMO-based algorithm is effective: Neon (rank-one Fanion) underperforms despite sharing the same theoretical convergence asymptotics as Muon in existing bounds.
-We suggest that future work on non-Euclidean LMO algorithms should explain in the corollaries to their convergence theorems the empirical superiority of Muon over other Fanions, and ideally account for the robustness of F-Muon and S-Muon as well.
-Without such refinements, it is difficult to believe that current theoretical bounds are relevant for practitioners.
+We suggest that future work on non-Euclidean LMO algorithms should explain in the corollaries to their convergence theorems the observed empirical superiority of Muon over other Fanions, and ideally account for the robustness of F-Muon and S-Muon as well.
 
